### 2024-11-25 manuscript/manuscript.tex
add images and manuscript

% \section{Abstract}
\begin{abstract}
The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread. The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods. For the first several years of the competition, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalization counts. Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons. % Modeling hospitalization data alone is limited to the relatively small amount of data, 
% but may be supplemented by existing ILI data which has been reported, 
% published at the state level, and successfully forecast for over a decade. 
In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data. The first component is for modeling ILI data using a dynamic nonlinear Bayesian hierarchical model. The second component is for modeling hospitalizations as a function of ILI. For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor. In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from \cite{osthus2019dynamic} and a nonlinear Bayesian hierarchical model from \cite{ulloa2019} are compared. Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model. Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared. We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year. We also found that other modeling decisions such as the exact nonlinear function to use in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
% We compare hospitalization forecasts from our model to other related models submitted to the CDC forecasting competition.

\keywords{Disease outbreak forecasting \and Bayesian hierarchical modeling \and Probabilistic forecasting \and Model discrepancy}

%\pagebreak %remove if needed

% Please include sections as the paper has, some of the following sections are meant as examples of what can be done, the bibliography should be made as given

\section{Introduction}

Every year the seasonal influenza outbreak burdens the public health system by infecting millions, causing an influx of primary care visits and hospitalizations and leading to between 290,000 and 650,000 deaths worldwide \cite[]{whoflufact2023}. \cite{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion. Accurate forecasting of infectious diseases can inform public decision making and ease the burden of an outbreak \cite[]{turtle2021accurate, lutz2019applying}.
There is a growing consensus that disease forecasts should be probabilistic in nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has been shown that reporting forecast uncertainty along with predictions may lead to better decision making \cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.

% \jarad{Test}
% \spencer{Test}

To better inform public decision making regarding the flu epidemic, in 2013 the United States Centers for Disease Control and Prevention (CDC) organized a national flu forecasting competition, also known as FluSight \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. Originally, over a dozen teams of researchers from academic and industry backgrounds participated in FluSight by contributing their own forecast models. Besides the 2020 season --or the flu season spanning the fall of 2020 and the winter of 2021-- FluSight has been operated annually and researchers outside the CDC have been invited to participate. Initially the target data for forecasts was influenza-like illness (ILI) data. ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}. 
The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, \cite{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework. An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.

The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications. As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}. The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded. The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}. Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition. Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.

% HHS hospitalization data for the state and national level has only been provided since 2021 whereas ILI data has been available since 2010. 
The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models. The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
Herein we use ILI models similar to those in \cite{osthus2019dynamic} and \cite{ulloa2019} for ILI forecasting.
The model of \cite{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework. Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by \cite{kennedy2001bayesian}. The model in \cite{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data. Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.

In line with the newer FluSight standard of forecasting hospitalizations, we model hospitalizations as a linear function of ILI. Thus forecasts produced herein target flu hospitalizations and are a mapping of ILI forecasts to hospitalizations. This allows for ILI data from many seasons to be exploited and for ILI forecasts to assist in forecasting hospitalizations, which has fewer seasons of data than ILI. Several modeling schemes and their forecasts for the 2023 flu hospitalization season are compared, and it is shown that the modeling decisions produce good forecast results for different states or times during the flu season.

In section \ref{sec:data} we review the ILI and hospitalization data provided by the CDC and targeted by FluSight. 
In section \ref{sec:functions} the modeling framework contributed by this manuscript is given. In the same section, functions similar to those used by \cite{osthus2019dynamic} and \cite{ulloa2019} are defined.
 These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively. Model fitting and implementation are described at the end of the section. Section \ref{sec:simulation2} is a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared. Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts. Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}, with the analysis being done using the conventions of FluSight.  
 % including comparing the forecasts to a baseline forecast under the proper scoring rule used for scoring interval forecasts \cite[]{mathis2024evaluation}. 
 Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.

\section{Flu outbreak data} \label{sec:data}
In this section we introduce and define ILI and hospitalization data and evaluate the data visually. ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season. Both of these data were collected at the state, territorial, and national level and were reported at least weekly. Overall the data is reported for 53 locations including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), and at the US national level. We will refer to forecast targets throughout this manuscript. A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a specific location and week during the season.

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers. Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases. An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}. Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza. To know whether or not a sick patient is infected with influenza would require a laboratory test.

In 2013, when FluSight began, the ILI data was the object of the forecasts. The data was released publicly at HHS region levels, and forecast teams were asked to provide forecasts of several ILI targets on the regional levels including season onset, 1-4 week ahead ILI levels, and the week of peak ILI activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}. Currently, the ILI data is collected by the CDC and published on an online portal for viewing at the national, HHS region, census, and state levels \cite[]{cdc2024fluviewportal}. To obtain ILI data, we used the \texttt{R} package \texttt{cdcfluview} which provides functions for downloading the data \cite[]{rudis2021cdcfluview}. Weekly ILI data from the national, HHS region, and census levels are available from the 1997 flu season until the current season. At the state level, data is available from the 2010 flu season to the current season.

Figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023. For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons. 
To better align with the flu behavior, week 1 is set as the first week of August and week 52 or 53 is the last week in July of the following year. For example week 1 of the 2013 season corresponds to the first week of August 2013, and week 52 of the same season corresponds to the last week of July 2014. This convention is used for the remainder of this manuscript.

\begin{figure}
    \centering
    \includegraphics[scale=.7]{Images/us_ili_seasons.png}
    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023. Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.}
    \label{fig:us_ili}
\end{figure}

Notable from the plots in figure \ref{fig:us_ili} is the regular trajectory of the ILI. With the exception of season 2020, the ILI begins low at week 1 and increases as the fall and winter progress until the ILI reaches a peak. As spring progresses to summer, the ILI decreases to low values. As \cite{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day. 
% The model formulated by \cite{osthus2019dynamic} was largely defined by the week 22 peak and influences the model we introduce in \ref{sec:model}. 
Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
\cite[]{ewing2017contact, garza2013effect}.
The only seasons when there was not a peak at week 22 were season 2022 where the season peak occurred particularly early and season 2020 which was greatly influenced by the COVID-19 pandemic.
% which would have affected patient visitations as well as reporting.

Figure \ref{fig:ili_vs_week} shows ILI data from five states and the District of Columbia, locations which received particular attention in \cite{osthus2021multiscale}. The plots include the ILI data for all seasons from 2010 to 2023 in grey, and the black line is the per week ILI average over seasons. The patterns in the individual states are similar to the national level plots in figure \ref{fig:us_ili} in that the ILI rises in the fall and winter until it peaks and descends as the spring and summer progress. For these locations ILI regularly peaks, either locally or globally, at or near week 22.

 \begin{figure}
    \centering
    \includegraphics[scale=.7]{Images/ili_vs_week.png}
    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023. Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.
     Plots include lines for ILI\% from the 2010 flu season to 2023 (grey) and for the weekly ILI averaged over all seasons (black).}
    \label{fig:ili_vs_week}
\end{figure}

\subsection{Hospitalization Data}

Hospital admission data, used as the object of FluSight forecasting for the 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported Patient Impact and Hospital Capacity by State Timeseries}. Several targets of respiratory illnesses including COVID-19, RSV, and influenza are reported weekly by most hospitals in the US. In February 2022 it became mandatory for all hospitals to report the number of COVID-19 and influenza hospitalizations, and since then reporting of hospitalizations has become widespread. These data were updated every Wednesday and Friday according to NHSN guidelines \cite[]{healthdata2024covidts}.

Figure \ref{fig:us_hosp} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons. These plots show similarities to the ILI plots in \ref{fig:us_ili} in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends. For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22. Figure \ref{fig:hosp_vs_week} shows the 2022 and 2023 weekly hospitalizations for the same states from \ref{fig:ili_vs_week}. Similar to the national data, the peak in 2022 came early compared to the peak of 2023.
% Comparisons similar to those of the national data may be made here. 
% \jarad{Did you forget to make comparisons?}

\begin{figure}
    \centering
    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
    \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
    \label{fig:us_hosp}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
    \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
    \label{fig:hosp_vs_week}
\end{figure}

Comparing figures \ref{fig:ili_vs_week} and \ref{fig:hosp_vs_week} shows that ILI and hospitalizations share the similar pattern of increasing to a peak in the winter and decreasing thereafter. Figure \ref{fig:ili_vs_hosp} shows scatter plots with ILI\% on the $x$-axes and hospitalizations on the $y$-axes, revealing a positive somewhat linear relationship between the two variables. This relationship motivates the forecast models outlined in the next section.

\begin{figure}
    \centering
    \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
    \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
    \label{fig:ili_vs_hosp}
\end{figure}

% \subsection{Simple hospitalization model}
% \jarad{This should be the start of the methods section, so it should be a new section not subsection.}
% The CDC forecast competition between 2013 and 2020 seasons led to a number of advances and effective forecast models of ILI, and compared to weekly hospitalization data there is an abundance of ILI data. One contribution of this paper is that data and modeling techniques from forecasting ILI data may be used in forecasting hospitalizations. 
% A relationship between the ILI and hospitalization data exists, and to illustrate this we consider the model in equation (\ref{eq:simple_hosp_reg}). The model represents the hospitalizations for all weeks of one season and one state. For any state, $H_w$ is the hospitalization count at week $w$ and $ILI_w$ is the ILI value at the same week. The model is a squared polynomial regression with an autoregressive component of order 1. This is also known as an autoregressive with exogenous variables model (ARX(1)).
% \jarad{reference}

% \begin{equation}
%     H_{w} = \alpha_0 + \alpha_1 ILI_{w} + \alpha_2 ILI_{w}^2 + \phi H_{w-1} + \epsilon_{w}; \quad \epsilon_{w} \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
%     \label{eq:simple_hosp_reg}
% \end{equation}
% \jarad{This model still doesn't make a lot of sense to me. I feel like we have talked about this before. Why isn't the H series centered around the ILI mean, e.g. 
% $\phi (H_{w-1} -\alpha_0 - \alpha_1 ILI_w + \alpha_2 ILI_w^2)$.}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.8]{Images/hosp_preds_ili_22.png}
%     \caption{Weekly hospitalization counts during the 2022 flu season for 5 states and DC (black dots) and predicted weekly hospitalizations from the fit of model (\ref{eq:simple_hosp_reg}) (red lines)}
%     \label{fig:hosp_preds_ili_22}
% \end{figure}
% \jarad{Make the relationship between the data in the previous plot and this plot clear.}

% We fit this model for states individually under the 2022 flu season using the \texttt{lm} function in \texttt{R}.
% \jarad{It doesn't seem very important tha twe used lm in R.}
% Figure \ref{fig:hosp_preds_ili_22} shows the predicted hospitalizations from the fit model for several states. The black points are the true hospitalization counts and the red lines are the predicted hospitalizations from the fit model. 
% \jarad{Predicted or estimated?}
% The simple model does a decent job predicting hospitalizations, but forecasting future hospitalizations is more difficult. 
% \jarad{Why?}
% This, however, demonstrates that it is reasonable to model the behavior of hospitalizations as a linear function of ILI. 
% \jarad{You have a quadratic function of ILI not a linear function.}
% In the following two sections, nonlinear functions used in modeling and the proposed model are described.

\section{ILI and hospitalization forecast modeling} \label{sec:functions}

% \jarad{Why not start with the ILI model from Osthus et al? It makes a lot more sense to me to start with what has already been done and then expand upon what they have done. Later you have a section called ILI Model, but isn't this the ILI model?}

The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows of a similar trajectory for modeling ILI. Compartmental models are standard mathematical models used for disease outbreaks. One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}. \cite{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data. The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy. In the first part of this section, we present an ILI model similar to the model in \cite{osthus2019dynamic}. With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.

With the aim of forecasting hospitalizations, we also introduce a linear model of hospitalizations data with ILI as a predictive covariate. To forecast hospitalizations, ILI data is first forecasted and the forecast is then plugged in as a covariate in the hospitalization model, thus producing hospitalization forecasts. The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.
% In this section, we review the SIR and ASG functions as well as the lack of either model to capture certain behavior. We call the systematic deviation between the functional model and the observed behavior \textit{discrepancy}.

\subsection{ILI Model} \label{sec:ili_model}

The proposed model for ILI for any location is given in (\ref{eq:ili_model}). Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given season. The ILI is a proportion, so the Beta random variable is a natural selection for modeling. Under the parameterization in of the Beta distribution used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a scale parameter. The nonlinear function $f_{\theta_s}(w)$ captures the trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the systematic patterns which $f_{\theta_s}(w)$ does not capture.

\begin{equation}
    \label{eq:ili_model}
    \begin{flalign*}
        &ILI_{s,w} \overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
        &\text{logit}(\pi_{s,w}) = f_{\theta_s}(w) + \gamma_w
    \end{flalign*}
\end{equation}
In \cite{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}. In \cite{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}. In \cite{ulloa2019} modeling is done hierarchically over seasons. 
% The SIR and ASG functions are defined in the next two subsections.
% and $\gamma_w$ is modeled as a reverse random walk as in (\ref{eq:rw}).

\subsection{Susceptible-Infectious-Recovered (SIR) compartmental model} \label{sec:sir_func}

The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by \cite{kermack1927contribution}. Since then, compartmental models have become standard for modeling infectious diseases \cite[]{allen2008mathematical}, and many extensions have been made and studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}. The SIR mathematical model includes three compartments and assumes that at any time $t>0$ every individual in a closed population belongs to exactly one compartment. The three compartments are susceptible ($S$), infectious ($I$), and recovered ($R$), and their interaction over the course of an outbreak is described by the differential equations in (\ref{eq:sir_diff}).

\begin{equation}
    \label{eq:sir_diff}
    \frac{dS}{dt} = -\beta SI, \quad \frac{dI}{dt} = \beta S I - \delta I, \quad \frac{dR}{dt} = \delta I
\end{equation}
Here $S$, $I$, and $R$ represent the proportion of the population in each compartment such that $S + I + R = 1$ for all $t$. The trajectory is determined by the disease transmission rate $\beta > 0$ and the recovery rate $\delta > 0$. Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual, and the expected rate of recovery to an immune state for a newly infected person. Whether or not a disease outbreak is classified as an epidemic is determined by the initial susceptible population $S_0$, or the susceptible population at time $0$, and the parameter $\rho = \delta/
\beta$. If $S_0/\rho > 1$, the outbreak is considered an epidemic. It is non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}. Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where the $S_0$ and $\rho$ were selected to match an outbreak that would classify as epidemic. In the case where $S_0 \leq \rho$, the trajectory for the $I$ compartment would never be increasing. The increase to a peak and subsequent decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is reasonable to model ILI by this compartment. Thus in modeling the ILI data, we consider the data to be analogous to the $I$ proportion of the population.

\begin{figure}
    \centering
    \includegraphics[scale=.6]{Images/sir_traj.png}
    \caption{Susceptible-infectious-recovered (SIR) model separated by compartments. The three compartments are the susceptible compartment (left), infectious (center), and recovered (right). In this example, $S_0/\rho > 1$.}
    \label{fig:sir_traj}
\end{figure}

\subsection{Asymmetric Gaussian (ASG) function} \label{sec:asg_func}
The ASG function is another example of a nonlinear function which can approximate the trajectory of the flu outbreak. The ASG was previously used by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used to model vegetation growth and satellite sensor data \cite[]{lewis2020extracting, jonsson2002seasonality, hird2009noise, beck2006improved, atkinson2012inter}. The ASG is a modification of the asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized by its rise to a peak and fall from that peak which may not occur at the same rate, as shown in figure \ref{fig:asg_function}. The ASG function is denoted as $ASG_\theta(w)$ where $\theta = (\lambda, \nu, \mu, \sigma_1^2, \sigma_2^2)$, $\nu > 0$, $\lambda > 0$, $\mu \in (-\infty, \infty)$, $\sigma_1, \sigma_2 > 0$ and $w \in (1, ..., W)$ is week. The function is defined in (\ref{eq:asg_function}).

\begin{figure}
    \centering
    \includegraphics[scale=.45]{Images/asg_function.png}
    \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
    \label{fig:asg_function}
\end{figure}

\begin{equation}
    \label{eq:asg_function}
    ASG_{\theta}(w) = 
    \begin{cases}
        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
    \end{cases}
\end{equation}
For modeling in this manuscript, we use a slightly reparameterized version of the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$. This constraint guarantees that the function has a peak greater than $\lambda$.

\begin{equation}
    \label{eq:asg_function_rep}
    ASG_{\theta}(w) = 
    \begin{cases}
        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
    \end{cases}
\end{equation}

\subsection{Model discrepancy}

The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as \cite{osthus2019dynamic} points out there may be systematic behavior that these or other possible functions may not capture. As noted in section \ref{sec:data}, figures \ref{fig:us_ili} and \ref{fig:ili_vs_week} show a regular peak at week 22 of the flu season. 
% It has been speculated that this drop may be the result of taking holiday from work or school or from the public wanting to avoid medical care around the holidays (??). 
Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to illustrate the systematic discrepancy from a fitted function. Figure \ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 to 2022 excluding 2020 with a best fit ASG function plotted over the ILI. The fits for each season were made by obtaining the maximum likelihood estimate (MLE) of a model given the ILI data where we assume the ASG function is the mean parameter of a Beta distributed random variable. Figure \ref{fig:discrepancy} shows the discrepancy between the model fit and the data for the same seasons. The grey lines show the difference between the data and the functions from figure \ref{fig:asg_fits} for each season, and the black line is the average by week over all seasons. The lines show that the ASG function typically underpredicts week 22 and overpredicts week 23. Perhaps for other weeks, like week 30 for example, there also tends to be systematic behavior not captured by the ASG function.

\begin{figure}
    \centering
    \includegraphics[scale=.5]{Images/asg_fits.png}
    \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
    \label{fig:asg_fits}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.45]{Images/discrepancy.png}
    \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
    \label{fig:discrepancy}
\end{figure}

The term $\gamma_w$, where $w$ is the season week, is included in (\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the function. Modeling discrepancy has been used in uncertainty analysis of simulators to capture systematic differences between mathematical models and 
reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,arendt2012improving,kennedy2001bayesian}. Modeling discrepancy can lead to overfitting, particularly in forecasting scenarios, and may also lead to identifiability issues. 
Thus, care must be taken in setting parameter constraints as well as in the selection of prior distributions \cite[]{osthus2019dynamic,brynjarsdottir2014learning}. 
Modeling of the discrepancy for ILI was done by \cite{osthus2019dynamic} during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}. As in Osthus's model, $\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}). 
\begin{equation}
    \label{eq:rw}
        \gamma_w|\gamma_{w + 1} &\overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
        \gamma_{W} &\sim N(0,\sigma^2_{\gamma_W})
\end{equation}
The idea for using the reverse random walk is that there are several previous seasons of ILI data, and assuming the random walk captures systematic behavior, fitting it hierarchically over seasons can assist in predicting future behavior in the current season. 
Reverse random walks have also been used with success in election forecasting and other flu forecasting models \cite[]{osthus2021multiscale, osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve identifiability.

\subsection{ILI forecasts}
Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating. Future ILI forecasts are obtained via the posterior predictive distribution where for week $w$, the predictive distribution is obtained by integrating over the parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ as in (\ref{eq:ili_post}) where $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ is the density function of the posterior distribution for the model parameters. If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ where $i$ is a positive integer.

\begin{equation}
    \label{eq:ili_post}
    p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) = \int \int \int \int p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
\end{equation}
% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
% \spencer{Isn't that captured in $d \pi_s$?}

% Below we consider four variations of model (\ref{eq:ili_model}). These include where $f_{\theta_s}(w) = ASG_{\theta}(w)$ and where $f_{\theta_s}(w) = \text{logit}(I_{s,w})$, $I_{s,w}$ being the infected compartment of an SIR model. For both function scenarios, models are fit where the discrepancy component $\gamma_w$ is included and excluded.

\subsection{Hospitalization model} \label{sec:hospital_model}

The second component for forecast modeling is the hospitalization model defined in (\ref{eq:final_hosp}). This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}. Here $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.  
% We further construct a forecast model for hospitalizations as a linear model of the ILI model. Similar to \ref{eq:simple_hosp_reg}, we select the model in (\ref{eq:final_hosp}). The desired outcome of this model is that the relationship between ILI and hospitalizations may be harnessed, and a future forecast of hospitalizations may be informed by a reasonable forecast for ILI as well as by past hospitalizations. 
\begin{equation}
    \begin{aligned}
    \label{eq:final_hosp}
    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, \omega_s) %N(0, \sigma_{\epsilon}^2)
    \end{aligned}
\end{equation}
This model is for any location, and for fitting purposes $ILI_{s,w}$ is always multiplied by $P$ which is proportional to the population of the state or territory, in this case the total population divided by 50,000. This is done as a means of scaling so that the prior distribution assigned to  $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, $\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all states.

% Like the ILI model, model (\ref{eq:final_hosp}) was fit via MCMC sampling.
Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also fit via Bayesian posterior updating. To obtain forecasts for $H_{s, w^* + i}$, the ILI posterior predictive distribution is used along with the posterior distribution for the parameters in (\ref{eq:final_hosp}). We considered three scenarios for model (\ref{eq:final_hosp}). A model where $D_s$ belongs to the normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, or $H_{s,w} + 1$ is lognormally distributed (LNORM). The population value $P$ is excluded from the LNORM model, and we set $\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting. In the LNORM model, if the linear parameters are not set the same for seasons 22 and 23, the final variance was more prone to be extreme. Besides varying the distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ or there is no quadratic ILI term.

\subsection{Prior selection}

The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in \cite{osthus2019dynamic} and \cite{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem. For model (\ref{eq:ili_model}), parameters which are common even when using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, $\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$. For the SIR function $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.

% \begin{equation}
%     \label{eq:full_prior}
%     \begin{align*}
%     p(\theta, \Sigma, \kappa_s, \sigma_{\gamma}^2, \sigma_{\gamma_W}^2, \Psi) &= p(\theta)p(\Sigma)p(\kappa_s)p(\sigma_{\gamma}^2) p(\sigma_{\gamma_W}^2)p(\Psi), \\
%     p(\Psi) &= p(\alpha_{0s})p(\alpha_{1s})p(\alpha_{2s})p(\phi)p(\sigma_{\epsilon_s})p(\omega_s)
%     \end{align*}
% \end{equation}
% \jarad{What purpose is this serving?}

% Where the component parameters are shared for both SIR and ASG models, the individual parameter prior distribution assignments when $D_s$ from \ref{eq:final_hosp} is a normal distribution are in (\ref{eq:shared_priors}).
% \jarad{You cannot start a sentence like this. I think you intend for "where" to be a continuation of the equation, but the equation may be anywhere relative to the text, i.e. you cannot count on the "where" being immediately after the equation.}
The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$. For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by \cite{gelman2006prior}, the priors were assigned to the variance parameters. Univariate parameters were assigned either a normal distribution prior if the support is on $\mathbb{R}$, a half-normal prior if the support is nonnegative, or a truncated-normal prior to match a more specific support. Under the ASG model, $\theta_s$ is modeled hierarchically over seasons so that for each season the transformed parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are assigned to $\theta$ and $\Sigma$.

% Because parameter identifiability is lost when modeling with a discrepancy term \cite[]{brynjarsdottir2014learning}, additional constraints on the prior distributions should be made to improve identifiability. 
Additional prior constraints were made to improve parameter identifiability. In \cite{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$. The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative priors. To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the parameters. A modular Bayesian approach involves multiple steps of parameter fitting where some parameters may be estimated without priors via maximum likelihood estimation or other means. Fitting the rest of the model parameters involves assigning priors and conditioning on the previously fit parameters. This has been done in computer modeling to improve identifiability and other issues, though \cite{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. We carried out the modular fit by first estimating the maximum likelihood estimate (MLE) for the parameter $\lambda_s$ for each season in (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.

% \spencer{Blue stuff below will be either included further on or will be removed entirely

% \begin{equation}
%     \label{eq:shared_priors}
%     \begin{align*}
%         \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
%         \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
%         \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
%         \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2) \\
%         \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2) \\
%         \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
%         \phi &\sim N(0, .4)\mathbbm{1}_{(-1,1)} \\
%         \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
%         \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
%     \end{align*}
% \end{equation}
% \jarad{Generally introduce the distribution families in the methods, but the distributions themselves when performing the data analysis. Here I think you just need to say that you are using normals for unrestricted parameters, positive normals for positive parameters and a normal truncated on (-1,1) for the correlation. Also mention they are all independent (which I guess was done before.}

% }

% \section{Data Model} \label{sec:model}

% The modeling framework we introduce in this section is two-fold. To produce a forecast, we fit a model of only ILI data and fit a second model of hospitalization data with ILI as a linear covariate. The two models are fit separately, and a probabilistic forecast of ILI is then used as a predictive covariate for hospitalizations in a second model. We chose this type of framework because of the apparent relationship between ILI and hospitalizations, the availability of several flu seasons of ILI data compared to only two seasons of hospitalizations data, and the success of the ILI models introduced by \cite{osthus2019dynamic} and \cite{ulloa2019}. In this section we outline the two data models, the Bayesian analysis used for fitting including prior selection, and assess the posterior samples.

\subsection{Parameter estimation and posterior predictive sampling}
\label{sec:implementation_posterior}

The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the \cite{stan2024manual} \cite[]{gabry2022stan}. Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides several diagnostic statistics for assessing the sampler.
As mentioned, most model parameter prior distributions were intended to be uninformative. Appendix \ref{app:B_prior} shows plots of posterior distributions for select parameters from ILI and hospitalization models.

We assessed the model fit for four ILI models. These included the SIR and ASG models and models with and without discrepancy modeling. When discrepancy is included, the models are denoted as SIRD and ASGD. These models were fit using US national data from 2010 to 2023 flu seasons, where data from the 2020 season was excluded because of the unique behavior during that season. Assessment for hospitalization modeling was done for six different models. These include NORM, LNORM, and LST models, and models where a quadratic ILI term is included or excluded. To assess posterior sampling convergence, models were fit to data where ILI and hospitalization data up to week 14 of the 2023 season was included. Sampling was done with four chains where from each chain 60,000 posterior draws were sampled, and the first 10,000 draws were discarded as a burn-in. The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for each parameter. Tables \ref{tab:ili_diagnostics} and \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum $ESS$ over all parameters for ILI and hospitalization models respectively. Forecast models for all other weeks of the season were fit using one chain of 60,000 draws where the first 10,000 draws were discarded as a burn-in. 
% \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
For parameters of the ASG models that were prone to cause trouble in posterior sampling, the starting values were set to be the MLEs.

\begin{table}
\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI models fitted on US data for week 14 of the 2023 season}
\label{tab:ili_diagnostics}
\begin{center}
    \begin{tabular}{c|cccc}
     & ASG & ASGD & SIR & SIRD \\
     \hline
     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
     $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
    \end{tabular}
    \end{center}
\end{table}

\begin{table}
\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six hospitalization models fitted on US data for week 14 of the 2023 season}
\label{tab:hsop_diagnostics}
\begin{center}
    \begin{tabular}{c|cccccc}
     & NORM & NORM^2 & LNORM & LNORM^2 &  LST &  LST^2 \\
     \hline
     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$\\
     $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
    \end{tabular}
    \end{center}
\end{table}

%  \begin{tabular}{ccccc}
% \hline
% \multicolumn{2}{c}{Set A} %%\\               %% <-- mistake
% %% \cline{1-2}                               %% <-- mistake
% &                                            %% <--  addition
% \multicolumn{2}{c}{Set B} \\\hline           %% <--  Changed
% Weight& Diameter    & Weight &Diameter      \\
% \hline
% \end{tabular}

To obtain forecast distributions of hospitalizations, draws from the posterior predictive distribution from the ILI model were used in conjunction with the posterior distribution of the hospitalizations model. When fitting model (\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were saved. Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal distributions for the model parameters were saved. To obtain forecast distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following steps are repeated $K$ times where $K$ is an integer for the number of desired samples. We set $K = 50,000$.

\begin{quote}
\begin{enumerate}[Step 1:]
  \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior distributions
  \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
  \begin{itemize}
    \item[] $\mu_{s,w + i}^* = \alpha_{0s}^* + \alpha_{1s}^* (ILI_{s,w + i}^* \times P) + \alpha_{2s}^* (ILI_{s,w + i}^* \times P)^2 + \phi^* H^*_{s,w + i - 1}$
  \end{itemize}
   \item Repeat step 3 for $i \in \{1,2,3,4\}$ to obtain $H^*_{s,(w + 1):(w + 4)}$
  \item Repeat steps 1-4 $K$ times
\end{enumerate}
\end{quote}
The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast for hospitalizations at week $w + i$. For the forecast competition analysis in section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were set to 0 to reflect realistic values of hospitalizations and comply with the FluSight forecasting rules.

% \subsection{Posterior predictive forecast} \label{posterior_forecast}

% Probability forecasts for state hospitalizations are obtained via the posterior predictive

% During the season, weekly hospitalization data was updated on Wednesdays, and forecast submissions were due that evening. ILI data was updated on Fridays leaving a lag of one week between the two sets of data when submissions were made. Because of this lag, for the most recent week $w^*$, $H_{s,w^*}$ is fit according to (\ref{eq:final_hosp}) but where $ILI_{s,w^*}$ is replaced by the posterior predictive distribution in (\ref{eq:ili_post}). For a given state, \textbf{ILI} is all observed ILI values the last value being $ILI_{s,w^*}$, and $\boldsymbol{\pi}_s = (\pi_{s,1}, ..., \pi_{s,w^*-1})$. Note that $w^*$ may also be replaced by any future $w + i$, $i \in (0, 1, ..., W)$.

% \begin{equation}
%     \label{eq:ili_post}
%     p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}, 
%     \textbf{H}) = \int_{\Pi \times \mathcal{K}} p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s) p(\boldsymbol{\pi}_s, \kappa_s | \textbf{ILI}, \textbf{H}) d(\boldsymbol{\pi}_s \times \kappa_s)
% \end{equation}

% Future hospitalization forecasts come from the posterior predictive distribution. We are interested in forecasting $\tilde{H}_{s,w + i}$ where $i \in (1, 2, 3, 4)$ meaning we forecast hospitalizations for 1, 2, 3, and 4-weeks ahead. The posterior predictive distribution of $\tilde{H}_{s,w + i}$ is defined in (\ref{eq:forecast_posterior}). Here $\Psi = (\beta_{0s}, \beta_{1s}, \beta_{2s}, \phi, \sigma_{\epsilon_s}, \nu_s)$ or the vector of all parameters in (\ref{eq:final_hosp}).
% %$\Psi = (\boldsymbol{\theta}, \boldsymbol{\beta})$

% \begin{equation}
%     \label{eq:forecast_posterior}
%     \begin{align*}
%         p(\tilde{H}_{s, w+i} |\textbf{H}, \textbf{ILI}) &= \\ \int_{\boldsymbol{\Psi} \times (0,1)} p&(\tilde{H}_{s, w+i} | \Psi) p(\Psi | \textbf{H}, \textbf{ILI}, \widetilde{ILI}_{s,w+i})  p(\widetilde{ILI}_{s,w+i} | \textbf{ILI})d (\Psi \times \widetilde{ILI}_{s,w+i}) 
%     \end{align*}
% \end{equation}

% After calculating (\ref{eq:forecast_posterior}) via MCMC sampling, the 23 quantiles requested by the CDC were calculated using the \texttt{quantile} function in \texttt{R} and truncated at 0. As stated, the focus of this paper is on forecasting hospitalizations when ILI is modeled according to the ASG function, but the work is heavily influenced by \cite{osthus2019dynamic}, so we compared forecasting ability to models where ILI is modeled similarly to the model proposed by Osthus.

% \begin{equation}
%     \label{eq:hosp_model}
%     \begin{flalign*}
%         \text{log}(H_w) = \alpha_0 + \alpha_1 \times \left[\text{logit}(I_{s,w})\right] + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
%     \end{flalign*}
% \end{equation}

% \begin{equation}
%     \label{eq:hosp_model}
%     \begin{flalign*}
%         \text{log}(H_w) = \alpha_0 + \alpha_1 \times \left[\text{log}(I_{s,w}) - \text{log}(1 - I_{s,w})\right] + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
%     \end{flalign*}
% \end{equation}

% \begin{equation}
% \label{eq:hosp_mod}
%     \begin{flalign*}
%         H_w &= \alpha_0_s + \alpha_1_s (I_{s,w} \times P) + \alpha_2_s (I_{s,w} \times P)^2 \\ &+ \alpha_3_s (I_{s,w} \times P)^3 + \phi_s H_{s, w-1} + \epsilon_{s,w} \\
%         \epsilon_{s,w} &\overset{iid}{\sim} N(0, \sigma^2_{\epsilon})  
%     \end{flalign*}

% \end{equation}

% \begin{equation}
% \label{eq:hosp_mod}
%     \begin{flalign*}
%         H_w &= \mu_s + \phi_s( H_{s, w-1} - \mu_s)+ \epsilon_{s,w} \\
%         \epsilon_{s,w} &\overset{iid}{\sim} N(0, \sigma^2_{\epsilon})  
%     \end{flalign*}

% \end{equation}

% where $\mu_s = \alpha_0_s + \alpha_1_s (I_{s,w} \times P) + \alpha_2_s (I_{s,w} \times P)^2 + \alpha_3_s (I_{s,w} \times P)^3$.

% Another option for modeling $H_w$ as a function of ILI is seen in

% \begin{equation}
%     \label{eq:hosp_model_sqrt}
%     \begin{flalign*}
%         \sqrt{H_w} = \alpha_0 + \alpha_1 \times I_{s,w} + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
%     \end{flalign*}
% \end{equation}

\section{Simulation Study} \label{sec:simulation2}
In this section, we present a simulation study conducted for comparing ILI models and further assessing the hospitalization forecast model. US ILI data is used, and hospitalization data is simulated. 
A leave-one-season-out (LOSO) approach was combined with a Monte Carlo simulation approach. For each replication, we simulated log-hospitalizations for all weeks during seasons 2010, ..., 2022, excluding 2020, using the existing ILI data as a predictive covariate. Each season was in turn ``left-out" and treated as if it was the most recent season which we desired to forecast. Fitting and forecasting was then done for weeks 14, 20, 26, 32, and 38 of the left out season, giving two weeks that tend to occur as flu cases increase, two as cases decrease, and one that occurs when the cases may be increasing or decreasing. Week 20 is a week leading up to the holiday week 22 where ILI typically has a local peak. We were particularly interested in how important modeling discrepancy is for forecasting at week 20.

For the simulation of hospitalizations, the parameters $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and $\sigma^2_{\epsilon_s}$ from the hospitalization model in (\ref{eq:final_hosp}) were considered the same across all seasons so that all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$. 
The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization data from the 2022 season. For fitting, the hospitalization data was first log-transformed. We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
and assigned the noninformative prior $p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$. The marginal posterior distribution $\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ was then the established posterior multivariate normal distribution and $\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior distribution \cite[]{gelman2013bayesian}. The posterior means of those parameters were used as the values from which log-hospitalizations were simulated. The number of replicates in the simulation was 500.

% Predictive coverage was assessed at the 0.5 and 0.95 levels and measures what proportion of forecasts were captured by the corresponding intervals.
Model comparison was done by calculating the continuous ranked probability score (CRPS) and the logarithmic score (LogS) for each forecast. These scores are both proper scoring rules which evaluate the forecast distribution and density functions respectively. Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}. The CRPS is defined in (\ref{eq:crps2}) and the LogS in (\ref{eq:logs2}). Here $F(\cdot)$ is the forecast distribution function, $f(\cdot)$ is the forecast density function, and $y^*$ is the observed targeted value of the forecast. The orientation of both scores is negative, meaning the smaller the score the better.

\begin{equation}
    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \mathbbm{1} (y^* \leq x))^2 dx
    \label{eq:crps2}
\end{equation}

% Having forecast distributions as posterior MCMC draws, the AE is calculated as the absolute difference between the sample median and the observed value.

\begin{equation}
    \text{LogS}(f, y^*) = -\text{log}(f(y^*))
    \label{eq:logs2}
\end{equation}
Both the CRPS and LogS are calculated using the \texttt{scoringRules} package in \texttt{R} \cite[]{jordan2019scoring}.
To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation. The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.

% The number of replicates was selected by assessing the Monte Carlo coverage of absolute predictive error AE for 150 replicates. The AE mean and standard deviation were calculated across season, week, and horizon for each season, week and horizon included in the simulation. For each of these levels, we wanted a Monte Carlo confidence width to be less than 25, so we solved for $R$ the following

% \[
%     25 = 2\times 1.96 * S_{m,s,w,h} / \sqrt{R}
% \]
% where $S_{m,s,w,h}$ is the SD for the MC AE for model $m$ at season $s$, week $w$, and horizon $h$. Across all $m, s, w, h$, the largest $R$ value was 1,030.533.

% Tables \ref{tab:horizon_coverage} and \ref{tab:seasonweek_coverage} show the nominal coverage of forecasts for the simulated data. For each forecast, 50\% predictive interval was calculated with by finding the $25^{th}$ and $75^{th}$ percentiles and the 95\% predictive interval by the $2.5^{th}$ and $97.5^{th}$ percentiles. Table \ref{tab:horizon_coverage} show the coverage for each model averaged over all horizons as well as for horizons of 1-4 weeks ahead. Table \ref{tab:seasonweek_coverage} shows coverage averaged over the season weeks 14, 20, 26, 32, and 38 forecasts. None of the four models possess the nominal coverage as in each case the coverage is below the 0.5 or 0.95 targeted proportion. This is due more to the forecasts' inability to nominally cover the true ILI data and not the simulated hospitalization data.

% \begin{table}
% \caption{}
% \label{tab:horizon_coverage}
% \begin{center}
%     \begin{tabular}{cccccc|ccccc}
%      &
%      \multicolumn{5}{c}{50\%}
%     & 
%     \multicolumn{5}{|c}{95\%} \\
%      Horizon week & Overall & 1 & 2 & 3 & 4 & Overall & 1 & 2 & 3 & 4\\ \hline
%      % Model & \\ \hline
%      ASG & \textbf{0.23} & 0.20 & 0.20 & \textbf{0.24} & \textbf{0.27} & 0.56 & 0.62 & \textbf{0.56} & 0.57 & \textbf{0.63} \\
%      SIR & 0.14 & 0.17 & 0.15 & 0.09 & 0.14 & 0.44 & 0.57 & 0.44 & 0.36 & 0.39\\
%      ASGD & 0.18 & 0.21 & 0.13 & 0.18 & 0.20 & 0.54 & 0.67 & 0.51 & 0.47 & 0.48\\ 
%      SIRD & 0.22 & \textbf{0.30} & \textbf{0.24} & 0.20 & 0.16 & \textbf{0.61} & \textbf{0.70} & 0.53 & \textbf{0.62} & 0.60\\ 
%     \end{tabular}
%     \end{center}
% \end{table}

% \begin{table}
% \caption{}
% \label{tab:seasonweek_coverage}
% \begin{center}
%     \begin{tabular}{cccccc|ccccc}
%      &
%      \multicolumn{5}{c}{50\%}
%     & 
%     \multicolumn{5}{|c}{95\%} \\
%      Season week & 14 & 20 & 26 & 32 & 38 & 14 & 20 & 26 & 32 & 38 \\ \hline
%      % Model & \\ \hline
%      ASG & \textbf{0.35} & 0.12 & 0.29 & 0.28 & \textbf{0.11} & 0.77 & 0.24 & 0.69 & \textbf{0.77} & \textbf{0.52} \\
%      SIR & 0.18 & 0.07 & 0.11 & 0.30 & 0.04 & 0.51 & 0.14 & 0.51 & 0.57 & 0.46\\
%      ASGD & 0.07 & 0.27 & \textbf{0.30} & 0.19 & 0.07 & 0.43 & 0.64 & \textbf{0.89} & 0.56 & 0.17\\ 
%      SIRD & 0.28 & \textbf{0.30} & 0.16 & \textbf{0.36} & 0.02 & \textbf{0.91} & \textbf{0.78} & 0.38 & 0.57 & 0.42\\ 
%     \end{tabular}
%     \end{center}
% \end{table}

Figures \ref{fig:overall_scores} - \ref{fig:logs_by_week_horizon} show boxplots of the CRPS and LogS for the four models. Figure \ref{fig:overall_scores} shows that the variation of overall CRPS is smallest for the ASG models and larger for the SIR models. The median scores for the SIR models also appears slightly higher than for the ASG models. 
When faceted by season in figure \ref{fig:crps_by_season}, the boxplots of the CRPS often show the same pattern for ASG and SIR models but not always. For example, the bulk of CRPS values for the SIRD model in 2019 appears to have smaller variation than the other models.
Figure \ref{fig:crps_by_week_horizon} shows CRPS boxplots faceted by week and horizon. Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, have CRPS values with lower medians and lower variation than the two models not accounting for ILI discrepancy. This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough.

\begin{figure}
\centering
% \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale=.35]{Images/overall_crps.png}
    % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale=.35]{Images/overall_logs.png}
    % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
% \end{subfigure}
\caption{Boxplots of the continuous ranked probability score (CRPS) (left) and logarithmic score (LogS) (right) for the four ILI models over all seasons, weeks, and horizons in the simulation study}
\label{fig:overall_scores}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.6]{Images/overall_crps.png}
%     \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
%     \label{fig:crps_overall}
% \end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.5]{Images/crps_by_season.png}
    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
    \label{fig:crps_by_season}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.5]{Images/crps_by_week_horizon.png}
    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis). Horizons include 1-4 week ahead forecastss and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
    \label{fig:crps_by_week_horizon}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.6]{Images/overall_logs.png}
%     \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
%     \label{fig:logs_overall}
% \end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.5]{Images/logs_by_season.png}
    \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
    \label{fig:logs_by_season}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.5]{Images/logs_by_week_horizon.png}
    \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis). Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
    \label{fig:logs_by_week_horizon}
\end{figure}

% This simulation study shows that the predictive coverage of each of the four models considered is well below the nominal coverage based on constructing intervals using quantiles. 
Figures \ref{fig:logs_by_season} and \ref{fig:logs_by_week_horizon} show boxplots of the LogS for the simulated forecasts. Note that in these plots only values of less than 17 are included to improve visualization. 
% Figure \ref{fig:overall_scores} shows less dramatic differences in the LogS variation than the CRPS. There are a few notable differences between what the LogS captures vs. the CRPS.
The LogS plots show similar results to the CRPS, though there are some differences. For example, SIRD in figure \ref{fig:logs_by_season} tends to show smaller LogS variation relative to the other three models than is seen by the CRPS of SIRD in figure \ref{fig:crps_by_season}. 
We also note the smaller relative LogS variation than CRPS variation for SIRD when comparing figure \ref{fig:logs_by_week_horizon} with figure \ref{fig:crps_by_week_horizon}, particularly at week 14. 
Among the four models, ASG tends to have the lowest values in CRPS and LogS though this is not always the case. When forecasting weeks 21-23, the two models which model discrepancy in the ILI, ASGD and SIRD, tend to outperform the models not modeling the discrepancy. This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22. 
% There are also slight differences in evaluation between the CRPS and the LogS. When scoring forecasts the two scores will not always recommend the same model(s). 
% Which of these or other scoring rules used in judging forecasts should reflect specific problem at hand.

\section{Analysis of forecasts for 2023 flu season}
\label{sec:analysis}

In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations. The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$. A single forecast is thus comprised of 11 predictive intervals and a median.
Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, and forecasts were made at the state and national level. The first week of forecasting took place during the week of October 7, 2023, and the final week was the week of April 27, 2024 making 29 total weeks of forecasts.
 The same format was used during the 2021 and 2022 seasons and for the COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
 Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS). The focus in this section will be on the LWIS (see appendix \ref{app:A_rwis} for RWIS based results).

The WIS is a proper scoring rule used for scoring quantile or interval forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all included quantiles, $B$ is the number of intervals,  $y^*$ is the observed value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a single interval as defined in (\ref{eq:is}). The goal of the forecaster is to minimize the WIS. The LWIS is the same as the WIS except that it is evaluated over the log of quantiles and the log of the observed value.

\begin{equation}
\label{eq:wis2}
        WIS_{0,B}(Q, y^*) = \frac{1}{B + 1/2} \times (w_0\times |y^* - median| + \sum_{b=1}^B \{w_k \times IS_{\alpha_b}(Q, y^*) \} )
\end{equation}

\begin{equation}
\label{eq:is}
        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\mathbbm{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \mathbbm{1}\{y^* > r\}
\end{equation}

We fit 24 forecast models for each location for all 29 weeks, and for each week forecast 1-4 week ahead hospitalizations. The 24 models included all combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST hospitalization models, and both quadratic and linear hospitalization models. 
The prior distributions under the SIR model are in (\ref{eq:sir_prior}). Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$. Here $\mathbbm{1}_{A}$ represents the indicator function for values within the set $A$.

\begin{equation}
\begin{aligned}
    \label{eq:sir_prior}
        I_{0s} &\sim N(0.005, 0.03) \mathbbm{1}_{(0, .1)} \\
        \beta_s &\sim N^+(0.8, 0.3) \\
        \rho_s &\sim N^+(0.68, 0.08)
\end{aligned}
\end{equation}
For the ASG model, the MLE $\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value. The remaining estimates were used as starting values for posterior sampling. The transformation $T(\theta_s) = (\text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ was made, and a prior distribution was assigned to $T(\theta_s)$.
The prior distributions for the ILI model under the ASG function are shown in (\ref{eq:ili_prior}). These are slightly informative priors because for most parameters we have an idea what reasonable values may be. Here $m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
where $\text{diag}(\cdot)$ 
is the diagonal matrix for the given entries.

\begin{equation}
\begin{aligned}
\label{eq:ili_prior}
                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
\end{aligned}
\end{equation}
Parameters shared by both the SIR and ASG models are the scale parameter $\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$. These priors are in (\ref{eq:shared_ili_priors}).

\begin{equation}
\begin{aligned}
    \label{eq:shared_ili_priors}
        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2)
\end{aligned}
\end{equation}
Because of the limited information for estimating $\sigma_{\gamma_W}^2$, we selected an informative prior distribution by first estimating  $\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first calculating the MLE for $\theta_s$. 
% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
$\widehat{ILI}_{s,w}$ was then predicted such that $\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over seasons of the differences $\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was calculated using the \texttt{mle2} function in the \texttt{bblme} package \cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} function.
The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).

\begin{equation}
\begin{aligned}
\label{eq:shared_hosp_priors}
        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
        \phi &\sim N(0, .4)\mathbbm{1}_{(-1,1)} \\
        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
\end{aligned}
\end{equation}

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth,height=6cm]{Images/normal_forecasts_us.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
\end{subfigure}
\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, and the hospitalization models are all normally distribution. The figure includes hospitalization forecasts where ILI is a linear predictor (left) and where ILI is a quadratic predictor (right). 50\% predictive intervals are pink and 95\% predictive intervals are red.}
\label{fig:normal_flu_forecasts}
\end{figure}

Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US flu hospitalizations during the 2023 season under the NORM hospitalization model. The forecasts shown are for weeks 14, 20, 26, and 32. The predictive bands are the 50\% and 95\% predictive intervals.
These plots show a tendency to often underpredict hospitalizations. The quadratic models appear to do a better job predicting  hospitalizations at the season peak but a poorer job predicting after the peak, whereas the linear models seem to predict well or slightly overpredict after the peak. 
% From these figures, it's difficult to say whether ASG or SIR models predict better, but at least for the quadractic models the week 22 predictions are maybe slightly better than for the linear models.

Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the season for all 24 models of US hospitalizations. LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling. Here, a darker value represents a larger LWIS and worse forecast compared to lighter values. The weeks around the holiday week 22 are highlighted by a red band. In most cases, there appears to be a turning point in performance at or near week 22. The models which include discrepancy appear to forecast better around week 22, though they may not outperform the models without discrepancy through the whole season. 
Such is the case for the NORM and LST quadratic models and the linear LNORM models. It's also notable that for the SIR model, including discrepancy appears to make for poorer forecasts after week 22 whereas for ASG, including discrepancy appears to improve forecasts.

\begin{figure}

    % \centering
    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom). Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor. Scores for models with an ASG ILI model are above while those with an SIR model are below. The lighter the shade, the lower the LWIS with low LWIS being better.}
    \label{fig:us_lwis}
\end{figure}

\begin{figure}

    \centering
    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season. Scores are averaged over all horizons 1-4 weeks ahead. Scores are faceted by ILI model function (columns) and by whether or not discrepancy modeling was included (rows). The lighter the shade, the lower the LWIS with low LWIS being better.}
    \label{fig:lwis_by_traj_loc}
\end{figure}

\begin{figure}

    \centering
    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season. Scores are averaged over all horizons 1-4 weeks ahead. Scores are faceted by hospitalization model distribution. The lighter the shade, the lower the LWIS with low LWIS being better.}
    \label{fig:lwis_by_dist_loc}
\end{figure}

Figures \ref{fig:lwis_by_traj_loc} and \ref{fig:lwis_by_dist_loc} also show weekly forecast performance by LWIS, but all 53 locations are included. Overall, the weeks leading up to week 22 are the most difficult to forecast. In figure \ref{fig:lwis_by_traj_loc}, it appears that both the SIR and ASG models which include discrepancy perform slightly better than the models which do not. Figure \ref{fig:lwis_by_dist_loc} shows the LWIS across the whole season faceted by hospitalization model distribution. The LNORM model appears to perform the best during the weeks leading up to week 22. 
The season overall score, calculated as the mean LWIS over all locations and weeks, for all 24 models is shown in table \ref{tab:forecast_scores}. The first main takeaway is that the top four performing models are SIR ILI models with the LNORM hospitalization models. The next main takeaway is that overall the models which include discrepancy outperformed the models without discrepancy.

\begin{table}
\caption{Overall scores for each of the 24 forecast models. The overall score is the log weighted interval score (LWIS) averaged over all locations, weeks, and horizons. The scores in the first two rows are for linear models, and the scores in the third and fourth rows are for quadratic models. The lowest WISs are bolded.}
\begin{tabular*}{\textwidth}
{@{\extracolsep{\fill}} 
    l 
    *{9}{c} 
    S[table-format=4.0]}
  & & \multicolumn{3}{c}{ASG} 
  & \multicolumn{3}{c}{SIR} \\ 
  \cmidrule{3-5} \cmidrule{6-8}
  & & LNORM & LST & NORM & LNORM & LST & NORM\\
  \midrule
  Linear & No Disc & 0.366 & 0.392 & 0.365 & \textbf{0.355} & 0.416 & 0.394 &\\ 
   & Disc & 0.358 & 0.391 & 0.365 & \textbf{0.343} & 0.382 & 0.354 &\\
  \midrule
  Quadratic & No Disc & 0.366 & 0.398 & 0.377 & \textbf{0.355} & 0.401 & 0.377 &\\ 
   & Disc & 0.358 & 0.390 & 0.369 & \textbf{0.343} & 0.378 & 0.356 &\\     
  \bottomrule
\end{tabular*}
\label{tab:forecast_scores}
\end{table}

It should not be assumed that these two takeaways will apply for future flu seasons. As suggested by the simulation study in the previous section, it may often be the case that the ASG ILI model is better for forecasting. Indeed, a close examination of figures \ref{fig:us_lwis}, \ref{fig:lwis_by_traj_loc}, and \ref{fig:lwis_by_dist_loc} shows that it is common for the ASG models to show better forecasting skill than the SIR models.

% Figure \ref{fig:us_lwis_box} shows boxplots of LWIS faceted by hospitalization model, separated by ILI model. The lognormal models tend to have higher LWIS scores than the LST and normal models. There doesn't appear to be a best model by LWIS, though it may be that overall discrepancy does improve forecasts made using the SIR function. Figure \ref{fig:state_lwis_box} is the same as figure \ref{fig:us_lwis_box} except that it shows plots for five states and DC. Interestingly the lognormal distribution for hospitalizations appears to outperform the LST and normal distributions. Again the discrepancy often appears to improve the forecasts and the SIR model. These boxplots also suggest that lognormal hospitalizations models and SIR models with discrepancy provide the best forecasts for Alabama, Montana, and Ohio.

% \begin{figure}

%     \centering
%     \includegraphics[width = 17cm, height = 5cm]{Images/us_lwis_box.png}
%     \caption{}
%     \label{fig:us_lwis_box}
% \end{figure}

% \begin{figure}

%     \centering
%     \includegraphics[scale = .6]{Images/state_lwis_box.png}
%     \caption{}
%     \label{fig:state_lwis_box}
% \end{figure}

% The lognormal and LST models at points during the season often displayed very wide and in some cases extreme predictive intervals. Interestingly the intervals for the LST models were not so wide when a square ILI term was included in the model, though it didn't necessarily lead to improved forecasts. This is seen in figure \ref{fig:t_sq_forecasts_us}.
% Figures \ref{fig:normal_sq_forecasts_us} and \ref{fig:log_sq_forecasts_us} also show forecasts for models excluding the square ILI term but for normal and lognormal distribution models.

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/normal_forecasts_us.png}
%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, and the models are all normal and exclude the square ILI term. 50\% predictive intervals are pink and 95\% predictive intervals are red.}
%     \label{fig:normal_forecasts_us}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/log_forecasts_us.png}
%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, and the models are all lognormal and exclude the square ILI term. 50\% predictive intervals are pink and 95\% predictive intervals are red.}
%     \label{fig:log_forecasts_us}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/t_forecasts_us.png}
%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, and the models are all LST and exclude the square ILI term. 50\% predictive intervals are pink and 95\% predictive intervals are red.}
%     \label{fig:t_forecasts_us}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/normal_sq_forecasts_us.png}
%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, and the models are all normal and include the square ILI term. 50\% predictive intervals are pink and 95\% predictive intervals are red.}
%     \label{fig:normal_sq_forecasts_us}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/log_sq_forecasts_us.png}
%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, and the models are all lognormal and include the square ILI term. 50\% predictive intervals are pink and 95\% predictive intervals are red.}
%     \label{fig:log_sq_forecasts_us}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/t_sq_forecasts_us.png}
%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, and the models are all LST and include the square ILI term. 50\% predictive intervals are pink and 95\% predictive intervals are red.}
%     \label{fig:t_sq_forecasts_us}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/asg_hosp_sq_ar1_rwis.png}
%     \caption{Overall median RWIS by location (black points) for forecasts made by a model including the ASG model without discrepancy, the normal hospitalization model with the square ILI term. The grey line at RWIS = 1 represents the RWIS for the baseline forecast.}
%     \label{fig:asg_sq_ar1_rwis}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/sir_disc_hosp_sq_ar1_rwis.png}
%     \caption{Overall median RWIS by location (black points) for forecasts made by a model including the SIR model with discrepancy, the normal hospitalization model with the square ILI term. The grey line at RWIS = 1 represents the RWIS for the baseline forecast.}
%     \label{fig:sir_disc_sq_rwis}
% \end{figure}

% \subsection{Model Comparison}

% Here the quantile forecasts of our models are compared with those of 10 teams who submitted forecasts during the 2023 flu season. The forecasts selected for comparison were chosen based on similarities shared with our in terms of data and methods used. The FluSight-baseline and FluSight-ensemble models are also included in the comparison. The table below summarizes the information for the selected models and is populated with information given by the forecast teams with their submissions (??).

% \begin{table}[htbp]
%     \centering
%     \begin{tabular}{ |p{3cm}||p{2.8cm}|p{6cm}|p{1.6cm}|  }
%          \hline
%          Model name & Additional Data & Model Summary & Ensemble\\
%          \hline
%          CU-ensemble  & ILI & Ensemble of mechanistic and statistical models& Yes\\
%          \hline
%          LosAlamos\_NAU &  & Bayesian Compartmental & No\\
%          \hline
%          MOBS &  & Metapopulation "SLIR" & No\\
%          \hline
%          UGA\_flucast-Copycat & ILIp, FluServ & Pattern matching/Nearest neighbor & No\\
%          \hline
%          UM-DeepOutbreak & Google symptom search data, ILI & Deep neural network & No\\
%          \hline
%          UMass-flusion & Flu\_Surv, ILI, WHO/NREVSS test & Ensemble of statistical/machine learning time series & Yes\\
%          \hline
%     \end{tabular}
%     \caption{Caption}
%     \label{tab:my_label}
% \end{table}

% The overall forecasts are compared visually using PIT histograms (??) and reliability diagrams (??). These plots are used to assess model calibration, which in probabilistic forecasting relies on the...

\section{Conclusion}
\label{sec:conclusion}

 In this manuscript we introduce a statistical modeling framework which allows for the incorporation of several ILI forecast modeling methods. Specifically, we built upon \cite{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy. We model flu hospitalizations by incorporating the ILI forecast model into a model forecasting hospitalizations where hospitalization predictions are a linear or quadratic function of ILI.

The simulation study in section \ref{sec:simulation2} suggests the ASG function in ILI modeling may slightly outperform the SIR model according to the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season forecasts, the SIR model was overall superior. The results from both the simulation study and the real data analysis suggest that the addition of a discrepancy component in ILI modeling may improve forecasts especially near the holiday week between Christmas and New Years day. It should not be assumed that these conclusions may be generalized for all locations in the US, weeks of a season, or for future flu seasons. 
% Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.

Forecasting the seasonal influenza outbreak remains a challenging task for forecasters. The general modeling framework in this manuscript is successful under diverse modeling conditions for all locations in the US and may contribute to future forecasting efforts. All forecast models were presented as separate from one another, but in the various locations and times of the flu season, different models perform better than others. To build on the work done here, a natural step forward would be to combine all or a few selected forecasts into an ensemble forecast. Such an ensemble may work to cancel out certain model biases or highlight model strengths, leading to more robust forecasts.

\section{Acknowledgments}
This work is partially supported by the National Science Foundation under Grant No. 2152117. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
% \spencer{I noticed that Luis acknowledged HPC@ISU for his work. Should I do the same?}

%%%%%%%%

%%%%
% Reference section comes before the appendix

%\bibliographystyle{acm} % use for numbered citations along with options given in preamble. Look at the main thesis.tex file
\bibliographystyle{apa}

\bibliography{master_bib}

% %\section{Bibliography}
% \bibliographystyle{apa}
% % \vspace{-20pt}
% \begingroup
%     \setlength{\bibsep}{13.2pt}
%     \linespread{1}\selectfont
%     \bibliography{master_bib}
% \endgroup
% \clearpage
% \pagebreak

%%%%%
%% Appendix
% This section may or may not be included
% Chapter 2 shows a double appendix example. Please use A or B as in Appendix A if there are multiple appendix. Use "Appendix A:" before writing the title
% Chapter 3 shows a single appendix example. Use "Appendix:" before writing the title

\begin{appendices}
\section{FluSight forecast competition scoring results} \label{app:A_rwis}

In the CDC flu forecast competition, a baseline forecast was made to which all other forecasts were compared. The baseline forecast for a given state and week had as a median the most recent observed hospitalization count. The uncertainty was based on differences between previous hospitalizations, and is similar to the baseline forecast used in previous flu forecasting seasons and in the COVID-19 hub \cite[]{mathis2024evaluation, cramer2022evaluation}. 
The RWIS for one model was calculated by first taking the ratio of the average LWIS paired with every other model. This was then diveded by the same ratio for the baseline forecasts (see methods in \cite{mathis2024evaluation} for details, and see also the supplementary material).
 An RWIS less than 1 indicates the forecast outperformed the baseline forecast.
 Figures \ref{fig:rwis_ili_tile} and \ref{fig:rwis_hosp_tile} show the mean RWIS of hospitalization forecasts over all 24 models from section \ref{sec:analysis} for all locations and weeks of the 2023 flu season. Some similar patterns to those noted in section \ref{sec:analysis} emerge, including a slightly better performance by forecasts including discrepancy modeling over those which do not include discrepancy modeling and better overall performance by the SIR models than by the ASG forecast models. Interestingly, figure \ref{fig:rwis_hosp_tile} shows the LNORM model showing especially poor RWIS performance about three quarters into the season. This was not seen in the LWIS in figure \ref{fig:lwis_by_dist_loc}.

\begin{figure}
    \centering
    \includegraphics[scale=.7]{Images/rwis_ili_mod.png}
    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks. The plot is separated by the four ILI models. Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
    \label{fig:rwis_ili_tile}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.6]{Images/rwis_hosp_mod.png}
    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks. The plot is separated by the three distribution choices for the hospitalization model. Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
    \label{fig:rwis_hosp_tile}
\end{figure}

Figure \ref{fig:state_median_points} shows the mean over weeks RWIS for the SIR and ASG forecast models with an RWIS less than 1 for the most locations. The plot on the left is from a forecast by an SIRD ILI model and a quadratic NORM hospitalization model. This forecast model had an RWIS less than 1 for 37 out of 53 locations. The plot on the right is from a forecast by an ASGD ILI model and linear NORM hospitalization model. This model had an RWIS less than 1 for 34 out of 53 locations.

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{Images/sir_disc_hosp_sq_ar1_rwis.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{Images/asg_disc_hosp_ar1_rwis.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
\end{subfigure}
\caption{RWIS over the whole 2023 flu season for all 53 locations for hospitalization forecasts from an SIRD quadratic hospitalization model with normal errors (left) and forecasts from an ASGD linear hospitalization model with normal errors (right). An RWIS less than 1 (left of vertical grey line) indicates the model forecasts outperformed the baseline forecasts for that particular region.}
\label{fig:state_median_points}
\end{figure}

\section{Posterior distribution plots for select parameters}
\label{app:B_prior}

The figures in this section show 95\% credible intervals for model parameters under the several modeling schemes along with the prior distributions assigned to the parameters. Figure \ref{fig:posterior_theta_sir} is for parameters unique to the SIR ILI model. Figures \ref{fig:posterior_theta} and \ref{fig:posterior_zeta} are for parameters unique to the ASG models. Figure \ref{fig:sir_asg_shared} is for parameters shared by SIR and ASG models, including parameters used for modeling discrepancy. Figures \ref{fig:hosp_lin_params} and \ref{fig:hosp_nus_param} are for parameters used in hospitalization modeling.

\begin{figure}
    \centering
    \includegraphics[scale=.6]{Images/posterior_theta_sir.png}
    \caption{Posterior 95\% credible intervals from ILI model for parameters of SIR differential equations. Shown are intervals from the US model of ILI for weeks 14, 20, and 26. The green is from the posterior where discrepancy is not modeled, and the blue is from the model where it is. The red interval is the 95\% interval of the prior distribution.}
    \label{fig:posterior_theta_sir}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.6]{Images/posterior_theta.png}
    \caption{Posterior 95\% credible intervals from ILI model for parameters of ASG function. Shown are intervals from the US model of ILI for weeks 14, 20, and 26. The red is from the posterior where discrepancy is not modeled, and the green is from the model where it is. The blue interval is the 95\% interval of the prior distribution.}
    \label{fig:posterior_theta}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=.6]{Images/posterior_zeta.png}
    \caption{Posterior 95\% credible intervals from ILI model for variance parameters of the ASG function. Shown are intervals from the US model of ILI for weeks 14, 20, and 26. The red is from the posterior where discrepancy is not modeled, and the green is from the model where it is. The blue interval is the 95\% interval of the prior distribution.}
    \label{fig:posterior_zeta}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{Images/posterior_kappa.png}
  % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution. Shown are intervals from the US model of ILI for weeks 14, 20, and 26. The blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD. The green interval is the 95\% interval of the prior distribution.}
  % \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{Images/posterior_sigma_gamma.png}
  % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy. Shown are intervals from the US model of ILI for weeks 14, 20, and 26. The blue is from the SIRD model and red from ASGD. The green interval is the 95\% interval of the prior distribution.}
  % \label{fig:sub2}
\end{subfigure}
\caption{Posterior 95\% credible intervals from ILI model for scale parameter $\kappa_s$. Blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD (left).  Posterior 95\% credible intervals from ILI model for scale parameter of modeled discrepancy $\sigma_{\gamma}$ (right). Shown are intervals from the US model of ILI for weeks 14, 20, and 26. Blue is from the SIRD model and red from ASGD. The green interval is the 95\% interval of the prior distribution.}
\label{fig:sir_asg_shared}
\end{figure}

\begin{figure}[hbt!]

\begin{subfigure}{.475\linewidth}
  \includegraphics[width=\linewidth]{Images/alpha0_post.png}
  % \caption{}
  % \label{MLEDdet}
\end{subfigure}\hfill % <-- "\hfill"
\begin{subfigure}{.475\linewidth}
  \includegraphics[width=\linewidth]{Images/alpha1_post.png}
  % \caption{}
  % \label{energydetPSK}
\end{subfigure}

\medskip % create some *vertical* separation between the graphs
\begin{subfigure}{.475\linewidth}
  \includegraphics[width=\linewidth]{Images/alpha2_post.png}
  % \caption{}
  % \label{velcomp}
\end{subfigure}\hfill % <-- "\hfill"
\begin{subfigure}{.475\linewidth}
  \includegraphics[width=\linewidth]{Images/phi_post.png}
  % \caption{}
  % \label{estcomp}
\end{subfigure}

\caption{Posterior 95\% credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top left). Prior distribution 95\% interval is also included.  95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top right). Prior distribution 95\% interval is also included.
95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (bottom left). Prior distribution 95\% interval is also included. 
95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included (bottom right). Prior distribution 95\% interval is also included. In each plot intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
\label{fig:hosp_lin_params}
\end{figure}

\begin{figure}[hbt!]

\begin{subfigure}{.475\linewidth}
  \includegraphics[width=\linewidth]{Images/sigma_epsilon_post.png}
  % \caption{}
  % \label{MLEDdet}
\end{subfigure}\hfill % <-- "\hfill"
\begin{subfigure}{.475\linewidth}
  \includegraphics[width=\linewidth]{Images/omega_post.png}
  % \caption{}
  % \label{energydetPSK}
\end{subfigure}

\caption{Posterior 95\% credible intervals for $\sigma_{\epsilon}$ for hospitalization models with and without the ILI squared term. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (left). Prior distribution 95\% interval is also included.
95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term (right). Prior distribution 95\% interval is also included. In all plots intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
\label{fig:hosp_nus_param}
\end{figure}

% \begin{figure}
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{Images/alpha0_post.png}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{Images/alpha1_post.png}
% \end{subfigure}
% \caption{(left) 95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. Prior distribution 95\% interval is also included. (right) 95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. Prior distribution 95\% interval is also included.}
% \label{fig:hosp_params_post1}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/alpha0_post.png}
%     \caption{95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. 95\% interval for prior distribution is also included}
%     \label{fig:alpha0_post}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.6]{Images/alpha1_post.png}
%     \caption{95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. 95\% interval for prior distribution is also included}
%     \label{fig:alpha1_post}
% \end{figure}

% \begin{figure}
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{Images/alpha2_post.png}
%   % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution. Shown are intervals from the US model of ILI for weeks 14, 20, and 26. The blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD. The green interval is the 95\% interval of the prior distribution.}
%   % \label{fig:sub1}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{Images/phi_post.png}

%   % \label{fig:sub2}
% \end{subfigure}

% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{Images/omega_post.png}

%   % \label{fig:sub2}

% \end{subfigure}
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{Images/sigma_epsilon_post.png}
% \end{subfigure}
% \caption{(top left) 95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. Prior distribution 95\% interval is also included. 
% (top right) 95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. Prior distribution 95\% interval is also included. (bottom) 95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. 95\% interval for prior distribution is also included}
% \label{fig:hosp_params_post2}
% \end{figure}

% \begin{figure}[hbt!]

% \begin{subfigure}{.475\linewidth}
%   \includegraphics[width=\linewidth]{Images/alpha0_post.png}
%   % \caption{}
%   % \label{MLEDdet}
% \end{subfigure}\hfill % <-- "\hfill"
% \begin{subfigure}{.475\linewidth}
%   \includegraphics[width=\linewidth]{Images/alpha1_post.png}
%   % \caption{}
%   % \label{energydetPSK}
% \end{subfigure}

% \medskip % create some *vertical* separation between the graphs
% \begin{subfigure}{.475\linewidth}
%   \includegraphics[width=\linewidth]{Images/alpha2_post.png}
%   % \caption{}
%   % \label{velcomp}
% \end{subfigure}\hfill % <-- "\hfill"
% \begin{subfigure}{.475\linewidth}
%   \includegraphics[width=\linewidth]{Images/phi_post.png}
%   % \caption{}
%   % \label{estcomp}
% \end{subfigure}

% \caption{A figure with four subfigures}
% \label{fig:roc}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.6]{Images/alpha2_post.png}
%     \caption{95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. 95\% interval for prior distribution is also included}
%     \label{fig:alpha2_post}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.6]{Images/phi_post.png}
%     \caption{95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. 95\% interval for prior distribution is also included}
%     \label{fig:phi_post}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.]{Images/omega_post.png}
%     \caption{95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term. Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown. 95\% interval for prior distribution is also included}
%     \label{fig:omega_post}
% \end{figure}

% This section may or may not be included
% \newpage
% \section{Appendix B: Additional tables showing median RWIS for location-model combinations}
% \label{app:B_tables}

% \begin{table}

% % \parbox{.45\linewidth}{
% \raggedleft
% \caption{Model with minimum median RWIS by location}
% \small
% \centering
% \begin{tabular}[t]{rrr}
% \label{tab:wis_by_state}
% \textbf{Location} & \textbf{Model} & \textbf{RWIS}\\

% Alabama & ASGD-NORM & 0.768\\

% Alaska & ASG-NORM-SQ & 0.886\\

% Arizona & ASG-NORM-SQ & 0.669\\

% Arkansas & SIRD-NORM-SQ & 0.734\\

% California & ASGD-NORM-SQ & 0.688\\

% Colorado & SIRD-NORM & 0.619\\

% Connecticut & ASG-NORM & 0.945\\

% Delaware & SIR-NORM-SQ & 0.836\\

% District of Columbia & ASGD-LST-SQ & 0.970\\

% Florida & ASGD-NORM-SQ & 0.680\\

% Georgia & ASG-LNORM & 0.595\\

% Hawaii & SIRD-LST & 0.979\\

% Idaho & ASGD-NORM & 0.895\\

% Illinois & SIRD-LST-SQ & 0.757\\

% Indiana & SIRD-NORM-SQ & 0.696\\

% Iowa & ASGD-NORM & 0.948\\

% Kansas & ASG-NORM-SQ & 0.753\\

% Kentucky & ASG-NORM-SQ & 0.697\\

% Louisiana & SIRD-LST & 0.533\\

% Maine & ASG-NORM-SQ & 0.879\\

% Maryland & SIRD-LNORM & 0.737\\

% Massachusetts & SIRD-NORM-SQ & 0.632\\

% Michigan & SIRD-LST-SQ & 0.733\\

% Minnesota & SIRD-LST-SQ & 0.673\\

% Mississippi & SIR-LNORM-SQ & 0.742\\

% Missouri & ASG-NORM-SQ & 0.797\\

% Montana & SIRD-LNORM-SQ & 0.855\\

% Nebraska & ASGD-NORM-SQ & 0.934\\

% Nevada & SIRD-NORM-SQ & 0.786\\

% New Hampshire & SIR-NORM & 0.766\\

% New Jersey & ASG-LST & 0.811\\

% New Mexico & SIRD-LNORM & 0.768\\

% New York & SIRD-NORM-SQ & 0.765\\

% North Carolina & ASGD-LST & 0.818\\

% North Dakota & ASG-NORM-SQ & 0.838\\

% Ohio & SIRD-NORM & 0.803\\

% Oklahoma & ASGD-NORM-SQ & 0.861\\

% Oregon & SIRD-LST & 0.838\\

% Pennsylvania & SIRD-LNORM & 0.725\\

% Puerto Rico & SIRD-LST & 0.966\\

% Rhode Island & ASG-NORM & 0.869\\

% South Carolina & ASG-LST & 0.670\\

% South Dakota & SIR-LNORM & 0.950\\

% Tennessee & ASG-LST & 0.610\\

% Texas & SIRD-LST-SQ & 0.737\\

% US & ASG-LST-SQ & 0.563\\

% Utah & ASG-LST & 0.702\\

% Vermont & SIRD-NORM-SQ & 0.960\\

% Virginia & ASG-LST & 0.828\\

% Washington & SIRD-NORM & 0.753\\

% West Virginia & SIRD-LNORM & 0.828\\

% Wisconsin & SIRD-LST-SQ & 0.793\\

% Wyoming & SIRD-LNORM & 0.836\\
% \hline
% \end{tabular}
% % \end{table}
% % }
% \end{table}

% \begin{table}
% \hfill
% % \parbox{.45\linewidth}{
% % \begin{table}
% % \raggedright
% \caption{Overall median RWIS and WIS by model}
% \small
% \centering
% \begin{tabular}[t]{c|cc}
% \label{tab:wis_by_model}
% \textbf{Model} & \textbf{RWIS} & \textbf{WIS}\\
% \hline
% SIRD-NORM-SQ & 0.898 & 15.884\\

% ASG-NORM-SQ & 0.903 & 17.152\\

% ASGD-NORM-SQ & 0.922 & 17.408\\

% SIRD-LST-SQ & 0.926 & 16.806\\

% SIRD-NORM & 0.928 & 16.371\\

% ASG-LST-SQ & 0.934 & 18.116\\

% ASGD-LST-SQ & 0.943 & 17.949\\

% SIRD-LST & 0.943 & 17.113\\

% SIRD-LNORM & 0.945 & 20.046\\

% ASG-NORM & 0.946 & 16.762\\

% SIRD-LNORM-SQ & 0.951 & 20.110\\

% ASG-LST & 0.957 & 17.180\\

% ASGD-NORM & 0.966 & 17.208\\

% ASGD-LST & 0.968 & 17.487\\

% ASG-LNORM & 0.968 & 20.628\\

% ASG-LNORM-SQ & 0.969 & 20.620\\

% ASGD-LNORM & 0.982 & 21.108\\

% ASGD-LNORM-SQ & 0.986 & 20.992\\

% FluSight-baseline & 1.000 & 17.585\\

% SIR-NORM-SQ & 1.002 & 18.156\\

% SIR-LNORM & 1.009 & 21.441\\

% SIR-LST-SQ & 1.009 & 18.547\\

% SIR-LNORM-SQ & 1.013 & 21.494\\

% SIR-LST & 1.034 & 19.000\\

% SIR-NORM & 1.067 & 19.904\\
% \hline
% \end{tabular}
% % }
% \end{table}

\end{appendices}

### 2024-11-25 make_images/hosp_prior_pictures.R

@@ -1,0 +1,350 @@
+setwd("~/flu_research/Prelim Content/")
+source("~/flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions.R")
+source("~/flu_research/nonlinear_flu_forecast/forecasts_23/mle_functions.R")
+library(ggplot2)
+library(dplyr)
+library(extraDistr)
+library(tidyr)
+library(ggpubr)
+
+
+
+
+################################################
+###############alphas, sigma####################
+################################################
+
+all_posts <- data.frame()
+for (week in c(14, 20, 26)) {
+  for (mod in c("", "_log", "_lst")) {
+    for (sq in c("", "_sq")) {
+      file_name <- paste0("hosp_fits/asg_disc_hosp", mod, sq, "_ar1_US_",
+                          week, "_posterior.csv")
+      
+      post <- read.csv(file_name) 
+      if (mod %in% c("", "_lst")) {
+        post <- post %>% 
+          select(contains("alpha") & contains(".2."), 
+                 contains("sigma_epsilon") & contains(".2.")) %>% 
+          select(-contains("pha2"))
+        colnames(post) <- c("alpha0", "alpha1", "alpha3", "sigma_epsilon")
+      } else if (mod == "_log") {
+        post <- post %>% 
+          select(contains("alpha"), contains("sigma")) %>% 
+          select(-contains("pha2"))
+      }
+      
+      post$model <- mod
+      post$sq <- sq
+      post$season_week <- week
+      
+      all_posts <- rbind(all_posts, post)
+      
+    }
+  }
+  
+}
+
+
+all_posts <- all_posts %>% 
+  mutate(modwk = paste0(model, season_week))
+
+
+posts_95 <- all_posts %>% 
+  group_by(modwk, model, sq, season_week) %>% 
+  reframe(alpha0_low95 = quantile(alpha0, probs = .025),
+          alpha0_upp95 = quantile(alpha0, probs = .975),
+          alpha1_low95 = quantile(alpha1, probs = .025),
+          alpha1_upp95 = quantile(alpha1, probs = .975),
+          phi_low95 = quantile(alpha3, probs = .025),
+          phi_upp95 = quantile(alpha3, probs = .975),
+          sigma_low95 = quantile(sigma_epsilon, probs = 0),
+          sigma_upp95 = quantile(sigma_epsilon, probs = .95))
+
+posts_95 <- posts_95 %>% 
[…]

### 2024-11-25 make_images/make_analysis_images.R

@@ -1,0 +1,573 @@
+setwd("~/flu_research/nonlinear_flu_forecast/")
+source("../../flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions.R")
+# setwd("./model-output/")
+# source("../target-data/get_target_data.R")
+library(stringr)
+library(epidatr)
+library(evalcast)
+library(dplyr)
+library(tidyr)
+library(scoringutils)
+both_flu <- comb_data(lag = 2)
+date_weeks <- both_flu %>%
+  select(date, season_week) %>% unique
+ILINet <- get_ili_data()
+
+
+setwd("~/flu_research/Prelim Content/models")
+
+#season_week 10 is "2023-10-14"
+
+target_data <- read.csv("../../../forecast-hub/FluSight-forecast-hub/target-data/target-hospital-admissions.csv") %>%
+  mutate(true_value = value) %>%
+  select(-value, -X)
+# setwd("~/forecast-hub/FluSight-forecast-hub/model-output/")
+folders <- list.files("./")
+models <- folders[-which(folders == "FluSight-ensemble")]
+
+sub_dates <- substr(list.files("./FluSight-baseline"), 1, 10)
+sub_date <- sub_dates[1]
+base <- "FluSight-baseline"
+
+baseline_forecasts <- data.frame()
+for (sub_date in sub_dates) {
+  baseline_file <- paste(base, "/", sub_date, "-",
+                         base, ".csv", sep = "")
+  
+  baseline_forecast <- read.csv(baseline_file)
+  baseline_forecasts <- rbind(baseline_forecasts, baseline_forecast)
+}
+
+baseline_wis <- baseline_forecasts %>%
+  filter(horizon %in% 0:3) %>%
+  mutate(target_end_date = as.character(date(target_end_date))) %>%
+  left_join(target_data,
+            by = c("location", "target_end_date" = "date")) %>%
+  mutate(output_type_id = as.numeric(output_type_id),
+         value = as.numeric(value),
+         true_value = as.numeric(true_value)) %>%
+  filter(is.na(location_name) == FALSE) %>%
+  group_by(reference_date, location_name, horizon) %>%
+  summarise(
+    bwis = weighted_interval_score(output_type_id,
+                                   value, true_value),
+    blwis = weighted_interval_score(output_type_id,
+                                    log(value + 1), log(true_value + 1))
+  ) %>%
+  mutate(model = base)
+
+team_scores <- data.frame()
[…]

### 2024-11-25 make_images/make_asg_function_plot.R

@@ -1,0 +1,134 @@
+
+source("~/flu_research/flu_models/mle_functions.R")
+library(ggplot2)
+library(deSolve)
+library(dplyr)
+
+
+asg <- function(x,lambda1,eta,mu,sigma1,sigma2) {
+  # eta <- lambda1 + exp(eta)
+  # eta <- exp(eta)
+  asg <- (lambda1 + (eta - lambda1)*exp(-(x - mu)^2/(2*exp(sigma1)^2)))*(x < mu) +
+    (lambda1 + (eta - lambda1)*exp(-(x - mu)^2/(2*exp(sigma2)^2)))*(x >= mu)
+
+  return(asg)
+}
+
+lambda1 = .2
+eta = .42
+mu = 24
+sigma1 = 1.5 #this will be the log of the standard deviance
+sigma2 = 2.3 #also the log
+
+x <- seq(3, 53, length.out = 1001)
+y <- asg(x, lambda1 = lambda1, eta = eta, mu = mu,
+         sigma1 = sigma1, sigma2 = sigma2)
+
+inf_left <- asg(mu - exp(sigma1), lambda1 = lambda1, eta = eta, mu = mu,
+                sigma1 = sigma1, sigma2 = sigma2)
+
+inf_right <- asg(mu + exp(sigma2), lambda1 = lambda1, eta = eta, mu = mu,
+                 sigma1 = sigma1, sigma2 = sigma2)
+
+plot(y ~ x, type = "l")
+
+data.frame(x, y) %>%
+  ggplot() +
+  geom_line(aes(x = x, y = y), size = 1.4) +
+  xlim(c(0, 53)) +
+  geom_segment(aes(x = mu - exp(sigma1),
+                   xend = mu + exp(sigma2),
+                   y = inf_left, yend = inf_right), linetype = 2,
+               size = 1.2) +
+
+  geom_segment(aes(x = mu, xend = mu, y = lambda1 + .01, yend = eta),
+               linetype = 2,
+               size = 1.2) +
+
+  geom_segment(aes(x = 7, xend = mu, y = eta, yend = eta),
+               linetype = 2,
+               size = 1.2) +
+
+  annotate("text", x = mu + exp(sigma2)/2, y = inf_right,
+           label = expression(sigma[2]), vjust = 1.6,
+           size = 10) +
+  annotate("text", x = mu - exp(sigma1)/2, y = inf_left,
+           label = expression(sigma[1]), vjust = 1.6,
+           size = 10) +
+
[…]

### 2024-11-25 make_images/make_forecast_assesment.R

@@ -1,0 +1,415 @@
+setwd("~/flu_research/Prelim Content/")
+source("../../flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions.R")
+# setwd("./model-output/")
+# source("../target-data/get_target_data.R")
+library(stringr)
+library(epidatr)
+library(evalcast)
+library(dplyr)
+library(tidyr)
+
+both_flu <- comb_data()
+date_weeks <- both_flu %>%
+  select(date, season_week) %>% unique
+ILINet <- get_ili_data()
+
+
+setwd("~/flu_research/Prelim Content/models")
+
+#season_week 10 is "2023-10-14"
+
+target_data <- read.csv("./target-data/target-hospital-admissions.csv") %>%
+  mutate(true_value = value) %>%
+  select(-value, -X)
+# setwd("~/forecast-hub/FluSight-forecast-hub/model-output/")
+folders <- list.files("./")
+models <- folders[folders != "old_models" & folders != "target-data"]
+sub_dates <- substr(list.files("./FluSight-baseline"), 1, 10)
+sub_date <- sub_dates[1]
+base <- "FluSight-baseline"
+
+baseline_forecasts <- data.frame()
+for (sub_date in sub_dates) {
+  baseline_file <- paste(base, "/", sub_date, "-",
+                         base, ".csv", sep = "")
+  
+  baseline_forecast <- read.csv(baseline_file)
+  baseline_forecasts <- rbind(baseline_forecasts, baseline_forecast)
+}
+
+baseline_wis <- baseline_forecasts %>%
+  filter(horizon %in% 0:3) %>%
+  mutate(target_end_date = as.character(date(target_end_date))) %>%
+  left_join(target_data,
+            by = c("location", "target_end_date" = "date")) %>%
+  mutate(output_type_id = as.numeric(output_type_id),
+         value = as.numeric(value),
+         true_value = as.numeric(true_value)) %>%
+  filter(is.na(location_name) == FALSE) %>%
+  group_by(reference_date, location_name, horizon) %>%
+  summarise(
+    bwis = weighted_interval_score(output_type_id,
+                                   value, true_value)
+  ) %>%
+  mutate(model = base)
+
+team_scores <- data.frame()
+for (sub_date in sub_dates) {
+  for (i in 1:length(models)) {
+    forecast_file <- paste(models[i], "/", sub_date, "-",
+                           models[i], ".csv", sep = "")
+    
[…]

### 2024-11-25 make_images/make_forecast_plots.R

@@ -1,0 +1,232 @@
+library(gridExtra)
+setwd("~/flu_research/Prelim Content/models")
+source("~/flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions.R")
+files <- list.files("./")
+
+state <- "US"
+
+
+# ILINet <- get_ili_data()
+# both_flu <- comb_data(lag = 2)
+
+plot_dates <- both_flu %>% ungroup() %>% filter(season_week %in% seq(14, 38, by = 6),
+                                  season == 2023) %>%
+  select(date) %>% unique() %>% unlist()
+
+locations <- both_flu %>% ungroup() %>%
+  select(location, location_name) %>% unique()
+
+
+#everything below can be changed for each image
+#uncomment everything above when first running the progam
+model <- "asg_hosp_ar1"
+models <- c("asg_hosp_ar1", "asg_disc_hosp_ar1", 
+            "sir_hosp_ar1", "sir_disc_hosp_ar1")
+# models <- "FluSight-baseline"
+labels <- c("ASG", "ASGD", "SIR", "SIRD")
+# labels <- "SIRD"
+# models <- c("sir_hosp_log_ar1", "sir_disc_hosp_log_ar1")
+# models <- "FluSight-baseline"
+# labels <- c("SIR", "SIRD")
+# model <- "FluSight-baseline"
+forecast_file1 <- paste(models, "/", plot_dates[1], "-",
+                       model, ".csv", sep = "")
+
+read_model <- function() {
+  read.csv() %>% 
+    mutate(model = model)
+}
+
+# forecast_file1 <- 
+#   do.call(rbind,
+#           lapply(paste0(models, "/", plot_dates[1], "-",
+#                         models, ".csv"), read.csv))
+
+forecast1 <- data.frame()
+for (mod in 1:length(models)) {
+  file_name <- paste0(models[mod], "/", plot_dates[1], "-",
+                 models[mod], ".csv") 
+  file <- read.csv(file_name) %>% 
+    mutate(model = labels[mod])
+  forecast1 <- rbind(forecast1, file)
+}
+
+forecast2 <- data.frame()
+for (mod in 1:length(models)) {
+  file_name <- paste0(models[mod], "/", plot_dates[2], "-",
+                      models[mod], ".csv") 
+  file <- read.csv(file_name) %>% 
+    mutate(model = labels[mod])
+  forecast2 <- rbind(forecast2, file)
+}
+
[…]

### 2024-11-25 make_images/make_images_for_data_section.R

@@ -1,0 +1,575 @@
+
+setwd("~/flu_research/nonlinear_flu_forecast/forecasts_23")
+source('~/flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions_new.R')
+source('~/flu_research/nonlinear_flu_forecast/mle_functions.R')
+
+
+
+both_flu <- comb_data(lag = 2) %>% 
+  mutate(uw_odd = (unweighted_ili/100)/(1 - (unweighted_ili/100)))
+ILINet <- get_ili_data()
+
+state <- "US"
+sample(unique(ILINet$region), 6)
+dat <- get_stan_data(ILINet, both_flu, s_region = "Texas")
+stan_dat <- dat[[1]]
+
+mean_ili <- ILINet %>%
+  filter(season %in% c(2010:2022)) %>%
+  group_by(region, season_week) %>%
+  summarise(
+    mean_ili = mean(unweighted_ili, na.rm = TRUE)
+  ) %>%
+  filter(region %in% c("District of Columbia", "Alabama", "Texas",
+                       "Ohio", "Iowa", "Montana")) %>%
+  mutate(region = factor(region, levels=c("District of Columbia", "Alabama",
+                                         "Texas", "Ohio", "Iowa", "Montana")))
+
+ILINet %>% filter(region == state, season >= 2010) %>% 
+  ggplot() +
+  geom_line(aes(x = season_week, y = unweighted_ili),
+            show.legend = FALSE, size = 1) +
+  facet_wrap(~season) +
+  ylab("ILI %") +
+  xlab("Week") +
+  # geom_vline(aes(xintercept = 22)) +
+  theme_bw() + 
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=15))
+
+both_flu %>% filter(region == "US", season >= 2022) %>% 
+  ggplot() +
+  geom_line(aes(x = season_week, y = value),
+            show.legend = FALSE, size = 1.4) +
+  facet_wrap(~season) +
+  ylab("Hospitalizations") +
+  xlab("Week") +
+  theme_bw() + 
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=15))
+
+ILINet %>%
+  filter(region %in% c("Alabama", "District of Columbia", "Texas",
+                       "Ohio", "Iowa", "Montana"),
+         season %in% c(2010:2023)) %>%
[…]

### 2024-11-25 make_images/make_prior_posterior_plots.R

@@ -1,0 +1,265 @@
+library(extraDistr)
+library(ggplot2)
+post <- readRDS("./flu_research/Prelim Content/US2024-02-03.rds")
+
+
+hist(post$`kappas[20]`)
+hist(post$sigma_gamma)
+hist(post$sigma_gamma_w) #N+(sigmahat, 1)
+hist(post$`alpha2[2]`)
+hist(post$`theta[1]`)
+
+
+#sigma_gamma_w, note the mean value sigma_gamma_hat is not included though should be
+x <- seq(0.0001, 4, length.out = 1001)
+y <- dhnorm(x, sigma=1)
+
+
+post %>% 
+  ggplot() +
+  geom_histogram(aes(x = sigma_gamma_w, y = after_stat(density)),
+                 fill = "lightgrey", colour = "black") +
+  geom_line(data = data.frame(x,y), aes(x = x, y = y), col = "red", size = 1) +
+  ylab("") +
+  xlab(~ sigma[gamma[W]]^2) +
+  theme_bw() + 
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=15))
+
+
+#sigma_gamma
+x <- seq(0.0001, .015, length.out = 1001)
+y <- dhnorm(x, sigma=.02)
+
+post %>% 
+  ggplot() +
+  geom_histogram(aes(x = `sigma_gamma`, y = after_stat(density)),
+                 fill = "lightgrey", colour = "black") +
+  geom_line(data = data.frame(x,y), aes(x = x, y = y), col = "red", size = 1) +
+  ylab("") +
+  xlab(~ sigma[gamma] ^ 2) +
+  theme_bw() + 
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=15))
+
+#kappa for 2023
+x <- seq(0.1, 31000, length.out = 1001)
+y <- dhnorm(x, sigma=10000)
+
+post %>% 
+  ggplot() +
+  geom_histogram(aes(x = `kappas[20]`, y = after_stat(density)),
+                 fill = "lightgrey", colour = "black") +
+  geom_line(data = data.frame(x,y), aes(x = x, y = y), col = "red", size = 1) +
+  ylab("") +
+  xlab(~ kappa[23]) +
+  theme_bw() + 
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=15))
+
+
+
+#beta0 2023
+
[…]

### 2024-11-26 bayesian_analysis_version/vtex-soft-texsupport.isba-ba-97c6706/nonlinear_forecast.tex
Commit with manuscripts

--- 2024-11-25 manuscript/manuscript.tex
@@
-% \section{Abstract}
+% \documentclass[ba,preprint]{imsart}% use this for supplement article
+\documentclass[ba]{imsart}
+
+%% Packages
+\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
+\RequirePackage[numbers]{natbib}
+%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
+\RequirePackage{graphicx}
+
+%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
+\usepackage{booktabs} %for the table in the analysis section
+\usepackage{enumerate} %for the short algorithm
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\pubyear{2024}
+\arxiv{2010.00000}
+\volume{TBA}
+\issue{TBA}
+\firstpage{1}
+\lastpage{1}
+
+\startlocaldefs
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Uncomment next line to change            %%
+%% the type of equation numbering           %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\numberwithin{equation}{section}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Axiom, Claim, Corollary, Hypothesis, %%
+%% Lemma, Theorem, Proposition              %%
+%% use \theoremstyle{plain}                 %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{plain}
+\newtheorem{axiom}{Axiom}
+\newtheorem{claim}[axiom]{Claim}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Assumption, Definition, Example,     %%
+%% Notation, Property, Remark, Fact         %%
+%% use \theoremstyle{definition}            %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem*{example}{Example}
+\newtheorem*{fact}{Fact}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Case use \theoremstyle{remark}       %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{remark}
+\newtheorem{case}{Case}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Please put your definitions here:        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\endlocaldefs
+
+\begin{document}
+
+\begin{frontmatter}
+\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy}
+%\title{A sample article title with some additional note\thanksref{t1}}
+\runtitle{Bayesian nonlinear flu hospitalizations forecasting}
+%\thankstext{T1}{A sample additional note to the title.}
+
+\begin{aug}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only one address is permitted per author. %%
+%% Only division, organization and e-mail is %%
+%% included in the address.                  %%
+%% Additional information can be included in %%
+%% the Acknowledgments section if necessary. %%
+%% ORCID can be inserted by command:         %%
+%% \orcid{0000-0000-0000-0000}               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author[A]{\fnms{Spencer Wadsworth}~\ead[label=e1]{sgw96@iastate.edu}}
+% \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}\orcid{0000-0000-0000-0000}}
+\and
+\author[A]{~\snm{Jarad Niemi}\ead[label=e3]{niemi@iastate.edu}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Addresses                                %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\address[A]{Department of Statistics,
+Iowa State University, Ames, IA 50011}
+
+\runauthor{Wadsworth and Niemi}
+\end{aug}
+
 \begin{abstract}
@@
 Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons.
-% Modeling hospitalization data alone is limited to the relatively small amount of data,
-% but may be supplemented by existing ILI data which has been reported, 
-% published at the state level, and successfully forecast for over a decade. 
 In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data.
@@
 We also found that other modeling decisions such as the exact nonlinear function to use in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
-% We compare hospitalization forecasts from our model to other related models submitted to the CDC forecasting competition.
-
-\keywords{Disease outbreak forecasting \and Bayesian hierarchical modeling \and Probabilistic forecasting \and Model discrepancy}
-
-%\pagebreak %remove if needed
-
-% Please include sections as the paper has, some of the following sections are meant as examples of what can be done, the bibliography should be made as given
-
+\end{abstract}
+
+% \begin{keyword}[class=MSC]
+% \kwd[Primary ]{00X00}
+% \kwd{00X00}
+% \kwd[; secondary ]{00X00}
+% \end{keyword}
+
+\begin{keyword}
+\kwd{Disease outbreak forecasting}
+\kwd{Bayesian hierarchical modeling}
+\kwd{Probabilistic forecasting}
+\kwd{Model discrepancy}
+\end{keyword}
+
+\end{frontmatter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% `\ ' is used here because TeX ignores    %%
+%% spaces after text commands.              %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-% HHS hospitalization data for the state and national level has only been provided since 2021 whereas ILI data has been available since 2010. 
 The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models.
@@
 Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}, with the analysis being done using the conventions of FluSight.
- % including comparing the forecasts to a baseline forecast under the proper scoring rule used for scoring interval forecasts \cite[]{mathis2024evaluation}. 
  Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.
@@
     \centering
-    \includegraphics[scale=.7]{Images/us_ili_seasons.png}
+    \includegraphics[scale=.7]{../../Images/us_ili_seasons.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
@@
 As \cite{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day.
-% The model formulated by \cite{osthus2019dynamic} was largely defined by the week 22 peak and influences the model we introduce in \ref{sec:model}. 
 Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
@@
 The only seasons when there was not a peak at week 22 were season 2022 where the season peak occurred particularly early and season 2020 which was greatly influenced by the COVID-19 pandemic.
-% which would have affected patient visitations as well as reporting.
 
@@
     \centering
-    \includegraphics[scale=.7]{Images/ili_vs_week.png}
+    \includegraphics[scale=.7]{../../Images/ili_vs_week.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
@@
 Similar to the national data, the peak in 2022 came early compared to the peak of 2023.
-% Comparisons similar to those of the national data may be made here. 
-% \jarad{Did you forget to make comparisons?}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.5]{../../Images/us_hospitalizations.png}
     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
@@
     \centering
-    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
+    \includegraphics[scale=.65]{../../Images/hosp_vs_week.png}
     \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
@@
     \centering
-    \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
+    \includegraphics[scale=.6]{../../Images/ili_vs_hosp.png}
     \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
@@
 \end{figure}
-
-% \subsection{Simple hospitalization model}
-% \jarad{This should be the start of the methods section, so it should be a new section not subsection.}
-% The CDC forecast competition between 2013 and 2020 seasons led to a number of advances and effective forecast models of ILI, and compared to weekly hospitalization data there is an abundance of ILI data.
-One contribution of this paper is that data and modeling techniques from forecasting ILI data may be used in forecasting hospitalizations.
-% A relationship between the ILI and hospitalization data exists, and to illustrate this we consider the model in equation (\ref{eq:simple_hosp_reg}).
-The model represents the hospitalizations for all weeks of one season and one state.
-For any state, $H_w$ is the hospitalization count at week $w$ and $ILI_w$ is the ILI value at the same week.
-The model is a squared polynomial regression with an autoregressive component of order 1.
-This is also known as an autoregressive with exogenous variables model (ARX(1)).
-% \jarad{reference}
-
-% \begin{equation}
-%     H_{w} = \alpha_0 + \alpha_1 ILI_{w} + \alpha_2 ILI_{w}^2 + \phi H_{w-1} + \epsilon_{w}; \quad \epsilon_{w} \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \label{eq:simple_hosp_reg}
-% \end{equation}
-% \jarad{This model still doesn't make a lot of sense to me. I feel like we have talked about this before. Why isn't the H series centered around the ILI mean, e.g. 
-% $\phi (H_{w-1} -\alpha_0 - \alpha_1 ILI_w + \alpha_2 ILI_w^2)$.}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.8]{Images/hosp_preds_ili_22.png}
-%     \caption{Weekly hospitalization counts during the 2022 flu season for 5 states and DC (black dots) and predicted weekly hospitalizations from the fit of model (\ref{eq:simple_hosp_reg}) (red lines)}
-%     \label{fig:hosp_preds_ili_22}
-% \end{figure}
-% \jarad{Make the relationship between the data in the previous plot and this plot clear.}
-
-% We fit this model for states individually under the 2022 flu season using the \texttt{lm} function in \texttt{R}.
-% \jarad{It doesn't seem very important tha twe used lm in R.}
-% Figure \ref{fig:hosp_preds_ili_22} shows the predicted hospitalizations from the fit model for several states.
-The black points are the true hospitalization counts and the red lines are the predicted hospitalizations from the fit model.
-% \jarad{Predicted or estimated?}
-% The simple model does a decent job predicting hospitalizations, but forecasting future hospitalizations is more difficult. 
-% \jarad{Why?}
-% This, however, demonstrates that it is reasonable to model the behavior of hospitalizations as a linear function of ILI. 
-% \jarad{You have a quadratic function of ILI not a linear function.}
-% In the following two sections, nonlinear functions used in modeling and the proposed model are described.
 
@@
 The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.
-% In this section, we review the SIR and ASG functions as well as the lack of either model to capture certain behavior.
-We call the systematic deviation between the functional model and the observed behavior \textit{discrepancy}.
 
@@
 The nonlinear function $f_{\theta_s}(w)$ captures the trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the systematic patterns which $f_{\theta_s}(w)$ does not capture.
-
-\begin{equation}
+\begin{equation}
+\begin{aligned}
     \label{eq:ili_model}
-    \begin{flalign*}
-        &ILI_{s,w} \overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
-        &\text{logit}(\pi_{s,w}) = f_{\theta_s}(w) + \gamma_w
-    \end{flalign*}
+    % \begin{flalign*}
+        ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
+        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w
+    % \end{flalign*}
+\end{aligned}
 \end{equation}
@@
 In \cite{ulloa2019} modeling is done hierarchically over seasons.
-% The SIR and ASG functions are defined in the next two subsections.
-% and $\gamma_w$ is modeled as a reverse random walk as in (\ref{eq:rw}).
 
@@
     \centering
-    \includegraphics[scale=.6]{Images/sir_traj.png}
+    \includegraphics[scale=.6]{../../Images/sir_traj.png}
     \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
@@
     \centering
-    \includegraphics[scale=.45]{Images/asg_function.png}
+    \includegraphics[scale=.45]{../../Images/asg_function.png}
     \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
@@
 As noted in section \ref{sec:data}, figures \ref{fig:us_ili} and \ref{fig:ili_vs_week} show a regular peak at week 22 of the flu season.
-% It has been speculated that this drop may be the result of taking holiday from work or school or from the public wanting to avoid medical care around the holidays (??). 
 Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to illustrate the systematic discrepancy from a fitted function.
@@
     \centering
-    \includegraphics[scale=.5]{Images/asg_fits.png}
+    \includegraphics[scale=.5]{../../Images/asg_fits.png}
     \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
@@
     \centering
-    \includegraphics[scale=.45]{Images/discrepancy.png}
+    \includegraphics[scale=.45]{../../Images/discrepancy.png}
     \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
@@
     \label{eq:rw}
-        \gamma_w|\gamma_{w + 1} &\overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
-        \gamma_{W} &\sim N(0,\sigma^2_{\gamma_W})
+        \gamma_w|\gamma_{w + 1} \overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
+        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W})
 \end{equation}
@@
 
-% Below we consider four variations of model (\ref{eq:ili_model}).
-These include where $f_{\theta_s}(w) = ASG_{\theta}(w)$ and where $f_{\theta_s}(w) = \text{logit}(I_{s,w})$, $I_{s,w}$ being the infected compartment of an SIR model.
-For both function scenarios, models are fit where the discrepancy component $\gamma_w$ is included and excluded.
-
 \subsection{Hospitalization model} \label{sec:hospital_model}
@@
 Here $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
-% We further construct a forecast model for hospitalizations as a linear model of the ILI model.
-Similar to \ref{eq:simple_hosp_reg}, we select the model in (\ref{eq:final_hosp}).
-The desired outcome of this model is that the relationship between ILI and hospitalizations may be harnessed, and a future forecast of hospitalizations may be informed by a reasonable forecast for ILI as well as by past hospitalizations.
 \begin{equation}
@@
 
-% Like the ILI model, model (\ref{eq:final_hosp}) was fit via MCMC sampling.
 Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also fit via Bayesian posterior updating.
@@
 
-% \begin{equation}
-%     \label{eq:full_prior}
-%     \begin{align*}
-%     p(\theta, \Sigma, \kappa_s, \sigma_{\gamma}^2, \sigma_{\gamma_W}^2, \Psi) &= p(\theta)p(\Sigma)p(\kappa_s)p(\sigma_{\gamma}^2) p(\sigma_{\gamma_W}^2)p(\Psi), \\
-%     p(\Psi) &= p(\alpha_{0s})p(\alpha_{1s})p(\alpha_{2s})p(\phi)p(\sigma_{\epsilon_s})p(\omega_s)
-%     \end{align*}
-% \end{equation}
-% \jarad{What purpose is this serving?}
-
-% Where the component parameters are shared for both SIR and ASG models, the individual parameter prior distribution assignments when $D_s$ from \ref{eq:final_hosp} is a normal distribution are in (\ref{eq:shared_priors}).
-% \jarad{You cannot start a sentence like this.
-I think you intend for "where" to be a continuation of the equation, but the equation may be anywhere relative to the text, i.e.
-you cannot count on the "where" being immediately after the equation.}
 The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
@@
 
-% Because parameter identifiability is lost when modeling with a discrepancy term \cite[]{brynjarsdottir2014learning}, additional constraints on the prior distributions should be made to improve identifiability.
 Additional prior constraints were made to improve parameter identifiability.
@@
 
-% \begin{equation}
-%     \label{eq:shared_priors}
-%     \begin{align*}
-%         \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
-%         \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-%         \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
-%         \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2) \\
-%         \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2) \\
-%         \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-%         \phi &\sim N(0, .4)\mathbbm{1}_{(-1,1)} \\
-%         \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
-%         \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
-%     \end{align*}
-% \end{equation}
-% \jarad{Generally introduce the distribution families in the methods, but the distributions themselves when performing the data analysis.
-Here I think you just need to say that you are using normals for unrestricted parameters, positive normals for positive parameters and a normal truncated on (-1,1) for the correlation.
-Also mention they are all independent (which I guess was done before.}
-
-% }
-
-% \section{Data Model} \label{sec:model}
-
-% The modeling framework we introduce in this section is two-fold.
-To produce a forecast, we fit a model of only ILI data and fit a second model of hospitalization data with ILI as a linear covariate.
-The two models are fit separately, and a probabilistic forecast of ILI is then used as a predictive covariate for hospitalizations in a second model.
-We chose this type of framework because of the apparent relationship between ILI and hospitalizations, the availability of several flu seasons of ILI data compared to only two seasons of hospitalizations data, and the success of the ILI models introduced by \cite{osthus2019dynamic} and \cite{ulloa2019}.
-In this section we outline the two data models, the Bayesian analysis used for fitting including prior selection, and assess the posterior samples.
-
 \subsection{Parameter estimation and posterior predictive sampling}
@@
 As mentioned, most model parameter prior distributions were intended to be uninformative.
-Appendix \ref{app:B_prior} shows plots of posterior distributions for select parameters from ILI and hospitalization models.
+Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material.
 
@@
     \begin{tabular}{c|cccccc}
-     & NORM & NORM^2 & LNORM & LNORM^2 &  LST &  LST^2 \\
+     & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
      \hline
@@
 \end{table}
-
-%  \begin{tabular}{ccccc}
-% \hline
-% \multicolumn{2}{c}{Set A} %%\\               %% <-- mistake
-% %% \cline{1-2}                               %% <-- mistake
-% &                                            %% <--  addition
-% \multicolumn{2}{c}{Set B} \\\hline           %% <--  Changed
-% Weight& Diameter    & Weight &Diameter      \\
-% \hline
-% \end{tabular}
 
@@
 
-% \subsection{Posterior predictive forecast} \label{posterior_forecast}
-
-% Probability forecasts for state hospitalizations are obtained via the posterior predictive
-
-% During the season, weekly hospitalization data was updated on Wednesdays, and forecast submissions were due that evening.
-ILI data was updated on Fridays leaving a lag of one week between the two sets of data when submissions were made.
-Because of this lag, for the most recent week $w^*$, $H_{s,w^*}$ is fit according to (\ref{eq:final_hosp}) but where $ILI_{s,w^*}$ is replaced by the posterior predictive distribution in (\ref{eq:ili_post}).
-For a given state, \textbf{ILI} is all observed ILI values the last value being $ILI_{s,w^*}$, and $\boldsymbol{\pi}_s = (\pi_{s,1}, ..., \pi_{s,w^*-1})$.
-Note that $w^*$ may also be replaced by any future $w + i$, $i \in (0, 1, ..., W)$.
-
-% \begin{equation}
-%     \label{eq:ili_post}
-%     p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}, 
-%     \textbf{H}) = \int_{\Pi \times \mathcal{K}} p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s) p(\boldsymbol{\pi}_s, \kappa_s | \textbf{ILI}, \textbf{H}) d(\boldsymbol{\pi}_s \times \kappa_s)
-% \end{equation}
-
-% Future hospitalization forecasts come from the posterior predictive distribution.
-We are interested in forecasting $\tilde{H}_{s,w + i}$ where $i \in (1, 2, 3, 4)$ meaning we forecast hospitalizations for 1, 2, 3, and 4-weeks ahead.
-The posterior predictive distribution of $\tilde{H}_{s,w + i}$ is defined in (\ref{eq:forecast_posterior}).
-Here $\Psi = (\beta_{0s}, \beta_{1s}, \beta_{2s}, \phi, \sigma_{\epsilon_s}, \nu_s)$ or the vector of all parameters in (\ref{eq:final_hosp}).
-% %$\Psi = (\boldsymbol{\theta}, \boldsymbol{\beta})$
-
-% \begin{equation}
-%     \label{eq:forecast_posterior}
-%     \begin{align*}
-%         p(\tilde{H}_{s, w+i} |\textbf{H}, \textbf{ILI}) &= \\ \int_{\boldsymbol{\Psi} \times (0,1)} p&(\tilde{H}_{s, w+i} | \Psi) p(\Psi | \textbf{H}, \textbf{ILI}, \widetilde{ILI}_{s,w+i})  p(\widetilde{ILI}_{s,w+i} | \textbf{ILI})d (\Psi \times \widetilde{ILI}_{s,w+i})
-%     \end{align*}
-% \end{equation}
-
-% After calculating (\ref{eq:forecast_posterior}) via MCMC sampling, the 23 quantiles requested by the CDC were calculated using the \texttt{quantile} function in \texttt{R} and truncated at 0.
-As stated, the focus of this paper is on forecasting hospitalizations when ILI is modeled according to the ASG function, but the work is heavily influenced by \cite{osthus2019dynamic}, so we compared forecasting ability to models where ILI is modeled similarly to the model proposed by Osthus.
-
-% \begin{equation}
-%     \label{eq:hosp_model}
-%     \begin{flalign*}
-%         \text{log}(H_w) = \alpha_0 + \alpha_1 \times \left[\text{logit}(I_{s,w})\right] + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \end{flalign*}
-% \end{equation}
-
-% \begin{equation}
-%     \label{eq:hosp_model}
-%     \begin{flalign*}
-%         \text{log}(H_w) = \alpha_0 + \alpha_1 \times \left[\text{log}(I_{s,w}) - \text{log}(1 - I_{s,w})\right] + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \end{flalign*}
-% \end{equation}
-
-% \begin{equation}
-% \label{eq:hosp_mod}
-%     \begin{flalign*}
-%         H_w &= \alpha_0_s + \alpha_1_s (I_{s,w} \times P) + \alpha_2_s (I_{s,w} \times P)^2 \\ &+ \alpha_3_s (I_{s,w} \times P)^3 + \phi_s H_{s, w-1} + \epsilon_{s,w} \\
-%         \epsilon_{s,w} &\overset{iid}{\sim} N(0, \sigma^2_{\epsilon})  
-%     \end{flalign*}
-
-% \end{equation}
-
-% \begin{equation}
-% \label{eq:hosp_mod}
-%     \begin{flalign*}
-%         H_w &= \mu_s + \phi_s( H_{s, w-1} - \mu_s)+ \epsilon_{s,w} \\
-%         \epsilon_{s,w} &\overset{iid}{\sim} N(0, \sigma^2_{\epsilon})  
-%     \end{flalign*}
-
-% \end{equation}
-
-% where $\mu_s = \alpha_0_s + \alpha_1_s (I_{s,w} \times P) + \alpha_2_s (I_{s,w} \times P)^2 + \alpha_3_s (I_{s,w} \times P)^3$.
-
-% Another option for modeling $H_w$ as a function of ILI is seen in
-
-% \begin{equation}
-%     \label{eq:hosp_model_sqrt}
-%     \begin{flalign*}
-%         \sqrt{H_w} = \alpha_0 + \alpha_1 \times I_{s,w} + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \end{flalign*}
-% \end{equation}
-
 \section{Simulation Study} \label{sec:simulation2}
@@
 
-% Predictive coverage was assessed at the 0.5 and 0.95 levels and measures what proportion of forecasts were captured by the corresponding intervals.
 Model comparison was done by calculating the continuous ranked probability score (CRPS) and the logarithmic score (LogS) for each forecast.
@@
 \begin{equation}
-    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \mathbbm{1} (y^* \leq x))^2 dx
+    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \mathbb{1} (y^* \leq x))^2 dx
     \label{eq:crps2}
 \end{equation}
-
-% Having forecast distributions as posterior MCMC draws, the AE is calculated as the absolute difference between the sample median and the observed value.
 
@@
 The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
-
-% The number of replicates was selected by assessing the Monte Carlo coverage of absolute predictive error AE for 150 replicates.
-The AE mean and standard deviation were calculated across season, week, and horizon for each season, week and horizon included in the simulation.
-For each of these levels, we wanted a Monte Carlo confidence width to be less than 25, so we solved for $R$ the following
-
-% \[
-%     25 = 2\times 1.96 * S_{m,s,w,h} / \sqrt{R}
-% \]
-% where $S_{m,s,w,h}$ is the SD for the MC AE for model $m$ at season $s$, week $w$, and horizon $h$. Across all $m, s, w, h$, the largest $R$ value was 1,030.533.
-
-% Tables \ref{tab:horizon_coverage} and \ref{tab:seasonweek_coverage} show the nominal coverage of forecasts for the simulated data.
-For each forecast, 50\% predictive interval was calculated with by finding the $25^{th}$ and $75^{th}$ percentiles and the 95\% predictive interval by the $2.5^{th}$ and $97.5^{th}$ percentiles.
-Table \ref{tab:horizon_coverage} show the coverage for each model averaged over all horizons as well as for horizons of 1-4 weeks ahead.
-Table \ref{tab:seasonweek_coverage} shows coverage averaged over the season weeks 14, 20, 26, 32, and 38 forecasts.
-None of the four models possess the nominal coverage as in each case the coverage is below the 0.5 or 0.95 targeted proportion.
-This is due more to the forecasts' inability to nominally cover the true ILI data and not the simulated hospitalization data.
-
-% \begin{table}
-% \caption{}
-% \label{tab:horizon_coverage}
-% \begin{center}
-%     \begin{tabular}{cccccc|ccccc}
-%      &
-%      \multicolumn{5}{c}{50\%}
-%     & 
-%     \multicolumn{5}{|c}{95\%} \\
-%      Horizon week & Overall & 1 & 2 & 3 & 4 & Overall & 1 & 2 & 3 & 4\\ \hline
-%      % Model & \\ \hline
-%      ASG & \textbf{0.23} & 0.20 & 0.20 & \textbf{0.24} & \textbf{0.27} & 0.56 & 0.62 & \textbf{0.56} & 0.57 & \textbf{0.63} \\
-%      SIR & 0.14 & 0.17 & 0.15 & 0.09 & 0.14 & 0.44 & 0.57 & 0.44 & 0.36 & 0.39\\
-%      ASGD & 0.18 & 0.21 & 0.13 & 0.18 & 0.20 & 0.54 & 0.67 & 0.51 & 0.47 & 0.48\\ 
-%      SIRD & 0.22 & \textbf{0.30} & \textbf{0.24} & 0.20 & 0.16 & \textbf{0.61} & \textbf{0.70} & 0.53 & \textbf{0.62} & 0.60\\ 
-%     \end{tabular}
-%     \end{center}
-% \end{table}
-
-% \begin{table}
-% \caption{}
-% \label{tab:seasonweek_coverage}
-% \begin{center}
-%     \begin{tabular}{cccccc|ccccc}
-%      &
-%      \multicolumn{5}{c}{50\%}
-%     & 
-%     \multicolumn{5}{|c}{95\%} \\
-%      Season week & 14 & 20 & 26 & 32 & 38 & 14 & 20 & 26 & 32 & 38 \\ \hline
-%      % Model & \\ \hline
-%      ASG & \textbf{0.35} & 0.12 & 0.29 & 0.28 & \textbf{0.11} & 0.77 & 0.24 & 0.69 & \textbf{0.77} & \textbf{0.52} \\
-%      SIR & 0.18 & 0.07 & 0.11 & 0.30 & 0.04 & 0.51 & 0.14 & 0.51 & 0.57 & 0.46\\
-%      ASGD & 0.07 & 0.27 & \textbf{0.30} & 0.19 & 0.07 & 0.43 & 0.64 & \textbf{0.89} & 0.56 & 0.17\\ 
-%      SIRD & 0.28 & \textbf{0.30} & 0.16 & \textbf{0.36} & 0.02 & \textbf{0.91} & \textbf{0.78} & 0.38 & 0.57 & 0.42\\ 
-%     \end{tabular}
-%     \end{center}
-% \end{table}
 
@@
   \centering
-  \includegraphics[scale=.35]{Images/overall_crps.png}
+  \includegraphics[scale=.35]{../../Images/overall_crps.png}
     % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
@@
   \centering
-  \includegraphics[scale=.35]{Images/overall_logs.png}
+  \includegraphics[scale=.35]{../../Images/overall_logs.png}
     % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
@@
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/overall_crps.png}
-%     \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-%     \label{fig:crps_overall}
-% \end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/crps_by_season.png}
+\begin{figure}
+    \centering
+    \includegraphics[scale=.5]{../../Images/crps_by_season.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
@@
     \centering
-    \includegraphics[scale=.5]{Images/crps_by_week_horizon.png}
+    \includegraphics[scale=.5]{../../Images/crps_by_week_horizon.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/overall_logs.png}
-%     \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
-%     \label{fig:logs_overall}
-% \end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/logs_by_season.png}
+\begin{figure}
+    \centering
+    \includegraphics[scale=.5]{../../Images/logs_by_season.png}
     \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
@@
     \centering
-    \includegraphics[scale=.5]{Images/logs_by_week_horizon.png}
+    \includegraphics[scale=.5]{../../Images/logs_by_week_horizon.png}
     \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
 
-% This simulation study shows that the predictive coverage of each of the four models considered is well below the nominal coverage based on constructing intervals using quantiles. 
 Figures \ref{fig:logs_by_season} and \ref{fig:logs_by_week_horizon} show boxplots of the LogS for the simulated forecasts.
 Note that in these plots only values of less than 17 are included to improve visualization.
-% Figure \ref{fig:overall_scores} shows less dramatic differences in the LogS variation than the CRPS. There are a few notable differences between what the LogS captures vs. the CRPS.
 The LogS plots show similar results to the CRPS, though there are some differences.
@@
 This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
-% There are also slight differences in evaluation between the CRPS and the LogS. When scoring forecasts the two scores will not always recommend the same model(s). 
-% Which of these or other scoring rules used in judging forecasts should reflect specific problem at hand.
 
@@
  Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS).
-The focus in this section will be on the LWIS (see appendix \ref{app:A_rwis} for RWIS based results).
+The focus in this section will be on the LWIS (see supplementary material for RWIS based results).
 
@@
 \label{eq:is}
-        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\mathbbm{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \mathbbm{1}\{y^* > r\}
+        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\mathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \mathbb{1}\{y^* > r\}
 \end{equation}
@@
 Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$.
-Here $\mathbbm{1}_{A}$ represents the indicator function for values within the set $A$.
+Here $\mathbb{1}_{A}$ represents the indicator function for values within the set $A$.
 
@@
     \label{eq:sir_prior}
-        I_{0s} &\sim N(0.005, 0.03) \mathbbm{1}_{(0, .1)} \\
+        I_{0s} &\sim N(0.005, 0.03) \mathbb{1}_{(0, .1)} \\
         \beta_s &\sim N^+(0.8, 0.3) \\
@@
         \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-        \phi &\sim N(0, .4)\mathbbm{1}_{(-1,1)} \\
+        \phi &\sim N(0, .4)\mathbb{1}_{(-1,1)} \\
         \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
@@
   \centering
-  \includegraphics[width=1\linewidth,height=6cm]{Images/normal_forecasts_us.png}
+  \includegraphics[width=1\linewidth,height=6cm]{../../Images/normal_forecasts_us.png}
   % \caption{A subfigure}
@@
   \centering
-  \includegraphics[width=1\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
+  \includegraphics[width=1\linewidth, height=6cm]{../../Images/normal_sq_forecasts_us.png}
   % \caption{A subfigure}
@@
     % \centering
-    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
+    \includegraphics[scale = .5]{../../Images/lwis_us_full_season.png}
     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
+    \includegraphics[scale = .7]{../../Images/lwis_by_traj_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
+    \includegraphics[scale = .7]{../../Images/lwis_by_dist_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
 {@{\extracolsep{\fill}} 
-    l 
-    *{9}{c} 
+    l*{9}{c} 
     S[table-format=4.0]}
@@
 
-% Figure \ref{fig:us_lwis_box} shows boxplots of LWIS faceted by hospitalization model, separated by ILI model.
-The lognormal models tend to have higher LWIS scores than the LST and normal models.
-There doesn't appear to be a best model by LWIS, though it may be that overall discrepancy does improve forecasts made using the SIR function.
-Figure \ref{fig:state_lwis_box} is the same as figure \ref{fig:us_lwis_box} except that it shows plots for five states and DC.
-Interestingly the lognormal distribution for hospitalizations appears to outperform the LST and normal distributions.
-Again the discrepancy often appears to improve the forecasts and the SIR model.
-These boxplots also suggest that lognormal hospitalizations models and SIR models with discrepancy provide the best forecasts for Alabama, Montana, and Ohio.
-
-% \begin{figure}
-
-%     \centering
-%     \includegraphics[width = 17cm, height = 5cm]{Images/us_lwis_box.png}
-%     \caption{}
-%     \label{fig:us_lwis_box}
-% \end{figure}
-
-% \begin{figure}
-
-%     \centering
-%     \includegraphics[scale = .6]{Images/state_lwis_box.png}
-%     \caption{}
-%     \label{fig:state_lwis_box}
-% \end{figure}
-
-% The lognormal and LST models at points during the season often displayed very wide and in some cases extreme predictive intervals.
-Interestingly the intervals for the LST models were not so wide when a square ILI term was included in the model, though it didn't necessarily lead to improved forecasts.
-This is seen in figure \ref{fig:t_sq_forecasts_us}.
-% Figures \ref{fig:normal_sq_forecasts_us} and \ref{fig:log_sq_forecasts_us} also show forecasts for models excluding the square ILI term but for normal and lognormal distribution models.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/normal_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all normal and exclude the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:normal_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/log_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all lognormal and exclude the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:log_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/t_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all LST and exclude the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:t_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/normal_sq_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all normal and include the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:normal_sq_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/log_sq_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all lognormal and include the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:log_sq_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/t_sq_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all LST and include the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:t_sq_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/asg_hosp_sq_ar1_rwis.png}
-%     \caption{Overall median RWIS by location (black points) for forecasts made by a model including the ASG model without discrepancy, the normal hospitalization model with the square ILI term.
-The grey line at RWIS = 1 represents the RWIS for the baseline forecast.}
-%     \label{fig:asg_sq_ar1_rwis}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/sir_disc_hosp_sq_ar1_rwis.png}
-%     \caption{Overall median RWIS by location (black points) for forecasts made by a model including the SIR model with discrepancy, the normal hospitalization model with the square ILI term.
-The grey line at RWIS = 1 represents the RWIS for the baseline forecast.}
-%     \label{fig:sir_disc_sq_rwis}
-% \end{figure}
-
-% \subsection{Model Comparison}
-
-% Here the quantile forecasts of our models are compared with those of 10 teams who submitted forecasts during the 2023 flu season.
-The forecasts selected for comparison were chosen based on similarities shared with our in terms of data and methods used.
-The FluSight-baseline and FluSight-ensemble models are also included in the comparison.
-The table below summarizes the information for the selected models and is populated with information given by the forecast teams with their submissions (??).
-
-% \begin{table}[htbp]
-%     \centering
-%     \begin{tabular}{ |p{3cm}||p{2.8cm}|p{6cm}|p{1.6cm}|  }
-%          \hline
-%          Model name & Additional Data & Model Summary & Ensemble\\
-%          \hline
-%          CU-ensemble  & ILI & Ensemble of mechanistic and statistical models& Yes\\
-%          \hline
-%          LosAlamos\_NAU &  & Bayesian Compartmental & No\\
-%          \hline
-%          MOBS &  & Metapopulation "SLIR" & No\\
-%          \hline
-%          UGA\_flucast-Copycat & ILIp, FluServ & Pattern matching/Nearest neighbor & No\\
-%          \hline
-%          UM-DeepOutbreak & Google symptom search data, ILI & Deep neural network & No\\
-%          \hline
-%          UMass-flusion & Flu\_Surv, ILI, WHO/NREVSS test & Ensemble of statistical/machine learning time series & Yes\\
-%          \hline
-%     \end{tabular}
-%     \caption{Caption}
-%     \label{tab:my_label}
-% \end{table}
-
-% The overall forecasts are compared visually using PIT histograms (??) and reliability diagrams (??).
-These plots are used to assess model calibration, which in probabilistic forecasting relies on the...
-
 \section{Conclusion}
@@
 
-\section{Acknowledgments}
+\begin{acks}[Acknowledgments]
 This work is partially supported by the National Science Foundation under Grant No.
@@
 Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
-% \spencer{I noticed that Luis acknowledged HPC@ISU for his work. Should I do the same?}
-
-%%%%%%%%
-
-%%%%
-% Reference section comes before the appendix
-
-%\bibliographystyle{acm} % use for numbered citations along with options given in preamble. Look at the main thesis.tex file
-\bibliographystyle{apa}
-
+\end{acks}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Funding information, if any,             %%
+%% should be provided in the                %%
+%% funding section.                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{funding}
+% The first author was supported by NSF Grant DMS-??-??????.
+% 
+% 
+% \end{funding}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Supplementary Material, including data   %%
+%% sets and code, should be provided in     %%
+%% {supplement} environment with title      %%
+%% and short description. It cannot be      %%
+%% available exclusively as external link.  %%
+%% All Supplementary Material must be       %%
+%% available to the reader on Project       %%
+%% Euclid with the published article.       %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{supplement}
+\stitle{Posterior distribution plots for select parameters}
+\sdescription{Plots of posterior distributions for parameters in both ILI and hospitalization models}
+\end{supplement}
+\begin{supplement}
+\stitle{FluSight forecast competition scoring results}
+\sdescription{Flu forecast results with relative weighted interval scoring rule similar to that used in  FluSight}
+\end{supplement}
+
+\bibliographystyle{ba}
 \bibliography{master_bib}
 
-% %\section{Bibliography}
-% \bibliographystyle{apa}
-% % \vspace{-20pt}
-% \begingroup
-%     \setlength{\bibsep}{13.2pt}
-%     \linespread{1}\selectfont
-%     \bibliography{master_bib}
-% \endgroup
-% \clearpage
-% \pagebreak
-
-%%%%%
-%% Appendix
-% This section may or may not be included
-% Chapter 2 shows a double appendix example. Please use A or B as in Appendix A if there are multiple appendix. Use "Appendix A:" before writing the title
-% Chapter 3 shows a single appendix example. Use "Appendix:" before writing the title
-
-\begin{appendices}
-\section{FluSight forecast competition scoring results} \label{app:A_rwis}
-
-In the CDC flu forecast competition, a baseline forecast was made to which all other forecasts were compared.
-The baseline forecast for a given state and week had as a median the most recent observed hospitalization count.
-The uncertainty was based on differences between previous hospitalizations, and is similar to the baseline forecast used in previous flu forecasting seasons and in the COVID-19 hub \cite[]{mathis2024evaluation, cramer2022evaluation}.
-The RWIS for one model was calculated by first taking the ratio of the average LWIS paired with every other model.
-This was then diveded by the same ratio for the baseline forecasts (see methods in \cite{mathis2024evaluation} for details, and see also the supplementary material).
- An RWIS less than 1 indicates the forecast outperformed the baseline forecast.
- Figures \ref{fig:rwis_ili_tile} and \ref{fig:rwis_hosp_tile} show the mean RWIS of hospitalization forecasts over all 24 models from section \ref{sec:analysis} for all locations and weeks of the 2023 flu season.
-Some similar patterns to those noted in section \ref{sec:analysis} emerge, including a slightly better performance by forecasts including discrepancy modeling over those which do not include discrepancy modeling and better overall performance by the SIR models than by the ASG forecast models.
-Interestingly, figure \ref{fig:rwis_hosp_tile} shows the LNORM model showing especially poor RWIS performance about three quarters into the season.
-This was not seen in the LWIS in figure \ref{fig:lwis_by_dist_loc}.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{Images/rwis_ili_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
-The plot is separated by the four ILI models.
-Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_ili_tile}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/rwis_hosp_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
-The plot is separated by the three distribution choices for the hospitalization model.
-Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_hosp_tile}
-\end{figure}
-
-Figure \ref{fig:state_median_points} shows the mean over weeks RWIS for the SIR and ASG forecast models with an RWIS less than 1 for the most locations.
-The plot on the left is from a forecast by an SIRD ILI model and a quadratic NORM hospitalization model.
-This forecast model had an RWIS less than 1 for 37 out of 53 locations.
-The plot on the right is from a forecast by an ASGD ILI model and linear NORM hospitalization model.
-This model had an RWIS less than 1 for 34 out of 53 locations.
-
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{Images/sir_disc_hosp_sq_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{Images/asg_disc_hosp_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{RWIS over the whole 2023 flu season for all 53 locations for hospitalization forecasts from an SIRD quadratic hospitalization model with normal errors (left) and forecasts from an ASGD linear hospitalization model with normal errors (right).
-An RWIS less than 1 (left of vertical grey line) indicates the model forecasts outperformed the baseline forecasts for that particular region.}
-\label{fig:state_median_points}
-\end{figure}
-
-\section{Posterior distribution plots for select parameters}
-\label{app:B_prior}
-
-The figures in this section show 95\% credible intervals for model parameters under the several modeling schemes along with the prior distributions assigned to the parameters.
-Figure \ref{fig:posterior_theta_sir} is for parameters unique to the SIR ILI model.
-Figures \ref{fig:posterior_theta} and \ref{fig:posterior_zeta} are for parameters unique to the ASG models.
-Figure \ref{fig:sir_asg_shared} is for parameters shared by SIR and ASG models, including parameters used for modeling discrepancy.
-Figures \ref{fig:hosp_lin_params} and \ref{fig:hosp_nus_param} are for parameters used in hospitalization modeling.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/posterior_theta_sir.png}
-    \caption{Posterior 95\% credible intervals from ILI model for parameters of SIR differential equations.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The green is from the posterior where discrepancy is not modeled, and the blue is from the model where it is.
-The red interval is the 95\% interval of the prior distribution.}
-    \label{fig:posterior_theta_sir}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/posterior_theta.png}
-    \caption{Posterior 95\% credible intervals from ILI model for parameters of ASG function.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The red is from the posterior where discrepancy is not modeled, and the green is from the model where it is.
-The blue interval is the 95\% interval of the prior distribution.}
-    \label{fig:posterior_theta}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/posterior_zeta.png}
-    \caption{Posterior 95\% credible intervals from ILI model for variance parameters of the ASG function.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The red is from the posterior where discrepancy is not modeled, and the green is from the model where it is.
-The blue interval is the 95\% interval of the prior distribution.}
-    \label{fig:posterior_zeta}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{Images/posterior_kappa.png}
-  % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{Images/posterior_sigma_gamma.png}
-  % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The blue is from the SIRD model and red from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Posterior 95\% credible intervals from ILI model for scale parameter $\kappa_s$.
-Blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD (left).
-Posterior 95\% credible intervals from ILI model for scale parameter of modeled discrepancy $\sigma_{\gamma}$ (right).
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-Blue is from the SIRD model and red from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-\label{fig:sir_asg_shared}
-\end{figure}
-
-\begin{figure}[hbt!]
-
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha0_post.png}
-  % \caption{}
-  % \label{MLEDdet}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha1_post.png}
-  % \caption{}
-  % \label{energydetPSK}
-\end{subfigure}
-
-\medskip % create some *vertical* separation between the graphs
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha2_post.png}
-  % \caption{}
-  % \label{velcomp}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/phi_post.png}
-  % \caption{}
-  % \label{estcomp}
-\end{subfigure}
-
-\caption{Posterior 95\% credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top left).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top right).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (bottom left).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included (bottom right).
-Prior distribution 95\% interval is also included.
-In each plot intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
-\label{fig:hosp_lin_params}
-\end{figure}
-
-\begin{figure}[hbt!]
-
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/sigma_epsilon_post.png}
-  % \caption{}
-  % \label{MLEDdet}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/omega_post.png}
-  % \caption{}
-  % \label{energydetPSK}
-\end{subfigure}
-
-\caption{Posterior 95\% credible intervals for $\sigma_{\epsilon}$ for hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (left).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term (right).
-Prior distribution 95\% interval is also included.
-In all plots intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
-\label{fig:hosp_nus_param}
-\end{figure}
-
-% \begin{figure}
-% \centering
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha0_post.png}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha1_post.png}
-% \end{subfigure}
-% \caption{(left) 95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-(right) 95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.}
-% \label{fig:hosp_params_post1}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/alpha0_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha0_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/alpha1_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha1_post}
-% \end{figure}
-
-% \begin{figure}
-% \centering
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha2_post.png}
-%   % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-%   % \label{fig:sub1}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/phi_post.png}
-
-%   % \label{fig:sub2}
-% \end{subfigure}
-
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/omega_post.png}
-
-%   % \label{fig:sub2}
-
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/sigma_epsilon_post.png}
-% \end{subfigure}
-% \caption{(top left) 95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-% (top right) 95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-(bottom) 95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-% \label{fig:hosp_params_post2}
-% \end{figure}
-
-% \begin{figure}[hbt!]
-
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha0_post.png}
-%   % \caption{}
-%   % \label{MLEDdet}
-% \end{subfigure}\hfill % <-- "\hfill"
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha1_post.png}
-%   % \caption{}
-%   % \label{energydetPSK}
-% \end{subfigure}
-
-% \medskip % create some *vertical* separation between the graphs
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha2_post.png}
-%   % \caption{}
-%   % \label{velcomp}
-% \end{subfigure}\hfill % <-- "\hfill"
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/phi_post.png}
-%   % \caption{}
-%   % \label{estcomp}
-% \end{subfigure}
-
-% \caption{A figure with four subfigures}
-% \label{fig:roc}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/alpha2_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha2_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/phi_post.png}
-%     \caption{95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:phi_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.]{Images/omega_post.png}
-%     \caption{95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:omega_post}
-% \end{figure}
-
-% This section may or may not be included
-% \newpage
-% \section{Appendix B: Additional tables showing median RWIS for location-model combinations}
-% \label{app:B_tables}
-
-% \begin{table}
-
-% % \parbox{.45\linewidth}{
-% \raggedleft
-% \caption{Model with minimum median RWIS by location}
-% \small
-% \centering
-% \begin{tabular}[t]{rrr}
-% \label{tab:wis_by_state}
-% \textbf{Location} & \textbf{Model} & \textbf{RWIS}\\
-
-% Alabama & ASGD-NORM & 0.768\\
-
-% Alaska & ASG-NORM-SQ & 0.886\\
-
-% Arizona & ASG-NORM-SQ & 0.669\\
-
-% Arkansas & SIRD-NORM-SQ & 0.734\\
-
-% California & ASGD-NORM-SQ & 0.688\\
-
-% Colorado & SIRD-NORM & 0.619\\
-
-% Connecticut & ASG-NORM & 0.945\\
-
-% Delaware & SIR-NORM-SQ & 0.836\\
-
-% District of Columbia & ASGD-LST-SQ & 0.970\\
-
-% Florida & ASGD-NORM-SQ & 0.680\\
-
-% Georgia & ASG-LNORM & 0.595\\
-
-% Hawaii & SIRD-LST & 0.979\\
-
-% Idaho & ASGD-NORM & 0.895\\
-
-% Illinois & SIRD-LST-SQ & 0.757\\
-
-% Indiana & SIRD-NORM-SQ & 0.696\\
-
-% Iowa & ASGD-NORM & 0.948\\
-
-% Kansas & ASG-NORM-SQ & 0.753\\
-
-% Kentucky & ASG-NORM-SQ & 0.697\\
-
-% Louisiana & SIRD-LST & 0.533\\
-
-% Maine & ASG-NORM-SQ & 0.879\\
-
-% Maryland & SIRD-LNORM & 0.737\\
-
-% Massachusetts & SIRD-NORM-SQ & 0.632\\
-
-% Michigan & SIRD-LST-SQ & 0.733\\
-
-% Minnesota & SIRD-LST-SQ & 0.673\\
-
-% Mississippi & SIR-LNORM-SQ & 0.742\\
-
-% Missouri & ASG-NORM-SQ & 0.797\\
-
-% Montana & SIRD-LNORM-SQ & 0.855\\
-
-% Nebraska & ASGD-NORM-SQ & 0.934\\
-
-% Nevada & SIRD-NORM-SQ & 0.786\\
-
-% New Hampshire & SIR-NORM & 0.766\\
-
-% New Jersey & ASG-LST & 0.811\\
-
-% New Mexico & SIRD-LNORM & 0.768\\
-
-% New York & SIRD-NORM-SQ & 0.765\\
-
-% North Carolina & ASGD-LST & 0.818\\
-
-% North Dakota & ASG-NORM-SQ & 0.838\\
-
-% Ohio & SIRD-NORM & 0.803\\
-
-% Oklahoma & ASGD-NORM-SQ & 0.861\\
-
-% Oregon & SIRD-LST & 0.838\\
-
-% Pennsylvania & SIRD-LNORM & 0.725\\
-
-% Puerto Rico & SIRD-LST & 0.966\\
-
-% Rhode Island & ASG-NORM & 0.869\\
-
-% South Carolina & ASG-LST & 0.670\\
-
-% South Dakota & SIR-LNORM & 0.950\\
-
-% Tennessee & ASG-LST & 0.610\\
-
-% Texas & SIRD-LST-SQ & 0.737\\
-
-% US & ASG-LST-SQ & 0.563\\
-
-% Utah & ASG-LST & 0.702\\
-
-% Vermont & SIRD-NORM-SQ & 0.960\\
-
-% Virginia & ASG-LST & 0.828\\
-
-% Washington & SIRD-NORM & 0.753\\
-
-% West Virginia & SIRD-LNORM & 0.828\\
-
-% Wisconsin & SIRD-LST-SQ & 0.793\\
-
-% Wyoming & SIRD-LNORM & 0.836\\
-% \hline
-% \end{tabular}
-% % \end{table}
-% % }
-% \end{table}
-
-% \begin{table}
-% \hfill
-% % \parbox{.45\linewidth}{
-% % \begin{table}
-% % \raggedright
-% \caption{Overall median RWIS and WIS by model}
-% \small
-% \centering
-% \begin{tabular}[t]{c|cc}
-% \label{tab:wis_by_model}
-% \textbf{Model} & \textbf{RWIS} & \textbf{WIS}\\
-% \hline
-% SIRD-NORM-SQ & 0.898 & 15.884\\
-
-% ASG-NORM-SQ & 0.903 & 17.152\\
-
-% ASGD-NORM-SQ & 0.922 & 17.408\\
-
-% SIRD-LST-SQ & 0.926 & 16.806\\
-
-% SIRD-NORM & 0.928 & 16.371\\
-
-% ASG-LST-SQ & 0.934 & 18.116\\
-
-% ASGD-LST-SQ & 0.943 & 17.949\\
-
-% SIRD-LST & 0.943 & 17.113\\
-
-% SIRD-LNORM & 0.945 & 20.046\\
-
-% ASG-NORM & 0.946 & 16.762\\
-
-% SIRD-LNORM-SQ & 0.951 & 20.110\\
-
-% ASG-LST & 0.957 & 17.180\\
-
-% ASGD-NORM & 0.966 & 17.208\\
-
-% ASGD-LST & 0.968 & 17.487\\
-
-% ASG-LNORM & 0.968 & 20.628\\
-
-% ASG-LNORM-SQ & 0.969 & 20.620\\
-
-% ASGD-LNORM & 0.982 & 21.108\\
-
-% ASGD-LNORM-SQ & 0.986 & 20.992\\
-
-% FluSight-baseline & 1.000 & 17.585\\
-
-% SIR-NORM-SQ & 1.002 & 18.156\\
-
-% SIR-LNORM & 1.009 & 21.441\\
-
-% SIR-LST-SQ & 1.009 & 18.547\\
-
-% SIR-LNORM-SQ & 1.013 & 21.494\\
-
-% SIR-LST & 1.034 & 19.000\\
-
-% SIR-NORM & 1.067 & 19.904\\
-% \hline
-% \end{tabular}
-% % }
-% \end{table}
-
-\end{appendices}
+\end{document}

### 2024-11-26 manuscript/supplementary/supplementary.tex
Commit with manuscripts

--- 2024-11-25 manuscript/manuscript.tex
@@
-% \section{Abstract}
-\begin{abstract}
-The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread.
-The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods.
-For the first several years of the competition, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalization counts.
-Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons.
-% Modeling hospitalization data alone is limited to the relatively small amount of data,
-% but may be supplemented by existing ILI data which has been reported, 
-% published at the state level, and successfully forecast for over a decade. 
-In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data.
-The first component is for modeling ILI data using a dynamic nonlinear Bayesian hierarchical model.
-The second component is for modeling hospitalizations as a function of ILI.
-For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor.
-In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from \cite{osthus2019dynamic} and a nonlinear Bayesian hierarchical model from \cite{ulloa2019} are compared.
-Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model.
-Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared.
-We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year.
-We also found that other modeling decisions such as the exact nonlinear function to use in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
-% We compare hospitalization forecasts from our model to other related models submitted to the CDC forecasting competition.
+\documentclass{article}
 
-\keywords{Disease outbreak forecasting \and Bayesian hierarchical modeling \and Probabilistic forecasting \and Model discrepancy}
+\usepackage{PRIMEarxiv}
 
-%\pagebreak %remove if needed
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage[linktocpage=true]{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{lipsum}
+\usepackage{fancyhdr}       % header
+\usepackage{graphicx}       % graphics
+\graphicspath{{media/}}     % organize your images and other figures under media/ folder
 
-% Please include sections as the paper has, some of the following sections are meant as examples of what can be done, the bibliography should be made as given
+% Packages added by Spencer
+\usepackage{amsmath,amsthm,amssymb,scrextend}
+\usepackage{appendix}
+\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}
+\usepackage{enumerate}
+\usepackage{csquotes}
+\usepackage{subcaption}
+\usepackage{amssymb}
+% \usepackage{bbm}
+\usepackage[bb=pazo]{mathalpha}
+\usepackage{bm}
 
-\section{Introduction}
+\pagestyle{empty}
 
-Every year the seasonal influenza outbreak burdens the public health system by infecting millions, causing an influx of primary care visits and hospitalizations and leading to between 290,000 and 650,000 deaths worldwide \cite[]{whoflufact2023}.
-\cite{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion.
-Accurate forecasting of infectious diseases can inform public decision making and ease the burden of an outbreak \cite[]{turtle2021accurate, lutz2019applying}.
-There is a growing consensus that disease forecasts should be probabilistic in nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has been shown that reporting forecast uncertainty along with predictions may lead to better decision making \cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.
+\newcommand\Bbbbone{%
+  \ifdefined\mathbbb%
+    \mathbbb{1}%
+  \else%
+    \boldsymbol{\mathbb{1}}%
+  \fi}
 
-% \jarad{Test}
-% \spencer{Test}
+\usepackage{xr}
+\externaldocument{supplementary}
 
-To better inform public decision making regarding the flu epidemic, in 2013 the United States Centers for Disease Control and Prevention (CDC) organized a national flu forecasting competition, also known as FluSight \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}.
-Originally, over a dozen teams of researchers from academic and industry backgrounds participated in FluSight by contributing their own forecast models.
-Besides the 2020 season --or the flu season spanning the fall of 2020 and the winter of 2021-- FluSight has been operated annually and researchers outside the CDC have been invited to participate.
-Initially the target data for forecasts was influenza-like illness (ILI) data.
-ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}.
-The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, \cite{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework.
-An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications.
-As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}.
-The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded.
-The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}.
-Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition.
-Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}.
-This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.
+% \usepackage[pdftex]{graphicx}
+% Standard, old-style thesis
+% \usepackage{isutraditional}   
+% \chaptertitle
+% Old-style, thesis numbering down to subsubsection
+\alternate
+\usepackage{rotating}
+% Bibliography without numbers or labels
+\usepackage{natbib}
+% Use the following line if you want square brackets and numbering system
+%\usepackage[square, numbers]{natbib}
 
-% HHS hospitalization data for the state and national level has only been provided since 2021 whereas ILI data has been available since 2010. 
-The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models.
-The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
-Herein we use ILI models similar to those in \cite{osthus2019dynamic} and \cite{ulloa2019} for ILI forecasting.
-The model of \cite{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework.
-Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by \cite{kennedy2001bayesian}.
-The model in \cite{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data.
-Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.
+% this package is for getting certain math scripts like the 1 needed for an indicator function
+\usepackage{bbm}
 
-In line with the newer FluSight standard of forecasting hospitalizations, we model hospitalizations as a linear function of ILI.
-Thus forecasts produced herein target flu hospitalizations and are a mapping of ILI forecasts to hospitalizations.
-This allows for ILI data from many seasons to be exploited and for ILI forecasts to assist in forecasting hospitalizations, which has fewer seasons of data than ILI.
-Several modeling schemes and their forecasts for the 2023 flu hospitalization season are compared, and it is shown that the modeling decisions produce good forecast results for different states or times during the flu season.
+% using this package to make a better looking list
+\usepackage{enumerate}
+% try this one for a list as well. may be better
+% \usepackage[font=itshape]{quoting}
 
-In section \ref{sec:data} we review the ILI and hospitalization data provided by the CDC and targeted by FluSight. 
-In section \ref{sec:functions} the modeling framework contributed by this manuscript is given.
-In the same section, functions similar to those used by \cite{osthus2019dynamic} and \cite{ulloa2019} are defined.
- These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively.
-Model fitting and implementation are described at the end of the section.
-Section \ref{sec:simulation2} is a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared.
-Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts.
-Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}, with the analysis being done using the conventions of FluSight.
- % including comparing the forecasts to a baseline forecast under the proper scoring rule used for scoring interval forecasts \cite[]{mathis2024evaluation}. 
- Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.
+% package for tables
 
-\section{Flu outbreak data} \label{sec:data}
-In this section we introduce and define ILI and hospitalization data and evaluate the data visually.
-ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season.
-Both of these data were collected at the state, territorial, and national level and were reported at least weekly.
-Overall the data is reported for 53 locations including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), and at the US national level.
-We will refer to forecast targets throughout this manuscript.
-A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a specific location and week during the season.
+\renewcommand{\bibfont}{\setstretch{1}\selectfont}
+\setlength{\bibsep}{13.2pt}
 
-\subsection{Influenza-like illness data}
-The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers.
-Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases.
-An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}.
-Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza.
-To know whether or not a sick patient is infected with influenza would require a laboratory test.
+\usepackage{chapterbib}
+%\renewcommand\bibsection{\section*{\ref name}}
+%\renewcommand\bibsection{\section*}
+\renewcommand{\bibsection}{\section{References}}
 
-In 2013, when FluSight began, the ILI data was the object of the forecasts.
-The data was released publicly at HHS region levels, and forecast teams were asked to provide forecasts of several ILI targets on the regional levels including season onset, 1-4 week ahead ILI levels, and the week of peak ILI activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}.
-Currently, the ILI data is collected by the CDC and published on an online portal for viewing at the national, HHS region, census, and state levels \cite[]{cdc2024fluviewportal}.
-To obtain ILI data, we used the \texttt{R} package \texttt{cdcfluview} which provides functions for downloading the data \cite[]{rudis2021cdcfluview}.
-Weekly ILI data from the national, HHS region, and census levels are available from the 1997 flu season until the current season.
-At the state level, data is available from the 2010 flu season to the current season.
+%\bibliographystyle{apa}  % changed from apa, abbrv, unsrtnat
+%\includeonly{titletoc,chapter1}
+%Optional Package to add PDF bookmarks and hypertext links
+% \usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
+\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
 
-Figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023.
-For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons.
-To better align with the flu behavior, week 1 is set as the first week of August and week 52 or 53 is the last week in July of the following year.
-For example week 1 of the 2013 season corresponds to the first week of August 2013, and week 52 of the same season corresponds to the last week of July 2014.
-This convention is used for the remainder of this manuscript.
+%\usepackage{titletoc}
+% \usepackage{hyperref}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{Images/us_ili_seasons.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
-Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.}
-    \label{fig:us_ili}
-\end{figure}
+% Packages/commands added by Jarad
+\usepackage{xcolor}
+\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
+\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}
 
-Notable from the plots in figure \ref{fig:us_ili} is the regular trajectory of the ILI.
-With the exception of season 2020, the ILI begins low at week 1 and increases as the fall and winter progress until the ILI reaches a peak.
-As spring progresses to summer, the ILI decreases to low values.
-As \cite{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day.
-% The model formulated by \cite{osthus2019dynamic} was largely defined by the week 22 peak and influences the model we introduce in \ref{sec:model}. 
-Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
-\cite[]{ewing2017contact, garza2013effect}.
-The only seasons when there was not a peak at week 22 were season 2022 where the season peak occurred particularly early and season 2020 which was greatly influenced by the COVID-19 pandemic.
-% which would have affected patient visitations as well as reporting.
+%Header
+\pagestyle{fancy}
+\thispagestyle{empty}
+\rhead{ \textit{ }}
 
-Figure \ref{fig:ili_vs_week} shows ILI data from five states and the District of Columbia, locations which received particular attention in \cite{osthus2021multiscale}.
-The plots include the ILI data for all seasons from 2010 to 2023 in grey, and the black line is the per week ILI average over seasons.
-The patterns in the individual states are similar to the national level plots in figure \ref{fig:us_ili} in that the ILI rises in the fall and winter until it peaks and descends as the spring and summer progress.
-For these locations ILI regularly peaks, either locally or globally, at or near week 22.
+% Update your Headers here
+\fancyhead[LO]{Bayesian nonlinear flu hospitalizations forecasting}
+% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}
 
- \begin{figure}
-    \centering
-    \includegraphics[scale=.7]{Images/ili_vs_week.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
-Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.
-     Plots include lines for ILI\% from the 2010 flu season to 2023 (grey) and for the weekly ILI averaged over all seasons (black).}
-    \label{fig:ili_vs_week}
-\end{figure}
+%% Title
+\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy
+%%%% Cite as
+%%%% Update your official citation here when published 
+% \thanks{\textit{\underline{Citation}}: 
+% \textbf{Authors. Title. Pages.... DOI:000000/11111.}} 
+}
 
-\subsection{Hospitalization Data}
+\author{
+  Spencer Wadsworth \\
+  Iowa State University \\
+  Ames, Iowa\\
+  \texttt{sgw96@iastate.edu} \\
+  %% examples of more authors
+   \And
+  Jarad Niemi \\
+  Iowa State University \\
+  Ames, Iowa\\
+  \texttt{niemi@iastate.edu} \\
+  %% \AND
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+}
 
-Hospital admission data, used as the object of FluSight forecasting for the 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported Patient Impact and Hospital Capacity by State Timeseries}.
-Several targets of respiratory illnesses including COVID-19, RSV, and influenza are reported weekly by most hospitals in the US.
-In February 2022 it became mandatory for all hospitals to report the number of COVID-19 and influenza hospitalizations, and since then reporting of hospitalizations has become widespread.
-These data were updated every Wednesday and Friday according to NHSN guidelines \cite[]{healthdata2024covidts}.
+\begin{document}
 
-Figure \ref{fig:us_hosp} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons.
-These plots show similarities to the ILI plots in \ref{fig:us_ili} in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends.
-For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22.
-Figure \ref{fig:hosp_vs_week} shows the 2022 and 2023 weekly hospitalizations for the same states from \ref{fig:ili_vs_week}.
-Similar to the national data, the peak in 2022 came early compared to the peak of 2023.
-% Comparisons similar to those of the national data may be made here. 
-% \jarad{Did you forget to make comparisons?}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
-    \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
-    \label{fig:us_hosp}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
-    \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
-    \label{fig:hosp_vs_week}
-\end{figure}
-
-Comparing figures \ref{fig:ili_vs_week} and \ref{fig:hosp_vs_week} shows that ILI and hospitalizations share the similar pattern of increasing to a peak in the winter and decreasing thereafter.
-Figure \ref{fig:ili_vs_hosp} shows scatter plots with ILI\% on the $x$-axes and hospitalizations on the $y$-axes, revealing a positive somewhat linear relationship between the two variables.
-This relationship motivates the forecast models outlined in the next section.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
-    \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
-    \label{fig:ili_vs_hosp}
-\end{figure}
-
-% \subsection{Simple hospitalization model}
-% \jarad{This should be the start of the methods section, so it should be a new section not subsection.}
-% The CDC forecast competition between 2013 and 2020 seasons led to a number of advances and effective forecast models of ILI, and compared to weekly hospitalization data there is an abundance of ILI data.
-One contribution of this paper is that data and modeling techniques from forecasting ILI data may be used in forecasting hospitalizations.
-% A relationship between the ILI and hospitalization data exists, and to illustrate this we consider the model in equation (\ref{eq:simple_hosp_reg}).
-The model represents the hospitalizations for all weeks of one season and one state.
-For any state, $H_w$ is the hospitalization count at week $w$ and $ILI_w$ is the ILI value at the same week.
-The model is a squared polynomial regression with an autoregressive component of order 1.
-This is also known as an autoregressive with exogenous variables model (ARX(1)).
-% \jarad{reference}
-
-% \begin{equation}
-%     H_{w} = \alpha_0 + \alpha_1 ILI_{w} + \alpha_2 ILI_{w}^2 + \phi H_{w-1} + \epsilon_{w}; \quad \epsilon_{w} \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \label{eq:simple_hosp_reg}
-% \end{equation}
-% \jarad{This model still doesn't make a lot of sense to me. I feel like we have talked about this before. Why isn't the H series centered around the ILI mean, e.g. 
-% $\phi (H_{w-1} -\alpha_0 - \alpha_1 ILI_w + \alpha_2 ILI_w^2)$.}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.8]{Images/hosp_preds_ili_22.png}
-%     \caption{Weekly hospitalization counts during the 2022 flu season for 5 states and DC (black dots) and predicted weekly hospitalizations from the fit of model (\ref{eq:simple_hosp_reg}) (red lines)}
-%     \label{fig:hosp_preds_ili_22}
-% \end{figure}
-% \jarad{Make the relationship between the data in the previous plot and this plot clear.}
-
-% We fit this model for states individually under the 2022 flu season using the \texttt{lm} function in \texttt{R}.
-% \jarad{It doesn't seem very important tha twe used lm in R.}
-% Figure \ref{fig:hosp_preds_ili_22} shows the predicted hospitalizations from the fit model for several states.
-The black points are the true hospitalization counts and the red lines are the predicted hospitalizations from the fit model.
-% \jarad{Predicted or estimated?}
-% The simple model does a decent job predicting hospitalizations, but forecasting future hospitalizations is more difficult. 
-% \jarad{Why?}
-% This, however, demonstrates that it is reasonable to model the behavior of hospitalizations as a linear function of ILI. 
-% \jarad{You have a quadratic function of ILI not a linear function.}
-% In the following two sections, nonlinear functions used in modeling and the proposed model are described.
-
-\section{ILI and hospitalization forecast modeling} \label{sec:functions}
-
-% \jarad{Why not start with the ILI model from Osthus et al?
-It makes a lot more sense to me to start with what has already been done and then expand upon what they have done.
-Later you have a section called ILI Model, but isn't this the ILI model?}
-
-The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows of a similar trajectory for modeling ILI.
-Compartmental models are standard mathematical models used for disease outbreaks.
-One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}.
-\cite{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data.
-The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy.
-In the first part of this section, we present an ILI model similar to the model in \cite{osthus2019dynamic}.
-With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.
-
-With the aim of forecasting hospitalizations, we also introduce a linear model of hospitalizations data with ILI as a predictive covariate.
-To forecast hospitalizations, ILI data is first forecasted and the forecast is then plugged in as a covariate in the hospitalization model, thus producing hospitalization forecasts.
-The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.
-% In this section, we review the SIR and ASG functions as well as the lack of either model to capture certain behavior.
-We call the systematic deviation between the functional model and the observed behavior \textit{discrepancy}.
-
-\subsection{ILI Model} \label{sec:ili_model}
-
-The proposed model for ILI for any location is given in (\ref{eq:ili_model}).
-Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given season.
-The ILI is a proportion, so the Beta random variable is a natural selection for modeling.
-Under the parameterization in of the Beta distribution used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a scale parameter.
-The nonlinear function $f_{\theta_s}(w)$ captures the trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the systematic patterns which $f_{\theta_s}(w)$ does not capture.
-
-\begin{equation}
-    \label{eq:ili_model}
-    \begin{flalign*}
-        &ILI_{s,w} \overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
-        &\text{logit}(\pi_{s,w}) = f_{\theta_s}(w) + \gamma_w
-    \end{flalign*}
-\end{equation}
-In \cite{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}.
-In \cite{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}.
-In \cite{ulloa2019} modeling is done hierarchically over seasons.
-% The SIR and ASG functions are defined in the next two subsections.
-% and $\gamma_w$ is modeled as a reverse random walk as in (\ref{eq:rw}).
-
-\subsection{Susceptible-Infectious-Recovered (SIR) compartmental model} \label{sec:sir_func}
-
-The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by \cite{kermack1927contribution}.
-Since then, compartmental models have become standard for modeling infectious diseases \cite[]{allen2008mathematical}, and many extensions have been made and studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}.
-The SIR mathematical model includes three compartments and assumes that at any time $t>0$ every individual in a closed population belongs to exactly one compartment.
-The three compartments are susceptible ($S$), infectious ($I$), and recovered ($R$), and their interaction over the course of an outbreak is described by the differential equations in (\ref{eq:sir_diff}).
-
-\begin{equation}
-    \label{eq:sir_diff}
-    \frac{dS}{dt} = -\beta SI, \quad \frac{dI}{dt} = \beta S I - \delta I, \quad \frac{dR}{dt} = \delta I
-\end{equation}
-Here $S$, $I$, and $R$ represent the proportion of the population in each compartment such that $S + I + R = 1$ for all $t$.
-The trajectory is determined by the disease transmission rate $\beta > 0$ and the recovery rate $\delta > 0$.
-Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual, and the expected rate of recovery to an immune state for a newly infected person.
-Whether or not a disease outbreak is classified as an epidemic is determined by the initial susceptible population $S_0$, or the susceptible population at time $0$, and the parameter $\rho = \delta/
-\beta$.
-If $S_0/\rho > 1$, the outbreak is considered an epidemic.
-It is non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}.
-Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where the $S_0$ and $\rho$ were selected to match an outbreak that would classify as epidemic.
-In the case where $S_0 \leq \rho$, the trajectory for the $I$ compartment would never be increasing.
-The increase to a peak and subsequent decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is reasonable to model ILI by this compartment.
-Thus in modeling the ILI data, we consider the data to be analogous to the $I$ proportion of the population.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/sir_traj.png}
-    \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
-The three compartments are the susceptible compartment (left), infectious (center), and recovered (right).
-In this example, $S_0/\rho > 1$.}
-    \label{fig:sir_traj}
-\end{figure}
-
-\subsection{Asymmetric Gaussian (ASG) function} \label{sec:asg_func}
-The ASG function is another example of a nonlinear function which can approximate the trajectory of the flu outbreak.
-The ASG was previously used by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used to model vegetation growth and satellite sensor data \cite[]{lewis2020extracting, jonsson2002seasonality, hird2009noise, beck2006improved, atkinson2012inter}.
-The ASG is a modification of the asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized by its rise to a peak and fall from that peak which may not occur at the same rate, as shown in figure \ref{fig:asg_function}.
-The ASG function is denoted as $ASG_\theta(w)$ where $\theta = (\lambda, \nu, \mu, \sigma_1^2, \sigma_2^2)$, $\nu > 0$, $\lambda > 0$, $\mu \in (-\infty, \infty)$, $\sigma_1, \sigma_2 > 0$ and $w \in (1, ..., W)$ is week.
-The function is defined in (\ref{eq:asg_function}).
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{Images/asg_function.png}
-    \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
-    \label{fig:asg_function}
-\end{figure}
-
-\begin{equation}
-    \label{eq:asg_function}
-    ASG_{\theta}(w) = 
-    \begin{cases}
-        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
-        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
-    \end{cases}
-\end{equation}
-For modeling in this manuscript, we use a slightly reparameterized version of the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$.
-This constraint guarantees that the function has a peak greater than $\lambda$.
-
-\begin{equation}
-    \label{eq:asg_function_rep}
-    ASG_{\theta}(w) = 
-    \begin{cases}
-        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
-        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
-    \end{cases}
-\end{equation}
-
-\subsection{Model discrepancy}
-
-The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as \cite{osthus2019dynamic} points out there may be systematic behavior that these or other possible functions may not capture.
-As noted in section \ref{sec:data}, figures \ref{fig:us_ili} and \ref{fig:ili_vs_week} show a regular peak at week 22 of the flu season.
-% It has been speculated that this drop may be the result of taking holiday from work or school or from the public wanting to avoid medical care around the holidays (??). 
-Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to illustrate the systematic discrepancy from a fitted function.
-Figure \ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 to 2022 excluding 2020 with a best fit ASG function plotted over the ILI.
-The fits for each season were made by obtaining the maximum likelihood estimate (MLE) of a model given the ILI data where we assume the ASG function is the mean parameter of a Beta distributed random variable.
-Figure \ref{fig:discrepancy} shows the discrepancy between the model fit and the data for the same seasons.
-The grey lines show the difference between the data and the functions from figure \ref{fig:asg_fits} for each season, and the black line is the average by week over all seasons.
-The lines show that the ASG function typically underpredicts week 22 and overpredicts week 23.
-Perhaps for other weeks, like week 30 for example, there also tends to be systematic behavior not captured by the ASG function.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/asg_fits.png}
-    \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
-    \label{fig:asg_fits}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{Images/discrepancy.png}
-    \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
-    \label{fig:discrepancy}
-\end{figure}
-
-The term $\gamma_w$, where $w$ is the season week, is included in (\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the function.
-Modeling discrepancy has been used in uncertainty analysis of simulators to capture systematic differences between mathematical models and
-reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,arendt2012improving,kennedy2001bayesian}.
-Modeling discrepancy can lead to overfitting, particularly in forecasting scenarios, and may also lead to identifiability issues.
-Thus, care must be taken in setting parameter constraints as well as in the selection of prior distributions \cite[]{osthus2019dynamic,brynjarsdottir2014learning}. 
-Modeling of the discrepancy for ILI was done by \cite{osthus2019dynamic} during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}.
-As in Osthus's model, $\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}).
-\begin{equation}
-    \label{eq:rw}
-        \gamma_w|\gamma_{w + 1} &\overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
-        \gamma_{W} &\sim N(0,\sigma^2_{\gamma_W})
-\end{equation}
-The idea for using the reverse random walk is that there are several previous seasons of ILI data, and assuming the random walk captures systematic behavior, fitting it hierarchically over seasons can assist in predicting future behavior in the current season.
-Reverse random walks have also been used with success in election forecasting and other flu forecasting models \cite[]{osthus2021multiscale, osthus2019dynamic, linzer2013dynamic}.
-The sum to zero constraint $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve identifiability.
-
-\subsection{ILI forecasts}
-Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating.
-Future ILI forecasts are obtained via the posterior predictive distribution where for week $w$, the predictive distribution is obtained by integrating over the parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ as in (\ref{eq:ili_post}) where $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ is the density function of the posterior distribution for the model parameters.
-If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ where $i$ is a positive integer.
-
-\begin{equation}
-    \label{eq:ili_post}
-    p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) = \int \int \int \int p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
-\end{equation}
-% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
-% \spencer{Isn't that captured in $d \pi_s$?}
-
-% Below we consider four variations of model (\ref{eq:ili_model}).
-These include where $f_{\theta_s}(w) = ASG_{\theta}(w)$ and where $f_{\theta_s}(w) = \text{logit}(I_{s,w})$, $I_{s,w}$ being the infected compartment of an SIR model.
-For both function scenarios, models are fit where the discrepancy component $\gamma_w$ is included and excluded.
-
-\subsection{Hospitalization model} \label{sec:hospital_model}
-
-The second component for forecast modeling is the hospitalization model defined in (\ref{eq:final_hosp}).
-This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}.
-Here $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
-% We further construct a forecast model for hospitalizations as a linear model of the ILI model.
-Similar to \ref{eq:simple_hosp_reg}, we select the model in (\ref{eq:final_hosp}).
-The desired outcome of this model is that the relationship between ILI and hospitalizations may be harnessed, and a future forecast of hospitalizations may be informed by a reasonable forecast for ILI as well as by past hospitalizations.
-\begin{equation}
-    \begin{aligned}
-    \label{eq:final_hosp}
-    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
-    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, \omega_s) %N(0, \sigma_{\epsilon}^2)
-    \end{aligned}
-\end{equation}
-This model is for any location, and for fitting purposes $ILI_{s,w}$ is always multiplied by $P$ which is proportional to the population of the state or territory, in this case the total population divided by 50,000.
-This is done as a means of scaling so that the prior distribution assigned to  $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, $\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all states.
-
-% Like the ILI model, model (\ref{eq:final_hosp}) was fit via MCMC sampling.
-Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also fit via Bayesian posterior updating.
-To obtain forecasts for $H_{s, w^* + i}$, the ILI posterior predictive distribution is used along with the posterior distribution for the parameters in (\ref{eq:final_hosp}).
-We considered three scenarios for model (\ref{eq:final_hosp}).
-A model where $D_s$ belongs to the normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, or $H_{s,w} + 1$ is lognormally distributed (LNORM).
-The population value $P$ is excluded from the LNORM model, and we set $\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting.
-In the LNORM model, if the linear parameters are not set the same for seasons 22 and 23, the final variance was more prone to be extreme.
-Besides varying the distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ or there is no quadratic ILI term.
-
-\subsection{Prior selection}
-
-The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in \cite{osthus2019dynamic} and \cite{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem.
-For model (\ref{eq:ili_model}), parameters which are common even when using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, $\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$.
-For the SIR function $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$.
-For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
-
-% \begin{equation}
-%     \label{eq:full_prior}
-%     \begin{align*}
-%     p(\theta, \Sigma, \kappa_s, \sigma_{\gamma}^2, \sigma_{\gamma_W}^2, \Psi) &= p(\theta)p(\Sigma)p(\kappa_s)p(\sigma_{\gamma}^2) p(\sigma_{\gamma_W}^2)p(\Psi), \\
-%     p(\Psi) &= p(\alpha_{0s})p(\alpha_{1s})p(\alpha_{2s})p(\phi)p(\sigma_{\epsilon_s})p(\omega_s)
-%     \end{align*}
-% \end{equation}
-% \jarad{What purpose is this serving?}
-
-% Where the component parameters are shared for both SIR and ASG models, the individual parameter prior distribution assignments when $D_s$ from \ref{eq:final_hosp} is a normal distribution are in (\ref{eq:shared_priors}).
-% \jarad{You cannot start a sentence like this.
-I think you intend for "where" to be a continuation of the equation, but the equation may be anywhere relative to the text, i.e.
-you cannot count on the "where" being immediately after the equation.}
-The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by \cite{gelman2006prior}, the priors were assigned to the variance parameters.
-Univariate parameters were assigned either a normal distribution prior if the support is on $\mathbb{R}$, a half-normal prior if the support is nonnegative, or a truncated-normal prior to match a more specific support.
-Under the ASG model, $\theta_s$ is modeled hierarchically over seasons so that for each season the transformed parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are assigned to $\theta$ and $\Sigma$.
-
-% Because parameter identifiability is lost when modeling with a discrepancy term \cite[]{brynjarsdottir2014learning}, additional constraints on the prior distributions should be made to improve identifiability.
-Additional prior constraints were made to improve parameter identifiability.
-In \cite{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$.
-The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative priors.
-To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the parameters.
-A modular Bayesian approach involves multiple steps of parameter fitting where some parameters may be estimated without priors via maximum likelihood estimation or other means.
-Fitting the rest of the model parameters involves assigning priors and conditioning on the previously fit parameters.
-This has been done in computer modeling to improve identifiability and other issues, though \cite{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}.
-We carried out the modular fit by first estimating the maximum likelihood estimate (MLE) for the parameter $\lambda_s$ for each season in (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
-
-% \spencer{Blue stuff below will be either included further on or will be removed entirely
-
-% \begin{equation}
-%     \label{eq:shared_priors}
-%     \begin{align*}
-%         \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
-%         \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-%         \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
-%         \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2) \\
-%         \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2) \\
-%         \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-%         \phi &\sim N(0, .4)\mathbbm{1}_{(-1,1)} \\
-%         \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
-%         \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
-%     \end{align*}
-% \end{equation}
-% \jarad{Generally introduce the distribution families in the methods, but the distributions themselves when performing the data analysis.
-Here I think you just need to say that you are using normals for unrestricted parameters, positive normals for positive parameters and a normal truncated on (-1,1) for the correlation.
-Also mention they are all independent (which I guess was done before.}
-
-% }
-
-% \section{Data Model} \label{sec:model}
-
-% The modeling framework we introduce in this section is two-fold.
-To produce a forecast, we fit a model of only ILI data and fit a second model of hospitalization data with ILI as a linear covariate.
-The two models are fit separately, and a probabilistic forecast of ILI is then used as a predictive covariate for hospitalizations in a second model.
-We chose this type of framework because of the apparent relationship between ILI and hospitalizations, the availability of several flu seasons of ILI data compared to only two seasons of hospitalizations data, and the success of the ILI models introduced by \cite{osthus2019dynamic} and \cite{ulloa2019}.
-In this section we outline the two data models, the Bayesian analysis used for fitting including prior selection, and assess the posterior samples.
-
-\subsection{Parameter estimation and posterior predictive sampling}
-\label{sec:implementation_posterior}
-
-The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the \cite{stan2024manual} \cite[]{gabry2022stan}.
-Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}.
-The \texttt{cmdstanr} package provides several diagnostic statistics for assessing the sampler.
-As mentioned, most model parameter prior distributions were intended to be uninformative.
-Appendix \ref{app:B_prior} shows plots of posterior distributions for select parameters from ILI and hospitalization models.
-
-We assessed the model fit for four ILI models.
-These included the SIR and ASG models and models with and without discrepancy modeling.
-When discrepancy is included, the models are denoted as SIRD and ASGD.
-These models were fit using US national data from 2010 to 2023 flu seasons, where data from the 2020 season was excluded because of the unique behavior during that season.
-Assessment for hospitalization modeling was done for six different models.
-These include NORM, LNORM, and LST models, and models where a quadratic ILI term is included or excluded.
-To assess posterior sampling convergence, models were fit to data where ILI and hospitalization data up to week 14 of the 2023 season was included.
-Sampling was done with four chains where from each chain 60,000 posterior draws were sampled, and the first 10,000 draws were discarded as a burn-in.
-The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for each parameter.
-Tables \ref{tab:ili_diagnostics} and \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum $ESS$ over all parameters for ILI and hospitalization models respectively.
-Forecast models for all other weeks of the season were fit using one chain of 60,000 draws where the first 10,000 draws were discarded as a burn-in.
-% \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
-For parameters of the ASG models that were prone to cause trouble in posterior sampling, the starting values were set to be the MLEs.
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI models fitted on US data for week 14 of the 2023 season}
-\label{tab:ili_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccc}
-     & ASG & ASGD & SIR & SIRD \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
-     $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
-    \end{tabular}
-    \end{center}
-\end{table}
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six hospitalization models fitted on US data for week 14 of the 2023 season}
-\label{tab:hsop_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccccc}
-     & NORM & NORM^2 & LNORM & LNORM^2 &  LST &  LST^2 \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$\\
-     $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
-    \end{tabular}
-    \end{center}
-\end{table}
-
-%  \begin{tabular}{ccccc}
-% \hline
-% \multicolumn{2}{c}{Set A} %%\\               %% <-- mistake
-% %% \cline{1-2}                               %% <-- mistake
-% &                                            %% <--  addition
-% \multicolumn{2}{c}{Set B} \\\hline           %% <--  Changed
-% Weight& Diameter    & Weight &Diameter      \\
-% \hline
-% \end{tabular}
-
-To obtain forecast distributions of hospitalizations, draws from the posterior predictive distribution from the ILI model were used in conjunction with the posterior distribution of the hospitalizations model.
-When fitting model (\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were saved.
-Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal distributions for the model parameters were saved.
-To obtain forecast distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following steps are repeated $K$ times where $K$ is an integer for the number of desired samples.
-We set $K = 50,000$.
-
-\begin{quote}
-\begin{enumerate}[Step 1:]
-  \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
-  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior distributions
-  \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
-  \begin{itemize}
-    \item[] $\mu_{s,w + i}^* = \alpha_{0s}^* + \alpha_{1s}^* (ILI_{s,w + i}^* \times P) + \alpha_{2s}^* (ILI_{s,w + i}^* \times P)^2 + \phi^* H^*_{s,w + i - 1}$
-  \end{itemize}
-   \item Repeat step 3 for $i \in \{1,2,3,4\}$ to obtain $H^*_{s,(w + 1):(w + 4)}$
-  \item Repeat steps 1-4 $K$ times
-\end{enumerate}
-\end{quote}
-The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast for hospitalizations at week $w + i$.
-For the forecast competition analysis in section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were set to 0 to reflect realistic values of hospitalizations and comply with the FluSight forecasting rules.
-
-% \subsection{Posterior predictive forecast} \label{posterior_forecast}
-
-% Probability forecasts for state hospitalizations are obtained via the posterior predictive
-
-% During the season, weekly hospitalization data was updated on Wednesdays, and forecast submissions were due that evening.
-ILI data was updated on Fridays leaving a lag of one week between the two sets of data when submissions were made.
-Because of this lag, for the most recent week $w^*$, $H_{s,w^*}$ is fit according to (\ref{eq:final_hosp}) but where $ILI_{s,w^*}$ is replaced by the posterior predictive distribution in (\ref{eq:ili_post}).
-For a given state, \textbf{ILI} is all observed ILI values the last value being $ILI_{s,w^*}$, and $\boldsymbol{\pi}_s = (\pi_{s,1}, ..., \pi_{s,w^*-1})$.
-Note that $w^*$ may also be replaced by any future $w + i$, $i \in (0, 1, ..., W)$.
-
-% \begin{equation}
-%     \label{eq:ili_post}
-%     p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}, 
-%     \textbf{H}) = \int_{\Pi \times \mathcal{K}} p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s) p(\boldsymbol{\pi}_s, \kappa_s | \textbf{ILI}, \textbf{H}) d(\boldsymbol{\pi}_s \times \kappa_s)
-% \end{equation}
-
-% Future hospitalization forecasts come from the posterior predictive distribution.
-We are interested in forecasting $\tilde{H}_{s,w + i}$ where $i \in (1, 2, 3, 4)$ meaning we forecast hospitalizations for 1, 2, 3, and 4-weeks ahead.
-The posterior predictive distribution of $\tilde{H}_{s,w + i}$ is defined in (\ref{eq:forecast_posterior}).
-Here $\Psi = (\beta_{0s}, \beta_{1s}, \beta_{2s}, \phi, \sigma_{\epsilon_s}, \nu_s)$ or the vector of all parameters in (\ref{eq:final_hosp}).
-% %$\Psi = (\boldsymbol{\theta}, \boldsymbol{\beta})$
-
-% \begin{equation}
-%     \label{eq:forecast_posterior}
-%     \begin{align*}
-%         p(\tilde{H}_{s, w+i} |\textbf{H}, \textbf{ILI}) &= \\ \int_{\boldsymbol{\Psi} \times (0,1)} p&(\tilde{H}_{s, w+i} | \Psi) p(\Psi | \textbf{H}, \textbf{ILI}, \widetilde{ILI}_{s,w+i})  p(\widetilde{ILI}_{s,w+i} | \textbf{ILI})d (\Psi \times \widetilde{ILI}_{s,w+i})
-%     \end{align*}
-% \end{equation}
-
-% After calculating (\ref{eq:forecast_posterior}) via MCMC sampling, the 23 quantiles requested by the CDC were calculated using the \texttt{quantile} function in \texttt{R} and truncated at 0.
-As stated, the focus of this paper is on forecasting hospitalizations when ILI is modeled according to the ASG function, but the work is heavily influenced by \cite{osthus2019dynamic}, so we compared forecasting ability to models where ILI is modeled similarly to the model proposed by Osthus.
-
-% \begin{equation}
-%     \label{eq:hosp_model}
-%     \begin{flalign*}
-%         \text{log}(H_w) = \alpha_0 + \alpha_1 \times \left[\text{logit}(I_{s,w})\right] + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \end{flalign*}
-% \end{equation}
-
-% \begin{equation}
-%     \label{eq:hosp_model}
-%     \begin{flalign*}
-%         \text{log}(H_w) = \alpha_0 + \alpha_1 \times \left[\text{log}(I_{s,w}) - \text{log}(1 - I_{s,w})\right] + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \end{flalign*}
-% \end{equation}
-
-% \begin{equation}
-% \label{eq:hosp_mod}
-%     \begin{flalign*}
-%         H_w &= \alpha_0_s + \alpha_1_s (I_{s,w} \times P) + \alpha_2_s (I_{s,w} \times P)^2 \\ &+ \alpha_3_s (I_{s,w} \times P)^3 + \phi_s H_{s, w-1} + \epsilon_{s,w} \\
-%         \epsilon_{s,w} &\overset{iid}{\sim} N(0, \sigma^2_{\epsilon})  
-%     \end{flalign*}
-
-% \end{equation}
-
-% \begin{equation}
-% \label{eq:hosp_mod}
-%     \begin{flalign*}
-%         H_w &= \mu_s + \phi_s( H_{s, w-1} - \mu_s)+ \epsilon_{s,w} \\
-%         \epsilon_{s,w} &\overset{iid}{\sim} N(0, \sigma^2_{\epsilon})  
-%     \end{flalign*}
-
-% \end{equation}
-
-% where $\mu_s = \alpha_0_s + \alpha_1_s (I_{s,w} \times P) + \alpha_2_s (I_{s,w} \times P)^2 + \alpha_3_s (I_{s,w} \times P)^3$.
-
-% Another option for modeling $H_w$ as a function of ILI is seen in
-
-% \begin{equation}
-%     \label{eq:hosp_model_sqrt}
-%     \begin{flalign*}
-%         \sqrt{H_w} = \alpha_0 + \alpha_1 \times I_{s,w} + \epsilon_w; \quad \epsilon_w \overset{iid}{\sim} N(0, \sigma_{\epsilon}^2)
-%     \end{flalign*}
-% \end{equation}
-
-\section{Simulation Study} \label{sec:simulation2}
-In this section, we present a simulation study conducted for comparing ILI models and further assessing the hospitalization forecast model. US ILI data is used, and hospitalization data is simulated. 
-A leave-one-season-out (LOSO) approach was combined with a Monte Carlo simulation approach.
-For each replication, we simulated log-hospitalizations for all weeks during seasons 2010, ..., 2022, excluding 2020, using the existing ILI data as a predictive covariate.
-Each season was in turn ``left-out" and treated as if it was the most recent season which we desired to forecast.
-Fitting and forecasting was then done for weeks 14, 20, 26, 32, and 38 of the left out season, giving two weeks that tend to occur as flu cases increase, two as cases decrease, and one that occurs when the cases may be increasing or decreasing.
-Week 20 is a week leading up to the holiday week 22 where ILI typically has a local peak.
-We were particularly interested in how important modeling discrepancy is for forecasting at week 20.
-
-For the simulation of hospitalizations, the parameters $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and $\sigma^2_{\epsilon_s}$ from the hospitalization model in (\ref{eq:final_hosp}) were considered the same across all seasons so that all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$.
-The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization data from the 2022 season.
-For fitting, the hospitalization data was first log-transformed.
-We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
-and assigned the noninformative prior $p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$.
-The marginal posterior distribution $\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ was then the established posterior multivariate normal distribution and $\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior distribution \cite[]{gelman2013bayesian}.
-The posterior means of those parameters were used as the values from which log-hospitalizations were simulated.
-The number of replicates in the simulation was 500.
-
-% Predictive coverage was assessed at the 0.5 and 0.95 levels and measures what proportion of forecasts were captured by the corresponding intervals.
-Model comparison was done by calculating the continuous ranked probability score (CRPS) and the logarithmic score (LogS) for each forecast.
-These scores are both proper scoring rules which evaluate the forecast distribution and density functions respectively.
-Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}.
-Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The CRPS is defined in (\ref{eq:crps2}) and the LogS in (\ref{eq:logs2}).
-Here $F(\cdot)$ is the forecast distribution function, $f(\cdot)$ is the forecast density function, and $y^*$ is the observed targeted value of the forecast.
-The orientation of both scores is negative, meaning the smaller the score the better.
-
-\begin{equation}
-    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \mathbbm{1} (y^* \leq x))^2 dx
-    \label{eq:crps2}
-\end{equation}
-
-% Having forecast distributions as posterior MCMC draws, the AE is calculated as the absolute difference between the sample median and the observed value.
-
-\begin{equation}
-    \text{LogS}(f, y^*) = -\text{log}(f(y^*))
-    \label{eq:logs2}
-\end{equation}
-Both the CRPS and LogS are calculated using the \texttt{scoringRules} package in \texttt{R} \cite[]{jordan2019scoring}.
-To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation.
-The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
-
-% The number of replicates was selected by assessing the Monte Carlo coverage of absolute predictive error AE for 150 replicates.
-The AE mean and standard deviation were calculated across season, week, and horizon for each season, week and horizon included in the simulation.
-For each of these levels, we wanted a Monte Carlo confidence width to be less than 25, so we solved for $R$ the following
-
-% \[
-%     25 = 2\times 1.96 * S_{m,s,w,h} / \sqrt{R}
-% \]
-% where $S_{m,s,w,h}$ is the SD for the MC AE for model $m$ at season $s$, week $w$, and horizon $h$. Across all $m, s, w, h$, the largest $R$ value was 1,030.533.
-
-% Tables \ref{tab:horizon_coverage} and \ref{tab:seasonweek_coverage} show the nominal coverage of forecasts for the simulated data.
-For each forecast, 50\% predictive interval was calculated with by finding the $25^{th}$ and $75^{th}$ percentiles and the 95\% predictive interval by the $2.5^{th}$ and $97.5^{th}$ percentiles.
-Table \ref{tab:horizon_coverage} show the coverage for each model averaged over all horizons as well as for horizons of 1-4 weeks ahead.
-Table \ref{tab:seasonweek_coverage} shows coverage averaged over the season weeks 14, 20, 26, 32, and 38 forecasts.
-None of the four models possess the nominal coverage as in each case the coverage is below the 0.5 or 0.95 targeted proportion.
-This is due more to the forecasts' inability to nominally cover the true ILI data and not the simulated hospitalization data.
-
-% \begin{table}
-% \caption{}
-% \label{tab:horizon_coverage}
-% \begin{center}
-%     \begin{tabular}{cccccc|ccccc}
-%      &
-%      \multicolumn{5}{c}{50\%}
-%     & 
-%     \multicolumn{5}{|c}{95\%} \\
-%      Horizon week & Overall & 1 & 2 & 3 & 4 & Overall & 1 & 2 & 3 & 4\\ \hline
-%      % Model & \\ \hline
-%      ASG & \textbf{0.23} & 0.20 & 0.20 & \textbf{0.24} & \textbf{0.27} & 0.56 & 0.62 & \textbf{0.56} & 0.57 & \textbf{0.63} \\
-%      SIR & 0.14 & 0.17 & 0.15 & 0.09 & 0.14 & 0.44 & 0.57 & 0.44 & 0.36 & 0.39\\
-%      ASGD & 0.18 & 0.21 & 0.13 & 0.18 & 0.20 & 0.54 & 0.67 & 0.51 & 0.47 & 0.48\\ 
-%      SIRD & 0.22 & \textbf{0.30} & \textbf{0.24} & 0.20 & 0.16 & \textbf{0.61} & \textbf{0.70} & 0.53 & \textbf{0.62} & 0.60\\ 
-%     \end{tabular}
-%     \end{center}
-% \end{table}
-
-% \begin{table}
-% \caption{}
-% \label{tab:seasonweek_coverage}
-% \begin{center}
-%     \begin{tabular}{cccccc|ccccc}
-%      &
-%      \multicolumn{5}{c}{50\%}
-%     & 
-%     \multicolumn{5}{|c}{95\%} \\
-%      Season week & 14 & 20 & 26 & 32 & 38 & 14 & 20 & 26 & 32 & 38 \\ \hline
-%      % Model & \\ \hline
-%      ASG & \textbf{0.35} & 0.12 & 0.29 & 0.28 & \textbf{0.11} & 0.77 & 0.24 & 0.69 & \textbf{0.77} & \textbf{0.52} \\
-%      SIR & 0.18 & 0.07 & 0.11 & 0.30 & 0.04 & 0.51 & 0.14 & 0.51 & 0.57 & 0.46\\
-%      ASGD & 0.07 & 0.27 & \textbf{0.30} & 0.19 & 0.07 & 0.43 & 0.64 & \textbf{0.89} & 0.56 & 0.17\\ 
-%      SIRD & 0.28 & \textbf{0.30} & 0.16 & \textbf{0.36} & 0.02 & \textbf{0.91} & \textbf{0.78} & 0.38 & 0.57 & 0.42\\ 
-%     \end{tabular}
-%     \end{center}
-% \end{table}
-
-Figures \ref{fig:overall_scores} - \ref{fig:logs_by_week_horizon} show boxplots of the CRPS and LogS for the four models.
-Figure \ref{fig:overall_scores} shows that the variation of overall CRPS is smallest for the ASG models and larger for the SIR models.
-The median scores for the SIR models also appears slightly higher than for the ASG models.
-When faceted by season in figure \ref{fig:crps_by_season}, the boxplots of the CRPS often show the same pattern for ASG and SIR models but not always.
-For example, the bulk of CRPS values for the SIRD model in 2019 appears to have smaller variation than the other models.
-Figure \ref{fig:crps_by_week_horizon} shows CRPS boxplots faceted by week and horizon.
-Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, have CRPS values with lower medians and lower variation than the two models not accounting for ILI discrepancy.
-This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough.
-
-\begin{figure}
-\centering
-% \begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[scale=.35]{Images/overall_crps.png}
-    % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[scale=.35]{Images/overall_logs.png}
-    % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
-% \end{subfigure}
-\caption{Boxplots of the continuous ranked probability score (CRPS) (left) and logarithmic score (LogS) (right) for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-\label{fig:overall_scores}
-\end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/overall_crps.png}
-%     \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-%     \label{fig:crps_overall}
-% \end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/crps_by_season.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
-    \label{fig:crps_by_season}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/crps_by_week_horizon.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecastss and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
-    \label{fig:crps_by_week_horizon}
-\end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/overall_logs.png}
-%     \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
-%     \label{fig:logs_overall}
-% \end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/logs_by_season.png}
-    \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
-    \label{fig:logs_by_season}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/logs_by_week_horizon.png}
-    \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
-    \label{fig:logs_by_week_horizon}
-\end{figure}
-
-% This simulation study shows that the predictive coverage of each of the four models considered is well below the nominal coverage based on constructing intervals using quantiles. 
-Figures \ref{fig:logs_by_season} and \ref{fig:logs_by_week_horizon} show boxplots of the LogS for the simulated forecasts.
-Note that in these plots only values of less than 17 are included to improve visualization.
-% Figure \ref{fig:overall_scores} shows less dramatic differences in the LogS variation than the CRPS. There are a few notable differences between what the LogS captures vs. the CRPS.
-The LogS plots show similar results to the CRPS, though there are some differences.
-For example, SIRD in figure \ref{fig:logs_by_season} tends to show smaller LogS variation relative to the other three models than is seen by the CRPS of SIRD in figure \ref{fig:crps_by_season}.
-We also note the smaller relative LogS variation than CRPS variation for SIRD when comparing figure \ref{fig:logs_by_week_horizon} with figure \ref{fig:crps_by_week_horizon}, particularly at week 14. 
-Among the four models, ASG tends to have the lowest values in CRPS and LogS though this is not always the case.
-When forecasting weeks 21-23, the two models which model discrepancy in the ILI, ASGD and SIRD, tend to outperform the models not modeling the discrepancy.
-This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
-% There are also slight differences in evaluation between the CRPS and the LogS. When scoring forecasts the two scores will not always recommend the same model(s). 
-% Which of these or other scoring rules used in judging forecasts should reflect specific problem at hand.
-
-\section{Analysis of forecasts for 2023 flu season}
-\label{sec:analysis}
-
-In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations.
-The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$.
-A single forecast is thus comprised of 11 predictive intervals and a median.
-Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, and forecasts were made at the state and national level.
-The first week of forecasting took place during the week of October 7, 2023, and the final week was the week of April 27, 2024 making 29 total weeks of forecasts.
- The same format was used during the 2021 and 2022 seasons and for the COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
-% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
- Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS).
-The focus in this section will be on the LWIS (see appendix \ref{app:A_rwis} for RWIS based results).
-
-The WIS is a proper scoring rule used for scoring quantile or interval forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all included quantiles, $B$ is the number of intervals,  $y^*$ is the observed value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ interval.
-$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a single interval as defined in (\ref{eq:is}).
-The goal of the forecaster is to minimize the WIS.
-The LWIS is the same as the WIS except that it is evaluated over the log of quantiles and the log of the observed value.
-
-\begin{equation}
-\label{eq:wis2}
-        WIS_{0,B}(Q, y^*) = \frac{1}{B + 1/2} \times (w_0\times |y^* - median| + \sum_{b=1}^B \{w_k \times IS_{\alpha_b}(Q, y^*) \} )
-\end{equation}
-
-\begin{equation}
-\label{eq:is}
-        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\mathbbm{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \mathbbm{1}\{y^* > r\}
-\end{equation}
-
-We fit 24 forecast models for each location for all 29 weeks, and for each week forecast 1-4 week ahead hospitalizations.
-The 24 models included all combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST hospitalization models, and both quadratic and linear hospitalization models.
-The prior distributions under the SIR model are in (\ref{eq:sir_prior}).
-Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$.
-Here $\mathbbm{1}_{A}$ represents the indicator function for values within the set $A$.
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:sir_prior}
-        I_{0s} &\sim N(0.005, 0.03) \mathbbm{1}_{(0, .1)} \\
-        \beta_s &\sim N^+(0.8, 0.3) \\
-        \rho_s &\sim N^+(0.68, 0.08)
-\end{aligned}
-\end{equation}
-For the ASG model, the MLE $\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value.
-The remaining estimates were used as starting values for posterior sampling.
-The transformation $T(\theta_s) = (\text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ was made, and a prior distribution was assigned to $T(\theta_s)$.
-The prior distributions for the ILI model under the ASG function are shown in (\ref{eq:ili_prior}).
-These are slightly informative priors because for most parameters we have an idea what reasonable values may be.
-Here $m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
-where $\text{diag}(\cdot)$ 
-is the diagonal matrix for the given entries.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:ili_prior}
-                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
-                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
-                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
-                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
-\end{aligned}
-\end{equation}
-Parameters shared by both the SIR and ASG models are the scale parameter $\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-These priors are in (\ref{eq:shared_ili_priors}).
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:shared_ili_priors}
-        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
-        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2)
-\end{aligned}
-\end{equation}
-Because of the limited information for estimating $\sigma_{\gamma_W}^2$, we selected an informative prior distribution by first estimating  $\hat{\sigma}_{\gamma_W}^2$.
-This was estimated for a given state by first calculating the MLE for $\theta_s$.
-% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
-%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
-$\widehat{ILI}_{s,w}$ was then predicted such that $\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season.
-Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over seasons of the differences $\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$.
-When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was calculated using the \texttt{mle2} function in the \texttt{bblme} package \cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} package \cite[]{soetaert2010desolve} function in \texttt{R}.
-Where the ASG function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} function.
-The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
-
-\begin{equation}
-\begin{aligned}
-\label{eq:shared_hosp_priors}
-        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-        \phi &\sim N(0, .4)\mathbbm{1}_{(-1,1)} \\
-        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
-        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
-\end{aligned}
-\end{equation}
-
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth,height=6cm]{Images/normal_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the hospitalization models are all normally distribution.
-The figure includes hospitalization forecasts where ILI is a linear predictor (left) and where ILI is a quadratic predictor (right).
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-\label{fig:normal_flu_forecasts}
-\end{figure}
-
-Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US flu hospitalizations during the 2023 season under the NORM hospitalization model.
-The forecasts shown are for weeks 14, 20, 26, and 32.
-The predictive bands are the 50\% and 95\% predictive intervals.
-These plots show a tendency to often underpredict hospitalizations.
-The quadratic models appear to do a better job predicting  hospitalizations at the season peak but a poorer job predicting after the peak, whereas the linear models seem to predict well or slightly overpredict after the peak.
-% From these figures, it's difficult to say whether ASG or SIR models predict better, but at least for the quadractic models the week 22 predictions are maybe slightly better than for the linear models.
-
-Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the season for all 24 models of US hospitalizations.
-LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling.
-Here, a darker value represents a larger LWIS and worse forecast compared to lighter values.
-The weeks around the holiday week 22 are highlighted by a red band.
-In most cases, there appears to be a turning point in performance at or near week 22.
-The models which include discrepancy appear to forecast better around week 22, though they may not outperform the models without discrepancy through the whole season.
-Such is the case for the NORM and LST quadratic models and the linear LNORM models.
-It's also notable that for the SIR model, including discrepancy appears to make for poorer forecasts after week 22 whereas for ASG, including discrepancy appears to improve forecasts.
-
-\begin{figure}
-
-    % \centering
-    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
-    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
-Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
-Scores for models with an ASG ILI model are above while those with an SIR model are below.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:us_lwis}
-\end{figure}
-
-\begin{figure}
-
-    \centering
-    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
-Scores are averaged over all horizons 1-4 weeks ahead.
-Scores are faceted by ILI model function (columns) and by whether or not discrepancy modeling was included (rows).
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:lwis_by_traj_loc}
-\end{figure}
-
-\begin{figure}
-
-    \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
-Scores are averaged over all horizons 1-4 weeks ahead.
-Scores are faceted by hospitalization model distribution.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:lwis_by_dist_loc}
-\end{figure}
-
-Figures \ref{fig:lwis_by_traj_loc} and \ref{fig:lwis_by_dist_loc} also show weekly forecast performance by LWIS, but all 53 locations are included.
-Overall, the weeks leading up to week 22 are the most difficult to forecast.
-In figure \ref{fig:lwis_by_traj_loc}, it appears that both the SIR and ASG models which include discrepancy perform slightly better than the models which do not.
-Figure \ref{fig:lwis_by_dist_loc} shows the LWIS across the whole season faceted by hospitalization model distribution.
-The LNORM model appears to perform the best during the weeks leading up to week 22.
-The season overall score, calculated as the mean LWIS over all locations and weeks, for all 24 models is shown in table \ref{tab:forecast_scores}.
-The first main takeaway is that the top four performing models are SIR ILI models with the LNORM hospitalization models.
-The next main takeaway is that overall the models which include discrepancy outperformed the models without discrepancy.
-
-\begin{table}
-\caption{Overall scores for each of the 24 forecast models.
-The overall score is the log weighted interval score (LWIS) averaged over all locations, weeks, and horizons.
-The scores in the first two rows are for linear models, and the scores in the third and fourth rows are for quadratic models.
-The lowest WISs are bolded.}
-\begin{tabular*}{\textwidth}
-{@{\extracolsep{\fill}} 
-    l 
-    *{9}{c} 
-    S[table-format=4.0]}
-  & & \multicolumn{3}{c}{ASG} 
-  & \multicolumn{3}{c}{SIR} \\ 
-  \cmidrule{3-5} \cmidrule{6-8}
-  & & LNORM & LST & NORM & LNORM & LST & NORM\\
-  \midrule
-  Linear & No Disc & 0.366 & 0.392 & 0.365 & \textbf{0.355} & 0.416 & 0.394 &\\ 
-   & Disc & 0.358 & 0.391 & 0.365 & \textbf{0.343} & 0.382 & 0.354 &\\
-  \midrule
-  Quadratic & No Disc & 0.366 & 0.398 & 0.377 & \textbf{0.355} & 0.401 & 0.377 &\\ 
-   & Disc & 0.358 & 0.390 & 0.369 & \textbf{0.343} & 0.378 & 0.356 &\\     
-  \bottomrule
-\end{tabular*}
-\label{tab:forecast_scores}
-\end{table}
-
-It should not be assumed that these two takeaways will apply for future flu seasons.
-As suggested by the simulation study in the previous section, it may often be the case that the ASG ILI model is better for forecasting.
-Indeed, a close examination of figures \ref{fig:us_lwis}, \ref{fig:lwis_by_traj_loc}, and \ref{fig:lwis_by_dist_loc} shows that it is common for the ASG models to show better forecasting skill than the SIR models.
-
-% Figure \ref{fig:us_lwis_box} shows boxplots of LWIS faceted by hospitalization model, separated by ILI model.
-The lognormal models tend to have higher LWIS scores than the LST and normal models.
-There doesn't appear to be a best model by LWIS, though it may be that overall discrepancy does improve forecasts made using the SIR function.
-Figure \ref{fig:state_lwis_box} is the same as figure \ref{fig:us_lwis_box} except that it shows plots for five states and DC.
-Interestingly the lognormal distribution for hospitalizations appears to outperform the LST and normal distributions.
-Again the discrepancy often appears to improve the forecasts and the SIR model.
-These boxplots also suggest that lognormal hospitalizations models and SIR models with discrepancy provide the best forecasts for Alabama, Montana, and Ohio.
-
-% \begin{figure}
-
-%     \centering
-%     \includegraphics[width = 17cm, height = 5cm]{Images/us_lwis_box.png}
-%     \caption{}
-%     \label{fig:us_lwis_box}
-% \end{figure}
-
-% \begin{figure}
-
-%     \centering
-%     \includegraphics[scale = .6]{Images/state_lwis_box.png}
-%     \caption{}
-%     \label{fig:state_lwis_box}
-% \end{figure}
-
-% The lognormal and LST models at points during the season often displayed very wide and in some cases extreme predictive intervals.
-Interestingly the intervals for the LST models were not so wide when a square ILI term was included in the model, though it didn't necessarily lead to improved forecasts.
-This is seen in figure \ref{fig:t_sq_forecasts_us}.
-% Figures \ref{fig:normal_sq_forecasts_us} and \ref{fig:log_sq_forecasts_us} also show forecasts for models excluding the square ILI term but for normal and lognormal distribution models.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/normal_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all normal and exclude the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:normal_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/log_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all lognormal and exclude the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:log_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/t_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all LST and exclude the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:t_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/normal_sq_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all normal and include the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:normal_sq_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/log_sq_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all lognormal and include the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:log_sq_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/t_sq_forecasts_us.png}
-%     \caption{1-4 week ahead forecasts for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the models are all LST and include the square ILI term.
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-%     \label{fig:t_sq_forecasts_us}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/asg_hosp_sq_ar1_rwis.png}
-%     \caption{Overall median RWIS by location (black points) for forecasts made by a model including the ASG model without discrepancy, the normal hospitalization model with the square ILI term.
-The grey line at RWIS = 1 represents the RWIS for the baseline forecast.}
-%     \label{fig:asg_sq_ar1_rwis}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/sir_disc_hosp_sq_ar1_rwis.png}
-%     \caption{Overall median RWIS by location (black points) for forecasts made by a model including the SIR model with discrepancy, the normal hospitalization model with the square ILI term.
-The grey line at RWIS = 1 represents the RWIS for the baseline forecast.}
-%     \label{fig:sir_disc_sq_rwis}
-% \end{figure}
-
-% \subsection{Model Comparison}
-
-% Here the quantile forecasts of our models are compared with those of 10 teams who submitted forecasts during the 2023 flu season.
-The forecasts selected for comparison were chosen based on similarities shared with our in terms of data and methods used.
-The FluSight-baseline and FluSight-ensemble models are also included in the comparison.
-The table below summarizes the information for the selected models and is populated with information given by the forecast teams with their submissions (??).
-
-% \begin{table}[htbp]
-%     \centering
-%     \begin{tabular}{ |p{3cm}||p{2.8cm}|p{6cm}|p{1.6cm}|  }
-%          \hline
-%          Model name & Additional Data & Model Summary & Ensemble\\
-%          \hline
-%          CU-ensemble  & ILI & Ensemble of mechanistic and statistical models& Yes\\
-%          \hline
-%          LosAlamos\_NAU &  & Bayesian Compartmental & No\\
-%          \hline
-%          MOBS &  & Metapopulation "SLIR" & No\\
-%          \hline
-%          UGA\_flucast-Copycat & ILIp, FluServ & Pattern matching/Nearest neighbor & No\\
-%          \hline
-%          UM-DeepOutbreak & Google symptom search data, ILI & Deep neural network & No\\
-%          \hline
-%          UMass-flusion & Flu\_Surv, ILI, WHO/NREVSS test & Ensemble of statistical/machine learning time series & Yes\\
-%          \hline
-%     \end{tabular}
-%     \caption{Caption}
-%     \label{tab:my_label}
-% \end{table}
-
-% The overall forecasts are compared visually using PIT histograms (??) and reliability diagrams (??).
-These plots are used to assess model calibration, which in probabilistic forecasting relies on the...
-
-\section{Conclusion}
-\label{sec:conclusion}
-
- In this manuscript we introduce a statistical modeling framework which allows for the incorporation of several ILI forecast modeling methods.
-Specifically, we built upon \cite{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy.
-We model flu hospitalizations by incorporating the ILI forecast model into a model forecasting hospitalizations where hospitalization predictions are a linear or quadratic function of ILI.
-
-The simulation study in section \ref{sec:simulation2} suggests the ASG function in ILI modeling may slightly outperform the SIR model according to the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season forecasts, the SIR model was overall superior.
-The results from both the simulation study and the real data analysis suggest that the addition of a discrepancy component in ILI modeling may improve forecasts especially near the holiday week between Christmas and New Years day.
-It should not be assumed that these conclusions may be generalized for all locations in the US, weeks of a season, or for future flu seasons.
-% Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.
-
-Forecasting the seasonal influenza outbreak remains a challenging task for forecasters.
-The general modeling framework in this manuscript is successful under diverse modeling conditions for all locations in the US and may contribute to future forecasting efforts.
-All forecast models were presented as separate from one another, but in the various locations and times of the flu season, different models perform better than others.
-To build on the work done here, a natural step forward would be to combine all or a few selected forecasts into an ensemble forecast.
-Such an ensemble may work to cancel out certain model biases or highlight model strengths, leading to more robust forecasts.
-
-\section{Acknowledgments}
-This work is partially supported by the National Science Foundation under Grant No.
-2152117.
-Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
-% \spencer{I noticed that Luis acknowledged HPC@ISU for his work. Should I do the same?}
-
-%%%%%%%%
-
-%%%%
-% Reference section comes before the appendix
-
-%\bibliographystyle{acm} % use for numbered citations along with options given in preamble. Look at the main thesis.tex file
-\bibliographystyle{apa}
-
-\bibliography{master_bib}
-
-% %\section{Bibliography}
-% \bibliographystyle{apa}
-% % \vspace{-20pt}
-% \begingroup
-%     \setlength{\bibsep}{13.2pt}
-%     \linespread{1}\selectfont
-%     \bibliography{master_bib}
-% \endgroup
-% \clearpage
-% \pagebreak
-
-%%%%%
-%% Appendix
-% This section may or may not be included
-% Chapter 2 shows a double appendix example. Please use A or B as in Appendix A if there are multiple appendix. Use "Appendix A:" before writing the title
-% Chapter 3 shows a single appendix example. Use "Appendix:" before writing the title
-
-\begin{appendices}
+% \begin{appendices}
 \section{FluSight forecast competition scoring results} \label{app:A_rwis}
@@
     \centering
-    \includegraphics[scale=.7]{Images/rwis_ili_mod.png}
+    \includegraphics[scale=.7]{../../Images/rwis_ili_mod.png}
     \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
@@
     \centering
-    \includegraphics[scale=.6]{Images/rwis_hosp_mod.png}
+    \includegraphics[scale=.6]{../../Images/rwis_hosp_mod.png}
     \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
@@
   \centering
-  \includegraphics[width=1\linewidth]{Images/sir_disc_hosp_sq_ar1_rwis.png}
+  \includegraphics[width=1\linewidth]{../../Images/sir_disc_hosp_sq_ar1_rwis.png}
   % \caption{A subfigure}
@@
   \centering
-  \includegraphics[width=1\linewidth]{Images/asg_disc_hosp_ar1_rwis.png}
+  \includegraphics[width=1\linewidth]{../../Images/asg_disc_hosp_ar1_rwis.png}
   % \caption{A subfigure}
@@
     \centering
-    \includegraphics[scale=.6]{Images/posterior_theta_sir.png}
+    \includegraphics[scale=.6]{../../Images/posterior_theta_sir.png}
     \caption{Posterior 95\% credible intervals from ILI model for parameters of SIR differential equations.
@@
     \centering
-    \includegraphics[scale=.6]{Images/posterior_theta.png}
+    \includegraphics[scale=.6]{../../Images/posterior_theta.png}
     \caption{Posterior 95\% credible intervals from ILI model for parameters of ASG function.
@@
     \centering
-    \includegraphics[scale=.6]{Images/posterior_zeta.png}
+    \includegraphics[scale=.6]{../../Images/posterior_zeta.png}
     \caption{Posterior 95\% credible intervals from ILI model for variance parameters of the ASG function.
@@
   \centering
-  \includegraphics[width=.8\linewidth]{Images/posterior_kappa.png}
+  \includegraphics[width=.8\linewidth]{../../Images/posterior_kappa.png}
   % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
@@
   \centering
-  \includegraphics[width=.8\linewidth]{Images/posterior_sigma_gamma.png}
+  \includegraphics[width=.8\linewidth]{../../Images/posterior_sigma_gamma.png}
   % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy.
@@
 \begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha0_post.png}
+  \includegraphics[width=\linewidth]{../../Images/alpha0_post.png}
   % \caption{}
@@
 \begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha1_post.png}
+  \includegraphics[width=\linewidth]{../../Images/alpha1_post.png}
   % \caption{}
@@
 \begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha2_post.png}
+  \includegraphics[width=\linewidth]{../../Images/alpha2_post.png}
   % \caption{}
@@
 \begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/phi_post.png}
+  \includegraphics[width=\linewidth]{../../Images/phi_post.png}
   % \caption{}
@@
 \begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/sigma_epsilon_post.png}
+  \includegraphics[width=\linewidth]{../../Images/sigma_epsilon_post.png}
   % \caption{}
@@
 \begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/omega_post.png}
+  \includegraphics[width=\linewidth]{../../Images/omega_post.png}
   % \caption{}
@@
 
-% \begin{figure}
-% \centering
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha0_post.png}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha1_post.png}
-% \end{subfigure}
-% \caption{(left) 95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-(right) 95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.}
-% \label{fig:hosp_params_post1}
-% \end{figure}
+% \end{appendices}
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/alpha0_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha0_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/alpha1_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha1_post}
-% \end{figure}
-
-% \begin{figure}
-% \centering
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha2_post.png}
-%   % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-%   % \label{fig:sub1}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/phi_post.png}
-
-%   % \label{fig:sub2}
-% \end{subfigure}
-
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/omega_post.png}
-
-%   % \label{fig:sub2}
-
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/sigma_epsilon_post.png}
-% \end{subfigure}
-% \caption{(top left) 95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-% (top right) 95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-(bottom) 95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-% \label{fig:hosp_params_post2}
-% \end{figure}
-
-% \begin{figure}[hbt!]
-
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha0_post.png}
-%   % \caption{}
-%   % \label{MLEDdet}
-% \end{subfigure}\hfill % <-- "\hfill"
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha1_post.png}
-%   % \caption{}
-%   % \label{energydetPSK}
-% \end{subfigure}
-
-% \medskip % create some *vertical* separation between the graphs
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha2_post.png}
-%   % \caption{}
-%   % \label{velcomp}
-% \end{subfigure}\hfill % <-- "\hfill"
-% \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/phi_post.png}
-%   % \caption{}
-%   % \label{estcomp}
-% \end{subfigure}
-
-% \caption{A figure with four subfigures}
-% \label{fig:roc}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/alpha2_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha2_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/phi_post.png}
-%     \caption{95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:phi_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.]{Images/omega_post.png}
-%     \caption{95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:omega_post}
-% \end{figure}
-
-% This section may or may not be included
-% \newpage
-% \section{Appendix B: Additional tables showing median RWIS for location-model combinations}
-% \label{app:B_tables}
-
-% \begin{table}
-
-% % \parbox{.45\linewidth}{
-% \raggedleft
-% \caption{Model with minimum median RWIS by location}
-% \small
-% \centering
-% \begin{tabular}[t]{rrr}
-% \label{tab:wis_by_state}
-% \textbf{Location} & \textbf{Model} & \textbf{RWIS}\\
-
-% Alabama & ASGD-NORM & 0.768\\
-
-% Alaska & ASG-NORM-SQ & 0.886\\
-
-% Arizona & ASG-NORM-SQ & 0.669\\
-
-% Arkansas & SIRD-NORM-SQ & 0.734\\
-
-% California & ASGD-NORM-SQ & 0.688\\
-
-% Colorado & SIRD-NORM & 0.619\\
-
-% Connecticut & ASG-NORM & 0.945\\
-
-% Delaware & SIR-NORM-SQ & 0.836\\
-
-% District of Columbia & ASGD-LST-SQ & 0.970\\
-
-% Florida & ASGD-NORM-SQ & 0.680\\
-
-% Georgia & ASG-LNORM & 0.595\\
-
-% Hawaii & SIRD-LST & 0.979\\
-
-% Idaho & ASGD-NORM & 0.895\\
-
-% Illinois & SIRD-LST-SQ & 0.757\\
-
-% Indiana & SIRD-NORM-SQ & 0.696\\
-
-% Iowa & ASGD-NORM & 0.948\\
-
-% Kansas & ASG-NORM-SQ & 0.753\\
-
-% Kentucky & ASG-NORM-SQ & 0.697\\
-
-% Louisiana & SIRD-LST & 0.533\\
-
-% Maine & ASG-NORM-SQ & 0.879\\
-
-% Maryland & SIRD-LNORM & 0.737\\
-
-% Massachusetts & SIRD-NORM-SQ & 0.632\\
-
-% Michigan & SIRD-LST-SQ & 0.733\\
-
-% Minnesota & SIRD-LST-SQ & 0.673\\
-
-% Mississippi & SIR-LNORM-SQ & 0.742\\
-
-% Missouri & ASG-NORM-SQ & 0.797\\
-
-% Montana & SIRD-LNORM-SQ & 0.855\\
-
-% Nebraska & ASGD-NORM-SQ & 0.934\\
-
-% Nevada & SIRD-NORM-SQ & 0.786\\
-
-% New Hampshire & SIR-NORM & 0.766\\
-
-% New Jersey & ASG-LST & 0.811\\
-
-% New Mexico & SIRD-LNORM & 0.768\\
-
-% New York & SIRD-NORM-SQ & 0.765\\
-
-% North Carolina & ASGD-LST & 0.818\\
-
-% North Dakota & ASG-NORM-SQ & 0.838\\
-
-% Ohio & SIRD-NORM & 0.803\\
-
-% Oklahoma & ASGD-NORM-SQ & 0.861\\
-
-% Oregon & SIRD-LST & 0.838\\
-
-% Pennsylvania & SIRD-LNORM & 0.725\\
-
-% Puerto Rico & SIRD-LST & 0.966\\
-
-% Rhode Island & ASG-NORM & 0.869\\
-
-% South Carolina & ASG-LST & 0.670\\
-
-% South Dakota & SIR-LNORM & 0.950\\
-
-% Tennessee & ASG-LST & 0.610\\
-
-% Texas & SIRD-LST-SQ & 0.737\\
-
-% US & ASG-LST-SQ & 0.563\\
-
-% Utah & ASG-LST & 0.702\\
-
-% Vermont & SIRD-NORM-SQ & 0.960\\
-
-% Virginia & ASG-LST & 0.828\\
-
-% Washington & SIRD-NORM & 0.753\\
-
-% West Virginia & SIRD-LNORM & 0.828\\
-
-% Wisconsin & SIRD-LST-SQ & 0.793\\
-
-% Wyoming & SIRD-LNORM & 0.836\\
-% \hline
-% \end{tabular}
-% % \end{table}
-% % }
-% \end{table}
-
-% \begin{table}
-% \hfill
-% % \parbox{.45\linewidth}{
-% % \begin{table}
-% % \raggedright
-% \caption{Overall median RWIS and WIS by model}
-% \small
-% \centering
-% \begin{tabular}[t]{c|cc}
-% \label{tab:wis_by_model}
-% \textbf{Model} & \textbf{RWIS} & \textbf{WIS}\\
-% \hline
-% SIRD-NORM-SQ & 0.898 & 15.884\\
-
-% ASG-NORM-SQ & 0.903 & 17.152\\
-
-% ASGD-NORM-SQ & 0.922 & 17.408\\
-
-% SIRD-LST-SQ & 0.926 & 16.806\\
-
-% SIRD-NORM & 0.928 & 16.371\\
-
-% ASG-LST-SQ & 0.934 & 18.116\\
-
-% ASGD-LST-SQ & 0.943 & 17.949\\
-
-% SIRD-LST & 0.943 & 17.113\\
-
-% SIRD-LNORM & 0.945 & 20.046\\
-
-% ASG-NORM & 0.946 & 16.762\\
-
-% SIRD-LNORM-SQ & 0.951 & 20.110\\
-
-% ASG-LST & 0.957 & 17.180\\
-
-% ASGD-NORM & 0.966 & 17.208\\
-
-% ASGD-LST & 0.968 & 17.487\\
-
-% ASG-LNORM & 0.968 & 20.628\\
-
-% ASG-LNORM-SQ & 0.969 & 20.620\\
-
-% ASGD-LNORM & 0.982 & 21.108\\
-
-% ASGD-LNORM-SQ & 0.986 & 20.992\\
-
-% FluSight-baseline & 1.000 & 17.585\\
-
-% SIR-NORM-SQ & 1.002 & 18.156\\
-
-% SIR-LNORM & 1.009 & 21.441\\
-
-% SIR-LST-SQ & 1.009 & 18.547\\
-
-% SIR-LNORM-SQ & 1.013 & 21.494\\
-
-% SIR-LST & 1.034 & 19.000\\
-
-% SIR-NORM & 1.067 & 19.904\\
-% \hline
-% \end{tabular}
-% % }
-% \end{table}
-
-\end{appendices}
+\end{document}

### 2024-12-03 bayesian_analysis_version/nonlinear_forecast.tex
adjust directories

--- 2024-11-26 bayesian_analysis_version/vtex-soft-texsupport.isba-ba-97c6706/nonlinear_forecast.tex
@@
     \centering
-    \includegraphics[scale=.7]{../../Images/us_ili_seasons.png}
+    \includegraphics[scale=.7]{../Images/us_ili_seasons.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.7]{../../Images/ili_vs_week.png}
+    \includegraphics[scale=.7]{../Images/ili_vs_week.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.5]{../../Images/us_hospitalizations.png}
+    \includegraphics[scale=.5]{../Images/us_hospitalizations.png}
     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
@@
     \centering
-    \includegraphics[scale=.65]{../../Images/hosp_vs_week.png}
+    \includegraphics[scale=.65]{../Images/hosp_vs_week.png}
     \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
@@
     \centering
-    \includegraphics[scale=.6]{../../Images/ili_vs_hosp.png}
+    \includegraphics[scale=.6]{../Images/ili_vs_hosp.png}
     \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
@@
     \centering
-    \includegraphics[scale=.6]{../../Images/sir_traj.png}
+    \includegraphics[scale=.6]{../Images/sir_traj.png}
     \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
@@
     \centering
-    \includegraphics[scale=.45]{../../Images/asg_function.png}
+    \includegraphics[scale=.45]{../Images/asg_function.png}
     \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
@@
     \centering
-    \includegraphics[scale=.5]{../../Images/asg_fits.png}
+    \includegraphics[scale=.5]{../Images/asg_fits.png}
     \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
@@
     \centering
-    \includegraphics[scale=.45]{../../Images/discrepancy.png}
+    \includegraphics[scale=.45]{../Images/discrepancy.png}
     \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
@@
   \centering
-  \includegraphics[scale=.35]{../../Images/overall_crps.png}
+  \includegraphics[scale=.35]{../Images/overall_crps.png}
     % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
@@
   \centering
-  \includegraphics[scale=.35]{../../Images/overall_logs.png}
+  \includegraphics[scale=.35]{../Images/overall_logs.png}
     % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
@@
     \centering
-    \includegraphics[scale=.5]{../../Images/crps_by_season.png}
+    \includegraphics[scale=.5]{../Images/crps_by_season.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
@@
     \centering
-    \includegraphics[scale=.5]{../../Images/crps_by_week_horizon.png}
+    \includegraphics[scale=.5]{../Images/crps_by_week_horizon.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
     \centering
-    \includegraphics[scale=.5]{../../Images/logs_by_season.png}
+    \includegraphics[scale=.5]{../Images/logs_by_season.png}
     \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
@@
     \centering
-    \includegraphics[scale=.5]{../../Images/logs_by_week_horizon.png}
+    \includegraphics[scale=.5]{../Images/logs_by_week_horizon.png}
     \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
   \centering
-  \includegraphics[width=1\linewidth,height=6cm]{../../Images/normal_forecasts_us.png}
+  \includegraphics[width=1\linewidth,height=6cm]{../Images/normal_forecasts_us.png}
   % \caption{A subfigure}
@@
   \centering
-  \includegraphics[width=1\linewidth, height=6cm]{../../Images/normal_sq_forecasts_us.png}
+  \includegraphics[width=1\linewidth, height=6cm]{../Images/normal_sq_forecasts_us.png}
   % \caption{A subfigure}
@@
     % \centering
-    \includegraphics[scale = .5]{../../Images/lwis_us_full_season.png}
+    \includegraphics[scale = .5]{../Images/lwis_us_full_season.png}
     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
@@
     \centering
-    \includegraphics[scale = .7]{../../Images/lwis_by_traj_loc.png}
+    \includegraphics[scale = .7]{../Images/lwis_by_traj_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
     \centering
-    \includegraphics[scale = .7]{../../Images/lwis_by_dist_loc.png}
+    \includegraphics[scale = .7]{../Images/lwis_by_dist_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.

### 2024-12-03 nonlinear_forecast.tex
fix latex issues, adjust directories

--- 2024-12-03 bayesian_analysis_version/nonlinear_forecast.tex
@@
 \usepackage{enumerate} %for the short algorithm
+\usepackage{subfigure} %for making plots with multiple images
+\usepackage{tabularx} %for the spacing in tabular
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \centering
-    \includegraphics[scale=.7]{../Images/us_ili_seasons.png}
+    \includegraphics[scale=.7]{Images/us_ili_seasons.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.7]{../Images/ili_vs_week.png}
+    \includegraphics[scale=.7]{Images/ili_vs_week.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.5]{../Images/us_hospitalizations.png}
+    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
@@
     \centering
-    \includegraphics[scale=.65]{../Images/hosp_vs_week.png}
+    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
     \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
@@
     \centering
-    \includegraphics[scale=.6]{../Images/ili_vs_hosp.png}
+    \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
     \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
@@
     \centering
-    \includegraphics[scale=.6]{../Images/sir_traj.png}
+    \includegraphics[scale=.6]{Images/sir_traj.png}
     \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
@@
     \centering
-    \includegraphics[scale=.45]{../Images/asg_function.png}
+    \includegraphics[scale=.45]{Images/asg_function.png}
     \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
@@
     \centering
-    \includegraphics[scale=.5]{../Images/asg_fits.png}
+    \includegraphics[scale=.5]{Images/asg_fits.png}
     \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
@@
     \centering
-    \includegraphics[scale=.45]{../Images/discrepancy.png}
+    \includegraphics[scale=.45]{Images/discrepancy.png}
     \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
@@
   \centering
-  \includegraphics[scale=.35]{../Images/overall_crps.png}
+  \includegraphics[scale=.35]{Images/overall_crps.png}
     % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
@@
   \centering
-  \includegraphics[scale=.35]{../Images/overall_logs.png}
+  \includegraphics[scale=.35]{Images/overall_logs.png}
     % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
@@
     \centering
-    \includegraphics[scale=.5]{../Images/crps_by_season.png}
+    \includegraphics[scale=.5]{Images/crps_by_season.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
@@
     \centering
-    \includegraphics[scale=.5]{../Images/crps_by_week_horizon.png}
+    \includegraphics[scale=.5]{Images/crps_by_week_horizon.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
     \centering
-    \includegraphics[scale=.5]{../Images/logs_by_season.png}
+    \includegraphics[scale=.5]{Images/logs_by_season.png}
     \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
@@
     \centering
-    \includegraphics[scale=.5]{../Images/logs_by_week_horizon.png}
+    \includegraphics[scale=.5]{Images/logs_by_week_horizon.png}
     \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}
   \centering
-  \includegraphics[width=1\linewidth,height=6cm]{../Images/normal_forecasts_us.png}
+  \includegraphics[width=1\linewidth,height=6cm]{Images/normal_forecasts_us.png}
   % \caption{A subfigure}
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}
   \centering
-  \includegraphics[width=1\linewidth, height=6cm]{../Images/normal_sq_forecasts_us.png}
+  \includegraphics[width=1\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
   % \caption{A subfigure}
@@
     % \centering
-    \includegraphics[scale = .5]{../Images/lwis_us_full_season.png}
+    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
@@
     \centering
-    \includegraphics[scale = .7]{../Images/lwis_by_traj_loc.png}
+    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
     \centering
-    \includegraphics[scale = .7]{../Images/lwis_by_dist_loc.png}
+    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
 {@{\extracolsep{\fill}} 
-    l*{9}{c} 
-    S[table-format=4.0]}
+    l*{9}{c}}
   & & \multicolumn{3}{c}{ASG} 

### 2024-12-03 functions_data/get_data_functions_new.R

@@ -1,0 +1,609 @@
+library(cmdstanr)
+library(lubridate)
+library(ggplot2)
+library(posterior)
+library(bayesplot)
+library(rstanarm)
+library(stringr)
+library(dplyr)
+library(cdcfluview)
+library(tidyr)
+
+check_error <- function() {print("error is here")}
+print("spencer is the coolest")
+get_hosp_data <- function(lag = 1) {
+  # raw_flu <- read.csv(
+  #   'https://raw.githubusercontent.com/cdcepi/Flusight-forecast-data/master/data-truth/truth-Incident%20Hospitalizations.csv')
+  
+  last_date <- floor_date(today() - 12, unit = "week", week_start = 6)
+  
+  #raw_url <- paste("https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_",
+  #last_date, ".csv", sep = "")
+  
+  raw_url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/target-data/target-hospital-admissions.csv"
+  
+  # url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_2023-09-23.csv"
+  
+  raw_flu <- read.csv(raw_url)
+  raw_flu <- raw_flu %>% 
+    mutate(year = year(date),week = week(date) - lag, 
+           week2 = week(date) - (lag - 1)) %>% 
+    mutate(week = week(date(date) - 1))
+  
+  return(raw_flu)
+}
+
+get_ili_data <- function(lag = 2) {
+  ILINet_state <- ilinet(region = 'state')
+  ILINet_us <- ilinet(region = 'national') %>% mutate(region = 'US')
+  ILINet <- rbind(ILINet_state,ILINet_us)
+  
+  # ILINet <- ILINet %>% 
+  #   mutate(year=year(week_start), week2=week(week_start),location_name=region)
+  #this allows to control the lag. let's default to lag = 2
+  ILINet <- ILINet %>%
+	 mutate(week_start = date(week_start) + 7*(lag - 1))
+
+  ILINet <- ILINet %>% 
+    mutate(year=year(week_start), week2=week(week_start),location_name=region) #%>% 
+  # filter(month(week_start) %in% c(1:6,8:12))
+  
+
+  
[…]

### 2024-12-03 functions_data/get_data_functions_new_hosp.R

@@ -1,0 +1,608 @@
+library(cmdstanr)
+library(lubridate)
+library(ggplot2)
+library(posterior)
+library(bayesplot)
+library(rstanarm)
+library(stringr)
+library(dplyr)
+library(cdcfluview)
+library(tidyr)
+
+check_error <- function() {print("error is here")}
+print("spencer is the coolest")
+get_hosp_data <- function(lag = 1) {
+  # raw_flu <- read.csv(
+  #   'https://raw.githubusercontent.com/cdcepi/Flusight-forecast-data/master/data-truth/truth-Incident%20Hospitalizations.csv')
+  
+  last_date <- floor_date(today() - 12, unit = "week", week_start = 6)
+  
+  #raw_url <- paste("https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_",
+  #last_date, ".csv", sep = "")
+  
+  raw_url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/target-data/target-hospital-admissions.csv"
+  
+  # url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_2023-09-23.csv"
+  
+  raw_flu <- read.csv(raw_url)
+  raw_flu <- raw_flu %>% 
+    mutate(year = year(date),week = week(date) - lag, 
+           week2 = week(date) - (lag - 1)) %>% 
+    mutate(week = week(date(date) - 1))
+  
+  return(raw_flu)
+}
+
+get_ili_data <- function(lag = 2) {
+  ILINet_state <- ilinet(region = 'state')
+  ILINet_us <- ilinet(region = 'national') %>% mutate(region = 'US')
+  ILINet <- rbind(ILINet_state,ILINet_us)
+  
+  # ILINet <- ILINet %>% 
+  #   mutate(year=year(week_start), week2=week(week_start),location_name=region)
+  #this allows to control the lag. let's default to lag = 2
+  ILINet <- ILINet %>%
+	 mutate(week_start = date(week_start) + 7*(lag - 1))
+
+  ILINet <- ILINet %>% 
+    mutate(year=year(week_start), week2=week(week_start),location_name=region) #%>% 
+  # filter(month(week_start) %in% c(1:6,8:12))
+  
+
+  
[…]

### 2024-12-03 get_data_functions_new.R

@@ -1,609 +1,0 @@
-library(cmdstanr)
-library(lubridate)
-library(ggplot2)
-library(posterior)
-library(bayesplot)
-library(rstanarm)
-library(stringr)
-library(dplyr)
-library(cdcfluview)
-library(tidyr)
-
-check_error <- function() {print("error is here")}
-print("spencer is the coolest")
-get_hosp_data <- function(lag = 1) {
-  # raw_flu <- read.csv(
-  #   'https://raw.githubusercontent.com/cdcepi/Flusight-forecast-data/master/data-truth/truth-Incident%20Hospitalizations.csv')
-  
-  last_date <- floor_date(today() - 12, unit = "week", week_start = 6)
-  
-  #raw_url <- paste("https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_",
-  #last_date, ".csv", sep = "")
-  
-  raw_url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/target-data/target-hospital-admissions.csv"
-  
-  # url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_2023-09-23.csv"
-  
-  raw_flu <- read.csv(raw_url)
-  raw_flu <- raw_flu %>% 
-    mutate(year = year(date),week = week(date) - lag, 
-           week2 = week(date) - (lag - 1)) %>% 
-    mutate(week = week(date(date) - 1))
-  
-  return(raw_flu)
-}
-
-get_ili_data <- function(lag = 2) {
-  ILINet_state <- ilinet(region = 'state')
-  ILINet_us <- ilinet(region = 'national') %>% mutate(region = 'US')
-  ILINet <- rbind(ILINet_state,ILINet_us)
-  
-  # ILINet <- ILINet %>% 
-  #   mutate(year=year(week_start), week2=week(week_start),location_name=region)
-  #this allows to control the lag. let's default to lag = 2
-  ILINet <- ILINet %>%
-	 mutate(week_start = date(week_start) + 7*(lag - 1))
-
-  ILINet <- ILINet %>% 
-    mutate(year=year(week_start), week2=week(week_start),location_name=region) #%>% 
-  # filter(month(week_start) %in% c(1:6,8:12))
-  
-
-  
[…]

### 2024-12-03 get_data_functions_new_hosp.R

@@ -1,608 +1,0 @@
-library(cmdstanr)
-library(lubridate)
-library(ggplot2)
-library(posterior)
-library(bayesplot)
-library(rstanarm)
-library(stringr)
-library(dplyr)
-library(cdcfluview)
-library(tidyr)
-
-check_error <- function() {print("error is here")}
-print("spencer is the coolest")
-get_hosp_data <- function(lag = 1) {
-  # raw_flu <- read.csv(
-  #   'https://raw.githubusercontent.com/cdcepi/Flusight-forecast-data/master/data-truth/truth-Incident%20Hospitalizations.csv')
-  
-  last_date <- floor_date(today() - 12, unit = "week", week_start = 6)
-  
-  #raw_url <- paste("https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_",
-  #last_date, ".csv", sep = "")
-  
-  raw_url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/target-data/target-hospital-admissions.csv"
-  
-  # url <- "https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/target-data-archive/target-hospital-admissions_2023-09-23.csv"
-  
-  raw_flu <- read.csv(raw_url)
-  raw_flu <- raw_flu %>% 
-    mutate(year = year(date),week = week(date) - lag, 
-           week2 = week(date) - (lag - 1)) %>% 
-    mutate(week = week(date(date) - 1))
-  
-  return(raw_flu)
-}
-
-get_ili_data <- function(lag = 2) {
-  ILINet_state <- ilinet(region = 'state')
-  ILINet_us <- ilinet(region = 'national') %>% mutate(region = 'US')
-  ILINet <- rbind(ILINet_state,ILINet_us)
-  
-  # ILINet <- ILINet %>% 
-  #   mutate(year=year(week_start), week2=week(week_start),location_name=region)
-  #this allows to control the lag. let's default to lag = 2
-  ILINet <- ILINet %>%
-	 mutate(week_start = date(week_start) + 7*(lag - 1))
-
-  ILINet <- ILINet %>% 
-    mutate(year=year(week_start), week2=week(week_start),location_name=region) #%>% 
-  # filter(month(week_start) %in% c(1:6,8:12))
-  
-
-  
[…]

### 2024-12-03 mle_functions.R

@@ -1,359 +1,0 @@
-library(deSolve)
-library(bbmle)
-
-
-
-####################################################
-####################SIR MLE Stuff###################
-####################################################
-
-sir <- function(t,x,parms){
-	S <- x[1]
-	I <- x[2]
-	R <- x[3]
-	with(as.list(parms),
-	     {
-	       dS <- -beta*S*I
-	       dI <- beta*S*I - gamma*I
-	       dR <- gamma*I
-	       res <- c(dS,dI,dR)
-	       list(res)
-	     })
-}
-
-sirLL <- function(lbeta, lgamma, S = .9, logI0, logkappa, season_week,
-		  unweighted_ili) {
-	parms <- c(beta = exp(lbeta), gamma = exp(lgamma))
-	I <- exp(logI0)
-	R <- 1 - (S + exp(logI0))
-	x0 <- c(S = S, I = I, R = R)
-	out <- ode(y = x0, season_week, sir, parms)
-	bdens <- dbetaprop(y = unweighted_ili, mu = out[,3], kappa = exp(logkappa))
-	-sum(log(bdens))
-}
-
-get_sir_mle <- function(season_week, unweighted_ili, lbeta = log(1.1),
-			lgamma = log(.2), logI0 = log(.05),
-			logkappa = log(4), S = .9, mit = 1e5) {
-	#print("here baby!")
-	fit <- mle2(sirLL,
-		    start = list(lbeta = lbeta,
-				 lgamma = lgamma,
-				 logI0 = logI0,
-				 logkappa = logkappa),
-		    data = list(season_week = season_week,
-				unweighted_ili = unweighted_ili,
-				S = S),
-		    method = "Nelder-Mead",
-		    control = list(maxit = mit, trace = 0),
-		    trace = FALSE)
-	
-        #print("finally here!")
-	params <- coef(fit)
-	beta <- exp(params[1])
-	gamma <- exp(params[2])
-	I0 <- exp(params[3])
-	kappa <- exp(params[4])
-
-	S0 <- S
-	R0 <- 1 - (S0 + I0)
-	
-	N <- 1
-	parms <- c(N = N, beta = as.numeric(beta), gamma = as.numeric(gamma))
-	x0 <- c(S0, I0, R0)
-	stateMatrix <- ode(y = x0, season_week, sir, parms)
-	colnames(stateMatrix) <- c("season_week", "S", "I", "R")
[…]

### 2024-12-03 manuscript/manuscript.tex
fix arxiv tex

@@
-% \section{Abstract}
 \begin{abstract}
@@
     \centering
-    \includegraphics[scale=.7]{Images/us_ili_seasons.png}
+    \includegraphics[scale=.7]{../Images/us_ili_seasons.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.7]{Images/ili_vs_week.png}
+    \includegraphics[scale=.7]{../Images/ili_vs_week.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
+    \includegraphics[scale=.5]{../Images/us_hospitalizations.png}
     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
@@
     \centering
-    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
+    \includegraphics[scale=.65]{../Images/hosp_vs_week.png}
     \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
@@
     \centering
-    \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
+    \includegraphics[scale=.6]{../Images/ili_vs_hosp.png}
     \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
@@
 %     \centering
-%     \includegraphics[scale=.8]{Images/hosp_preds_ili_22.png}
+%     \includegraphics[scale=.8]{../Images/hosp_preds_ili_22.png}
 %     \caption{Weekly hospitalization counts during the 2022 flu season for 5 states and DC (black dots) and predicted weekly hospitalizations from the fit of model (\ref{eq:simple_hosp_reg}) (red lines)}
@@
     \centering
-    \includegraphics[scale=.6]{Images/sir_traj.png}
+    \includegraphics[scale=.6]{../Images/sir_traj.png}
     \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
@@
     \centering
-    \includegraphics[scale=.45]{Images/asg_function.png}
+    \includegraphics[scale=.45]{../Images/asg_function.png}
     \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
@@
     \centering
-    \includegraphics[scale=.5]{Images/asg_fits.png}
+    \includegraphics[scale=.5]{../Images/asg_fits.png}
     \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
@@
     \centering
-    \includegraphics[scale=.45]{Images/discrepancy.png}
+    \includegraphics[scale=.45]{../Images/discrepancy.png}
     \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
@@
 %         \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-%         \phi &\sim N(0, .4)\mathbbm{1}_{(-1,1)} \\
+%         \phi &\sim N(0, .4)\mathbb{1}_{(-1,1)} \\
 %         \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
@@
 As mentioned, most model parameter prior distributions were intended to be uninformative. 
-Appendix \ref{app:B_prior} shows plots of posterior distributions for select parameters from ILI and hospitalization models.
+% Figures \ref{fig:posterior_theta_sir} - \ref{fig:hosp_nus_param} in appendix \ref{app:B_prior} shows plots of posterior distributions for select parameters from ILI and hospitalization models.
+Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material.
 
@@
 \begin{equation}
-    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \mathbbm{1} (y^* \leq x))^2 dx
+    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \mathbb{1} (y^* \leq x))^2 dx
     \label{eq:crps2}
@@
   \centering
-  \includegraphics[scale=.35]{Images/overall_crps.png}
+  \includegraphics[scale=.35]{../Images/overall_crps.png}
     % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
@@
   \centering
-  \includegraphics[scale=.35]{Images/overall_logs.png}
+  \includegraphics[scale=.35]{../Images/overall_logs.png}
     % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
@@
 %     \centering
-%     \includegraphics[scale=.6]{Images/overall_crps.png}
+%     \includegraphics[scale=.6]{../Images/overall_crps.png}
 %     \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
@@
     \centering
-    \includegraphics[scale=.5]{Images/crps_by_season.png}
+    \includegraphics[scale=.5]{../Images/crps_by_season.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
@@
     \centering
-    \includegraphics[scale=.5]{Images/crps_by_week_horizon.png}
+    \includegraphics[scale=.5]{../Images/crps_by_week_horizon.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
 %     \centering
-%     \includegraphics[scale=.6]{Images/overall_logs.png}
+%     \includegraphics[scale=.6]{../Images/overall_logs.png}
 %     \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
@@
     \centering
-    \includegraphics[scale=.5]{Images/logs_by_season.png}
+    \includegraphics[scale=.5]{../Images/logs_by_season.png}
     \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
@@
     \centering
-    \includegraphics[scale=.5]{Images/logs_by_week_horizon.png}
+    \includegraphics[scale=.5]{../Images/logs_by_week_horizon.png}
     \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
   \centering
-  \includegraphics[width=1\linewidth,height=6cm]{Images/normal_forecasts_us.png}
+  \includegraphics[width=1\linewidth,height=6cm]{../Images/normal_forecasts_us.png}
   % \caption{A subfigure}
@@
   \centering
-  \includegraphics[width=1\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
+  \includegraphics[width=1\linewidth, height=6cm]{../Images/normal_sq_forecasts_us.png}
   % \caption{A subfigure}
@@
     % \centering
-    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
+    \includegraphics[scale = .5]{../Images/lwis_us_full_season.png}
     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
+    \includegraphics[scale = .7]{../Images/lwis_by_traj_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
+    \includegraphics[scale = .7]{../Images/lwis_by_dist_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
 {@{\extracolsep{\fill}} 
-    l 
-    *{9}{c} 
-    S[table-format=4.0]}
+    l*{9}{c}}
   & & \multicolumn{3}{c}{ASG} 
@@
 \bibliographystyle{apa}
-
 \bibliography{master_bib}
@@
 
-\begin{appendices}
-\section{FluSight forecast competition scoring results} \label{app:A_rwis}
-
-In the CDC flu forecast competition, a baseline forecast was made to which all other forecasts were compared.
+% \begin{appendices}
+% \section{FluSight forecast competition scoring results} \label{app:A_rwis}
+% 
+% In the CDC flu forecast competition, a baseline forecast was made to which all other forecasts were compared.
 The baseline forecast for a given state and week had as a median the most recent observed hospitalization count.
 The uncertainty was based on differences between previous hospitalizations, and is similar to the baseline forecast used in previous flu forecasting seasons and in the COVID-19 hub \cite[]{mathis2024evaluation, cramer2022evaluation}.
-The RWIS for one model was calculated by first taking the ratio of the average LWIS paired with every other model.
+% The RWIS for one model was calculated by first taking the ratio of the average LWIS paired with every other model.
 This was then diveded by the same ratio for the baseline forecasts (see methods in \cite{mathis2024evaluation} for details, and see also the supplementary material).
- An RWIS less than 1 indicates the forecast outperformed the baseline forecast.
- Figures \ref{fig:rwis_ili_tile} and \ref{fig:rwis_hosp_tile} show the mean RWIS of hospitalization forecasts over all 24 models from section \ref{sec:analysis} for all locations and weeks of the 2023 flu season.
+%  An RWIS less than 1 indicates the forecast outperformed the baseline forecast.
+%  Figures \ref{fig:rwis_ili_tile} and \ref{fig:rwis_hosp_tile} show the mean RWIS of hospitalization forecasts over all 24 models from section \ref{sec:analysis} for all locations and weeks of the 2023 flu season.
 Some similar patterns to those noted in section \ref{sec:analysis} emerge, including a slightly better performance by forecasts including discrepancy modeling over those which do not include discrepancy modeling and better overall performance by the SIR models than by the ASG forecast models.
@@
 This was not seen in the LWIS in figure \ref{fig:lwis_by_dist_loc}.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{Images/rwis_ili_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
+% 
+% 
+% \begin{figure}
+%     \centering
+%     \includegraphics[scale=.7]{../Images/rwis_ili_mod.png}
+%     \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
 The plot is separated by the four ILI models.
 Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_ili_tile}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/rwis_hosp_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
+%     \label{fig:rwis_ili_tile}
+% \end{figure}
+% 
+% 
+% \begin{figure}
+%     \centering
+%     \includegraphics[scale=.6]{../Images/rwis_hosp_mod.png}
+%     \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
 The plot is separated by the three distribution choices for the hospitalization model.
 Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_hosp_tile}
-\end{figure}
-
-Figure \ref{fig:state_median_points} shows the mean over weeks RWIS for the SIR and ASG forecast models with an RWIS less than 1 for the most locations.
+%     \label{fig:rwis_hosp_tile}
+% \end{figure}
+% 
+% 
+% Figure \ref{fig:state_median_points} shows the mean over weeks RWIS for the SIR and ASG forecast models with an RWIS less than 1 for the most locations.
 The plot on the left is from a forecast by an SIRD ILI model and a quadratic NORM hospitalization model.
@@
 This model had an RWIS less than 1 for 34 out of 53 locations.
-
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{Images/sir_disc_hosp_sq_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{Images/asg_disc_hosp_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{RWIS over the whole 2023 flu season for all 53 locations for hospitalization forecasts from an SIRD quadratic hospitalization model with normal errors (left) and forecasts from an ASGD linear hospitalization model with normal errors (right).
+% 
+% \begin{figure}
+% \centering
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=1\linewidth]{../Images/sir_disc_hosp_sq_ar1_rwis.png}
+%   % \caption{A subfigure}
+%   % \label{fig:sub1}
+% \end{subfigure}%
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=1\linewidth]{../Images/asg_disc_hosp_ar1_rwis.png}
+%   % \caption{A subfigure}
+%   % \label{fig:sub2}
+% \end{subfigure}
+% \caption{RWIS over the whole 2023 flu season for all 53 locations for hospitalization forecasts from an SIRD quadratic hospitalization model with normal errors (left) and forecasts from an ASGD linear hospitalization model with normal errors (right).
 An RWIS less than 1 (left of vertical grey line) indicates the model forecasts outperformed the baseline forecasts for that particular region.}
-\label{fig:state_median_points}
-\end{figure}
-
-\section{Posterior distribution plots for select parameters}
-\label{app:B_prior}
-
-The figures in this section show 95\% credible intervals for model parameters under the several modeling schemes along with the prior distributions assigned to the parameters.
+% \label{fig:state_median_points}
+% \end{figure}
+% 
+% 
+% 
+% 
+% 
+% 
+% \section{Posterior distribution plots for select parameters}
+% \label{app:B_prior}
+% 
+% The figures in this section show 95\% credible intervals for model parameters under the several modeling schemes along with the prior distributions assigned to the parameters.
 Figure \ref{fig:posterior_theta_sir} is for parameters unique to the SIR ILI model.
@@
 Figures \ref{fig:hosp_lin_params} and \ref{fig:hosp_nus_param} are for parameters used in hospitalization modeling.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/posterior_theta_sir.png}
-    \caption{Posterior 95\% credible intervals from ILI model for parameters of SIR differential equations.
+% 
+% 
+% \begin{figure}
+%     \centering
+%     \includegraphics[scale=.6]{../Images/posterior_theta_sir.png}
+%     \caption{Posterior 95\% credible intervals from ILI model for parameters of SIR differential equations.
 Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
@@
 The red interval is the 95\% interval of the prior distribution.}
-    \label{fig:posterior_theta_sir}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/posterior_theta.png}
-    \caption{Posterior 95\% credible intervals from ILI model for parameters of ASG function.
+%     \label{fig:posterior_theta_sir}
+% \end{figure}
+% 
+% \begin{figure}
+%     \centering
+%     \includegraphics[scale=.6]{../Images/posterior_theta.png}
+%     \caption{Posterior 95\% credible intervals from ILI model for parameters of ASG function.
 Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
@@
 The blue interval is the 95\% interval of the prior distribution.}
-    \label{fig:posterior_theta}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/posterior_zeta.png}
-    \caption{Posterior 95\% credible intervals from ILI model for variance parameters of the ASG function.
+%     \label{fig:posterior_theta}
+% \end{figure}
+% 
+% 
+% \begin{figure}
+%     \centering
+%     \includegraphics[scale=.6]{../Images/posterior_zeta.png}
+%     \caption{Posterior 95\% credible intervals from ILI model for variance parameters of the ASG function.
 Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
@@
 The blue interval is the 95\% interval of the prior distribution.}
-    \label{fig:posterior_zeta}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{Images/posterior_kappa.png}
-  % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{Images/posterior_sigma_gamma.png}
-  % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The blue is from the SIRD model and red from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Posterior 95\% credible intervals from ILI model for scale parameter $\kappa_s$.
-Blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD (left).
-Posterior 95\% credible intervals from ILI model for scale parameter of modeled discrepancy $\sigma_{\gamma}$ (right).
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-Blue is from the SIRD model and red from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-\label{fig:sir_asg_shared}
-\end{figure}
-
-\begin{figure}[hbt!]
-
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha0_post.png}
-  % \caption{}
-  % \label{MLEDdet}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha1_post.png}
-  % \caption{}
-  % \label{energydetPSK}
-\end{subfigure}
-
-\medskip % create some *vertical* separation between the graphs
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/alpha2_post.png}
-  % \caption{}
-  % \label{velcomp}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/phi_post.png}
-  % \caption{}
-  % \label{estcomp}
-\end{subfigure}
-
-\caption{Posterior 95\% credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top left).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top right).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (bottom left).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included (bottom right).
-Prior distribution 95\% interval is also included.
-In each plot intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
-\label{fig:hosp_lin_params}
-\end{figure}
-
-\begin{figure}[hbt!]
-
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/sigma_epsilon_post.png}
-  % \caption{}
-  % \label{MLEDdet}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{Images/omega_post.png}
-  % \caption{}
-  % \label{energydetPSK}
-\end{subfigure}
-
-\caption{Posterior 95\% credible intervals for $\sigma_{\epsilon}$ for hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (left).
-Prior distribution 95\% interval is also included.
-95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term (right).
-Prior distribution 95\% interval is also included.
-In all plots intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
-\label{fig:hosp_nus_param}
-\end{figure}
-
+%     \label{fig:posterior_zeta}
+% \end{figure}
+% 
 % \begin{figure}
@@
 %   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha0_post.png}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha1_post.png}
-% \end{subfigure}
-% \caption{(left) 95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-(right) 95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.}
-% \label{fig:hosp_params_post1}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.5]{Images/alpha0_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha0_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/alpha1_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha1_post}
-% \end{figure}
-
-% \begin{figure}
-% \centering
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/alpha2_post.png}
+%   \includegraphics[width=.8\linewidth]{../Images/posterior_kappa.png}
 %   % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
@@
 %   \centering
-%   \includegraphics[width=.8\linewidth]{Images/phi_post.png}
-
+%   \includegraphics[width=.8\linewidth]{../Images/posterior_sigma_gamma.png}
+%   % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy.
+Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
+The blue is from the SIRD model and red from ASGD.
+The green interval is the 95\% interval of the prior distribution.}
 %   % \label{fig:sub2}
 % \end{subfigure}
-
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/omega_post.png}
-
-%   % \label{fig:sub2}
-
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}
-%   \centering
-%   \includegraphics[width=.8\linewidth]{Images/sigma_epsilon_post.png}
-% \end{subfigure}
-% \caption{(top left) 95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-% (top right) 95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-Prior distribution 95\% interval is also included.
-(bottom) 95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-% \label{fig:hosp_params_post2}
+% \caption{Posterior 95\% credible intervals from ILI model for scale parameter $\kappa_s$.
+Blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD (left).
+Posterior 95\% credible intervals from ILI model for scale parameter of modeled discrepancy $\sigma_{\gamma}$ (right).
+Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
+Blue is from the SIRD model and red from ASGD.
+The green interval is the 95\% interval of the prior distribution.}
+% \label{fig:sir_asg_shared}
 % \end{figure}
-
+% 
 % \begin{figure}[hbt!]
-
+% 
 % \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha0_post.png}
+%   \includegraphics[width=\linewidth]{../Images/alpha0_post.png}
 %   % \caption{}
@@
 % \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha1_post.png}
+%   \includegraphics[width=\linewidth]{../Images/alpha1_post.png}
 %   % \caption{}
@@
 % \end{subfigure}
-
+% 
 % \medskip % create some *vertical* separation between the graphs
 % \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/alpha2_post.png}
+%   \includegraphics[width=\linewidth]{../Images/alpha2_post.png}
 %   % \caption{}
@@
 % \begin{subfigure}{.475\linewidth}
-%   \includegraphics[width=\linewidth]{Images/phi_post.png}
+%   \includegraphics[width=\linewidth]{../Images/phi_post.png}
 %   % \caption{}
@@
 % \end{subfigure}
-
-% \caption{A figure with four subfigures}
-% \label{fig:roc}
+% 
+% \caption{Posterior 95\% credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
+Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top left).
+Prior distribution 95\% interval is also included.
+95\% posterior credible intervals for $\alpha_1$ for the three hospitalization models separated by whether or not the squared ILI term was included.
+Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (top right).
+Prior distribution 95\% interval is also included.
+% 95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
+Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (bottom left).
+Prior distribution 95\% interval is also included.
+% 95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included (bottom right).
+Prior distribution 95\% interval is also included.
+In each plot intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
+% \label{fig:hosp_lin_params}
 % \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/alpha2_post.png}
-%     \caption{95\% posterior credible intervals for $\alpha_2$ for the three hospitalization models.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:alpha2_post}
+% 
+% 
+% \begin{figure}[hbt!]
+% 
+% \begin{subfigure}{.475\linewidth}
+%   \includegraphics[width=\linewidth]{../Images/sigma_epsilon_post.png}
+%   % \caption{}
+%   % \label{MLEDdet}
+% \end{subfigure}\hfill % <-- "\hfill"
+% \begin{subfigure}{.475\linewidth}
+%   \includegraphics[width=\linewidth]{../Images/omega_post.png}
+%   % \caption{}
+%   % \label{energydetPSK}
+% \end{subfigure}
+% 
+% \caption{Posterior 95\% credible intervals for $\sigma_{\epsilon}$ for hospitalization models with and without the ILI squared term.
+Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (left).
+Prior distribution 95\% interval is also included.
+% 95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term (right).
+Prior distribution 95\% interval is also included.
+In all plots intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
+% \label{fig:hosp_nus_param}
 % \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.6]{Images/phi_post.png}
-%     \caption{95\% posterior credible intervals for $\phi$ for the three hospitalization models separated by whether or not the squared ILI term was included.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:phi_post}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[scale=.]{Images/omega_post.png}
-%     \caption{95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term.
-Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.
-95\% interval for prior distribution is also included}
-%     \label{fig:omega_post}
-% \end{figure}
-
-% This section may or may not be included
-% \newpage
-% \section{Appendix B: Additional tables showing median RWIS for location-model combinations}
-% \label{app:B_tables}
-
-% \begin{table}
-
-% % \parbox{.45\linewidth}{
-% \raggedleft
-% \caption{Model with minimum median RWIS by location}
-% \small
-% \centering
-% \begin{tabular}[t]{rrr}
-% \label{tab:wis_by_state}
-% \textbf{Location} & \textbf{Model} & \textbf{RWIS}\\
-
-% Alabama & ASGD-NORM & 0.768\\
-
-% Alaska & ASG-NORM-SQ & 0.886\\
-
-% Arizona & ASG-NORM-SQ & 0.669\\
-
-% Arkansas & SIRD-NORM-SQ & 0.734\\
-
-% California & ASGD-NORM-SQ & 0.688\\
-
-% Colorado & SIRD-NORM & 0.619\\
-
-% Connecticut & ASG-NORM & 0.945\\
-
-% Delaware & SIR-NORM-SQ & 0.836\\
-
-% District of Columbia & ASGD-LST-SQ & 0.970\\
-
-% Florida & ASGD-NORM-SQ & 0.680\\
-
-% Georgia & ASG-LNORM & 0.595\\
-
-% Hawaii & SIRD-LST & 0.979\\
-
-% Idaho & ASGD-NORM & 0.895\\
-
-% Illinois & SIRD-LST-SQ & 0.757\\
-
-% Indiana & SIRD-NORM-SQ & 0.696\\
-
-% Iowa & ASGD-NORM & 0.948\\
-
-% Kansas & ASG-NORM-SQ & 0.753\\
-
-% Kentucky & ASG-NORM-SQ & 0.697\\
-
-% Louisiana & SIRD-LST & 0.533\\
-
-% Maine & ASG-NORM-SQ & 0.879\\
-
-% Maryland & SIRD-LNORM & 0.737\\
-
-% Massachusetts & SIRD-NORM-SQ & 0.632\\
-
-% Michigan & SIRD-LST-SQ & 0.733\\
-
-% Minnesota & SIRD-LST-SQ & 0.673\\
-
-% Mississippi & SIR-LNORM-SQ & 0.742\\
-
-% Missouri & ASG-NORM-SQ & 0.797\\
-
-% Montana & SIRD-LNORM-SQ & 0.855\\
-
-% Nebraska & ASGD-NORM-SQ & 0.934\\
-
-% Nevada & SIRD-NORM-SQ & 0.786\\
-
-% New Hampshire & SIR-NORM & 0.766\\
-
-% New Jersey & ASG-LST & 0.811\\
-
-% New Mexico & SIRD-LNORM & 0.768\\
-
-% New York & SIRD-NORM-SQ & 0.765\\
-
-% North Carolina & ASGD-LST & 0.818\\
-
-% North Dakota & ASG-NORM-SQ & 0.838\\
-
-% Ohio & SIRD-NORM & 0.803\\
-
-% Oklahoma & ASGD-NORM-SQ & 0.861\\
-
-% Oregon & SIRD-LST & 0.838\\
-
-% Pennsylvania & SIRD-LNORM & 0.725\\
-
-% Puerto Rico & SIRD-LST & 0.966\\
-
-% Rhode Island & ASG-NORM & 0.869\\
-
-% South Carolina & ASG-LST & 0.670\\
-
-% South Dakota & SIR-LNORM & 0.950\\
-
-% Tennessee & ASG-LST & 0.610\\
-
-% Texas & SIRD-LST-SQ & 0.737\\
-
-% US & ASG-LST-SQ & 0.563\\
-
-% Utah & ASG-LST & 0.702\\
-
-% Vermont & SIRD-NORM-SQ & 0.960\\
-
-% Virginia & ASG-LST & 0.828\\
-
-% Washington & SIRD-NORM & 0.753\\
-
-% West Virginia & SIRD-LNORM & 0.828\\
-
-% Wisconsin & SIRD-LST-SQ & 0.793\\
-
-% Wyoming & SIRD-LNORM & 0.836\\
-% \hline
-% \end{tabular}
-% % \end{table}
-% % }
-% \end{table}
-
-% \begin{table}
-% \hfill
-% % \parbox{.45\linewidth}{
-% % \begin{table}
-% % \raggedright
-% \caption{Overall median RWIS and WIS by model}
-% \small
-% \centering
-% \begin{tabular}[t]{c|cc}
-% \label{tab:wis_by_model}
-% \textbf{Model} & \textbf{RWIS} & \textbf{WIS}\\
-% \hline
-% SIRD-NORM-SQ & 0.898 & 15.884\\
-
-% ASG-NORM-SQ & 0.903 & 17.152\\
-
-% ASGD-NORM-SQ & 0.922 & 17.408\\
-
-% SIRD-LST-SQ & 0.926 & 16.806\\
-
-% SIRD-NORM & 0.928 & 16.371\\
-
-% ASG-LST-SQ & 0.934 & 18.116\\
-
-% ASGD-LST-SQ & 0.943 & 17.949\\
-
-% SIRD-LST & 0.943 & 17.113\\
-
-% SIRD-LNORM & 0.945 & 20.046\\
-
-% ASG-NORM & 0.946 & 16.762\\
-
-% SIRD-LNORM-SQ & 0.951 & 20.110\\
-
-% ASG-LST & 0.957 & 17.180\\
-
-% ASGD-NORM & 0.966 & 17.208\\
-
-% ASGD-LST & 0.968 & 17.487\\
-
-% ASG-LNORM & 0.968 & 20.628\\
-
-% ASG-LNORM-SQ & 0.969 & 20.620\\
-
-% ASGD-LNORM & 0.982 & 21.108\\
-
-% ASGD-LNORM-SQ & 0.986 & 20.992\\
-
-% FluSight-baseline & 1.000 & 17.585\\
-
-% SIR-NORM-SQ & 1.002 & 18.156\\
-
-% SIR-LNORM & 1.009 & 21.441\\
-
-% SIR-LST-SQ & 1.009 & 18.547\\
-
-% SIR-LNORM-SQ & 1.013 & 21.494\\
-
-% SIR-LST & 1.034 & 19.000\\
-
-% SIR-NORM & 1.067 & 19.904\\
-% \hline
-% \end{tabular}
-% % }
-% \end{table}
-
-\end{appendices}
+% 
+% 
+% 
+% \end{appendices}
+% \end{document}

### 2024-12-03 nonlinear_forecast.tex
fix supplemental

@@
 As mentioned, most model parameter prior distributions were intended to be uninformative.
-Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material.
+Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material \cite[]{wadsworth2024bas}.
 
@@
   \centering
-  \includegraphics[scale=.35]{Images/overall_crps.png}
+  \includegraphics[scale=.3]{Images/overall_crps.png}
     % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
@@
   \centering
-  \includegraphics[scale=.35]{Images/overall_logs.png}
+  \includegraphics[scale=.3]{Images/overall_logs.png}
     % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
@@
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecastss and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
+Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
     \label{fig:crps_by_week_horizon}
@@
  Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS).
-The focus in this section will be on the LWIS (see supplementary material for RWIS based results).
+The focus in this section will be on the LWIS (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
 
@@
   \centering
-  \includegraphics[width=1\linewidth,height=6cm]{Images/normal_forecasts_us.png}
+  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
   % \caption{A subfigure}
@@
   \centering
-  \includegraphics[width=1\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
+  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
   % \caption{A subfigure}
@@
 \begin{supplement}
-\stitle{Posterior distribution plots for select parameters}
+\stitle{1. Posterior distribution plots for select parameters}
 \sdescription{Plots of posterior distributions for parameters in both ILI and hospitalization models}
@@
 \begin{supplement}
-\stitle{FluSight forecast competition scoring results}
+\stitle{2. FluSight forecast competition scoring results}
 \sdescription{Flu forecast results with relative weighted interval scoring rule similar to that used in  FluSight}

### 2024-12-03 supplemental.tex
fix supplemental

--- 2024-11-26 manuscript/supplementary/supplementary.tex
@@
-\documentclass{article}
+% \documentclass[ba,preprint]{imsart}% use this for supplement article
+\documentclass[ba]{imsart}
 
-\usepackage{PRIMEarxiv}
+%% Packages
+\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
+\RequirePackage[numbers]{natbib}
+%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
+\RequirePackage{graphicx}
 
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage[linktocpage=true]{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{lipsum}
-\usepackage{fancyhdr}       % header
-\usepackage{graphicx}       % graphics
-\graphicspath{{media/}}     % organize your images and other figures under media/ folder
+%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
+\usepackage{booktabs} %for the table in the analysis section
+\usepackage{enumerate} %for the short algorithm
+\usepackage{subfigure} %for making plots with multiple images
+\usepackage{tabularx} %for the spacing in tabular
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Packages added by Spencer
-\usepackage{amsmath,amsthm,amssymb,scrextend}
-\usepackage{appendix}
-\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}
-\usepackage{enumerate}
-\usepackage{csquotes}
-\usepackage{subcaption}
-\usepackage{amssymb}
-% \usepackage{bbm}
-\usepackage[bb=pazo]{mathalpha}
-\usepackage{bm}
+\pubyear{2024}
+\arxiv{2010.00000}
+\volume{TBA}
+\issue{TBA}
+\firstpage{1}
+\lastpage{1}
 
-\pagestyle{empty}
-
-\newcommand\Bbbbone{%
-  \ifdefined\mathbbb%
-    \mathbbb{1}%
-  \else%
-    \boldsymbol{\mathbb{1}}%
-  \fi}
-
-\usepackage{xr}
-\externaldocument{supplementary}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \usepackage[pdftex]{graphicx}
-% Standard, old-style thesis
-% \usepackage{isutraditional}   
-% \chaptertitle
-% Old-style, thesis numbering down to subsubsection
-\alternate
-\usepackage{rotating}
-% Bibliography without numbers or labels
-\usepackage{natbib}
-% Use the following line if you want square brackets and numbering system
-%\usepackage[square, numbers]{natbib}
-
-% this package is for getting certain math scripts like the 1 needed for an indicator function
-\usepackage{bbm}
-
-% using this package to make a better looking list
-\usepackage{enumerate}
-% try this one for a list as well. may be better
-% \usepackage[font=itshape]{quoting}
-
-% package for tables
-
-\renewcommand{\bibfont}{\setstretch{1}\selectfont}
-\setlength{\bibsep}{13.2pt}
-
-\usepackage{chapterbib}
-%\renewcommand\bibsection{\section*{\ref name}}
-%\renewcommand\bibsection{\section*}
-\renewcommand{\bibsection}{\section{References}}
-
-%\bibliographystyle{apa}  % changed from apa, abbrv, unsrtnat
-%\includeonly{titletoc,chapter1}
-%Optional Package to add PDF bookmarks and hypertext links
-% \usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
-\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
-
-%\usepackage{titletoc}
-% \usepackage{hyperref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Packages/commands added by Jarad
-\usepackage{xcolor}
-\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
-\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}
-
-%Header
-\pagestyle{fancy}
-\thispagestyle{empty}
-\rhead{ \textit{ }}
-
-% Update your Headers here
-\fancyhead[LO]{Bayesian nonlinear flu hospitalizations forecasting}
-% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}
-
-%% Title
-\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy
-%%%% Cite as
-%%%% Update your official citation here when published 
-% \thanks{\textit{\underline{Citation}}: 
-% \textbf{Authors. Title. Pages.... DOI:000000/11111.}} 
-}
-
-\author{
-  Spencer Wadsworth \\
-  Iowa State University \\
-  Ames, Iowa\\
-  \texttt{sgw96@iastate.edu} \\
-  %% examples of more authors
-   \And
-  Jarad Niemi \\
-  Iowa State University \\
-  Ames, Iowa\\
-  \texttt{niemi@iastate.edu} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-}
+\startlocaldefs
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Uncomment next line to change            %%
+%% the type of equation numbering           %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\numberwithin{equation}{section}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Axiom, Claim, Corollary, Hypothesis, %%
+%% Lemma, Theorem, Proposition              %%
+%% use \theoremstyle{plain}                 %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{plain}
+\newtheorem{axiom}{Axiom}
+\newtheorem{claim}[axiom]{Claim}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Assumption, Definition, Example,     %%
+%% Notation, Property, Remark, Fact         %%
+%% use \theoremstyle{definition}            %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem*{example}{Example}
+\newtheorem*{fact}{Fact}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Case use \theoremstyle{remark}       %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{remark}
+\newtheorem{case}{Case}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Please put your definitions here:        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\endlocaldefs
 
@@
 
-% \begin{appendices}
-\section{FluSight forecast competition scoring results} \label{app:A_rwis}
-
-In the CDC flu forecast competition, a baseline forecast was made to which all other forecasts were compared.
-The baseline forecast for a given state and week had as a median the most recent observed hospitalization count.
-The uncertainty was based on differences between previous hospitalizations, and is similar to the baseline forecast used in previous flu forecasting seasons and in the COVID-19 hub \cite[]{mathis2024evaluation, cramer2022evaluation}.
-The RWIS for one model was calculated by first taking the ratio of the average LWIS paired with every other model.
-This was then diveded by the same ratio for the baseline forecasts (see methods in \cite{mathis2024evaluation} for details, and see also the supplementary material).
- An RWIS less than 1 indicates the forecast outperformed the baseline forecast.
- Figures \ref{fig:rwis_ili_tile} and \ref{fig:rwis_hosp_tile} show the mean RWIS of hospitalization forecasts over all 24 models from section \ref{sec:analysis} for all locations and weeks of the 2023 flu season.
-Some similar patterns to those noted in section \ref{sec:analysis} emerge, including a slightly better performance by forecasts including discrepancy modeling over those which do not include discrepancy modeling and better overall performance by the SIR models than by the ASG forecast models.
-Interestingly, figure \ref{fig:rwis_hosp_tile} shows the LNORM model showing especially poor RWIS performance about three quarters into the season.
-This was not seen in the LWIS in figure \ref{fig:lwis_by_dist_loc}.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{../../Images/rwis_ili_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
-The plot is separated by the four ILI models.
-Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_ili_tile}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{../../Images/rwis_hosp_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
-The plot is separated by the three distribution choices for the hospitalization model.
-Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_hosp_tile}
-\end{figure}
-
-Figure \ref{fig:state_median_points} shows the mean over weeks RWIS for the SIR and ASG forecast models with an RWIS less than 1 for the most locations.
-The plot on the left is from a forecast by an SIRD ILI model and a quadratic NORM hospitalization model.
-This forecast model had an RWIS less than 1 for 37 out of 53 locations.
-The plot on the right is from a forecast by an ASGD ILI model and linear NORM hospitalization model.
-This model had an RWIS less than 1 for 34 out of 53 locations.
-
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{../../Images/sir_disc_hosp_sq_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{../../Images/asg_disc_hosp_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{RWIS over the whole 2023 flu season for all 53 locations for hospitalization forecasts from an SIRD quadratic hospitalization model with normal errors (left) and forecasts from an ASGD linear hospitalization model with normal errors (right).
-An RWIS less than 1 (left of vertical grey line) indicates the model forecasts outperformed the baseline forecasts for that particular region.}
-\label{fig:state_median_points}
-\end{figure}
+\begin{supplement}
 
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
     \centering
-    \includegraphics[scale=.6]{../../Images/posterior_theta_sir.png}
+    \includegraphics[scale=.6]{Images/posterior_theta_sir.png}
     \caption{Posterior 95\% credible intervals from ILI model for parameters of SIR differential equations.
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
     \centering
-    \includegraphics[scale=.6]{../../Images/posterior_theta.png}
+    \includegraphics[scale=.6]{Images/posterior_theta.png}
     \caption{Posterior 95\% credible intervals from ILI model for parameters of ASG function.
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
     \centering
-    \includegraphics[scale=.6]{../../Images/posterior_zeta.png}
+    \includegraphics[scale=.6]{Images/posterior_zeta.png}
     \caption{Posterior 95\% credible intervals from ILI model for variance parameters of the ASG function.
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}
   \centering
-  \includegraphics[width=.8\linewidth]{../../Images/posterior_kappa.png}
+  \includegraphics[width=.49\linewidth]{Images/posterior_kappa.png}
   % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}
   \centering
-  \includegraphics[width=.8\linewidth]{../../Images/posterior_sigma_gamma.png}
+  \includegraphics[width=.49\linewidth]{Images/posterior_sigma_gamma.png}
   % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy.
@@
 
+% \begin{figure}[hbt!]
+% 
+% \begin{subfigure}
+%   \includegraphics[width=.49\linewidth]{Images/alpha0_post.png}
+%   % \caption{}
+%   % \label{MLEDdet}
+% \end{subfigure}\hfill % <-- "\hfill"
+% \begin{subfigure}
+%   \includegraphics[width=.49\linewidth]{Images/alpha1_post.png}
+%   % \caption{}
+%   % \label{energydetPSK}
+% \end{subfigure}
+% 
+% \medskip % create some *vertical* separation between the graphs
+% \begin{subfigure}
+%   \includegraphics[width=.49\linewidth]{Images/alpha2_post.png}
+%   % \caption{}
+%   % \label{velcomp}
+% \end{subfigure}\hfill % <-- "\hfill"
+% \begin{subfigure}
+%   \includegraphics[width=.49\linewidth]{Images/phi_post.png}
+%   % \caption{}
+%   % \label{estcomp}
+% \end{subfigure}
+
 \begin{figure}[hbt!]
-
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{../../Images/alpha0_post.png}
-  % \caption{}
-  % \label{MLEDdet}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{../../Images/alpha1_post.png}
-  % \caption{}
-  % \label{energydetPSK}
+\centering
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/alpha0_post.png}
+\end{subfigure}%
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/alpha1_post.png}
 \end{subfigure}
-
-\medskip % create some *vertical* separation between the graphs
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{../../Images/alpha2_post.png}
-  % \caption{}
-  % \label{velcomp}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{../../Images/phi_post.png}
-  % \caption{}
-  % \label{estcomp}
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/alpha2_post.png}
+\end{subfigure}%
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/phi_post.png}
 \end{subfigure}
-
 \caption{Posterior 95\% credible intervals for $\alpha_0$ for the three hospitalization models separated by whether or not the squared ILI term was included.
@@
 
+% \begin{figure}[hbt!]
+% 
+% \begin{subfigure}
+%   \includegraphics[width=\linewidth]{Images/sigma_epsilon_post.png}
+%   % \caption{}
+%   % \label{MLEDdet}
+% \end{subfigure}\hfill % <-- "\hfill"
+% \begin{subfigure}
+%   \includegraphics[width=\linewidth]{Images/omega_post.png}
+%   % \caption{}
+%   % \label{energydetPSK}
+% \end{subfigure}
+% 
+% \caption{Posterior 95\% credible intervals for $\sigma_{\epsilon}$ for hospitalization models with and without the ILI squared term.
+Intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown (left).
+Prior distribution 95\% interval is also included.
+% 95\% posterior credible intervals for $\omega$ for LST hospitalization models with and without the ILI squared term (right).
+Prior distribution 95\% interval is also included.
+In all plots intervals are for the US hospitalizations and weeks 14, 20 and 26 are shown.}
+% \label{fig:hosp_nus_param}
+% \end{figure}
+
 \begin{figure}[hbt!]
-
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{../../Images/sigma_epsilon_post.png}
-  % \caption{}
-  % \label{MLEDdet}
-\end{subfigure}\hfill % <-- "\hfill"
-\begin{subfigure}{.475\linewidth}
-  \includegraphics[width=\linewidth]{../../Images/omega_post.png}
-  % \caption{}
-  % \label{energydetPSK}
+\centering
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/sigma_epsilon_post.png}
+  % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
+Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
+The blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD.
+The green interval is the 95\% interval of the prior distribution.}
+  % \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/omega_post.png}
+  % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy.
+Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
+The blue is from the SIRD model and red from ASGD.
+The green interval is the 95\% interval of the prior distribution.}
+  % \label{fig:sub2}
 \end{subfigure}
-
 \caption{Posterior 95\% credible intervals for $\sigma_{\epsilon}$ for hospitalization models with and without the ILI squared term.
@@
 
-% \end{appendices}
+\newpage
+\section{FluSight forecast competition scoring results} \label{app:A_rwis}
+
+In the CDC flu forecast competition, a baseline forecast was made to which all other forecasts were compared.
+The baseline forecast for a given state and week had as a median the most recent observed hospitalization count.
+The uncertainty was based on differences between previous hospitalizations, and is similar to the baseline forecast used in previous flu forecasting seasons and in the COVID-19 hub \cite[]{mathis2024evaluation, cramer2022evaluation}.
+The RWIS for one model was calculated by first taking the ratio of the average LWIS paired with every other model.
+This was then diveded by the same ratio for the baseline forecasts (see methods in \cite{mathis2024evaluation} for details).
+ An RWIS less than 1 indicates the forecast outperformed the baseline forecast.
+ Figures \ref{fig:rwis_ili_tile} and \ref{fig:rwis_hosp_tile} show the mean RWIS of hospitalization forecasts over all 24 models from the main manuscript for all locations and weeks of the 2023 flu season.
+Some similar patterns to those noted in section the manuscript emerge, including a slightly better performance by forecasts including discrepancy modeling over those which do not include discrepancy modeling and better overall performance by the SIR models than by the ASG forecast models.
+Interestingly, figure \ref{fig:rwis_hosp_tile} shows the LNORM model showing especially poor RWIS performance about three quarters into the season.
+This was not seen in the LWIS in the main manuscript.
+%figure \ref{fig:lwis_by_dist_loc}.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.7]{Images/rwis_ili_mod.png}
+    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
+The plot is separated by the four ILI models.
+Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
+    \label{fig:rwis_ili_tile}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.6]{Images/rwis_hosp_mod.png}
+    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
+The plot is separated by the three distribution choices for the hospitalization model.
+Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
+    \label{fig:rwis_hosp_tile}
+\end{figure}
+
+Figure \ref{fig:state_median_points} shows the mean over weeks RWIS for the SIR and ASG forecast models with an RWIS less than 1 for the most locations.
+The plot on the left is from a forecast by an SIRD ILI model and a quadratic NORM hospitalization model.
+This forecast model had an RWIS less than 1 for 37 out of 53 locations.
+The plot on the right is from a forecast by an ASGD ILI model and linear NORM hospitalization model.
+This model had an RWIS less than 1 for 34 out of 53 locations.
+
+\begin{figure}
+\centering
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/sir_disc_hosp_sq_ar1_rwis.png}
+  % \caption{A subfigure}
+  % \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth]{Images/asg_disc_hosp_ar1_rwis.png}
+  % \caption{A subfigure}
+  % \label{fig:sub2}
+\end{subfigure}
+\caption{RWIS over the whole 2023 flu season for all 53 locations for hospitalization forecasts from an SIRD quadratic hospitalization model with normal errors (left) and forecasts from an ASGD linear hospitalization model with normal errors (right).
+An RWIS less than 1 (left of vertical grey line) indicates the model forecasts outperformed the baseline forecasts for that particular region.}
+\label{fig:state_median_points}
+\end{figure}
+\end{supplement}
+
+\bibliographystyle{ba}
+\bibliography{master_bib}
 

### 2024-12-10 nonlinear_forecast.tex
Use graphicspath

@@
 \RequirePackage{graphicx}
+
+\graphicspath{{Images/}}
 
@@
     \centering
-    \includegraphics[scale=.7]{Images/us_ili_seasons.png}
+    \includegraphics[scale=.7]{us_ili_seasons.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.7]{Images/ili_vs_week.png}
+    \includegraphics[scale=.7]{ili_vs_week.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
+    \includegraphics[scale=.5]{us_hospitalizations.png}
     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
@@
     \centering
-    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
+    \includegraphics[scale=.65]{hosp_vs_week.png}
     \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
@@
     \centering
-    \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
+    \includegraphics[scale=.6]{ili_vs_hosp.png}
     \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
@@
     \centering
-    \includegraphics[scale=.6]{Images/sir_traj.png}
+    \includegraphics[scale=.6]{sir_traj.png}
     \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
@@
     \centering
-    \includegraphics[scale=.45]{Images/asg_function.png}
+    \includegraphics[scale=.45]{asg_function.png}
     \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
@@
     \centering
-    \includegraphics[scale=.5]{Images/asg_fits.png}
+    \includegraphics[scale=.5]{asg_fits.png}
     \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
@@
     \centering
-    \includegraphics[scale=.45]{Images/discrepancy.png}
+    \includegraphics[scale=.45]{discrepancy.png}
     \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
@@
   \centering
-  \includegraphics[scale=.3]{Images/overall_crps.png}
+  \includegraphics[scale=.3]{overall_crps.png}
     % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
@@
   \centering
-  \includegraphics[scale=.3]{Images/overall_logs.png}
+  \includegraphics[scale=.3]{overall_logs.png}
     % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
@@
     \centering
-    \includegraphics[scale=.5]{Images/crps_by_season.png}
+    \includegraphics[scale=.5]{crps_by_season.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
@@
     \centering
-    \includegraphics[scale=.5]{Images/crps_by_week_horizon.png}
+    \includegraphics[scale=.5]{crps_by_week_horizon.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
     \centering
-    \includegraphics[scale=.5]{Images/logs_by_season.png}
+    \includegraphics[scale=.5]{logs_by_season.png}
     \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
@@
     \centering
-    \includegraphics[scale=.5]{Images/logs_by_week_horizon.png}
+    \includegraphics[scale=.5]{logs_by_week_horizon.png}
     \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
   \centering
-  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
+  \includegraphics[width=.49\linewidth,height=6cm]{normal_forecasts_us.png}
   % \caption{A subfigure}
@@
   \centering
-  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
+  \includegraphics[width=.49\linewidth, height=6cm]{normal_sq_forecasts_us.png}
   % \caption{A subfigure}
@@
     % \centering
-    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
-    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
-Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
-Scores for models with an ASG ILI model are above while those with an SIR model are below.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
+\includegraphics[scale = .5]{lwis_us_full_season.png}
+\caption{Each plot shows the log weighted interval score (LWIS, lighter is better) for every week of the 2023 flu season with scores for models with (bottom) and without (top) including discrepancy.
+Facets distinguish hospitalization distribution family and ILI as a linear or quadratic predictor (column facets, quadratic indicated with 2) with asymmetric Gaussian or SIR ILI models (row facets).}
     \label{fig:us_lwis}
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
+    \includegraphics[scale = .7]{lwis_by_traj_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
+    \includegraphics[scale = .7]{lwis_by_dist_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.

### 2024-12-10 nonlinear_forecast.tex
Push some changes

@@
 Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases.
-An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}.
+An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat.'' Prior to the 2021 season, the definition included ``without a known cause other than influenza'' \cite[]{cdcfluview2023}.
 Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza.
@@
     \includegraphics[scale=.5]{crps_by_season.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
+    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
     \label{fig:crps_by_season}
@@
 Here, a darker value represents a larger LWIS and worse forecast compared to lighter values.
-The weeks around the holiday week 22 are highlighted by a red band.
 In most cases, there appears to be a turning point in performance at or near week 22.
@@
     % \centering
-\includegraphics[scale = .5]{lwis_us_full_season.png}
+\includegraphics[scale = .5]{lwis_us_full_season}
 \caption{Each plot shows the log weighted interval score (LWIS, lighter is better) for every week of the 2023 flu season with scores for models with (bottom) and without (top) including discrepancy.
-Facets distinguish hospitalization distribution family and ILI as a linear or quadratic predictor (column facets, quadratic indicated with 2) with asymmetric Gaussian or SIR ILI models (row facets).}
+Facets distinguish hospitalization distribution family and ILI as a linear or quadratic predictor (column facets, quadratic indicated with 2) with asymmetric Gaussian or SIR ILI models (row facets).
+The weeks around the holiday week 22 are highlighted by a red band.}
     \label{fig:us_lwis}

### 2024-12-13 nonlinear_forecast.tex
update images and supplemental material

--- 2024-12-03 nonlinear_forecast.tex
@@
 \usepackage{tabularx} %for the spacing in tabular
+\usepackage{newtxmath} %for making indicator function
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author[A]{\fnms{Spencer Wadsworth}~\ead[label=e1]{sgw96@iastate.edu}}
+\author[A]{\snm{Spencer Wadsworth}~\ead[label=e1]{sgw96@iastate.edu}}
 % \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}\orcid{0000-0000-0000-0000}}
 \and
-\author[A]{~\snm{Jarad Niemi}\ead[label=e3]{niemi@iastate.edu}}
+\author[A]{~\snm{Jarad Niemi}\ead[label=e2]{niemi@iastate.edu}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \centering
-    \includegraphics[scale=.7]{Images/us_ili_seasons.png}
+    \includegraphics[scale=.5]{Images/us_ili_seasons.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
@@
 
-Figure \ref{fig:ili_vs_week} shows ILI data from five states and the District of Columbia, locations which received particular attention in \cite{osthus2021multiscale}.
-The plots include the ILI data for all seasons from 2010 to 2023 in grey, and the black line is the per week ILI average over seasons.
-The patterns in the individual states are similar to the national level plots in figure \ref{fig:us_ili} in that the ILI rises in the fall and winter until it peaks and descends as the spring and summer progress.
-For these locations ILI regularly peaks, either locally or globally, at or near week 22.
-
- \begin{figure}
-    \centering
-    \includegraphics[scale=.7]{Images/ili_vs_week.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
-Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.
-     Plots include lines for ILI\% from the 2010 flu season to 2023 (grey) and for the weekly ILI averaged over all seasons (black).}
-    \label{fig:ili_vs_week}
-\end{figure}
-
 \subsection{Hospitalization Data}
@@
 For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22.
-Figure \ref{fig:hosp_vs_week} shows the 2022 and 2023 weekly hospitalizations for the same states from \ref{fig:ili_vs_week}.
-Similar to the national data, the peak in 2022 came early compared to the peak of 2023.
+Additional plots of ILI data and hospitalization
+data for select states are included in the supplementary material 
+\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar to 
+those in figures \ref{fig:us_ili} and \ref{fig:us_hosp} with activity increasing
+to a peak and then decreasing.
 
@@
     \label{fig:us_hosp}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
-    \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
-    \label{fig:hosp_vs_week}
-\end{figure}
-
-Comparing figures \ref{fig:ili_vs_week} and \ref{fig:hosp_vs_week} shows that ILI and hospitalizations share the similar pattern of increasing to a peak in the winter and decreasing thereafter.
-Figure \ref{fig:ili_vs_hosp} shows scatter plots with ILI\% on the $x$-axes and hospitalizations on the $y$-axes, revealing a positive somewhat linear relationship between the two variables.
-This relationship motivates the forecast models outlined in the next section.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
-    \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
-    \label{fig:ili_vs_hosp}
 \end{figure}
@@
 The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as \cite{osthus2019dynamic} points out there may be systematic behavior that these or other possible functions may not capture.
-As noted in section \ref{sec:data}, figures \ref{fig:us_ili} and \ref{fig:ili_vs_week} show a regular peak at week 22 of the flu season.
+As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a regular peak at week 22 of the flu season.
 Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to illustrate the systematic discrepancy from a fitted function.
@@
 
-The second component for forecast modeling is the hospitalization model defined in (\ref{eq:final_hosp}).
-This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}.
-Here $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
+The second component for forecast modeling is based on the relationship between hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
+This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}. Figure \ref{fig:lag_scatter}
+shows scatterplots of the difference between hospitalizations and scaled 1 week 
+lags by ILI percentage at the US national level. The points in the left plot are
+from hospitalization differences whereas in the right the hospitalizations were
+log transformed. The positive and slightly curved relationship suggests model
+(\ref{eq:final_hosp}) is a reasonable choice.
+
+\begin{figure}
+    \includegraphics[scale=.46]{Images/lagdiff_vs_ili.png}
+    \caption{Scatterplots of national hospitalizations (left) or log hospitalizations (right) lag differences by ILI \% where the lag is scaled by $\phi$.
+Points are colored by season.
+For hospitalizations $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$.
+}
+    \label{fig:lag_scatter}
+\end{figure}
+
 \begin{equation}
@@
 \end{equation}
+In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
+
 This model is for any location, and for fitting purposes $ILI_{s,w}$ is always multiplied by $P$ which is proportional to the population of the state or territory, in this case the total population divided by 50,000.
@@
 The orientation of both scores is negative, meaning the smaller the score the better.
+Here $\vmathbb{1}\{\cdot\}$ is the indicator function.
 
 \begin{equation}
-    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \mathbb{1} (y^* \leq x))^2 dx
+    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
     \label{eq:crps2}
@@
 
-Figures \ref{fig:overall_scores} - \ref{fig:logs_by_week_horizon} show boxplots of the CRPS and LogS for the four models.
-Figure \ref{fig:overall_scores} shows that the variation of overall CRPS is smallest for the ASG models and larger for the SIR models.
-The median scores for the SIR models also appears slightly higher than for the ASG models.
-When faceted by season in figure \ref{fig:crps_by_season}, the boxplots of the CRPS often show the same pattern for ASG and SIR models but not always.
-For example, the bulk of CRPS values for the SIRD model in 2019 appears to have smaller variation than the other models.
-Figure \ref{fig:crps_by_week_horizon} shows CRPS boxplots faceted by week and horizon.
-Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, have CRPS values with lower medians and lower variation than the two models not accounting for ILI discrepancy.
+Figures \ref{fig:crps_by_week_horizon} and \ref{fig:logs_by_week_horizon} show 
+boxplots of results from the simulation. The figures respectively
+show boxplots of the CRPS and LogS for the 500 simulation replicates faceted by week and horizon.
+Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
+Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with lower medians and lower variation than the two models not accounting for ILI discrepancy.
 This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough.
+This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
+Among the four models, ASG tends to have the lowest values in CRPS and LogS though this is not always the case.
+This suggests that the additional flexibility of the five parameter ASG function is often better for forecasting than that of the three parameter SIR function.
+Additional
+plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and LogS.
 
@@
     \centering
-% \begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[scale=.3]{Images/overall_crps.png}
-    % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[scale=.3]{Images/overall_logs.png}
-    % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
-% \end{subfigure}
-\caption{Boxplots of the continuous ranked probability score (CRPS) (left) and logarithmic score (LogS) (right) for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-\label{fig:overall_scores}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/crps_by_season.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
-    \label{fig:crps_by_season}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/crps_by_week_horizon.png}
+    \includegraphics[scale=.4]{Images/crps_by_week_horizon.png}
     \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
@@
     \centering
-    \includegraphics[scale=.5]{Images/logs_by_season.png}
-    \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
-    \label{fig:logs_by_season}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/logs_by_week_horizon.png}
-    \caption{Boxplots of logarithmic (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
+    \includegraphics[scale=.4]{Images/logs_by_week_horizon.png}
+    \caption{Boxplots of the logarithmic score (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
 Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
@@
 \end{figure}
-
-Figures \ref{fig:logs_by_season} and \ref{fig:logs_by_week_horizon} show boxplots of the LogS for the simulated forecasts.
-Note that in these plots only values of less than 17 are included to improve visualization.
-The LogS plots show similar results to the CRPS, though there are some differences.
-For example, SIRD in figure \ref{fig:logs_by_season} tends to show smaller LogS variation relative to the other three models than is seen by the CRPS of SIRD in figure \ref{fig:crps_by_season}.
-We also note the smaller relative LogS variation than CRPS variation for SIRD when comparing figure \ref{fig:logs_by_week_horizon} with figure \ref{fig:crps_by_week_horizon}, particularly at week 14. 
-Among the four models, ASG tends to have the lowest values in CRPS and LogS though this is not always the case.
-When forecasting weeks 21-23, the two models which model discrepancy in the ILI, ASGD and SIRD, tend to outperform the models not modeling the discrepancy.
-This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
 
@@
 \label{eq:is}
-        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\mathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \mathbb{1}\{y^* > r\}
+        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\vmathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \vmathbb{1}\{y^* > r\}
 \end{equation}
@@
 Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$.
-Here $\mathbb{1}_{A}$ represents the indicator function for values within the set $A$.
+Here $\vmathbb{1}_{A}$ represents the indicator function for values within the set $A$.
 
@@
     \label{eq:sir_prior}
-        I_{0s} &\sim N(0.005, 0.03) \mathbb{1}_{(0, .1)} \\
+        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
         \beta_s &\sim N^+(0.8, 0.3) \\
@@
         \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-        \phi &\sim N(0, .4)\mathbb{1}_{(-1,1)} \\
+        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
         \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
-Scores are averaged over all horizons 1-4 weeks ahead.
-Scores are faceted by ILI model function (columns) and by whether or not discrepancy modeling was included (rows).
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:lwis_by_traj_loc}
-\end{figure}
-
-\begin{figure}
-
-    \centering
     \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
@@
 
-Figures \ref{fig:lwis_by_traj_loc} and \ref{fig:lwis_by_dist_loc} also show weekly forecast performance by LWIS, but all 53 locations are included.
-Overall, the weeks leading up to week 22 are the most difficult to forecast.
-In figure \ref{fig:lwis_by_traj_loc}, it appears that both the SIR and ASG models which include discrepancy perform slightly better than the models which do not.
-Figure \ref{fig:lwis_by_dist_loc} shows the LWIS across the whole season faceted by hospitalization model distribution.
-The LNORM model appears to perform the best during the weeks leading up to week 22.
-The season overall score, calculated as the mean LWIS over all locations and weeks, for all 24 models is shown in table \ref{tab:forecast_scores}.
+Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by LWIS, but all 53 locations are included.
+The LWIS is shown across the whole season faceted by hospitalization model distribution.
+Overall, the weeks leading up to week 22 are the most difficult to forecast, but the LNORM model appears to perform the better than the other models during the weeks around the holiday week 22.
+The overall, calculated as the mean LWIS over all locations and weeks, for all 24 models is shown in table \ref{tab:forecast_scores}.
 The first main takeaway is that the top four performing models are SIR ILI models with the LNORM hospitalization models.
@@
 As suggested by the simulation study in the previous section, it may often be the case that the ASG ILI model is better for forecasting.
-Indeed, a close examination of figures \ref{fig:us_lwis}, \ref{fig:lwis_by_traj_loc}, and \ref{fig:lwis_by_dist_loc} shows that it is common for the ASG models to show better forecasting skill than the SIR models.
+Indeed, a close examination of figures \ref{fig:us_lwis} and \ref{fig:lwis_by_dist_loc} shows that it is common for the ASG models to show better forecasting skill than the SIR models.
 

### 2024-12-13 supplemental.tex
update images and supplemental material

@@
 \begin{supplement}
+
+\section{ILI \& Hospitalization data for some regions}
+
+Figure \ref{fig:ili_vs_week} shows ILI data from five states and the District of Columbia, locations which received particular attention in \cite{osthus2021multiscale}.
+The plots include the ILI data for all seasons from 2010 to 2023 in grey, and the black line is the per week ILI average over seasons.
+The patterns in the individual states are similar to the national level plots in that the ILI rises in the fall and winter until it peaks and descends as the spring and summer progress.
+For these locations ILI regularly peaks, either locally or globally, at or near week 22.
+
+ \begin{figure}[hbt!]
+    \centering
+    \includegraphics[scale=.7]{Images/ili_vs_week.png}
+    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
+Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.
+     Plots include lines for ILI\% from the 2010 flu season to 2023 (grey) and for the weekly ILI averaged over all seasons (black).}
+    \label{fig:ili_vs_week}
+\end{figure}
+
+Figure \ref{fig:hosp_vs_week} shows the 2022 and 2023 weekly hospitalizations for the same states from \ref{fig:ili_vs_week}.
+Similar to the national data, the peak in 2022 came early compared to the peak of 2023.
+Comparing figures \ref{fig:ili_vs_week} and \ref{fig:hosp_vs_week} shows that ILI and hospitalizations share the similar pattern of increasing to a peak in the winter and decreasing thereafter.
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
+    \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
+    \label{fig:hosp_vs_week}
+\end{figure}
+
+% Figure \ref{fig:ili_vs_hosp} shows scatter plots with ILI\% on the $x$-axes and hospitalizations on the $y$-axes, revealing a positive somewhat linear relationship between the two variables.
+This relationship motivates the forecast models outlined in the next section.
+% 
+% \begin{figure}[hbt!]
+%     \centering
+%     \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
+%     \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
+%     \label{fig:ili_vs_hosp}
+% \end{figure}
+
+\section{Additional results plots from simulation study}
+
+Figures \ref{fig:overall_scores} - \ref{fig:logs_by_season} show boxplots of the CRPS and LogS for the four models.
+Figure \ref{fig:overall_scores} shows that the variation of overall CRPS is smallest for the ASG models and larger for the SIR models.
+The median scores for the SIR models also appears slightly higher than for the ASG models.
+When faceted by season in figure \ref{fig:crps_by_season}, the boxplots of the CRPS often show the same pattern for ASG and SIR models but not always.
+For example, the bulk of CRPS values for the SIRD model in 2019 appears to have smaller variation than the other models.
+The LogS plots show similar results to the CRPS plots, though there are some differences.
+For example, SIRD in figure \ref{fig:logs_by_season} tends to show smaller LogS variation relative to the other three models than is seen by the CRPS of SIRD in figure \ref{fig:crps_by_season}.
+
+\begin{figure}[hbt!]
+\centering
+% \begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[scale=.3]{Images/overall_crps.png}
+    % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
+% \end{subfigure}%
+% \begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[scale=.3]{Images/overall_logs.png}
+    % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
+% \end{subfigure}
+\caption{Boxplots of the continuous ranked probability score (CRPS) (left) and logarithmic score (LogS) (right) for the four ILI models over all seasons, weeks, and horizons in the simulation study}
+\label{fig:overall_scores}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[scale=.5]{Images/crps_by_season.png}
+    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
+    \label{fig:crps_by_season}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[scale=.5]{Images/logs_by_season.png}
+    \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
+    \label{fig:logs_by_season}
+\end{figure}
+
+\section{Additional results plots for 2023-24 flu analysis}
+
+In figure \ref{fig:lwis_by_traj_loc}, it appears that both the SIR and ASG models which include discrepancy perform slightly better than the models which do not.
+
+\begin{figure}[hbt!]
+
+    \centering
+    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
+    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
+Scores are averaged over all horizons 1-4 weeks ahead.
+Scores are faceted by ILI model function (columns) and by whether or not discrepancy modeling was included (rows).
+The lighter the shade, the lower the LWIS with low LWIS being better.}
+    \label{fig:lwis_by_traj_loc}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
     \centering
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
 \centering

### 2024-12-13 functions_data/get_data_functions_new.R

@@ -107,7 +107,9 @@ get_ili_data <- function(lag = 2) {
   pops <- read.csv("https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/locations.csv")
   # colnames(pops)[1] <- "region"
   pops <- pops %>% 
-    select(location_name, population, count_rate1, count_rate2, count_rate3,
+    select(location_name, population, count_rate1, 
+           # count_rate2, 
+           count_rate3,
            count_rate4, count_rate5)
   colnames(pops)[1] <- "region"
   
@@ -128,7 +130,9 @@ comb_data <- function(lag = 2, s_region = NA, s_season = 2022) {
   
   ILINet <- get_ili_data(lag = lag) %>% 
     select(region, unweighted_ili, week_start, location_name, season, place,
-           season_week, population, count_rate1, count_rate2, count_rate3,
+           season_week, population, count_rate1, 
+           # count_rate2, 
+           count_rate3,
            count_rate4, count_rate5) %>%
     select(-place)
   

### 2024-12-13 make_images/make_images_for_data_section.R

@@ -1,7 +1,7 @@
 
-setwd("~/flu_research/nonlinear_flu_forecast/forecasts_23")
-source('~/flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions_new.R')
-source('~/flu_research/nonlinear_flu_forecast/mle_functions.R')
+setwd("~/flu_research/nonlinear_flu_forecast/")
+source('~/flu_research/nonlinear_flu_forecast/functions_data/get_data_functions_new.R')
+source('~/flu_research/nonlinear_flu_forecast/functions_data/mle_functions.R')
 
 
 
@@ -193,6 +193,31 @@ get_fit_flu <- function(x, y, square = FALSE) {
   return(exp(preds) - 1)
 }
 
+
+
+get_lag_fits <- function(x, y, square = FALSE, log = FALSE) {
+  x2 <- x^2
+  x3 <- x^3
+  if (log == TRUE) {y <- log(y + 1)}
+  if (square == FALSE) {
+    mod <- lm(y ~ lag(y) + x)
+    ar1 <- coef(mod)[2]
+  }
+  # else if (square == 3) {
+  #   mod <- lm(y ~ x + x2 + x3)
+  #   preds <- predict(mod)
+  # }
+  else {
+    mod <- lm(y ~ lag(y) + x + x2)
+    mod <- lm(y ~ lag(y) + x)
+    ar1 <- coef(mod)[2]
+  }
+  # ar1 <- .64
+  lag_val <- y - ar1*lag(y)
+  print(ar1)
+  return(lag_val)
+}
+
 both_flu1 <- comb_data(lag = 1)
 both_flu2 <- comb_data(lag = 2)
 
@@ -227,9 +252,56 @@ both_flu2 %>%
   ggplot(aes(x  = unweighted_ili, y = value)) +
   geom_point()
   
+both_test <- both_flu1 %>% 
+  filter(region == "US", season == 2023) %>% 
+  mutate(lag_val = lag(value), uili2 = unweighted_ili^2) %>% 
+  mutate(ar1_val = value- .64*lag_val)
+mod <- lm(value ~ lag_val + unweighted_ili, data = both_test)
+
+plot(ar1_val ~ unweighted_ili^2, data = both_test)
+plot(value ~ unweighted_ili, data = both_test)
+
+
+library(latex2exp)
+both_flu1 %>% 
+  filter(region %in% c("US"),
+         # c("Idaho", "Utah", "Oregon", "Minnesota", "California",
[…]

### 2024-12-13 nonlinear_forecast.tex
update figures and formatting

@@
 \begin{supplement}
-\stitle{1. Posterior distribution plots for select parameters}
-\sdescription{Plots of posterior distributions for parameters in both ILI and hospitalization models}
+Supplement to "Forecasting Influenza Hospitalizations Using a
+Bayesian Hierarchical Nonlinear Model with
+Discrepancy"\\
+1. Additional plots for ILI \& Hospitalization data for select regions, results from simulation study, and 2023-24 flu analysis\\
+% \end{supplement}
+2. Posterior distribution plots for select parameters\\
+% \sdescription{2. Plots of posterior distributions for parameters in both ILI and hospitalization models}
+3. FluSight forecast competition scoring results
 \end{supplement}
-\begin{supplement}
-\stitle{2. FluSight forecast competition scoring results}
-\sdescription{Flu forecast results with relative weighted interval scoring rule similar to that used in  FluSight}
-\end{supplement}
+% \begin{supplement}
+% \stitle{3. FluSight forecast competition scoring results}
+% \sdescription{Flu forecast results with relative weighted interval scoring rule similar to that used in  FluSight}
+% \end{supplement}
 

### 2024-12-13 supplemental.tex
update figures and formatting

@@
 
-\section{ILI \& Hospitalization data for some regions}
+\section{Additional plots for ILI \& Hospitalization data for select regions,
+         results from simulation study, and 2023-24 flu analysis}
 
@@
 
-\section{Additional results plots from simulation study}
+% \section{Additional results plots from simulation study}
 
@@
 
-\section{Additional results plots for 2023-24 flu analysis}
+% \section{Additional results plots for 2023-24 flu analysis}
 

### 2024-12-18 nonlinear_forecast/manuscript.tex
update for submissions

--- 2024-12-13 nonlinear_forecast.tex
@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
+\documentclass{article}
 
-%% Packages
-\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
-\RequirePackage[numbers]{natbib}
-%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
-\RequirePackage{graphicx}
+\usepackage{PRIMEarxiv}
 
-%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
+\usepackage[utf8]{inputenc} % allow utf-8 input
+% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage[linktocpage=true]{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+% \usepackage{lipsum}
+\usepackage{fancyhdr}       % header
+\usepackage{graphicx}       % graphics
+\graphicspath{{media/}}     % organize your images and other figures under media/ folder
+
+% Packages added by Spencer
+\usepackage{amsmath,amsthm,amssymb,scrextend}
+\usepackage{appendix}
+\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}
+\usepackage{enumerate}
+\usepackage{csquotes}
+% \usepackage{subcaption}
+\usepackage{amssymb}
+% \usepackage{bbm}
+\usepackage[bb=pazo]{mathalpha}
+\usepackage{bm}
+\usepackage{newtxmath} %for making indicator function
 \usepackage{booktabs} %for the table in the analysis section
-\usepackage{enumerate} %for the short algorithm
+% \usepackage{enumerate} %for the short algorithm
 \usepackage{subfigure} %for making plots with multiple images
 \usepackage{tabularx} %for the spacing in tabular
-\usepackage{newtxmath} %for making indicator function
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\pubyear{2024}
-\arxiv{2010.00000}
-\volume{TBA}
-\issue{TBA}
-\firstpage{1}
-\lastpage{1}
+\pagestyle{empty}
 
-\startlocaldefs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Uncomment next line to change            %%
-%% the type of equation numbering           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\numberwithin{equation}{section}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Axiom, Claim, Corollary, Hypothesis, %%
-%% Lemma, Theorem, Proposition              %%
-%% use \theoremstyle{plain}                 %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{plain}
-\newtheorem{axiom}{Axiom}
-\newtheorem{claim}[axiom]{Claim}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Assumption, Definition, Example,     %%
-%% Notation, Property, Remark, Fact         %%
-%% use \theoremstyle{definition}            %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem*{example}{Example}
-\newtheorem*{fact}{Fact}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Case use \theoremstyle{remark}       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{remark}
-\newtheorem{case}{Case}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Please put your definitions here:        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\endlocaldefs
+\newcommand\Bbbbone{%
+  \ifdefined\mathbbb%
+    \mathbbb{1}%
+  \else%
+    \boldsymbol{\mathbb{1}}%
+  \fi}
+
+% \usepackage{xr}
+% \externaldocument{supplementary}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \usepackage[pdftex]{graphicx}
+% Standard, old-style thesis
+% \usepackage{isutraditional}   
+% \chaptertitle
+% Old-style, thesis numbering down to subsubsection
+% \alternate
+% \usepackage{rotating}
+% Bibliography without numbers or labels
+\usepackage{natbib}
+% Use the following line if you want square brackets and numbering system
+%\usepackage[square, numbers]{natbib}
+
+% this package is for getting certain math scripts like the 1 needed for an indicator function
+% \usepackage{bbm}
+
+% using this package to make a better looking list
+\usepackage{enumerate}
+% try this one for a list as well. may be better
+% \usepackage[font=itshape]{quoting}
+
+% package for tables
+
+% \renewcommand{\bibfont}{\setstretch{1}\selectfont}
+% \setlength{\bibsep}{13.2pt}
+
+\usepackage{chapterbib}
+%\renewcommand\bibsection{\section*{\ref name}}
+%\renewcommand\bibsection{\section*}
+\renewcommand{\bibsection}{\section{References}}
+
+%\bibliographystyle{apa}  % changed from apa, abbrv, unsrtnat
+%\includeonly{titletoc,chapter1}
+%Optional Package to add PDF bookmarks and hypertext links
+% \usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
+\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
+
+%\usepackage{titletoc}
+% \usepackage{hyperref}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Packages/commands added by Jarad
+\usepackage{xcolor}
+\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
+\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}
+
+%Header
+\pagestyle{fancy}
+\thispagestyle{empty}
+\rhead{ \textit{ }}
+
+% Update your Headers here
+\fancyhead[LO]{Bayesian nonlinear flu hospitalizations forecasting}
+% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}
+
+%% Title
+\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy
+%%%% Cite as
+%%%% Update your official citation here when published 
+% \thanks{\textit{\underline{Citation}}: 
+% \textbf{Authors. Title. Pages.... DOI:000000/11111.}} 
+}
+
+\author{
+  Spencer Wadsworth \\
+  Iowa State University \\
+  Ames, Iowa\\
+  \texttt{sgw96@iastate.edu} \\
+  %% examples of more authors
+   \And
+  Jarad Niemi \\
+  Iowa State University \\
+  Ames, Iowa\\
+  \texttt{niemi@iastate.edu} \\
+  %% \AND
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+}
 
 \begin{document}
-
-\begin{frontmatter}
-\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy}
-%\title{A sample article title with some additional note\thanksref{t1}}
-\runtitle{Bayesian nonlinear flu hospitalizations forecasting}
-%\thankstext{T1}{A sample additional note to the title.}
-
-\begin{aug}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Only one address is permitted per author. %%
-%% Only division, organization and e-mail is %%
-%% included in the address.                  %%
-%% Additional information can be included in %%
-%% the Acknowledgments section if necessary. %%
-%% ORCID can be inserted by command:         %%
-%% \orcid{0000-0000-0000-0000}               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author[A]{\snm{Spencer Wadsworth}~\ead[label=e1]{sgw96@iastate.edu}}
-% \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}\orcid{0000-0000-0000-0000}}
-\and
-\author[A]{~\snm{Jarad Niemi}\ead[label=e2]{niemi@iastate.edu}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Addresses                                %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\address[A]{Department of Statistics,
-Iowa State University, Ames, IA 50011}
-
-\runauthor{Wadsworth and Niemi}
-\end{aug}
-
+\maketitle
 \begin{abstract}
@@
 For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor.
-In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from \cite{osthus2019dynamic} and a nonlinear Bayesian hierarchical model from \cite{ulloa2019} are compared.
+In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from Osthus et al.
+% \cite[]{osthus2019dynamic} 
+% 
+and a nonlinear Bayesian hierarchical model from Ulloa 
+% \cite[]{ulloa2019}
+are compared.
 Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model.
@@
 We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year.
-We also found that other modeling decisions such as the exact nonlinear function to use in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
+We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
 \end{abstract}
+
+% \begin{abstract}
+% The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread.
+The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods.
+Beginning in 2013, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalizations.
+Reliable hospitalization data has only been available since 2021, but ILI data has been available since 2010 and has been successfully forecast for several seasons.
+In this manuscript, we introduce a two component modeling framework for forecasting hospitalizations utilizing both hospitalization and ILI data.
+The first component is for modeling ILI data using a nonlinear Bayesian model.
+The second component is for modeling hospitalizations as a function of ILI.
+For hospitalization forecasts, ILI is first forecast then hospitalizations are forecast with ILI forecasts used as a predictor.
+In a simulation study, the hospitalization forecast model is assessed and two previously successful ILI forecast models are compared.
+Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model.
+Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared.
+We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year.
+We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
+% \end{abstract}
 
@@
 % \end{keyword}
-
-\begin{keyword}
-\kwd{Disease outbreak forecasting}
-\kwd{Bayesian hierarchical modeling}
-\kwd{Probabilistic forecasting}
-\kwd{Model discrepancy}
-\end{keyword}
-
-\end{frontmatter}
 
@@
 Every year the seasonal influenza outbreak burdens the public health system by infecting millions, causing an influx of primary care visits and hospitalizations and leading to between 290,000 and 650,000 deaths worldwide \cite[]{whoflufact2023}.
-\cite{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion.
+Molinari et al.
+\cite[]{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion.
 Accurate forecasting of infectious diseases can inform public decision making and ease the burden of an outbreak \cite[]{turtle2021accurate, lutz2019applying}.
@@
 ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}.
-The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, \cite{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework.
+The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, Osthus et al.
+\cite[]{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework.
 An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
@@
 The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
-Herein we use ILI models similar to those in \cite{osthus2019dynamic} and \cite{ulloa2019} for ILI forecasting.
-The model of \cite{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework.
-Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by \cite{kennedy2001bayesian}.
-The model in \cite{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data.
+Herein we use ILI models similar to those in Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
+The model of Osthus et al.
+\cite[]{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework.
+Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by Kennedy and O'Hagan \cite[]{kennedy2001bayesian}.
+The model in Ulloa \cite[]{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data.
 Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.
@@
 In section \ref{sec:functions} the modeling framework contributed by this manuscript is given.
-In the same section, functions similar to those used by \cite{osthus2019dynamic} and \cite{ulloa2019} are defined.
+In the same section, functions similar to those used by Osthus et al.
+\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} are defined.
  These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively.
 Model fitting and implementation are described at the end of the section.
-Section \ref{sec:simulation2} is a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared.
+Section \ref{sec:simulation2} includes a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared.
 Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts.
-Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}, with the analysis being done using the conventions of FluSight.
+Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}.
  Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.
@@
  \section{Flu outbreak data} \label{sec:data}
-In this section we introduce and define ILI and hospitalization data and evaluate the data visually.
+In this section we introduce and define ILI and hospitalization data and visually evaluate the data.
 ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season.
@@
 Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza.
-To know whether or not a sick patient is infected with influenza would require a laboratory test.
+To know whether or not a sick patient is infected with influenza requires a laboratory test.
 
@@
 
-Figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023.
+The top of figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023.
 For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons.
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
     \centering
     \includegraphics[scale=.5]{Images/us_ili_seasons.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
+
+    \centering
+    \includegraphics[scale=.4]{Images/us_hospitalizations.png}
+    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023 (top) and weekly flu confirmed hospitalization counts at the national level for 2022 and 2023 flu seasons (bottom).
 Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.}
@@
 
-Notable from the plots in figure \ref{fig:us_ili} is the regular trajectory of the ILI.
+Notable from the plots in the top of figure \ref{fig:us_ili} is the regular trajectory of the ILI.
 With the exception of season 2020, the ILI begins low at week 1 and increases as the fall and winter progress until the ILI reaches a peak.
 As spring progresses to summer, the ILI decreases to low values.
-As \cite{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day.
+As Osthus et al.
+\cite[]{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day.
 Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
@@
 
-Figure \ref{fig:us_hosp} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons.
-These plots show similarities to the ILI plots in \ref{fig:us_ili} in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends.
+The bottom of figure \ref{fig:us_ili} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons.
+These plots show similarities to the ILI plots in the top of the figure in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends.
 For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22.
@@
 \cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar to 
-those in figures \ref{fig:us_ili} and \ref{fig:us_hosp} with activity increasing
+those in figure \ref{fig:us_ili} with activity increasing
 to a peak and then decreasing.
 
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
-    \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
-    \label{fig:us_hosp}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[scale=.5]{Images/us_hospitalizations.png}
+%     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
+%     \label{fig:us_hosp}
+% \end{figure}
 
@@
 
-The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows of a similar trajectory for modeling ILI.
-Compartmental models are standard mathematical models used for disease outbreaks.
+The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows a similar trajectory.
+Compartmental models are standard mathematical models used for modelling disease outbreaks.
 One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}.
-\cite{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data.
+Ulloa \cite[]{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data.
 The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy.
-In the first part of this section, we present an ILI model similar to the model in \cite{osthus2019dynamic}.
+In the first part of this section, we present an ILI model similar to the model in Osthus et al.
+\cite[]{osthus2019dynamic}.
 With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.
@@
 \end{equation}
-In \cite{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}.
-In \cite{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}.
-In \cite{ulloa2019} modeling is done hierarchically over seasons.
+In Osthus et al.
+\cite[]{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}.
+In Ulloa \cite[]{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}.
+In Ulloa \cite[]{ulloa2019} modeling is done hierarchically over seasons.
 
@@
 
-The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by \cite{kermack1927contribution}.
+The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by Kermack and McKendrick \cite[]{kermack1927contribution}.
 Since then, compartmental models have become standard for modeling infectious diseases \cite[]{allen2008mathematical}, and many extensions have been made and studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}.
@@
 The trajectory is determined by the disease transmission rate $\beta > 0$ and the recovery rate $\delta > 0$.
-Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual, and the expected rate of recovery to an immune state for a newly infected person.
+Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual and the expected rate of recovery to an immune state for a newly infected person.
 Whether or not a disease outbreak is classified as an epidemic is determined by the initial susceptible population $S_0$, or the susceptible population at time $0$, and the parameter $\rho = \delta/
@@
 It is non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}.
-Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where the $S_0$ and $\rho$ were selected to match an outbreak that would classify as epidemic.
+Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where $S_0$ and $\rho$ were selected to match an outbreak that is an epidemic.
 In the case where $S_0 \leq \rho$, the trajectory for the $I$ compartment would never be increasing.
@@
 
-The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as \cite{osthus2019dynamic} points out there may be systematic behavior that these or other possible functions may not capture.
+The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as Osthsu et al.
+\cite[]{osthus2019dynamic} point out there may be systematic behavior in the ILI that these or other possible functions may not capture.
 As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a regular peak at week 22 of the flu season.
@@
 Thus, care must be taken in setting parameter constraints as well as in the selection of prior distributions \cite[]{osthus2019dynamic,brynjarsdottir2014learning}. 
-Modeling of the discrepancy for ILI was done by \cite{osthus2019dynamic} during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}.
+Modeling of the discrepancy for ILI was done by Osthus et al.
+during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}.
 As in Osthus's model, $\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}).
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
     \includegraphics[scale=.46]{Images/lagdiff_vs_ili.png}
@@
 
-The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in \cite{osthus2019dynamic} and \cite{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem.
+The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in Osthus et al.
+\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem.
 For model (\ref{eq:ili_model}), parameters which are common even when using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, $\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$.
@@
 The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by \cite{gelman2006prior}, the priors were assigned to the variance parameters.
+For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by Gelman \cite[]{gelman2006prior}, the priors were assigned to the variance parameters.
 Univariate parameters were assigned either a normal distribution prior if the support is on $\mathbb{R}$, a half-normal prior if the support is nonnegative, or a truncated-normal prior to match a more specific support.
@@
 Additional prior constraints were made to improve parameter identifiability.
-In \cite{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$.
+In Osthus et al.
+\cite[]{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$.
 The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative priors.
@@
 Fitting the rest of the model parameters involves assigning priors and conditioning on the previously fit parameters.
-This has been done in computer modeling to improve identifiability and other issues, though \cite{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}.
+This has been done in computer modeling to improve identifiability and other issues, though Liu et al.
+\cite[]{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}.
 We carried out the modular fit by first estimating the maximum likelihood estimate (MLE) for the parameter $\lambda_s$ for each season in (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
@@
 
-The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the \cite{stan2024manual} \cite[]{gabry2022stan}.
+The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the Stan Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}.
 Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}.
@@
 
-Model comparison was done by calculating the continuous ranked probability score (CRPS) and the logarithmic score (LogS) for each forecast.
-These scores are both proper scoring rules which evaluate the forecast distribution and density functions respectively.
+Model comparison was done by calculating the continuous ranked probability score (CRPS) for each forecast.
+The CRPS is a proper scoring rule which evaluates the forecast distribution function.
 Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}.
 Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The CRPS is defined in (\ref{eq:crps2}) and the LogS in (\ref{eq:logs2}).
-Here $F(\cdot)$ is the forecast distribution function, $f(\cdot)$ is the forecast density function, and $y^*$ is the observed targeted value of the forecast.
-The orientation of both scores is negative, meaning the smaller the score the better.
+The CRPS is defined in (\ref{eq:crps2}).
+Here $F(\cdot)$ is the forecast distribution function, and $y^*$ is the observed targeted value of the forecast.
+The orientation of the CRPS is negative, meaning the smaller the score the better.
 Here $\vmathbb{1}\{\cdot\}$ is the indicator function.
+The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
 
@@
 
-\begin{equation}
-    \text{LogS}(f, y^*) = -\text{log}(f(y^*))
-    \label{eq:logs2}
-\end{equation}
-Both the CRPS and LogS are calculated using the \texttt{scoringRules} package in \texttt{R} \cite[]{jordan2019scoring}.
-To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation.
-The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
+% \begin{equation}
+%     \text{LogS}(f, y^*) = -\text{log}(f(y^*))
+%     \label{eq:logs2}
+% \end{equation}
+% Both the CRPS and LogS are calculated using the \texttt{scoringRules} package in \texttt{R} \cite[]{jordan2019scoring}.
+% To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation.
 
-Figures \ref{fig:crps_by_week_horizon} and \ref{fig:logs_by_week_horizon} show 
-boxplots of results from the simulation. The figures respectively
-show boxplots of the CRPS and LogS for the 500 simulation replicates faceted by week and horizon.
-Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
+Figure \ref{fig:crps_by_week_horizon} shows 
+boxplots of the CRPS for the 500 simulation replicates faceted by week and horizon. 
+% Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
 Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with lower medians and lower variation than the two models not accounting for ILI discrepancy.
@@
 This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
-Among the four models, ASG tends to have the lowest values in CRPS and LogS though this is not always the case.
+Among the four models, ASG tends to have the lowest CRPS values though this is not always the case.
 This suggests that the additional flexibility of the five parameter ASG function is often better for forecasting than that of the three parameter SIR function.
 Additional
-plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and LogS.
+plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and another common proper scoring rule known as the logarithmic score.
 
-\begin{figure}
+\begin{figure}[hbt!]
     \centering
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{Images/logs_by_week_horizon.png}
-    \caption{Boxplots of the logarithmic score (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
-    \label{fig:logs_by_week_horizon}
-\end{figure}
-
 \section{Analysis of forecasts for 2023 flu season}
@@
 In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations.
-The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$.
+The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$.
 A single forecast is thus comprised of 11 predictive intervals and a median.
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
 \centering
-\begin{subfigure}
+% \begin{subfigure}
   \centering
@@
   % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}
+% \end{subfigure}%
+% \begin{subfigure}
   \centering
@@
   % \label{fig:sub2}
-\end{subfigure}
+% \end{subfigure}
 \caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
@@
 LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling.
-Here, a darker value represents a larger LWIS and worse forecast compared to lighter values.
-The weeks around the holiday week 22 are highlighted by a red band.
-In most cases, there appears to be a turning point in performance at or near week 22.
-The models which include discrepancy appear to forecast better around week 22, though they may not outperform the models without discrepancy through the whole season.
-Such is the case for the NORM and LST quadratic models and the linear LNORM models.
-It's also notable that for the SIR model, including discrepancy appears to make for poorer forecasts after week 22 whereas for ASG, including discrepancy appears to improve forecasts.
+Here, the smaller the LWIS the better.
+The weeks around the holiday week 22 are highlighted by a grey band.
+In several cases, there appears to be a turning point in performance at or near week 22.
+For most ASG models, the models which include discrepancy tend to outperform those which do not beginning near week 22 and for the rest of the season.
+For the SIR models, the discrepancy does not always lead to better forecasts, but in for the LNORM hospitalization models, including discrepancy greatly improves forecasts around the holiday week 22.
+The LNORM LWIS scores tend to be higher than for LST and NORM hospitalization models, but figure \ref{fig:lwis_by_dist_loc} and table \ref{tab:forecast_scores} suggest that the LNORM hospitalization model outperforms LST and NORM for most states and territories.
 
-\begin{figure}
+\begin{figure}[hbt!]
 
     % \centering
-    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
-    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
+    \includegraphics[scale = .6]{Images/lwis_ln_us_full_season.png}
+    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including and exluding discrepancy in the ILI model.
+Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
+Scores for models with an ASG ILI model are above while those with an SIR model are below.
+The lower the LWIS the better the forecast.}
+    \label{fig:us_lwis}
+\end{figure}
+
+% \begin{figure}
+%     
+%     % \centering
+%     \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
+%     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
 Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
@@
 The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:us_lwis}
-\end{figure}
+%     \label{fig:us_lwis}
+% \end{figure}
 
-\begin{figure}
+\begin{figure}[hbt!]
 
@@
  In this manuscript we introduce a statistical modeling framework which allows for the incorporation of several ILI forecast modeling methods.
-Specifically, we built upon \cite{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy.
+Specifically, we built upon Osthus et al.
+\cite[]{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy.
 We model flu hospitalizations by incorporating the ILI forecast model into a model forecasting hospitalizations where hospitalization predictions are a linear or quadratic function of ILI.
@@
 
-\begin{acks}[Acknowledgments]
+\section{Acknowledgments}
 This work is partially supported by the National Science Foundation under Grant No.
@@
 Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
-\end{acks}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Funding information, if any,             %%
-%% should be provided in the                %%
-%% funding section.                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{funding}
-% The first author was supported by NSF Grant DMS-??-??????.
-% 
-% 
-% \end{funding}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Supplementary Material, including data   %%
-%% sets and code, should be provided in     %%
-%% {supplement} environment with title      %%
-%% and short description. It cannot be      %%
-%% available exclusively as external link.  %%
-%% All Supplementary Material must be       %%
-%% available to the reader on Project       %%
-%% Euclid with the published article.       %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{supplement}
+\section{Supplementary Material}
 Supplement to "Forecasting Influenza Hospitalizations Using a
@@
 Discrepancy"\\
-1. Additional plots for ILI \& Hospitalization data for select regions, results from simulation study, and 2023-24 flu analysis\\
+A. Additional plots for ILI \& Hospitalization data for select regions, results from simulation study, and 2023-24 flu analysis\\
 % \end{supplement}
-2. Posterior distribution plots for select parameters\\
+B. Posterior distribution plots for select parameters\\
 % \sdescription{2. Plots of posterior distributions for parameters in both ILI and hospitalization models}
-3. FluSight forecast competition scoring results
-\end{supplement}
-% \begin{supplement}
-% \stitle{3. FluSight forecast competition scoring results}
-% \sdescription{Flu forecast results with relative weighted interval scoring rule similar to that used in  FluSight}
-% \end{supplement}
+C. FluSight forecast competition scoring results
 
-\bibliographystyle{ba}
+%Bibliography
+% \bibliographystyle{unsrt}  
+% \bibliography{references}
+
+\bibliographystyle{apa}
 \bibliography{master_bib}

### 2024-12-18 nonlinear_forecast/supplement.tex
update for submissions

--- 2024-12-13 supplemental.tex
@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
+\documentclass{article}
 
-%% Packages
-\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
-\RequirePackage[numbers]{natbib}
-%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
-\RequirePackage{graphicx}
+\usepackage{PRIMEarxiv}
 
-%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
+\usepackage[utf8]{inputenc} % allow utf-8 input
+% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage[linktocpage=true]{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+% \usepackage{lipsum}
+\usepackage{fancyhdr}       % header
+\usepackage{graphicx}       % graphics
+\graphicspath{{media/}}     % organize your images and other figures under media/ folder
+
+% Packages added by Spencer
+\usepackage{amsmath,amsthm,amssymb,scrextend}
+\usepackage{appendix}
+\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}
+\usepackage{enumerate}
+\usepackage{csquotes}
+% \usepackage{subcaption}
+\usepackage{amssymb}
+% \usepackage{bbm}
+\usepackage[bb=pazo]{mathalpha}
+\usepackage{bm}
+\usepackage{newtxmath} %for making indicator function
 \usepackage{booktabs} %for the table in the analysis section
-\usepackage{enumerate} %for the short algorithm
+% \usepackage{enumerate} %for the short algorithm
 \usepackage{subfigure} %for making plots with multiple images
 \usepackage{tabularx} %for the spacing in tabular
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\pubyear{2024}
-\arxiv{2010.00000}
-\volume{TBA}
-\issue{TBA}
-\firstpage{1}
-\lastpage{1}
+\pagestyle{empty}
 
-\startlocaldefs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Uncomment next line to change            %%
-%% the type of equation numbering           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\numberwithin{equation}{section}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Axiom, Claim, Corollary, Hypothesis, %%
-%% Lemma, Theorem, Proposition              %%
-%% use \theoremstyle{plain}                 %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{plain}
-\newtheorem{axiom}{Axiom}
-\newtheorem{claim}[axiom]{Claim}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Assumption, Definition, Example,     %%
-%% Notation, Property, Remark, Fact         %%
-%% use \theoremstyle{definition}            %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem*{example}{Example}
-\newtheorem*{fact}{Fact}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Case use \theoremstyle{remark}       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{remark}
-\newtheorem{case}{Case}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Please put your definitions here:        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\endlocaldefs
+\newcommand\Bbbbone{%
+  \ifdefined\mathbbb%
+    \mathbbb{1}%
+  \else%
+    \boldsymbol{\mathbb{1}}%
+  \fi}
+
+% \usepackage{xr}
+% \externaldocument{supplementary}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \usepackage[pdftex]{graphicx}
+% Standard, old-style thesis
+% \usepackage{isutraditional}   
+% \chaptertitle
+% Old-style, thesis numbering down to subsubsection
+% \alternate
+% \usepackage{rotating}
+% Bibliography without numbers or labels
+\usepackage{natbib}
+% Use the following line if you want square brackets and numbering system
+%\usepackage[square, numbers]{natbib}
+
+% this package is for getting certain math scripts like the 1 needed for an indicator function
+% \usepackage{bbm}
+
+% using this package to make a better looking list
+\usepackage{enumerate}
+% try this one for a list as well. may be better
+% \usepackage[font=itshape]{quoting}
+
+% package for tables
+
+% \renewcommand{\bibfont}{\setstretch{1}\selectfont}
+% \setlength{\bibsep}{13.2pt}
+
+\usepackage{chapterbib}
+%\renewcommand\bibsection{\section*{\ref name}}
+%\renewcommand\bibsection{\section*}
+\renewcommand{\bibsection}{\section{References}}
+
+\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
 
@@
 
-\begin{supplement}
+\renewcommand{\thesection}{\Alph{section}}
 
@@
     \centering
-    \includegraphics[scale=.7]{Images/ili_vs_week.png}
+    \includegraphics[scale=.5]{Images/ili_vs_week.png}
     \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
@@
     \centering
-    \includegraphics[scale=.65]{Images/hosp_vs_week.png}
+    \includegraphics[scale=.55]{Images/hosp_vs_week.png}
     \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
@@
 
-Figures \ref{fig:overall_scores} - \ref{fig:logs_by_season} show boxplots of the CRPS and LogS for the four models.
+Figures \ref{fig:overall_scores} - \ref{fig:logs_by_season} show boxplots of the CRPS and logarithmic score (LogS) for the four models for 500 simulation replicates.
+Figure \ref{fig:logs_by_week_horizon} shows
+boxplots of the LogS faceted by week and horizon.
 Figure \ref{fig:overall_scores} shows that the variation of overall CRPS is smallest for the ASG models and larger for the SIR models.
@@
 For example, SIRD in figure \ref{fig:logs_by_season} tends to show smaller LogS variation relative to the other three models than is seen by the CRPS of SIRD in figure \ref{fig:crps_by_season}.
+
+The LogS is defined in (\ref{eq:logs2}) where $f(\cdot)$ is the forecast density function and $y^*$ is the observed value.
+To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation, and the \texttt{scoringRules} package in \texttt{R} is used for the calculation \cite[]{jordan2019scoring}.
+
+\begin{equation}
+    \text{LogS}(f, y^*) = -\text{log}(f(y^*))
+    \label{eq:logs2}
+\end{equation}
 
@@
     \label{fig:crps_by_season}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[scale=.4]{Images/logs_by_week_horizon.png}
+    \caption{Boxplots of the logarithmic score (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
+Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
+    \label{fig:logs_by_week_horizon}
 \end{figure}
@@
 \end{figure}
-\end{supplement}
 
-\bibliographystyle{ba}
+\bibliographystyle{apa}
 \bibliography{master_bib}

### 2024-12-18 flu_forecast_23/get_data_functions_new.R

@@ -107,7 +107,9 @@ get_ili_data <- function(lag = 2) {
   pops <- read.csv("https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/locations.csv")
   # colnames(pops)[1] <- "region"
   pops <- pops %>% 
-    select(location_name, population, count_rate1, count_rate2, count_rate3,
+    select(location_name, population, count_rate1, 
+           #count_rate2, 
+           count_rate3,
            count_rate4, count_rate5)
   colnames(pops)[1] <- "region"
   
@@ -128,7 +130,9 @@ comb_data <- function(lag = 2, s_region = NA, s_season = 2022) {
   
   ILINet <- get_ili_data(lag = lag) %>% 
     select(region, unweighted_ili, week_start, location_name, season, place,
-           season_week, population, count_rate1, count_rate2, count_rate3,
+           season_week, population, count_rate1, 
+           #count_rate2, 
+           count_rate3,
            count_rate4, count_rate5) %>%
     select(-place)
   

### 2024-12-18 make_images/make_analysis_images.R

@@ -1,5 +1,5 @@
 setwd("~/flu_research/nonlinear_flu_forecast/")
-source("../../flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions.R")
+source("./flu_forecast_23/get_data_functions_new.R")
 # setwd("./model-output/")
 # source("../target-data/get_target_data.R")
 library(stringr)
@@ -20,7 +20,7 @@ setwd("~/flu_research/Prelim Content/models")
 
 target_data <- read.csv("../../../forecast-hub/FluSight-forecast-hub/target-data/target-hospital-admissions.csv") %>%
   mutate(true_value = value) %>%
-  select(-value, -X)
+  select(-value)
 # setwd("~/forecast-hub/FluSight-forecast-hub/model-output/")
 folders <- list.files("./")
 models <- folders[-which(folders == "FluSight-ensemble")]
@@ -194,6 +194,45 @@ team_scores %>%
         legend.text = element_text(size = 12), 
         legend.title = element_text(size = 14),
         legend.position = "none")
+
+
+team_scores %>% 
+  # filter(horizon == 0) %>%
+  # filter(dist == "LST") %>%
+  mutate(distq = ifelse(sq == "SQ", paste0(dist,2),
+                        dist),
+         disc = ifelse(disc == "D", "Disc","No Disc")) %>% 
+  filter(location_name == "US") %>%
+  filter(!(model %in% c("FluSight-ensemble", "FluSight-baseline"))) %>% 
+  mutate(forecast_date = date(reference_date) + 7*as.numeric(horizon)) %>%
+  ungroup() %>% 
+  left_join(both_flu %>% 
+              filter(season == 2023) %>% 
+              select(season_week, date) %>% 
+              mutate(date = date(date)) %>% 
+              unique(),
+            by = join_by(forecast_date == date), multiple = "all") %>% 
+  group_by(season_week, disc, traj, distq) %>% 
+  summarise(mrwis = mean(lwis)) %>% 
+  rowwise() %>% 
+  mutate(minwis = mean(mrwis)) %>% 
+  ggplot() +
+  geom_vline(xintercept = 22, size = 8, colour = "lightgrey") +
+  geom_line(aes(x = season_week, y = minwis, colour = disc),
+            size = 1) +
+  scale_color_manual(values = c("darkgrey", "black")) +
+  facet_grid(traj~distq) +
+  ylab("LWIS") +
+  xlab("Week") +
+  theme_bw() +
[…]

### 2024-12-18 make_images/make_images_for_data_section.R

@@ -31,11 +31,12 @@ ILINet %>% filter(region == state, season >= 2010) %>%
             show.legend = FALSE, size = 1) +
   facet_wrap(~season) +
   ylab("ILI %") +
-  xlab("Week") +
+  xlab("") +
   # geom_vline(aes(xintercept = 22)) +
   theme_bw() + 
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=15))
+  theme(axis.text=element_text(size=15),
+        axis.title=element_text(size=18),
+        strip.text=element_text(size = 15))
 
 both_flu %>% filter(region == "US", season >= 2022) %>% 
   ggplot() +
@@ -45,8 +46,9 @@ both_flu %>% filter(region == "US", season >= 2022) %>%
   ylab("Hospitalizations") +
   xlab("Week") +
   theme_bw() + 
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=15))
+  theme(axis.text=element_text(size=16),
+        axis.title=element_text(size=19),
+        strip.text=element_text(size = 17))
 
 ILINet %>%
   filter(region %in% c("Alabama", "District of Columbia", "Texas",
@@ -64,7 +66,8 @@ ILINet %>%
   xlab("Week") +
   theme_bw() +
   theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=15))
+        axis.title=element_text(size=17),
+        strip.text=element_text(size=13))
 
 both_flu %>%
   filter(region %in% c("District of Columbia", "Alabama", "Texas",
@@ -81,7 +84,8 @@ both_flu %>%
   xlab("Week") +
   theme_bw() +
   theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=15))
+        axis.title=element_text(size=17),
+        strip.text=element_text(size=13))
 
 
 both_flu %>%

### 2025-04-22 response.tex
create response document

\documentclass{article}

\usepackage{enumerate}
\usepackage{hyperref}

\begin{document}

\section*{Review BA2412-010}
\subsection*{Overall assessment:}
This paper by Wadsworth and Niemi adapts two models originally designed to forecast outpatient
ILI data in order to apply them to influenza hospitalization data. This is motivated by a switch in the
target definition of the US CDC FluSight challenge. The paper is well-written and pleasant to read.
The overall idea of leveraging models from previous FluSight cycles for the new target is appealing.
However, several points require clarification. Given the focus of the journal, the Bayesian inference
aspect should be developed more thoroughly as it currently seems a little ad hoc. Moreover, it
currently does not become clear to which degree the proposed methods offer improved forecasts
compared to existing (and potentially much simpler) models. I provide some specific comments in
the following.

\subsection*{Specific comments:}

\begin{enumerate}[1.]
\item The entire Bayesian inference scheme seems somewhat heuristic.

\begin{enumerate}[a.]
\item The ILI model is run separately and then feeds into the hospitalization model. So there is no
flow of information from the hospitalization data to the ILI model, which there would be in a fully
Bayesian procedure. Is this done on purpose and if so what is the motivation? The proposed
procedure reminds me of the "Bayesian cut" in the version described by Plummer
\href{(https://link.springer.com/article/10.1007/s11222-014-9503-z)}{(https://link.springer.com/article/10.1007/s11222-014-9503-z)}.
This may be a useful reference.
\item The authors employ a "modular Bayesian approach" mixing maximum likelihood and
Bayesian estimation. The authors should provide more ample references to the literature concerning
this type of procedure (is "modular Bayesian" a standard term?).
\item Also, it should be argued explicitly why this approach was chosen over a standard fully
Bayesian approach in the setting at hand. Was a fully Bayesian approach also explored? Which
difficulties were encountered?
\item Notably it would be helpful to comment on how this approach affects the uncertainty
propagation / overall quantification of forecast uncertainty.

\end{enumerate}

\item Several comments on the data and general modelling choices:

\begin{enumerate}[a.]
\item Influenza-like illness (ILI) is not a single pathogen, but a mix of several pathogens causing
similar symptoms. How reasonable is it to assume SIR dynamics for such an indicator? Do we
really assume that the tipping point due to susceptible depletion can be captured reasonably here? In
my understanding this would be the only reason to include an S compartment.

\item My impression is that the SIR does not really have any mechanistic meaning here and is just a
way of parameterizing curves that look qualitatively like ILI seasons. The ASG approach can be
seen as an alternative parameterization for such curves. Could the authors comment on what
practical differences exist between the two parameterizations?

\item Is the hospitalization indicator specific to influenza? My understanding is that it is, but this
should be more clear in Section 2.2. If it is influenza-specific: to which degree should we expect ILI
to be a helpful predictor given that it also contains COVID, RSV etc.? The influenza dynamics are
likely somewhat hidden in the ILI indicator.

\item Does the hospitalization indicator refer to incident (new) hospitalization or prevalence, i.e.,
bed occupancy?

\item Is there any "mechanistic" rationale behind the ARX formulation for the hospitalizations? If we
predict prevalence / bed occupancy I can sort of see why the differences between weeks would be
explained by ILI. If we predict new hospitalizations I am not sure what motivates this ARX
formulation.

\item Currently the ILI value from the same week is used in equation (7). Would there be a point in
exploring different lags?

\item On page 14 it says: "all negative values of [...] were set to 0 to reflect realistic values of
hospitalizations". Could this be handled more naturally by appropriate distributional assumptions?

\end{enumerate}

\item And some remarks on the evaluation.

\begin{enumerate}[a.]

\item It would be interesting to have comparisons of performance beyond the variations of the two
proposed models. There should definitely be some naive baseline model (e.g., a persistence
forecast) and some out-of-the-box model applied directly to the hospitalizations (e.g. seasonal
ARIMA). Given the forecasting task was taken from the CDC challenge it would also be interesting
to see how the proposed models fare against the other participants of the challenge.

\item The measure the authors call LWIS tends to give a lot of weight to forecast tasks with low
(expected) incidence as the relative differences between predictions and observations can be quite
large then. The "regular" WIS gives most weight to weeks with high incidence. So in a sense both
are complementary and it may be helpful to also discuss regular WIS a little more.

\item Why was the CRPS used in the simulation and WIS in the data example? These two are sort of
the same (WIS approximates CRPS), so things would be more coherent if one was used throughout.

\item Some assessment of calibration, e.g. interval coverage, should be added.

\end{enumerate}

\item The authors should at least discuss the possibility of sharing certain parameters across geographic
units as done e.g. in
\href{https://www.nature.com/articles/s41467-021-23234-5}{https://www.nature.com/articles/s41467-021-23234-5}
by Osthus and Moran.

\end{enumerate}

\section*{Review BA2412-010R1AE1}

The present paper proposes a novel two component framework for projecting Health 
and Human Services (HHS) hospitalization, based on hospitalization and 
influenza-like illness (ILI) data. In a first, step ILI data are modelled which 
are then used as a predictive covariate in the modelling and projection of the 
hospitalisation data. Probabilistic forecasts are assessed in a simulation study 
using proper scoring rules.

\begin{enumerate}[1.]

\item Since this paper is submitted to Bayesian analysis, the authors should 
underline more the modelling aspects and why this is an important contribution 
for the Bayesian community. What is the impact and importance of priors. Does 
the model borrow information between different model parts, propagate 
uncertainty ... In the introduction none of this is mentioned so the question 
arises whether Bayesian Analysis is the right journal.

\item As I understand one challenge of forecasting hospitalisation data is that the 
start of the collection of these data is very recent, so that a model only 
based on these data has not much historic data to be based on. Thus, the idea 
of using ILI data is natural. The authors could emphasize more and discuss 
potential alternative models that could even be included in the simulation 
study. I am missing a comparison to models that are not the newly proposed 
ones within the paper.

\item Why is it sensible to include ILI as a linear function when modelling 
hospitalisation. Please motivate more.

\item The authors provide also state level predictions. However, I do not see 
where a spatial component is included in the model. Is this only via P or are 
all states just modelled independently. For influenza I would think that it 
might be reasonable to assume that neighbouring states show similar behaviour 
in both ILI and hosptialisation so that borrowing strength between states 
within one model seems sensible. Please discuss.

\end{enumerate}

\subsection*{Specific comments:}

\begin{enumerate}[-]

\item Section 2, page 3: please be more clear on what you mean by the 
"first seven season". Provide a time frame.

\item Figure 1: I was wondering whether the R-package geofacet could be useful to 
show also a spatial trend across US states, in addition to temporal trend, for 
certain years,  similar to the second plot provided here:

https://hafen.github.io/geofacet/

I could image that the curves, onset and general behaviour, varies across the 
states. The authors refer to this a bit in the discussion. In addition I would 
prefer to also have months names in the axis labels, to avoid confusion that 
week 1 is not the first week in January.

Can the hospitalization figures also be shown on state level at least in an appendix.

\item Section 3.5: Please be clear what the strength of a Bayesian inference scheme is.

\item Equation 6): What does the tilde mean?

\item Page 11: Please motivate the choice of an autoregressive process of order 1 
more clearly. Why not use a higher order? Would this be beneficial?

\item Page 14: Please discuss the assumption that all components of $\alpha_s$ and 
$\sigma_\epsilon^2$ are assumed to be same across all seasons. Is this realistic?

\item Figure 7: Please make clear in the caption what the columns represent in an 
intuitive manner.

\item Equation 11, 12, 13, 14: Please motivate. The authors write in Section 3.4 
that the modelling of discrepancy might require careful selection of priors. 
Please underline more how this is taken care of in this manuscript.

\item Please check for typos.

\end{enumerate}
\end{document}

### 2025-04-22 response.tex
add some comments for response to reviews

@@
 \usepackage{hyperref}
+
+\usepackage{xcolor}
+\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
+\newcommand{\spencer}[1]{{\color{green} Spencer: #1}}
 
@@
 difficulties were encountered?
+
+\spencer{Without assuming a value for the $\lambda$ parameter in the ASG model, there were
+serious problems with parameter identifiability. 
+Osthus et al. in their SIR model doe something similar by setting $S_0 = 0.9$}
+
 \item Notably it would be helpful to comment on how this approach affects the uncertainty
@@
 
+\spencer{I agree with these comments on the SIR model. Perhaps we emphasized the SIR model more than we 
+should have. Osthus et al. use the SIR model in combination with a data driven model
+in an attempt to capture some aspect of the science in their forecast model.
+More important in the argument of this paper than that the specific SIR model is
+a superior in some way for modeling ILI is that the ASG model may
+be a reasonable alternative to existing compartmental models at least in terms
+of forecast performance.}
+
 \item My impression is that the SIR does not really have any mechanistic meaning here and is just a
@@
 
+\spencer{Practially, the reason to use the SIR model is generally because it is meant
+to capture some of the scientific understanding of a disease outbreak. Again, we argue 
+that the ASG is a reasonable alternative to SIR models in the case of the seasonal flu
+outbreak. The ASG shape with a rise to a peak and a subsequent fall is similar in shape
+to the regular ILI trajectory. It has also been used successfully in forecasting of 
+other pheneomena with similar rises and falls (i.e. crop moisture)}.
+
 \item Is the hospitalization indicator specific to influenza? My understanding is that it is, but this
@@
 likely somewhat hidden in the ILI indicator.
+\spencer{Yes, hospitalizations here are specific to lab confirmed influenza patients. 
+I believe figure 6 in the text gives reason to believe ILI is a good predictor. Maybe 
+we should add some additional EDA with an $R^2$ value to further defend this?}
 
@@
 bed occupancy?
+\spencer{The weekly hospitalization indicator refers to new incident hospitalizations}
 
@@
 exploring different lags?
+\spencer{Exploring different lags is an important idea here, especially considering there
+may be a reporting lag. However, the reporting lag is not as big an issue as it has
+been in the past. We assessed ILI lags for both one and two weeks previous to the current
+week and found that ILI had less predictive power. }
 
@@
 the same (WIS approximates CRPS), so things would be more coherent if one was used throughout.
+\spencer{Probably use CRPS throughout and not WIS at all.}
 
 \item Some assessment of calibration, e.g. interval coverage, should be added.
+\spencer{We can add this. I don't think it will look good, but maybe we recommend
+further research involving calibration of posterior credible regions (Syring \& Martin 2019)
+or generalized predictive distributions (Wu \& Martin 2021).}
 
@@
 
+\spencer{The ILI models are hierarchical and intended to be made so that information
+accross all previous seasons may be borrowed for modeling the current season.}
+
 \item As I understand one challenge of forecasting hospitalisation data is that the 
@@
 ones within the paper.
+\spencer{Should include comparison to a naive baseline (probably a random walk) 
+and an out of the box ARMA or ARIMA, and maybe exponential smoothing.}
 
@@
 hospitalisation. Please motivate more.
+
+\spencer{I think this is supported in figure 6 in the paper, though as mentioned 
+in response to previous reviewer, maybe additional EDA would be helpful.}
 
@@
 \item Equation 6): What does the tilde mean?
+\spencer{The tilde indicates forecast ILI. Maybe it just shouldn't be there?}
 
@@
 more clearly. Why not use a higher order? Would this be beneficial?
+\spencer{We explored higher orders and found that they didn't provide any more
+predictive power than order 1 did.}
 

### 2025-04-23 flu_forecast_23/full_model_analysis.R
new analysis

@@ -1,0 +1,131 @@
+setwd("./flu_forecast_23/")
+library(tidyr)
+library(readr)
+source("./get_data_functions_new.R")
+source("./mle_functions.R")
+library(evalcast)
+
+
+both_flu <- comb_data(lag = 1) %>%
+  filter(season >= 2010) #%>% select(-X)
+both_flu2 <- comb_data(lag = 2) %>%
+  filter(season >= 2010) #%>% select(-X)
+
+ILINet <- get_ili_data() %>%
+  filter(season >= 2010)
+
+
+mod <- cmdstan_model(stan_file = "../stan_models/asg_hosp_log_sq_ar1.stan")
+mod2 <- cmdstan_model(stan_file = "../stan_models/hosp_log_ar1.stan")
+
+#dir.create(paste0("./hosp_fits/", model))
+select_regions <- unique(ILINet$region)
+#select_regions <- c("Alabama", "Alaska")#, "Idaho")
+#select_regions <- c("US")
+hosp_log <- ifelse(str_detect(model, "log"), TRUE, FALSE)
+hosp_lst <- ifelse(str_detect(model, "lst"), TRUE, FALSE)
+sqr <- ifelse(str_detect(model, "sq"), TRUE, FALSE)
+if (hosp_log == TRUE) {both_flu$value <- log(both_flu$value + 1)}
+if (hosp_log == TRUE) {both_flu2$value <- log(both_flu2$value + 1)}
+
+
+# all_forecasts <- foreach(j = select_regions,
+#                          .packages = c("tidyr", "dplyr", "evalcast")
+#                          ,.errorhandling = "remove"
+#                          ,.combine = rbind) %:% 
+#   #k <- 14
+#   foreach(k = 
+#             #c(14, 20, 26), .combine = rbind) %dopar% {
+#             9:38
+#           , .combine = rbind) %dopar% {
+            # j <- "Georgia"; k <- 21 
+j <- "US"
+all_qforcs_res <- data.frame()
+for (k in 13:38) {
+            both_flu_hold <- both_flu %>%
+              filter(region == j, season == 2022 | (season == 2023 & season_week <= k)) 
+            
+            both_flu2_hold <- both_flu2 %>%
+              filter(region == j, season == 2022 | (season == 2023 & season_week <= k))
+            
+            stan_dat <- list(HM = nrow(both_flu_hold),
+                             n_seasons_hosp = length(unique(both_flu_hold$season)),
+                             ili_ps = both_flu_hold$unweighted_ili/100,
[…]

### 2025-04-23 make_images/make_images_for_data_section.R

@@ -1,7 +1,7 @@
 
-setwd("~/flu_research/nonlinear_flu_forecast/")
-source('~/flu_research/nonlinear_flu_forecast/functions_data/get_data_functions_new.R')
-source('~/flu_research/nonlinear_flu_forecast/functions_data/mle_functions.R')
+setwd("./nonlinear_flu_forecast/")
+source('./functions_data/get_data_functions_new.R')
+source('./nonlinear_flu_forecast/functions_data/mle_functions.R')
 
 
 
@@ -165,7 +165,7 @@ both_flu <- both_flu %>%
   mutate(uwili2 = count_rate2*unweighted_ili^2, 
          cruili = unweighted_ili*count_rate2)
 
-sample_flu <- both_flu %>% filter(region == "Iowa", season == 2022:2023)
+sample_flu <- both_flu %>% filter(region == "Iowa", season %in% 2022:2023)
 y <- sample_flu$value; x = sample_flu$unweighted_ili*sample_flu$count_rate1
 x2 <- x^2
 mod <- lm(y ~ x)

### 2025-04-28 response.tex
add some comments
Base analysis material

@@
 \usepackage{hyperref}
+\usepackage{amsmath}
+% \usepackage{amsfonts}
 
@@
 \newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
-\newcommand{\spencer}[1]{{\color{green} Spencer: #1}}
+\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}
 
@@
 This may be a useful reference.
+
+\spencer{In the forecasting competition, the data comes generally by the afternoon and
+forecasts are due by the evening. We initially did a full Bayesian procedure
+(minus the MLE selection of the $\lambda$ parameter in the ASG ILI model), but fit 
+was very slow and made it difficutl to get forecasts submitted in time. We switched
+to using two separate models because of the time saved in model fitting. We also
+found that in terms of the proper scoring rules and predictive interval coverage,
+there wasn't a drop in performance although the parameter estimation did 
+change.
+
+\begin{equation}
+\begin{aligned}
+    p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}) &= \int 
+       p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) 
+       p(\boldsymbol{\alpha} | \textbf{ILI}, \textbf{H}) d\boldsymbol{\alpha} \\
+    p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) &= 
+       \int p(\tilde{H}_{s,w^*} | \textbf{H}, \textbf{ILI}, 
+       \widetilde{ILI}_{s,w^*}, \boldsymbol{\alpha}) 
+       p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) d\widetilde{ILI}_{s,w^*}
+\end{aligned}
+\end{equation}
+}
+
 \item The authors employ a "modular Bayesian approach" mixing maximum likelihood and
@@
 
+\spencer{It was precisely the inclusion of the discrepancy that caused the serious problems
+with identifiability which motivated the modular Bayesian approach. }
+
 \item Please check for typos.

### 2025-04-28 flu_forecast_23/base_forecasts.R

@@ -1,0 +1,137 @@
+setwd("./flu_forecast_23/")
+library(tidyr)
+library(readr)
+source("./get_data_functions_new.R")
+source("./mle_functions.R")
+library(evalcast)
+library(forecast)
+
+
+both_flu <- comb_data(lag = 1) %>%
+  filter(season >= 2010) #%>% select(-X)
+both_flu2 <- comb_data(lag = 2) %>%
+  filter(season >= 2010) #%>% select(-X)
+
+ILINet <- get_ili_data() %>%
+  filter(season >= 2010)
+
+
+#dir.create(paste0("./hosp_fits/", model))
+select_regions <- unique(ILINet$region)
+#select_regions <- c("Alabama", "Alaska")#, "Idaho")
+#select_regions <- c("US")
+hosp_log <- ifelse(str_detect(model, "log"), TRUE, FALSE)
+hosp_lst <- ifelse(str_detect(model, "lst"), TRUE, FALSE)
+sqr <- ifelse(str_detect(model, "sq"), TRUE, FALSE)
+if (hosp_log == TRUE) {both_flu$value <- log(both_flu$value + 1)}
+if (hosp_log == TRUE) {both_flu2$value <- log(both_flu2$value + 1)}
+
+
+# all_forecasts <- foreach(j = select_regions,
+#                          .packages = c("tidyr", "dplyr", "evalcast")
+#                          ,.errorhandling = "remove"
+#                          ,.combine = rbind) %:% 
+#   #k <- 14
+#   foreach(k = 
+#             #c(14, 20, 26), .combine = rbind) %dopar% {
+#             9:38
+#           , .combine = rbind) %dopar% {
+# j <- "Georgia"; k <- 21 
+N <- 5000
+all_qforcs_res <- data.frame()
+for (j in select_regions) { #[47:length(select_regions)]) {
+
+  for (k in 9:38) {
+    both_flu_hold <- both_flu %>%
+      filter(region == j, season == 2022 | (season == 2023 & season_week <= k)) 
+    
+    forecast_date <- date(max(both_flu_hold$date))
+  
+    
+    probs <- c(.01, .025, seq(.05, .95, by = .05), .975, .99)
+    
+    both_flu_reg <- both_flu %>% 
+      filter(region == j) %>% 
+      mutate(true_value = value) %>% 
+      ungroup() %>% 
+      select(date, true_value) %>% 
+      mutate(date = date(date))
+    
+    basemod <- naive(both_flu_hold$value)
[…]

### 2025-04-28 flu_forecast_23/full_model_analysis.R

@@ -41,7 +41,7 @@ if (hosp_log == TRUE) {both_flu2$value <- log(both_flu2$value + 1)}
             # j <- "Georgia"; k <- 21 
 j <- "US"
 all_qforcs_res <- data.frame()
-for (k in 13:38) {
+for (k in 25:38) {
             both_flu_hold <- both_flu %>%
               filter(region == j, season == 2022 | (season == 2023 & season_week <= k)) 
             
@@ -126,6 +126,7 @@ all_qforcs_res %>%
   group_by(date) %>%
   summarise(mwis = mean(wis)) %>%
   ggplot() +
-  geom_line(aes(x = date, y = mwis))
+  geom_line(aes(x = date, y = mwis)) +
+  geom_hline(yintercept = .9)
 
-
+write.csv(all_qforcs_res, "res.csv", row.names = FALSE)

### 2025-04-28 flu_forecast_23/get_data_functions_new.R

@@ -108,7 +108,7 @@ get_ili_data <- function(lag = 2) {
   # colnames(pops)[1] <- "region"
   pops <- pops %>% 
     select(location_name, population, count_rate1, 
-           #count_rate2, 
+           count_rate2,
            count_rate3,
            count_rate4, count_rate5)
   colnames(pops)[1] <- "region"

### 2025-04-28 make_images/make_analysis_images.R

@@ -1,4 +1,4 @@
-setwd("~/flu_research/nonlinear_flu_forecast/")
+setwd("../")
 source("./flu_forecast_23/get_data_functions_new.R")
 # setwd("./model-output/")
 # source("../target-data/get_target_data.R")

### 2025-04-29 flu_forecast_23/get_data_functions_new.R
fix ili data

@@ -107,6 +107,7 @@ get_ili_data <- function(lag = 2) {
   pops <- read.csv("https://raw.githubusercontent.com/cdcepi/FluSight-forecast-hub/main/auxiliary-data/locations.csv")
   # colnames(pops)[1] <- "region"
   pops <- pops %>% 
+    mutate(count_rate2 = count_rate1*2) %>% 
     select(location_name, population, count_rate1, 
            count_rate2,
            count_rate3,
@@ -131,7 +132,7 @@ comb_data <- function(lag = 2, s_region = NA, s_season = 2022) {
   ILINet <- get_ili_data(lag = lag) %>% 
     select(region, unweighted_ili, week_start, location_name, season, place,
            season_week, population, count_rate1, 
-           #count_rate2, 
+           count_rate2,
            count_rate3,
            count_rate4, count_rate5) %>%
     select(-place)

### 2025-06-30 flu_forecast_23/full_fit_hosp.R
add full fit

@@ -39,14 +39,14 @@ sqr <- ifelse(str_detect(model, "sq"), TRUE, FALSE)
 if (hosp_log == TRUE) {both_flu$value <- log(both_flu$value + 1)}
 if (hosp_log == TRUE) {both_flu2$value <- log(both_flu2$value + 1)}
 
-
+#select_regions <- "US"
 all_forecasts <- foreach(j = select_regions,
 	.packages = c("tidyr", "dplyr", "evalcast")
 	,.errorhandling = "remove"
 	,.combine = rbind) %:% 
 	#k <- 14
   foreach(k = 
-	  #c(14, 20, 26), .combine = rbind) %dopar% {
+	  #c(14, 20, 26)
 	  9:38
   	, .combine = rbind) %dopar% {
 # j <- "Georgia"; k <- 21 
@@ -70,7 +70,7 @@ all_forecasts <- foreach(j = select_regions,
 		     sigma_nu = 15
 		 )
     location <- unique(both_flu_hold$location)
-    #if (j != "US" & !(k %in% c(14, 20, 26))) {
+    if (j != "US" & !(k %in% c(14, 20, 26))) {
             samps <- mod$sample(data = stan_dat,
 				chains = 1,
 				adapt_delta = .9999,
@@ -78,27 +78,27 @@ all_forecasts <- foreach(j = select_regions,
 				iter_sampling = 50000)
 
     	    draws <- samps$draws(format = "df")
-    #}
-    #if (j == "US" & k %in% c(14, 20, 26)) {
-    #        samps <- mod$sample(data = stan_dat,
-    #    			chains = 4,
-    #    			adapt_delta = .9999,
-    #    			iter_warmup = 10000,
-    #    			iter_sampling = 50000)
-    #        draws <- samps$draws(format = "df")
+    }
+    if (j == "US" & k %in% c(14, 20, 26)) {
+            samps <- mod$sample(data = stan_dat,
+        			chains = 4,
+        			adapt_delta = .9999,
+        			iter_warmup = 10000,
+        			iter_sampling = 50000)
+            draws <- samps$draws(format = "df")
 
-    #        summary <- samps$summary()
-    #        diag_list <- lapply(samps$diagnostic_summary(), FUN = mean)
-    #        diag <- diag_list %>% data.frame()
-    #        diag$max_rhat <- max(summary$rhat, na.rm = TRUE)
-    #        diag$min_ess <- min(summary$ess_bulk, na.rm = TRUE)
[…]

### 2025-06-30 abstract.tex
fix full fit

--- 2024-12-13 nonlinear_forecast.tex
@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
-
-%% Packages
-\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
-\RequirePackage[numbers]{natbib}
-%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
-\RequirePackage{graphicx}
-
-%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
-\usepackage{booktabs} %for the table in the analysis section
-\usepackage{enumerate} %for the short algorithm
-\usepackage{subfigure} %for making plots with multiple images
-\usepackage{tabularx} %for the spacing in tabular
-\usepackage{newtxmath} %for making indicator function
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\pubyear{2024}
-\arxiv{2010.00000}
-\volume{TBA}
-\issue{TBA}
-\firstpage{1}
-\lastpage{1}
-
-\startlocaldefs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Uncomment next line to change            %%
-%% the type of equation numbering           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\numberwithin{equation}{section}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Axiom, Claim, Corollary, Hypothesis, %%
-%% Lemma, Theorem, Proposition              %%
-%% use \theoremstyle{plain}                 %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{plain}
-\newtheorem{axiom}{Axiom}
-\newtheorem{claim}[axiom]{Claim}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Assumption, Definition, Example,     %%
-%% Notation, Property, Remark, Fact         %%
-%% use \theoremstyle{definition}            %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem*{example}{Example}
-\newtheorem*{fact}{Fact}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Case use \theoremstyle{remark}       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{remark}
-\newtheorem{case}{Case}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Please put your definitions here:        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\endlocaldefs
-
-\begin{document}
-
-\begin{frontmatter}
-\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy}
-%\title{A sample article title with some additional note\thanksref{t1}}
-\runtitle{Bayesian nonlinear flu hospitalizations forecasting}
-%\thankstext{T1}{A sample additional note to the title.}
-
-\begin{aug}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Only one address is permitted per author. %%
-%% Only division, organization and e-mail is %%
-%% included in the address.                  %%
-%% Additional information can be included in %%
-%% the Acknowledgments section if necessary. %%
-%% ORCID can be inserted by command:         %%
-%% \orcid{0000-0000-0000-0000}               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author[A]{\snm{Spencer Wadsworth}~\ead[label=e1]{sgw96@iastate.edu}}
-% \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}\orcid{0000-0000-0000-0000}}
-\and
-\author[A]{~\snm{Jarad Niemi}\ead[label=e2]{niemi@iastate.edu}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Addresses                                %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\address[A]{Department of Statistics,
-Iowa State University, Ames, IA 50011}
-
-\runauthor{Wadsworth and Niemi}
-\end{aug}
-
 \begin{abstract}
-The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread.
-The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods.
-For the first several years of the competition, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalization counts.
-Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons.
-In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data.
-The first component is for modeling ILI data using a dynamic nonlinear Bayesian hierarchical model.
-The second component is for modeling hospitalizations as a function of ILI.
-For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor.
-In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from \cite{osthus2019dynamic} and a nonlinear Bayesian hierarchical model from \cite{ulloa2019} are compared.
-Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model.
-Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared.
-We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year.
-We also found that other modeling decisions such as the exact nonlinear function to use in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
+The annual influenza outbreak leads to significant public health and economic 
+burdens making it desirable to have prompt and accurate probabilistic forecasts 
+of the disease spread. The United States Centers for Disease Control and 
+Prevention (CDC) hosts annually a national flu forecasting competition which has 
+led to the development of a variety of flu forecast modeling methods. For the 
+first several years of the competition, the target to be forecast was weekly 
+percentage of patients with an influenza-like illness (ILI), but in 2021 the 
+target was changed to weekly hospitalization counts. Reliable state and national 
+hospitalization data has only been available since 2021, but for ILI the data 
+has been available since 2010 and has been successfully forecast for several 
+seasons.
+In this manuscript, we introduce a two component modeling framework for 
+forecasting weekly hospitalizations utilizing both hospitalization data and ILI 
+data. The first component is for modeling ILI data using a dynamic nonlinear 
+Bayesian hierarchical model. The second component is for modeling 
+hospitalizations as a function of ILI. For hospitalization forecasts, ILI is 
+first forecasted and then hospitalizations are forecast with ILI forecasts used 
+as a linear or quadratic predictor. In a simulation study, two ILI forecast 
+models, including one similar to the winning model for two seasons of the CDC 
+forecast competition from Osthus et al.
+% \cite[]{osthus2019dynamic} 
+% 
+and a nonlinear Bayesian hierarchical model from Ulloa 
+% \cite[]{ulloa2019}
+are compared. Also assessed is the usefulness of including a systematic model 
+discrepancy term in the ILI model. Forecasts of state and national 
+hospitalizations for the 2023-24 flu season are made, and different modeling 
+decisions are compared. We found that including a discrepancy component in the 
+ILI model tends to improve forecasts during certain weeks of the year. We also 
+found that other modeling decisions such as the exact nonlinear function to be 
+used in the ILI model or the error distribution for hospitalization models may 
+or may not be better than other decisions, depending on the season, location, 
+or week of the forecast.
 \end{abstract}
-
-% \begin{keyword}[class=MSC]
-% \kwd[Primary ]{00X00}
-% \kwd{00X00}
-% \kwd[; secondary ]{00X00}
-% \end{keyword}
-
-\begin{keyword}
-\kwd{Disease outbreak forecasting}
-\kwd{Bayesian hierarchical modeling}
-\kwd{Probabilistic forecasting}
-\kwd{Model discrepancy}
-\end{keyword}
-
-\end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% `\ ' is used here because TeX ignores    %%
-%% spaces after text commands.              %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-
-Every year the seasonal influenza outbreak burdens the public health system by infecting millions, causing an influx of primary care visits and hospitalizations and leading to between 290,000 and 650,000 deaths worldwide \cite[]{whoflufact2023}.
-\cite{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion.
-Accurate forecasting of infectious diseases can inform public decision making and ease the burden of an outbreak \cite[]{turtle2021accurate, lutz2019applying}.
-There is a growing consensus that disease forecasts should be probabilistic in nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has been shown that reporting forecast uncertainty along with predictions may lead to better decision making \cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.
-
-% \jarad{Test}
-% \spencer{Test}
-
-To better inform public decision making regarding the flu epidemic, in 2013 the United States Centers for Disease Control and Prevention (CDC) organized a national flu forecasting competition, also known as FluSight \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}.
-Originally, over a dozen teams of researchers from academic and industry backgrounds participated in FluSight by contributing their own forecast models.
-Besides the 2020 season --or the flu season spanning the fall of 2020 and the winter of 2021-- FluSight has been operated annually and researchers outside the CDC have been invited to participate.
-Initially the target data for forecasts was influenza-like illness (ILI) data.
-ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}.
-The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, \cite{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework.
-An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
-
-The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications.
-As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}.
-The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded.
-The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}.
-Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition.
-Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}.
-This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.
-
-The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models.
-The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
-Herein we use ILI models similar to those in \cite{osthus2019dynamic} and \cite{ulloa2019} for ILI forecasting.
-The model of \cite{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework.
-Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by \cite{kennedy2001bayesian}.
-The model in \cite{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data.
-Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.
-
-In line with the newer FluSight standard of forecasting hospitalizations, we model hospitalizations as a linear function of ILI.
-Thus forecasts produced herein target flu hospitalizations and are a mapping of ILI forecasts to hospitalizations.
-This allows for ILI data from many seasons to be exploited and for ILI forecasts to assist in forecasting hospitalizations, which has fewer seasons of data than ILI.
-Several modeling schemes and their forecasts for the 2023 flu hospitalization season are compared, and it is shown that the modeling decisions produce good forecast results for different states or times during the flu season.
-
-In section \ref{sec:data} we review the ILI and hospitalization data provided by the CDC and targeted by FluSight. 
-In section \ref{sec:functions} the modeling framework contributed by this manuscript is given.
-In the same section, functions similar to those used by \cite{osthus2019dynamic} and \cite{ulloa2019} are defined.
- These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively.
-Model fitting and implementation are described at the end of the section.
-Section \ref{sec:simulation2} is a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared.
-Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts.
-Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}, with the analysis being done using the conventions of FluSight.
- Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.
-
- \section{Flu outbreak data} \label{sec:data}
-In this section we introduce and define ILI and hospitalization data and evaluate the data visually.
-ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season.
-Both of these data were collected at the state, territorial, and national level and were reported at least weekly.
-Overall the data is reported for 53 locations including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), and at the US national level.
-We will refer to forecast targets throughout this manuscript.
-A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a specific location and week during the season.
-
-\subsection{Influenza-like illness data}
-The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers.
-Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases.
-An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}.
-Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza.
-To know whether or not a sick patient is infected with influenza would require a laboratory test.
-
-In 2013, when FluSight began, the ILI data was the object of the forecasts.
-The data was released publicly at HHS region levels, and forecast teams were asked to provide forecasts of several ILI targets on the regional levels including season onset, 1-4 week ahead ILI levels, and the week of peak ILI activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}.
-Currently, the ILI data is collected by the CDC and published on an online portal for viewing at the national, HHS region, census, and state levels \cite[]{cdc2024fluviewportal}.
-To obtain ILI data, we used the \texttt{R} package \texttt{cdcfluview} which provides functions for downloading the data \cite[]{rudis2021cdcfluview}.
-Weekly ILI data from the national, HHS region, and census levels are available from the 1997 flu season until the current season.
-At the state level, data is available from the 2010 flu season to the current season.
-
-Figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023.
-For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons.
-To better align with the flu behavior, week 1 is set as the first week of August and week 52 or 53 is the last week in July of the following year.
-For example week 1 of the 2013 season corresponds to the first week of August 2013, and week 52 of the same season corresponds to the last week of July 2014.
-This convention is used for the remainder of this manuscript.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/us_ili_seasons.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
-Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.}
-    \label{fig:us_ili}
-\end{figure}
-
-Notable from the plots in figure \ref{fig:us_ili} is the regular trajectory of the ILI.
-With the exception of season 2020, the ILI begins low at week 1 and increases as the fall and winter progress until the ILI reaches a peak.
-As spring progresses to summer, the ILI decreases to low values.
-As \cite{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day.
-Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
-\cite[]{ewing2017contact, garza2013effect}.
-The only seasons when there was not a peak at week 22 were season 2022 where the season peak occurred particularly early and season 2020 which was greatly influenced by the COVID-19 pandemic.
-
-\subsection{Hospitalization Data}
-
-Hospital admission data, used as the object of FluSight forecasting for the 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported Patient Impact and Hospital Capacity by State Timeseries}.
-Several targets of respiratory illnesses including COVID-19, RSV, and influenza are reported weekly by most hospitals in the US.
-In February 2022 it became mandatory for all hospitals to report the number of COVID-19 and influenza hospitalizations, and since then reporting of hospitalizations has become widespread.
-These data were updated every Wednesday and Friday according to NHSN guidelines \cite[]{healthdata2024covidts}.
-
-Figure \ref{fig:us_hosp} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons.
-These plots show similarities to the ILI plots in \ref{fig:us_ili} in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends.
-For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22.
-Additional plots of ILI data and hospitalization
-data for select states are included in the supplementary material 
-\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar to 
-those in figures \ref{fig:us_ili} and \ref{fig:us_hosp} with activity increasing
-to a peak and then decreasing.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
-    \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
-    \label{fig:us_hosp}
-\end{figure}
-
-\section{ILI and hospitalization forecast modeling} \label{sec:functions}
-
-% \jarad{Why not start with the ILI model from Osthus et al?
-It makes a lot more sense to me to start with what has already been done and then expand upon what they have done.
-Later you have a section called ILI Model, but isn't this the ILI model?}
-
-The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows of a similar trajectory for modeling ILI.
-Compartmental models are standard mathematical models used for disease outbreaks.
-One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}.
-\cite{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data.
-The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy.
-In the first part of this section, we present an ILI model similar to the model in \cite{osthus2019dynamic}.
-With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.
-
-With the aim of forecasting hospitalizations, we also introduce a linear model of hospitalizations data with ILI as a predictive covariate.
-To forecast hospitalizations, ILI data is first forecasted and the forecast is then plugged in as a covariate in the hospitalization model, thus producing hospitalization forecasts.
-The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.
-
-\subsection{ILI Model} \label{sec:ili_model}
-
-The proposed model for ILI for any location is given in (\ref{eq:ili_model}).
-Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given season.
-The ILI is a proportion, so the Beta random variable is a natural selection for modeling.
-Under the parameterization in of the Beta distribution used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a scale parameter.
-The nonlinear function $f_{\theta_s}(w)$ captures the trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the systematic patterns which $f_{\theta_s}(w)$ does not capture.
-\begin{equation}
-\begin{aligned}
-    \label{eq:ili_model}
-    % \begin{flalign*}
-        ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
-        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w
-    % \end{flalign*}
-\end{aligned}
-\end{equation}
-In \cite{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}.
-In \cite{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}.
-In \cite{ulloa2019} modeling is done hierarchically over seasons.
-
-\subsection{Susceptible-Infectious-Recovered (SIR) compartmental model} \label{sec:sir_func}
-
-The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by \cite{kermack1927contribution}.
-Since then, compartmental models have become standard for modeling infectious diseases \cite[]{allen2008mathematical}, and many extensions have been made and studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}.
-The SIR mathematical model includes three compartments and assumes that at any time $t>0$ every individual in a closed population belongs to exactly one compartment.
-The three compartments are susceptible ($S$), infectious ($I$), and recovered ($R$), and their interaction over the course of an outbreak is described by the differential equations in (\ref{eq:sir_diff}).
-
-\begin{equation}
-    \label{eq:sir_diff}
-    \frac{dS}{dt} = -\beta SI, \quad \frac{dI}{dt} = \beta S I - \delta I, \quad \frac{dR}{dt} = \delta I
-\end{equation}
-Here $S$, $I$, and $R$ represent the proportion of the population in each compartment such that $S + I + R = 1$ for all $t$.
-The trajectory is determined by the disease transmission rate $\beta > 0$ and the recovery rate $\delta > 0$.
-Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual, and the expected rate of recovery to an immune state for a newly infected person.
-Whether or not a disease outbreak is classified as an epidemic is determined by the initial susceptible population $S_0$, or the susceptible population at time $0$, and the parameter $\rho = \delta/
-\beta$.
-If $S_0/\rho > 1$, the outbreak is considered an epidemic.
-It is non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}.
-Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where the $S_0$ and $\rho$ were selected to match an outbreak that would classify as epidemic.
-In the case where $S_0 \leq \rho$, the trajectory for the $I$ compartment would never be increasing.
-The increase to a peak and subsequent decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is reasonable to model ILI by this compartment.
-Thus in modeling the ILI data, we consider the data to be analogous to the $I$ proportion of the population.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/sir_traj.png}
-    \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
-The three compartments are the susceptible compartment (left), infectious (center), and recovered (right).
-In this example, $S_0/\rho > 1$.}
-    \label{fig:sir_traj}
-\end{figure}
-
-\subsection{Asymmetric Gaussian (ASG) function} \label{sec:asg_func}
-The ASG function is another example of a nonlinear function which can approximate the trajectory of the flu outbreak.
-The ASG was previously used by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used to model vegetation growth and satellite sensor data \cite[]{lewis2020extracting, jonsson2002seasonality, hird2009noise, beck2006improved, atkinson2012inter}.
-The ASG is a modification of the asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized by its rise to a peak and fall from that peak which may not occur at the same rate, as shown in figure \ref{fig:asg_function}.
-The ASG function is denoted as $ASG_\theta(w)$ where $\theta = (\lambda, \nu, \mu, \sigma_1^2, \sigma_2^2)$, $\nu > 0$, $\lambda > 0$, $\mu \in (-\infty, \infty)$, $\sigma_1, \sigma_2 > 0$ and $w \in (1, ..., W)$ is week.
-The function is defined in (\ref{eq:asg_function}).
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{Images/asg_function.png}
-    \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
-    \label{fig:asg_function}
-\end{figure}
-
-\begin{equation}
-    \label{eq:asg_function}
-    ASG_{\theta}(w) = 
-    \begin{cases}
-        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
-        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
-    \end{cases}
-\end{equation}
-For modeling in this manuscript, we use a slightly reparameterized version of the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$.
-This constraint guarantees that the function has a peak greater than $\lambda$.
-
-\begin{equation}
-    \label{eq:asg_function_rep}
-    ASG_{\theta}(w) = 
-    \begin{cases}
-        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
-        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
-    \end{cases}
-\end{equation}
-
-\subsection{Model discrepancy}
-
-The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as \cite{osthus2019dynamic} points out there may be systematic behavior that these or other possible functions may not capture.
-As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a regular peak at week 22 of the flu season.
-Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to illustrate the systematic discrepancy from a fitted function.
-Figure \ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 to 2022 excluding 2020 with a best fit ASG function plotted over the ILI.
-The fits for each season were made by obtaining the maximum likelihood estimate (MLE) of a model given the ILI data where we assume the ASG function is the mean parameter of a Beta distributed random variable.
-Figure \ref{fig:discrepancy} shows the discrepancy between the model fit and the data for the same seasons.
-The grey lines show the difference between the data and the functions from figure \ref{fig:asg_fits} for each season, and the black line is the average by week over all seasons.
-The lines show that the ASG function typically underpredicts week 22 and overpredicts week 23.
-Perhaps for other weeks, like week 30 for example, there also tends to be systematic behavior not captured by the ASG function.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/asg_fits.png}
-    \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
-    \label{fig:asg_fits}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{Images/discrepancy.png}
-    \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
-    \label{fig:discrepancy}
-\end{figure}
-
-The term $\gamma_w$, where $w$ is the season week, is included in (\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the function.
-Modeling discrepancy has been used in uncertainty analysis of simulators to capture systematic differences between mathematical models and
-reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,arendt2012improving,kennedy2001bayesian}.
-Modeling discrepancy can lead to overfitting, particularly in forecasting scenarios, and may also lead to identifiability issues.
-Thus, care must be taken in setting parameter constraints as well as in the selection of prior distributions \cite[]{osthus2019dynamic,brynjarsdottir2014learning}. 
-Modeling of the discrepancy for ILI was done by \cite{osthus2019dynamic} during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}.
-As in Osthus's model, $\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}).
-\begin{equation}
-    \label{eq:rw}
-        \gamma_w|\gamma_{w + 1} \overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
-        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W})
-\end{equation}
-The idea for using the reverse random walk is that there are several previous seasons of ILI data, and assuming the random walk captures systematic behavior, fitting it hierarchically over seasons can assist in predicting future behavior in the current season.
-Reverse random walks have also been used with success in election forecasting and other flu forecasting models \cite[]{osthus2021multiscale, osthus2019dynamic, linzer2013dynamic}.
-The sum to zero constraint $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve identifiability.
-
-\subsection{ILI forecasts}
-Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating.
-Future ILI forecasts are obtained via the posterior predictive distribution where for week $w$, the predictive distribution is obtained by integrating over the parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ as in (\ref{eq:ili_post}) where $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ is the density function of the posterior distribution for the model parameters.
-If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ where $i$ is a positive integer.
-
-\begin{equation}
-    \label{eq:ili_post}
-    p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) = \int \int \int \int p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
-\end{equation}
-% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
-% \spencer{Isn't that captured in $d \pi_s$?}
-
-\subsection{Hospitalization model} \label{sec:hospital_model}
-
-The second component for forecast modeling is based on the relationship between hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
-This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}. Figure \ref{fig:lag_scatter}
-shows scatterplots of the difference between hospitalizations and scaled 1 week 
-lags by ILI percentage at the US national level. The points in the left plot are
-from hospitalization differences whereas in the right the hospitalizations were
-log transformed. The positive and slightly curved relationship suggests model
-(\ref{eq:final_hosp}) is a reasonable choice.
-
-\begin{figure}
-    \includegraphics[scale=.46]{Images/lagdiff_vs_ili.png}
-    \caption{Scatterplots of national hospitalizations (left) or log hospitalizations (right) lag differences by ILI \% where the lag is scaled by $\phi$.
-Points are colored by season.
-For hospitalizations $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$.
-}
-    \label{fig:lag_scatter}
-\end{figure}
-
-\begin{equation}
-    \begin{aligned}
-    \label{eq:final_hosp}
-    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
-    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, \omega_s) %N(0, \sigma_{\epsilon}^2)
-    \end{aligned}
-\end{equation}
-In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
-
-This model is for any location, and for fitting purposes $ILI_{s,w}$ is always multiplied by $P$ which is proportional to the population of the state or territory, in this case the total population divided by 50,000.
-This is done as a means of scaling so that the prior distribution assigned to  $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, $\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all states.
-
-Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also fit via Bayesian posterior updating.
-To obtain forecasts for $H_{s, w^* + i}$, the ILI posterior predictive distribution is used along with the posterior distribution for the parameters in (\ref{eq:final_hosp}).
-We considered three scenarios for model (\ref{eq:final_hosp}).
-A model where $D_s$ belongs to the normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, or $H_{s,w} + 1$ is lognormally distributed (LNORM).
-The population value $P$ is excluded from the LNORM model, and we set $\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting.
-In the LNORM model, if the linear parameters are not set the same for seasons 22 and 23, the final variance was more prone to be extreme.
-Besides varying the distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ or there is no quadratic ILI term.
-
-\subsection{Prior selection}
-
-The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in \cite{osthus2019dynamic} and \cite{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem.
-For model (\ref{eq:ili_model}), parameters which are common even when using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, $\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$.
-For the SIR function $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$.
-For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
-
-The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by \cite{gelman2006prior}, the priors were assigned to the variance parameters.
-Univariate parameters were assigned either a normal distribution prior if the support is on $\mathbb{R}$, a half-normal prior if the support is nonnegative, or a truncated-normal prior to match a more specific support.
-Under the ASG model, $\theta_s$ is modeled hierarchically over seasons so that for each season the transformed parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are assigned to $\theta$ and $\Sigma$.
-
-Additional prior constraints were made to improve parameter identifiability.
-In \cite{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$.
-The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative priors.
-To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the parameters.
-A modular Bayesian approach involves multiple steps of parameter fitting where some parameters may be estimated without priors via maximum likelihood estimation or other means.
-Fitting the rest of the model parameters involves assigning priors and conditioning on the previously fit parameters.
-This has been done in computer modeling to improve identifiability and other issues, though \cite{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}.
-We carried out the modular fit by first estimating the maximum likelihood estimate (MLE) for the parameter $\lambda_s$ for each season in (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
-
-% \spencer{Blue stuff below will be either included further on or will be removed entirely
-
-\subsection{Parameter estimation and posterior predictive sampling}
-\label{sec:implementation_posterior}
-
-The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the \cite{stan2024manual} \cite[]{gabry2022stan}.
-Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}.
-The \texttt{cmdstanr} package provides several diagnostic statistics for assessing the sampler.
-As mentioned, most model parameter prior distributions were intended to be uninformative.
-Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material \cite[]{wadsworth2024bas}.
-
-We assessed the model fit for four ILI models.
-These included the SIR and ASG models and models with and without discrepancy modeling.
-When discrepancy is included, the models are denoted as SIRD and ASGD.
-These models were fit using US national data from 2010 to 2023 flu seasons, where data from the 2020 season was excluded because of the unique behavior during that season.
-Assessment for hospitalization modeling was done for six different models.
-These include NORM, LNORM, and LST models, and models where a quadratic ILI term is included or excluded.
-To assess posterior sampling convergence, models were fit to data where ILI and hospitalization data up to week 14 of the 2023 season was included.
-Sampling was done with four chains where from each chain 60,000 posterior draws were sampled, and the first 10,000 draws were discarded as a burn-in.
-The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for each parameter.
-Tables \ref{tab:ili_diagnostics} and \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum $ESS$ over all parameters for ILI and hospitalization models respectively.
-Forecast models for all other weeks of the season were fit using one chain of 60,000 draws where the first 10,000 draws were discarded as a burn-in.
-% \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
-For parameters of the ASG models that were prone to cause trouble in posterior sampling, the starting values were set to be the MLEs.
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI models fitted on US data for week 14 of the 2023 season}
-\label{tab:ili_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccc}
-     & ASG & ASGD & SIR & SIRD \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
-     $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
-    \end{tabular}
-    \end{center}
-\end{table}
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six hospitalization models fitted on US data for week 14 of the 2023 season}
-\label{tab:hsop_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccccc}
-     & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$\\
-     $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
-    \end{tabular}
-    \end{center}
-\end{table}
-
-To obtain forecast distributions of hospitalizations, draws from the posterior predictive distribution from the ILI model were used in conjunction with the posterior distribution of the hospitalizations model.
-When fitting model (\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were saved.
-Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal distributions for the model parameters were saved.
-To obtain forecast distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following steps are repeated $K$ times where $K$ is an integer for the number of desired samples.
-We set $K = 50,000$.
-
-\begin{quote}
-\begin{enumerate}[Step 1:]
-  \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
-  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior distributions
-  \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
-  \begin{itemize}
-    \item[] $\mu_{s,w + i}^* = \alpha_{0s}^* + \alpha_{1s}^* (ILI_{s,w + i}^* \times P) + \alpha_{2s}^* (ILI_{s,w + i}^* \times P)^2 + \phi^* H^*_{s,w + i - 1}$
-  \end{itemize}
-   \item Repeat step 3 for $i \in \{1,2,3,4\}$ to obtain $H^*_{s,(w + 1):(w + 4)}$
-  \item Repeat steps 1-4 $K$ times
-\end{enumerate}
-\end{quote}
-The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast for hospitalizations at week $w + i$.
-For the forecast competition analysis in section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were set to 0 to reflect realistic values of hospitalizations and comply with the FluSight forecasting rules.
-
-\section{Simulation Study} \label{sec:simulation2}
-In this section, we present a simulation study conducted for comparing ILI models and further assessing the hospitalization forecast model. US ILI data is used, and hospitalization data is simulated. 
-A leave-one-season-out (LOSO) approach was combined with a Monte Carlo simulation approach.
-For each replication, we simulated log-hospitalizations for all weeks during seasons 2010, ..., 2022, excluding 2020, using the existing ILI data as a predictive covariate.
-Each season was in turn ``left-out" and treated as if it was the most recent season which we desired to forecast.
-Fitting and forecasting was then done for weeks 14, 20, 26, 32, and 38 of the left out season, giving two weeks that tend to occur as flu cases increase, two as cases decrease, and one that occurs when the cases may be increasing or decreasing.
-Week 20 is a week leading up to the holiday week 22 where ILI typically has a local peak.
-We were particularly interested in how important modeling discrepancy is for forecasting at week 20.
-
-For the simulation of hospitalizations, the parameters $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and $\sigma^2_{\epsilon_s}$ from the hospitalization model in (\ref{eq:final_hosp}) were considered the same across all seasons so that all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$.
-The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization data from the 2022 season.
-For fitting, the hospitalization data was first log-transformed.
-We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
-and assigned the noninformative prior $p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$.
-The marginal posterior distribution $\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ was then the established posterior multivariate normal distribution and $\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior distribution \cite[]{gelman2013bayesian}.
-The posterior means of those parameters were used as the values from which log-hospitalizations were simulated.
-The number of replicates in the simulation was 500.
-
-Model comparison was done by calculating the continuous ranked probability score (CRPS) and the logarithmic score (LogS) for each forecast.
-These scores are both proper scoring rules which evaluate the forecast distribution and density functions respectively.
-Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}.
-Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The CRPS is defined in (\ref{eq:crps2}) and the LogS in (\ref{eq:logs2}).
-Here $F(\cdot)$ is the forecast distribution function, $f(\cdot)$ is the forecast density function, and $y^*$ is the observed targeted value of the forecast.
-The orientation of both scores is negative, meaning the smaller the score the better.
-Here $\vmathbb{1}\{\cdot\}$ is the indicator function.
-
-\begin{equation}
-    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
-    \label{eq:crps2}
-\end{equation}
-
-\begin{equation}
-    \text{LogS}(f, y^*) = -\text{log}(f(y^*))
-    \label{eq:logs2}
-\end{equation}
-Both the CRPS and LogS are calculated using the \texttt{scoringRules} package in \texttt{R} \cite[]{jordan2019scoring}.
-To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation.
-The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
-
-Figures \ref{fig:crps_by_week_horizon} and \ref{fig:logs_by_week_horizon} show 
-boxplots of results from the simulation. The figures respectively
-show boxplots of the CRPS and LogS for the 500 simulation replicates faceted by week and horizon.
-Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
-Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with lower medians and lower variation than the two models not accounting for ILI discrepancy.
-This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough.
-This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
-Among the four models, ASG tends to have the lowest values in CRPS and LogS though this is not always the case.
-This suggests that the additional flexibility of the five parameter ASG function is often better for forecasting than that of the three parameter SIR function.
-Additional
-plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and LogS.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{Images/crps_by_week_horizon.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
-    \label{fig:crps_by_week_horizon}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{Images/logs_by_week_horizon.png}
-    \caption{Boxplots of the logarithmic score (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
-    \label{fig:logs_by_week_horizon}
-\end{figure}
-
-\section{Analysis of forecasts for 2023 flu season}
-\label{sec:analysis}
-
-In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations.
-The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$.
-A single forecast is thus comprised of 11 predictive intervals and a median.
-Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, and forecasts were made at the state and national level.
-The first week of forecasting took place during the week of October 7, 2023, and the final week was the week of April 27, 2024 making 29 total weeks of forecasts.
- The same format was used during the 2021 and 2022 seasons and for the COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
-% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
- Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS).
-The focus in this section will be on the LWIS (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
-
-The WIS is a proper scoring rule used for scoring quantile or interval forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all included quantiles, $B$ is the number of intervals,  $y^*$ is the observed value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ interval.
-$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a single interval as defined in (\ref{eq:is}).
-The goal of the forecaster is to minimize the WIS.
-The LWIS is the same as the WIS except that it is evaluated over the log of quantiles and the log of the observed value.
-
-\begin{equation}
-\label{eq:wis2}
-        WIS_{0,B}(Q, y^*) = \frac{1}{B + 1/2} \times (w_0\times |y^* - median| + \sum_{b=1}^B \{w_k \times IS_{\alpha_b}(Q, y^*) \} )
-\end{equation}
-
-\begin{equation}
-\label{eq:is}
-        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\vmathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \vmathbb{1}\{y^* > r\}
-\end{equation}
-
-We fit 24 forecast models for each location for all 29 weeks, and for each week forecast 1-4 week ahead hospitalizations.
-The 24 models included all combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST hospitalization models, and both quadratic and linear hospitalization models.
-The prior distributions under the SIR model are in (\ref{eq:sir_prior}).
-Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$.
-Here $\vmathbb{1}_{A}$ represents the indicator function for values within the set $A$.
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:sir_prior}
-        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
-        \beta_s &\sim N^+(0.8, 0.3) \\
-        \rho_s &\sim N^+(0.68, 0.08)
-\end{aligned}
-\end{equation}
-For the ASG model, the MLE $\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value.
-The remaining estimates were used as starting values for posterior sampling.
-The transformation $T(\theta_s) = (\text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ was made, and a prior distribution was assigned to $T(\theta_s)$.
-The prior distributions for the ILI model under the ASG function are shown in (\ref{eq:ili_prior}).
-These are slightly informative priors because for most parameters we have an idea what reasonable values may be.
-Here $m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
-where $\text{diag}(\cdot)$ 
-is the diagonal matrix for the given entries.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:ili_prior}
-                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
-                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
-                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
-                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
-\end{aligned}
-\end{equation}
-Parameters shared by both the SIR and ASG models are the scale parameter $\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-These priors are in (\ref{eq:shared_ili_priors}).
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:shared_ili_priors}
-        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
-        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2)
-\end{aligned}
-\end{equation}
-Because of the limited information for estimating $\sigma_{\gamma_W}^2$, we selected an informative prior distribution by first estimating  $\hat{\sigma}_{\gamma_W}^2$.
-This was estimated for a given state by first calculating the MLE for $\theta_s$.
-% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
-%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
-$\widehat{ILI}_{s,w}$ was then predicted such that $\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season.
-Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over seasons of the differences $\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$.
-When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was calculated using the \texttt{mle2} function in the \texttt{bblme} package \cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} package \cite[]{soetaert2010desolve} function in \texttt{R}.
-Where the ASG function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} function.
-The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
-
-\begin{equation}
-\begin{aligned}
-\label{eq:shared_hosp_priors}
-        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
-        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
-        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
-\end{aligned}
-\end{equation}
-
-\begin{figure}
-\centering
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the hospitalization models are all normally distribution.
-The figure includes hospitalization forecasts where ILI is a linear predictor (left) and where ILI is a quadratic predictor (right).
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-\label{fig:normal_flu_forecasts}
-\end{figure}
-
-Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US flu hospitalizations during the 2023 season under the NORM hospitalization model.
-The forecasts shown are for weeks 14, 20, 26, and 32.
-The predictive bands are the 50\% and 95\% predictive intervals.
-These plots show a tendency to often underpredict hospitalizations.
-The quadratic models appear to do a better job predicting  hospitalizations at the season peak but a poorer job predicting after the peak, whereas the linear models seem to predict well or slightly overpredict after the peak.
-% From these figures, it's difficult to say whether ASG or SIR models predict better, but at least for the quadractic models the week 22 predictions are maybe slightly better than for the linear models.
-
-Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the season for all 24 models of US hospitalizations.
-LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling.
-Here, a darker value represents a larger LWIS and worse forecast compared to lighter values.
-The weeks around the holiday week 22 are highlighted by a red band.
-In most cases, there appears to be a turning point in performance at or near week 22.
-The models which include discrepancy appear to forecast better around week 22, though they may not outperform the models without discrepancy through the whole season.
-Such is the case for the NORM and LST quadratic models and the linear LNORM models.
-It's also notable that for the SIR model, including discrepancy appears to make for poorer forecasts after week 22 whereas for ASG, including discrepancy appears to improve forecasts.
-
-\begin{figure}
-
-    % \centering
-    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
-    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
-Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
-Scores for models with an ASG ILI model are above while those with an SIR model are below.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:us_lwis}
-\end{figure}
-
-\begin{figure}
-
-    \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
-Scores are averaged over all horizons 1-4 weeks ahead.
-Scores are faceted by hospitalization model distribution.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:lwis_by_dist_loc}
-\end{figure}
-
-Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by LWIS, but all 53 locations are included.
-The LWIS is shown across the whole season faceted by hospitalization model distribution.
-Overall, the weeks leading up to week 22 are the most difficult to forecast, but the LNORM model appears to perform the better than the other models during the weeks around the holiday week 22.
-The overall, calculated as the mean LWIS over all locations and weeks, for all 24 models is shown in table \ref{tab:forecast_scores}.
-The first main takeaway is that the top four performing models are SIR ILI models with the LNORM hospitalization models.
-The next main takeaway is that overall the models which include discrepancy outperformed the models without discrepancy.
-
-\begin{table}
-\caption{Overall scores for each of the 24 forecast models.
-The overall score is the log weighted interval score (LWIS) averaged over all locations, weeks, and horizons.
-The scores in the first two rows are for linear models, and the scores in the third and fourth rows are for quadratic models.
-The lowest WISs are bolded.}
-\begin{tabular*}{\textwidth}
-{@{\extracolsep{\fill}} 
-    l*{9}{c}}
-  & & \multicolumn{3}{c}{ASG} 
-  & \multicolumn{3}{c}{SIR} \\ 
-  \cmidrule{3-5} \cmidrule{6-8}
-  & & LNORM & LST & NORM & LNORM & LST & NORM\\
-  \midrule
-  Linear & No Disc & 0.366 & 0.392 & 0.365 & \textbf{0.355} & 0.416 & 0.394 &\\ 
-   & Disc & 0.358 & 0.391 & 0.365 & \textbf{0.343} & 0.382 & 0.354 &\\
-  \midrule
-  Quadratic & No Disc & 0.366 & 0.398 & 0.377 & \textbf{0.355} & 0.401 & 0.377 &\\ 
-   & Disc & 0.358 & 0.390 & 0.369 & \textbf{0.343} & 0.378 & 0.356 &\\     
-  \bottomrule
-\end{tabular*}
-\label{tab:forecast_scores}
-\end{table}
-
-It should not be assumed that these two takeaways will apply for future flu seasons.
-As suggested by the simulation study in the previous section, it may often be the case that the ASG ILI model is better for forecasting.
-Indeed, a close examination of figures \ref{fig:us_lwis} and \ref{fig:lwis_by_dist_loc} shows that it is common for the ASG models to show better forecasting skill than the SIR models.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
- In this manuscript we introduce a statistical modeling framework which allows for the incorporation of several ILI forecast modeling methods.
-Specifically, we built upon \cite{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy.
-We model flu hospitalizations by incorporating the ILI forecast model into a model forecasting hospitalizations where hospitalization predictions are a linear or quadratic function of ILI.
-
-The simulation study in section \ref{sec:simulation2} suggests the ASG function in ILI modeling may slightly outperform the SIR model according to the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season forecasts, the SIR model was overall superior.
-The results from both the simulation study and the real data analysis suggest that the addition of a discrepancy component in ILI modeling may improve forecasts especially near the holiday week between Christmas and New Years day.
-It should not be assumed that these conclusions may be generalized for all locations in the US, weeks of a season, or for future flu seasons.
-% Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.
-
-Forecasting the seasonal influenza outbreak remains a challenging task for forecasters.
-The general modeling framework in this manuscript is successful under diverse modeling conditions for all locations in the US and may contribute to future forecasting efforts.
-All forecast models were presented as separate from one another, but in the various locations and times of the flu season, different models perform better than others.
-To build on the work done here, a natural step forward would be to combine all or a few selected forecasts into an ensemble forecast.
-Such an ensemble may work to cancel out certain model biases or highlight model strengths, leading to more robust forecasts.
-
-\begin{acks}[Acknowledgments]
-This work is partially supported by the National Science Foundation under Grant No.
-2152117.
-Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
-\end{acks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Funding information, if any,             %%
-%% should be provided in the                %%
-%% funding section.                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{funding}
-% The first author was supported by NSF Grant DMS-??-??????.
-% 
-% 
-% \end{funding}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Supplementary Material, including data   %%
-%% sets and code, should be provided in     %%
-%% {supplement} environment with title      %%
-%% and short description. It cannot be      %%
-%% available exclusively as external link.  %%
-%% All Supplementary Material must be       %%
-%% available to the reader on Project       %%
-%% Euclid with the published article.       %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{supplement}
-\stitle{1. Posterior distribution plots for select parameters}
-\sdescription{Plots of posterior distributions for parameters in both ILI and hospitalization models}
-\end{supplement}
-\begin{supplement}
-\stitle{2. FluSight forecast competition scoring results}
-\sdescription{Flu forecast results with relative weighted interval scoring rule similar to that used in  FluSight}
-\end{supplement}
-
-\bibliographystyle{ba}
-\bibliography{master_bib}
-
-\end{document}

### 2025-06-30 manuscript_res.tex
fix full fit

--- 2024-12-18 nonlinear_forecast/manuscript.tex
@@
-\documentclass{article}
-
-\usepackage{PRIMEarxiv}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage[linktocpage=true]{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-% \usepackage{lipsum}
-\usepackage{fancyhdr}       % header
-\usepackage{graphicx}       % graphics
-\graphicspath{{media/}}     % organize your images and other figures under media/ folder
-
-% Packages added by Spencer
-\usepackage{amsmath,amsthm,amssymb,scrextend}
-\usepackage{appendix}
-\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}
-\usepackage{enumerate}
-\usepackage{csquotes}
-% \usepackage{subcaption}
-\usepackage{amssymb}
-% \usepackage{bbm}
-\usepackage[bb=pazo]{mathalpha}
-\usepackage{bm}
-\usepackage{newtxmath} %for making indicator function
-\usepackage{booktabs} %for the table in the analysis section
-% \usepackage{enumerate} %for the short algorithm
-\usepackage{subfigure} %for making plots with multiple images
-\usepackage{tabularx} %for the spacing in tabular
-
-\pagestyle{empty}
-
-\newcommand\Bbbbone{%
-  \ifdefined\mathbbb%
-    \mathbbb{1}%
-  \else%
-    \boldsymbol{\mathbb{1}}%
-  \fi}
-
-% \usepackage{xr}
-% \externaldocument{supplementary}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \usepackage[pdftex]{graphicx}
-% Standard, old-style thesis
-% \usepackage{isutraditional}   
-% \chaptertitle
-% Old-style, thesis numbering down to subsubsection
-% \alternate
-% \usepackage{rotating}
-% Bibliography without numbers or labels
-\usepackage{natbib}
-% Use the following line if you want square brackets and numbering system
-%\usepackage[square, numbers]{natbib}
-
-% this package is for getting certain math scripts like the 1 needed for an indicator function
-% \usepackage{bbm}
-
-% using this package to make a better looking list
-\usepackage{enumerate}
-% try this one for a list as well. may be better
-% \usepackage[font=itshape]{quoting}
-
-% package for tables
-
-% \renewcommand{\bibfont}{\setstretch{1}\selectfont}
-% \setlength{\bibsep}{13.2pt}
-
-\usepackage{chapterbib}
-%\renewcommand\bibsection{\section*{\ref name}}
-%\renewcommand\bibsection{\section*}
-\renewcommand{\bibsection}{\section{References}}
-
-%\bibliographystyle{apa}  % changed from apa, abbrv, unsrtnat
-%\includeonly{titletoc,chapter1}
-%Optional Package to add PDF bookmarks and hypertext links
-% \usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
-\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
-
-%\usepackage{titletoc}
-% \usepackage{hyperref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Packages/commands added by Jarad
-\usepackage{xcolor}
-\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
-\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}
-
-%Header
-\pagestyle{fancy}
-\thispagestyle{empty}
-\rhead{ \textit{ }}
-
-% Update your Headers here
-\fancyhead[LO]{Bayesian nonlinear flu hospitalizations forecasting}
-% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}
-
-%% Title
-\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy
-%%%% Cite as
-%%%% Update your official citation here when published 
-% \thanks{\textit{\underline{Citation}}: 
-% \textbf{Authors. Title. Pages.... DOI:000000/11111.}} 
-}
-
-\author{
-  Spencer Wadsworth \\
-  Iowa State University \\
-  Ames, Iowa\\
-  \texttt{sgw96@iastate.edu} \\
-  %% examples of more authors
-   \And
-  Jarad Niemi \\
-  Iowa State University \\
-  Ames, Iowa\\
-  \texttt{niemi@iastate.edu} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread.
-The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods.
-For the first several years of the competition, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalization counts.
-Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons.
-In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data.
-The first component is for modeling ILI data using a dynamic nonlinear Bayesian hierarchical model.
-The second component is for modeling hospitalizations as a function of ILI.
-For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor.
-In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from Osthus et al.
-% \cite[]{osthus2019dynamic} 
-% 
-and a nonlinear Bayesian hierarchical model from Ulloa 
-% \cite[]{ulloa2019}
-are compared.
-Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model.
-Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared.
-We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year.
-We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
-\end{abstract}
-
-% \begin{abstract}
-% The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread.
-The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods.
-Beginning in 2013, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalizations.
-Reliable hospitalization data has only been available since 2021, but ILI data has been available since 2010 and has been successfully forecast for several seasons.
-In this manuscript, we introduce a two component modeling framework for forecasting hospitalizations utilizing both hospitalization and ILI data.
-The first component is for modeling ILI data using a nonlinear Bayesian model.
-The second component is for modeling hospitalizations as a function of ILI.
-For hospitalization forecasts, ILI is first forecast then hospitalizations are forecast with ILI forecasts used as a predictor.
-In a simulation study, the hospitalization forecast model is assessed and two previously successful ILI forecast models are compared.
-Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model.
-Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared.
-We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year.
-We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
-% \end{abstract}
-
-% \begin{keyword}[class=MSC]
-% \kwd[Primary ]{00X00}
-% \kwd{00X00}
-% \kwd[; secondary ]{00X00}
-% \end{keyword}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% `\ ' is used here because TeX ignores    %%
-%% spaces after text commands.              %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 
-Every year the seasonal influenza outbreak burdens the public health system by infecting millions, causing an influx of primary care visits and hospitalizations and leading to between 290,000 and 650,000 deaths worldwide \cite[]{whoflufact2023}.
-Molinari et al.
-\cite[]{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion.
-Accurate forecasting of infectious diseases can inform public decision making and ease the burden of an outbreak \cite[]{turtle2021accurate, lutz2019applying}.
-There is a growing consensus that disease forecasts should be probabilistic in nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has been shown that reporting forecast uncertainty along with predictions may lead to better decision making \cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.
+Every year the seasonal influenza outbreak burdens the public health system by 
+infecting millions, causing an influx of primary care visits and 
+hospitalizations and leading to between 290,000 and 650,000 deaths worldwide 
+\cite[]{whoflufact2023}. Molinari et al. \cite[]{molinari2007annual} estimated 
+the United States' annual economic burden from medical costs, loss of income, 
+and deaths to be over \$87 billion. Accurate forecasting of infectious diseases 
+can inform public decision making and ease the burden of an outbreak 
+\cite[]{turtle2021accurate, lutz2019applying}.
+There is a growing consensus that disease forecasts should be probabilistic in 
+nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has 
+been shown that reporting forecast uncertainty along with predictions may lead 
+to better decision making 
+\cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.
 
@@
 
-To better inform public decision making regarding the flu epidemic, in 2013 the United States Centers for Disease Control and Prevention (CDC) organized a national flu forecasting competition, also known as FluSight \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}.
-Originally, over a dozen teams of researchers from academic and industry backgrounds participated in FluSight by contributing their own forecast models.
-Besides the 2020 season --or the flu season spanning the fall of 2020 and the winter of 2021-- FluSight has been operated annually and researchers outside the CDC have been invited to participate.
-Initially the target data for forecasts was influenza-like illness (ILI) data.
-ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}.
-The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, Osthus et al.
-\cite[]{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework.
-An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
-
-The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications.
-As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}.
-The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded.
-The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}.
-Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition.
-Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}.
-This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.
-
-The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models.
-The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
-Herein we use ILI models similar to those in Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
-The model of Osthus et al.
-\cite[]{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework.
-Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by Kennedy and O'Hagan \cite[]{kennedy2001bayesian}.
-The model in Ulloa \cite[]{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data.
-Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.
-
-In line with the newer FluSight standard of forecasting hospitalizations, we model hospitalizations as a linear function of ILI.
-Thus forecasts produced herein target flu hospitalizations and are a mapping of ILI forecasts to hospitalizations.
-This allows for ILI data from many seasons to be exploited and for ILI forecasts to assist in forecasting hospitalizations, which has fewer seasons of data than ILI.
-Several modeling schemes and their forecasts for the 2023 flu hospitalization season are compared, and it is shown that the modeling decisions produce good forecast results for different states or times during the flu season.
-
-In section \ref{sec:data} we review the ILI and hospitalization data provided by the CDC and targeted by FluSight. 
-In section \ref{sec:functions} the modeling framework contributed by this manuscript is given.
-In the same section, functions similar to those used by Osthus et al.
-\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} are defined.
- These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively.
-Model fitting and implementation are described at the end of the section.
-Section \ref{sec:simulation2} includes a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared.
-Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts.
-Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}.
- Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.
+To better inform public decision making regarding the flu epidemic, in 2013 the 
+United States Centers for Disease Control and Prevention (CDC) organized a 
+national flu forecasting competition, also known as FluSight 
+\cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. 
+Originally, over a dozen teams of researchers from academic and industry 
+backgrounds participated in FluSight by contributing their own forecast models. 
+Besides the 2020 season --or the flu season spanning the fall of 2020 and the 
+winter of 2021-- FluSight has been operated annually and researchers outside 
+the CDC have been invited to participate. Initially the target data for 
+forecasts was influenza-like illness (ILI) data. ILI is the proportion of 
+patients who meet a healthcare provider and who display flu like symptoms, 
+and ILI data has been available at the state and national level since the 2010 
+flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}. 
+The collaborative ILI forecasting effort has led to a number of modeling 
+developments in flu forecasting 
+\cite[see references therein for more examples]{mcandrew2021adaptively, 
+osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's 
+introduction, Osthus et al.  \cite[]{osthus2019dynamic} categorized the most 
+commonly used flu forecasting models into four classes including mechanistic 
+models based on differential equation compartmental models, agent based models 
+based on population simulation, machine learning/regression models including 
+data driven machine learning and statistical models, and data assimilation 
+models which are constructed by assimilating mechanistic models into a 
+probabilistic framework. An additional forecast model used in FluSight 
+involves the combination of several forecasts into a single ensemble forecast, 
+which has been shown to perform well relative to individual models 
+\cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
+
+The administration of FluSight saw few changes during the first seven seasons
+between 2013 and 2019, 
+but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 
+forecasting led to major modifications. As a result of the COVID-19 pandemic 
+which began during the 2019 flu season, the typical flu outbreak behavior was 
+altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}. 
+The COVID-19 pandemic led to the creation of the Health and Human Services 
+(HHS) Patient Impact and Hospital Capacity Data System 
+\cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization 
+data, and the COVID-19 Forecast Hub was founded. The COVID-19 Forecast Hub 
+was based on FluSight but with certain major adjustments including how the 
+forecast uncertainty is represented and the addition of the weekly publication 
+of a multi-model ensemble forecast as the official forecast of the CDC 
+\cite[]{bracher2021evaluating, Cramer2022-hub-dataset}. Using estimated 
+quantiles for representing forecast uncertainty and creating a multi-model 
+ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by 
+the flu forecast competition. Additionally the target of the flu forecasts 
+changed from being ILI data to being HHS hospitalization data, which reports 
+the number of hospitalizations due to a laboratory confirmed flu infection  
+\cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of 
+having COVID-19 cases in the population making ILI data, already only a proxy 
+for flu behavior, more difficult to interpret.
+
+The contribution of this manuscript is to introduce a two component framework 
+for modeling HHS hospitalization forecasts where hospitalization data and 
+years of ILI data are used to inform forecast models. The first modeling 
+component is a model of ILI data and the second is a model of hospitalization 
+data with ILI as a predictive covariate.
+Herein we use ILI models similar to those in Osthus et al. 
+\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
+The model of Osthus et al. \cite[]{osthus2019dynamic} is a combined data 
+assimilation and statistical regression model which involves a compartmental 
+model in a probabilistic framework. Their model also includes an additional 
+component for capturing a systematic discrepancy between the deterministic 
+part of the model and the actual data, an idea which was first introduced by 
+Kennedy and O'Hagan \cite[]{kennedy2001bayesian}. The model in Ulloa 
+\cite[]{ulloa2019} is a Bayesian hierarchical regression model with an 
+underlying function intended to capture the trajectory of the seasonal 
+ILI data. Herein, we provide a framework under which discrepancy modeling may 
+be used along with a general function modeling ILI data, and we show the 
+effectiveness of including discrepancy modeling during certain periods of 
+the flu season.
+
+In line with the newer FluSight standard of forecasting hospitalizations, we 
+model hospitalizations as a linear function of ILI. Thus forecasts produced 
+herein target flu hospitalizations and are a mapping of ILI forecasts to 
+hospitalizations. This allows for ILI data from many seasons to be exploited 
+and for ILI forecasts to assist in forecasting hospitalizations, which has 
+fewer seasons of data than ILI. Several modeling schemes and their forecasts 
+for the 2023 flu hospitalization season are compared, and it is shown that the 
+modeling decisions produce good forecast results for different states or times 
+during the flu season.
+
+In section \ref{sec:data} we review the ILI and hospitalization data provided 
+by the CDC and targeted by FluSight. 
+In section \ref{sec:functions} the modeling framework contributed by this 
+manuscript is given. In the same section, functions similar to those used by 
+Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} are 
+defined.
+ These functions are the susceptible-infectious-recovered (SIR) compartmental 
+ model and the asymmetric Gaussian (ASG) function  respectively. Model fitting 
+ and implementation are described at the end of the section. Section 
+ \ref{sec:simulation2} includes a simulation study where four ILI forecast 
+ models and their use in forecasting hospitalizations are compared. Commonly 
+ used proper scoring rules \cite[]{gneiting2007strictly}, which are also 
+ introduced and defined in section \ref{sec:simulation2}, are used for 
+ comparing the forecasts. Forecasting of the 2023 flu outbreak along with 
+ assessment and comparison under several selected models is performed in 
+ section \ref{sec:analysis}. 
+ Finally, the manuscript is concluded in section \ref{sec:conclusion} with 
+ general observations and some discussion.
 
  \section{Flu outbreak data} \label{sec:data}
-In this section we introduce and define ILI and hospitalization data and visually evaluate the data.
-ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season.
-Both of these data were collected at the state, territorial, and national level and were reported at least weekly.
-Overall the data is reported for 53 locations including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), and at the US national level.
-We will refer to forecast targets throughout this manuscript.
-A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a specific location and week during the season.
+In this section we introduce and define ILI and hospitalization data and 
+visually evaluate the data. ILI and hospitalization data have been the object 
+of forecasting for FluSight with ILI being the target for the first seven 
+seasons and hospitalizations being the target since the 2022 season. Both of 
+these data were collected at the state, territorial, and national level and 
+were reported at least weekly. Overall the data is reported for 53 locations 
+including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), 
+and at the US national level. We will refer to forecast targets throughout this 
+manuscript. A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a 
+specific location and week during the season.
 
 \subsection{Influenza-like illness data}
-The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers.
-Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases.
-An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}.
-Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza.
-To know whether or not a sick patient is infected with influenza requires a laboratory test.
+The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects 
+information on respiratory illness from outpatient visits to health care 
+providers. Over 3,400 outpatient health care providers in all 50 US states, PR, 
+DC, and the US Virgin Islands report each week the total number of outpatient 
+visits along with the number of ILI cases. An ILI case is defined as a 
+``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
+and/or a sore throat." Prior to the 2021 season, the definition included 
+``without a known cause other than influenza" \cite[]{cdcfluview2023}. Because 
+other illnesses such as COVID-19, RSV, and the common cold may induce similar 
+respiratory symptoms, ILI may include patients infected with some disease 
+other than influenza. To know whether or not a sick patient is infected with 
+influenza requires a laboratory test.
 
 In 2013, when FluSight began, the ILI data was the object of the forecasts. 
-The data was released publicly at HHS region levels, and forecast teams were asked to provide forecasts of several ILI targets on the regional levels including season onset, 1-4 week ahead ILI levels, and the week of peak ILI activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}.
-Currently, the ILI data is collected by the CDC and published on an online portal for viewing at the national, HHS region, census, and state levels \cite[]{cdc2024fluviewportal}.
-To obtain ILI data, we used the \texttt{R} package \texttt{cdcfluview} which provides functions for downloading the data \cite[]{rudis2021cdcfluview}.
-Weekly ILI data from the national, HHS region, and census levels are available from the 1997 flu season until the current season.
-At the state level, data is available from the 2010 flu season to the current season.
-
-The top of figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023.
-For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons.
-To better align with the flu behavior, week 1 is set as the first week of August and week 52 or 53 is the last week in July of the following year.
-For example week 1 of the 2013 season corresponds to the first week of August 2013, and week 52 of the same season corresponds to the last week of July 2014.
-This convention is used for the remainder of this manuscript.
+The data was released publicly at HHS region levels, and forecast teams were 
+asked to provide forecasts of several ILI targets on the regional levels 
+including season onset, 1-4 week ahead ILI levels, and the week of peak ILI 
+activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}. Currently, 
+the ILI data is collected by the CDC and published on an online portal for 
+viewing at the national, HHS region, census, and state levels 
+\cite[]{cdc2024fluviewportal}. To obtain ILI data, we used the \texttt{R} 
+package \texttt{cdcfluview} which provides functions for downloading the data 
+\cite[]{rudis2021cdcfluview}. Weekly ILI data from the national, HHS region, 
+and census levels are available from the 1997 flu season until the current 
+season. At the state level, data is available from the 2010 flu season to the 
+current season.
+
+The top of figure \ref{fig:us_ili} shows the ILI data at the national level for 
+flu seasons 2010 to 2023. For most seasons there are 52 weeks listed, but for 
+the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 
+Sundays during those seasons. 
+To better align with the flu behavior, week 1 is set as the first week of 
+August and week 52 or 53 is the last week in July of the following year. For 
+example week 1 of the 2013 season corresponds to the first week of August 2013, 
+and week 52 of the same season corresponds to the last week of July 2014. This 
+convention is used for the remainder of this manuscript.
 
@@
     \includegraphics[scale=.4]{Images/us_hospitalizations.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023 (top) and weekly flu confirmed hospitalization counts at the national level for 2022 and 2023 flu seasons (bottom).
-Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.}
+    \caption{Percentage of outpatient visits with an influenza-like illness 
+    (ILI) in the US for seasons 2010 to 2023 (top) and weekly flu confirmed 
+    hospitalization counts at the national level for 2022 and 2023 flu seasons 
+    (bottom). Week 1 is the first week of August of the year the flu season 
+    begins and the last week of the season is the last week of July of the 
+    following year.}
     \label{fig:us_ili}
@@
 
-Notable from the plots in the top of figure \ref{fig:us_ili} is the regular trajectory of the ILI.
-With the exception of season 2020, the ILI begins low at week 1 and increases as the fall and winter progress until the ILI reaches a peak.
-As spring progresses to summer, the ILI decreases to low values.
-As Osthus et al.
-\cite[]{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day.
-Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
+Notable from the plots in the top of figure \ref{fig:us_ili} is the regular 
+trajectory of the ILI. With the exception of season 2020, the ILI begins low 
+at week 1 and increases as the fall and winter progress until the ILI reaches 
+a peak. As spring progresses to summer, the ILI decreases to low values. As 
+Osthus et al. \cite[]{osthus2019dynamic} point out, there is nearly always 
+either a global or local peak at week 22 which typically corresponds to the 
+week between Christmas and New Year's day. 
+Whether local or global, the ILI holiday peak is generally expected and thought 
+to be due to widespread holiday travel, school closure, or other unique social 
+behavior   
 \cite[]{ewing2017contact, garza2013effect}.
-The only seasons when there was not a peak at week 22 were season 2022 where the season peak occurred particularly early and season 2020 which was greatly influenced by the COVID-19 pandemic.
+The only seasons when there was not a peak at week 22 were season 2022 where 
+the season peak occurred particularly early and season 2020 which was greatly 
+influenced by the COVID-19 pandemic.
 
@@
 
-Hospital admission data, used as the object of FluSight forecasting for the 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported Patient Impact and Hospital Capacity by State Timeseries}.
-Several targets of respiratory illnesses including COVID-19, RSV, and influenza are reported weekly by most hospitals in the US.
-In February 2022 it became mandatory for all hospitals to report the number of COVID-19 and influenza hospitalizations, and since then reporting of hospitalizations has become widespread.
-These data were updated every Wednesday and Friday according to NHSN guidelines \cite[]{healthdata2024covidts}.
-
-The bottom of figure \ref{fig:us_ili} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons.
-These plots show similarities to the ILI plots in the top of the figure in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends.
-For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22.
-Additional plots of ILI data and hospitalization
+Hospital admission data, used as the object of FluSight forecasting for the 
+2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
+Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
+Patient Impact and Hospital Capacity by State Timeseries}. Several targets of 
+respiratory illnesses including COVID-19, RSV, and influenza are reported 
+weekly by most hospitals in the US. In February 2022 it became mandatory for 
+all hospitals to report the number of COVID-19 and influenza hospitalizations, 
+and since then reporting of hospitalizations has become widespread. These data 
+were updated every Wednesday and Friday according to NHSN guidelines 
+\cite[]{healthdata2024covidts}.
+
+The bottom of figure \ref{fig:us_ili} shows the weekly national 
+hospitalizations for the 2022 and 2023 flu seasons. These plots show 
+similarities to the ILI plots in the top of the figure in that at the early 
+weeks of the season hospitalizations are low, but they increase in the fall 
+to a peak after which they decrease until the flu outbreak ends. For both the 
+2022 and 2023 seasons, the hospitalizations peaked during the same week as 
+ILI, and in 2023 that peak occurred during the holiday week 22. Additional 
+plots of ILI data and hospitalization
 data for select states are included in the supplementary material 
-\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar to 
+\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar 
+to 
 those in figure \ref{fig:us_ili} with activity increasing
@@
 
-The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows a similar trajectory.
-Compartmental models are standard mathematical models used for modelling disease outbreaks.
-One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}.
-Ulloa \cite[]{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data.
-The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy.
-In the first part of this section, we present an ILI model similar to the model in Osthus et al.
-\cite[]{osthus2019dynamic}.
-With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.
-
-With the aim of forecasting hospitalizations, we also introduce a linear model of hospitalizations data with ILI as a predictive covariate.
-To forecast hospitalizations, ILI data is first forecasted and the forecast is then plugged in as a covariate in the hospitalization model, thus producing hospitalization forecasts.
-The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.
+The typical behavior of the ILI data which starts at lower values in the late 
+summer and fall but which increases to a peak, usually in December or January, 
+followed by a decline motivates the use of a nonlinear function which follows 
+a similar trajectory. Compartmental models are standard mathematical models 
+used for modelling disease outbreaks. One important compartmental model is the 
+susceptible-infectious-recovered (SIR) model, which is used by some to model 
+ILI data \cite[]{osthus2019dynamic, allen2017primer}. Ulloa \cite[]{ulloa2019} 
+chose to use the asymmetric Gaussian (ASG) function to model ILI data. The 
+SIR and ASG models may both be appropriate to describe ILI behavior over the 
+course of a flu season, however there may also be systematic behavior not 
+captured by either, necessitating an additional model component to capture 
+the discrepancy. In the first part of this section, we present an ILI model 
+similar to the model in Osthus et al. \cite[]{osthus2019dynamic}. With some 
+generalization, the model may incorporate any appropriate nonlinear function, 
+though the focus here is on the SIR and ASG functions which are defined.
+
+With the aim of forecasting hospitalizations, we also introduce a linear 
+model of hospitalizations data with ILI as a predictive covariate. To forecast 
+hospitalizations, ILI data is first forecasted and the forecast is then 
+plugged in as a covariate in the hospitalization model, thus producing 
+hospitalization forecasts. The hospitalization model is also defined in this 
+section, and the section is concluded with descriptions of selected prior 
+distributions, model implementation, and posterior sampling.
 
@@
 The proposed model for ILI for any location is given in (\ref{eq:ili_model}). 
-Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given season.
-The ILI is a proportion, so the Beta random variable is a natural selection for modeling.
-Under the parameterization in of the Beta distribution used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a scale parameter.
-The nonlinear function $f_{\theta_s}(w)$ captures the trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the systematic patterns which $f_{\theta_s}(w)$ does not capture.
+Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, 
+where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given 
+season. The ILI is a proportion, so the Beta random variable is a natural 
+selection for modeling. Under the parameterization in of the Beta distribution 
+used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the 
+variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a 
+scale parameter. The nonlinear function $f_{\theta_s}(w)$ captures the 
+trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the 
+systematic patterns which $f_{\theta_s}(w)$ does not capture. 
 \begin{equation}
@@
 \end{equation}
-In Osthus et al.
-\cite[]{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}.
-In Ulloa \cite[]{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}.
-In Ulloa \cite[]{ulloa2019} modeling is done hierarchically over seasons.
+In Osthus et al. \cite[]{osthus2019dynamic} 
+$f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious 
+compartment of the SIR model from (\ref{eq:sir_diff}) in section 
+\ref{sec:sir_func}. In Ulloa \cite[]{ulloa2019} 
+$f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section 
+\ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically 
+over seasons.
 
@@
 
-The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by Kermack and McKendrick \cite[]{kermack1927contribution}.
-Since then, compartmental models have become standard for modeling infectious diseases \cite[]{allen2008mathematical}, and many extensions have been made and studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}.
-The SIR mathematical model includes three compartments and assumes that at any time $t>0$ every individual in a closed population belongs to exactly one compartment.
-The three compartments are susceptible ($S$), infectious ($I$), and recovered ($R$), and their interaction over the course of an outbreak is described by the differential equations in (\ref{eq:sir_diff}).
+The SIR compartmental model is a mathematical model used for modeling disease 
+outbreaks and was introduced by Kermack and McKendrick 
+\cite[]{kermack1927contribution}. Since then, compartmental models have 
+become standard for modeling infectious diseases 
+\cite[]{allen2008mathematical}, and many extensions have been made and 
+studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}. 
+The SIR mathematical model includes three compartments and assumes that at 
+any time $t>0$ every individual in a closed population belongs to exactly 
+one compartment. The three compartments are susceptible ($S$), infectious ($I$), 
+and recovered ($R$), and their interaction over the course of an outbreak is 
+described by the differential equations in (\ref{eq:sir_diff}).
 
@@
 \end{equation}
-Here $S$, $I$, and $R$ represent the proportion of the population in each compartment such that $S + I + R = 1$ for all $t$.
-The trajectory is determined by the disease transmission rate $\beta > 0$ and the recovery rate $\delta > 0$.
-Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual and the expected rate of recovery to an immune state for a newly infected person.
-Whether or not a disease outbreak is classified as an epidemic is determined by the initial susceptible population $S_0$, or the susceptible population at time $0$, and the parameter $\rho = \delta/
-\beta$.
-If $S_0/\rho > 1$, the outbreak is considered an epidemic.
-It is non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}.
-Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where $S_0$ and $\rho$ were selected to match an outbreak that is an epidemic.
-In the case where $S_0 \leq \rho$, the trajectory for the $I$ compartment would never be increasing.
-The increase to a peak and subsequent decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is reasonable to model ILI by this compartment.
-Thus in modeling the ILI data, we consider the data to be analogous to the $I$ proportion of the population.
+Here $S$, $I$, and $R$ represent the proportion of the population in each 
+compartment such that $S + I + R = 1$ for all $t$. The trajectory is 
+determined by the disease transmission rate $\beta > 0$ and the recovery 
+rate $\delta > 0$. Respectively, these may be thought of as the expected 
+proportion of susceptible individuals who will be infected by an infectious 
+individual and the expected rate of recovery to an immune state for a newly 
+infected person. Whether or not a disease outbreak is classified as an 
+epidemic is determined by the initial susceptible population $S_0$, or the 
+susceptible population at time $0$, and the parameter $\rho = \delta/
+\beta$. If $S_0/\rho > 1$, the outbreak is considered an epidemic. It is 
+non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}. Figure 
+\ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR 
+model where $S_0$ and $\rho$ were selected to match an outbreak that is an 
+epidemic. In the case where $S_0 \leq \rho$, the trajectory for the $I$ 
+compartment would never be increasing. The increase to a peak and subsequent 
+decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is 
+reasonable to model ILI by this compartment. Thus in modeling the ILI data, 
+we consider the data to be analogous to the $I$ proportion of the population.
 
@@
     \includegraphics[scale=.6]{Images/sir_traj.png}
-    \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
-The three compartments are the susceptible compartment (left), infectious (center), and recovered (right).
-In this example, $S_0/\rho > 1$.}
+    \caption{Susceptible-infectious-recovered (SIR) model separated by 
+    compartments. The three compartments are the susceptible compartment 
+    (left), infectious (center), and recovered (right). In this 
+    example, $S_0/\rho > 1$.}
     \label{fig:sir_traj}
@@
 \subsection{Asymmetric Gaussian (ASG) function} \label{sec:asg_func}
-The ASG function is another example of a nonlinear function which can approximate the trajectory of the flu outbreak.
-The ASG was previously used by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used to model vegetation growth and satellite sensor data \cite[]{lewis2020extracting, jonsson2002seasonality, hird2009noise, beck2006improved, atkinson2012inter}.
-The ASG is a modification of the asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized by its rise to a peak and fall from that peak which may not occur at the same rate, as shown in figure \ref{fig:asg_function}.
-The ASG function is denoted as $ASG_\theta(w)$ where $\theta = (\lambda, \nu, \mu, \sigma_1^2, \sigma_2^2)$, $\nu > 0$, $\lambda > 0$, $\mu \in (-\infty, \infty)$, $\sigma_1, \sigma_2 > 0$ and $w \in (1, ..., W)$ is week.
-The function is defined in (\ref{eq:asg_function}).
+The ASG function is another example of a nonlinear function which can 
+approximate the trajectory of the flu outbreak. The ASG was previously used 
+by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used 
+to model vegetation growth and satellite sensor data 
+\cite[]{lewis2020extracting, jonsson2002seasonality, hird2009noise, 
+beck2006improved, atkinson2012inter}. The ASG is a modification of the 
+asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized 
+by its rise to a peak and fall from that peak which may not occur at the same 
+rate, as shown in figure \ref{fig:asg_function}. The ASG function is denoted 
+as $ASG_\theta(w)$ where 
+$\theta = (\lambda, \nu, \mu, \sigma_1^2, \sigma_2^2)$, $\nu > 0$, 
+$\lambda > 0$, $\mu \in (-\infty, \infty)$, $\sigma_1, \sigma_2 > 0$ and 
+$w \in (1, ..., W)$ is week. The function is defined in 
+(\ref{eq:asg_function}).
 
@@
     \includegraphics[scale=.45]{Images/asg_function.png}
-    \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
+    \caption{Example plot of asymmetric Gaussian (ASG) function showing the 
+    shape of the function in relation to the parameters 
+    $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
     \label{fig:asg_function}
@@
 \end{equation}
-For modeling in this manuscript, we use a slightly reparameterized version of the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$.
+For modeling in this manuscript, we use a slightly reparameterized version of 
+the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$. 
 This constraint guarantees that the function has a peak greater than $\lambda$.
@@
 
-The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as Osthsu et al.
-\cite[]{osthus2019dynamic} point out there may be systematic behavior in the ILI that these or other possible functions may not capture.
-As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a regular peak at week 22 of the flu season.
-Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to illustrate the systematic discrepancy from a fitted function.
-Figure \ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 to 2022 excluding 2020 with a best fit ASG function plotted over the ILI.
-The fits for each season were made by obtaining the maximum likelihood estimate (MLE) of a model given the ILI data where we assume the ASG function is the mean parameter of a Beta distributed random variable.
-Figure \ref{fig:discrepancy} shows the discrepancy between the model fit and the data for the same seasons.
-The grey lines show the difference between the data and the functions from figure \ref{fig:asg_fits} for each season, and the black line is the average by week over all seasons.
-The lines show that the ASG function typically underpredicts week 22 and overpredicts week 23.
-Perhaps for other weeks, like week 30 for example, there also tends to be systematic behavior not captured by the ASG function.
+The SIR and ASG functions are useful for capturing the main trend of the ILI 
+data, but as Osthsu et al. \cite[]{osthus2019dynamic} point out there may be 
+systematic behavior in the ILI that these or other possible functions may not 
+capture. As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a 
+regular peak at week 22 of the flu season. 
+Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to 
+illustrate the systematic discrepancy from a fitted function. Figure 
+\ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 
+to 2022 excluding 2020 with a best fit ASG function plotted over the ILI. The 
+fits for each season were made by obtaining the maximum likelihood estimate 
+(MLE) of a model given the ILI data where we assume the ASG function is the 
+mean parameter of a Beta distributed random variable. Figure 
+\ref{fig:discrepancy} shows the discrepancy between the model fit and the data 
+for the same seasons. The grey lines show the difference between the data and 
+the functions from figure \ref{fig:asg_fits} for each season, and the black 
+line is the average by week over all seasons. The lines show that the ASG 
+function typically underpredicts week 22 and overpredicts week 23. Perhaps 
+for other weeks, like week 30 for example, there also tends to be systematic 
+behavior not captured by the ASG function.
 
@@
     \includegraphics[scale=.5]{Images/asg_fits.png}
-    \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
+    \caption{Observed US national influenza-like illness (ILI) percentage for 
+    seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an 
+    asymmetric Gaussian (ASG) model for the ILI data (black)}
     \label{fig:asg_fits}
@@
     \includegraphics[scale=.45]{Images/discrepancy.png}
-    \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
+    \caption{Difference between observed US national influenza-like illness 
+    (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 
+    2010 to 2022 excluding 2020 (grey) and the average difference of all 
+    seasons (black)}
     \label{fig:discrepancy}
@@
 
-The term $\gamma_w$, where $w$ is the season week, is included in (\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the function.
-Modeling discrepancy has been used in uncertainty analysis of simulators to capture systematic differences between mathematical models and
-reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,arendt2012improving,kennedy2001bayesian}.
-Modeling discrepancy can lead to overfitting, particularly in forecasting scenarios, and may also lead to identifiability issues.
-Thus, care must be taken in setting parameter constraints as well as in the selection of prior distributions \cite[]{osthus2019dynamic,brynjarsdottir2014learning}. 
-Modeling of the discrepancy for ILI was done by Osthus et al.
-during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}.
-As in Osthus's model, $\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}).
+The term $\gamma_w$, where $w$ is the season week, is included in 
+(\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the 
+function. Modeling discrepancy has been used in uncertainty analysis of 
+simulators to capture systematic differences between mathematical models and 
+reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,
+arendt2012improving,kennedy2001bayesian}. Modeling discrepancy can lead to 
+overfitting, particularly in forecasting scenarios, and may also lead to 
+identifiability issues. 
+Thus, care must be taken in setting parameter constraints as well as in the 
+selection of prior distributions \cite[]{osthus2019dynamic,
+brynjarsdottir2014learning}. 
+Modeling of the discrepancy for ILI was done by Osthus et al. during the 2015 
+and 2016 flu seasons where their model outperformed all others in the CDC flu 
+forecasting challenge \cite[]{osthus2019dynamic}. As in Osthus's model, 
+$\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}). 
 \begin{equation}
@@
 \end{equation}
-The idea for using the reverse random walk is that there are several previous seasons of ILI data, and assuming the random walk captures systematic behavior, fitting it hierarchically over seasons can assist in predicting future behavior in the current season.
-Reverse random walks have also been used with success in election forecasting and other flu forecasting models \cite[]{osthus2021multiscale, osthus2019dynamic, linzer2013dynamic}.
-The sum to zero constraint $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve identifiability.
-
-\subsection{ILI forecasts}
-Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating.
-Future ILI forecasts are obtained via the posterior predictive distribution where for week $w$, the predictive distribution is obtained by integrating over the parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ as in (\ref{eq:ili_post}) where $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ is the density function of the posterior distribution for the model parameters.
-If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ where $i$ is a positive integer.
+The idea for using the reverse random walk is that there are several previous 
+seasons of ILI data, and assuming the random walk captures systematic behavior, 
+fitting it hierarchically over seasons can assist in predicting future behavior 
+in the current season. 
+Reverse random walks have also been used with success in election forecasting 
+and other flu forecasting models \cite[]{osthus2021multiscale, 
+osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint 
+$-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
+identifiability.
+
+\subsection{Hospitalization model} \label{sec:hospital_model}
+
+The second component for forecast modeling is based on the relationship between 
+hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
+This is an example of an autoregressive model with exogenous variables where 
+the autoregressive lag is one (ARX(1)) 
+\cite[]{raftery2010online,ljung1987system}.
+
+\begin{equation}
+    \begin{aligned}
+    \label{eq:final_hosp}
+    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} 
+    (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
+    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, 
+    \omega_s) %N(0, \sigma_{\epsilon}^2)
+    \end{aligned}
+\end{equation}
+
+Figure \ref{fig:lag_scatter} helps
+illustrate why this modeling decision was made.
+shows scatterplots of ILI percentage and hospitalizations on the top row.
+The bottom row shows scatterplots of the difference between hospitalizations and 
+scaled 1 week 
+lags by ILI percentage at the US national level. The plots in the right column
+are from the logarithm of hospitalizations. The lines through the points is
+drawn by connecting hospitalizations for consecutive weeks during the flu 
+season.
+These plots show that a linear or quadratic relationship between ILI percentage 
+and hospitalizations does not capture much of the temporal correlation in 
+hospitalizations as shown by the obvious loops drawn in the plot. 
+However, when including a 
+single lag of hospitalizations in the model, much of the time correlation is
+accounted for as shown by the tighter and more frequent crossing of lines shown
+in the bottom row. Higher than order 1 autoregressions showed little or no 
+improvement over order 1, hence we consider the ARX(1) reasonably well 
+suited for modeling ILI and hospitalzations.
+
+\begin{figure}[hbt!]
+    \includegraphics[scale=.2]{Images/ili_v_svat_vs_lagdiff.png}
+    \caption{Scatterplots of national hospitalizations (left) or log 
+    hospitalizations (right) lag differences by ILI \% where the lag is 
+    scaled by $\phi$. Points are colored by season. For hospitalizations 
+    $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$. }
+    \label{fig:lag_scatter}
+\end{figure}
+
+In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for 
+week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed 
+according to some distribution $D_s$ with mean parameter 0, scale parameter 
+$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees 
+of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
+
+This model is for any location, and for fitting purposes $ILI_{s,w}$ is always 
+multiplied by $P$ which is proportional to the population of the state or 
+territory, in this case the total population divided by 50,000. This is done 
+as a means of scaling so that the prior distribution assigned to  
+$\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, 
+$\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all 
+states.
+
+Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also 
+fit via Bayesian posterior updating. To obtain forecasts for $H_{s, w^* + i}$, 
+the ILI posterior predictive distribution is used along with the posterior 
+distribution for the parameters in (\ref{eq:final_hosp}). We considered three 
+scenarios for model (\ref{eq:final_hosp}). A model where $D_s$ belongs to the 
+normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ 
+is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, 
+or $H_{s,w} + 1$ is lognormally distributed (LNORM). The population value $P$ 
+is excluded from the LNORM model, and we set 
+$\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting. In 
+the LNORM model, if the linear parameters are not set the same for seasons 22 
+and 23, the final variance was more prone to be extreme. Besides varying the 
+distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ 
+or there is no quadratic ILI term.
+
+\subsection{Prior selection}
+
+The priors selected for the ILI data model under both the SIR and ASG models 
+largely follow the prior selections in Osthus et al. \cite[]{osthus2019dynamic} 
+and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved 
+numerical stability and/or we felt the adjusted prior made more sense for the 
+problem. For model (\ref{eq:ili_model}), parameters which are common even when 
+using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
+$\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$. For the SIR function 
+$\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
+function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
+For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
+estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
+
+The priors assigned were mostly noninformative, though in certain cases the 
+prior distributions were selected for numerical stability as was the case 
+for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$. For these two scale 
+parameters only, rather than assigning a half-normal prior to the standard 
+deviation parameters, as recommended by Gelman \cite[]{gelman2006prior}, the 
+priors were assigned to the variance parameters. Univariate parameters were 
+assigned either a normal distribution prior if the support is on $\mathbb{R}$, 
+a half-normal prior if the support is nonnegative, or a truncated-normal 
+prior to match a more specific support. Under the ASG model, $\theta_s$ is 
+modeled hierarchically over seasons so that for each season the transformed 
+parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are 
+assigned to $\theta$ and $\Sigma$.
+
+Additional prior constraints were made to improve parameter identifiability. 
+In Osthus et al. \cite[]{osthus2019dynamic} the initial value of the 
+susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
+The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
+priors. To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
+in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the 
+parameters. A modular Bayesian approach involves multiple steps of parameter 
+fitting where some parameters may be estimated without priors via maximum 
+likelihood estimation or other means. Fitting the rest of the model parameters 
+involves assigning priors and conditioning on the previously fit parameters. 
+This has been done in computer modeling to improve identifiability and other 
+issues, though Liu et al. \cite[]{liu2009modularization} warn this approach is 
+not probabilistically sound if parameter inference is a priority 
+\cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. 
+We carried out the modular fit by first estimating the maximum likelihood 
+estimate (MLE) for the parameter $\lambda_s$ for each season in 
+(\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
+
+\subsection{Hospitalization Forecasts Using the Bayesian Cut}
+
+Rather than combining the ILI model in equation (\ref{eq:ili_model}) with the 
+hospitalization model in equation (\ref{eq:final_hosp}) and computing a fully Bayesian
+posterior distribution, both models were fit separately and 1 to 4 week ahead
+ILI forecasts from model (\ref{eq:ili_model}) were combined with the posterior
+predictive distribution of model (\ref{eq:final_hosp}) to obtain 1 to 4 week
+ahead hospitalization forecasts. In this scheme, the posterior distribution of 
+model (\ref{eq:ili_model}) parameters was not directly influenced by the 
+hospitalization data and the posterior distribution of model 
+(\ref{eq:final_hosp}) parameters was not directly influenced by model 
+(\ref{eq:ili_model}) parameters. Hence there was limited feedback between the
+two components, but propagation of uncertainty in ILI forecasts was still 
+permitted in the estimating of hospitalization forecasts.
+
+This two-component modeling is remeniscent of the "Bayesian cut" as 
+described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
+into two component models or "modules" may be justfied for several reasons
+including to reduce time to fit the model, avoiding mixing issues of posterior
+sampling distributions, improving predictive performance or to prevent
+model misspecification in one model to influence the other 
+\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, plummer2015cuts}.
+
+The decision to cut the model into two components was made for several reasons.
+When fitting the fully Bayesian joint model, the time to fit a single model using
+Markov chain Monte Carlo (MCMC) sampling and
+obtain forecasts was sometimes upwards of 4 hours whereas fitting the two models
+separately usually took less than 1 hour. When assessing convergence of model
+parameters, it was rare that posterior sampling chains for the joint model
+showed convergence even after 55,000 iterations, whereas the component model 
+fits gave no signs to be concerened with convergence. And most importantly,
+the forecasts from the component models generally outperformed those of the 
+joing model in terms of minimizing proper scoring rules. 
+The process for obtaining forecasts is described in the following paragraphs 
+and computation is described in section \ref{sec:implementation_posterior}.
+
+Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating. Future ILI 
+forecasts are obtained via the posterior predictive distribution where for
+week $w$, the predictive distribution is obtained by integrating over the 
+parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
+as in (\ref{eq:ili_post}) where 
+$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ 
+is the density function of the posterior distribution for the model parameters. 
+If the current week is $w$ then the desired forecasts are for weeks $w + i$ 
+where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
+predicted or forecast ILI at $i$ weeks in the future.
 
@@
     \label{eq:ili_post}
-    p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) = \int \int \int \int p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+    p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}) = \int \int \int \int 
+    p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
+    \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
+    \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) 
+    d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
 \end{equation}
@@
 % \spencer{Isn't that captured in $d \pi_s$?}
-
-\subsection{Hospitalization model} \label{sec:hospital_model}
-
-The second component for forecast modeling is based on the relationship between hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
-This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}. Figure \ref{fig:lag_scatter}
-shows scatterplots of the difference between hospitalizations and scaled 1 week 
-lags by ILI percentage at the US national level. The points in the left plot are
-from hospitalization differences whereas in the right the hospitalizations were
-log transformed. The positive and slightly curved relationship suggests model
-(\ref{eq:final_hosp}) is a reasonable choice.
-
-\begin{figure}[hbt!]
-    \includegraphics[scale=.46]{Images/lagdiff_vs_ili.png}
-    \caption{Scatterplots of national hospitalizations (left) or log hospitalizations (right) lag differences by ILI \% where the lag is scaled by $\phi$.
-Points are colored by season.
-For hospitalizations $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$.
-}
-    \label{fig:lag_scatter}
-\end{figure}
+Hospitalization forecasts $\widetilde{H}_{s,w + i}$ at week $w + i$ are then
+obtained by combining the posterior predictive distribution of model 
+(\ref{eq:final_hosp}) with (\ref{eq:ili_post}) and integrating over 
+$\widetilde{ILI}_{s,w + i}$ as in (\ref{eq:hosp_post_forc}). Here 
+$\boldsymbol{\alpha}$ represents all regression and variance parameters in
+model (\ref{eq:final_hosp}).
 
@@
 \begin{aligned}
-    \label{eq:final_hosp}
-    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
-    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, \omega_s) %N(0, \sigma_{\epsilon}^2)
+\label{eq:hosp_post_forc}
+    p(\tilde{H}_{s,w + i} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) &= 
+       \int p(\tilde{H}_{s,w + i} | \textbf{H}, \textbf{ILI}, 
+       \widetilde{ILI}_{s,w + i}, \boldsymbol{\alpha}) 
+       p(\widetilde{ILI}_{s,w + i} | \textbf{ILI}) d\widetilde{ILI}_{s,w + i}
 \end{aligned}
 \end{equation}
-In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
-
-This model is for any location, and for fitting purposes $ILI_{s,w}$ is always multiplied by $P$ which is proportional to the population of the state or territory, in this case the total population divided by 50,000.
-This is done as a means of scaling so that the prior distribution assigned to  $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, $\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all states.
-
-Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also fit via Bayesian posterior updating.
-To obtain forecasts for $H_{s, w^* + i}$, the ILI posterior predictive distribution is used along with the posterior distribution for the parameters in (\ref{eq:final_hosp}).
-We considered three scenarios for model (\ref{eq:final_hosp}).
-A model where $D_s$ belongs to the normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, or $H_{s,w} + 1$ is lognormally distributed (LNORM).
-The population value $P$ is excluded from the LNORM model, and we set $\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting.
-In the LNORM model, if the linear parameters are not set the same for seasons 22 and 23, the final variance was more prone to be extreme.
-Besides varying the distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ or there is no quadratic ILI term.
-
-\subsection{Prior selection}
-
-The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in Osthus et al.
-\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem.
-For model (\ref{eq:ili_model}), parameters which are common even when using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, $\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$.
-For the SIR function $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$.
-For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
-
-The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by Gelman \cite[]{gelman2006prior}, the priors were assigned to the variance parameters.
-Univariate parameters were assigned either a normal distribution prior if the support is on $\mathbb{R}$, a half-normal prior if the support is nonnegative, or a truncated-normal prior to match a more specific support.
-Under the ASG model, $\theta_s$ is modeled hierarchically over seasons so that for each season the transformed parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are assigned to $\theta$ and $\Sigma$.
-
-Additional prior constraints were made to improve parameter identifiability.
-In Osthus et al.
-\cite[]{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$.
-The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative priors.
-To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the parameters.
-A modular Bayesian approach involves multiple steps of parameter fitting where some parameters may be estimated without priors via maximum likelihood estimation or other means.
-Fitting the rest of the model parameters involves assigning priors and conditioning on the previously fit parameters.
-This has been done in computer modeling to improve identifiability and other issues, though Liu et al.
-\cite[]{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}.
-We carried out the modular fit by first estimating the maximum likelihood estimate (MLE) for the parameter $\lambda_s$ for each season in (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
-
-% \spencer{Blue stuff below will be either included further on or will be removed entirely
 
@@
 
-The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the Stan Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}.
-Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}.
-The \texttt{cmdstanr} package provides several diagnostic statistics for assessing the sampler.
-As mentioned, most model parameter prior distributions were intended to be uninformative.
-Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material \cite[]{wadsworth2024bas}.
-
-We assessed the model fit for four ILI models.
-These included the SIR and ASG models and models with and without discrepancy modeling.
-When discrepancy is included, the models are denoted as SIRD and ASGD.
-These models were fit using US national data from 2010 to 2023 flu seasons, where data from the 2020 season was excluded because of the unique behavior during that season.
+The models were fit via MCMC sampling using the 
+\texttt{cmdstanr} package which was developed and is maintained by the Stan 
+Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
+implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn 
+sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
+several diagnostic statistics for assessing the sampler.
+As mentioned, most model parameter prior distributions were intended to be 
+uninformative.
+Plots of posterior distributions for select parameters from ILI and 
+hospitalization models are shown in the supplementary material 
+\cite[]{wadsworth2024bas}.
+
+We assessed the model fit for four ILI models. These included the SIR and 
+ASG models and models with and without discrepancy modeling. When discrepancy 
+is included, the models are denoted as SIRD and ASGD. These models were fit 
+using US national data from 2010 to 2023 flu seasons, where data from the 
+2020 season was excluded because of the unique behavior during that season. 
 Assessment for hospitalization modeling was done for six different models. 
-These include NORM, LNORM, and LST models, and models where a quadratic ILI term is included or excluded.
-To assess posterior sampling convergence, models were fit to data where ILI and hospitalization data up to week 14 of the 2023 season was included.
-Sampling was done with four chains where from each chain 60,000 posterior draws were sampled, and the first 10,000 draws were discarded as a burn-in.
-The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for each parameter.
-Tables \ref{tab:ili_diagnostics} and \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum $ESS$ over all parameters for ILI and hospitalization models respectively.
-Forecast models for all other weeks of the season were fit using one chain of 60,000 draws where the first 10,000 draws were discarded as a burn-in.
+These include NORM, LNORM, and LST models, and models where a quadratic ILI 
+term is included or excluded. To assess posterior sampling convergence, models 
+were fit to data where ILI and hospitalization data up to week 14 of the 2023 
+season was included. Sampling was done with four chains where from each chain 
+60,000 posterior draws were sampled, and the first 10,000 draws were discarded 
+as a burn-in. The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the 
+effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for 
+each parameter. Tables \ref{tab:ili_diagnostics} and 
+\ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum 
+$ESS$ over all parameters for ILI and hospitalization models respectively. 
+Forecast models for all other weeks of the season were fit using one chain of 
+60,000 draws where the first 10,000 draws were discarded as a burn-in. 
 % \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
-For parameters of the ASG models that were prone to cause trouble in posterior sampling, the starting values were set to be the MLEs.
+For parameters of the ASG models that were prone to cause trouble in posterior 
+sampling, the starting values were set to be the MLEs.
 
 \begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI models fitted on US data for week 14 of the 2023 season}
+\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI 
+models fitted on US data for week 14 of the 2023 season}
 \label{tab:ili_diagnostics}
@@
 \begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six hospitalization models fitted on US data for week 14 of the 2023 season}
+\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six 
+hospitalization models fitted on US data for week 14 of the 2023 season}
 \label{tab:hsop_diagnostics}
@@
      \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$\\
+     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & 
+     $< 1.001$\\
      $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
@@
 
-To obtain forecast distributions of hospitalizations, draws from the posterior predictive distribution from the ILI model were used in conjunction with the posterior distribution of the hospitalizations model.
-When fitting model (\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were saved.
-Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal distributions for the model parameters were saved.
-To obtain forecast distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following steps are repeated $K$ times where $K$ is an integer for the number of desired samples.
-We set $K = 50,000$.
+To obtain forecast distributions of hospitalizations, draws from the posterior 
+predictive distribution from the ILI model were used in conjunction with the 
+posterior distribution of the hospitalizations model. When fitting model 
+(\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were 
+saved. Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal 
+distributions for the model parameters were saved. To obtain forecast 
+distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following 
+steps are repeated $K$ times where $K$ is an integer for the number of desired 
+samples. We set $K = 50,000$.
 
@@
   \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
-  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior distributions
+  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, 
+  $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior 
+  distributions
   \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
@@
 \end{quote}
-The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast for hospitalizations at week $w + i$.
-For the forecast competition analysis in section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were set to 0 to reflect realistic values of hospitalizations and comply with the FluSight forecasting rules.
+The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
+for hospitalizations at week $w + i$. For the forecast competition analysis in 
+section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
+set to 0 to reflect realistic values of hospitalizations and comply with the 
+FluSight forecasting rules.
 
 \section{Simulation Study} \label{sec:simulation2}
-In this section, we present a simulation study conducted for comparing ILI models and further assessing the hospitalization forecast model. US ILI data is used, and hospitalization data is simulated. 
-A leave-one-season-out (LOSO) approach was combined with a Monte Carlo simulation approach.
-For each replication, we simulated log-hospitalizations for all weeks during seasons 2010, ..., 2022, excluding 2020, using the existing ILI data as a predictive covariate.
-Each season was in turn ``left-out" and treated as if it was the most recent season which we desired to forecast.
-Fitting and forecasting was then done for weeks 14, 20, 26, 32, and 38 of the left out season, giving two weeks that tend to occur as flu cases increase, two as cases decrease, and one that occurs when the cases may be increasing or decreasing.
-Week 20 is a week leading up to the holiday week 22 where ILI typically has a local peak.
-We were particularly interested in how important modeling discrepancy is for forecasting at week 20.
-
-For the simulation of hospitalizations, the parameters $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and $\sigma^2_{\epsilon_s}$ from the hospitalization model in (\ref{eq:final_hosp}) were considered the same across all seasons so that all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$.
-The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization data from the 2022 season.
-For fitting, the hospitalization data was first log-transformed.
-We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
-and assigned the noninformative prior $p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$.
-The marginal posterior distribution $\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ was then the established posterior multivariate normal distribution and $\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior distribution \cite[]{gelman2013bayesian}.
-The posterior means of those parameters were used as the values from which log-hospitalizations were simulated.
-The number of replicates in the simulation was 500.
-
-Model comparison was done by calculating the continuous ranked probability score (CRPS) for each forecast.
-The CRPS is a proper scoring rule which evaluates the forecast distribution function.
-Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}.
-Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The CRPS is defined in (\ref{eq:crps2}).
-Here $F(\cdot)$ is the forecast distribution function, and $y^*$ is the observed targeted value of the forecast.
-The orientation of the CRPS is negative, meaning the smaller the score the better.
-Here $\vmathbb{1}\{\cdot\}$ is the indicator function.
-The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
-
-\begin{equation}
-    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
-    \label{eq:crps2}
-\end{equation}
+In this section, we present a simulation study conducted for comparing ILI 
+models and further assessing the hospitalization forecast model. US ILI data 
+is used, and hospitalization data is simulated. 
+A leave-one-season-out (LOSO) approach was combined with a Monte Carlo 
+simulation approach. For each replication, we simulated log-hospitalizations 
+for all weeks during seasons 2010, ..., 2022, excluding 2020, using the 
+existing ILI data as a predictive covariate. Each season was in turn 
+``left-out" and treated as if it was the most recent season which we desired 
+to forecast. Fitting and forecasting was then done for weeks 14, 20, 26, 32, 
+and 38 of the left out season, giving two weeks that tend to occur as flu 
+cases increase, two as cases decrease, and one that occurs when the cases may 
+be increasing or decreasing. Week 20 is a week leading up to the holiday week 
+22 where ILI typically has a local peak. We were particularly interested in 
+how important modeling discrepancy is for forecasting at week 20.
+
+Besides the 4 ILI models, we also included forecasts for a naive baseline
+(BASE) model and an autoregressive integrated moving average (ARIMA) model. 
+The BASE model is a simple random walk with drift model, and the ARIMA model
+is fit by selecting the best possible model give the data according to a 
+fitting criterion. Both the BASE and ARIMA ILI models were fit using the 
+\texttt{forecast} package in \texttt{R} \cite{hyndman2008forecast}. Samples
+were then drawn from the forecast distributions, and these were then used
+along with the model in (\ref{eq:final_hosp}) to make hospitalization 
+forecast distributions.
+
+For the simulation of hospitalizations, the parameters 
+$\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and 
+$\sigma^2_{\epsilon_s}$ from the hospitalization model in 
+(\ref{eq:final_hosp}) were considered the same across all seasons so that 
+all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$. 
+The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were 
+estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization 
+data from the 2022 season. For fitting, the hospitalization data was first 
+log-transformed. We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
+and assigned the noninformative prior 
+$p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$. 
+The marginal posterior distribution 
+$\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ 
+was then the established posterior multivariate normal distribution and 
+$\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior 
+distribution \cite[]{gelman2013bayesian}. The posterior means of those 
+parameters were used as the values from which log-hospitalizations were 
+simulated. The number of replicates in the simulation was 500.
+
+Model comparison was done by calculating a proper scoring rule for each 
+forecast.
+Proper scoring rules are the 
+current standard for comparing performance between probabilistic forecasts 
+and selecting the best forecasts according to the notion of maximizing 
+sharpness subject to (auto-)calibration 
+\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
+rules are commonly used in forecast comparison and have the property that a 
+forecaster is incentivized to be honest in the reporting of their forecasts 
+\cite[]{gneiting2007strictly, gneiting2014probabilistic}.
+The proper scoring rule primarily used in the FluSight competition and the 
+COVID-19 Forecast Hub and which we will use herein is the 
+weighted interval score (WIS)
+\cite[]{mathis2024evaluation, bracher2021evaluating}.
+The WIS is used for scoring quantile or interval 
+forecasts 
+\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} 
+and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
+included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
+value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
+weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
+interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
+single interval as defined in (\ref{eq:is}). Here 
+$\vmathbb{1}\{\cdot\}$ is the indicator function.
+The goal of the forecaster is to 
+minimize the WIS.
+
+\begin{equation}
+\label{eq:wis2}
+        WIS_{0,B}(Q, y^*) = \frac{1}{B + 1/2} \times (w_0\times |y^* - median| + \sum_{b=1}^B \{w_k \times IS_{\alpha_b}(Q, y^*) \} )
+\end{equation}
+
+\begin{equation}
+\label{eq:is}
+        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\vmathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \vmathbb{1}\{y^* > r\}
+\end{equation}
+
+%  The CRPS is defined 
+% in (\ref{eq:crps2}). Here $F(\cdot)$ is the forecast distribution function, 
+% and $y^*$ is the observed targeted value of the forecast. The orientation of 
+% the CRPS is negative, meaning the smaller the score the better. Here 
+% $\vmathbb{1}\{\cdot\}$ is the indicator function.
+% The CRPS is calculated using a quantile decomposition from 
+% \cite[]{laio2007verification}. 
+% 
+% \begin{equation}
+%     \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
+%     \label{eq:crps2}
+% \end{equation}
 
@@
 
-Figure \ref{fig:crps_by_week_horizon} shows 
-boxplots of the CRPS for the 500 simulation replicates faceted by week and horizon. 
+Figure \ref{fig:wis_by_week_horizon} shows 
+boxplots of the WIS for the 500 simulation replicates faceted by season week 
+on the y-axis and horizon on the x-axis. 
 % Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
-Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with lower medians and lower variation than the two models not accounting for ILI discrepancy.
-This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough.
-This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
-Among the four models, ASG tends to have the lowest CRPS values though this is not always the case.
-This suggests that the additional flexibility of the five parameter ASG function is often better for forecasting than that of the three parameter SIR function.
-Additional
-plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and another common proper scoring rule known as the logarithmic score.
+In most plots the ASGD model shows the best performance, 
+especially during week 20 where
+forecasts are near weeks 22 and 23 which show a seasonal peak and trough in 
+figure \ref{fig:discrepancy}. The SIRD also shows improved performance over the
+SIR model, though its performance does not match that of the ASG or ASGD.
+Here the ASGD model regularly outperforms
+the BASE and ARIMA models, but the three other nonlinear models do not. 
+The ASG model
+tends to outperform the SIR and SIRD models suggesting that the additional 
+flexibility of the five parameters ASG function 
+is often better for forecasting than that of the three parameter SIR function.
+But the best performance comes with the combination of the ASG function and
+the discrepancy modeling.
+
+% Notable plots here are for week 20 where the two models accounting for the 
+% ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with 
+% lower medians and lower variation than the two models not accounting for ILI 
+% discrepancy. This makes sense given the forecasts at week 20 are forecasting 
+% weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and 
+% trough. This confirms that there is value in modeling discrepancy at least 
+% during around the holiday weeks near week 22. Among the four models, ASG 
+% tends to have the lowest CRPS values though this is not always the case. 
+% This suggests that the additional flexibility of the five parameter ASG 
+% function is often better for forecasting than that of the three parameter SIR 
+% function. 
+% Additional
+% plots in the supplemtary material \cite[]{wadsworth2024bas} further show that 
+% the ASG models often but not always outperform the SIR models according to the 
+% CRPS and another common proper scoring rule known as the logarithmic score.
 
@@
     \centering
-    \includegraphics[scale=.4]{Images/crps_by_week_horizon.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
-    \label{fig:crps_by_week_horizon}
+    \includegraphics[scale=.4]{Images/wis_by_week_horizon.png}
+    \caption{Boxplots of weighted interval score (WIS) for the 
+    four Bayesian nonlinear ILI models, a baseline model (BASE) and
+    an autoregressive integrated moving average model (ARIMA) 
+    over all seasons in the simulation study faceted by horizon 
+    (x-axis) and season week (y-axis). Horizons include 1-4 week ahead forecasts 
+    and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
+    \label{fig:wis_by_week_horizon}
 \end{figure}
@@
 
-In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations.
-The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$.
-A single forecast is thus comprised of 11 predictive intervals and a median.
-Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, and forecasts were made at the state and national level.
-The first week of forecasting took place during the week of October 7, 2023, and the final week was the week of April 27, 2024 making 29 total weeks of forecasts.
- The same format was used during the 2021 and 2022 seasons and for the COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
+In this section we apply the forecast models to make forecasts of the 2023 
+flu season weekly hospitalizations. The scoring of the forecasts is in the 
+context of the FluSight competition where each competing forecast was 
+submitted as a set of 23 quantiles corresponding to the given probability 
+levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$. A 
+single forecast is thus comprised of 11 predictive intervals and a median.
+Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
+and forecasts were made at the state and national level. The first week of 
+forecasting took place during the week of October 7, 2023, and the final week 
+was the week of April 27, 2024 making 29 total weeks of forecasts.
+ The same format was used during the 2021 and 2022 seasons and for the 
+ COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
 % Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
- Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS).
-The focus in this section will be on the LWIS (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
-
-The WIS is a proper scoring rule used for scoring quantile or interval forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all included quantiles, $B$ is the number of intervals,  $y^*$ is the observed value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ interval.
-$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a single interval as defined in (\ref{eq:is}).
-The goal of the forecaster is to minimize the WIS.
-The LWIS is the same as the WIS except that it is evaluated over the log of quantiles and the log of the observed value.
-
-\begin{equation}
-\label{eq:wis2}
-        WIS_{0,B}(Q, y^*) = \frac{1}{B + 1/2} \times (w_0\times |y^* - median| + \sum_{b=1}^B \{w_k \times IS_{\alpha_b}(Q, y^*) \} )
-\end{equation}
-
-\begin{equation}
-\label{eq:is}
-        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\vmathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \vmathbb{1}\{y^* > r\}
-\end{equation}
-
-We fit 24 forecast models for each location for all 29 weeks, and for each week forecast 1-4 week ahead hospitalizations.
-The 24 models included all combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST hospitalization models, and both quadratic and linear hospitalization models.
+ Primary scores for evaluating each forecast were the WIS, 
+ the log-weighted interval score (LWIS), and the relative weighted 
+ interval score (RWIS). The focus in this section will be on the LWIS 
+ (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
+ The LWIS is the same as the WIS except that it is evaluated 
+over the log of quantiles and the log of the observed value.
+
+We fit 24 forecast models for each location for all 29 weeks, and for each 
+week forecast 1-4 week ahead hospitalizations. The 24 models included all 
+combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
+hospitalization models, and both quadratic and linear hospitalization models. 
 The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
-Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$.
-Here $\vmathbb{1}_{A}$ represents the indicator function for values within the set $A$.
+Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
+$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
+rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
+indicator function for values within the set $A$.
 
@@
 \end{equation}
-For the ASG model, the MLE $\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value.
+For the ASG model, the MLE 
+$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value.
 The remaining estimates were used as starting values for posterior sampling.
 The transformation $T(\theta_s) = (\text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ was made, and a prior distribution was assigned to $T(\theta_s)$.
-The prior distributions for the ILI model under the ASG function are shown in (\ref{eq:ili_prior}).
-These are slightly informative priors because for most parameters we have an idea what reasonable values may be.
-Here $m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
+The prior distributions for the ILI model under the ASG function are shown 
+in (\ref{eq:ili_prior}). These are slightly informative priors because for 
+most parameters we have an idea what reasonable values may be. Here 
+$m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
 where $\text{diag}(\cdot)$ 
@@
 \end{equation}
-Parameters shared by both the SIR and ASG models are the scale parameter $\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-These priors are in (\ref{eq:shared_ili_priors}).
+Parameters shared by both the SIR and ASG models are the scale parameter 
+$\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and 
+$\sigma_{\gamma_W}^2$. These priors are in (\ref{eq:shared_ili_priors}).
 
@@
 \end{equation}
-Because of the limited information for estimating $\sigma_{\gamma_W}^2$, we selected an informative prior distribution by first estimating  $\hat{\sigma}_{\gamma_W}^2$.
-This was estimated for a given state by first calculating the MLE for $\theta_s$.
+Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
+we selected an informative prior distribution by first estimating  
+$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
+calculating the MLE for $\theta_s$. 
 % We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
 %     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
-$\widehat{ILI}_{s,w}$ was then predicted such that $\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season.
-Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over seasons of the differences $\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$.
-When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was calculated using the \texttt{mle2} function in the \texttt{bblme} package \cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} package \cite[]{soetaert2010desolve} function in \texttt{R}.
-Where the ASG function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} function.
+$\widehat{ILI}_{s,w}$ was then predicted such that 
+$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
+Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
+seasons of the differences 
+$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
+When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
+calculated using the \texttt{mle2} function in the \texttt{bblme} package 
+\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
+package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
+function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
+function.
 The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
@@
 \centering
-% \begin{subfigure}
+\begin{subfigure}
   \centering
@@
   % \label{fig:sub1}
-% \end{subfigure}%
-% \begin{subfigure}
+\end{subfigure}%
+\begin{subfigure}
   \centering
@@
   % \label{fig:sub2}
-% \end{subfigure}
-\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the hospitalization models are all normally distribution.
-The figure includes hospitalization forecasts where ILI is a linear predictor (left) and where ILI is a quadratic predictor (right).
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
+\end{subfigure}
+\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
+season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, 
+and the hospitalization models are all normally distribution. The figure 
+includes hospitalization forecasts where ILI is a linear predictor (left) 
+and where ILI is a quadratic predictor (right). 50\% predictive intervals 
+are pink and 95\% predictive intervals are red.}
 \label{fig:normal_flu_forecasts}
@@
 
-Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US flu hospitalizations during the 2023 season under the NORM hospitalization model.
-The forecasts shown are for weeks 14, 20, 26, and 32.
-The predictive bands are the 50\% and 95\% predictive intervals.
-These plots show a tendency to often underpredict hospitalizations.
-The quadratic models appear to do a better job predicting  hospitalizations at the season peak but a poorer job predicting after the peak, whereas the linear models seem to predict well or slightly overpredict after the peak.
+Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US 
+flu hospitalizations during the 2023 season under the NORM hospitalization 
+model. The forecasts shown are for weeks 14, 20, 26, and 32. The predictive 
+bands are the 50\% and 95\% predictive intervals.
+These plots show a tendency to often underpredict hospitalizations. The 
+quadratic models appear to do a better job predicting  hospitalizations at 
+the season peak but a poorer job predicting after the peak, whereas the linear 
+models seem to predict well or slightly overpredict after the peak. 
 % From these figures, it's difficult to say whether ASG or SIR models predict better, but at least for the quadractic models the week 22 predictions are maybe slightly better than for the linear models.
 
-Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the season for all 24 models of US hospitalizations.
-LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling.
-Here, the smaller the LWIS the better.
-The weeks around the holiday week 22 are highlighted by a grey band.
-In several cases, there appears to be a turning point in performance at or near week 22.
-For most ASG models, the models which include discrepancy tend to outperform those which do not beginning near week 22 and for the rest of the season.
-For the SIR models, the discrepancy does not always lead to better forecasts, but in for the LNORM hospitalization models, including discrepancy greatly improves forecasts around the holiday week 22.
-The LNORM LWIS scores tend to be higher than for LST and NORM hospitalization models, but figure \ref{fig:lwis_by_dist_loc} and table \ref{tab:forecast_scores} suggest that the LNORM hospitalization model outperforms LST and NORM for most states and territories.
+Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the 
+season for all 24 models of US hospitalizations. LWIS scores are grouped by 
+ILI model, hospitalization model distribution, and by the linear or quadratic 
+modeling. Here, the smaller the LWIS the better. The weeks around the holiday 
+week 22 are highlighted by a grey band. 
+In several cases, there appears to be a turning point in performance at or 
+near week 22. For most ASG models, the models which include discrepancy tend 
+to outperform those which do not beginning near week 22 and for the rest of 
+the season.
+For the SIR models, the discrepancy does not always lead to better forecasts, 
+but in for the LNORM hospitalization models, including discrepancy greatly 
+improves forecasts around the holiday week 22. The LNORM LWIS scores tend to 
+be higher than for LST and NORM hospitalization models, but figure 
+\ref{fig:lwis_by_dist_loc} and table \ref{tab:forecast_scores} suggest that 
+the LNORM hospitalization model outperforms LST and NORM for most states and 
+territories.
 
@@
     \includegraphics[scale = .6]{Images/lwis_ln_us_full_season.png}
-    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including and exluding discrepancy in the ILI model.
-Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
-Scores for models with an ASG ILI model are above while those with an SIR model are below.
-The lower the LWIS the better the forecast.}
+    \caption{Each plot shows the log weighted interval score (LWIS) for every 
+    week of the 2023 flu season with scores for models including and exluding 
+    discrepancy in the ILI model. Scores are separated by hospitalization 
+    distribution family and by ILI as a linear or quadratic predictor. Scores 
+    for models with an ASG ILI model are above while those with an SIR model 
+    are below. The lower the LWIS the better the forecast.}
     \label{fig:us_lwis}
@@
     \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
-Scores are averaged over all horizons 1-4 weeks ahead.
-Scores are faceted by hospitalization model distribution.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
+    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 
+    US states, PR, DC, and national level forecasts at each week during the 
+    2023 flu season. Scores are averaged over all horizons 1-4 weeks ahead. 
+    Scores are faceted by hospitalization model distribution. The lighter the 
+    shade, the lower the LWIS with low LWIS being better.}
     \label{fig:lwis_by_dist_loc}
@@
 
-Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by LWIS, but all 53 locations are included.
-The LWIS is shown across the whole season faceted by hospitalization model distribution.
-Overall, the weeks leading up to week 22 are the most difficult to forecast, but the LNORM model appears to perform the better than the other models during the weeks around the holiday week 22.
-The overall, calculated as the mean LWIS over all locations and weeks, for all 24 models is shown in table \ref{tab:forecast_scores}.
-The first main takeaway is that the top four performing models are SIR ILI models with the LNORM hospitalization models.
-The next main takeaway is that overall the models which include discrepancy outperformed the models without discrepancy.
+Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by 
+LWIS, but all 53 locations are included. The LWIS is shown across the whole 
+season faceted by hospitalization model distribution. Overall, the weeks 
+leading up to week 22 are the most difficult to forecast, but the LNORM model 
+appears to perform the better than the other models during the weeks around 
+the holiday week 22. 
+The overall, calculated as the mean LWIS over all locations and weeks, for all 
+24 models is shown in table \ref{tab:forecast_scores}. The first main takeaway 
+is that the top four performing models are SIR ILI models with the LNORM 
+hospitalization models. The next main takeaway is that overall the models 
+which include discrepancy outperformed the models without discrepancy.
 
 \begin{table}
-\caption{Overall scores for each of the 24 forecast models.
-The overall score is the log weighted interval score (LWIS) averaged over all locations, weeks, and horizons.
-The scores in the first two rows are for linear models, and the scores in the third and fourth rows are for quadratic models.
-The lowest WISs are bolded.}
+\caption{Overall scores for each of the 24 forecast models. The overall score 
+is the log weighted interval score (LWIS) averaged over all locations, weeks, 
+and horizons. The scores in the first two rows are for linear models, and the 
+scores in the third and fourth rows are for quadratic models. The lowest WISs 
+are bolded.}
 \begin{tabular*}{\textwidth}
@@
 
-It should not be assumed that these two takeaways will apply for future flu seasons.
-As suggested by the simulation study in the previous section, it may often be the case that the ASG ILI model is better for forecasting.
-Indeed, a close examination of figures \ref{fig:us_lwis} and \ref{fig:lwis_by_dist_loc} shows that it is common for the ASG models to show better forecasting skill than the SIR models.
+It should not be assumed that these two takeaways will apply for future flu 
+seasons. As suggested by the simulation study in the previous section, it may 
+often be the case that the ASG ILI model is better for forecasting. Indeed, a 
+close examination of figures \ref{fig:us_lwis} and \ref{fig:lwis_by_dist_loc} 
+shows that it is common for the ASG models to show better forecasting skill 
+than the SIR models.
 
@@
 
- In this manuscript we introduce a statistical modeling framework which allows for the incorporation of several ILI forecast modeling methods.
-Specifically, we built upon Osthus et al.
-\cite[]{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy.
-We model flu hospitalizations by incorporating the ILI forecast model into a model forecasting hospitalizations where hospitalization predictions are a linear or quadratic function of ILI.
-
-The simulation study in section \ref{sec:simulation2} suggests the ASG function in ILI modeling may slightly outperform the SIR model according to the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season forecasts, the SIR model was overall superior.
-The results from both the simulation study and the real data analysis suggest that the addition of a discrepancy component in ILI modeling may improve forecasts especially near the holiday week between Christmas and New Years day.
-It should not be assumed that these conclusions may be generalized for all locations in the US, weeks of a season, or for future flu seasons.
+ In this manuscript we introduce a statistical modeling framework which allows 
+ for the incorporation of several ILI forecast modeling methods. Specifically, 
+ we built upon Osthus et al. \cite[]{osthus2019dynamic} and introduced a 
+ framework for modeling ILI which includes the use of an arbitrary function 
+ for modeling the main trajectory of ILI along with modeling the discrepancy. 
+ We model flu hospitalizations by incorporating the ILI forecast model into a 
+ model forecasting hospitalizations where hospitalization predictions are a 
+ linear or quadratic function of ILI.
+
+The simulation study in section \ref{sec:simulation2} suggests the ASG 
+function in ILI modeling may slightly outperform the SIR model according to 
+the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season 
+forecasts, the SIR model was overall superior. The results from both the 
+simulation study and the real data analysis suggest that the addition of a 
+discrepancy component in ILI modeling may improve forecasts especially near 
+the holiday week between Christmas and New Years day. It should not be assumed 
+that these conclusions may be generalized for all locations in the US, weeks 
+of a season, or for future flu seasons. 
 % Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.
 
-Forecasting the seasonal influenza outbreak remains a challenging task for forecasters.
-The general modeling framework in this manuscript is successful under diverse modeling conditions for all locations in the US and may contribute to future forecasting efforts.
-All forecast models were presented as separate from one another, but in the various locations and times of the flu season, different models perform better than others.
-To build on the work done here, a natural step forward would be to combine all or a few selected forecasts into an ensemble forecast.
-Such an ensemble may work to cancel out certain model biases or highlight model strengths, leading to more robust forecasts.
-
-\section{Acknowledgments}
-This work is partially supported by the National Science Foundation under Grant No.
-2152117.
-Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
-
-\section{Supplementary Material}
-Supplement to "Forecasting Influenza Hospitalizations Using a
-Bayesian Hierarchical Nonlinear Model with
-Discrepancy"\\
-A. Additional plots for ILI \& Hospitalization data for select regions, results from simulation study, and 2023-24 flu analysis\\
-% \end{supplement}
-B. Posterior distribution plots for select parameters\\
-% \sdescription{2. Plots of posterior distributions for parameters in both ILI and hospitalization models}
-C. FluSight forecast competition scoring results
-
-%Bibliography
-% \bibliographystyle{unsrt}  
-% \bibliography{references}
-
-\bibliographystyle{apa}
-\bibliography{master_bib}
-
-\end{document}
+Forecasting the seasonal influenza outbreak remains a challenging task for 
+forecasters. The general modeling framework in this manuscript is successful 
+under diverse modeling conditions for all locations in the US and may 
+contribute to future forecasting efforts. All forecast models were presented 
+as separate from one another, but in the various locations and times of the 
+flu season, different models perform better than others. To build on the work 
+done here, a natural step forward would be to combine all or a few selected 
+forecasts into an ensemble forecast. Such an ensemble may work to cancel out 
+certain model biases or highlight model strengths, leading to more robust 
+forecasts.

### 2025-06-30 nonlinear_forecast.tex
reformat sections
fix full fit

@@
 \begin{frontmatter}
-\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy}
+\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical 
+Nonlinear Model with Discrepancy}
 %\title{A sample article title with some additional note\thanksref{t1}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author[A]{\snm{Spencer Wadsworth}~\ead[label=e1]{sgw96@iastate.edu}}
+\author[A]{\snm{Spencer Wadsworth}\ead[label=e1]{sgw96@iastate.edu}}
 % \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}\orcid{0000-0000-0000-0000}}
@@
 
-\begin{abstract}
-The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread.
-The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods.
-For the first several years of the competition, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalization counts.
-Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons.
-In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data.
-The first component is for modeling ILI data using a dynamic nonlinear Bayesian hierarchical model.
-The second component is for modeling hospitalizations as a function of ILI.
-For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor.
-In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from \cite{osthus2019dynamic} and a nonlinear Bayesian hierarchical model from \cite{ulloa2019} are compared.
-Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model.
-Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared.
-We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year.
-We also found that other modeling decisions such as the exact nonlinear function to use in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
-\end{abstract}
+\input{abstract}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
 
-Every year the seasonal influenza outbreak burdens the public health system by infecting millions, causing an influx of primary care visits and hospitalizations and leading to between 290,000 and 650,000 deaths worldwide \cite[]{whoflufact2023}.
-\cite{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion.
-Accurate forecasting of infectious diseases can inform public decision making and ease the burden of an outbreak \cite[]{turtle2021accurate, lutz2019applying}.
-There is a growing consensus that disease forecasts should be probabilistic in nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has been shown that reporting forecast uncertainty along with predictions may lead to better decision making \cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.
-
-% \jarad{Test}
-% \spencer{Test}
-
-To better inform public decision making regarding the flu epidemic, in 2013 the United States Centers for Disease Control and Prevention (CDC) organized a national flu forecasting competition, also known as FluSight \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}.
-Originally, over a dozen teams of researchers from academic and industry backgrounds participated in FluSight by contributing their own forecast models.
-Besides the 2020 season --or the flu season spanning the fall of 2020 and the winter of 2021-- FluSight has been operated annually and researchers outside the CDC have been invited to participate.
-Initially the target data for forecasts was influenza-like illness (ILI) data.
-ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}.
-The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, \cite{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework.
-An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
-
-The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications.
-As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}.
-The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded.
-The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}.
-Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition.
-Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}.
-This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.
-
-The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models.
-The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
-Herein we use ILI models similar to those in \cite{osthus2019dynamic} and \cite{ulloa2019} for ILI forecasting.
-The model of \cite{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework.
-Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by \cite{kennedy2001bayesian}.
-The model in \cite{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data.
-Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.
-
-In line with the newer FluSight standard of forecasting hospitalizations, we model hospitalizations as a linear function of ILI.
-Thus forecasts produced herein target flu hospitalizations and are a mapping of ILI forecasts to hospitalizations.
-This allows for ILI data from many seasons to be exploited and for ILI forecasts to assist in forecasting hospitalizations, which has fewer seasons of data than ILI.
-Several modeling schemes and their forecasts for the 2023 flu hospitalization season are compared, and it is shown that the modeling decisions produce good forecast results for different states or times during the flu season.
-
-In section \ref{sec:data} we review the ILI and hospitalization data provided by the CDC and targeted by FluSight. 
-In section \ref{sec:functions} the modeling framework contributed by this manuscript is given.
-In the same section, functions similar to those used by \cite{osthus2019dynamic} and \cite{ulloa2019} are defined.
- These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively.
-Model fitting and implementation are described at the end of the section.
-Section \ref{sec:simulation2} is a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared.
-Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts.
-Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}, with the analysis being done using the conventions of FluSight.
- Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.
-
- \section{Flu outbreak data} \label{sec:data}
-In this section we introduce and define ILI and hospitalization data and evaluate the data visually.
-ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season.
-Both of these data were collected at the state, territorial, and national level and were reported at least weekly.
-Overall the data is reported for 53 locations including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), and at the US national level.
-We will refer to forecast targets throughout this manuscript.
-A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a specific location and week during the season.
-
-\subsection{Influenza-like illness data}
-The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers.
-Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases.
-An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}.
-Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza.
-To know whether or not a sick patient is infected with influenza would require a laboratory test.
-
-In 2013, when FluSight began, the ILI data was the object of the forecasts.
-The data was released publicly at HHS region levels, and forecast teams were asked to provide forecasts of several ILI targets on the regional levels including season onset, 1-4 week ahead ILI levels, and the week of peak ILI activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}.
-Currently, the ILI data is collected by the CDC and published on an online portal for viewing at the national, HHS region, census, and state levels \cite[]{cdc2024fluviewportal}.
-To obtain ILI data, we used the \texttt{R} package \texttt{cdcfluview} which provides functions for downloading the data \cite[]{rudis2021cdcfluview}.
-Weekly ILI data from the national, HHS region, and census levels are available from the 1997 flu season until the current season.
-At the state level, data is available from the 2010 flu season to the current season.
-
-Figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023.
-For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons.
-To better align with the flu behavior, week 1 is set as the first week of August and week 52 or 53 is the last week in July of the following year.
-For example week 1 of the 2013 season corresponds to the first week of August 2013, and week 52 of the same season corresponds to the last week of July 2014.
-This convention is used for the remainder of this manuscript.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/us_ili_seasons.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023.
-Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.}
-    \label{fig:us_ili}
-\end{figure}
-
-Notable from the plots in figure \ref{fig:us_ili} is the regular trajectory of the ILI.
-With the exception of season 2020, the ILI begins low at week 1 and increases as the fall and winter progress until the ILI reaches a peak.
-As spring progresses to summer, the ILI decreases to low values.
-As \cite{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day.
-Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
-\cite[]{ewing2017contact, garza2013effect}.
-The only seasons when there was not a peak at week 22 were season 2022 where the season peak occurred particularly early and season 2020 which was greatly influenced by the COVID-19 pandemic.
-
-\subsection{Hospitalization Data}
-
-Hospital admission data, used as the object of FluSight forecasting for the 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported Patient Impact and Hospital Capacity by State Timeseries}.
-Several targets of respiratory illnesses including COVID-19, RSV, and influenza are reported weekly by most hospitals in the US.
-In February 2022 it became mandatory for all hospitals to report the number of COVID-19 and influenza hospitalizations, and since then reporting of hospitalizations has become widespread.
-These data were updated every Wednesday and Friday according to NHSN guidelines \cite[]{healthdata2024covidts}.
-
-Figure \ref{fig:us_hosp} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons.
-These plots show similarities to the ILI plots in \ref{fig:us_ili} in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends.
-For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22.
-Additional plots of ILI data and hospitalization
-data for select states are included in the supplementary material 
-\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar to 
-those in figures \ref{fig:us_ili} and \ref{fig:us_hosp} with activity increasing
-to a peak and then decreasing.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/us_hospitalizations.png}
-    \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
-    \label{fig:us_hosp}
-\end{figure}
-
-\section{ILI and hospitalization forecast modeling} \label{sec:functions}
-
-% \jarad{Why not start with the ILI model from Osthus et al?
-It makes a lot more sense to me to start with what has already been done and then expand upon what they have done.
-Later you have a section called ILI Model, but isn't this the ILI model?}
-
-The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows of a similar trajectory for modeling ILI.
-Compartmental models are standard mathematical models used for disease outbreaks.
-One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}.
-\cite{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data.
-The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy.
-In the first part of this section, we present an ILI model similar to the model in \cite{osthus2019dynamic}.
-With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.
-
-With the aim of forecasting hospitalizations, we also introduce a linear model of hospitalizations data with ILI as a predictive covariate.
-To forecast hospitalizations, ILI data is first forecasted and the forecast is then plugged in as a covariate in the hospitalization model, thus producing hospitalization forecasts.
-The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.
-
-\subsection{ILI Model} \label{sec:ili_model}
-
-The proposed model for ILI for any location is given in (\ref{eq:ili_model}).
-Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given season.
-The ILI is a proportion, so the Beta random variable is a natural selection for modeling.
-Under the parameterization in of the Beta distribution used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a scale parameter.
-The nonlinear function $f_{\theta_s}(w)$ captures the trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the systematic patterns which $f_{\theta_s}(w)$ does not capture.
-\begin{equation}
-\begin{aligned}
-    \label{eq:ili_model}
-    % \begin{flalign*}
-        ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
-        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w
-    % \end{flalign*}
-\end{aligned}
-\end{equation}
-In \cite{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}.
-In \cite{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}.
-In \cite{ulloa2019} modeling is done hierarchically over seasons.
-
-\subsection{Susceptible-Infectious-Recovered (SIR) compartmental model} \label{sec:sir_func}
-
-The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by \cite{kermack1927contribution}.
-Since then, compartmental models have become standard for modeling infectious diseases \cite[]{allen2008mathematical}, and many extensions have been made and studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}.
-The SIR mathematical model includes three compartments and assumes that at any time $t>0$ every individual in a closed population belongs to exactly one compartment.
-The three compartments are susceptible ($S$), infectious ($I$), and recovered ($R$), and their interaction over the course of an outbreak is described by the differential equations in (\ref{eq:sir_diff}).
-
-\begin{equation}
-    \label{eq:sir_diff}
-    \frac{dS}{dt} = -\beta SI, \quad \frac{dI}{dt} = \beta S I - \delta I, \quad \frac{dR}{dt} = \delta I
-\end{equation}
-Here $S$, $I$, and $R$ represent the proportion of the population in each compartment such that $S + I + R = 1$ for all $t$.
-The trajectory is determined by the disease transmission rate $\beta > 0$ and the recovery rate $\delta > 0$.
-Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual, and the expected rate of recovery to an immune state for a newly infected person.
-Whether or not a disease outbreak is classified as an epidemic is determined by the initial susceptible population $S_0$, or the susceptible population at time $0$, and the parameter $\rho = \delta/
-\beta$.
-If $S_0/\rho > 1$, the outbreak is considered an epidemic.
-It is non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}.
-Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where the $S_0$ and $\rho$ were selected to match an outbreak that would classify as epidemic.
-In the case where $S_0 \leq \rho$, the trajectory for the $I$ compartment would never be increasing.
-The increase to a peak and subsequent decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is reasonable to model ILI by this compartment.
-Thus in modeling the ILI data, we consider the data to be analogous to the $I$ proportion of the population.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.6]{Images/sir_traj.png}
-    \caption{Susceptible-infectious-recovered (SIR) model separated by compartments.
-The three compartments are the susceptible compartment (left), infectious (center), and recovered (right).
-In this example, $S_0/\rho > 1$.}
-    \label{fig:sir_traj}
-\end{figure}
-
-\subsection{Asymmetric Gaussian (ASG) function} \label{sec:asg_func}
-The ASG function is another example of a nonlinear function which can approximate the trajectory of the flu outbreak.
-The ASG was previously used by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used to model vegetation growth and satellite sensor data \cite[]{lewis2020extracting, jonsson2002seasonality, hird2009noise, beck2006improved, atkinson2012inter}.
-The ASG is a modification of the asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized by its rise to a peak and fall from that peak which may not occur at the same rate, as shown in figure \ref{fig:asg_function}.
-The ASG function is denoted as $ASG_\theta(w)$ where $\theta = (\lambda, \nu, \mu, \sigma_1^2, \sigma_2^2)$, $\nu > 0$, $\lambda > 0$, $\mu \in (-\infty, \infty)$, $\sigma_1, \sigma_2 > 0$ and $w \in (1, ..., W)$ is week.
-The function is defined in (\ref{eq:asg_function}).
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{Images/asg_function.png}
-    \caption{Example plot of asymmetric Gaussian (ASG) function showing the shape of the function in relation to the parameters $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
-    \label{fig:asg_function}
-\end{figure}
-
-\begin{equation}
-    \label{eq:asg_function}
-    ASG_{\theta}(w) = 
-    \begin{cases}
-        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
-        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
-    \end{cases}
-\end{equation}
-For modeling in this manuscript, we use a slightly reparameterized version of the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$.
-This constraint guarantees that the function has a peak greater than $\lambda$.
-
-\begin{equation}
-    \label{eq:asg_function_rep}
-    ASG_{\theta}(w) = 
-    \begin{cases}
-        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
-        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
-    \end{cases}
-\end{equation}
-
-\subsection{Model discrepancy}
-
-The SIR and ASG functions are useful for capturing the main trend of the ILI data, but as \cite{osthus2019dynamic} points out there may be systematic behavior that these or other possible functions may not capture.
-As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a regular peak at week 22 of the flu season.
-Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to illustrate the systematic discrepancy from a fitted function.
-Figure \ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 to 2022 excluding 2020 with a best fit ASG function plotted over the ILI.
-The fits for each season were made by obtaining the maximum likelihood estimate (MLE) of a model given the ILI data where we assume the ASG function is the mean parameter of a Beta distributed random variable.
-Figure \ref{fig:discrepancy} shows the discrepancy between the model fit and the data for the same seasons.
-The grey lines show the difference between the data and the functions from figure \ref{fig:asg_fits} for each season, and the black line is the average by week over all seasons.
-The lines show that the ASG function typically underpredicts week 22 and overpredicts week 23.
-Perhaps for other weeks, like week 30 for example, there also tends to be systematic behavior not captured by the ASG function.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.5]{Images/asg_fits.png}
-    \caption{Observed US national influenza-like illness (ILI) percentage for seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an asymmetric Gaussian (ASG) model for the ILI data (black)}
-    \label{fig:asg_fits}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{Images/discrepancy.png}
-    \caption{Difference between observed US national influenza-like illness (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 2010 to 2022 excluding 2020 (grey) and the average difference of all seasons (black)}
-    \label{fig:discrepancy}
-\end{figure}
-
-The term $\gamma_w$, where $w$ is the season week, is included in (\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the function.
-Modeling discrepancy has been used in uncertainty analysis of simulators to capture systematic differences between mathematical models and
-reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,arendt2012improving,kennedy2001bayesian}.
-Modeling discrepancy can lead to overfitting, particularly in forecasting scenarios, and may also lead to identifiability issues.
-Thus, care must be taken in setting parameter constraints as well as in the selection of prior distributions \cite[]{osthus2019dynamic,brynjarsdottir2014learning}. 
-Modeling of the discrepancy for ILI was done by \cite{osthus2019dynamic} during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}.
-As in Osthus's model, $\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}).
-\begin{equation}
-    \label{eq:rw}
-        \gamma_w|\gamma_{w + 1} \overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
-        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W})
-\end{equation}
-The idea for using the reverse random walk is that there are several previous seasons of ILI data, and assuming the random walk captures systematic behavior, fitting it hierarchically over seasons can assist in predicting future behavior in the current season.
-Reverse random walks have also been used with success in election forecasting and other flu forecasting models \cite[]{osthus2021multiscale, osthus2019dynamic, linzer2013dynamic}.
-The sum to zero constraint $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve identifiability.
-
-\subsection{ILI forecasts}
-Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating.
-Future ILI forecasts are obtained via the posterior predictive distribution where for week $w$, the predictive distribution is obtained by integrating over the parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ as in (\ref{eq:ili_post}) where $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ is the density function of the posterior distribution for the model parameters.
-If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ where $i$ is a positive integer.
-
-\begin{equation}
-    \label{eq:ili_post}
-    p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) = \int \int \int \int p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
-\end{equation}
-% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
-% \spencer{Isn't that captured in $d \pi_s$?}
-
-\subsection{Hospitalization model} \label{sec:hospital_model}
-
-The second component for forecast modeling is based on the relationship between hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
-This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}. Figure \ref{fig:lag_scatter}
-shows scatterplots of the difference between hospitalizations and scaled 1 week 
-lags by ILI percentage at the US national level. The points in the left plot are
-from hospitalization differences whereas in the right the hospitalizations were
-log transformed. The positive and slightly curved relationship suggests model
-(\ref{eq:final_hosp}) is a reasonable choice.
-
-\begin{figure}
-    \includegraphics[scale=.46]{Images/lagdiff_vs_ili.png}
-    \caption{Scatterplots of national hospitalizations (left) or log hospitalizations (right) lag differences by ILI \% where the lag is scaled by $\phi$.
-Points are colored by season.
-For hospitalizations $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$.
-}
-    \label{fig:lag_scatter}
-\end{figure}
-
-\begin{equation}
-    \begin{aligned}
-    \label{eq:final_hosp}
-    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
-    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, \omega_s) %N(0, \sigma_{\epsilon}^2)
-    \end{aligned}
-\end{equation}
-In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
-
-This model is for any location, and for fitting purposes $ILI_{s,w}$ is always multiplied by $P$ which is proportional to the population of the state or territory, in this case the total population divided by 50,000.
-This is done as a means of scaling so that the prior distribution assigned to  $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, $\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all states.
-
-Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also fit via Bayesian posterior updating.
-To obtain forecasts for $H_{s, w^* + i}$, the ILI posterior predictive distribution is used along with the posterior distribution for the parameters in (\ref{eq:final_hosp}).
-We considered three scenarios for model (\ref{eq:final_hosp}).
-A model where $D_s$ belongs to the normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, or $H_{s,w} + 1$ is lognormally distributed (LNORM).
-The population value $P$ is excluded from the LNORM model, and we set $\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting.
-In the LNORM model, if the linear parameters are not set the same for seasons 22 and 23, the final variance was more prone to be extreme.
-Besides varying the distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ or there is no quadratic ILI term.
-
-\subsection{Prior selection}
-
-The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in \cite{osthus2019dynamic} and \cite{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem.
-For model (\ref{eq:ili_model}), parameters which are common even when using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, $\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$.
-For the SIR function $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$.
-For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
-
-The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by \cite{gelman2006prior}, the priors were assigned to the variance parameters.
-Univariate parameters were assigned either a normal distribution prior if the support is on $\mathbb{R}$, a half-normal prior if the support is nonnegative, or a truncated-normal prior to match a more specific support.
-Under the ASG model, $\theta_s$ is modeled hierarchically over seasons so that for each season the transformed parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are assigned to $\theta$ and $\Sigma$.
-
-Additional prior constraints were made to improve parameter identifiability.
-In \cite{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$.
-The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative priors.
-To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the parameters.
-A modular Bayesian approach involves multiple steps of parameter fitting where some parameters may be estimated without priors via maximum likelihood estimation or other means.
-Fitting the rest of the model parameters involves assigning priors and conditioning on the previously fit parameters.
-This has been done in computer modeling to improve identifiability and other issues, though \cite{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}.
-We carried out the modular fit by first estimating the maximum likelihood estimate (MLE) for the parameter $\lambda_s$ for each season in (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
-
-% \spencer{Blue stuff below will be either included further on or will be removed entirely
-
-\subsection{Parameter estimation and posterior predictive sampling}
-\label{sec:implementation_posterior}
-
-The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the \cite{stan2024manual} \cite[]{gabry2022stan}.
-Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}.
-The \texttt{cmdstanr} package provides several diagnostic statistics for assessing the sampler.
-As mentioned, most model parameter prior distributions were intended to be uninformative.
-Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material \cite[]{wadsworth2024bas}.
-
-We assessed the model fit for four ILI models.
-These included the SIR and ASG models and models with and without discrepancy modeling.
-When discrepancy is included, the models are denoted as SIRD and ASGD.
-These models were fit using US national data from 2010 to 2023 flu seasons, where data from the 2020 season was excluded because of the unique behavior during that season.
-Assessment for hospitalization modeling was done for six different models.
-These include NORM, LNORM, and LST models, and models where a quadratic ILI term is included or excluded.
-To assess posterior sampling convergence, models were fit to data where ILI and hospitalization data up to week 14 of the 2023 season was included.
-Sampling was done with four chains where from each chain 60,000 posterior draws were sampled, and the first 10,000 draws were discarded as a burn-in.
-The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for each parameter.
-Tables \ref{tab:ili_diagnostics} and \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum $ESS$ over all parameters for ILI and hospitalization models respectively.
-Forecast models for all other weeks of the season were fit using one chain of 60,000 draws where the first 10,000 draws were discarded as a burn-in.
-% \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
-For parameters of the ASG models that were prone to cause trouble in posterior sampling, the starting values were set to be the MLEs.
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI models fitted on US data for week 14 of the 2023 season}
-\label{tab:ili_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccc}
-     & ASG & ASGD & SIR & SIRD \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
-     $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
-    \end{tabular}
-    \end{center}
-\end{table}
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six hospitalization models fitted on US data for week 14 of the 2023 season}
-\label{tab:hsop_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccccc}
-     & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$\\
-     $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
-    \end{tabular}
-    \end{center}
-\end{table}
-
-To obtain forecast distributions of hospitalizations, draws from the posterior predictive distribution from the ILI model were used in conjunction with the posterior distribution of the hospitalizations model.
-When fitting model (\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were saved.
-Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal distributions for the model parameters were saved.
-To obtain forecast distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following steps are repeated $K$ times where $K$ is an integer for the number of desired samples.
-We set $K = 50,000$.
-
-\begin{quote}
-\begin{enumerate}[Step 1:]
-  \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
-  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior distributions
-  \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
-  \begin{itemize}
-    \item[] $\mu_{s,w + i}^* = \alpha_{0s}^* + \alpha_{1s}^* (ILI_{s,w + i}^* \times P) + \alpha_{2s}^* (ILI_{s,w + i}^* \times P)^2 + \phi^* H^*_{s,w + i - 1}$
-  \end{itemize}
-   \item Repeat step 3 for $i \in \{1,2,3,4\}$ to obtain $H^*_{s,(w + 1):(w + 4)}$
-  \item Repeat steps 1-4 $K$ times
-\end{enumerate}
-\end{quote}
-The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast for hospitalizations at week $w + i$.
-For the forecast competition analysis in section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were set to 0 to reflect realistic values of hospitalizations and comply with the FluSight forecasting rules.
-
-\section{Simulation Study} \label{sec:simulation2}
-In this section, we present a simulation study conducted for comparing ILI models and further assessing the hospitalization forecast model. US ILI data is used, and hospitalization data is simulated. 
-A leave-one-season-out (LOSO) approach was combined with a Monte Carlo simulation approach.
-For each replication, we simulated log-hospitalizations for all weeks during seasons 2010, ..., 2022, excluding 2020, using the existing ILI data as a predictive covariate.
-Each season was in turn ``left-out" and treated as if it was the most recent season which we desired to forecast.
-Fitting and forecasting was then done for weeks 14, 20, 26, 32, and 38 of the left out season, giving two weeks that tend to occur as flu cases increase, two as cases decrease, and one that occurs when the cases may be increasing or decreasing.
-Week 20 is a week leading up to the holiday week 22 where ILI typically has a local peak.
-We were particularly interested in how important modeling discrepancy is for forecasting at week 20.
-
-For the simulation of hospitalizations, the parameters $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and $\sigma^2_{\epsilon_s}$ from the hospitalization model in (\ref{eq:final_hosp}) were considered the same across all seasons so that all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$.
-The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization data from the 2022 season.
-For fitting, the hospitalization data was first log-transformed.
-We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
-and assigned the noninformative prior $p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$.
-The marginal posterior distribution $\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ was then the established posterior multivariate normal distribution and $\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior distribution \cite[]{gelman2013bayesian}.
-The posterior means of those parameters were used as the values from which log-hospitalizations were simulated.
-The number of replicates in the simulation was 500.
-
-Model comparison was done by calculating the continuous ranked probability score (CRPS) and the logarithmic score (LogS) for each forecast.
-These scores are both proper scoring rules which evaluate the forecast distribution and density functions respectively.
-Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}.
-Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The CRPS is defined in (\ref{eq:crps2}) and the LogS in (\ref{eq:logs2}).
-Here $F(\cdot)$ is the forecast distribution function, $f(\cdot)$ is the forecast density function, and $y^*$ is the observed targeted value of the forecast.
-The orientation of both scores is negative, meaning the smaller the score the better.
-Here $\vmathbb{1}\{\cdot\}$ is the indicator function.
-
-\begin{equation}
-    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
-    \label{eq:crps2}
-\end{equation}
-
-\begin{equation}
-    \text{LogS}(f, y^*) = -\text{log}(f(y^*))
-    \label{eq:logs2}
-\end{equation}
-Both the CRPS and LogS are calculated using the \texttt{scoringRules} package in \texttt{R} \cite[]{jordan2019scoring}.
-To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation.
-The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
-
-Figures \ref{fig:crps_by_week_horizon} and \ref{fig:logs_by_week_horizon} show 
-boxplots of results from the simulation. The figures respectively
-show boxplots of the CRPS and LogS for the 500 simulation replicates faceted by week and horizon.
-Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
-Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with lower medians and lower variation than the two models not accounting for ILI discrepancy.
-This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough.
-This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
-Among the four models, ASG tends to have the lowest values in CRPS and LogS though this is not always the case.
-This suggests that the additional flexibility of the five parameter ASG function is often better for forecasting than that of the three parameter SIR function.
-Additional
-plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and LogS.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{Images/crps_by_week_horizon.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
-    \label{fig:crps_by_week_horizon}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{Images/logs_by_week_horizon.png}
-    \caption{Boxplots of the logarithmic score (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
-    \label{fig:logs_by_week_horizon}
-\end{figure}
-
-\section{Analysis of forecasts for 2023 flu season}
-\label{sec:analysis}
-
-In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations.
-The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$.
-A single forecast is thus comprised of 11 predictive intervals and a median.
-Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, and forecasts were made at the state and national level.
-The first week of forecasting took place during the week of October 7, 2023, and the final week was the week of April 27, 2024 making 29 total weeks of forecasts.
- The same format was used during the 2021 and 2022 seasons and for the COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
-% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
- Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS).
-The focus in this section will be on the LWIS (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
-
-The WIS is a proper scoring rule used for scoring quantile or interval forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all included quantiles, $B$ is the number of intervals,  $y^*$ is the observed value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ interval.
-$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a single interval as defined in (\ref{eq:is}).
-The goal of the forecaster is to minimize the WIS.
-The LWIS is the same as the WIS except that it is evaluated over the log of quantiles and the log of the observed value.
-
-\begin{equation}
-\label{eq:wis2}
-        WIS_{0,B}(Q, y^*) = \frac{1}{B + 1/2} \times (w_0\times |y^* - median| + \sum_{b=1}^B \{w_k \times IS_{\alpha_b}(Q, y^*) \} )
-\end{equation}
-
-\begin{equation}
-\label{eq:is}
-        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\vmathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \vmathbb{1}\{y^* > r\}
-\end{equation}
-
-We fit 24 forecast models for each location for all 29 weeks, and for each week forecast 1-4 week ahead hospitalizations.
-The 24 models included all combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST hospitalization models, and both quadratic and linear hospitalization models.
-The prior distributions under the SIR model are in (\ref{eq:sir_prior}).
-Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$.
-Here $\vmathbb{1}_{A}$ represents the indicator function for values within the set $A$.
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:sir_prior}
-        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
-        \beta_s &\sim N^+(0.8, 0.3) \\
-        \rho_s &\sim N^+(0.68, 0.08)
-\end{aligned}
-\end{equation}
-For the ASG model, the MLE $\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value.
-The remaining estimates were used as starting values for posterior sampling.
-The transformation $T(\theta_s) = (\text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ was made, and a prior distribution was assigned to $T(\theta_s)$.
-The prior distributions for the ILI model under the ASG function are shown in (\ref{eq:ili_prior}).
-These are slightly informative priors because for most parameters we have an idea what reasonable values may be.
-Here $m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
-where $\text{diag}(\cdot)$ 
-is the diagonal matrix for the given entries.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:ili_prior}
-                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
-                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
-                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
-                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
-\end{aligned}
-\end{equation}
-Parameters shared by both the SIR and ASG models are the scale parameter $\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-These priors are in (\ref{eq:shared_ili_priors}).
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:shared_ili_priors}
-        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
-        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2)
-\end{aligned}
-\end{equation}
-Because of the limited information for estimating $\sigma_{\gamma_W}^2$, we selected an informative prior distribution by first estimating  $\hat{\sigma}_{\gamma_W}^2$.
-This was estimated for a given state by first calculating the MLE for $\theta_s$.
-% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
-%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
-$\widehat{ILI}_{s,w}$ was then predicted such that $\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season.
-Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over seasons of the differences $\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$.
-When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was calculated using the \texttt{mle2} function in the \texttt{bblme} package \cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} package \cite[]{soetaert2010desolve} function in \texttt{R}.
-Where the ASG function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} function.
-The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
-
-\begin{equation}
-\begin{aligned}
-\label{eq:shared_hosp_priors}
-        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
-        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
-        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
-\end{aligned}
-\end{equation}
-
-\begin{figure}
-\centering
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 season for weeks 14, 20, 26, and 32.
-Forecasts are separated by ILI model, and the hospitalization models are all normally distribution.
-The figure includes hospitalization forecasts where ILI is a linear predictor (left) and where ILI is a quadratic predictor (right).
-50\% predictive intervals are pink and 95\% predictive intervals are red.}
-\label{fig:normal_flu_forecasts}
-\end{figure}
-
-Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US flu hospitalizations during the 2023 season under the NORM hospitalization model.
-The forecasts shown are for weeks 14, 20, 26, and 32.
-The predictive bands are the 50\% and 95\% predictive intervals.
-These plots show a tendency to often underpredict hospitalizations.
-The quadratic models appear to do a better job predicting  hospitalizations at the season peak but a poorer job predicting after the peak, whereas the linear models seem to predict well or slightly overpredict after the peak.
-% From these figures, it's difficult to say whether ASG or SIR models predict better, but at least for the quadractic models the week 22 predictions are maybe slightly better than for the linear models.
-
-Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the season for all 24 models of US hospitalizations.
-LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling.
-Here, a darker value represents a larger LWIS and worse forecast compared to lighter values.
-The weeks around the holiday week 22 are highlighted by a red band.
-In most cases, there appears to be a turning point in performance at or near week 22.
-The models which include discrepancy appear to forecast better around week 22, though they may not outperform the models without discrepancy through the whole season.
-Such is the case for the NORM and LST quadratic models and the linear LNORM models.
-It's also notable that for the SIR model, including discrepancy appears to make for poorer forecasts after week 22 whereas for ASG, including discrepancy appears to improve forecasts.
-
-\begin{figure}
-
-    % \centering
-    \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
-    \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
-Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
-Scores for models with an ASG ILI model are above while those with an SIR model are below.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:us_lwis}
-\end{figure}
-
-\begin{figure}
-
-    \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
-Scores are averaged over all horizons 1-4 weeks ahead.
-Scores are faceted by hospitalization model distribution.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:lwis_by_dist_loc}
-\end{figure}
-
-Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by LWIS, but all 53 locations are included.
-The LWIS is shown across the whole season faceted by hospitalization model distribution.
-Overall, the weeks leading up to week 22 are the most difficult to forecast, but the LNORM model appears to perform the better than the other models during the weeks around the holiday week 22.
-The overall, calculated as the mean LWIS over all locations and weeks, for all 24 models is shown in table \ref{tab:forecast_scores}.
-The first main takeaway is that the top four performing models are SIR ILI models with the LNORM hospitalization models.
-The next main takeaway is that overall the models which include discrepancy outperformed the models without discrepancy.
-
-\begin{table}
-\caption{Overall scores for each of the 24 forecast models.
-The overall score is the log weighted interval score (LWIS) averaged over all locations, weeks, and horizons.
-The scores in the first two rows are for linear models, and the scores in the third and fourth rows are for quadratic models.
-The lowest WISs are bolded.}
-\begin{tabular*}{\textwidth}
-{@{\extracolsep{\fill}} 
-    l*{9}{c}}
-  & & \multicolumn{3}{c}{ASG} 
-  & \multicolumn{3}{c}{SIR} \\ 
-  \cmidrule{3-5} \cmidrule{6-8}
-  & & LNORM & LST & NORM & LNORM & LST & NORM\\
-  \midrule
-  Linear & No Disc & 0.366 & 0.392 & 0.365 & \textbf{0.355} & 0.416 & 0.394 &\\ 
-   & Disc & 0.358 & 0.391 & 0.365 & \textbf{0.343} & 0.382 & 0.354 &\\
-  \midrule
-  Quadratic & No Disc & 0.366 & 0.398 & 0.377 & \textbf{0.355} & 0.401 & 0.377 &\\ 
-   & Disc & 0.358 & 0.390 & 0.369 & \textbf{0.343} & 0.378 & 0.356 &\\     
-  \bottomrule
-\end{tabular*}
-\label{tab:forecast_scores}
-\end{table}
-
-It should not be assumed that these two takeaways will apply for future flu seasons.
-As suggested by the simulation study in the previous section, it may often be the case that the ASG ILI model is better for forecasting.
-Indeed, a close examination of figures \ref{fig:us_lwis} and \ref{fig:lwis_by_dist_loc} shows that it is common for the ASG models to show better forecasting skill than the SIR models.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
- In this manuscript we introduce a statistical modeling framework which allows for the incorporation of several ILI forecast modeling methods.
-Specifically, we built upon \cite{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy.
-We model flu hospitalizations by incorporating the ILI forecast model into a model forecasting hospitalizations where hospitalization predictions are a linear or quadratic function of ILI.
-
-The simulation study in section \ref{sec:simulation2} suggests the ASG function in ILI modeling may slightly outperform the SIR model according to the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season forecasts, the SIR model was overall superior.
-The results from both the simulation study and the real data analysis suggest that the addition of a discrepancy component in ILI modeling may improve forecasts especially near the holiday week between Christmas and New Years day.
-It should not be assumed that these conclusions may be generalized for all locations in the US, weeks of a season, or for future flu seasons.
-% Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.
-
-Forecasting the seasonal influenza outbreak remains a challenging task for forecasters.
-The general modeling framework in this manuscript is successful under diverse modeling conditions for all locations in the US and may contribute to future forecasting efforts.
-All forecast models were presented as separate from one another, but in the various locations and times of the flu season, different models perform better than others.
-To build on the work done here, a natural step forward would be to combine all or a few selected forecasts into an ensemble forecast.
-Such an ensemble may work to cancel out certain model biases or highlight model strengths, leading to more robust forecasts.
+\input{manuscript_res}
 
 \begin{acks}[Acknowledgments]
-This work is partially supported by the National Science Foundation under Grant No.
-2152117.
-Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
+This work is partially supported by the National Science Foundation under 
+Grant No. 2152117. Any opinions, findings, and conclusions or recommendations 
+expressed in this material are those of the author(s) and do not necessarily 
+reflect the views of the National Science Foundation.
 \end{acks}
@@
 Discrepancy"\\
-1. Additional plots for ILI \& Hospitalization data for select regions, results from simulation study, and 2023-24 flu analysis\\
+A. Additional plots for ILI \& Hospitalization data for select regions, results 
+from simulation study, and 2023-24 flu analysis\\
 % \end{supplement}
-2. Posterior distribution plots for select parameters\\
+B. Posterior distribution plots for select parameters\\
 % \sdescription{2. Plots of posterior distributions for parameters in both ILI and hospitalization models}
-3. FluSight forecast competition scoring results
+C. FluSight forecast competition scoring results
 \end{supplement}

### 2025-06-30 flu_forecast_23/full_fit_hosp.R

@@ -115,6 +115,9 @@ all_forecasts <- foreach(j = select_regions,
    ili_forc <- read.csv(paste0("./ili_fits/", ili_model, "/reference_date_", date(reference_date), 
 			"_week", k, "/", j, ".csv"))
    
+   ili_forc <- ili_forc %>%
+     filter(if_all(everything(), ~ !is.na(.)))
+   
    forecasts <- data.frame()
    count_rate <- unique(both_flu_hold$count_rate2)
    count_rate_sig <- count_rate
@@ -183,7 +186,10 @@ all_forecasts <- foreach(j = select_regions,
 
 	}
 	forecasts <- rbind(forecasts, hosp_forecast)
-   }	   
+   }
+   forecasts <- forecasts %>%
+     filter(if_all(everything(), ~ !is.na(.)))
+   
    colnames(forecasts) <- paste0("ahead", -1:3)
    #print(head(forecasts))
    probs <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)

### 2025-06-30 flu_forecast_23/full_fit_hosp_li.R
add li

@@ -1,0 +1,240 @@
+setwd("./")
+library(tidyr)
+library(readr)
+source("./get_data_functions_new.R")
+source("./mle_functions.R")
+library(parallel)
+library(doParallel)
+library(doMC)
+library(evalcast)
+#n.cores <- detectCores()
+n.cores <- 64
+my.cluster <- makeCluster(n.cores, type = "PSOCK")
+doParallel::registerDoParallel(cl = my.cluster)
+foreach::getDoParRegistered()
+foreach::getDoParWorkers()
+registerDoMC(cores = n.cores)
+
+both_flu <- comb_data(lag = 1) %>%
+    filter(season >= 2010) #%>% select(-X)
+both_flu2 <- comb_data(lag = 2) %>%
+    filter(season >= 2010) #%>% select(-X)
+
+ILINet <- get_ili_data() %>%
+    filter(season >= 2010)
+
+args <- commandArgs()
+model <- args[6]
+ili_model <- args[7]
+mod_string <- paste("../stan_models/", model, ".stan", sep = "")
+mod <- cmdstan_model(stan_file = mod_string)
+
+#dir.create(paste0("./hosp_fits/", model))
+select_regions <- unique(ILINet$region)
+#select_regions <- c("Alabama", "Alaska")#, "Idaho")
+#select_regions <- c("US")
+hosp_log <- ifelse(str_detect(model, "log"), TRUE, FALSE)
+hosp_lst <- ifelse(str_detect(model, "lst"), TRUE, FALSE)
+sqr <- ifelse(str_detect(model, "sq"), TRUE, FALSE)
+if (hosp_log == TRUE) {both_flu$value <- log(both_flu$value + 1)}
+if (hosp_log == TRUE) {both_flu2$value <- log(both_flu2$value + 1)}
+
+#select_regions <- "US"
+all_forecasts <- foreach(j = select_regions,
+	.packages = c("tidyr", "dplyr", "evalcast")
+	,.errorhandling = "remove"
+	,.combine = rbind) %:% 
+	#k <- 14
+  foreach(k = 
+	  #c(14, 20, 26)
+	  9:38
+  	, .combine = rbind) %dopar% {
+# j <- "Georgia"; k <- 21 
+       
+    both_flu_hold <- both_flu %>%
+	   filter(region == j, season == 2022 | (season == 2023 & season_week <= k)) 
+      
+    both_flu2_hold <- both_flu2 %>%
[…]

### 2025-06-30 flu_forecast_23/full_fit_hosp_li.R
add forecasts

@@ -115,6 +115,9 @@ all_forecasts <- foreach(j = select_regions,
    ili_forc <- read.csv(paste0("./ili_fits/", ili_model, "/reference_date_", date(reference_date), 
 			"_week", k, "/", j, ".csv"))
    
+   ili_forc <- ili_forc %>%
+     filter(if_all(everything(), ~ !is.na(.)))
+   
    forecasts <- data.frame()
    count_rate <- unique(both_flu_hold$count_rate2)
    count_rate_sig <- count_rate
@@ -185,6 +188,10 @@ all_forecasts <- foreach(j = select_regions,
 	}
 	forecasts <- rbind(forecasts, hosp_forecast)
    }	   
+   
+   forecasts <- forecasts %>%
+     filter(if_all(everything(), ~ !is.na(.)))
+   
    colnames(forecasts) <- paste0("ahead", -1:3)
    #print(head(forecasts))
    probs <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)

### 2025-07-01 flu_forecast_23/full_fit_hosp.R
fix forecasts

@@ -70,35 +70,35 @@ all_forecasts <- foreach(j = select_regions,
 		     sigma_nu = 15
 		 )
     location <- unique(both_flu_hold$location)
-    if (j != "US" & !(k %in% c(14, 20, 26))) {
-            samps <- mod$sample(data = stan_dat,
+    # if (j != "US" & !(k %in% c(14, 20, 26))) {
+        samps <- mod$sample(data = stan_dat,
 				chains = 1,
 				adapt_delta = .9999,
 				iter_warmup = 10000,
 				iter_sampling = 50000)
 
     	    draws <- samps$draws(format = "df")
-    }
-    if (j == "US" & k %in% c(14, 20, 26)) {
-            samps <- mod$sample(data = stan_dat,
-        			chains = 4,
-        			adapt_delta = .9999,
-        			iter_warmup = 10000,
-        			iter_sampling = 50000)
-            draws <- samps$draws(format = "df")
-
-            summary <- samps$summary()
-            diag_list <- lapply(samps$diagnostic_summary(), FUN = mean)
-            diag <- diag_list %>% data.frame()
-            diag$max_rhat <- max(summary$rhat, na.rm = TRUE)
-            diag$min_ess <- min(summary$ess_bulk, na.rm = TRUE)
-            diag$min_esst <- min(summary$ess_tail, na.rm = TRUE)
-            write.csv(diag, paste0("./hosp_fits/", ili_model, "_", model, "_", j, "_", k,
-        			   "_diagnostics.csv"), row.names = FALSE)
-
-            write.csv(draws, paste0("./hosp_fits/", ili_model, "_", model, "_", j, "_", k,
-        			    "_posterior.csv"), row.names = FALSE)
-    } 
+    # }
+    # if (j == "US" & k %in% c(14, 20, 26)) {
+    #         samps <- mod$sample(data = stan_dat,
+    #     			chains = 4,
+    #     			adapt_delta = .9999,
+    #     			iter_warmup = 10000,
+    #     			iter_sampling = 50000)
+    #         draws <- samps$draws(format = "df")
+    # 
+    #         summary <- samps$summary()
+    #         diag_list <- lapply(samps$diagnostic_summary(), FUN = mean)
+    #         diag <- diag_list %>% data.frame()
+    #         diag$max_rhat <- max(summary$rhat, na.rm = TRUE)
[…]

### 2025-07-01 flu_forecast_23/full_fit_hosp_li.R

@@ -70,7 +70,7 @@ all_forecasts <- foreach(j = select_regions,
 		     sigma_nu = 15
 		 )
     location <- unique(both_flu_hold$location)
-    if (j != "US" & !(k %in% c(14, 20, 26))) {
+    # if (j != "US" & !(k %in% c(14, 20, 26))) {
             samps <- mod$sample(data = stan_dat,
 				chains = 1,
 				adapt_delta = .9999,
@@ -78,27 +78,27 @@ all_forecasts <- foreach(j = select_regions,
 				iter_sampling = 50000)
 
     	    draws <- samps$draws(format = "df")
-    }
-    if (j == "US" & k %in% c(14, 20, 26)) {
-            samps <- mod$sample(data = stan_dat,
-        			chains = 4,
-        			adapt_delta = .9999,
-        			iter_warmup = 10000,
-        			iter_sampling = 50000)
-            draws <- samps$draws(format = "df")
-
-            summary <- samps$summary()
-            diag_list <- lapply(samps$diagnostic_summary(), FUN = mean)
-            diag <- diag_list %>% data.frame()
-            diag$max_rhat <- max(summary$rhat, na.rm = TRUE)
-            diag$min_ess <- min(summary$ess_bulk, na.rm = TRUE)
-            diag$min_esst <- min(summary$ess_tail, na.rm = TRUE)
-            write.csv(diag, paste0("./hosp_fits/", ili_model, "_", model, "_", j, "_", k,
-        			   "_diagnostics.csv"), row.names = FALSE)
-
-            write.csv(draws, paste0("./hosp_fits/", ili_model, "_", model, "_", j, "_", k,
-        			    "_posterior.csv"), row.names = FALSE)
-    } 
+    # }
+    # if (j == "US" & k %in% c(14, 20, 26)) {
+    #         samps <- mod$sample(data = stan_dat,
+    #     			chains = 4,
+    #     			adapt_delta = .9999,
+    #     			iter_warmup = 10000,
+    #     			iter_sampling = 50000)
+    #         draws <- samps$draws(format = "df")
+    # 
+    #         summary <- samps$summary()
+    #         diag_list <- lapply(samps$diagnostic_summary(), FUN = mean)
+    #         diag <- diag_list %>% data.frame()
+    #         diag$max_rhat <- max(summary$rhat, na.rm = TRUE)
+    #         diag$min_ess <- min(summary$ess_bulk, na.rm = TRUE)
[…]

### 2025-07-01 flu_forecast_23/full_fit_hosp.R
fix stuff

@@ -46,8 +46,8 @@ all_forecasts <- foreach(j = select_regions,
 	,.combine = rbind) %:% 
 	#k <- 14
   foreach(k = 
-	  #c(14, 20, 26)
-	  9:38
+	  c(14, 20, 26)
+	  #9:38
   	, .combine = rbind) %dopar% {
 # j <- "Georgia"; k <- 21 
        

### 2025-07-01 flu_forecast_23/full_fit_hosp_tog.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+#SBATCH --time=5-23:59:00
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=64
+#SBATCH --mem=0
+
+
+module load gcc
+module load r
+module load udunits
+module load proj
+module load r-rgdal
+#Rscript ./comp_fit_hosp.R "$1"
+Rscript ./full_fit_hosp.R "$1" "$2" 

### 2025-08-04 manuscript_res.tex
add some score plots

@@
         ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
-        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w
+        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w + \upsilon_{s,w}
     % \end{flalign*}
@@
 
-The term $\gamma_w$, where $w$ is the season week, is included in 
-(\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the 
-function. Modeling discrepancy has been used in uncertainty analysis of 
+Modeling discrepancy has been used in uncertainty analysis of 
 simulators to capture systematic differences between mathematical models and 
@@
 and 2016 flu seasons where their model outperformed all others in the CDC flu 
-forecasting challenge \cite[]{osthus2019dynamic}. As in Osthus's model, 
-$\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}). 
-\begin{equation}
+forecasting challenge \cite[]{osthus2019dynamic}.
+The terms $\gamma_w$ and $\upsilon_{s,w}$, where $s$ is the season and $w$ is the 
+season week, are included in 
+(\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the 
+function. Here $\gamma_w$ is shared across all seasons, whereas $\upsilon_{s,w}$
+is specific to season $s$. In fact, to assist in identifying $\gamma_w$,
+$\upsilon_{s,w}$ is only included for the current season being forecast and none of
+the previous seasons.
+ As in Osthus's model, 
+$\gamma_w$ and $\upsilon_{s,w}$ are modeled as reverse random walks, as shown 
+in (\ref{eq:rw}). 
+\begin{equation}
+    \begin{aligned}
     \label{eq:rw}
-        \gamma_w|\gamma_{w + 1} \overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
-        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W})
-\end{equation}
+        \gamma_w|\gamma_{w + 1} &\overset{ind}{\sim} N(\gamma_{w+1},\sigma^2_{\gamma}), \quad
+        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W}) \\
+        \upsilon_{s,w}|\upsilon_{s, w + 1} &\overset{ind}{\sim} N(\upsilon_{s, w+1},\sigma^2_{\upsilon})
+    \end{aligned}
+\end{equation}
+
 The idea for using the reverse random walk is that there are several previous 
@@
 $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
-identifiability.
+identifiability. An additional constraint is $\upsilon_{s, W} = 0$.
 
@@
 $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
-function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
+function $\theta_s = (\lambda_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
 For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
@@
 The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
-priors. To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
-in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the 
-parameters. A modular Bayesian approach involves multiple steps of parameter 
-fitting where some parameters may be estimated without priors via maximum 
-likelihood estimation or other means. Fitting the rest of the model parameters 
-involves assigning priors and conditioning on the previously fit parameters. 
-This has been done in computer modeling to improve identifiability and other 
-issues, though Liu et al. \cite[]{liu2009modularization} warn this approach is 
-not probabilistically sound if parameter inference is a priority 
-\cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. 
-We carried out the modular fit by first estimating the maximum likelihood 
-estimate (MLE) for the parameter $\lambda_s$ for each season in 
-(\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
+priors. 
+% To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
+% in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the 
+% parameters. A modular Bayesian approach involves multiple steps of parameter 
+% fitting where some parameters may be estimated without priors via maximum 
+% likelihood estimation or other means. Fitting the rest of the model parameters 
+% involves assigning priors and conditioning on the previously fit parameters. 
+% This has been done in computer modeling to improve identifiability and other 
+% issues, though Liu et al. \cite[]{liu2009modularization} warn this approach is 
+% not probabilistically sound if parameter inference is a priority 
+% \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. 
+% We carried out the modular fit by first estimating the maximum likelihood 
+% estimate (MLE) for the parameter $\lambda_s$ for each season in 
+% (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
 
@@
 \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The proper scoring rule primarily used in the FluSight competition and the 
-COVID-19 Forecast Hub and which we will use herein is the 
-weighted interval score (WIS)
+The proper scoring rules primarily used in the FluSight competition and the 
+COVID-19 Forecast Hub and which we will use herein are the 
+weighted interval score (WIS) and the log-weighted interval score (LWIS).
 \cite[]{mathis2024evaluation, bracher2021evaluating}.
@@
 $\vmathbb{1}\{\cdot\}$ is the indicator function.
-The goal of the forecaster is to 
-minimize the WIS.
+The LWIS is the same as the WIS except that it is evaluated 
+over the log of quantiles and the log of the observed value.
 
@@
 
+The forecasts used in this section and in the subsequent section follow the 
+FluSight convention of being quantiles corresponding to the 23 given 
+probabilities $\\(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$
+making each forecast a set of 11 predictive intervals and a predictive median.
 Figure \ref{fig:wis_by_week_horizon} shows 
-boxplots of the WIS for the 500 simulation replicates faceted by season week 
+boxplots of the WIS for the 500 simulation replicates faceted by week 
 on the y-axis and horizon on the x-axis. 
@@
 Here the ASGD model regularly outperforms
-the BASE and ARIMA models, but the three other nonlinear models do not. 
+the BASE and ARIMA models. The three other nonlinear models do not perform as
+well.
 The ASG model
@@
 is often better for forecasting than that of the three parameter SIR function.
-But the best performance comes with the combination of the ASG function and
-the discrepancy modeling.
+The model with the best performance is the ASGD.The ASGD model shows 
+especially good forecast skill at week 20 around where we would expect the 
+discrepancy to be most useful.
 
@@
     \centering
-    \includegraphics[scale=.4]{Images/wis_by_week_horizon.png}
+    \includegraphics[scale=.57]{Images/wis_by_week_horizon.png}
     \caption{Boxplots of weighted interval score (WIS) for the 
@@
 
+\begin{table}[hbt!]
+\centering
+\begin{tabular}{lrrrr}
+Model & WIS & MAE & 50\%Coverage  & 95\% Coverage \\ 
+  \hline
+ASGD & 0.09 & 0.15 & 42 & 91 \\ 
+  ASG & 0.13 & 0.18 & 33 & 71 \\ 
+  ARIMA & 0.13 & 0.18 & 35 & 68 \\ 
+  BASE & 0.15 & 0.20 & 33 & 65 \\ 
+  SIRD & 0.15 & 0.23 & 33 & 78 \\ 
+  SIR & 0.21 & 0.29 & 20 & 51 \\ 
+\end{tabular}
+\end{table}
+
 \section{Analysis of forecasts for 2023 flu season}
@@
 
-In this section we apply the forecast models to make forecasts of the 2023 
-flu season weekly hospitalizations. The scoring of the forecasts is in the 
+In this section we apply the forecast models to make forecasts for the 2023 
+flu season weekly hospitalizations and compare the results with those of 
+forecasts submitted to FluSight. The scoring of the forecasts is in the 
 context of the FluSight competition where each competing forecast was 
-submitted as a set of 23 quantiles corresponding to the given probability 
-levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$. A 
-single forecast is thus comprised of 11 predictive intervals and a median.
+submitted as a set of 23 quantiles corresponding to give probabilities.
 Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
@@
 forecasting took place during the week of October 7, 2023, and the final week 
-was the week of April 27, 2024 making 29 total weeks of forecasts.
+was the week of May 4, 2024 making 30 total weeks of forecasts.
  The same format was used during the 2021 and 2022 seasons and for the 
@@
 % Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
- Primary scores for evaluating each forecast were the WIS, 
- the log-weighted interval score (LWIS), and the relative weighted 
- interval score (RWIS). The focus in this section will be on the LWIS 
- (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
- The LWIS is the same as the WIS except that it is evaluated 
-over the log of quantiles and the log of the observed value.
-
-We fit 24 forecast models for each location for all 29 weeks, and for each 
-week forecast 1-4 week ahead hospitalizations. The 24 models included all 
-combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
-hospitalization models, and both quadratic and linear hospitalization models. 
+ Primary scores for evaluating each forecast were the WIS and LWIS
+
 The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
@@
 For the ASG model, the MLE 
-$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value.
-The remaining estimates were used as starting values for posterior sampling.
-The transformation $T(\theta_s) = (\text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ was made, and a prior distribution was assigned to $T(\theta_s)$.
+$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ 
+was calculated. % and $\hat{\lambda}_s$ was accepted as a fixed value. 
+These estimates were used as starting values for posterior sampling. 
+The transformation 
+$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ 
+was made, and a prior distribution was assigned to $T(\theta_s)$.
 The prior distributions for the ILI model under the ASG function are shown 
@@
 most parameters we have an idea what reasonable values may be. Here 
-$m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
+$m = (-4.5, 0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.3, 0.2, 5, 2, 2)$
 where $\text{diag}(\cdot)$ 
@@
         \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2)
+        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
+        \sigma_{\upsilon_{s,w}} &\sim N^+(0, .1^2)
 \end{aligned}
@@
 
-\begin{figure}[hbt!]
-\centering
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
-season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, 
-and the hospitalization models are all normally distribution. The figure 
-includes hospitalization forecasts where ILI is a linear predictor (left) 
-and where ILI is a quadratic predictor (right). 50\% predictive intervals 
-are pink and 95\% predictive intervals are red.}
-\label{fig:normal_flu_forecasts}
-\end{figure}
-
+We fit 24 forecast models for each location for all 30 weeks, and for each 
+week forecast 1-4 week ahead hospitalizations. The 24 models included all 
+combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
+hospitalization models, and both quadratic and linear hospitalization models.
 Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US 
@@
 models seem to predict well or slightly overpredict after the peak.
-% From these figures, it's difficult to say whether ASG or SIR models predict better, but at least for the quadractic models the week 22 predictions are maybe slightly better than for the linear models.
-
-Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the 
-season for all 24 models of US hospitalizations. LWIS scores are grouped by 
-ILI model, hospitalization model distribution, and by the linear or quadratic 
-modeling. Here, the smaller the LWIS the better. The weeks around the holiday 
-week 22 are highlighted by a grey band. 
-In several cases, there appears to be a turning point in performance at or 
-near week 22. For most ASG models, the models which include discrepancy tend 
-to outperform those which do not beginning near week 22 and for the rest of 
-the season.
-For the SIR models, the discrepancy does not always lead to better forecasts, 
-but in for the LNORM hospitalization models, including discrepancy greatly 
-improves forecasts around the holiday week 22. The LNORM LWIS scores tend to 
-be higher than for LST and NORM hospitalization models, but figure 
-\ref{fig:lwis_by_dist_loc} and table \ref{tab:forecast_scores} suggest that 
-the LNORM hospitalization model outperforms LST and NORM for most states and 
-territories.
+
+Of the 24 models, the one which produced the best forecasts was that using
+the ASGD ILI model and the NORM quadratic hospitalization model. The results of
+that analysis may be found in the supplementary materials ??. For the remainder 
+of this section, forecast performance from the best model, which we denote
+as ASG\_NORM2, is compared to that of forecasts from 20 models from which
+submissions were made to FluSight in 2023. 
+The models selected are those which were not ensemble models. Among the models
+was the FluSight baseline. The baseline forecast had as a median the most 
+recently observed 
+hospitalization count, and the forecast uncertainty was based on 
+differences between 
+previous hospitalizations. The baseline forecasts was
+similar to the baseline forecast used in
+previous flu forecasting seasons and in the COVID-19 hub 
+\cite[]{mathis2024evaluation, Cramer2022-hub-dataset}.
+
+Figure \ref{fig:lwis_cover_sum} shows forecast results for 2023 flu 
+hospitalization forecasts for all 21 models being compared but with results
+for the ASG\_NORM2 and FluSight baseline models highlighted. 
+For each week of the 2023 season, the  
+LWIS averaged over location and horizon 
+for each model is shown in a). Boxplots of LWIS for 1, 2, 3, and 4-week ahead
+horizons for the ASG\_NORM2 and baseline models are shown in b). And 
+overall empirical coverage of the 11 predictive intervals for all models is 
+shown in c) with ASG\_NORM2 and the baseline model again highlighted. 
+These plots show that over the season the ASG\_NORM2 model tended to 
+outperform most others in terms of LWIS, and in general is superior to the 
+baseline model. The coverage plot shows that the ASG\_NORM2 model is well 
+calibrated at most levels, with slight overcoverage at the lower levels
+up to about 80\%, and has better coverage than all other models. Similar results
+for the WIS are shown in the supplementary materials ??.
 
@@
     % \centering
-    \includegraphics[scale = .6]{Images/lwis_ln_us_full_season.png}
-    \caption{Each plot shows the log weighted interval score (LWIS) for every 
-    week of the 2023 flu season with scores for models including and exluding 
-    discrepancy in the ILI model. Scores are separated by hospitalization 
-    distribution family and by ILI as a linear or quadratic predictor. Scores 
-    for models with an ASG ILI model are above while those with an SIR model 
-    are below. The lower the LWIS the better the forecast.}
-    \label{fig:us_lwis}
+    \includegraphics[scale = .6]{Images/lwis_cover_sum.png}
+    \caption{a) Averaged over location and horizon log-weighted interval score
+    (LWIS) for each week during the season for the 21 models being compared. 
+    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
+    FluSight-baseline (red) are colored to stand out.
+    b) Boxplots of all LWIS scores for 1, 2, 3, and 4-week ahead horizons 
+    for the ASG\_NORM2 (blue) and FluSight-baseline (red) models.
+    c) Plot of empirical coverage against theoretical coverage for 11 
+    predictive intervals from the 10\% nominal level to the 98\%. The black
+    line $x = y$ shows where optimal coverage would fall.
+    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
+    FluSight-baseline (red) are colored to stand out.}
+    \label{fig:lwis_cover_sum}
 \end{figure}
 
-% \begin{figure}
-%     
-%     % \centering
-%     \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
-%     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
-Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
-Scores for models with an ASG ILI model are above while those with an SIR model are below.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-%     \label{fig:us_lwis}
-% \end{figure}
+Figure \ref{fig:state_and_date_lwis} shows the average LWIS for each model
+relative to the ASG\_NORM2 model by location in the top tile plot and by 
+season week in the bottom tile plot. That is, the value for each tile is the 
+ratio of the average ASG\_NORM2 LWIS to the average for the labeled model at the
+labeled location or week. A value less than 1 shows superior performance by
+ASG\_NORM2 with increasing superiority as the value goes to 0. A value larger
+than 1 shows inferior performance, and a value of 1 shows equal performance.
+Tiles are grey where the labeled forecast model made no forecast submissions.
+Viewed this way, only UGA\_flucast-INFLAenza appears to potentially outperform
+ASG\_NORM2.
 
 \begin{figure}[hbt!]
-
+\begin{adjustwidth}{-3cm}{-1.5cm}
+    % \begin{minipage}{\linewidth}
+    % \centering
+    \makebox[\linewidth][l]{
+    \includegraphics[scale = .45]{Images/state_and_date_lwis.png}
+    }
+    \end{adjustwidth}
+    \caption{}
+    \label{fig:state_and_date_lwis}
+    % \end{minipage}
+
+\end{figure}
+
+\begin{table}[ht]
 \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 
-    US states, PR, DC, and national level forecasts at each week during the 
-    2023 flu season. Scores are averaged over all horizons 1-4 weeks ahead. 
-    Scores are faceted by hospitalization model distribution. The lighter the 
-    shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:lwis_by_dist_loc}
-\end{figure}
-
-Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by 
-LWIS, but all 53 locations are included. The LWIS is shown across the whole 
-season faceted by hospitalization model distribution. Overall, the weeks 
-leading up to week 22 are the most difficult to forecast, but the LNORM model 
-appears to perform the better than the other models during the weeks around 
-the holiday week 22. 
-The overall, calculated as the mean LWIS over all locations and weeks, for all 
-24 models is shown in table \ref{tab:forecast_scores}. The first main takeaway 
-is that the top four performing models are SIR ILI models with the LNORM 
-hospitalization models. The next main takeaway is that overall the models 
-which include discrepancy outperformed the models without discrepancy.
-
-\begin{table}
-\caption{Overall scores for each of the 24 forecast models. The overall score 
-is the log weighted interval score (LWIS) averaged over all locations, weeks, 
-and horizons. The scores in the first two rows are for linear models, and the 
-scores in the third and fourth rows are for quadratic models. The lowest WISs 
-are bolded.}
-\begin{tabular*}{\textwidth}
-{@{\extracolsep{\fill}} 
-    l*{9}{c}}
-  & & \multicolumn{3}{c}{ASG} 
-  & \multicolumn{3}{c}{SIR} \\ 
-  \cmidrule{3-5} \cmidrule{6-8}
-  & & LNORM & LST & NORM & LNORM & LST & NORM\\
-  \midrule
-  Linear & No Disc & 0.366 & 0.392 & 0.365 & \textbf{0.355} & 0.416 & 0.394 &\\ 
-   & Disc & 0.358 & 0.391 & 0.365 & \textbf{0.343} & 0.382 & 0.354 &\\
-  \midrule
-  Quadratic & No Disc & 0.366 & 0.398 & 0.377 & \textbf{0.355} & 0.401 & 0.377 &\\ 
-   & Disc & 0.358 & 0.390 & 0.369 & \textbf{0.343} & 0.378 & 0.356 &\\     
-  \bottomrule
-\end{tabular*}
-\label{tab:forecast_scores}
+\begin{adjustwidth}{-3cm}{-1.5cm}
+\begin{tabular}{lrrrrrrrr}
+  % \hline
+ & MWIS & MLWIS & MedWIS & MedLWIS & MSEC & \% Forcs & RWIS & RLWIS\\ 
+  \hline
+ASG & 62 & 0.34 & 14 & 0.23 & 0.001 & 100 & 1 & 1 \\ 
+  PSI-PROF & 63 & 0.35 & 15 & 0.23 & 0.003 & 100 & 0.90 & 0.97 \\ 
+  CEPH-Rtrend\_fluH & 69 & 0.39 & 13 & 0.28 & 0.004 & 100 & 0.81 & 0.86 \\ 
+  FluSight-baseline & 85 & 0.41 & 16 & 0.32 & 0.023 & 100 & 0.63 & 0.82 \\ 
+  UM-DeepOutbreak & 83 & 0.46 & 21 & 0.39 & 0.010 & 99 & 0.64 & 0.72 \\ 
+  cfarenewal-cfaepimlight & 76 & 0.37 & 15 & 0.26 & 0.001 & 99 & 0.72 & 0.92 \\ 
+  MOBS-GLEAM\_FLUH & 66 & 0.35 & 15 & 0.22 & 0.021 & 98 & 0.87 & 0.95 \\ 
+  LosAlamos\_NAU-CModel\_Flu & 242 & 0.64 & 24 & 0.42 & 0.153 & 93 & 0.21 & 0.53 \\ 
+  UGuelph-CompositeCurve & 132 & 0.52 & 25 & 0.41 & 0.018 & 93 & 0.43 & 0.62 \\ 
+  LUcompUncertLab-chimera & 84 & 0.43 & 17 & 0.33 & 0.004 & 92 & 0.67 & 0.71 \\ 
+  UGA\_flucast-INFLAenza & 74 & 0.32 & 15 & 0.20 & 0.003 & 90 & 0.77 & 1.07 \\ 
+  cfa-flumech & 101 & 0.41 & 17 & 0.27 & 0.019 & 89 & 0.58 & 0.85 \\ 
+  UGA\_flucast-Copycat & 76 & 0.35 & 16 & 0.21 & 0.012 & 87 & 0.81 & 0.98 \\ 
+  GT-FluFNP & 105 & 0.48 & 18 & 0.38 & 0.046 & 86 & 0.45 & 0.73 \\ 
+  UNC\_IDD-InfluPaint & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
+  NU\_UCSD-GLEAM\_AI\_FLUH & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
+  SigSci-CREG & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
+  PSI-PROF\_beta & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
+  NIH-Flu\_ARIMA & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
+  GH-model & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
+  UGA\_flucast-OKeeffe & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
+\end{tabular}
+\end{adjustwidth}
 \end{table}

### 2025-08-04 make_images/make_final_analysis_images.R

@@ -1,10 +1,24 @@
-source("./flu_forecast_23/get_data_functions_new.R")
-team_scores <- readRDS("./flu_forecast_23/hub_scores_no_ens.rds")
-ili_scores <- readRDS("./flu_forecast_23/nl_scores.rds")
+source("../flu_forecast_23/get_data_functions_new.R")
+library(datasets)
+state.name <- c(state.name, "Puerto Rico", "District of Columbia", "US")
+state.abb <- c(state.abb, "PR", "DC", "US")
+states <- data.frame(loc_abb = state.abb, location_name = state.name)
+
+
+team_scores <- readRDS("../flu_forecast_23/hub_scores_no_ens.rds") %>% 
+  left_join(states, by = "location_name") %>% 
+  mutate(model = ifelse(model == "FluSight-baseline", 
+                        # "<b><i>FluSight-baseline</i></b>", 
+                        "<span style='font-size:12pt'><b><i>FluSight-baseline</i></b></span>",
+                        model))
+ili_scores <- readRDS("../flu_forecast_23/nl_scores.rds") %>% 
+  left_join(states, by = "location_name")
+  
+
 
 rbind(team_scores, ili_scores) %>% 
-  filter(str_detect(model, "asg")) %>%
-  filter(location_name != "Florida") %>% 
+  # filter(str_detect(model, "asg")) %>%
+  # filter(location_name != "Florida") %>% 
   group_by(model) %>% 
   summarise(m = mean(lwis, na.rm = TRUE),
             n = n()) %>% 
@@ -14,71 +28,429 @@ rbind(team_scores, ili_scores) %>%
 rbind(team_scores, ili_scores) %>% 
   filter(model %in% c("asg_disc2_nm_hosp_sq_ar1", "FluSight-baseline")) %>% 
   ggplot() +
-  geom_boxplot(aes(x = factor(horizon), y = wis, colour = model)) +
+  geom_boxplot(aes(x = factor(horizon), y = lwis, colour = model)) #+
   scale_y_log10() 
 
-team_scores <- rbind(team_scores, ili_scores %>% filter(model == "asg_disc2_nm_hosp_sq_ar1"))
+team_scores <- rbind(team_scores, 
+                     ili_scores %>% 
+                       filter(model == "asg_disc2_nm_hosp_sq_ar1")) 
 
 best_nl <- ili_scores %>% 
   filter(model == "asg_disc2_nm_hosp_sq_ar1") %>% 
[…]

### 2025-08-04 make_images/make_scores_comp.R

@@ -237,23 +237,97 @@ team_coverage %>%
   stat_function(fun = function(x) x + 0, color = "blue", size = 1.2)
   
 
-
+team_scores <- readRDS("~/flu_research/nonlinear_flu_forecast/flu_forecast_23/hub_scores_no_ens.rds")
 
 setwd("../../../flu_research/nonlinear_flu_forecast/flu_forecast_23/fin_hosp_forecasts/")
 
 folders <- list.files("./")
-folders <- folders[!str_detect(folders, "full") & 
+folders <- folders[!str_detect(folders, "full") &
                    !str_detect(folders, "_disc_")]
-folders <- folders[!str_detect(folders, "disc2") | 
-                     str_detect(folders, "disc2_nm")]
+# folders <- folders[!str_detect(folders, "disc2") |
+#                      str_detect(folders, "disc2_nm")]
 
+folders <- folders[(str_detect(folders, "asg_nm")
+                   | str_detect(folders, "asg_disc2_nm")
+                   | str_detect(folders, "sir_hosp")
+                   | str_detect(folders, "sir_disc2"))
+                   & !str_detect(folders, "logi")
+                   & !str_detect(folders, "tr")]
+
+# folders <- folders[str_detect(folders, "asg_disc2_ho") & !str_detect(folders, "tr")]
+# folders <- folders[str_detect(folders, "asg_disc2_nm") & !str_detect(folders, "tr")]
 models <- folders
+# models <- "asg_disc2_nm_hosp_lst_ar1"
+# most_coverage <- ili_coverage
+ili_coverage <- data.frame()
 
+for (sub_date in sub_dates) {
+  for (i in 1:length(models)) {
+    forecast_file <- paste(models[i], "/", sub_date, "-",
+                           models[i], ".csv", sep = "")
+    
+    if (file.exists(forecast_file) == FALSE) next
+    forecast <- read.csv(forecast_file) %>%
+      mutate(location = as.character(location),
+             output_type_id = as.character(output_type_id)) %>%
+      mutate(location = ifelse(nchar(location) < 2, paste0("0",location),
+                               location)) %>%
+      mutate(value = ifelse(value < 0, 0, value)) %>%
+      filter(output_type == "quantile") %>%
+      mutate(location = as.character(location)) %>%
[…]

### 2025-08-06 manuscript_res.tex
most recent plotting updates

@@
 \begin{figure}[hbt!]
-    \centering
-    \includegraphics[scale=.57]{Images/wis_by_week_horizon.png}
+    % \centering
+    \includegraphics[scale=.5]{Images/all_sim_res.png}
     \caption{Boxplots of weighted interval score (WIS) for the 
@@
 
-\begin{table}[hbt!]
+% \begin{table}[hbt!]
+% \centering
+% \begin{tabular}{lrrrr}
+% Model & WIS & MAE & 50\%Coverage  & 95\% Coverage \\ 
+%   \hline
+% ASGD & 0.09 & 0.15 & 42 & 91 \\ 
+%   ASG & 0.13 & 0.18 & 33 & 71 \\ 
+%   ARIMA & 0.13 & 0.18 & 35 & 68 \\ 
+%   BASE & 0.15 & 0.20 & 33 & 65 \\ 
+%   SIRD & 0.15 & 0.23 & 33 & 78 \\ 
+%   SIR & 0.21 & 0.29 & 20 & 51 \\ 
+% \end{tabular}
+% \end{table}
+
+\begin{table}[ht]
 \centering
-\begin{tabular}{lrrrr}
-Model & WIS & MAE & 50\%Coverage  & 95\% Coverage \\ 
+\begin{tabular}{lrrr}
+Model & WIS & 50\%Coverage  & 95\% Coverage \\ 
   \hline
-ASGD & 0.09 & 0.15 & 42 & 91 \\ 
-  ASG & 0.13 & 0.18 & 33 & 71 \\ 
-  ARIMA & 0.13 & 0.18 & 35 & 68 \\ 
-  BASE & 0.15 & 0.20 & 33 & 65 \\ 
-  SIRD & 0.15 & 0.23 & 33 & 78 \\ 
-  SIR & 0.21 & 0.29 & 20 & 51 \\ 
+ASGD & 0.02 & 0.42 & 0.91 \\ 
+  ASG & 0.03 & 0.33 & 0.71 \\ 
+  ARIMA & 0.03 & 0.35 & 0.68 \\ 
+  BASE & 0.03 & 0.33 & 0.65 \\ 
+  SIRD & 0.03 & 0.33 & 0.78 \\ 
+  SIR & 0.05 & 0.20 & 0.51 \\
 \end{tabular}
@@
 
+\begin{figure}[hbt!]
+
+  \centering
+  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
+\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
+season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, 
+and the hospitalization models are all normally distribution. The figure 
+includes hospitalization forecasts where ILI is a linear predictor (left) 
+and where ILI is a quadratic predictor (right). 50\% predictive intervals 
+are pink and 95\% predictive intervals are red.}
+\label{fig:normal_flu_forecasts}
+\end{figure}
+
 Of the 24 models, the one which produced the best forecasts was that using
@@
     % \begin{minipage}{\linewidth}
-    % \centering
+    \centering
     \makebox[\linewidth][l]{

### 2025-08-06 supplemental_res.tex
most recent plotting updates

--- 2024-12-13 supplemental.tex
@@
 \begin{supplement}
+\renewcommand{\thesection}{\Alph{section}}
 
@@
 \newpage
+\section{Model forecast comparison}
+
+\begin{figure}[hbt!]
+\centering
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
+  % \caption{A subfigure}
+  % \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}
+  \centering
+  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
+  % \caption{A subfigure}
+  % \label{fig:sub2}
+\end{subfigure}
+\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
+season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, 
+and the hospitalization models are all normally distribution. The figure 
+includes hospitalization forecasts where ILI is a linear predictor (left) 
+and where ILI is a quadratic predictor (right). 50\% predictive intervals 
+are pink and 95\% predictive intervals are red.}
+\label{fig:normal_flu_forecasts}
+\end{figure}
+
+\begin{figure}[hbt!]
+
+    % \centering
+    \includegraphics[scale = .6]{Images/lwis_ln_us_full_season.png}
+    \caption{Each plot shows the log weighted interval score (LWIS) for every 
+    week of the 2023 flu season with scores for models including and exluding 
+    discrepancy in the ILI model. Scores are separated by hospitalization 
+    distribution family and by ILI as a linear or quadratic predictor. Scores 
+    for models with an ASG ILI model are above while those with an SIR model 
+    are below. The lower the LWIS the better the forecast.}
+    \label{fig:us_lwis}
+\end{figure}
+
+% \begin{figure}
+%     
+%     % \centering
+%     \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
+%     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
+Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
+Scores for models with an ASG ILI model are above while those with an SIR model are below.
+The lighter the shade, the lower the LWIS with low LWIS being better.}
+%     \label{fig:us_lwis}
+% \end{figure}
+
+\begin{figure}[hbt!]
+
+    \centering
+    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
+    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 
+    US states, PR, DC, and national level forecasts at each week during the 
+    2023 flu season. Scores are averaged over all horizons 1-4 weeks ahead. 
+    Scores are faceted by hospitalization model distribution. The lighter the 
+    shade, the lower the LWIS with low LWIS being better.}
+    \label{fig:lwis_by_dist_loc}
+\end{figure}
+
+Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by 
+LWIS, but all 53 locations are included. The LWIS is shown across the whole 
+season faceted by hospitalization model distribution. Overall, the weeks 
+leading up to week 22 are the most difficult to forecast, but the LNORM model 
+appears to perform the better than the other models during the weeks around 
+the holiday week 22. 
+The overall, calculated as the mean LWIS over all locations and weeks, for all 
+24 models is shown in table \ref{tab:forecast_scores}. The first main takeaway 
+is that the top four performing models are SIR ILI models with the LNORM 
+hospitalization models. The next main takeaway is that overall the models 
+which include discrepancy outperformed the models without discrepancy.
+
+\begin{table}
+\caption{Overall scores for each of the 24 forecast models. The overall score 
+is the log weighted interval score (LWIS) averaged over all locations, weeks, 
+and horizons. The scores in the first two rows are for linear models, and the 
+scores in the third and fourth rows are for quadratic models. The lowest WISs 
+are bolded.}
+\begin{tabular*}{\textwidth}
+{@{\extracolsep{\fill}} 
+    l*{9}{c}}
+  & & \multicolumn{3}{c}{ASG} 
+  & \multicolumn{3}{c}{SIR} \\ 
+  \cmidrule{3-5} \cmidrule{6-8}
+  & & LNORM & LST & NORM & LNORM & LST & NORM\\
+  \midrule
+  Linear & No Disc & 0.366 & 0.392 & 0.365 & \textbf{0.355} & 0.416 & 0.394 &\\ 
+   & Disc & 0.358 & 0.391 & 0.365 & \textbf{0.343} & 0.382 & 0.354 &\\
+  \midrule
+  Quadratic & No Disc & 0.366 & 0.398 & 0.377 & \textbf{0.355} & 0.401 & 0.377 &\\ 
+   & Disc & 0.358 & 0.390 & 0.369 & \textbf{0.343} & 0.378 & 0.356 &\\     
+  \bottomrule
+\end{tabular*}
+\label{tab:forecast_scores}
+\end{table}
+
 \section{FluSight forecast competition scoring results} \label{app:A_rwis}

### 2025-08-06 make_images/make_final_analysis_images.R

@@ -1,5 +1,6 @@
 source("../flu_forecast_23/get_data_functions_new.R")
 library(datasets)
+library(ggtext)
 state.name <- c(state.name, "Puerto Rico", "District of Columbia", "US")
 state.abb <- c(state.abb, "PR", "DC", "US")
 states <- data.frame(loc_abb = state.abb, location_name = state.name)
@@ -10,9 +11,11 @@ team_scores <- readRDS("../flu_forecast_23/hub_scores_no_ens.rds") %>%
   mutate(model = ifelse(model == "FluSight-baseline", 
                         # "<b><i>FluSight-baseline</i></b>", 
                         "<span style='font-size:12pt'><b><i>FluSight-baseline</i></b></span>",
-                        model))
+                        model)) %>% 
+  unique()
 ili_scores <- readRDS("../flu_forecast_23/nl_scores.rds") %>% 
-  left_join(states, by = "location_name")
+  left_join(states, by = "location_name") %>% 
+  unique()
   
 
 
@@ -33,7 +36,8 @@ rbind(team_scores, ili_scores) %>%
 
 team_scores <- rbind(team_scores, 
                      ili_scores %>% 
-                       filter(model == "asg_disc2_nm_hosp_sq_ar1")) 
+                       filter(model == "asg_disc2_nm_hosp_sq_ar1")) %>% 
+  unique()
 
 best_nl <- ili_scores %>% 
   filter(model == "asg_disc2_nm_hosp_sq_ar1") %>% 
@@ -49,7 +53,7 @@ team_scores %>%
   # filter(model == "FluSight-baseline") %>%
   ggplot() +
   geom_point(aes(x = nl_mwis, y = mwis, colour = model)) +
-  stat_function(fun = function(x) x + 0, color = "blue") +
+  stat_function(fun = function(x) x + 0, color = "blue3") +
   scale_y_log10() +
   scale_x_log10() +
   theme(legend.position = "none")
@@ -117,17 +121,16 @@ state_ldiff <- team_diffs %>%
                 x = loc_abb,
                 fill = sldiff)) +
   scale_fill_gradient2(
-    low = "royalblue", high = "red", mid = "white",
+    low = "blue3", high = "red1", mid = "white",
     midpoint = 1) +
   scale_x_discrete(expand = c(0,0)) +
   scale_y_discrete(expand = c(0,0)) +
-  xlab("") +
-  ylab("LWIS") +
+  xlab("Location") +
+  ylab("") +
[…]

### 2025-08-06 make_images/make_forecast_plots.R

@@ -1,15 +1,15 @@
 library(gridExtra)
-setwd("~/flu_research/Prelim Content/models")
-source("~/flu_research/nonlinear_flu_forecast/forecasts_23/get_data_functions.R")
+setwd("~/flu_research/nonlinear_flu_forecast/flu_forecast_23/fin_hosp_forecasts/")
+source("~/flu_research/nonlinear_flu_forecast/flu_forecast_23/get_data_functions_new.R")
 files <- list.files("./")
 
 state <- "US"
 
 
-# ILINet <- get_ili_data()
-# both_flu <- comb_data(lag = 2)
+ILINet <- get_ili_data()
+both_flu <- comb_data(lag = 1)
 
-plot_dates <- both_flu %>% ungroup() %>% filter(season_week %in% seq(14, 38, by = 6),
+plot_dates <- both_flu %>% ungroup() %>% filter(season_week %in% seq(13, 37, by = 6),
                                   season == 2023) %>%
   select(date) %>% unique() %>% unlist()
 
@@ -19,11 +19,14 @@ locations <- both_flu %>% ungroup() %>%
 
 #everything below can be changed for each image
 #uncomment everything above when first running the progam
-model <- "asg_hosp_ar1"
-models <- c("asg_hosp_ar1", "asg_disc_hosp_ar1", 
-            "sir_hosp_ar1", "sir_disc_hosp_ar1")
+model <- "asg_disc2_nm_hosp_ar1"
+models <- c("asg_nm_hosp_sq_ar1", "asg_disc2_nm_hosp_sq_ar1", 
+            "sir_hosp_sq_ar1", "sir_disc2_hosp_sq_ar1")
 # models <- "FluSight-baseline"
 labels <- c("ASG", "ASGD", "SIR", "SIRD")
+labels <- factor(labels, levels = c("ASGD", "ASG", "SIRD", "SIR"))
+# models <- "asg_disc2_nm_hosp_sq_ar1"
+# lables <- "ASGD"
 # labels <- "SIRD"
 # models <- c("sir_hosp_log_ar1", "sir_disc_hosp_log_ar1")
 # models <- "FluSight-baseline"
@@ -188,7 +191,8 @@ forecast5 <- forecast5 %>%
 
 both_flu %>% 
   # filter(location_name != "US") %>% 
-  filter(season == 2023, location_name %in% state, date > "2023-09-30") %>% 
+  filter(season == 2023, location_name %in% state, date > "2023-10-07",
[…]

### 2025-08-06 make_images/make_scores_comp.R

@@ -384,14 +384,15 @@ for (sub_date in sub_dates) {
 }
 
 # saveRDS(ili_scores, "~/flu_research/nonlinear_flu_forecast/flu_forecast_23/nl_scores.rds")
-# rbind(ili_scores, team_scores) %>% 
-ili_scores %>% 
+rbind(ili_scores, team_scores) %>%
+# ili_scores %>% 
+  filter(location_name == "US") %>% 
   group_by(model) %>% 
   filter(horizon != -1) %>% 
   unique() %>% 
   filter(!str_detect(model, "tr_")) %>% #, location_name != "Florida") %>% 
   filter(!is.na(wis)) %>% 
-  summarise(m = mean(wis, na.rm = TRUE),
+  summarise(m = mean(lwis, na.rm = TRUE),
             n = n()) %>%
   # filter(str_detect(model, "asg_nm")) %>%
   arrange(m) %>% 

### 2025-08-06 make_images/make_simulation_images.R

@@ -2,6 +2,7 @@ library(ggplot2)
 library(dplyr)
 library(stringr)
 library(ggforce)
+library(tidyr)
 
 
 
@@ -22,31 +23,140 @@ for (p in 1:length(paths)) {
 }
 
 
-scores %>% 
-  mutate(model = factor(model, levels = c("ASGD", "ASG", "ARIMA", 
-                                   "BASE", "SIRD", "SIR"))) %>% 
-  mutate(week_ahead = week_ahead + 1) %>% 
-  mutate(week_ahead = ifelse(week_ahead == 1, paste(week_ahead, 
-                                                     "week", sep = " "),
-                              paste(week_ahead, "weeks", sep = " "))) %>% 
-  ggplot() +
-  geom_boxplot(aes(x = model, y = wis)) +
-  ggh4x::facet_grid2(last_week~week_ahead, 
-                     scales = "free_y", independent = "y") +
-  ylab("WIS") +
-  xlab("Model") +
+scores <- scores %>% 
+  mutate(model = factor(model, levels = c("ASGD", "SIRD", "ASG", "SIR",
+                                          "ARIMA", "BASE"))) %>% 
+  mutate(week_ahead = factor(week_ahead + 1))# %>% 
+  # mutate(week_ahead = ifelse(week_ahead == 1, paste(week_ahead, 
+  #                                                    "week", sep = " "),
+  #                             paste(week_ahead, "weeks", sep = " "))) 
+
+
+sim_cov <- scores %>% 
+  select(model, contains("cover")) %>% 
+  pivot_longer(contains("cover"), , names_to = "coverage", 
+               values_to = "covered") %>% 
+  mutate(coverage = .01*as.numeric(str_extract(coverage, "\\d+"))) %>% 
+  group_by(model, coverage) %>% 
+  summarise(pcover = mean(covered)) %>% 
+  # filter(model == "FluSight-baseline") %>% 
+  ggplot(aes(x = coverage, y = pcover, linetype = model, colour = model)) +
+  geom_line(size = 1.4) +
+  geom_abline(intercept = 0, slope = 1, size = 1.2) +
+  scale_colour_manual(values = c("ASGD" = "#E69F00",  # orange
+                                 "BASE" = "#E66100",
+                                 "ASG" = "#56B4E9",  # sky blue
+                                 "ARIMA" = "#009E73",  # bluish green
+                                 "SIR" = "#D6C840",  # yellow
+                                 "SIRD" = "#0072B2"   # blue
+  )) +
[…]

### 2025-08-07 make_images/make_simulation_images.R
fix sim images

@@ -3,7 +3,7 @@ library(dplyr)
 library(stringr)
 library(ggforce)
 library(tidyr)
-
+#make quick fix
 
 
 score_loc <- "../simulation/hosp_sim/scores_"

### 2025-08-07 make_images/make_simulation_images.R
delete old file

@@ -1,198 +1,0 @@
-library(ggplot2)
-library(dplyr)
-library(stringr)
-library(ggforce)
-library(tidyr)
-#make quick fix
-
-
-score_loc <- "../simulation/hosp_sim/scores_"
-models <- c("arima_full", "base_full", "asg_nm", 
-            "asg_disc2_nm", "sir", "sir_disc2")
-
-
-paths <- paste0(score_loc, models, ".csv")
-names <- c("ARIMA", "BASE", "ASG", "ASGD", "SIR", "SIRD")
-
-scores <- data.frame()
-for (p in 1:length(paths)) {
-  score <- read.csv(paths[p])
-  score$model <- names[p]
-  
-  scores <- rbind(scores, score)
-}
-
-
-scores <- scores %>% 
-  mutate(model = factor(model, levels = c("ASGD", "SIRD", "ASG", "SIR",
-                                          "ARIMA", "BASE"))) %>% 
-  mutate(week_ahead = factor(week_ahead + 1))# %>% 
-  # mutate(week_ahead = ifelse(week_ahead == 1, paste(week_ahead, 
-  #                                                    "week", sep = " "),
-  #                             paste(week_ahead, "weeks", sep = " "))) 
-
-
-sim_cov <- scores %>% 
-  select(model, contains("cover")) %>% 
-  pivot_longer(contains("cover"), , names_to = "coverage", 
-               values_to = "covered") %>% 
-  mutate(coverage = .01*as.numeric(str_extract(coverage, "\\d+"))) %>% 
-  group_by(model, coverage) %>% 
-  summarise(pcover = mean(covered)) %>% 
-  # filter(model == "FluSight-baseline") %>% 
-  ggplot(aes(x = coverage, y = pcover, linetype = model, colour = model)) +
-  geom_line(size = 1.4) +
-  geom_abline(intercept = 0, slope = 1, size = 1.2) +
-  scale_colour_manual(values = c("ASGD" = "#E69F00",  # orange
-                                 "BASE" = "#E66100",
-                                 "ASG" = "#56B4E9",  # sky blue
-                                 "ARIMA" = "#009E73",  # bluish green
-                                 "SIR" = "#D6C840",  # yellow
-                                 "SIRD" = "#0072B2"   # blue
-  )) +
-  scale_linetype_manual(values = c("ASGD" = "solid",  # orange
-                                   "BASE" = "12345678",
-                                   "ASG" = "longdash",  # sky blue
[…]

### 2025-08-08 response.tex
finish responses to reviewers

--- 2025-04-22 response.tex
@@
 \usepackage{hyperref}
+\usepackage{amsmath}
+% \usepackage{amsfonts}
+
+\usepackage{xcolor}
+\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
+\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}
 
@@
 This may be a useful reference.
+
+\spencer{We appreciated this comment and the included reference to the Bayesian
+cut. This led to several new insights. 
+We added a couple paragraphs of discussion to address this. They appear in
+the revised manuscript as below.
+
+``This two-component modeling is remeniscent of the "Bayesian cut" as 
+described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
+into two component models or "modules" may be justfied for several reasons
+including to reduce time to fit the model, avoiding mixing issues of posterior
+sampling distributions, improving predictive performance or to prevent
+model misspecification in one model to influence the other 
+\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, plummer2015cuts}.
+
+The decision to cut the model into two components was made for several reasons.
+When fitting the fully Bayesian joint model, the time to fit a single model using
+Markov chain Monte Carlo (MCMC) sampling and
+obtain forecasts was sometimes upwards of 4 hours whereas fitting the two models
+separately usually took less than 1 hour. When assessing convergence of model
+parameters, it was rare that posterior sampling chains for the joint model
+showed convergence even after 55,000 iterations, whereas the component model 
+fits gave no signs which raised concern for convergence. And most importantly,
+the forecasts from the component models generally outperformed those of the 
+joing model in terms of minimizing proper scoring rules. 
+The process for obtaining forecasts is described in the following paragraphs 
+and computation is described in section \ref{sec:implementation_posterior}."}
+
+% \spencer{In the forecasting competition, the data comes generally by the afternoon and
+% forecasts are due by the evening. We initially did a full Bayesian procedure
+% (minus the MLE selection of the $\lambda$ parameter in the ASG ILI model), but fit 
+% was very slow and made it difficutl to get forecasts submitted in time. We switched
+% to using two separate models because of the time saved in model fitting. We also
+% found that in terms of the proper scoring rules and predictive interval coverage,
+% there wasn't a drop in performance although the parameter estimation did 
+% change.
+% 
+% \begin{equation}
+% \begin{aligned}
+%     p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}) &= \int 
+%        p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) 
+%        p(\boldsymbol{\alpha} | \textbf{ILI}, \textbf{H}) d\boldsymbol{\alpha} \\
+%     p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) &= 
+%        \int p(\tilde{H}_{s,w^*} | \textbf{H}, \textbf{ILI}, 
+%        \widetilde{ILI}_{s,w^*}, \boldsymbol{\alpha}) 
+%        p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) d\widetilde{ILI}_{s,w^*}
+% \end{aligned}
+% \end{equation}
+% }
+
 \item The authors employ a "modular Bayesian approach" mixing maximum likelihood and
@@
 this type of procedure (is "modular Bayesian" a standard term?).
+
+\spencer{In the revised manuscript we no longer use the modular approach
+where some parameters are set as their MLEs before fitting a Bayesian 
+model. Instead
+we assign a strong prior to the parameter which was giving us trouble
+and fit a full Bayesian model. This is
+described in the following statement added to revised manuscript:
+
+``In a previous version of this manuscript, when fitting an ASG model
+with discrepancy components,
+rather than estimating all parameters in a Bayesian way,
+identifiability was encouraged by setting
+the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})
+to be equal to its MLE. In this updated manuscript, 
+however, we use empirical Bayes and center the first component of the 
+hierarichal prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
+all seasons and set the variance parameter so the prior was tight around the
+mean. This led to improved mixing of posterior draws and better forecasts."}
+
 \item Also, it should be argued explicitly why this approach was chosen over a standard fully
@@
 difficulties were encountered?
+
+\spencer{See response to comment 1b.}
+
 \item Notably it would be helpful to comment on how this approach affects the uncertainty
 propagation / overall quantification of forecast uncertainty.
+
+\spencer{See response to comment 1b.}
 
@@
 
+\spencer{Obviously the SIR dynamics are not a perfect match for ILI data, but
+the disease transmission mechanism of influenza is at least partially 
+responsible for ILI. We touch on this a little more directly in the revised 
+manuscript by stating:
+
+``The SIR is a mechanistic model used to capture transmission of a particular
+disease, and the ASG is a smooth function which may approximate the 
+same shape as the ILI trajectory. As ILI may be influenced by more 
+diseases than influenza, the disease transmission described by the SIR model
+can only partially contribute to the true ILI data-generating mechanism[44].
+Though the ASG function may be able to roughly capture ILI, there may still be 
+systematic behavior which it cannot capture. Thus when using either the SIR
+or ASG models, an additional component for modeling discrepancy should be 
+included."
+
+We cite Osthus et al. 2019, who also touch on this.}
+
 \item My impression is that the SIR does not really have any mechanistic meaning here and is just a
@@
 practical differences exist between the two parameterizations?
+
+\spencer{Practially, the reason to use the SIR model is generally because it is meant
+to capture some of the scientific understanding of a disease outbreak. Again, we argue 
+that the ASG is a reasonable alternative to SIR models in the case of the seasonal flu
+outbreak. The ASG shape with a rise to a peak and a subsequent fall is similar in shape
+to the regular ILI trajectory. It has also been used successfully in forecasting of 
+other pheneomena with similar rises and falls (i.e. crop moisture)}.
 
@@
 
+\spencer{Yes, the hospitalization indicator is specific to influenza. See 
+response to the next comment (2d.) for how it was addressed in the manuscript.
+
+For your second point on how well ILI can be a helpful predictor of hospitalizations,
+we addressed this in the manuscript by adding the following discussion about
+figure 6. See also
+
+``Figure 6 helps
+illustrate why this modeling decision was made.
+The figure shows scatterplots of ILI percentage and 
+hospitalizations on the top row.
+The bottom row shows scatterplots of the difference between hospitalizations and 
+1 week lags scaled 
+by ILI percentage at the US national level. The plots in the right column
+are of the logarithm of hospitalizations. The lines through the points are
+drawn by connecting hospitalizations for consecutive weeks during the flu 
+season.
+These plots show that a linear or quadratic relationship between ILI percentage 
+and hospitalizations does not capture much of the temporal correlation in 
+hospitalizations as shown by the obvious loops drawn in the plot. 
+However, when including a 
+single autoregressive lag of hospitalizations in the model, much of the time 
+correlation is
+accounted for as shown by the tighter and more frequent crossing of lines
+in the bottom row. Higher than order 1 autoregressions showed little or no 
+improvement over order 1, hence we consider the ARX(1) reasonably well 
+suited for modeling ILI and hospitalzations."}
+
 \item Does the hospitalization indicator refer to incident (new) hospitalization or prevalence, i.e.,
 bed occupancy?
+
+\spencer{We revised the first paragraph in section 2.2 to read:
+
+``Weekly hospital admission data, which was used as the object of FluSight 
+forecasting for the 
+2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
+Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
+Patient Impact and Hospital Capacity by State Timeseries}. 
+In February 2022 it became mandatory for 
+all hospitals to report the number of COVID-19 and influenza hospitalizations, 
+and since then reporting of hospitalizations has become widespread. These data 
+were updated every Wednesday and Friday according to NHSN guidelines 
+\cite[]{healthdata2024covidts}. Weekly influenza hospitalization data is
+defined as the number of newly hospitalized patients with a confirmed diagnosis 
+of influenza. From hereon, influenza hospitalizations
+will be referrred to as hospitalizations."}
 
@@
 
+\spencer{ILI is useful as a predictor of hospitalizations, but it fails to 
+capture the correlation between consecutive hospitalizations. By adding the 
+AR1 component to a model with ILI as a predictor, much of the correlation
+structure is then accounted for. 
+See the second part of our response to comment 2c. above for how this is 
+addressed in the revised manuscript.}
+
 \item Currently the ILI value from the same week is used in equation (7). Would there be a point in
@@
 
+\spencer{Exploring different lags is an important idea here, especially considering there
+may be a reporting lag. However, the reporting lag is not as big an issue as it has
+been in the past. We assessed ILI lags for both one and two weeks previous to the current
+week and found that ILI had less predictive power. }
+
 \item On page 14 it says: "all negative values of [...] were set to 0 to reflect realistic values of
 hospitalizations". Could this be handled more naturally by appropriate distributional assumptions?
+
+\spencer{We tried both fitting log-hospitalizations as well as assuming 
+hospitalizations followed a distribution truncated at 0, but this led to
+poorer forecasts or problems fitting the posterior distribution. We make this
+statement in the revised manuscript:
+
+``The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
+for hospitalizations at week $w + i$. 
+Occasionally $\{H^*_{s,w + i}\}^K$ included a small number of negative values,
+which do not make sense when the distribution is meant to forecast
+hospitalizations, a nonnegative number. We attempted to alleviate this problem
+by modeling log-hospitalizations or by assuming hospitalizations followed a
+distribution truncated at 0, but the cost of doing so led to poorer forecasts
+or problems drawing from the posterior distribution.
+For the forecast competition analysis in 
+section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
+set to 0 to reflect realistic values of hospitalizations and comply with the 
+FluSight forecasting rules. This ultimately made almost no difference in 
+results."}
 
@@
 
+\spencer{In the revised manuscript, 
+both the simulation study and the real data analysis were updated to include
+comparison of our models to others. In the simulation study, we included out 
+of the box methods including an ARIMA model and a random walk baseline model
+(similar to the baseline used in the FluSight competition).
+In the real data analysis, our model forecasts are compared to those of 20
+non-ensemble models submitted to the FluSight competition during the 2023-24
+season. In both the simulation study and the real data analysis, 
+our models show very good performance when compared to the other models. 
+The original real data analysis where we compare several models that
+belong to our framework is now in the supplementary materials.}
+
 \item The measure the authors call LWIS tends to give a lot of weight to forecast tasks with low
@@
 
+\spencer{In the revised manuscript, we discuss both the WIS and LWIS in the 
+real data analysis section and report both measures. The results for both are
+very similar, so to save space we focus a bit more on the LWIS and included 
+a little more discussion, including a few plots,
+on WIS results in the supplementrary materials.}
+
 \item Why was the CRPS used in the simulation and WIS in the data example? These two are sort of
@@
 
+\spencer{In the revised manuscript we do use the CRPS, but rather 
+use the WIS in both the simulation study
+and the real data analysis. We decided on the WIS because that was the measure
+used in the FluSight competition and the forecasts which we }
+
 \item Some assessment of calibration, e.g. interval coverage, should be added.
+
+\spencer{We updated figures 7 and 9, which show results for the simulation
+study and the real data analysis respectively, to include a plot showing
+empirical predictive coverage for several nominal levels. This showed that 
+our forecasts are reasonably well calibrated relative to others which we 
+compare them to.}
 
@@
 by Osthus and Moran.
+
+\spencer{Add statement to discussion on this ??}
 
@@
 
+\spencer{
+
+Statement at the beginning of section 3.
+
+``With the goal of producing a probablistic forecast, Bayesian modeling is a 
+natural framework for describing forecast uncertainty. 
+Though our main interest is in effeciently producing a probabilistic 
+forecast and not 
+necessarily in estimating model parameters, decisions on certain prior
+distributions and how the models were fit proved critical in producing good 
+forecasts. After defining the ILI and hospitalization models in this section, 
+we describe the 
+selection of prior distributions, model fitting, and producing forecasts
+from the posterior predictive distribution by combining ILI and hospitalization
+models in a way similar to using the 
+Bayesian cut [46]."
+
+Here's an example of one of a 
+few other statements we added which describes the imporance of 
+an informative prior for a 
+specific parameter.
+
+``Modeling discrepancy can lead to 
+overfitting, particularly in forecasting scenarios, and may also lead to 
+identifiability issues. 
+Thus, care must be taken in setting parameter constraints as well as in the 
+selection of prior distributions \cite[]{osthus2019dynamic,
+brynjarsdottir2014learning}. 
+For instance, when setting $f_{\theta_s}(w)$ to
+be the ASG function
+in model (\ref{eq:ili_model}), we found the parameter $\lambda$ from 
+in (\ref{eq:asg_function_rep}) difficult to consistently 
+estimate precisely because of lack of identifiability. This was resolved by 
+assigning to $\lambda$ a more informative prior distribution."
+?? incomplete}
+
 \item As I understand one challenge of forecasting hospitalisation data is that the 
@@
 
+\spencer{In the revised manuscript, 
+both the simulation study and the real data analysis were updated to include
+comparison of our models to others. In the simulation study, we included out 
+of the box methods including an ARIMA model and a random walk baseline model
+(similar to the baseline used in the FluSight competition).
+In the real data analysis, our model forecasts are compared to those of 20
+non-ensemble models submitted to the FluSight competition during the 2023-24
+season. In both the simulation study and the real data analysis, 
+our models show very good performance when compared to the other models. 
+The original real data analysis where we compare several models that
+belong to our framework is now in the supplementary materials.}
+
 \item Why is it sensible to include ILI as a linear function when modelling 
 hospitalisation. Please motivate more.
+
+\spencer{In the original manuscript, figure 6 and some surrounding discussion
+were intended to get across the point that hospitalizations can reasonably
+taken as a linear predictor of ILI along with an autoregressive component.
+Obviously it didn't get the point across very well. 
+In the revised manuscript, figure 6 has been replaced with an updated and more
+descriptive figure (also titled "figure 6") along with the in manuscript
+text.
+
+``Figure 6 helps
+illustrate why this modeling decision was made.
+The figure shows scatterplots of ILI percentage and 
+hospitalizations on the top row.
+The bottom row shows scatterplots of the difference between hospitalizations and 
+1 week 
+lags scaled 
+by ILI percentage at the US national level. The plots in the right column
+are of the logarithm of hospitalizations. The lines through the points are
+drawn by connecting hospitalizations for consecutive weeks during the flu 
+season.
+These plots show that a linear or quadratic relationship between ILI percentage 
+and hospitalizations does not capture much of the temporal correlation in 
+hospitalizations as shown by the obvious loops drawn in the plot. 
+However, when including a 
+single autoregressive lag of hospitalizations in the model, much of the time 
+correlation is
+accounted for as shown by the tighter and more frequent crossing of lines
+in the bottom row. Higher than order 1 autoregressions showed little or no 
+improvement over order 1, hence we consider the ARX(1) reasonably well 
+suited for modeling ILI and hospitalzations."}
 
@@
 within one model seems sensible. Please discuss.
+
+\spencer{Yes, borrowing strength between states would be reasonable, but 
+fitting such a model would likely take much more time than was give by FluSight
+to produce forecasts. In the revised manuscript, after defining the ILI model
+we state:
+
+``Model (\ref{eq:ili_model}) was fit 
+independently for each location. This decision was made largely due to the 
+limited time given by FluSight to produce forecasts. That is,
+during the competition, data were typically released on Wednesday afternoons and
+forecasts were due by the evening of the same day giving only a few hours to 
+produce forecasts. A natural extension of the model would be to allow for states
+to borrow information from neighboring states." ??Add something to discussion}
 
@@
 "first seven season". Provide a time frame.
+
+\spencer{The ``first seven seasons" was replaced with 
+`` the first seven seasons, 2013 to 2019,".}
 
@@
 
+\spencer{I made the changes to the x-axes so that they are labeled by date, 
+listed as month, rather than by week. I also added in the 
+supplementary material the geofacet plots showing 
+ILI and hospitalizations for each state organized by geographic location. Some
+discussion about the relation between neighboring states is made there.}
+
 \item Section 3.5: Please be clear what the strength of a Bayesian inference scheme is.
@@
 
+\spencer{In the revised manuscript we state:
+
+``Here $\widetilde{ILI}_{s,w + i}$ represents the 
+predicted or forecast ILI at $i$ weeks in the future."}
+
 \item Page 11: Please motivate the choice of an autoregressive process of order 1 
 more clearly. Why not use a higher order? Would this be beneficial?
+
+\spencer{We explored higher orders and found that they didn't provide any more
+predictive power than order 1 did. See our response to your general comment 3.}
 
@@
 
+\spencer{We do not generally make this assumption in our model. The only time
+we make this assumption is when modeling log-hospitalizations instead of 
+hospitalizations and then we do it becuase otherwise the predictive variance
+was unreasonably large for lack of hospitalization data.
+Where this is pertinent in the supplementary materials, we note ??.
+In the simulation study,
+we simulate hospitalization data which we attempt to forecast for each left
+out season in the leave one season out scheme. For clarification, we added 
+the following.
+
+``Note that this is not
+a model assumption we make but only a simplyfying assumption for simiulating
+the data."}
+
 \item Figure 7: Please make clear in the caption what the columns represent in an 
 intuitive manner.
+
+\spencer{In the updated manuscript, this figure was replaced with another figure
+which included 3 plots we believe is much more intuitive and descriptive than
+the original figure. The title of the new figure remains ``figure 7" and 
+the caption is hopefully understandable.}
 
@@
 
+\spencer{We decided to put equations 11 - 14 in the supplementary materials, and
+give more of an over description of the prior distributions in section 3.6. Here
+is one updated paragraph describing how a prior distribution was assigned to
+improve identifiability issues introduced by including discprepancy modeling.
+
+``Additional prior constraints were made to improve parameter identifiability. 
+Just as was done in Osthus et al. \cite[]{osthus2019dynamic},
+we set the initial value of the 
+susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
+The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
+priors. In a previous version of this manuscript, when fitting an ASG model
+with discrepancy components,
+rather than estimating all parameters in a Bayesian way,
+we set the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})
+to be equal to its MLE. In this updated manuscript, 
+however, we use empirical Bayes and center the first component of the 
+hierarichal prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
+all seasons and set the variance parameter so the prior was tight around the
+mean."}
+
 \item Please check for typos.

### 2025-08-08 manuscript_res.tex
finish responses to reviewers

@@
 of forecasting for FluSight with ILI being the target for the first seven
-seasons and hospitalizations being the target since the 2022 season. Both of 
+seasons, 2013 to 2019, and hospitalizations being the target since the 
+2022 season. Both of 
 these data were collected at the state, territorial, and national level and 
@@
 information on respiratory illness from outpatient visits to health care 
-providers. Over 3,400 outpatient health care providers in all 50 US states, PR, 
-DC, and the US Virgin Islands report each week the total number of outpatient 
+providers. Each week, over 3,400 outpatient health care providers in all 50 
+US states, PR, 
+DC, and the US Virgin Islands report the total number of outpatient 
 visits along with the number of ILI cases. An ILI case is defined as a 
@@
     \centering
-    \includegraphics[scale=.4]{Images/us_hospitalizations.png}
+    \includegraphics[scale=.42]{Images/us_hospitalizations.png}
     \caption{Percentage of outpatient visits with an influenza-like illness 
@@
 
-Hospital admission data, used as the object of FluSight forecasting for the 
+Weekly hospital admission data, which was used as the object of FluSight 
+forecasting for the 
 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
 Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
-Patient Impact and Hospital Capacity by State Timeseries}. Several targets of 
-respiratory illnesses including COVID-19, RSV, and influenza are reported 
-weekly by most hospitals in the US. In February 2022 it became mandatory for 
+Patient Impact and Hospital Capacity by State Timeseries}. 
+In February 2022 it became mandatory for 
 all hospitals to report the number of COVID-19 and influenza hospitalizations, 
@@
 were updated every Wednesday and Friday according to NHSN guidelines 
-\cite[]{healthdata2024covidts}.
+\cite[]{healthdata2024covidts}. Weekly influenza hospitalization data is
+defined as the number of newly hospitalized patients with a confirmed diagnosis 
+of influenza. From hereon, influenza hospitalizations
+will be referrred to as hospitalizations.
 
@@
 plots of ILI data and hospitalization
-data for select states are included in the supplementary material 
+data for all 50 US states are included in the supplementary material 
 \cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar 
@@
 used for modelling disease outbreaks. One important compartmental model is the 
-susceptible-infectious-recovered (SIR) model, which is used by some to model 
+susceptible-infectious-recovered (SIR) 
+disease transmission model, which has been used by some to model 
 ILI data \cite[]{osthus2019dynamic, allen2017primer}. Ulloa \cite[]{ulloa2019} 
-chose to use the asymmetric Gaussian (ASG) function to model ILI data. The 
-SIR and ASG models may both be appropriate to describe ILI behavior over the 
-course of a flu season, however there may also be systematic behavior not 
-captured by either, necessitating an additional model component to capture 
-the discrepancy. In the first part of this section, we present an ILI model 
+chose to use the asymmetric Gaussian (ASG) function to model ILI data. 
+The SIR is a mechanistic model used to capture transmission of a particular
+disease, and the ASG is a smooth function which may approximate the 
+same shape as the ILI trajectory. As ILI may be influenced by more 
+diseases than influenza, the disease transmission described by the SIR model
+can only partially contribute to the true ILI data-generating mechanism
+\cite[]{osthus2019dynamic}.
+Though the ASG function may be able to roughly capture ILI, there may still be 
+systematic behavior which it cannot capture. Thus when using either the SIR
+or ASG models, an additional component for modeling discrepancy should be 
+included.
+
+In the first part of this section, we present an ILI model 
 similar to the model in Osthus et al. \cite[]{osthus2019dynamic}. With some 
@@
 though the focus here is on the SIR and ASG functions which are defined. 
-
 With the aim of forecasting hospitalizations, we also introduce a linear 
@@
 hospitalization forecasts. The hospitalization model is also defined in this 
-section, and the section is concluded with descriptions of selected prior 
-distributions, model implementation, and posterior sampling.
+section.
+
+The forecasts produced should be probabilistic, making Bayesian modeling a 
+natural framework for describing forecast uncertainty. 
+Though our main interest is in effeciently producing a probabilistic 
+forecast and not 
+necessarily in estimating model parameters, decisions on certain prior
+distributions and how the models were fit proved critical in producing good 
+forecasts. After defining the ILI and hospitalization models in this section, 
+we describe the 
+selection of prior distributions, model fitting, and producing forecasts
+from the posterior predictive distribution by combining ILI and hospitalization
+models in a way similar to using the 
+Bayesian cut \cite[]{plummer2015cuts}.
 
@@
 \ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically 
-over seasons.
+over seasons. Herein we also adopt the hierarchical modeling as it proved a 
+powerful tool for improving forecasts.
+
+Model (\ref{eq:ili_model}) was fit 
+independently for each location. This decision was made largely due to the 
+limited time given by FluSight to produce forecasts. That is,
+during the competition, data were typically released on Wednesday afternoons and
+forecasts were due by the evening of the same day giving only a few hours to 
+produce forecasts. A natural extension of the model would be to allow for states
+to borrow information from neighboring states.
 
@@
 brynjarsdottir2014learning}. 
+For instance, when setting $f_{\theta_s}(w)$ to
+be the ASG function
+in model (\ref{eq:ili_model}), we found the parameter $\lambda$ from 
+in (\ref{eq:asg_function_rep}) difficult to consistently 
+estimate precisely because of lack of identifiability. This was resolved by 
+assigning to $\lambda$ a more informative prior distribution.
 Modeling of the discrepancy for ILI was done by Osthus et al. during the 2015 
@@
 illustrate why this modeling decision was made.
-shows scatterplots of ILI percentage and hospitalizations on the top row.
+The figure shows scatterplots of ILI percentage and 
+hospitalizations on the top row.
 The bottom row shows scatterplots of the difference between hospitalizations and 
-scaled 1 week 
-lags by ILI percentage at the US national level. The plots in the right column
-are from the logarithm of hospitalizations. The lines through the points is
+1 week lags scaled 
+by ILI percentage at the US national level. The plots in the right column
+are of the logarithm of hospitalizations. The lines through the points are
 drawn by connecting hospitalizations for consecutive weeks during the flu 
@@
 However, when including a 
-single lag of hospitalizations in the model, much of the time correlation is
-accounted for as shown by the tighter and more frequent crossing of lines shown
+single autoregressive lag of hospitalizations in the model, much of the time 
+correlation is
+accounted for as shown by the tighter and more frequent crossing of lines
 in the bottom row. Higher than order 1 autoregressions showed little or no 
@@
 according to some distribution $D_s$ with mean parameter 0, scale parameter 
-$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees 
-of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
-
-This model is for any location, and for fitting purposes $ILI_{s,w}$ is always 
+$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ which 
+may be used as degrees of freedom
+when $D_s$ belongs to the location-scale t (LST) family.
+
+Like the ILI model (\ref{eq:ili_model}), 
+the hospitalization model was fit independently fore each location.
+For fitting purposes $ILI_{s,w}$ is always 
 multiplied by $P$ which is proportional to the population of the state or 
@@
 the ILI posterior predictive distribution is used along with the posterior 
-distribution for the parameters in (\ref{eq:final_hosp}). We considered three 
-scenarios for model (\ref{eq:final_hosp}). A model where $D_s$ belongs to the 
-normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ 
-is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, 
-or $H_{s,w} + 1$ is lognormally distributed (LNORM). The population value $P$ 
-is excluded from the LNORM model, and we set 
-$\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting. In 
-the LNORM model, if the linear parameters are not set the same for seasons 22 
-and 23, the final variance was more prone to be extreme. Besides varying the 
-distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ 
-or there is no quadratic ILI term.
+distribution for the parameters in (\ref{eq:final_hosp}). 
+We let $D_s$ belong to the normal family, though in the supplementary materials
+?? we show a real data analysis where we considered other distribution families
+including the LST and lognormal families.
+
+% We considered three 
+% scenarios for model (\ref{eq:final_hosp}). A model where $D_s$ belongs to the 
+% normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ 
+% is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, 
+% or $H_{s,w} + 1$ is lognormally distributed (LNORM). The population value $P$ 
+% is excluded from the LNORM model, and we set 
+% $\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting. In 
+% the LNORM model, if the linear parameters are not set the same for seasons 22 
+% and 23, the final variance was more prone to be extreme. Besides varying the 
+% distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ 
+% or there is no quadratic ILI term.
 
@@
 The priors selected for the ILI data model under both the SIR and ASG models 
-largely follow the prior selections in Osthus et al. \cite[]{osthus2019dynamic} 
-and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved 
-numerical stability and/or we felt the adjusted prior made more sense for the 
+largely follow the prior distribution 
+selections in Osthus et al. \cite[]{osthus2019dynamic} 
+and Ulloa \cite[]{ulloa2019} with a few exceptions where selecting
+different priors improved 
+identifiability, 
+numerical stability, and/or we felt the adjusted prior made more sense for the 
 problem. For model (\ref{eq:ili_model}), parameters which are common even when 
 using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
-$\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$. For the SIR function 
+$\sigma_{\gamma}^2$, $\sigma_{\gamma_W}^2$, and $\sigma_{\upsilon}$. For the SIR function 
 $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
@@
 a half-normal prior if the support is nonnegative, or a truncated-normal 
-prior to match a more specific support. Under the ASG model, $\theta_s$ is 
+prior to match a more specific support. Under the ASG model, $\theta_s$ was 
 modeled hierarchically over seasons so that for each season the transformed 
 parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are 
-assigned to $\theta$ and $\Sigma$.
+assigned to $\theta$ and $\Sigma$. Here 
+$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2)$.
 
 Additional prior constraints were made to improve parameter identifiability. 
-In Osthus et al. \cite[]{osthus2019dynamic} the initial value of the 
+Just as was done in Osthus et al. \cite[]{osthus2019dynamic},
+we set the initial value of the 
 susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
 The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
-priors. 
+priors. In a previous version of this manuscript, when fitting an ASG model
+with discrepancy components,
+rather than estimating all parameters in a Bayesian way,
+identifiability was encouraged by setting
+the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})
+to be equal to its MLE. In this updated manuscript, 
+however, we use empirical Bayes and center the first component of the 
+hierarichal prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
+all seasons and set the variance parameter so the prior was tight around the
+mean. This led to improved mixing of posterior draws and better forecasts.
+The 
+specific prior distributions assigned to each parameter are listed in the 
+supplementary materials ??.
+
 % To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
@@
 Rather than combining the ILI model in equation (\ref{eq:ili_model}) with the 
-hospitalization model in equation (\ref{eq:final_hosp}) and computing a fully Bayesian
+hospitalization model in equation (\ref{eq:final_hosp}) and fitting a single
 posterior distribution, both models were fit separately and 1 to 4 week ahead
@@
 showed convergence even after 55,000 iterations, whereas the component model 
-fits gave no signs to be concerened with convergence. And most importantly,
+fits gave no signs which raised concern for convergence. And most importantly,
 the forecasts from the component models generally outperformed those of the 
@@
 
-We assessed the model fit for four ILI models. These included the SIR and 
+We assessed the model fit for four ILI models and for 
+the hospitalization model. The ILI models included the SIR and 
 ASG models and models with and without discrepancy modeling. When discrepancy 
@@
 2020 season was excluded because of the unique behavior during that season. 
-Assessment for hospitalization modeling was done for six different models. 
-These include NORM, LNORM, and LST models, and models where a quadratic ILI 
-term is included or excluded. To assess posterior sampling convergence, models 
-were fit to data where ILI and hospitalization data up to week 14 of the 2023 
-season was included. Sampling was done with four chains where from each chain 
+% Assessment for hospitalization modeling was done for six different models. 
+% These include NORM, LNORM, and LST models, and models where a quadratic ILI 
+% term is included or excluded. 
+To assess posterior sampling convergence, models 
+were fit to data where ILI and hospitalization data up to weeks 14, 20, and 26
+of the 2023 
+season were included. Sampling was done with four chains where from each chain 
 60,000 posterior draws were sampled, and the first 10,000 draws were discarded 
@@
 effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for 
-each parameter. Tables \ref{tab:ili_diagnostics} and 
-\ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum 
-$ESS$ over all parameters for ILI and hospitalization models respectively. 
-Forecast models for all other weeks of the season were fit using one chain of 
-60,000 draws where the first 10,000 draws were discarded as a burn-in. 
-% \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
-For parameters of the ASG models that were prone to cause trouble in posterior 
-sampling, the starting values were set to be the MLEs.
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI 
-models fitted on US data for week 14 of the 2023 season}
-\label{tab:ili_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccc}
-     & ASG & ASGD & SIR & SIRD \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
-     $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
-    \end{tabular}
-    \end{center}
-\end{table}
-
-\begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six 
-hospitalization models fitted on US data for week 14 of the 2023 season}
-\label{tab:hsop_diagnostics}
-\begin{center}
-    \begin{tabular}{c|cccccc}
-     & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
-     \hline
-     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & 
-     $< 1.001$\\
-     $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
-    \end{tabular}
-    \end{center}
-\end{table}
+each parameter, and for all models fit, the largest $\hat{R}$ was below the 
+recommend threshold of 1.05, and the smallest $ESS$ was over 1,000.
+% Tables \ref{tab:ili_diagnostics} and 
+% \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum 
+% $ESS$ over all parameters for ILI and hospitalization models respectively. 
+% Forecast models for all other weeks of the season were fit using one chain of 
+% 60,000 draws where the first 10,000 draws were discarded as a burn-in. 
+% % \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
+% For parameters of the ASG models that were prone to cause trouble in posterior 
+% sampling, the starting values were set to be the MLEs.
+
+% \begin{table}
+% \caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI 
+% models fitted on US data for week 14 of the 2023 season}
+% \label{tab:ili_diagnostics}
+% \begin{center}
+%     \begin{tabular}{c|cccc}
+%      & ASG & ASGD & SIR & SIRD \\
+%      \hline
+%      $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
+%      $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
+%     \end{tabular}
+%     \end{center}
+% \end{table}
+
+% \begin{table}
+% \caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six 
+% hospitalization models fitted on US data for week 14 of the 2023 season}
+% \label{tab:hsop_diagnostics}
+% \begin{center}
+%     \begin{tabular}{c|cccccc}
+%      & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
+%      \hline
+%      $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & 
+%      $< 1.001$\\
+%      $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
+%     \end{tabular}
+%     \end{center}
+% \end{table}
 
@@
 The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
-for hospitalizations at week $w + i$. For the forecast competition analysis in 
+for hospitalizations at week $w + i$. 
+Occasionally $\{H^*_{s,w + i}\}^K$ included a small number of negative values,
+which do not make sense when the distribution is meant to forecast
+hospitalizations, a nonnegative number. We attempted to alleviate this problem
+by modeling log-hospitalizations or by assuming hospitalizations followed a
+distribution truncated at 0, but the cost of doing so led to poorer forecasts
+or problems drawing from the posterior distribution.
+For the forecast competition analysis in 
 section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
 set to 0 to reflect realistic values of hospitalizations and comply with the 
-FluSight forecasting rules.
+FluSight forecasting rules. This ultimately made almost no difference in 
+results.
 
@@
 existing ILI data as a predictive covariate. Each season was in turn 
-``left-out" and treated as if it was the most recent season which we desired 
-to forecast. Fitting and forecasting was then done for weeks 14, 20, 26, 32, 
+``left-out" and treated as if it was the most recent season for which forecasts
+are desired. Fitting and forecasting was then done for weeks 14, 20, 26, 32, 
 and 38 of the left out season, giving two weeks that tend to occur as flu 
@@
 fitting criterion. Both the BASE and ARIMA ILI models were fit using the 
-\texttt{forecast} package in \texttt{R} \cite{hyndman2008forecast}. Samples
-were then drawn from the forecast distributions, and these were then used
-along with the model in (\ref{eq:final_hosp}) to make hospitalization 
-forecast distributions.
+\texttt{forecast} package in \texttt{R} \cite{hyndman2008forecast}.
 
@@
 (\ref{eq:final_hosp}) were considered the same across all seasons so that 
-all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$. 
+all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$. Note that this is not
+a model assumption we make but only a simplyfying assumption for simulating
+the data.
 The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were 
@@
 The proper scoring rules primarily used in the FluSight competition and the 
-COVID-19 Forecast Hub and which we will use herein are the 
-weighted interval score (WIS) and the log-weighted interval score (LWIS).
+COVID-19 Forecast Hub and which we will use herein are the negatively oriented
+(smaller is better)
+weighted interval score (WIS)
 \cite[]{mathis2024evaluation, bracher2021evaluating}.
@@
 weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
-interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
-single interval as defined in (\ref{eq:is}). Here 
+interval where $b \in \{1, ..., B\}$. 
+$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
+single interval, defined in (\ref{eq:is}). Here 
 $\vmathbb{1}\{\cdot\}$ is the indicator function.
-The LWIS is the same as the WIS except that it is evaluated 
-over the log of quantiles and the log of the observed value.
 
@@
 making each forecast a set of 11 predictive intervals and a predictive median.
-Figure \ref{fig:wis_by_week_horizon} shows 
-boxplots of the WIS for the 500 simulation replicates faceted by week 
-on the y-axis and horizon on the x-axis. 
-% Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
-In most plots the ASGD model shows the best performance, 
-especially during week 20 where
-forecasts are near weeks 22 and 23 which show a seasonal peak and trough in 
-figure \ref{fig:discrepancy}. The SIRD also shows improved performance over the
-SIR model, though its performance does not match that of the ASG or ASGD.
-Here the ASGD model regularly outperforms
-the BASE and ARIMA models. The three other nonlinear models do not perform as
-well.
-The ASG model
-tends to outperform the SIR and SIRD models suggesting that the additional 
-flexibility of the five parameters ASG function 
-is often better for forecasting than that of the three parameter SIR function.
-The model with the best performance is the ASGD.The ASGD model shows 
-especially good forecast skill at week 20 around where we would expect the 
-discrepancy to be most useful.
+Figure \ref{fig:wis_by_week_horizon} gives three plots showing results of the 
+study. Plot a) shows WIS scores averaged over each left out season and each 
+of the 500 simulation replicates. The ASGD forecasts on average clearly 
+outperformed the other models considered for almost each week. The ASGD does
+especially well at week 20 which contains forecasts of the holiday week 22, 
+showing the particular importance of modeling discrepancy at and around that 
+week. The SIRD outperforms the SIR model, again showing the importance of 
+modeling discrepancy, but it is still outperformed by the ASG and even for a
+few weeks by the BASE and ARIMA models. Plot b) shows boxplots of all forecast
+scores by model and separated by horizon forecasts of 1 to 4 weeks ahead. Here
+also the ASGD appears to perform the best with the other variations of the 
+nonlinear models not always outperforming the BASE or ARIMA models. 
+To assess calibration, plot c)
+shows the empirical coverage of each model for 11 predictive intervals from 
+10\% to 98\% nominal levels. Each model is was undercalibrated, but the ASGD 
+model is the closest to being well calibrated, especially at the hightest 
+nominal levels. Table \ref{tab:sim_wis_res} shows summaries of the mean WIS and
+coverage for 50\% and 95\% predictive intervals.
 
@@
 
+% \begin{minipage}
 \begin{figure}[hbt!]
     % \centering
-    \includegraphics[scale=.5]{Images/all_sim_res.png}
-    \caption{Boxplots of weighted interval score (WIS) for the 
-    four Bayesian nonlinear ILI models, a baseline model (BASE) and
-    an autoregressive integrated moving average model (ARIMA) 
-    over all seasons in the simulation study faceted by horizon 
-    (x-axis) and season week (y-axis). Horizons include 1-4 week ahead forecasts 
-    and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
+    \begin{adjustwidth}{-.65cm}{-1.5cm}
+    \includegraphics[scale=.45]{Images/all_sim_res.png}
+    \end{adjustwidth}
+    \caption{Results of forecasts for simulated hospitalizations for ASGD, ASG,
+    SIRD, SIR, BASE, and ARIMA models. Plot a), WIS averaged over all
+    left out seasons and 500 simulation replicated for each of weeks 14, 20, 
+    26, 32, and 38, coloured by model. Plot b), boxplots of all WIS 
+    scores separated by forecast horizon of 1 to 4 weeks ahead. Plot c),
+    empirical calibration of each forecast model as the overall 
+    forecast coverage of predictive intervals for the 11 nominal levels
+    10\%, 20\%, ..., 90\%, 95\%, 98\%.}
     \label{fig:wis_by_week_horizon}
 \end{figure}
+% \end{minipage}
 
@@
 \centering
+\caption{Summary values of forecast performance in simulation study for 
+    ASGD, ASG,
+    SIRD, SIR, BASE, and ARIMA models. WIS, 50\% and 95\% predictive interval
+    coverage averaged over all seasons and 500 simulation replicates.}
 \begin{tabular}{lrrr}
@@
 \end{tabular}
+\label{tab:sim_wis_res}
 \end{table}
+
+The results of this simulation study demonstrate the effectiveness of using the
+assymetric Gaussian function along with modeling discrepancy for modeling
+ILI and making flu hospitalization
+forecasts in the US. The following section shows similar effectiveness of
+coupling an effective ILI flu model with a hospitalization model to forecast
+hospitalizations during the 2023 flu season.
 
@@
 context of the FluSight competition where each competing forecast was 
-submitted as a set of 23 quantiles corresponding to give probabilities.
+submitted as a set of 23 quantiles corresponding to given probabilities.
 Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
@@
 % Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
- Primary scores for evaluating each forecast were the WIS and LWIS
-
-The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
-Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
-$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
-rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
-indicator function for values within the set $A$.
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:sir_prior}
-        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
-        \beta_s &\sim N^+(0.8, 0.3) \\
-        \rho_s &\sim N^+(0.68, 0.08)
-\end{aligned}
-\end{equation}
-For the ASG model, the MLE 
-$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ 
-was calculated. % and $\hat{\lambda}_s$ was accepted as a fixed value. 
-These estimates were used as starting values for posterior sampling. 
-The transformation 
-$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ 
-was made, and a prior distribution was assigned to $T(\theta_s)$.
-The prior distributions for the ILI model under the ASG function are shown 
-in (\ref{eq:ili_prior}). These are slightly informative priors because for 
-most parameters we have an idea what reasonable values may be. Here 
-$m = (-4.5, 0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.3, 0.2, 5, 2, 2)$
-where $\text{diag}(\cdot)$ 
-is the diagonal matrix for the given entries.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:ili_prior}
-                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
-                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
-                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
-                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
-\end{aligned}
-\end{equation}
-Parameters shared by both the SIR and ASG models are the scale parameter 
-$\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and 
-$\sigma_{\gamma_W}^2$. These priors are in (\ref{eq:shared_ili_priors}).
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:shared_ili_priors}
-        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
-        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
-        \sigma_{\upsilon_{s,w}} &\sim N^+(0, .1^2)
-\end{aligned}
-\end{equation}
-Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
-we selected an informative prior distribution by first estimating  
-$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
-calculating the MLE for $\theta_s$. 
-% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
-%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
-$\widehat{ILI}_{s,w}$ was then predicted such that 
-$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
-Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
-seasons of the differences 
-$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
-When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
-calculated using the \texttt{mle2} function in the \texttt{bblme} package 
-\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
-package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
-function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
-function.
-The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
-
-\begin{equation}
-\begin{aligned}
-\label{eq:shared_hosp_priors}
-        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
-        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
-        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
-\end{aligned}
-\end{equation}
-
-We fit 24 forecast models for each location for all 30 weeks, and for each 
-week forecast 1-4 week ahead hospitalizations. The 24 models included all 
-combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
-hospitalization models, and both quadratic and linear hospitalization models.
-Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US 
-flu hospitalizations during the 2023 season under the NORM hospitalization 
-model. The forecasts shown are for weeks 14, 20, 26, and 32. The predictive 
-bands are the 50\% and 95\% predictive intervals.
-These plots show a tendency to often underpredict hospitalizations. The 
-quadratic models appear to do a better job predicting  hospitalizations at 
-the season peak but a poorer job predicting after the peak, whereas the linear 
-models seem to predict well or slightly overpredict after the peak.
+ Primary scores for evaluating each forecast were the WIS and the 
+ log-weighted intveral score (LWIS). The LWIS is defined the same as the WIS
+ but where the log of the observation as well as the forecast quatiles are taken
+ prior to calculating the score. We primarily use the LWIS in this section,
+ though forecasting results in terms of the WIS may also be found in the 
+ supplementary materials ??.
+
+Figure \ref{fig:normal_flu_forecasts} shows hospitalization forecasts at 
+four different weeks during the season for ASGD, ASG, SIRD, and SIR ILI models
+where hospitalizations are modeled according to model (\ref{eq:final_hosp})
+where ILI is a quadratic predictor of hospitalizations, and hospitalizations
+are assumed to be normally distributed. Of these forecasts, the ASGD appears 
+to do the best job at capturing the trajectory of hospitalizations as well
+as usually making the sharpest predictions, particularly later in the season.
 
@@
   \centering
-  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
+  \begin{adjustwidth}{-1cm}{-1.9cm}
+  \includegraphics[scale=.45]{Images/normal_sq_forecasts_us.png}
+  \end{adjustwidth}
 \caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
 season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, 
-and the hospitalization models are all normally distribution. The figure 
-includes hospitalization forecasts where ILI is a linear predictor (left) 
-and where ILI is a quadratic predictor (right). 50\% predictive intervals 
-are pink and 95\% predictive intervals are red.}
+and the hospitalization models are all normally distributed and ILI is a 
+quadratic predictor.}
 \label{fig:normal_flu_forecasts}
@@
 
-Of the 24 models, the one which produced the best forecasts was that using
-the ASGD ILI model and the NORM quadratic hospitalization model. The results of
-that analysis may be found in the supplementary materials ??. For the remainder 
-of this section, forecast performance from the best model, which we denote
-as ASG\_NORM2, is compared to that of forecasts from 20 models from which
-submissions were made to FluSight in 2023. 
-The models selected are those which were not ensemble models. Among the models
+For the
+remainder of this section, focus will be made on forecasts made by coupling the
+ASGD model with normal hospitalization model where ILI is a quadratic predictor.
+We denote this model as ASGD\_NORM2. 
+This decision was based on the ASGD's performance in the simulation study in 
+section \ref{sec:simulation2} as well as an analysis of several other forecast
+models including ASG, SIRD, SIR ILI models coupled with other variations of the
+hospitalization model where the ASGD\_NORM2 model showed the best performance.
+Details
+of that analysis are found in the supplementary materials ??.
+
+The FluSight 
+models to which the ASGD\_NORM2 model are compared are the 
+20 non-ensemble models
+submitted to FluSight during the 2023 season.
+The submitted forecasts may be found at \cite{mathis2023flusight}.
+Among the models
 was the FluSight baseline. The baseline forecast had as a median the most 
@@
 hospitalization forecasts for all 21 models being compared but with results
-for the ASG\_NORM2 and FluSight baseline models highlighted. 
+for the ASGD\_NORM2 and FluSight baseline models highlighted. 
 For each week of the 2023 season, the  
@@
 for each model is shown in a). Boxplots of LWIS for 1, 2, 3, and 4-week ahead
-horizons for the ASG\_NORM2 and baseline models are shown in b). And 
+horizons for the ASGD\_NORM2 and baseline models are shown in b). And 
 overall empirical coverage of the 11 predictive intervals for all models is 
-shown in c) with ASG\_NORM2 and the baseline model again highlighted. 
-These plots show that over the season the ASG\_NORM2 model tended to 
+shown in c) with ASGD\_NORM2 and the baseline model again highlighted. 
+These plots show that over the season the ASGD\_NORM2 model tended to 
 outperform most others in terms of LWIS, and in general is superior to the 
-baseline model. The coverage plot shows that the ASG\_NORM2 model is well 
+baseline model. The coverage plot shows that the ASGD\_NORM2 model is well 
 calibrated at most levels, with slight overcoverage at the lower levels
@@
     % \centering
-    \includegraphics[scale = .6]{Images/lwis_cover_sum.png}
+    \begin{adjustwidth}{-1.2cm}{-1.5cm}
+    \includegraphics[scale = .54]{Images/lwis_cover_sum.png}
     \caption{a) Averaged over location and horizon log-weighted interval score
     (LWIS) for each week during the season for the 21 models being compared. 
-    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
+    Each grey line represents one model, but the ASGD\_NORM2 (blue) and 
     FluSight-baseline (red) are colored to stand out.
     b) Boxplots of all LWIS scores for 1, 2, 3, and 4-week ahead horizons 
-    for the ASG\_NORM2 (blue) and FluSight-baseline (red) models.
+    for the ASGD\_NORM2 (blue) and FluSight-baseline (red) models.
     c) Plot of empirical coverage against theoretical coverage for 11 
@@
     line $x = y$ shows where optimal coverage would fall.
-    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
+    Each grey line represents one model, but the ASGD\_NORM2 (blue) and 
     FluSight-baseline (red) are colored to stand out.}
+    \end{adjustwidth}
     \label{fig:lwis_cover_sum}
@@
 
-Figure \ref{fig:state_and_date_lwis} shows the average LWIS for each model
-relative to the ASG\_NORM2 model by location in the top tile plot and by 
-season week in the bottom tile plot. That is, the value for each tile is the 
-ratio of the average ASG\_NORM2 LWIS to the average for the labeled model at the
-labeled location or week. A value less than 1 shows superior performance by
-ASG\_NORM2 with increasing superiority as the value goes to 0. A value larger
-than 1 shows inferior performance, and a value of 1 shows equal performance.
-Tiles are grey where the labeled forecast model made no forecast submissions.
-Viewed this way, only UGA\_flucast-INFLAenza appears to potentially outperform
-ASG\_NORM2.
+During the FluSight competition, most participating teams would occasionally 
+miss a forecast submission for some week, location, or horizon. This makes a 
+simple average of scores an imperfect measure of comparison, as forecasts
+of certain targets included for one model may be excluded by another model.
+Thus in figure \ref{fig:state_and_date_lwis} 
+we compare the ASGD\_NORM2 model with all 
+other models directly by comparing LWIS scores averaged over only shared
+forecasts. Each tile in the plots represents the ratio of average ASGD\_NORM2
+LWIS to that of one FluSight model by location or by week.
+We call this the relative LWIS or RLWIS. A value smaller than
+1 indicates the ASGD\_NORM2 model outperforms the other model with 
+increasing outperformance as the value goes to 0. A value 
+higher than 1 indicateds the ASGD\_NORM2 is outperformed. A grey tile indicates
+the competing model had no forecasts matching the specified target.
+Similar plots showing WIS scores are found in the supplementary materials ??.
+
+Our model outperforms most other models with only the 
+UGA\_flucast-INFLAenza model appearing to outperform ours. 
+Table \ref{tab:fin_analysis_stats} gives summary results for mean and median
+WIS and LWIS scores, predictive coverage, and relative RWIS and RLWIS scores
+for each model. These summary values show the ASGD\_NORM2 as performing very
+well compared to all other models, and outperforming all models except
+UGA\_flucast-INFLAenza by RWIS and RLWIS. 
+The takeaway of this section, like those of section \ref{sec:simulation2}, show
+the combination of well performing model of ILI, the 
+hierarchical Bayesian ASGD, with a 
+straightforward hospitalization model lead to good forecasts which 
+outperformed almost all other non-ensemble forecasts during of the 2023
+FluSight competition.
 
 \begin{figure}[hbt!]
-\begin{adjustwidth}{-3cm}{-1.5cm}
+\begin{adjustwidth}{-1.2cm}{-1.5cm}
     % \begin{minipage}{\linewidth}
@@
     \makebox[\linewidth][l]{
-    \includegraphics[scale = .45]{Images/state_and_date_lwis.png}
+    \includegraphics[scale = .5]{Images/state_and_date_lwis.png}
     }
     \end{adjustwidth}
-    \caption{}
+    \caption{Ratio of average log-weighted interval score for ASGD\_NORM2 
+    to that of on competing FluSight model. Scores are averaged
+    over all targets within indicated location (top) or week (bottom) and 
+    only include forecasts shared by both models. Lower scores (blue) indicate
+    superior performance by the ASGD\_NORM2 model and higher scores (red)
+    indicate superior performance by competing model.}
     \label{fig:state_and_date_lwis}
@@
 \begin{adjustwidth}{-3cm}{-1.5cm}
+\caption{Overall summary scores for each of the 20 non-ensemble FluSight
+models and the ASGD\_NORM2 model. Summaries include mean weighted interval
+and log-weighted interval scoreS (MWIS/MLWIS), 
+median weighted interval and log-weighted interval scores (MedWIS/MedLWIS), 
+mean squared error
+difference between predictive model empirical and theoretical coverage (MSEC)
+see figure \ref{fig:lwis_cover_sum} c)), percentage of forecasts for all 
+targets during the 2023 season (\% Forcs), and relative weighted interval
+and log-weighted interval scores (RWIS/RLWIS).}
 \begin{tabular}{lrrrrrrrr}
@@
   \hline
-ASG & 62 & 0.34 & 14 & 0.23 & 0.001 & 100 & 1 & 1 \\ 
+ASGD\_NORM2 & 62 & 0.34 & 14 & 0.23 & 0.001 & 100 & 1 & 1 \\ 
   PSI-PROF & 63 & 0.35 & 15 & 0.23 & 0.003 & 100 & 0.90 & 0.97 \\ 
@@
 \end{adjustwidth}
+\label{tab:fin_analysis_stats}
 \end{table}
-
-It should not be assumed that these two takeaways will apply for future flu 
-seasons. As suggested by the simulation study in the previous section, it may 
-often be the case that the ASG ILI model is better for forecasting. Indeed, a 
-close examination of figures \ref{fig:us_lwis} and \ref{fig:lwis_by_dist_loc} 
-shows that it is common for the ASG models to show better forecasting skill 
-than the SIR models.
 

### 2025-08-08 supplemental_res.tex
finish responses to reviewers

@@
 \end{figure}
+
+\section{Prior distributions}
+
+The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
+Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
+$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
+rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
+indicator function for values within the set $A$.
+
+\begin{equation}
+\begin{aligned}
+    \label{eq:sir_prior}
+        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
+        \beta_s &\sim N^+(0.8, 0.3) \\
+        \rho_s &\sim N^+(0.68, 0.08)
+\end{aligned}
+\end{equation}
+For the ASG model, the MLE 
+$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ 
+was calculated. % and $\hat{\lambda}_s$ was accepted as a fixed value. 
+These estimates were used as starting values for posterior sampling. 
+The transformation 
+$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ 
+was made, and a prior distribution was assigned to $T(\theta_s)$.
+The prior distributions for the ILI model under the ASG function are shown 
+in (\ref{eq:ili_prior}). These are slightly informative priors because for 
+most parameters we have an idea what reasonable values may be. Here 
+$m = (-4.5, 0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.3, 0.2, 5, 2, 2)$
+where $\text{diag}(\cdot)$ 
+is the diagonal matrix for the given entries.
+
+\begin{equation}
+\begin{aligned}
+\label{eq:ili_prior}
+                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
+                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
+                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
+                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
+\end{aligned}
+\end{equation}
+Parameters shared by both the SIR and ASG models are the scale parameter 
+$\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and 
+$\sigma_{\gamma_W}^2$. These priors are in (\ref{eq:shared_ili_priors}).
+
+\begin{equation}
+\begin{aligned}
+    \label{eq:shared_ili_priors}
+        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
+        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
+        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
+        \sigma_{\upsilon_{s,w}} &\sim N^+(0, .1^2)
+\end{aligned}
+\end{equation}
+Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
+we selected an informative prior distribution by first estimating  
+$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
+calculating the MLE for $\theta_s$. 
+% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
+%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
+$\widehat{ILI}_{s,w}$ was then predicted such that 
+$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
+Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
+seasons of the differences 
+$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
+When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
+calculated using the \texttt{mle2} function in the \texttt{bblme} package 
+\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
+package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
+function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
+function.
+The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
+
+\begin{equation}
+\begin{aligned}
+\label{eq:shared_hosp_priors}
+        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
+        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
+        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
+        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
+        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
+        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
+\end{aligned}
+\end{equation}
 \end{supplement}

### 2025-08-08 make_images/make_images_for_data_section.R

@@ -1,13 +1,21 @@
 
 setwd("./nonlinear_flu_forecast/")
-source('./functions_data/get_data_functions_new.R')
-source('./nonlinear_flu_forecast/functions_data/mle_functions.R')
+source('./flu_forecast_23/get_data_functions_new.R')
+source('./flu_forecast_23/mle_functions.R')
+library(geofacet)
+library(datasets)
+library(ggtext)
+state.name <- c(state.name, "Puerto Rico", "District of Columbia", "US")
+state.abb <- c(state.abb, "PR", "DC", "US")
+states <- data.frame(loc_abb = state.abb, region = state.name)
 
 
-
-both_flu <- comb_data(lag = 2) %>% 
+both_flu <- comb_data(lag = 1) %>% 
   mutate(uw_odd = (unweighted_ili/100)/(1 - (unweighted_ili/100)))
 ILINet <- get_ili_data()
+
+both_flu <- readRDS("./flu_forecast_23/both_flu.rds")
+ILINet <- readRDS("./flu_forecast_23/ILINet.rds")
 
 state <- "US"
 sample(unique(ILINet$region), 6)
@@ -15,21 +23,30 @@ dat <- get_stan_data(ILINet, both_flu, s_region = "Texas")
 stan_dat <- dat[[1]]
 
 mean_ili <- ILINet %>%
+  left_join(states, by = "region") %>% 
   filter(season %in% c(2010:2022)) %>%
-  group_by(region, season_week) %>%
+  group_by(loc_abb, season_week) %>%
   summarise(
     mean_ili = mean(unweighted_ili, na.rm = TRUE)
   ) %>%
-  filter(region %in% c("District of Columbia", "Alabama", "Texas",
-                       "Ohio", "Iowa", "Montana")) %>%
-  mutate(region = factor(region, levels=c("District of Columbia", "Alabama",
-                                         "Texas", "Ohio", "Iowa", "Montana")))
+  filter(!is.na(loc_abb))
+  # filter(region %in% c("District of Columbia", "Alabama", "Texas",
+  #                      "Ohio", "Iowa", "Montana")) %>%
+  mutate(region = factor(region))
+  # mutate(region = factor(region, levels=c("District of Columbia", "Alabama",
+  #                                        "Texas", "Ohio", "Iowa", "Montana")))
 
 ILINet %>% filter(region == state, season >= 2010) %>% 
   ggplot() +
   geom_line(aes(x = season_week, y = unweighted_ili),
[…]

### 2025-08-11 make_images/ili_prior_pictures.R
ili priors plots update

@@ -5,6 +5,7 @@ library(truncnorm)
 
 weeks <- c(14, 20, 26)
 mods <- c("asg_nm", "asg_disc2_nm", "sir", "sir_disc2")
+model <- c("ASG", "ASGD", "SIR", "SIRD")
 
 
 asg14 <- readRDS("../flu_forecast_23/asg_nm_summary.rds") %>% 
@@ -26,7 +27,28 @@ asg_model_post <- rbind(asg14, asgd14) %>%
 sir_model_post <- rbind(sir14, sird14) %>% 
   select(model, variable, q2.5, q97.5) %>% 
   mutate(variable = as.character(variable)) 
-  
+
+
+for_dir <- "../flu_forecast_23/"
+all_model_post <- data.frame()
+
+for (i in 1:length(weeks)) {
+  for (j in 1:length(mods)) {
+    dir <- paste(for_dir, mods[j], "_", weeks[i], "_", "summary.rds", sep = "")
+    post <- readRDS(dir)
+    post$model <- model[j]
+    post$week <- as.character(weeks[i])
+    all_model_post <- rbind(all_model_post, post)
+  }
+}  
+
+all_model_post %>% 
+  group_by(model, week) %>% 
+  summarise(maw = round(max(rhat, na.rm = TRUE), 4),
+            mis = min(ess_bulk, na.rm = TRUE))
+
+all_posts <- all_model_post %>% 
+  select(model, week, variable, q2.5, q97.5)
 
 variables <- c("I0", "beta", "rho")
 sm <- c(.005, .8, .68)
@@ -47,15 +69,19 @@ sir_bounds <- apply(sir_samps, MARGIN = 2,
 colnames(sir_bounds) <- c("q2.5", "q97.5")
 sir_bounds <- as.data.frame(sir_bounds)
 sir_priors <- cbind(model = "prior", variable = variables, sir_bounds)
+sir_priors$week <- "prior"
+sir_priors <- sir_priors %>% 
+  select(model, week, variable, q2.5, q97.5)
 
+# one <- rbind(sir_model_post); one$week <- "14"
+# two <- rbind(sir_model_post); two$week <- "20"
+# three <- rbind(sir_model_post, sir_priors); three$week <- "26"
 
-one <- rbind(sir_model_post); one$week <- "14"
-two <- rbind(sir_model_post); two$week <- "20"
-three <- rbind(sir_model_post, sir_priors); three$week <- "26"
[…]

### 2025-08-11 supplemental_res.tex
update supplementary material

@@
 \usepackage{tabularx} %for the spacing in tabular
+\usepackage{newtxmath} %for making indicator function
+\usepackage{changepage} %for adjusting the margins
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Figure \ref{fig:ili_vs_week} shows ILI data from five states and the District of Columbia, locations which received particular attention in \cite{osthus2021multiscale}.
-The plots include the ILI data for all seasons from 2010 to 2023 in grey, and the black line is the per week ILI average over seasons.
-The patterns in the individual states are similar to the national level plots in that the ILI rises in the fall and winter until it peaks and descends as the spring and summer progress.
-For these locations ILI regularly peaks, either locally or globally, at or near week 22.
+Figure \ref{fig:ili_vs_week} shows ILI data all 50 states in the US. 
+The plots include the ILI data for all seasons from 2010 to 2023 in grey, 
+and the black line is the per week ILI average over seasons. 
+The patterns in the individual states are similar to the national 
+level plots in that the ILI rises in the fall and winter until it peaks 
+and descends as the spring and summer progress. 
+It can also be seen that neighboring states tend to have similar ILI levels.
+At nearly all states, the ILI typically peaks, either locally or globally, 
+at or near 
+week 22.
 
@@
     \centering
-    \includegraphics[scale=.5]{Images/ili_vs_week.png}
-    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in five different states and the District of Columbia for seasons 2010 to 2023.
-Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.
-     Plots include lines for ILI\% from the 2010 flu season to 2023 (grey) and for the weekly ILI averaged over all seasons (black).}
+    \includegraphics[scale=.45]{Images/ili_vs_week.png}
+    \caption{Percentage of outpatient visits with an influenza-like illness 
+    (ILI) for the 50 US states seasons 2010 to 2023. Week 1 is the first week 
+    of August of the year the flu season begins and the last week of the 
+    season is the last week of July of the following year.
+     Plots include lines for ILI\% from the 2010 flu season to 2023 (grey) 
+     and for the weekly ILI averaged over all seasons (black).}
     \label{fig:ili_vs_week}
@@
 
-Figure \ref{fig:hosp_vs_week} shows the 2022 and 2023 weekly hospitalizations for the same states from \ref{fig:ili_vs_week}.
-Similar to the national data, the peak in 2022 came early compared to the peak of 2023.
-Comparing figures \ref{fig:ili_vs_week} and \ref{fig:hosp_vs_week} shows that ILI and hospitalizations share the similar pattern of increasing to a peak in the winter and decreasing thereafter.
+Figure \ref{fig:hosp_vs_week} shows the 2022 and 2023 weekly hospitalizations 
+for the 50 US states. Similar to the US national data, 
+the state peaks in 2022 came early compared to the peak of 2023. 
+This figure shows hospitalization levels may similar for neighboring states,
+but also that state population plays a role in the number of hospiatlizations.
+Comparing figures 
+\ref{fig:ili_vs_week} and \ref{fig:hosp_vs_week} shows that ILI and 
+hospitalizations share the similar pattern of increasing to a peak in the 
+winter and decreasing thereafter.
 
@@
     \centering
-    \includegraphics[scale=.55]{Images/hosp_vs_week.png}
-    \caption{Weekly hospitalization counts for five states and DC for the 2022 (grey) and 2023 (black) flu seasons}
+    \includegraphics[scale=.5]{Images/hosp_vs_week.png}
+    \caption{Weekly hospitalization counts for the 50 US states for the 2022 
+    (grey) and 2023 (black) flu seasons}
     \label{fig:hosp_vs_week}
@@
 
-% Figure \ref{fig:ili_vs_hosp} shows scatter plots with ILI\% on the $x$-axes and hospitalizations on the $y$-axes, revealing a positive somewhat linear relationship between the two variables.
-This relationship motivates the forecast models outlined in the next section.
-% 
-% \begin{figure}[hbt!]
-%     \centering
-%     \includegraphics[scale=.6]{Images/ili_vs_hosp.png}
-%     \caption{Weekly flu confirmed hospitalization counts ($y$-axis) for five states and the District of Columbia for the 2022 and 2023 flu seasons plotted against ILI\% ($x$-axis)}
-%     \label{fig:ili_vs_hosp}
-% \end{figure}
+\section{Prior distributions} \label{app:B_prior}
 
-% \section{Additional results plots from simulation study}
-
-Figures \ref{fig:overall_scores} - \ref{fig:logs_by_season} show boxplots of the CRPS and logarithmic score (LogS) for the four models for 500 simulation replicates.
-Figure \ref{fig:logs_by_week_horizon} shows
-boxplots of the LogS faceted by week and horizon.
-Figure \ref{fig:overall_scores} shows that the variation of overall CRPS is smallest for the ASG models and larger for the SIR models.
-The median scores for the SIR models also appears slightly higher than for the ASG models.
-When faceted by season in figure \ref{fig:crps_by_season}, the boxplots of the CRPS often show the same pattern for ASG and SIR models but not always.
-For example, the bulk of CRPS values for the SIRD model in 2019 appears to have smaller variation than the other models.
-The LogS plots show similar results to the CRPS plots, though there are some differences.
-For example, SIRD in figure \ref{fig:logs_by_season} tends to show smaller LogS variation relative to the other three models than is seen by the CRPS of SIRD in figure \ref{fig:crps_by_season}.
-
-The LogS is defined in (\ref{eq:logs2}) where $f(\cdot)$ is the forecast density function and $y^*$ is the observed value.
-To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation, and the \texttt{scoringRules} package in \texttt{R} is used for the calculation \cite[]{jordan2019scoring}.
+The prior distributions under the SIR model are in (\ref{eq:sir_prior}).
+Because we set $S_{0s} = 0.9$, prior distributions were assigned only to
+$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery
+rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the
+indicator function for values within the set $A$.
 
 \begin{equation}
-    \text{LogS}(f, y^*) = -\text{log}(f(y^*))
-    \label{eq:logs2}
+\begin{aligned}
+    \label{eq:sir_prior}
+        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
+        \beta_s &\sim N^+(0.8, 0.3) \\
+        \rho_s &\sim N^+(0.68, 0.08)
+\end{aligned}
 \end{equation}
+For the ASG model, the MLE 
+$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ 
+was calculated. % and $\hat{\lambda}_s$ was accepted as a fixed value. 
+These estimates were used as starting values for posterior sampling. 
+The transformation 
+$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ 
+was made, and a prior distribution was assigned to $T(\theta_s)$.
+The prior distributions for the ILI model under the ASG function 
+without discrepancy are shown 
+in (\ref{eq:ili_prior}). These are slightly informative priors because for 
+most parameters we have an idea what reasonable values may be. Here 
+$m = (-4.5, 0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.3, 0.2, 5, 2, 2)$
+where $\text{diag}(\cdot)$ 
+is the diagonal matrix for the given entries. When modeling ASG with 
+discrepancy, the first entry in $m$ was replaced with $\hat{\lambda}$ the 
+mean over seasons of MLEs for that parameter, and in the matrix $C$
+the first diagonal 
+variance parameter of $0.3$ was replaced with $0.01$. This was done to alleviate
+the identifiability issues of $\lambda_s$ that are indtroduced when including 
+discrepancy in the ILI model.
+
+\begin{equation}
+\begin{aligned}
+\label{eq:ili_prior}
+                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
+                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
+                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
+                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
+\end{aligned}
+\end{equation}
+Parameters shared by both the SIR and ASG models are the scale parameter 
+$\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and 
+$\sigma_{\gamma_W}^2$. These priors are in (\ref{eq:shared_ili_priors}).
+
+\begin{equation}
+\begin{aligned}
+    \label{eq:shared_ili_priors}
+        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
+        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
+        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
+        \sigma_{\upsilon_{s,w}} &\sim N^+(0, .1^2)
+\end{aligned}
+\end{equation}
+Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
+we selected an informative prior distribution by first estimating  
+$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
+calculating the MLE for $\theta_s$. 
+% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
+%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
+$\widehat{ILI}_{s,w}$ was then predicted such that 
+$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
+Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
+seasons of the differences 
+$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
+When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
+calculated using the \texttt{mle2} function in the \texttt{bblme} package 
+\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
+package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
+function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
+function.
+The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
+
+\begin{equation}
+\begin{aligned}
+\label{eq:shared_hosp_priors}
+        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
+        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
+        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
+        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
+        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
+        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
+\end{aligned}
+\end{equation}
+
+\section{Posterior distribution plots for select parameters}
+\label{app:C_post}
+
+The figures in this section show 95\% credible intervals for model parameters 
+under the several modeling schemes along with the prior distributions assigned
+to the parameters. Figure \ref{fig:posterior_theta_sir} is for parameters 
+unique to the SIR ILI model. Figure \ref{fig:posterior_theta}  
+show parameters unique to the ASG models. 
+Figure \ref{fig:sir_asg_shared} is for parameters shared by SIR and ASG 
+models, including parameters used for modeling discrepancy. Figures 
+\ref{fig:hosp_lin_params} and \ref{fig:hosp_nus_param} are for parameters 
+used in hospitalization modeling.
 
@@
     \centering
-% \begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[scale=.3]{Images/overall_crps.png}
-    % \caption{Boxplots of CRPS for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-% \end{subfigure}%
-% \begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[scale=.3]{Images/overall_logs.png}
-    % \caption{Boxplots of LogS for the four ILI models over all seasons, weeks, and horizons in the simulation study}}
-% \end{subfigure}
-\caption{Boxplots of the continuous ranked probability score (CRPS) (left) and logarithmic score (LogS) (right) for the four ILI models over all seasons, weeks, and horizons in the simulation study}
-\label{fig:overall_scores}
-\end{figure}
-
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[scale=.5]{Images/crps_by_season.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
-    \label{fig:crps_by_season}
-\end{figure}
-
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[scale=.4]{Images/logs_by_week_horizon.png}
-    \caption{Boxplots of the logarithmic score (LogS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season.}
-    \label{fig:logs_by_week_horizon}
-\end{figure}
-
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[scale=.5]{Images/logs_by_season.png}
-    \caption{Boxplots of logarithmic score (LogS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
-    \label{fig:logs_by_season}
-\end{figure}
-
-% \section{Additional results plots for 2023-24 flu analysis}
-
-In figure \ref{fig:lwis_by_traj_loc}, it appears that both the SIR and ASG models which include discrepancy perform slightly better than the models which do not.
-
-\begin{figure}[hbt!]
-
-    \centering
-    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
-    \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 US states, PR, DC, and national level forecasts at each week during the 2023 flu season.
-Scores are averaged over all horizons 1-4 weeks ahead.
-Scores are faceted by ILI model function (columns) and by whether or not discrepancy modeling was included (rows).
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-    \label{fig:lwis_by_traj_loc}
-\end{figure}
-
-\section{Posterior distribution plots for select parameters}
-\label{app:B_prior}
-
-The figures in this section show 95\% credible intervals for model parameters under the several modeling schemes along with the prior distributions assigned to the parameters.
-Figure \ref{fig:posterior_theta_sir} is for parameters unique to the SIR ILI model.
-Figures \ref{fig:posterior_theta} and \ref{fig:posterior_zeta} are for parameters unique to the ASG models.
-Figure \ref{fig:sir_asg_shared} is for parameters shared by SIR and ASG models, including parameters used for modeling discrepancy.
-Figures \ref{fig:hosp_lin_params} and \ref{fig:hosp_nus_param} are for parameters used in hospitalization modeling.
-
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[scale=.6]{Images/posterior_theta_sir.png}
-    \caption{Posterior 95\% credible intervals from ILI model for parameters of SIR differential equations.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The green is from the posterior where discrepancy is not modeled, and the blue is from the model where it is.
-The red interval is the 95\% interval of the prior distribution.}
+    \includegraphics[scale=.6]{Images/posterior_sir.png}
+    \caption{Posterior 95\% credible intervals from ILI model for 
+    parameters of SIR differential equations. Shown are intervals 
+    from the US model of ILI for weeks 14, 20, and 26 and coloured by 
+    SIR and SIRD
+    models and the assigned prior distributions.}
     \label{fig:posterior_theta_sir}
@@
     \centering
-    \includegraphics[scale=.6]{Images/posterior_theta.png}
-    \caption{Posterior 95\% credible intervals from ILI model for parameters of ASG function.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The red is from the posterior where discrepancy is not modeled, and the green is from the model where it is.
-The blue interval is the 95\% interval of the prior distribution.}
+    \includegraphics[scale=.6]{Images/posterior_asg.png}
+    \caption{Posterior 95\% credible intervals from ILI model for parameters of 
+    ASG function. Shown are intervals from the US model of ILI for weeks 14, 
+    20, and 26 and coloured by ASG and ASGD models and the assigned 
+    prior distributions.}
     \label{fig:posterior_theta}
@@
 \centering
-    \includegraphics[scale=.6]{Images/posterior_zeta.png}
-    \caption{Posterior 95\% credible intervals from ILI model for variance parameters of the ASG function.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The red is from the posterior where discrepancy is not modeled, and the green is from the model where it is.
-The blue interval is the 95\% interval of the prior distribution.}
-    \label{fig:posterior_zeta}
-\end{figure}
-
-\begin{figure}[hbt!]
   \centering
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth]{Images/posterior_kappa.png}
+  \includegraphics[width=.95\linewidth]{Images/posterior_comb.png}
   % \caption{95\% posterior credible intervals from ILI model for scale parameter of the model distribution.
@@
   % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth]{Images/posterior_sigma_gamma.png}
-  % \caption{95\% posterior credible intervals from ILI model for scale parameter of modeled discrepancy.
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-The blue is from the SIRD model and red from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Posterior 95\% credible intervals from ILI model for scale parameter $\kappa_s$.
-Blue is from the SIR model, purple from SIRD, red from ASG and yellow from ASGD (left).
-Posterior 95\% credible intervals from ILI model for scale parameter of modeled discrepancy $\sigma_{\gamma}$ (right).
-Shown are intervals from the US model of ILI for weeks 14, 20, and 26.
-Blue is from the SIRD model and red from ASGD.
-The green interval is the 95\% interval of the prior distribution.}
+\caption{Posterior 95\% credible intervals from ILI model for scale 
+parameter $\kappa_s$. 
+Blue is from the SIR model, purple from SIRD, 
+red from ASG and yellow from ASGD (left).  
+Posterior 95\% credible intervals from ILI model for scale parameter of 
+modeled discrepancy $\sigma_{\gamma}$ (right). Shown are intervals from the 
+US model of ILI for weeks 14, 20, and 26. Blue is from the SIRD model and red 
+from ASGD. The green interval is the 95\% interval of the prior distribution.}
 \label{fig:sir_asg_shared}
 \end{figure}
-
-% \begin{figure}[hbt!]
-% 
-% \begin{subfigure}
-%   \includegraphics[width=.49\linewidth]{Images/alpha0_post.png}
-%   % \caption{}
-%   % \label{MLEDdet}
-% \end{subfigure}\hfill % <-- "\hfill"
-% \begin{subfigure}
-%   \includegraphics[width=.49\linewidth]{Images/alpha1_post.png}
-%   % \caption{}
-%   % \label{energydetPSK}
-% \end{subfigure}
-% 
-% \medskip % create some *vertical* separation between the graphs
-% \begin{subfigure}
-%   \includegraphics[width=.49\linewidth]{Images/alpha2_post.png}
-%   % \caption{}
-%   % \label{velcomp}
-% \end{subfigure}\hfill % <-- "\hfill"
-% \begin{subfigure}
-%   \includegraphics[width=.49\linewidth]{Images/phi_post.png}
-%   % \caption{}
-%   % \label{estcomp}
-% \end{subfigure}
 
@@
 \newpage
-\section{Model forecast comparison}
+\section{Model combinations forecast comparison}
 
-\begin{figure}[hbt!]
-\centering
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
-season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, 
-and the hospitalization models are all normally distribution. The figure 
-includes hospitalization forecasts where ILI is a linear predictor (left) 
-and where ILI is a quadratic predictor (right). 50\% predictive intervals 
-are pink and 95\% predictive intervals are red.}
-\label{fig:normal_flu_forecasts}
-\end{figure}
+We fit 24 forecast models for each location for all 30 weeks, and for each 
+week forecast 1-4 week ahead hospitalizations. The 24 models included all 
+combinations of ASG, ASGD, SIR, and SIRD ILI models, hospitalization
+models where the assumed distribution of 
+hospitalizations was normal (NORM), lognormal (LNORM), 
+and location-scale t (LST)
+hospitalization models, and where ILI was assumed to be a linear or 
+quadratic predictor of hospitalizations.
+
+Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the 
+season for all 24 models of US hospitalizations. LWIS scores are grouped by 
+ILI model, hospitalization model distribution, and by the linear or quadratic 
+modeling. Here, the smaller the LWIS the better. The weeks around the holiday 
+week 22 are highlighted by a grey band. 
+In several cases, there appears to be a turning point in performance at or 
+near week 22. For most ASG models, the models which include discrepancy tend 
+to outperform those which do not beginning near week 22 and for the rest of 
+the season.
+For the SIR models, the discrepancy does not always lead to better forecasts, 
+but in for the LNORM hospitalization models, including discrepancy greatly 
+improves forecasts around the holiday week 22. The LNORM LWIS scores tend to 
+be higher than for LST and NORM hospitalization models, but figure 
+\ref{fig:lwis_by_dist_loc} and table \ref{tab:forecast_scores} suggest that 
+the LNORM hospitalization model outperforms LST and NORM for most states and 
+territories.
 
@@
     % \centering
-    \includegraphics[scale = .6]{Images/lwis_ln_us_full_season.png}
+    \includegraphics[scale = .5]{Images/lwis_ln_us_full_season.png}
     \caption{Each plot shows the log weighted interval score (LWIS) for every 
@@
 
-% \begin{figure}
-%     
-%     % \centering
-%     \includegraphics[scale = .5]{Images/lwis_us_full_season.png}
-%     \caption{Each plot shows the log weighted interval score (LWIS) for every week of the 2023 flu season with scores for models including discrepancy in the ILI model (top) and excluding discrepancy (bottom).
-Scores are separated by hospitalization distribution family and by ILI as a linear or quadratic predictor.
-Scores for models with an ASG ILI model are above while those with an SIR model are below.
-The lighter the shade, the lower the LWIS with low LWIS being better.}
-%     \label{fig:us_lwis}
-% \end{figure}
+Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by 
+LWIS, but all 53 locations are included. The LWIS is shown across the whole 
+season faceted by hospitalization model distribution. Overall, the weeks 
+leading up to week 22 are the most difficult to forecast, but the LNORM model 
+appears to perform the better than the other models during the weeks around 
+the holiday week 22. 
+The overall LWIS, calculated as the mean LWIS over all locations and weeks, 
+for all 
+24 models is shown in table \ref{tab:forecast_scores}. The overall best 
+performing forecast model combination is ASGD with ILI as a quadratic 
+predictor of normally distributed hospitalizations. Forecasts from this 
+model combination are compared with forecasts from FluSight in the main text.
+
+In figure \ref{fig:lwis_by_traj_loc}, it appears that both the SIR and ASG 
+models which include discrepancy perform better than the models 
+which do not include discrepancy.
 
@@
     \centering
-    \includegraphics[scale = .7]{Images/lwis_by_dist_loc.png}
+    \includegraphics[scale = .6]{Images/lwis_by_dist_loc.png}
     \caption{Each plot shows the log weighted interval scores (LWIS) for all 50 
@@
 
-Figure \ref{fig:lwis_by_dist_loc} also shows weekly forecast performance by 
-LWIS, but all 53 locations are included. The LWIS is shown across the whole 
-season faceted by hospitalization model distribution. Overall, the weeks 
-leading up to week 22 are the most difficult to forecast, but the LNORM model 
-appears to perform the better than the other models during the weeks around 
-the holiday week 22. 
-The overall, calculated as the mean LWIS over all locations and weeks, for all 
-24 models is shown in table \ref{tab:forecast_scores}. The first main takeaway 
-is that the top four performing models are SIR ILI models with the LNORM 
-hospitalization models. The next main takeaway is that overall the models 
-which include discrepancy outperformed the models without discrepancy.
+\begin{figure}[hbt!]
+
+    \centering
+    \includegraphics[scale = .7]{Images/lwis_by_traj_loc.png}
+    \caption{Each plot shows the log weighted interval scores (LWIS) 
+    for all 50 US states, PR, DC, and national level forecasts at each week 
+    during the 2023 flu season. Scores are averaged over all horizons 1-4 
+    weeks ahead. Scores are faceted by ILI model function (columns) and by 
+    whether or not discrepancy modeling was included (rows). The lighter the 
+    shade, the lower the LWIS with low LWIS being better.}
+    \label{fig:lwis_by_traj_loc}
+\end{figure}
 
@@
 and horizons. The scores in the first two rows are for linear models, and the 
-scores in the third and fourth rows are for quadratic models. The lowest WISs 
+scores in the third and fourth rows are for quadratic models. The lowest LWISs 
 are bolded.}
@@
   \midrule
-  Linear & No Disc & 0.366 & 0.392 & 0.365 & \textbf{0.355} & 0.416 & 0.394 &\\ 
-   & Disc & 0.358 & 0.391 & 0.365 & \textbf{0.343} & 0.382 & 0.354 &\\
+  Linear & No Disc & 0.390 & 0.419 & 0.411 & 0.382 & 0.447& 0.442 &\\ 
+   & Disc & 0.358 & 0.400 & 0.391 & 0.378 & 0.413 & 0.396 &\\
   \midrule
-  Quadratic & No Disc & 0.366 & 0.398 & 0.377 & \textbf{0.355} & 0.401 & 0.377 &\\ 
-   & Disc & 0.358 & 0.390 & 0.369 & \textbf{0.343} & 0.378 & 0.356 &\\     
+  Quadratic & No Disc & 0.390 & 0.416 & 0.384 & 0.381 & 0.423 & 0.394 &\\ 
+   & Disc & 0.358 & 0.371 & \textbf{0.341} & 0.379 & 0.398 & 0.360 &\\     
   \bottomrule
@@
 
-\section{FluSight forecast competition scoring results} \label{app:A_rwis}
+\section{FluSight forecasts WIS comparison}
 
-In the CDC flu forecast competition, a baseline forecast was made to which all other forecasts were compared.
-The baseline forecast for a given state and week had as a median the most recent observed hospitalization count.
-The uncertainty was based on differences between previous hospitalizations, and is similar to the baseline forecast used in previous flu forecasting seasons and in the COVID-19 hub \cite[]{mathis2024evaluation, cramer2022evaluation}.
-The RWIS for one model was calculated by first taking the ratio of the average LWIS paired with every other model.
-This was then diveded by the same ratio for the baseline forecasts (see methods in \cite{mathis2024evaluation} for details).
- An RWIS less than 1 indicates the forecast outperformed the baseline forecast.
- Figures \ref{fig:rwis_ili_tile} and \ref{fig:rwis_hosp_tile} show the mean RWIS of hospitalization forecasts over all 24 models from the main manuscript for all locations and weeks of the 2023 flu season.
-Some similar patterns to those noted in section the manuscript emerge, including a slightly better performance by forecasts including discrepancy modeling over those which do not include discrepancy modeling and better overall performance by the SIR models than by the ASG forecast models.
-Interestingly, figure \ref{fig:rwis_hosp_tile} shows the LNORM model showing especially poor RWIS performance about three quarters into the season.
-This was not seen in the LWIS in the main manuscript.
-%figure \ref{fig:lwis_by_dist_loc}.
+Figures \ref{fig:wis_cover_sum} and \ref{fig:state_and_date_wis} are 
+similar to figures \ref{fig:lwis_cover_sum} and \ref{fig:state_and_date_lwis}
+except that they show WIS scores instead of LWIS scores. The results of 
+foreacst comparison for WIS scores is very similar to those of the LWIS in 
+terms of how well the $ASGD_NORM2$ forecasts did relative to the others from
+the FluSight competition.
 
 \begin{figure}[hbt!]
-    \centering
-    \includegraphics[scale=.7]{Images/rwis_ili_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
-The plot is separated by the four ILI models.
-Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_ili_tile}
+
+    % \centering
+    \begin{adjustwidth}{-1.2cm}{-1.5cm}
+    \includegraphics[scale = .54]{Images/wis_cover_sum.png}
+    \caption{a) Averaged over location and horizon weighted interval score
+    (WIS) for each week during the season for the 21 models being compared. 
+    Each grey line represents one model, but the ASGD\_NORM2 (blue) and 
+    FluSight-baseline (red) are colored to stand out.
+    b) Boxplots of all WIS scores for 1, 2, 3, and 4-week ahead horizons 
+    for the ASGD\_NORM2 (blue) and FluSight-baseline (red) models. 
+    To improve vizualization, the $y$-axis
+    is on the log scale.}
+    \end{adjustwidth}
+    \label{fig:wis_cover_sum}
 \end{figure}
@@
 \begin{figure}[hbt!]
+\begin{adjustwidth}{-1.2cm}{-1.5cm}
+    % \begin{minipage}{\linewidth}
     \centering
-    \includegraphics[scale=.6]{Images/rwis_hosp_mod.png}
-    \caption{RWIS averaged over all 24 models and horizons for forecasts across all locations and weeks.
-The plot is separated by the three distribution choices for the hospitalization model.
-Dark blue is lower RWIS, and dark tan is higher RWIS where white is RWIS = 1}
-    \label{fig:rwis_hosp_tile}
+    \makebox[\linewidth][l]{
+    \includegraphics[scale = .5]{Images/state_and_date_wis.png}
+    }
+    \end{adjustwidth}
+    \caption{Ratio of average weighted interval score (WIS) for ASGD\_NORM2 
+    to that of on competing FluSight model. Scores are averaged
+    over all targets within indicated location (top) or week (bottom) and 
+    only include forecasts shared by both models. Lower scores (blue) indicate
+    superior performance by the ASGD\_NORM2 model and higher scores (red)
+    indicate superior performance by competing model.}
+    \label{fig:state_and_date_wis}
+    % \end{minipage}
+
 \end{figure}
 
-Figure \ref{fig:state_median_points} shows the mean over weeks RWIS for the SIR and ASG forecast models with an RWIS less than 1 for the most locations.
-The plot on the left is from a forecast by an SIRD ILI model and a quadratic NORM hospitalization model.
-This forecast model had an RWIS less than 1 for 37 out of 53 locations.
-The plot on the right is from a forecast by an ASGD ILI model and linear NORM hospitalization model.
-This model had an RWIS less than 1 for 34 out of 53 locations.
-
-\begin{figure}[hbt!]
-\centering
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth]{Images/sir_disc_hosp_sq_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}
-  \centering
-  \includegraphics[width=.49\linewidth]{Images/asg_disc_hosp_ar1_rwis.png}
-  % \caption{A subfigure}
-  % \label{fig:sub2}
-\end{subfigure}
-\caption{RWIS over the whole 2023 flu season for all 53 locations for hospitalization forecasts from an SIRD quadratic hospitalization model with normal errors (left) and forecasts from an ASGD linear hospitalization model with normal errors (right).
-An RWIS less than 1 (left of vertical grey line) indicates the model forecasts outperformed the baseline forecasts for that particular region.}
-\label{fig:state_median_points}
-\end{figure}
-
-\section{Prior distributions}
-
-The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
-Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
-$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
-rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
-indicator function for values within the set $A$.
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:sir_prior}
-        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
-        \beta_s &\sim N^+(0.8, 0.3) \\
-        \rho_s &\sim N^+(0.68, 0.08)
-\end{aligned}
-\end{equation}
-For the ASG model, the MLE 
-$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ 
-was calculated. % and $\hat{\lambda}_s$ was accepted as a fixed value. 
-These estimates were used as starting values for posterior sampling. 
-The transformation 
-$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ 
-was made, and a prior distribution was assigned to $T(\theta_s)$.
-The prior distributions for the ILI model under the ASG function are shown 
-in (\ref{eq:ili_prior}). These are slightly informative priors because for 
-most parameters we have an idea what reasonable values may be. Here 
-$m = (-4.5, 0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.3, 0.2, 5, 2, 2)$
-where $\text{diag}(\cdot)$ 
-is the diagonal matrix for the given entries.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:ili_prior}
-                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
-                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
-                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
-                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
-\end{aligned}
-\end{equation}
-Parameters shared by both the SIR and ASG models are the scale parameter 
-$\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and 
-$\sigma_{\gamma_W}^2$. These priors are in (\ref{eq:shared_ili_priors}).
-
-\begin{equation}
-\begin{aligned}
-    \label{eq:shared_ili_priors}
-        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
-        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
-        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
-        \sigma_{\upsilon_{s,w}} &\sim N^+(0, .1^2)
-\end{aligned}
-\end{equation}
-Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
-we selected an informative prior distribution by first estimating  
-$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
-calculating the MLE for $\theta_s$. 
-% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
-%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
-$\widehat{ILI}_{s,w}$ was then predicted such that 
-$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
-Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
-seasons of the differences 
-$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
-When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
-calculated using the \texttt{mle2} function in the \texttt{bblme} package 
-\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
-package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
-function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
-function.
-The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).
-
-\begin{equation}
-\begin{aligned}
-\label{eq:shared_hosp_priors}
-        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
-        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
-        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
-        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
-        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
-\end{aligned}
-\end{equation}
 \end{supplement}

### 2025-08-11 make_images/make_analysis_images.R

@@ -14,8 +14,7 @@ date_weeks <- both_flu %>%
 ILINet <- get_ili_data()
 
 
-setwd("~/Documents/Research/nonlinear_flu_forecast/models")
-
+setwd("./flu_forecast_23/fin_hosp_forecasts/")
 #season_week 10 is "2023-10-14"
 
 target_data <- read.csv("../../FluSight-forecast-hub/target-data/target-hospital-admissions.csv") %>%
@@ -56,65 +55,66 @@ baseline_wis <- baseline_forecasts %>%
   ) %>%
   mutate(model = base)
 
-team_scores <- data.frame()
-for (sub_date in sub_dates) {
-  for (i in 1:length(models)) {
-    forecast_file <- paste(models[i], "/", sub_date, "-",
-                           models[i], ".csv", sep = "")
-    
-    if (file.exists(forecast_file) == FALSE) next
-    forecast <- read.csv(forecast_file) %>%
-      mutate(location = as.character(location),
-             output_type_id = as.character(output_type_id)) %>%
-      mutate(location = ifelse(nchar(location) < 2, paste0("0",location),
-                               location)) %>%
-      filter(output_type == "quantile", horizon != -1) %>%
-      mutate(location = as.character(location)) %>% # %>% select(reference_date, horizon, target, target_end_date, location, output_type, output_type_id, value)
-      mutate(value = ifelse(value < 0, 0, value)) #%>%
-    # mutate(target_end_date = as.character(date(target_end_date) - 7))
-    
-    for_wis <- forecast %>%
-      filter(horizon %in% 0:3) %>%
-      left_join(target_data,
-                by = c("location", "target_end_date" = "date")) %>%
-      mutate(output_type_id = as.numeric(output_type_id),
-             value = as.numeric(value),
-             true_value = as.numeric(true_value))
-    
-    # if (mean(is.na(for_wis$location_name)) == 0) next
-    if (nrow(for_wis) == 0) next
-    wis <- for_wis %>%
-      filter(is.na(location_name) == FALSE) %>%
-      group_by(reference_date, location_name, horizon) %>%
-      reframe(
-        wis = weighted_interval_score(output_type_id,
-                                      value, true_value),
-        lwis = weighted_interval_score(output_type_id,
-                                       log(value + 1), log(true_value + 1)),
[…]

### 2025-08-13 response.tex
draft editing

@@
 with discrepancy components,
-rather than estimating all parameters in a Bayesian way,
+rather than estimating all parameters through posterior updating,
 identifiability was encouraged by setting
@@
 ``The SIR is a mechanistic model used to capture transmission of a particular
-disease, and the ASG is a smooth function which may approximate the 
-same shape as the ILI trajectory. As ILI may be influenced by more 
+disease, and since ILI may be influenced by more 
 diseases than influenza, the disease transmission described by the SIR model
-can only partially contribute to the true ILI data-generating mechanism[44].
-Though the ASG function may be able to roughly capture ILI, there may still be 
+can only partially contribute to the true ILI data-generating mechanism
+[44].
+
+Ulloa [53]
+chose to use the asymmetric Gaussian (ASG) function to model ILI data.
+The ASG is a smooth function which may approximate the 
+same shape as the ILI trajectory.
+The ASG function has more parameters and is thus slightly more flexible than
+the SIR model, and thus may be able to roughly capture ILI.
+However, there may still be 
 systematic behavior which it cannot capture. Thus when using either the SIR
@@
 The bottom row shows scatterplots of the difference between hospitalizations and 
-1 week lags scaled 
+1 week lags
 by ILI percentage at the US national level. The plots in the right column
@@
 However, when including a 
-single autoregressive lag of hospitalizations in the model, much of the time 
+single AR lag of hospitalizations in the model, much of the time 
 correlation is
 accounted for as shown by the tighter and more frequent crossing of lines
-in the bottom row. Higher than order 1 autoregressions showed little or no 
-improvement over order 1, hence we consider the ARX(1) reasonably well 
-suited for modeling ILI and hospitalzations."}
+in the bottom row. Models with more than 1 AR lag showed little or no 
+improvement over 1 lag."}
 
@@
 by modeling log-hospitalizations or by assuming hospitalizations followed a
-distribution truncated at 0, but the cost of doing so led to poorer forecasts
-or problems drawing from the posterior distribution.
+distribution truncated at 0, but these models led to poorer forecasts
+and issues drawing from the posterior distribution.
 For the forecast competition analysis in 
@@
 set to 0 to reflect realistic values of hospitalizations and comply with the 
-FluSight forecasting rules. This ultimately made almost no difference in 
-results."}
+FluSight forecasting rules."}
 
@@
 
-``With the goal of producing a probablistic forecast, Bayesian modeling is a 
+``Because we desire probabilistic forecasts, Bayesian modeling is a 
 natural framework for describing forecast uncertainty. 
@@
 forecast and not 
-necessarily in estimating model parameters, decisions on certain prior
-distributions and how the models were fit proved critical in producing good 
-forecasts. After defining the ILI and hospitalization models in this section, 
+necessarily in estimating model parameters, certain modeling decisions and 
+prior distribution selections proved critical in producing good 
+forecasts. After defining the ILI and hospitalization models, 
 we describe the 
@@
 1 week 
-lags scaled 
+lags
 by ILI percentage at the US national level. The plots in the right column
@@
 with discrepancy components,
-rather than estimating all parameters in a Bayesian way,
+rather than estimating all parameters through posterior updating,
 we set the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})

### 2025-08-13 manuscript_res.tex
update supplementary material
draft editing

@@
 flu seasons 2010 to 2023. For most seasons there are 52 weeks listed, but for 
-the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 
-Sundays during those seasons. 
+the 2010, 2015, and 2021 seasons there are 53 weeks because those
+seasons had 53 Sundays. 
 To better align with the flu behavior, week 1 is set as the first week of 
 August and week 52 or 53 is the last week in July of the following year. For 
-example week 1 of the 2013 season corresponds to the first week of August 2013, 
+example, week 1 of the 2013 season corresponds to the first week of August 2013, 
 and week 52 of the same season corresponds to the last week of July 2014. This 
@@
 \cite[]{ewing2017contact, garza2013effect}.
-The only seasons when there was not a peak at week 22 were season 2022 where 
+The only seasons when there was not a peak at week 22 were season 2022 when 
 the season peak occurred particularly early and season 2020 which was greatly 
@@
 The bottom of figure \ref{fig:us_ili} shows the weekly national 
-hospitalizations for the 2022 and 2023 flu seasons. These plots show 
-similarities to the ILI plots in the top of the figure in that at the early 
+hospitalizations for the 2022 and 2023 flu seasons. These plots show a 
+trend similar to the ILI plots in the top of the figure in that during the early 
 weeks of the season hospitalizations are low, but they increase in the fall 
-to a peak after which they decrease until the flu outbreak ends. For both the 
+to a peak and then decrease until the flu outbreak ends. For both the 
 2022 and 2023 seasons, the hospitalizations peaked during the same week as 
@@
 data for all 50 US states are included in the supplementary material 
-\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar 
-to 
-those in figure \ref{fig:us_ili} with activity increasing
-to a peak and then decreasing.
+\cite[]{wadsworth2024bas}. These plots show similarly flu outbreak trajectories 
+similar to those in figure \ref{fig:us_ili}.
 
@@
 The typical behavior of the ILI data which starts at lower values in the late 
-summer and fall but which increases to a peak, usually in December or January, 
-followed by a decline motivates the use of a nonlinear function which follows 
+summer, increases to a peak in the winter, 
+and declines in the spring motivates the use of a nonlinear function which follows 
 a similar trajectory. Compartmental models are standard mathematical models 
@@
 disease transmission model, which has been used by some to model 
-ILI data \cite[]{osthus2019dynamic, allen2017primer}. Ulloa \cite[]{ulloa2019} 
-chose to use the asymmetric Gaussian (ASG) function to model ILI data. 
+ILI data \cite[]{osthus2019dynamic, allen2017primer}.  
 The SIR is a mechanistic model used to capture transmission of a particular
-disease, and the ASG is a smooth function which may approximate the 
-same shape as the ILI trajectory. As ILI may be influenced by more 
+disease, and since ILI may be influenced by more 
 diseases than influenza, the disease transmission described by the SIR model
@@
 \cite[]{osthus2019dynamic}.
-Though the ASG function may be able to roughly capture ILI, there may still be 
+
+Ulloa \cite[]{ulloa2019} 
+chose to use the asymmetric Gaussian (ASG) function to model ILI data.
+The ASG is a smooth function which may approximate the 
+same shape as the ILI trajectory.
+The ASG function has more parameters and is thus slightly more flexible than
+the SIR model, and thus may be able to roughly capture ILI.
+However, there may still be 
 systematic behavior which it cannot capture. Thus when using either the SIR
@@
 generalization, the model may incorporate any appropriate nonlinear function, 
-though the focus here is on the SIR and ASG functions which are defined. 
+though the focus here is on the SIR and ASG functions, defined later
+in this section. 
 With the aim of forecasting hospitalizations, we also introduce a linear 
-model of hospitalizations data with ILI as a predictive covariate. To forecast 
-hospitalizations, ILI data is first forecasted and the forecast is then 
-plugged in as a covariate in the hospitalization model, thus producing 
-hospitalization forecasts. The hospitalization model is also defined in this 
-section.
-
-The forecasts produced should be probabilistic, making Bayesian modeling a 
+model of hospitalization data with ILI as a predictive covariate. To forecast 
+hospitalizations, ILI data is first forecasted and the forecast is then,
+as a covariate,
+plugged into the hospitalization model to produce
+hospitalization forecasts.
+
+Because we desire probabilistic forecasts, Bayesian modeling is a 
 natural framework for describing forecast uncertainty. 
@@
 scale parameter. The nonlinear function $f_{\theta_s}(w)$ captures the 
-trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the 
-systematic patterns which $f_{\theta_s}(w)$ does not capture. 
+trajectory of the ILI, and $\gamma_w$ is a discrepancy term shared by 
+all seasons which captures 
+systematic patterns not captured by $f_{\theta_s}(w)$. The additional 
+discrepancy term $\upsilon_{s,w}$ captures additional season 
+specific patterns.
+
 \begin{equation}
@@
 \ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically 
-over seasons. Herein we also adopt the hierarchical modeling as it proved a 
-powerful tool for improving forecasts.
+over seasons. Herein we also adopt the hierarchical modeling so as to exploit 
+information contained in the data across all seasons.
 
@@
 The SIR compartmental model is a mathematical model used for modeling disease 
-outbreaks and was introduced by Kermack and McKendrick 
+outbreaks and was introduced in 1927 by Kermack and McKendrick 
 \cite[]{kermack1927contribution}. Since then, compartmental models have 
-become standard for modeling infectious diseases 
+become standard in modeling infectious diseases 
 \cite[]{allen2008mathematical}, and many extensions have been made and 
@@
 individual and the expected rate of recovery to an immune state for a newly 
-infected person. Whether or not a disease outbreak is classified as an 
+infected individual. Whether or not a disease outbreak is classified as an 
 epidemic is determined by the initial susceptible population $S_0$, or the 
@@
 epidemic. In the case where $S_0 \leq \rho$, the trajectory for the $I$ 
-compartment would never be increasing. The increase to a peak and subsequent 
-decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is 
+compartment will be to never increase. The increase to a peak and subsequent 
+decrease in the $I$ compartment of figure \ref{fig:sir_traj} illustrate why
+it is 
 reasonable to model ILI by this compartment. Thus in modeling the ILI data, 
-we consider the data to be analogous to the $I$ proportion of the population.
+we consider the data to be rouchly 
+analogous to the $I$ proportion of the population.
 
@@
 approximate the trajectory of the flu outbreak. The ASG was previously used 
-by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used 
+by Ulloa to model and forecast ILI \cite[]{ulloa2019}. It has also been used 
 to model vegetation growth and satellite sensor data 
@@
 asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized 
-by its rise to a peak and fall from that peak which may not occur at the same 
+by its rise to a peak and fall from that peak, which rise and fall 
+may not occur at the same 
 rate, as shown in figure \ref{fig:asg_function}. The ASG function is denoted 
@@
 \end{equation}
-For modeling in this manuscript, we use a slightly reparameterized version of 
-the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$. 
+In this manuscript, we use a slightly reparameterized version of function
+(\ref{eq:asg_function}) shown in (\ref{eq:asg_function_rep}), 
+where $\eta = \nu - \lambda > 0$. 
 This constraint guarantees that the function has a peak greater than $\lambda$.
@@
 function typically underpredicts week 22 and overpredicts week 23. Perhaps 
-for other weeks, like week 30 for example, there also tends to be systematic 
+for other weeks, week 30 for example, there tends to be systematic 
 behavior not captured by the ASG function.
@@
 function. Here $\gamma_w$ is shared across all seasons, whereas $\upsilon_{s,w}$
-is specific to season $s$. In fact, to assist in identifying $\gamma_w$,
-$\upsilon_{s,w}$ is only included for the current season being forecast and none of
-the previous seasons.
+is specific to season $s$. To assist in identifying $\gamma_w$,
+$\upsilon_{s,w}$ is only included for the current season being forecast and 
+none of the previous seasons.
  As in Osthus's model, 
@@
 $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
-identifiability. An additional constraint is $\upsilon_{s, W} = 0$.
+identifiability as well as the constraint $\upsilon_{s, W} = 0$.
 
@@
 
-Figure \ref{fig:lag_scatter} helps
+Figure \ref{fig:lag_scatter} shows US national level data which helps
 illustrate why this modeling decision was made.
@@
 The bottom row shows scatterplots of the difference between hospitalizations and 
-1 week lags scaled 
-by ILI percentage at the US national level. The plots in the right column
+1 week lags
+by ILI percentage. The plots in the right column
 are of the logarithm of hospitalizations. The lines through the points are
@@
 However, when including a 
-single autoregressive lag of hospitalizations in the model, much of the time 
+single AR lag of hospitalizations in the model, much of the time 
 correlation is
 accounted for as shown by the tighter and more frequent crossing of lines
-in the bottom row. Higher than order 1 autoregressions showed little or no 
-improvement over order 1, hence we consider the ARX(1) reasonably well 
-suited for modeling ILI and hospitalzations.
+in the bottom row. Models with more than 1 AR lag showed little or no 
+improvement over 1 lag.
 
@@
 Like the ILI model (\ref{eq:ili_model}), 
-the hospitalization model was fit independently fore each location.
+the hospitalization model was fit independently for each location.
 For fitting purposes $ILI_{s,w}$ is always 
@@
 $\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all 
-states.
+locations.
 
@@
 We let $D_s$ belong to the normal family, though in the supplementary materials
-?? we show a real data analysis where we considered other distribution families
+we present a real forecast
+analysis where we considered other distribution families
 including the LST and lognormal families.
@@
 priors were assigned to the variance parameters. Univariate parameters were 
-assigned either a normal distribution prior if the support is on $\mathbb{R}$, 
+assigned either a normal distribution prior if the support was on $\mathbb{R}$, 
 a half-normal prior if the support is nonnegative, or a truncated-normal 
@@
 parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are 
-assigned to $\theta$ and $\Sigma$. Here 
+assigned to $\theta$ and $\Sigma$ where
 $T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2)$.
@@
 specific prior distributions assigned to each parameter are listed in the 
-supplementary materials ??.
+supplementary materials.
 
@@
 
-The decision to cut the model into two components was made for several reasons.
+The decision to cut the forecast modelinto two component models
+was made for several reasons.
 When fitting the fully Bayesian joint model, the time to fit a single model using
 Markov chain Monte Carlo (MCMC) sampling and
-obtain forecasts was sometimes upwards of 4 hours whereas fitting the two models
-separately usually took less than 1 hour. When assessing convergence of model
+obtain forecasts was sometimes many hours whereas fitting the two models
+separately took a fraction of the time. When assessing convergence of model
 parameters, it was rare that posterior sampling chains for the joint model
@@
 
-Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating. Future ILI 
-forecasts are obtained via the posterior predictive distribution where for
-week $w$, the predictive distribution is obtained by integrating over the 
+Model (\ref{eq:ili_model}) was fit via Bayesian posterior updating. Future ILI 
+forecasts were obtained via the posterior predictive distribution where for
+week $w$, the predictive distribution was obtained by integrating over the 
 parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
@@
 is the density function of the posterior distribution for the model parameters. 
-If the current week is $w$ then the desired forecasts are for weeks $w + i$ 
+If the current week was $w$ then the desired forecasts were for weeks $w + i$ 
 where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
-predicted or forecast ILI at $i$ weeks in the future.
+predicted or forecasted ILI at $i$ weeks in the future.
 
@@
 % \spencer{Isn't that captured in $d \pi_s$?}
-Hospitalization forecasts $\widetilde{H}_{s,w + i}$ at week $w + i$ are then
+Hospitalization forecasts $\widetilde{H}_{s,w + i}$ at week $w + i$ were then
 obtained by combining the posterior predictive distribution of model 
 (\ref{eq:final_hosp}) with (\ref{eq:ili_post}) and integrating over 
-$\widetilde{ILI}_{s,w + i}$ as in (\ref{eq:hosp_post_forc}). Here 
+$\widetilde{ILI}_{s,w + i}$ as in (\ref{eq:hosp_post_forc}). Here,
 $\boldsymbol{\alpha}$ represents all regression and variance parameters in
@@
 several diagnostic statistics for assessing the sampler.
-As mentioned, most model parameter prior distributions were intended to be 
-uninformative.
-Plots of posterior distributions for select parameters from ILI and 
-hospitalization models are shown in the supplementary material 
+Plots of prior and posterior distributions for select parameters from ILI and 
+hospitalization models are included in the supplementary material 
 \cite[]{wadsworth2024bas}.
 
-We assessed the model fit for four ILI models and for 
+Model fit was assessed for four ILI models and for 
 the hospitalization model. The ILI models included the SIR and 
@@
 % term is included or excluded. 
-To assess posterior sampling convergence, models 
-were fit to data where ILI and hospitalization data up to weeks 14, 20, and 26
-of the 2023 
-season were included. Sampling was done with four chains where from each chain 
+Sampling was done with four chains where from each chain 
 60,000 posterior draws were sampled, and the first 10,000 draws were discarded 
@@
 each parameter, and for all models fit, the largest $\hat{R}$ was below the 
-recommend threshold of 1.05, and the smallest $ESS$ was over 1,000.
+recommend threshold of 1.05, and the smallest $ESS$ was over 500.
 % Tables \ref{tab:ili_diagnostics} and 
@@
 set to 0 to reflect realistic values of hospitalizations and comply with the 
-FluSight forecasting rules. This ultimately made almost no difference in 
-results.
+FluSight forecasting rules.
 
@@
     \end{adjustwidth}
-    \caption{Ratio of average log-weighted interval score for ASGD\_NORM2 
+    \caption{Ratio of average log-weighted interval score (LWIS) for ASGD\_NORM2 
     to that of on competing FluSight model. Scores are averaged
@@
   PSI-PROF & 63 & 0.35 & 15 & 0.23 & 0.003 & 100 & 0.90 & 0.97 \\ 
+  SIRD\_NORM2 & 67 & 0.36 & 16 & 0.26 & 0.001 & 100 & 0.93 & 0.95 \\
   CEPH-Rtrend\_fluH & 69 & 0.39 & 13 & 0.28 & 0.004 & 100 & 0.81 & 0.86 \\ 

### 2025-08-13 make_images/make_final_analysis_images.R

@@ -17,22 +17,16 @@ ili_scores <- readRDS("../flu_forecast_23/nl_scores.rds") %>%
   left_join(states, by = "location_name") %>% 
   unique()
   
+team_scores <- rbind(team_scores, ili_scores %>% 
+                       filter(model == "sir_disc2_hosp_sq_ar1")) %>% 
+  mutate(model = ifelse(model == "sir_disc2_hosp_sq_ar1", 
+                        "<span style='font-size:12pt'><b><i>SIRD_NORM2</i></b></span>",
+                        model))
 
 
-rbind(team_scores, ili_scores) %>% 
-  # filter(str_detect(model, "asg")) %>%
-  # filter(location_name != "Florida") %>% 
-  group_by(model) %>% 
-  summarise(m = mean(lwis, na.rm = TRUE),
-            n = n()) %>% 
-  arrange(m)
 
 
-rbind(team_scores, ili_scores) %>% 
-  filter(model %in% c("asg_disc2_nm_hosp_sq_ar1", "FluSight-baseline")) %>% 
-  ggplot() +
-  geom_boxplot(aes(x = factor(horizon), y = lwis, colour = model)) #+
-  scale_y_log10() 
+
 
 team_scores <- rbind(team_scores, 
                      ili_scores %>% 
@@ -289,8 +283,12 @@ state_diff / date_diff
 #############Coverage#############
 ##################################
 # saveRDS(ili_coverage, "./nl_cover.rds")
-team_coverage <- readRDS("./hub_cover_no_ens.rds")
-ili_coverage <- readRDS("./nl_cover.rds")
+team_coverage <- readRDS("../flu_forecast_23/hub_cover_no_ens.rds")
+ili_coverage <- readRDS("../flu_forecast_23/nl_cover.rds")
+# team_coverage <- rbind(team_coverage, ili_coverage %>% 
+#                          filter(model == "sir_disc2_hosp_sq_ar1")) %>% 
+#   mutate(model = ifelse(model == "sir_disc2_hosp_sq_ar1", "SIRD_NORM2",
+#                         model))
 pcover <- team_coverage %>% 
   filter(!str_detect(model, "FluSight-baseline")) %>% 
   dplyr::select(reference_date, horizon, location, model, contains("cover")) %>% 
@@ -316,6 +314,18 @@ pcover <- team_coverage %>%
               summarise(pcover = mean(covered)),
[…]

### 2025-08-14 manuscript_res.tex
almost finished edits

@@
 
-The contribution of this manuscript is to introduce a two component framework 
+The contribution of this manuscript is to introduce a two component 
+Bayesian model framework 
 for modeling HHS hospitalization forecasts where hospitalization data and 
 years of ILI data are used to inform forecast models. The first modeling 
-component is a model of ILI data and the second is a model of hospitalization 
+component is a hierarchical model of ILI data. 
+The second is a model of hospitalization 
 data with ILI as a predictive covariate.
-Herein we use ILI models similar to those in Osthus et al. 
+Herein we use Bayesian ILI models similar to those in Osthus et al. 
 \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
@@
 ILI data. Herein, we provide a framework under which discrepancy modeling may 
-be used along with a general function modeling ILI data, and we show the 
-effectiveness of including discrepancy modeling during certain periods of 
-the flu season.
+be used along with a general function modeling ILI data. Modeling discrepancy
+introduces some challenges and requires thoughtful selection of model 
+constraints and prior distributions to be used effectively.
 
 In line with the newer FluSight standard of forecasting hospitalizations, we 
-model hospitalizations as a linear function of ILI. Thus forecasts produced 
-herein target flu hospitalizations and are a mapping of ILI forecasts to 
-hospitalizations. This allows for ILI data from many seasons to be exploited 
-and for ILI forecasts to assist in forecasting hospitalizations, which has 
-fewer seasons of data than ILI. Several modeling schemes and their forecasts 
-for the 2023 flu hospitalization season are compared, and it is shown that the 
-modeling decisions produce good forecast results for different states or times 
-during the flu season.
+model hospitalizations as a linear function of ILI. The Bayesian scheme 
+produces a posterior predictive distribution of hospitalizations, which is 
+shown to be a reasonably well calibrated probabilistic forecast.
+The flu hospitalization forecasts are a mapping of ILI forecasts to 
+hospitalizations, and this allows for ILI data from many seasons to be exploited 
+and for ILI forecasts to assist in forecasting hospitalizations.
+At this state of national flu forecasting, using ILI data can be especially
+useful since there are many more seasons of available ILI data than 
+hospitalization data. 
+The forecasts produced in this manuscript are compared with out of the box 
+forecast models in a simulation study, and in a real data analysis the 
+forecasts are compared with 20 other forecasts from the 2023 FluSight forecast
+competition. The proposed models outperform both the out of the box methods, and
+they also outperform all but one of the FluSight models.
 
@@
 \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
-rules are commonly used in forecast comparison and have the property that a 
+rules are commonly used in forecast comparison, and they are deisgned such 
+that a 
 forecaster is incentivized to be honest in the reporting of their forecasts 
 \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The proper scoring rules primarily used in the FluSight competition and the 
-COVID-19 Forecast Hub and which we will use herein are the negatively oriented
+The proper scoring rule primarily used in the FluSight competition and the 
+COVID-19 Forecast Hub, and which is used in this simulation study,
+is the negatively oriented
 (smaller is better)
@@
 The WIS is used for scoring quantile or interval 
-forecasts 
-\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} 
-and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
+forecasts, or forecasts made up of predicitive intervals of several nominal 
+levels
+\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating}. 
+The WIS defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
 included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
@@
 of the 500 simulation replicates. The ASGD forecasts on average clearly 
-outperformed the other models considered for almost each week. The ASGD does
-especially well at week 20 which contains forecasts of the holiday week 22, 
-showing the particular importance of modeling discrepancy at and around that 
+outperformed the other models considered in almost every week. The ASGD does
+especially well at week 20 which contains forecasts of the holiday week 22. 
+This demonstrates the
+particular importance of modeling discrepancy at and around that 
 week. The SIRD outperforms the SIR model, again showing the importance of 
@@
 scores by model and separated by horizon forecasts of 1 to 4 weeks ahead. Here
-also the ASGD appears to perform the best with the other variations of the 
+also the ASGD performs the best with the other variations of the 
 nonlinear models not always outperforming the BASE or ARIMA models. 
@@
 shows the empirical coverage of each model for 11 predictive intervals from 
-10\% to 98\% nominal levels. Each model is was undercalibrated, but the ASGD 
+10\% to 98\% nominal levels. Each model is undercalibrated, but the ASGD 
 model is the closest to being well calibrated, especially at the hightest 
@@
  log-weighted intveral score (LWIS). The LWIS is defined the same as the WIS
- but where the log of the observation as well as the forecast quatiles are taken
+ but where the log of the observation as well as the 
+ log forecast quantiles are taken
  prior to calculating the score. We primarily use the LWIS in this section,
- though forecasting results in terms of the WIS may also be found in the 
- supplementary materials ??.
+ though we including results for the WIS in the 
+ supplementary materials.
 
@@
 remainder of this section, focus will be made on forecasts made by coupling the
-ASGD model with normal hospitalization model where ILI is a quadratic predictor.
-We denote this model as ASGD\_NORM2. 
-This decision was based on the ASGD's performance in the simulation study in 
-section \ref{sec:simulation2} as well as an analysis of several other forecast
+ASGD and SIRD models with the 
+normal hospitalization model where ILI is a quadratic predictor.
+These forecast schemes are denoted as ASGD\_NORM2 and SIRD\_NOMR2. 
+In the supplementary materials, we include analysis of 2023 hospitalization
+forecasts for several other forecast
 models including ASG, SIRD, SIR ILI models coupled with other variations of the
-hospitalization model where the ASGD\_NORM2 model showed the best performance.
-Details
-of that analysis are found in the supplementary materials ??.
+hospitalization model where the ASGD\_NORM2 and SIRD\_NORM2 models
+showed the best and second best results respectively.
 
 The FluSight 
-models to which the ASGD\_NORM2 model are compared are the 
+models to which are used for comparison are
 20 non-ensemble models
@@
 Figure \ref{fig:lwis_cover_sum} shows forecast results for 2023 flu 
-hospitalization forecasts for all 21 models being compared but with results
-for the ASGD\_NORM2 and FluSight baseline models highlighted. 
+hospitalization forecasts for all 22 models being compared but with results
+for the ASGD\_NORM2, SIRD\_NORM2, and FluSight baseline models highlighted. 
 For each week of the 2023 season, the  
@@
 for each model is shown in a). Boxplots of LWIS for 1, 2, 3, and 4-week ahead
-horizons for the ASGD\_NORM2 and baseline models are shown in b). And 
+horizons for the ASGD\_NORM2, SIRD\_NORM2, and baseline models are shown in b). 
+And 
 overall empirical coverage of the 11 predictive intervals for all models is 
-shown in c) with ASGD\_NORM2 and the baseline model again highlighted. 
+shown in c) with ASGD\_NORM2, SIRD\_NORM2
+and the baseline model again highlighted. 
 These plots show that over the season the ASGD\_NORM2 model tended to 
@@
 up to about 80\%, and has better coverage than all other models. Similar results
-for the WIS are shown in the supplementary materials ??.
+for the WIS are shown in the supplementary materials.
 
@@
 miss a forecast submission for some week, location, or horizon. This makes a 
-simple average of scores an imperfect measure of comparison, as forecasts
+simple average of scores an imperfect measure for comparison, as forecasts
 of certain targets included for one model may be excluded by another model.
 Thus in figure \ref{fig:state_and_date_lwis} 
-we compare the ASGD\_NORM2 model with all 
-other models directly by comparing LWIS scores averaged over only shared
-forecasts. Each tile in the plots represents the ratio of average ASGD\_NORM2
+the ASGD\_NORM2 model is compared with all 
+other models individually using LWIS scores averaged over only the 
+forecasts shared by the two models.
+Each tile represents the ratio of average ASGD\_NORM2
 LWIS to that of one FluSight model by location or by week.
 We call this the relative LWIS or RLWIS. A value smaller than
-1 indicates the ASGD\_NORM2 model outperforms the other model with 
-increasing outperformance as the value goes to 0. A value 
+1 indicates the ASGD\_NORM2 model outperforms the competing model. A value 
 higher than 1 indicateds the ASGD\_NORM2 is outperformed. A grey tile indicates
 the competing model had no forecasts matching the specified target.
-Similar plots showing WIS scores are found in the supplementary materials ??.
-
-Our model outperforms most other models with only the 
-UGA\_flucast-INFLAenza model appearing to outperform ours. 
+Similar plots showing WIS scores are found in the supplementary materials. 
+The ASGD\_NORM2 outperforms most other competing models more often than not, but
+there are some weeks or locations where forecasting was poorer. For example, 
+the plot shows the ASGD\_NORM2 performed poorly in Montana relative to the other
+models.
+
 Table \ref{tab:fin_analysis_stats} gives summary results for mean and median
 WIS and LWIS scores, predictive coverage, and relative RWIS and RLWIS scores
-for each model. These summary values show the ASGD\_NORM2 as performing very
-well compared to all other models, and outperforming all models except
-UGA\_flucast-INFLAenza by RWIS and RLWIS. 
-The takeaway of this section, like those of section \ref{sec:simulation2}, show
-the combination of well performing model of ILI, the 
-hierarchical Bayesian ASGD, with a 
-straightforward hospitalization model lead to good forecasts which 
-outperformed almost all other non-ensemble forecasts during of the 2023
-FluSight competition.
+for each model. Along with the plots in this section, it 
+show that over the course of the 
+2023 season, 
+the ASGD\_NORM2 outperformed every other model except the 
+UGA\_flucast-INFLAenza. The SIRD\_NORM2 is also one of the top performing 
+forecast models.
 
@@
  In this manuscript we introduce a statistical modeling framework which allows 
- for the incorporation of several ILI forecast modeling methods. Specifically, 
+ for the incorporation of several nonlinear
+ ILI forecast modeling methods. Specifically, 
  we built upon Osthus et al. \cite[]{osthus2019dynamic} and introduced a 
  framework for modeling ILI which includes the use of an arbitrary function 
- for modeling the main trajectory of ILI along with modeling the discrepancy. 
+ for modeling the main trajectory of ILI along with modeling the discrepancy and
+ specifically incorporate the smooth ASG function and compare forecasts using 
+ the ASG with those using SIR. 
  We model flu hospitalizations by incorporating the ILI forecast model into a 
  model forecasting hospitalizations where hospitalization predictions are a 
- linear or quadratic function of ILI.
-
-The simulation study in section \ref{sec:simulation2} suggests the ASG 
-function in ILI modeling may slightly outperform the SIR model according to 
-the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season 
-forecasts, the SIR model was overall superior. The results from both the 
-simulation study and the real data analysis suggest that the addition of a 
-discrepancy component in ILI modeling may improve forecasts especially near 
-the holiday week between Christmas and New Years day. It should not be assumed 
-that these conclusions may be generalized for all locations in the US, weeks 
-of a season, or for future flu seasons. 
+ linear or quadratic function of ILI. Utlizing the ILI data which has been
+ available for many more years than flu hospitalization, greatly improves
+ one's ability to forecast hospitalizations, and the borrowing information 
+ from all previous seasons through hierarhcical modeling is able to further
+ exploit the ILI data. Including model iscrepancy was an important piece of 
+ successful forecasting, but it required making thoughtful model constraints
+ and selection of prior distributions.
+
+The simulation study in section \ref{sec:simulation2} suggests the ASGD 
+function in ILI modeling outperforms the SIRD model and out of the box 
+forecast methods according to 
+the WIS scoring rule, and in the analysis of 2023 FluSight forecasts the 
+ASGD\_NORM2 modeling scheme outperformed all but one of the competing
+forecasts when looking at the whole season. Though the ASGD\_NORM2 did very 
+well across all seasons, for some weeks and locations its perfomance was not
+as good. It should also not be assumed that ASGD\_NORM2 forecasts would perform
+as well in every season given that each flu season will come with its own
+challenges which the modeling scheme may or may not address.
 % Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.
 
-Forecasting the seasonal influenza outbreak remains a challenging task for 
-forecasters. The general modeling framework in this manuscript is successful 
-under diverse modeling conditions for all locations in the US and may 
-contribute to future forecasting efforts. All forecast models were presented 
-as separate from one another, but in the various locations and times of the 
-flu season, different models perform better than others. To build on the work 
-done here, a natural step forward would be to combine all or a few selected 
-forecasts into an ensemble forecast. Such an ensemble may work to cancel out 
-certain model biases or highlight model strengths, leading to more robust 
-forecasts.
+The models in this manuscript were such that the data in each
+location were modeled as
+independent from all other locations. It is reasonable to assume that a
+disease outbreak in one state will share characterstics with the outbreak in 
+neighboring states, and indeed the behavior of ILI and hospitalizations is more
+closely related for neighboring states as shown in data plots in the
+supplementary materials. The one 2023 FluSight forecast model which outperformed
+the ASGD\_NORM2, the UGA\_flucast-INFLAenza, was listed as being a spatial
+model. And in past seasons, some of the top performing forecast models also 
+explicitly modeled spatial relationships \cite[]{osthus2021multiscale}.
+Likewise, the forecasts to which ours was compared were all non-ensemble 
+forecasts which were all outperformed by the ensemble forecasts in FluSight.
+Futher extensions of the models herein could include explicit modeling of 
+spatial relationships between states and inclusion in an ensemble of other
+forecast models.

### 2025-08-14 abstract.tex
update abstract

@@
 of the disease spread. The United States Centers for Disease Control and 
-Prevention (CDC) hosts annually a national flu forecasting competition which has 
+Prevention hosts annually a national flu forecasting competition which has 
 led to the development of a variety of flu forecast modeling methods. For the 
@@
 seasons.
-In this manuscript, we introduce a two component modeling framework for 
+In this manuscript, we introduce a two component 
+Bayesian modeling framework for 
 forecasting weekly hospitalizations utilizing both hospitalization data and ILI 
 data. The first component is for modeling ILI data using a dynamic nonlinear 
-Bayesian hierarchical model. The second component is for modeling 
+Bayesian hierarchical model. This component includes modeling discrepancy 
+between a functional model and the observed which requires thoughtful decisions
+for modeling constraints and selecting prior parameter distributions.
+The second component is for modeling 
 hospitalizations as a function of ILI. For hospitalization forecasts, ILI is 
-first forecasted and then hospitalizations are forecast with ILI forecasts used 
-as a linear or quadratic predictor. In a simulation study, two ILI forecast 
+first forecasted and then hospitalizations forecasts are produced using
+ILI forecasts 
+as a linear or quadratic predictor. The combination of the models is done in a
+manner similar to using the Bayesian cut.
+In a simulation study, two ILI forecast 
 models, including one similar to the winning model for two seasons of the CDC 
@@
 % \cite[]{ulloa2019}
-are compared. Also assessed is the usefulness of including a systematic model 
-discrepancy term in the ILI model. Forecasts of state and national 
-hospitalizations for the 2023-24 flu season are made, and different modeling 
-decisions are compared. We found that including a discrepancy component in the 
-ILI model tends to improve forecasts during certain weeks of the year. We also 
-found that other modeling decisions such as the exact nonlinear function to be 
-used in the ILI model or the error distribution for hospitalization models may 
-or may not be better than other decisions, depending on the season, location, 
-or week of the forecast.
+are compared with other standard forecasts models.
+The usefulness of including a systematic model 
+discrepancy term in the ILI model is specifically addressed and proves 
+to greatly improve forecasts. In a real data analysis, 
+forecasts of state and national 
+hospitalizations for the 2023-24 flu season are made and compared with forecasts
+from 20 other competing models. The forecasts from the proposed methodology 
+outperformed all but one of 20 competing models. 
 \end{abstract}

### 2025-08-14 response.tex
editing almost complete

@@
 produce forecasts. A natural extension of the model would be to allow for states
-to borrow information from neighboring states." ??Add something to discussion}
+to borrow information from neighboring states."
+
+We also added the following to the discussion:
+
+``The models in this manuscript were such that the data in each
+location were modeled as
+independent from all other locations. It is reasonable to assume that a
+disease outbreak in one state will share characterstics with the outbreak in 
+neighboring states, and indeed the behavior of ILI and hospitalizations is more
+closely related for neighboring states as shown in data plots in the
+supplementary materials. The one 2023 FluSight forecast model which outperformed
+the ASGD\_NORM2, the UGA\_flucast-INFLAenza, was listed as being a spatial
+model. And in past seasons, some of the top performing forecast models also 
+explicitly modeled spatial relationships."}
 

### 2025-08-14 manuscript_res.tex
editing almost complete

@@
 models including ASG, SIRD, SIR ILI models coupled with other variations of the
-hospitalization model where the ASGD\_NORM2 and SIRD\_NORM2 models
-showed the best and second best results respectively.
+hospitalization model where the ASGD\_NORM2 showed the best results
+and the SIRD\_NORM2 performed nearly as well.
 
@@
     (LWIS) for each week during the season for the 21 models being compared. 
-    Each grey line represents one model, but the ASGD\_NORM2 (blue) and 
+    Each grey line represents one model, but the ASGD\_NORM2 (blue), 
+    SIRD\_NORM2 (green), and 
     FluSight-baseline (red) are colored to stand out.
@@
     \caption{Ratio of average log-weighted interval score (LWIS) for ASGD\_NORM2 
-    to that of on competing FluSight model. Scores are averaged
+    to that of each competing FluSight model. Scores are averaged
     over all targets within indicated location (top) or week (bottom) and 

### 2025-08-14 supplemental_res.tex
editing almost complete

@@
     (WIS) for each week during the season for the 21 models being compared. 
-    Each grey line represents one model, but the ASGD\_NORM2 (blue) and 
+    Each grey line represents one model, but the ASGD\_NORM2 (blue),
+    SIRD\_NORM2 (green), and 
     FluSight-baseline (red) are colored to stand out.
@@
     \caption{Ratio of average weighted interval score (WIS) for ASGD\_NORM2 
-    to that of on competing FluSight model. Scores are averaged
+    to that of each competing FluSight model. Scores are averaged
     over all targets within indicated location (top) or week (bottom) and 

### 2025-08-28 response.tex
last revisions

@@
 ``This two-component modeling is remeniscent of the "Bayesian cut" as 
-described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
-into two component models or "modules" may be justfied for several reasons
+described in [44] and [41]. Cutting a model
+into two component models or ``modules`` may be justfied for several reasons
 including to reduce time to fit the model, avoiding mixing issues of posterior
@@
 model misspecification in one model to influence the other 
-\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, plummer2015cuts}.
+[41, 27, 26, 44].
 
-The decision to cut the model into two components was made for several reasons.
+The decision to cut the forecast model into two component models
+was made for several reasons.
 When fitting the fully Bayesian joint model, the time to fit a single model using
 Markov chain Monte Carlo (MCMC) sampling and
-obtain forecasts was sometimes upwards of 4 hours whereas fitting the two models
-separately usually took less than 1 hour. When assessing convergence of model
+obtain forecasts was sometimes many hours whereas fitting the two models
+separately took a fraction of the time. When assessing convergence of model
 parameters, it was rare that posterior sampling chains for the joint model
@@
 the forecasts from the component models generally outperformed those of the 
-joing model in terms of minimizing proper scoring rules. 
-The process for obtaining forecasts is described in the following paragraphs 
-and computation is described in section \ref{sec:implementation_posterior}."}
+joint model in terms of minimizing proper scoring rules.``}
 
@@
 identifiability was encouraged by setting
-the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})
+the parameter $\lambda_s$ from (4)
 to be equal to its MLE. In this updated manuscript, 
@@
 all seasons and set the variance parameter so the prior was tight around the
-mean. This led to improved mixing of posterior draws and better forecasts."}
+mean. This led to improved mixing of posterior draws and better forecasts.``}
 
@@
 we addressed this in the manuscript by adding the following discussion about
-figure 6. See also
+figure 6.
 
@@
 However, when including a 
-single AR lag of hospitalizations in the model, much of the time 
+single autoregressive lag of hospitalizations in the model, much of the time 
 correlation is
 accounted for as shown by the tighter and more frequent crossing of lines
-in the bottom row. Models with more than 1 AR lag showed little or no 
-improvement over 1 lag."}
+in the bottom row. Models with more than 1 autoregressive
+lag showed little or no 
+improvement over 1 lag.``}
 
@@
 
-\spencer{We revised the first paragraph in section 2.2 to read:
+\spencer{It refers to new hospitalizations. 
+We revised the first paragraph in section 2.2 to read:
 
@@
 were updated every Wednesday and Friday according to NHSN guidelines 
-\cite[]{healthdata2024covidts}. Weekly influenza hospitalization data is
+[22]. Weekly influenza hospitalization data is
 defined as the number of newly hospitalized patients with a confirmed diagnosis 
 of influenza. From hereon, influenza hospitalizations
-will be referrred to as hospitalizations."}
+will be referrred to as hospitalizations.``}
 
@@
 been in the past. We assessed ILI lags for both one and two weeks previous to the current
-week and found that ILI had less predictive power. }
+week and found that ILI had less predictive power. We note this in the
+revised manuscript by saying:
+
+`` We found that models
+ with more than 1 autoregressive lag showed 
+ little or no forecast improvement over 1 lag.``}
 
@@
 For the forecast competition analysis in 
-section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
+section 5, all negative values of $\{H^*_{s,w + i}\}^K$ were 
 set to 0 to reflect realistic values of hospitalizations and comply with the 
-FluSight forecasting rules."}
+FluSight forecasting rules.``}
 
@@
 and the real data analysis. We decided on the WIS because that was the measure
-used in the FluSight competition and the forecasts which we }
+used in the FluSight competition.}
 
@@
 
-\spencer{Add statement to discussion on this ??}
+\spencer{The following two statements were added. The first in the main body
+and the second in the conslusion.
+
+``Model (1) was fit 
+independently for each location. This decision was made largely due to the 
+limited time given by FluSight to produce forecasts. That is,
+during the competition, data were typically released on Wednesday afternoons and
+forecasts were due by the evening of the same day giving only a few hours to 
+produce forecasts. A natural extension of the model would be to allow for states
+to borrow information from neighboring states.``
+
+``The models in this manuscript were such that the data in each
+location were modeled as
+independent from all other locations. It is reasonable to assume that a
+disease outbreak in one state will share characterstics with the outbreak in 
+neighboring states, and indeed the behavior of ILI and hospitalizations is more
+closely related for neighboring states as shown in data plots in the
+supplementary materials. Notably, the one 2023 FluSight forecast model which 
+outperformed
+the ASGD\_NORM2, the UGA\_flucast-INFLAenza, was listed as being a spatial
+model. And in past seasons, some of the top performing forecast models also 
+explicitly modeled spatial relationships [43].``}
 
@@
 Thus, care must be taken in setting parameter constraints as well as in the 
-selection of prior distributions \cite[]{osthus2019dynamic,
-brynjarsdottir2014learning}. 
+selection of prior distributions [42,8]. 
 For instance, when setting $f_{\theta_s}(w)$ to
 be the ASG function
-in model (\ref{eq:ili_model}), we found the parameter $\lambda$ from 
+in model (1), we found the parameter $\lambda$ from 
 in (\ref{eq:asg_function_rep}) difficult to consistently 
 estimate precisely because of lack of identifiability. This was resolved by 
-assigning to $\lambda$ a more informative prior distribution."
-?? incomplete}
+assigning to $\lambda$ a more informative prior distribution.``}
 
@@
 
-``Model (\ref{eq:ili_model}) was fit 
+``Model (1) was fit 
 independently for each location. This decision was made largely due to the 
@@
 produce forecasts. A natural extension of the model would be to allow for states
-to borrow information from neighboring states."
+to borrow information from neighboring states.``
 
@@
 closely related for neighboring states as shown in data plots in the
-supplementary materials. The one 2023 FluSight forecast model which outperformed
+supplementary materials. Notably, the one 2023 FluSight forecast model which 
+outperformed
 the ASGD\_NORM2, the UGA\_flucast-INFLAenza, was listed as being a spatial
 model. And in past seasons, some of the top performing forecast models also 
-explicitly modeled spatial relationships."}
+explicitly modeled spatial relationships [43].``}
 
@@
 ``Additional prior constraints were made to improve parameter identifiability. 
-Just as was done in Osthus et al. \cite[]{osthus2019dynamic},
+Just as was done in Osthus et al. [42],
 we set the initial value of the 
@@
 rather than estimating all parameters through posterior updating,
-we set the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})
+we set the parameter $\lambda_s$ from (4)
 to be equal to its MLE. In this updated manuscript, 

### 2025-08-28 abstract.tex
last revisions

@@
 forecasting weekly hospitalizations utilizing both hospitalization data and ILI 
-data. The first component is for modeling ILI data using a dynamic nonlinear 
+data. The first component is for modeling ILI data using a nonlinear 
 Bayesian hierarchical model. This component includes modeling discrepancy 
-between a functional model and the observed which requires thoughtful decisions
+between a nonlinear functional model and the observed ILI which requires 
+thoughtful decisions
 for modeling constraints and selecting prior parameter distributions.

### 2025-08-28 manuscript_res.tex
last revisions

@@
 \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. 
-Originally, over a dozen teams of researchers from academic and industry 
+The competition lasted from the fall of 2013 through spring of 2014, and
+originally over a dozen teams of researchers from academic and industry 
 backgrounds participated in FluSight by contributing their own forecast models. 
-Besides the 2020 season --or the flu season spanning the fall of 2020 and the 
-winter of 2021-- FluSight has been operated annually and researchers outside 
-the CDC have been invited to participate. Initially the target data for 
+Besides the 2020-21 season, FluSight has been operated annually and 
+researchers outside 
+the CDC have been invited to participate. From hereon, we denote a 
+flu season by the year in which the outbreak began, i.e. the 2020 season 
+season outbreak began in 2020 and lasted until spring 2021.
+Initially the target data for 
 forecasts was influenza-like illness (ILI) data. ILI is the proportion of 
@@
 \cite[see references therein for more examples]{mcandrew2021adaptively, 
-osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's 
+osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their 
 introduction, Osthus et al.  \cite[]{osthus2019dynamic} categorized the most 
@@
 models which are constructed by assimilating mechanistic models into a 
-probabilistic framework. An additional forecast model used in FluSight 
+probabilistic framework. An additional forecast modeling class used in FluSight 
 involves the combination of several forecasts into a single ensemble forecast, 
@@
 ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by 
-the flu forecast competition. Additionally the target of the flu forecasts 
+FluSight. Additionally the target of the flu forecasts 
 changed from being ILI data to being HHS hospitalization data, which reports 
 the number of hospitalizations due to a laboratory confirmed flu infection  
-\cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of 
+\cite[]{mathis2024evaluation,healthdata2024covidts}. This in part was 
+as a result of 
 having COVID-19 cases in the population making ILI data, already only a proxy 
@@
 The contribution of this manuscript is to introduce a two component 
-Bayesian model framework 
+Bayesian modeling framework 
 for modeling HHS hospitalization forecasts where hospitalization data and 
@@
 be used along with a general function modeling ILI data. Modeling discrepancy
-introduces some challenges and requires thoughtful selection of model 
+introduced challenges which required thoughtful selection of model 
 constraints and prior distributions to be used effectively.
@@
 In line with the newer FluSight standard of forecasting hospitalizations, we 
-model hospitalizations as a linear function of ILI. The Bayesian scheme 
-produces a posterior predictive distribution of hospitalizations, which is 
-shown to be a reasonably well calibrated probabilistic forecast.
-The flu hospitalization forecasts are a mapping of ILI forecasts to 
-hospitalizations, and this allows for ILI data from many seasons to be exploited 
-and for ILI forecasts to assist in forecasting hospitalizations.
-At this state of national flu forecasting, using ILI data can be especially
-useful since there are many more seasons of available ILI data than 
-hospitalization data. 
+model hospitalizations as a linear function of ILI. 
+The flu hospitalization forecasts are a linear mapping of ILI forecasts
+allowing for ILI data from many seasons to be exploited
+and assist in forecasting hospitalizations.
+To produce hospitalization forecasts, a scheme similar to 
+the Bayesian cut is used where forecasts from the posterior predictive
+distribution of an ILI model are plugged into the hospitalization model
+to produce a posterior predictive hospitalization forecast distribution.
+We show that this scheme produces accurate and reasonably well calibrated
+probabilistic forecasts.
 The forecasts produced in this manuscript are compared with out of the box 
@@
 competition. The proposed models outperform both the out of the box methods, and
-they also outperform all but one of the FluSight models.
+they outperform all but one of the FluSight models.
 
@@
  models and their use in forecasting hospitalizations are compared. Commonly 
- used proper scoring rules \cite[]{gneiting2007strictly}, which are also 
- introduced and defined in section \ref{sec:simulation2}, are used for 
+ used proper scoring rules \cite[]{gneiting2007strictly} are used for 
  comparing the forecasts. Forecasting of the 2023 flu outbreak along with 
- assessment and comparison under several selected models is performed in 
+ assessment and comparison of 20 FluSight models is performed in 
  section \ref{sec:analysis}. 
@@
  \section{Flu outbreak data} \label{sec:data}
-In this section we introduce and define ILI and hospitalization data and 
-visually evaluate the data. ILI and hospitalization data have been the object 
+In this section we introduce, define, and visually evaluate
+ILI and hospitalization 
+data. ILI and hospitalization data have been the object 
 of forecasting for FluSight with ILI being the target for the first seven
@@
 2022 season. Both of 
-these data were collected at the state, territorial, and national level and 
+these data were collected at the state, territorial, and national levels and 
 were reported at least weekly. Overall the data is reported for 53 locations 
 including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), 
-and at the US national level. We will refer to forecast targets throughout this 
+and the US national level. We will refer to forecast targets throughout this 
 manuscript. A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a 
@@
 \subsection{Influenza-like illness data}
-The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects 
+The US Outpatient Influenza-like Illness Surveillance Network collects 
 information on respiratory illness from outpatient visits to health care 
@@
 ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
-and/or a sore throat." Prior to the 2021 season, the definition included 
-``without a known cause other than influenza" \cite[]{cdcfluview2023}. Because 
+and/or a sore throat.`` Prior to the 2021 season, the definition included 
+``without a known cause other than influenza`` \cite[]{cdcfluview2023}. Because 
 other illnesses such as COVID-19, RSV, and the common cold may induce similar 
@@
 \cite[]{rudis2021cdcfluview}. Weekly ILI data from the national, HHS region, 
-and census levels are available from the 1997 flu season until the current 
+and census levels are available from the 1997 flu season up to the current 
 season. At the state level, data is available from the 2010 flu season to the 
@@
 \cite[]{ewing2017contact, garza2013effect}.
-The only seasons when there was not a peak at week 22 were season 2022 when 
-the season peak occurred particularly early and season 2020 which was greatly 
+The only seasons when there was not a peak at week 22 were 2022, when 
+the season peak occurred particularly early, and 2020 which was greatly 
 influenced by the COVID-19 pandemic.
@@
 hospitalizations for the 2022 and 2023 flu seasons. These plots show a 
-trend similar to the ILI plots in the top of the figure in that during the early 
+trend similar to the ILI plots in that during the early 
 weeks of the season hospitalizations are low, but they increase in the fall 
@@
 data for all 50 US states are included in the supplementary material 
-\cite[]{wadsworth2024bas}. These plots show similarly flu outbreak trajectories 
+\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories 
 similar to those in figure \ref{fig:us_ili}.
@@
 
-The typical behavior of the ILI data which starts at lower values in the late 
-summer, increases to a peak in the winter, 
-and declines in the spring motivates the use of a nonlinear function which follows 
+The typical behavior of the ILI data to begin low, rise to a peak, and then 
+fall motivates the use of a nonlinear function which follows 
 a similar trajectory. Compartmental models are standard mathematical models 
-used for modelling disease outbreaks. One important compartmental model is the 
+used for modelling disease outbreaks which may capture ILI behavior. 
+One important compartmental model is the 
 susceptible-infectious-recovered (SIR) 
@@
 similar to the model in Osthus et al. \cite[]{osthus2019dynamic}. With some 
-generalization, the model may incorporate any appropriate nonlinear function, 
+generalization, the model may incorporate any appropriate linear or
+nonlinear function, 
 though the focus here is on the SIR and ASG functions, defined later
@@
 from the posterior predictive distribution by combining ILI and hospitalization
-models in a way similar to using the 
+models using a scheme similar to the 
 Bayesian cut \cite[]{plummer2015cuts}.
@@
 season. The ILI is a proportion, so the Beta random variable is a natural 
-selection for modeling. Under the parameterization in of the Beta distribution 
+selection for modeling. Under the parameterization of the Beta distribution 
 used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the 
@@
 compartment of the SIR model from (\ref{eq:sir_diff}) in section 
-\ref{sec:sir_func}. In Ulloa \cite[]{ulloa2019} 
+\ref{sec:sir_func}. In Ulloa \cite[]{ulloa2019}, 
 $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section 
-\ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically 
-over seasons. Herein we also adopt the hierarchical modeling so as to exploit 
+\ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019}, modeling is done hierarchically 
+over seasons. Herein we also adopt the hierarchical modeling so as to borrow 
 information contained in the data across all seasons.
@@
 it is 
-reasonable to model ILI by this compartment. Thus in modeling the ILI data, 
-we consider the data to be rouchly 
-analogous to the $I$ proportion of the population.
+reasonable to use the SIR model to partially capture ILI, although it is
+unlikely to completely capture the true data generating mechanism. Thus the 
+use of modeling discrepancy.
 
@@
 systematic behavior in the ILI that these or other possible functions may not 
-capture. As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a 
-regular peak at week 22 of the flu season. 
+capture. 
 Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to 
-illustrate the systematic discrepancy from a fitted function. Figure 
+illustrate the systematic discrepancy from a fitted function to ILI  data.
+Figure 
 \ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 
 to 2022 excluding 2020 with a best fit ASG function plotted over the ILI. The 
-fits for each season were made by obtaining the maximum likelihood estimate 
+fits for each season were made by computing the maximum likelihood estimate 
 (MLE) of a model given the ILI data where we assume the ASG function is the 
@@
 the functions from figure \ref{fig:asg_fits} for each season, and the black 
-line is the average by week over all seasons. The lines show that the ASG 
+line is the average discrepancy over all seasons. The lines show that the ASG 
 function typically underpredicts week 22 and overpredicts week 23. Perhaps 
-for other weeks, week 30 for example, there tends to be systematic 
+for other weeks, week 30 for example, there tends to be additional systematic 
 behavior not captured by the ASG function.
@@
 in model (\ref{eq:ili_model}), we found the parameter $\lambda$ from 
-in (\ref{eq:asg_function_rep}) difficult to consistently 
+(\ref{eq:asg_function_rep}) difficult to 
 estimate precisely because of lack of identifiability. This was resolved by 
@@
 function. Here $\gamma_w$ is shared across all seasons, whereas $\upsilon_{s,w}$
-is specific to season $s$. To assist in identifying $\gamma_w$,
+is specific to season $s$. To assist in estimating $\gamma_w$,
 $\upsilon_{s,w}$ is only included for the current season being forecast and 
@@
 However, when including a 
-single AR lag of hospitalizations in the model, much of the time 
+single autoregressive lag of hospitalizations in the model, much of the time 
 correlation is
 accounted for as shown by the tighter and more frequent crossing of lines
-in the bottom row. Models with more than 1 AR lag showed little or no 
+in the bottom row. We found that models with more than 1 autoregressive
+lag showed little or no forecast
 improvement over 1 lag.
@@
     scaled by $\phi$. Points are colored by season. For hospitalizations 
-    $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$. }
+    $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$. Lines through
+    the points follow the weekly progression of influenza.}
     \label{fig:lag_scatter}
@@
 
-The priors selected for the ILI data model under both the SIR and ASG models 
+The prior distributions
+selected for the ILI data model under both the SIR and ASG models 
 largely follow the prior distribution 
@@
 assigned to $\theta$ and $\Sigma$ where
-$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2)$.
+$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$.
 
@@
 
-This two-component modeling is remeniscent of the "Bayesian cut" as 
+This two-component modeling is remeniscent of the Bayesian cut 
 described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
-into two component models or "modules" may be justfied for several reasons
+into two component models or ``modules`` may be justfied for several reasons
 including to reduce time to fit the model, avoiding mixing issues of posterior
@@
 
-The decision to cut the forecast modelinto two component models
+The decision to cut the forecast model into two component models
 was made for several reasons.
@@
 the forecasts from the component models generally outperformed those of the 
-joing model in terms of minimizing proper scoring rules. 
-The process for obtaining forecasts is described in the following paragraphs 
-and computation is described in section \ref{sec:implementation_posterior}.
+joint model in terms of minimizing proper scoring rules.
 
@@
 distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following 
-steps are repeated $K$ times where $K$ is an integer for the number of desired 
+steps were repeated $K$ times where $K$ is an integer for the number of desired 
 samples. We set $K = 50,000$.
@@
 existing ILI data as a predictive covariate. Each season was in turn 
-``left-out" and treated as if it was the most recent season for which forecasts
+``left-out`` and treated as if it was the most recent season for which forecasts
 are desired. Fitting and forecasting was then done for weeks 14, 20, 26, 32, 
@@
 \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
-rules are commonly used in forecast comparison, and they are deisgned such 
+rules are commonly used in forecast comparison, and they are designed such 
 that a 
@@
 Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
-and forecasts were made at the state and national level. The first week of 
+and forecasts were made at the state and national levels. The first week of 
 forecasting took place during the week of October 7, 2023, and the final week 
@@
 to do the best job at capturing the trajectory of hospitalizations as well
-as usually making the sharpest predictions, particularly later in the season.
+as most frequently
+making the sharpest predictions, particularly later in the season.
 
@@
 forecasts for several other forecast
-models including ASG, SIRD, SIR ILI models coupled with other variations of the
+models including ASG and SIR ILI models coupled with other variations of the
 hospitalization model where the ASGD\_NORM2 showed the best results
-and the SIRD\_NORM2 performed nearly as well.
+and the SIRD\_NORM2 performed nearly as well as the ASGD\_NORM2.
 
 The FluSight 
-models to which are used for comparison are
+models which are used for comparison are the
 20 non-ensemble models
@@
 differences between 
-previous hospitalizations. The baseline forecasts was
-similar to the baseline forecast used in
+previous hospitalizations. The baseline forecast model was
+similar to the baseline forecast model used in
 previous flu forecasting seasons and in the COVID-19 hub 
@@
 
-Figure \ref{fig:lwis_cover_sum} shows forecast results for 2023 flu 
+Figure \ref{fig:lwis_cover_sum} shows forecast results for the 2023 flu 
 hospitalization forecasts for all 22 models being compared but with results
@@
 shown in c) with ASGD\_NORM2, SIRD\_NORM2
-and the baseline model again highlighted. 
+and the baseline model highlighted. 
 These plots show that over the season the ASGD\_NORM2 model tended to 
@@
     % \centering
-    \begin{adjustwidth}{-1.2cm}{-1.5cm}
+    \begin{adjustwidth}{0cm}{0cm}
     \includegraphics[scale = .54]{Images/lwis_cover_sum.png}
-    \caption{a) Averaged over location and horizon log-weighted interval score
-    (LWIS) for each week during the season for the 21 models being compared. 
+    \caption{a) Log-weighted interval score (LWIS) averaged 
+    over location and horizon for each week during the season for the 21 models being compared. 
     Each grey line represents one model, but the ASGD\_NORM2 (blue), 
@@
     line $x = y$ shows where optimal coverage would fall.
-    Each grey line represents one model, but the ASGD\_NORM2 (blue) and 
+    Each grey line represents one model, but the ASGD\_NORM2 (blue),
+    SIRD\_NORM2 and 
     FluSight-baseline (red) are colored to stand out.}
@@
 LWIS to that of one FluSight model by location or by week.
-We call this the relative LWIS or RLWIS. A value smaller than
+We call this the relative LWIS (RLWIS). A value smaller than
 1 indicates the ASGD\_NORM2 model outperforms the competing model. A value 
@@
 there are some weeks or locations where forecasting was poorer. For example, 
-the plot shows the ASGD\_NORM2 performed poorly in Montana relative to the other
+the plot shows the ASGD\_NORM2 performed poorly in Montana relative to most other
 models.
@@
 for each model. Along with the plots in this section, it 
-show that over the course of the 
+shows that over the course of the 
 2023 season, 
@@
 models and the ASGD\_NORM2 model. Summaries include mean weighted interval
-and log-weighted interval scoreS (MWIS/MLWIS), 
+and log-weighted interval scores (MWIS/MLWIS), 
 median weighted interval and log-weighted interval scores (MedWIS/MedLWIS), 
@@
 and log-weighted interval scores (RWIS/RLWIS).}
+\small
 \begin{tabular}{lrrrrrrrr}
@@
  for modeling the main trajectory of ILI along with modeling the discrepancy and
- specifically incorporate the smooth ASG function and compare forecasts using 
- the ASG with those using SIR. 
+ specifically incorporate the smooth ASG function.
  We model flu hospitalizations by incorporating the ILI forecast model into a 
@@
  from all previous seasons through hierarhcical modeling is able to further
- exploit the ILI data. Including model iscrepancy was an important piece of 
+ exploit the ILI data. Including model discrepancy was an important piece of 
  successful forecasting, but it required making thoughtful model constraints
@@
 forecasts when looking at the whole season. Though the ASGD\_NORM2 did very 
-well across all seasons, for some weeks and locations its perfomance was not
-as good. It should also not be assumed that ASGD\_NORM2 forecasts would perform
+well across the whole season, there were a small number of weeks or locations
+where performance was not as good.
+It should not be assumed that ASGD\_NORM2 forecasts would perform
 as well in every season given that each flu season will come with its own
@@
 closely related for neighboring states as shown in data plots in the
-supplementary materials. The one 2023 FluSight forecast model which outperformed
+supplementary materials. Notably, the one 2023 FluSight forecast model which 
+outperformed
 the ASGD\_NORM2, the UGA\_flucast-INFLAenza, was listed as being a spatial

### 2025-08-28 nonlinear_forecast.tex
add some score plots
finish responses to reviewers
update supplementary material
fix up arxiv manuscript

@@
 \usepackage{newtxmath} %for making indicator function
+\usepackage{changepage} %for adjusting the margins
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\pubyear{2024}
-\arxiv{2010.00000}
+\pubyear{2025}
+\arxiv{2412.14339}
 \volume{TBA}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author[A]{\snm{Spencer Wadsworth}\ead[label=e1]{sgw96@iastate.edu}}
+\author[A]{\snm{Spencer Wadsworth}\ead[label=e1]{iac25002@uconn.edu}}
 % \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}\orcid{0000-0000-0000-0000}}
 \and
-\author[A]{~\snm{Jarad Niemi}\ead[label=e2]{niemi@iastate.edu}}
+\author[B]{~\snm{Jarad Niemi}\ead[label=e2]{niemi@iastate.edu}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \address[A]{Department of Statistics,
+University of Connecticut, Storrs, CT 06269}
+\address[B]{Department of Statistics,
 Iowa State University, Ames, IA 50011}
@@
 % \end{supplement}
-B. Posterior distribution plots for select parameters\\
+B. Prior distribution assignments\\
+C. Posterior distribution plots for select parameters\\
 % \sdescription{2. Plots of posterior distributions for parameters in both ILI and hospitalization models}
-C. FluSight forecast competition scoring results
+D. Comparison of ILI/hospitalization forecast model combinations\\
+E. FluSight forecast competition scoring results in terms of WIS
 \end{supplement}

### 2025-08-28 supplemental_res.tex
fix up arxiv manuscript

@@
     \caption{Weekly hospitalization counts for the 50 US states for the 2022 
-    (grey) and 2023 (black) flu seasons}
+    (grey) and 2023 (black) flu seasons.}
     \label{fig:hosp_vs_week}
@@
 
-Figures \ref{fig:wis_cover_sum} and \ref{fig:state_and_date_wis} are 
-similar to figures \ref{fig:lwis_cover_sum} and \ref{fig:state_and_date_lwis}
+Figures \ref{fig:wis_cover_sum1} and \ref{fig:state_and_date_wis} are 
+similar to figures the main figures in the 2023 flu forecast assessment in the 
+main text
 except that they show WIS scores instead of LWIS scores. The results of 
 foreacst comparison for WIS scores is very similar to those of the LWIS in 
-terms of how well the $ASGD_NORM2$ forecasts did relative to the others from
+terms of how well the ASGD\_NORM2 forecasts did relative to the others from
 the FluSight competition.
@@
 \begin{figure}[hbt!]
-
-    % \centering
 \begin{adjustwidth}{-1.2cm}{-1.5cm}
+    % \begin{minipage}{\linewidth}
+    \centering
+    \makebox[\linewidth][l]{
     \includegraphics[scale = .54]{Images/wis_cover_sum.png}
+    }
+    \end{adjustwidth}
     \caption{a) Averaged over location and horizon weighted interval score
@@
     b) Boxplots of all WIS scores for 1, 2, 3, and 4-week ahead horizons
-    for the ASGD\_NORM2 (blue) and FluSight-baseline (red) models. 
+    for the ASGD\_NORM2 (blue), SIRD\_NORM2 (green), and
+    FluSight-baseline (red) models.
     To improve vizualization, the $y$-axis
     is on the log scale.}
-    \end{adjustwidth}
-    \label{fig:wis_cover_sum}
+    \label{fig:wis_cover_sum1}
+    % \end{minipage}
+
 \end{figure}
@@
 
+% \begin{figure}[hbt!]
+%     \label{fig:wis_cover_sum1}
+%     % \centering
+%     \begin{adjustwidth}{-1.2cm}{-1.5cm}
+%     \includegraphics[scale = .54]{Images/wis_cover_sum.png}
+%     \caption{a) Averaged over location and horizon weighted interval score
+%     (WIS) for each week during the season for the 21 models being compared. 
+%     Each grey line represents one model, but the ASGD\_NORM2 (blue),
+%     SIRD\_NORM2 (green), and 
+%     FluSight-baseline (red) are colored to stand out.
+%     b) Boxplots of all WIS scores for 1, 2, 3, and 4-week ahead horizons 
+%     for the ASGD\_NORM2 (blue), SIRD\_NORM2 (green), and
+%     FluSight-baseline (red) models. 
+%     To improve vizualization, the $y$-axis
+%     is on the log scale.}
+%     \end{adjustwidth}
+%     
+% \end{figure}
+
 \end{supplement}

### 2025-11-28 response_round2.tex
Start responses

\documentclass[ba]{imsart}

\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage[numbers]{natbib}
%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
\RequirePackage{graphicx}

%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
\usepackage{booktabs} %for the table in the analysis section
\usepackage{enumerate} %for the short algorithm
\usepackage{subfigure} %for making plots with multiple images
\usepackage{tabularx} %for the spacing in tabular
\usepackage{newtxmath} %for making indicator function
\usepackage{changepage} %for adjusting the margins
\usepackage{dsfont} %for getting the indicator function
\usepackage{mathtools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\1}[1]{\mathds{1}\left[#1\right]}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{amsmath}

\usepackage{xcolor}
\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
\newcommand{\spencer}[1]{{\color{red} Response: #1}}

\begin{document}

\Huge
\noindent AE1
\normalsize
\\~\\

I think the authors did a great job in this newly submitted version and 
addressed most of the earlier raised comments in a nice and convincing way. 
The paper is well written and interesting to read.

I only have two remaining comments.

\begin{enumerate}[-]

  \item Section 3.7. End of of first paragraph: ``..., but propagation of 
  uncertainty   in ILI forecasts was stil permitted in the estimating of 
  hospitalisation forecasts.``

  Please clarify more explicitly how uncertainty was propagated between the two 
  model components. My understanding is that this was achieved by incorporating 
  not only point estimates of the parameters but also samples from the predictive 
  distribution of the ILI model. If this is correct, please state it clearly 
  already at this stage and consider adding a brief discussion of the advantages 
  and disadvantages of this approach.
  \\~\\
  \item I recommend including the statement raised by the reviewer in the first 
  submission—that WIS approximates CRPS (ideally supported with a reference)—and 
  therefore leads to essentially the same conclusions. I think the motivation for 
  using WIS instead of CRPS is well explained, especially given that WIS is the 
  primary evaluation measure used in the FluSight competition 
  \\~\\

  \emph{The following statements were added to page 17.}
  \begin{quote}
  \emph{
      Commonly used proper scoring rules for continuous distribution forecasts
      include the log score and the continuously
      ranked probability score (CRPS).}
      \\~\\  
  \emph{
      We note that as the number of forecast intervals $B$ increases, the WIS becomes 
      arbitrarily close to
      the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
      \cite[]{bracher2021evaluating}.}
      \\~\\
\end{quote}

\end{enumerate}

\Huge
\noindent R1
\normalsize
\\~\\
\noindent The authors have addressed all my comments from the first round. I only have two remaining points
on the extended empirical analysis on FluSight.

\begin{enumerate}[1.]

  \item The authors mention the difficulty of missing submissions during the FluSight season (hence the
    display chosen for Figure 10). How was this handled for Table 2? Do the averages refer to different
    sets of forecasting tasks for each model? If so the results are not fully comparable. The referenced
    paper by Cramer et al contains a (heuristic) scheme to compare scores in such a setting, which
    resembles what was done for Figure 10. This may be more appropriate than reporting raw averages.
    Models with very small submission fractions (maybe those with less than 80\%, definitely those with
    less than 50\%) in my opinion should just be removed altogether.
    \\~\\
    \emph{The columns RLWIS and RWIS in table 2 correspond directly to the 
          values plotted in figure 10. We rearranged the table so that RLWIS and
          RWIS come first. The updated table, shown as "Table 1" is below.
          To clarify what's given in table 2 we stated the
          following.}

    \begin{quote}
      Table 2 gives the numerical values for RLWIS 
      and relative WIS (RWIS) over
      all locations and weeks for each model, relative to the ASGD\_NORM2 model.
      It also includes the non relative mean WIS and LWIS for each model, that is the 
      overall mean scores for all forecasts submitted regardless of submission overlap
      with another model.
    \end{quote}

    \begin{quote}
      During the FluSight competition, most participating teams would occasionally 
      miss a forecast submission for some week, location, or horizon.
      The FluSight 
      models which are used for comparison are the
      13 non-ensemble models which submitted forecasts for at least 80\% of 
      FluSight targets during the 2023 season.

      Because forecast submissions for some targets were occassionally missed
      by participating teams, for any two models there was often a mismatch between
      targets for which forecasts were included or exlcuded. This makes a
      simple average of scores is an imperfect measure for comparison.
    \end{quote}

  \item It is uncommon to report median scores as their statistical properties are unclear. As mean score
    are also there I'm not fundamentally opposed to reporting them, but in my opinion mean scores
    alone are sufficient.  
    \\~\\
    \emph{Thank you for this comment. Because of it, we removed the median scores
    from the table which improved the aesthetic of the table as well as simplified
    it. See the new table below.}
\\~\\    
\end{enumerate}

\begin{table}[ht]
\centering
% \begin{adjustwidth}{-3cm}{-1.5cm}
\caption{Overall summary scores for each of the 13 non-ensemble FluSight
models with over 80\% submissions and 
the ASGD\_NORM2 and SIRD\_NORM2 models.
Summaries include 
mean weighted interval the relative log-weighted interval
and weighted interval scores (RLWIS/RWIS),
the overall mean log-weighted and weighted interval scores (MLWIS/MWIS), 
mean squared error
difference between predictive model empirical and theoretical coverage (MSEC)
see figure 9 c)), and percentage of forecasts 
submitted for all 
targets during the 2023 season (\% Forcs).}
\small
\begin{tabular}{lrrrrrrrr}
  % \hline
                              & RLWIS & RWIS & MLWIS & MWIS & MSEC  & \% Forcs\\ 
  \hline 
  ASGD\_NORM2                 & 1     & 1    & 0.34  & 62   & 0.001 & 100 \\ 
  PSI-PROF                    & 0.98  & 0.98 & 0.35  & 63   & 0.003 & 100 \\ 
  SIRD\_NORM2                 & 0.95  & 0.93 & 0.36  & 67   & 0.001 & 100 \\ 
  CEPH-Rtrend\_fluH           & 0.87  & 0.89 & 0.39  & 69   & 0.004 & 100 \\ 
  FluSight-baseline           & 0.83  & 0.73 & 0.41  & 85   & 0.023 & 100 \\ 
  UM-DeepOutbreak             & 0.74  & 0.73 & 0.46  & 83   & 0.010 & 99 \\ 
  cfarenewal-cfaepimlight     & 0.93  & 0.82 & 0.37  & 76   & 0.001 & 99 \\ 
  MOBS-GLEAM\_FLUH            & 0.97  & 0.95 & 0.35  & 66   & 0.021 & 98 \\ 
  LosAlamos\_NAU-CModel\_Flu  & 0.54  & 0.26 & 0.64  & 242  & 0.153 & 93 \\ 
  UGuelph-CompositeCurve      & 0.66  & 0.50 & 0.52  & 132  & 0.018 & 93 \\ 
  LUcompUncertLab-chimera     & 0.73  & 0.76 & 0.43  & 84   & 0.004 & 92 \\ 
  UGA\_flucast-INFLAenza      & 1.09  & 0.90 & 0.32  & 74   & 0.003 & 90 \\ 
  cfa-flumech                 & 0.86  & 0.65 & 0.41  & 101  & 0.019 & 89 \\ 
  UGA\_flucast-Copycat        & 0.99  & 0.89 & 0.35  & 76   & 0.012 & 87 \\ 
  GT-FluFNP                   & 0.74  & 0.56 & 0.48  & 105  & 0.046 & 86 \\ 
  % UNC\_IDD-InfluPaint         & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
  % NU\_UCSD-GLEAM\_AI\_FLUH    & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
  % SigSci-CREG                 & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
  % PSI-PROF\_beta              & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
  % NIH-Flu\_ARIMA              & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
  % GH-model                    & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
  % UGA\_flucast-OKeeffe        & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
\end{tabular}
% \end{adjustwidth}
\label{tab:fin_analysis_stats}
\end{table}

\newpage

\Huge
\noindent R2
\normalsize
\\~\\
Forecasting Influenza Hospitaliztions Using a Bayesian Hierarchical Nonlinear 
Model with Discrepancy

This manuscript seems to be a very reasonable attempt at forecasting influenza 
induced hospitalizations. The
manuscript also puts the methodology in the context of the FluSight competition 
organised by the CDC in the US.
Apologies, I have found it hard to review as it is conceptually very hard to do 
so in my opinion. Firstly,
I do not find comfortable modelling the proportion, rather a natural outcome to 
model will be the number of hospitalization using a Generalized Linear model 
(Poisson distribution). Secondly, I find it un-intuitive to model the
flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model 
we focus on two aspects: the systematic part and the random part. Here, as far 
as  I understand the systematic part is the ASG function. The random
part has several components, which are hidden in Equation (5). My difficulty in 
modelling using the ASG
function lies in my inability to relate any meaningful meteorological variables 
(which are very much drivers of the seasonal flu outbreaks). See for example 
the paper,  "A hierarchical Bayesian model for improving short-term forecasting 
of hospital demand by including meteorological information" by  SK Sahu, 
B Baffour, PR Harper, JH Minty, C Sarran
Journal of the Royal Statistical Society, Series A (2014).

In spite of the above criticisms, I appreciate the hierarchical modeling 
approach of fitting the two models separately in components. But I wonder how 
the uncertainty has been propagated from one component model to another. 
Cutting the feedback idea is well understood but one still needs to integrate 
uncertainty to get an overall measure of the uncertainty in the forecasts. 
I did not clearly understand this issue from my reading of  the paper. I can 
suggest one way to propagate this uncertainty is by using MCMC samples 
(and methods), see for example, the paper,  "A rigorous statistical framework 
for spatio-temporal pollution prediction and estimation of its long-term impact 
on health" by 
D Lee, S Mukhopadhyay, A Rushworth, and SK Sahu, in Biostatistics (2017).

In conclusion, I do think that the authors have done a good job of analysis and 
simulation based demonstrations of the results. These can be further improved 
by including some better measures of assessing prediction/forecasting accuracy, 
e.g. the CRPS (Continuous Ranked Probability Score), see for example Chapter 6 
of the book
"Bayesian modeling of spatio-temporal data with R" by Sahu 
(Chapman and Hall, 2022). 
\\~\\
\emph{In the simulation study in section 4 of the first version of this paper, 
scoring was
done using the CRPS, but in the FluSight analysis in section 5 scoring was done
using the WIS because the available forecasts are not amenable to being scored
the with the CRPS. A recommendation from one of the reviewers was to simplify 
by choosing one or the other scores, and we chose the WIS. To respond to this 
point, however, we modified the first paragraph on page 17. The paragraph is 
below and the added statements are bolded.
}

\begin{quote}
Model comparison was done by calculating a proper scoring rule for each 
forecast.
Proper scoring rules are the 
current standard for comparing performance between probabilistic forecasts 
and selecting the best forecasts according to the notion of maximizing 
sharpness subject to (auto-)calibration 
\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
rules are commonly used in forecast comparison, and they are designed such 
that a 
forecaster is incentivized to be honest in the reporting of their forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic}.
\textbf{Commonly used proper scoring rules for continuous distribution forecasts
include the log score and the continuously
ranked probability score (CRPS).} 
The proper scoring rule primarily used in the FluSight competition and the 
COVID-19 Forecast Hub, and which is used in this simulation study,
is the negatively oriented
(smaller is better)
weighted interval score (WIS)
\cite[]{mathis2024evaluation, bracher2021evaluating}.
The WIS is used for scoring quantile or interval 
forecasts \textbf{--the forecast format used in both FluSight and the 
COVID-19 Forecast Hub--}
or forecasts made up of predicitive intervals of several nominal 
levels
\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating}. 
The WIS is 
defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
interval where $b \in \{1, ..., B\}$. 
$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
single interval, defined in (\ref{eq:is}). Here 
$\vmathbb{1}\{\cdot\}$ is the indicator function.
\textbf{We note that as the number of forecast intervals $B$ increases, the WIS becomes 
arbitrarily close to
the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
\cite[]{bracher2021evaluating}.}
\end{quote}

\bibliographystyle{ba}
\bibliography{master_bib}

\end{document}

### 2025-11-28 supplemental_res.tex
Start responses

@@
     \makebox[\linewidth][l]{
-    \includegraphics[scale = .54]{Images/wis_cover_sum.png}
+    \includegraphics[scale = .54]{Images/wis_cover_sum2.png}
     }
@@
     \caption{a) Averaged over location and horizon weighted interval score
-    (WIS) for each week during the season for the 21 models being compared.
+    (WIS) for each week during the season for the 15 models being compared.
     Each grey line represents one model, but the ASGD\_NORM2 (blue),
@@
     \makebox[\linewidth][l]{
-    \includegraphics[scale = .5]{Images/state_and_date_wis.png}
+    \includegraphics[scale = .5]{Images/state_and_date_wis2.png}
     }

### 2025-11-28 make_images/make_final_analysis_images.R

@@ -102,6 +102,9 @@ state_ord <- team_scores %>%
   arrange(sldiff)
 
 state_ldiff <- team_diffs %>% 
+  filter(!(model %in% c("UNC_IDD-InfluPaint", "NU_UCSD-GLEAM_AI_FLUH",
+                        "SigSci-CREG", "PSI-PROF_beta", "NIH-Flu_ARIMA",
+                        "GH-model", "UGA_flucast-OKeeffe"))) %>% 
   filter(model != "asg_disc2_nm_hosp_sq_ar1") %>%
   mutate(model = factor(model, levels = scores_ord$model),
          loc_abb = factor(loc_abb, levels = state_ord$loc_abb)) %>% 
@@ -134,6 +137,9 @@ state_ldiff <- team_diffs %>%
 
 
 state_diff <- team_diffs %>% 
+  filter(!(model %in% c("UNC_IDD-InfluPaint", "NU_UCSD-GLEAM_AI_FLUH",
+                        "SigSci-CREG", "PSI-PROF_beta", "NIH-Flu_ARIMA",
+                        "GH-model", "UGA_flucast-OKeeffe"))) %>% 
   filter(model != "asg_disc2_nm_hosp_sq_ar1") %>%
   mutate(model = factor(model, levels = scores_ord$model),
          loc_abb = factor(loc_abb, levels = state_ord$loc_abb)) %>% 
@@ -177,6 +183,9 @@ state_diff <- team_diffs %>%
 
 
 team_ddiffs <- team_scores %>% 
+  filter(!(model %in% c("UNC_IDD-InfluPaint", "NU_UCSD-GLEAM_AI_FLUH",
+                        "SigSci-CREG", "PSI-PROF_beta", "NIH-Flu_ARIMA",
+                        "GH-model", "UGA_flucast-OKeeffe"))) %>% 
   filter(model != "asg_disc2_nm_hosp_sq_ar1") %>% 
   left_join(best_nl, by = c("reference_date", "loc_abb", "horizon")) %>% 
   group_by(model, reference_date) %>% 
@@ -201,6 +210,9 @@ library(ggtext)
 # labels <- levels(team_ddiffs$model)
 # labels[labels == "FluSight-baseline"] <- "<b>FluSight-baseline</b>"
 date_ldiff <- team_ddiffs %>%
+  filter(!(model %in% c("UNC_IDD-InfluPaint", "NU_UCSD-GLEAM_AI_FLUH",
[…]

### 2025-12-01 CopyOfmanuscript_res.tex
fix section 3.7

--- 2025-08-28 manuscript_res.tex
@@
 \begin{figure}[hbt!]
+    \centering
     \includegraphics[scale=.2]{Images/ili_v_svat_vs_lagdiff.png}
@@
 described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
-into two component models or ``modules`` may be justfied for several reasons
+into two component models or ``modules`` may be justified for several reasons
 including to reduce time to fit the model, avoiding mixing issues of posterior
@@
 The decision to cut the forecast model into two component models
-was made for several reasons.
+was made for three major reasons.
 When fitting the fully Bayesian joint model, the time to fit a single model using
@@
 \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
+Commonly used proper scoring rules for continuous distribution forecasts
+include the log score and the continuously
+ranked probability score (CRPS). 
 The proper scoring rule primarily used in the FluSight competition and the 
@@
 The WIS is used for scoring quantile or interval 
-forecasts, or forecasts made up of predicitive intervals of several nominal 
+forecasts --the forecast format used in both FluSight and the 
+COVID-19 Forecast Hub--
+or forecasts made up of predicitive intervals of several nominal 
 levels
 \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating}. 
-The WIS defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
+The WIS is 
+defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
 included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
@@
 $\vmathbb{1}\{\cdot\}$ is the indicator function.
+We note that as the number of forecast intervals $B$ increases, the WIS becomes 
+arbitrarily close to
+the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
+\cite[]{bracher2021evaluating}.
 
@@
 \begin{figure}[hbt!]
-    % \centering
-    \begin{adjustwidth}{-.65cm}{-1.5cm}
+    \centering
+    % \begin{adjustwidth}{-.65cm}{-1.5cm}
     \includegraphics[scale=.45]{Images/all_sim_res.png}
-    \end{adjustwidth}
+    % \end{adjustwidth}
     \caption{Results of forecasts for simulated hospitalizations for ASGD, ASG,
@@
   \centering
-  \begin{adjustwidth}{-1cm}{-1.9cm}
+  % \begin{adjustwidth}{-1cm}{-1.9cm}
   \includegraphics[scale=.45]{Images/normal_sq_forecasts_us.png}
-  \end{adjustwidth}
+  % \end{adjustwidth}
 \caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
@@
 For the
-remainder of this section, focus will be made on forecasts made by coupling the
+remainder of this section, the focus is on forecasts made by coupling the
 ASGD and SIRD models with the 
@@
 
+During the FluSight competition, most participating teams would occasionally 
+miss a forecast submission for some week, location, or horizon.
 The FluSight 
 models which are used for comparison are the
-20 non-ensemble models
-submitted to FluSight during the 2023 season.
+13 non-ensemble models which submitted forecasts for at least 80\% of 
+FluSight targets during the 2023 season. 
 The submitted forecasts may be found at \cite{mathis2023flusight}.
@@
 Figure \ref{fig:lwis_cover_sum} shows forecast results for the 2023 flu 
-hospitalization forecasts for all 22 models being compared but with results
+hospitalization forecasts for all 15 models being compared but with results
 for the ASGD\_NORM2, SIRD\_NORM2, and FluSight baseline models highlighted. 
@@
 
-    % \centering
-    \begin{adjustwidth}{0cm}{0cm}
-    \includegraphics[scale = .54]{Images/lwis_cover_sum.png}
+    \centering
+    % \begin{adjustwidth}{0cm}{0cm}
+    \includegraphics[scale = .6]{Images/lwis_cover_sum2.png}
     \caption{a) Log-weighted interval score (LWIS) averaged 
@@
     FluSight-baseline (red) are colored to stand out.}
-    \end{adjustwidth}
+    % \end{adjustwidth}
     \label{fig:lwis_cover_sum}
@@
 
-During the FluSight competition, most participating teams would occasionally 
-miss a forecast submission for some week, location, or horizon. This makes a 
-simple average of scores an imperfect measure for comparison, as forecasts
-of certain targets included for one model may be excluded by another model.
-Thus in figure \ref{fig:state_and_date_lwis} 
+Because forecast submissions for some targets were occassionally missed
+by participating teams, for any two models there was often a mismatch between
+targets for which forecasts were included or exlcuded. This makes a
+simple average of scores is an imperfect measure for comparison.
+Thus in figure \ref{fig:state_and_date_lwis}, 
 the ASGD\_NORM2 model is compared with all 
@@
 
-Table \ref{tab:fin_analysis_stats} gives summary results for mean and median
-WIS and LWIS scores, predictive coverage, and relative RWIS and RLWIS scores
-for each model. Along with the plots in this section, it 
+Table \ref{tab:fin_analysis_stats} gives the numerical values for RLWIS 
+and relative WIS (RWIS) over
+all locations and weeks for each model, relative to the ASGD\_NORM2 model.
+It also includes the non relative mean WIS and LWIS for each model, that is the 
+overall mean scores for all forecasts submitted regardless of submission overlap
+with another model.
+Along with the plots in this section, it 
 shows that over the course of the 
@@
 \begin{figure}[hbt!]
-\begin{adjustwidth}{-1.2cm}{-1.5cm}
+% \begin{adjustwidth}{-1.2cm}{-1.5cm}
     % \begin{minipage}{\linewidth}
@@
     \makebox[\linewidth][l]{
-    \includegraphics[scale = .5]{Images/state_and_date_lwis.png}
+    \includegraphics[scale = .5]{Images/state_and_date_lwis2.png}
     }
-    \end{adjustwidth}
+    % \end{adjustwidth}
     \caption{Ratio of average log-weighted interval score (LWIS) for ASGD\_NORM2 
@@
 \centering
-\begin{adjustwidth}{-3cm}{-1.5cm}
-\caption{Overall summary scores for each of the 20 non-ensemble FluSight
-models and the ASGD\_NORM2 model. Summaries include mean weighted interval
+% \begin{adjustwidth}{-3cm}{-1.5cm}
+\caption{Overall summary scores for each of the 13 non-ensemble FluSight
+models with over 80\% submissions and 
+the ASGD\_NORM2 and SIRD\_NORM2 models.
+Summaries include 
+mean weighted interval the relative log-weighted interval
+and weighted interval scores (RLWIS/RWIS),
+the overall mean log-weighted and weighted interval scores (MLWIS/MWIS), 
+mean squared error
+difference between predictive model empirical and theoretical coverage (MSEC)
+see figure \ref{fig:lwis_cover_sum} c)), and percentage of forecasts 
+submitted for all 
+targets during the 2023 season (\% Forcs).}
+\small
+\begin{tabular}{lrrrrrrrr}
+  % \hline
+                              & RLWIS & RWIS & MLWIS & MWIS & MSEC  & \% Forcs\\ 
+  \hline 
+  ASGD\_NORM2                 & 1     & 1    & 0.34  & 62   & 0.001 & 100 \\ 
+  PSI-PROF                    & 0.98  & 0.98 & 0.35  & 63   & 0.003 & 100 \\ 
+  SIRD\_NORM2                 & 0.95  & 0.93 & 0.36  & 67   & 0.001 & 100 \\ 
+  CEPH-Rtrend\_fluH           & 0.87  & 0.89 & 0.39  & 69   & 0.004 & 100 \\ 
+  FluSight-baseline           & 0.83  & 0.73 & 0.41  & 85   & 0.023 & 100 \\ 
+  UM-DeepOutbreak             & 0.74  & 0.73 & 0.46  & 83   & 0.010 & 99 \\ 
+  cfarenewal-cfaepimlight     & 0.93  & 0.82 & 0.37  & 76   & 0.001 & 99 \\ 
+  MOBS-GLEAM\_FLUH            & 0.97  & 0.95 & 0.35  & 66   & 0.021 & 98 \\ 
+  LosAlamos\_NAU-CModel\_Flu  & 0.54  & 0.26 & 0.64  & 242  & 0.153 & 93 \\ 
+  UGuelph-CompositeCurve      & 0.66  & 0.50 & 0.52  & 132  & 0.018 & 93 \\ 
+  LUcompUncertLab-chimera     & 0.73  & 0.76 & 0.43  & 84   & 0.004 & 92 \\ 
+  UGA\_flucast-INFLAenza      & 1.09  & 0.90 & 0.32  & 74   & 0.003 & 90 \\ 
+  cfa-flumech                 & 0.86  & 0.65 & 0.41  & 101  & 0.019 & 89 \\ 
+  UGA\_flucast-Copycat        & 0.99  & 0.89 & 0.35  & 76   & 0.012 & 87 \\ 
+  GT-FluFNP                   & 0.74  & 0.56 & 0.48  & 105  & 0.046 & 86 \\ 
+  % UNC\_IDD-InfluPaint         & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
+  % NU\_UCSD-GLEAM\_AI\_FLUH    & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
+  % SigSci-CREG                 & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
+  % PSI-PROF\_beta              & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
+  % NIH-Flu\_ARIMA              & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
+  % GH-model                    & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
+  % UGA\_flucast-OKeeffe        & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
+\end{tabular}
+% \end{adjustwidth}
+\label{tab:fin_analysis_stats}
+\end{table}
+
+\begin{table}[ht]
+\centering
+% \begin{adjustwidth}{-3cm}{-1.5cm}
+\caption{Overall summary scores for each of the 13 non-ensemble FluSight
+models with over 80\% submissions and 
+the ASGD\_NORM2 and SIRD\_NORM2 models. 
+Summaries include mean weighted interval
 and log-weighted interval scores (MWIS/MLWIS), 
-median weighted interval and log-weighted interval scores (MedWIS/MedLWIS), 
 mean squared error
@@
 targets during the 2023 season (\% Forcs), and relative weighted interval
-and log-weighted interval scores (RWIS/RLWIS).}
+and log-weighted interval scores (RWIS/RLWIS).
+\\}
 \small
@@
   % \hline
- & MWIS & MLWIS & MedWIS & MedLWIS & MSEC & \% Forcs & RWIS & RLWIS\\ 
+                              & MWIS & MLWIS & MSEC  & \% Forcs & RWIS & RLWIS\\ 
   \hline 
-ASGD\_NORM2 & 62 & 0.34 & 14 & 0.23 & 0.001 & 100 & 1 & 1 \\ 
-  PSI-PROF & 63 & 0.35 & 15 & 0.23 & 0.003 & 100 & 0.90 & 0.97 \\ 
-  SIRD\_NORM2 & 67 & 0.36 & 16 & 0.26 & 0.001 & 100 & 0.93 & 0.95 \\
-  CEPH-Rtrend\_fluH & 69 & 0.39 & 13 & 0.28 & 0.004 & 100 & 0.81 & 0.86 \\ 
-  FluSight-baseline & 85 & 0.41 & 16 & 0.32 & 0.023 & 100 & 0.63 & 0.82 \\ 
-  UM-DeepOutbreak & 83 & 0.46 & 21 & 0.39 & 0.010 & 99 & 0.64 & 0.72 \\ 
-  cfarenewal-cfaepimlight & 76 & 0.37 & 15 & 0.26 & 0.001 & 99 & 0.72 & 0.92 \\ 
-  MOBS-GLEAM\_FLUH & 66 & 0.35 & 15 & 0.22 & 0.021 & 98 & 0.87 & 0.95 \\ 
-  LosAlamos\_NAU-CModel\_Flu & 242 & 0.64 & 24 & 0.42 & 0.153 & 93 & 0.21 & 0.53 \\ 
-  UGuelph-CompositeCurve & 132 & 0.52 & 25 & 0.41 & 0.018 & 93 & 0.43 & 0.62 \\ 
-  LUcompUncertLab-chimera & 84 & 0.43 & 17 & 0.33 & 0.004 & 92 & 0.67 & 0.71 \\ 
-  UGA\_flucast-INFLAenza & 74 & 0.32 & 15 & 0.20 & 0.003 & 90 & 0.77 & 1.07 \\ 
-  cfa-flumech & 101 & 0.41 & 17 & 0.27 & 0.019 & 89 & 0.58 & 0.85 \\ 
-  UGA\_flucast-Copycat & 76 & 0.35 & 16 & 0.21 & 0.012 & 87 & 0.81 & 0.98 \\ 
-  GT-FluFNP & 105 & 0.48 & 18 & 0.38 & 0.046 & 86 & 0.45 & 0.73 \\ 
-  UNC\_IDD-InfluPaint & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
-  NU\_UCSD-GLEAM\_AI\_FLUH & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
-  SigSci-CREG & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
-  PSI-PROF\_beta & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
-  NIH-Flu\_ARIMA & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
-  GH-model & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
-  UGA\_flucast-OKeeffe & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
+ASGD\_NORM2                   & 62   & 0.34  & 0.001 & 100      & 1    & 1 \\ 
+  PSI-PROF                    & 63   & 0.35  & 0.003 & 100      & 0.90 & 0.97 \\ 
+  SIRD\_NORM2                 & 67   & 0.36  & 0.001 & 100      & 0.93 & 0.95 \\
+  CEPH-Rtrend\_fluH           & 69   & 0.39  & 0.004 & 100      & 0.81 & 0.86 \\ 
+  FluSight-baseline           & 85   & 0.41  & 0.023 & 100      & 0.63 & 0.82 \\ 
+  UM-DeepOutbreak             & 83   & 0.46  & 0.010 & 99       & 0.64 & 0.72 \\ 
+  cfarenewal-cfaepimlight     & 76   & 0.37  & 0.001 & 99       & 0.72 & 0.92 \\ 
+  MOBS-GLEAM\_FLUH            & 66   & 0.35  & 0.021 & 98       & 0.87 & 0.95 \\ 
+  LosAlamos\_NAU-CModel\_Flu  & 242  & 0.64  & 0.153 & 93       & 0.21 & 0.53 \\ 
+  UGuelph-CompositeCurve      & 132  & 0.52  & 0.018 & 93       & 0.43 & 0.62 \\ 
+  LUcompUncertLab-chimera     & 84   & 0.43  & 0.004 & 92       & 0.67 & 0.71 \\ 
+  UGA\_flucast-INFLAenza      & 74   & 0.32  & 0.003 & 90       & 0.77 & 1.07 \\ 
+  cfa-flumech                 & 101  & 0.41  & 0.019 & 89       & 0.58 & 0.85 \\ 
+  UGA\_flucast-Copycat        & 76   & 0.35  & 0.012 & 87       & 0.81 & 0.98 \\ 
+  GT-FluFNP                   & 105  & 0.48  & 0.046 & 86       & 0.45 & 0.73 \\ 
+  % UNC\_IDD-InfluPaint         & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
+  % NU\_UCSD-GLEAM\_AI\_FLUH    & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
+  % SigSci-CREG                 & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
+  % PSI-PROF\_beta              & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
+  % NIH-Flu\_ARIMA              & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
+  % GH-model                    & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
+  % UGA\_flucast-OKeeffe        & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
 \end{tabular}
-\end{adjustwidth}
+% \end{adjustwidth}
 \label{tab:fin_analysis_stats}
 \end{table}
+
+% \begin{table}[ht]
+% \centering
+% % \begin{adjustwidth}{-3cm}{-1.5cm}
+% \caption{Overall summary scores for each of the 13 non-ensemble FluSight
+% models with over 80\% submissions and 
+% the ASGD\_NORM2 and SIRD\_NORM2 models. 
+% Summaries include mean weighted interval
+% and log-weighted interval scores (MWIS/MLWIS), 
+% median weighted interval and log-weighted interval scores (MedWIS/MedLWIS), 
+% mean squared error
+% difference between predictive model empirical and theoretical coverage (MSEC)
+% see figure \ref{fig:lwis_cover_sum} c)), percentage of forecasts for all 
+% targets during the 2023 season (\% Forcs), and relative weighted interval
+% and log-weighted interval scores (RWIS/RLWIS).}
+% \small
+% \begin{tabular}{lrrrrrrrr}
+%   % \hline
+%                               & MWIS & MLWIS & MedWIS & MedLWIS & MSEC  & \% Forcs & RWIS & RLWIS\\ 
+%   \hline 
+% ASGD\_NORM2                   & 62   & 0.34  & 14     & 0.23    & 0.001 & 100      & 1    & 1 \\ 
+%   PSI-PROF                    & 63   & 0.35  & 15     & 0.23    & 0.003 & 100      & 0.90 & 0.97 \\ 
+%   SIRD\_NORM2                 & 67   & 0.36  & 16     & 0.26    & 0.001 & 100      & 0.93 & 0.95 \\
+%   CEPH-Rtrend\_fluH           & 69   & 0.39  & 13     & 0.28    & 0.004 & 100      & 0.81 & 0.86 \\ 
+%   FluSight-baseline           & 85   & 0.41  & 16     & 0.32    & 0.023 & 100      & 0.63 & 0.82 \\ 
+%   UM-DeepOutbreak             & 83   & 0.46  & 21     & 0.39    & 0.010 & 99       & 0.64 & 0.72 \\ 
+%   cfarenewal-cfaepimlight     & 76   & 0.37  & 15     & 0.26    & 0.001 & 99       & 0.72 & 0.92 \\ 
+%   MOBS-GLEAM\_FLUH            & 66   & 0.35  & 15     & 0.22    & 0.021 & 98       & 0.87 & 0.95 \\ 
+%   LosAlamos\_NAU-CModel\_Flu  & 242  & 0.64  & 24     & 0.42    & 0.153 & 93       & 0.21 & 0.53 \\ 
+%   UGuelph-CompositeCurve      & 132  & 0.52  & 25     & 0.41    & 0.018 & 93       & 0.43 & 0.62 \\ 
+%   LUcompUncertLab-chimera     & 84   & 0.43  & 17     & 0.33    & 0.004 & 92       & 0.67 & 0.71 \\ 
+%   UGA\_flucast-INFLAenza      & 74   & 0.32  & 15     & 0.20    & 0.003 & 90       & 0.77 & 1.07 \\ 
+%   cfa-flumech                 & 101  & 0.41  & 17     & 0.27    & 0.019 & 89       & 0.58 & 0.85 \\ 
+%   UGA\_flucast-Copycat        & 76   & 0.35  & 16     & 0.21    & 0.012 & 87       & 0.81 & 0.98 \\ 
+%   GT-FluFNP                   & 105  & 0.48  & 18     & 0.38    & 0.046 & 86       & 0.45 & 0.73 \\ 
+%   % UNC\_IDD-InfluPaint         & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
+%   % NU\_UCSD-GLEAM\_AI\_FLUH    & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
+%   % SigSci-CREG                 & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
+%   % PSI-PROF\_beta              & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
+%   % NIH-Flu\_ARIMA              & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
+%   % GH-model                    & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
+%   % UGA\_flucast-OKeeffe        & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
+% \end{tabular}
+% % \end{adjustwidth}
+% \label{tab:fin_analysis_stats}
+% \end{table}
 

### 2025-12-01 response_round1.tex
Fix up reviewer comments

--- 2025-11-28 response_round2.tex
@@
 \begin{document}
+\Huge
+\noindent Response to reviewers
+\\~\\
+\normalsize
+We are grateful to the three reviewers and their insightful comments and 
+questions. Overall, they made a huge difference in improving the content of
+our manuscript. In the most recent revision, we are particularly grateful for 
+the questions on uncertainty propogation. Obviously the explanation we gave
+for how uncertainty propogation wasn't clear enough, so in this iteration of 
+ we reworked sections 3.7 and 3.8 to make more clear how 
+uncertainty propogation was done. Specifics are found below and in the 
+revised manuscript. Our responses to each reviewers particular comments are
+\emph{italicized}, and specific changes/additions to the main manuscript are 
+indented.
+\\~\\
 
@@
   \\~\\
+  \emph{To clarify how uncertainty was propogated, we consildated sections
+  3.7 and 3.8, reordered the main ideas, and removed redundant information. We
+  also added some advantages/disadvantages of our approach. The changes now
+  read:}
+  \begin{quote}
+        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
+        \label{sec:implementation_posterior}
+        \emph{
+        Rather than combine the ILI and hopsitalization models into a single model,
+        (1) and 
+        (6) were fit separately via Bayesian posterior updating 
+        in two stages and then fused together to produce hospitalization 
+        forecasts. In the first stage, future ILI 
+        forecasts were obtained from the posterior predictive distribution where for
+        week $w$, the predictive distribution was obtained by integrating over the 
+        parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
+        as in (\ref{eq:ili_post}) where 
+        $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
+        is the density function of the posterior distribution for the model parameters
+        and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$. 
+        The desired forecasts were for weeks $w + i$ 
+        where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
+        predicted or forecasted ILI at $i$ weeks in the future.}
+
+        \begin{equation}
+            \label{eq:ili_post}
+            \tag{7}
+            p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
+            p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
+            \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
+            \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
+            d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+        \end{equation}
+        % \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
+        % \spencer{Isn't that captured in $d \pi_s$?}
+
+        \emph{\indent In the second stage, the posterior distribution
+        for parameters of model (6) given observed
+        hospitalization data $\textbf{H}_w$, observed ILI data, and
+        denoted as 
+        $p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
+        was obtained.
+        The posterior predictive in (\ref{eq:ili_post}) was then combined with 
+        the hospitalization parameter posterior distribution
+        to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
+        for week $w + i$.
+        Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
+        parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
+        That is, the first stage component model produced $K$ draws
+        $(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
+        and the second stage component model produced
+        $K$ parameter draws 
+        $(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
+        $(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
+        ILI and parameter posterior distribution draws was then plugged into equation
+        (6) to produce}
+
+        \[
+         \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
+                                          \alpha_{1s}^{(k)} 
+                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
+                                          \alpha_{2s}^{(k)} 
+                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
+                                          \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
+                                          \epsilon_{s,w + i}^{(k)}
+        \]
+        \emph{where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
+        $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
+        from which
+        the probabilistic hospitalization forecast was estimated. In practice we set
+        $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
+        \\~\\
+        \indent Occasionally $\{H^{(k)}_{s,w + i}\}^K$ included a small number of negative 
+        values,
+        which do not make sense when the distribution is meant to forecast
+        hospitalizations, a nonnegative number. We attempted to alleviate this problem
+        by modeling log-hospitalizations or by assuming hospitalizations followed a
+        distribution truncated at 0, but these models led to poorer forecasts
+        and issues drawing from the posterior distribution.
+        For the forecast competition analysis in 
+        section 5, all negative values of $\{H^{(k)}_{s,w + i}\}^K$ 
+        were 
+        set to 0 to reflect realistic values of hospitalizations and comply with the 
+        FluSight forecasting rules.}
+        \\~\\
+        \emph{
+        \indent In the two-stage scheme, the posterior distribution of 
+        model (1) parameters was not directly influenced by the 
+        hospitalization data and the posterior distribution of model 
+        (6) parameters was not directly influenced by model 
+        (1) parameters. Hence there was limited feedback between the
+        two components, but propagation of uncertainty in ILI forecasts was still 
+        permitted in the estimating of hospitalization forecasts.
+        This two-stage modeling is remeniscent of the Bayesian cut 
+        described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
+        into two component models or ``modules`` may be justified for several reasons
+        including to reduce time to fit the model, avoiding mixing issues of posterior
+        sampling distributions, improving predictive performance or to prevent
+        model misspecification in one model to influence the other 
+        \cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, 
+        lee2017rigorous, plummer2015cuts}. A drawback
+        from cutting a model is potential loss of useful information. 
+        In our current 
+        application for example, the ILI model parameters receive no potentially 
+        useful information from hospiatlization data.
+        \\~\\
+        \indent The decision to cut the forecast model into two component models
+        was made for three major reasons.
+        When fitting the fully Bayesian joint model, the time to fit a single model 
+        using
+        MCMC sampling and
+        obtain forecasts was sometimes many hours whereas fitting the two models
+        separately took a fraction of the time. When assessing convergence of model
+        parameters, it was rare that posterior sampling chains for the joint model
+        showed convergence even after 55,000 iterations, whereas the component model 
+        fits gave no signs which raised concern for convergence. And most importantly,
+        the forecasts from the component models generally outperformed those of the 
+        joint model in terms of minimizing proper scoring rules.}
+        \\~\\
+  \end{quote}
+
   \item I recommend including the statement raised by the reviewer in the first 
@@
           following.}
-
+     \emph{     
     \begin{quote}
@@
     \end{quote}
+    }
 
     \begin{quote}
+    \emph{
       During the FluSight competition, most participating teams would occasionally 
@@
       FluSight targets during the 2023 season. 
-
+      }
+      \emph{
       Because forecast submissions for some targets were occassionally missed
@@
       targets for which forecasts were included or exlcuded. This makes a
-      simple average of scores is an imperfect measure for comparison.
+      simple average of scores is an imperfect measure for comparison.}
     \end{quote}
@@
 \\~\\
-Forecasting Influenza Hospitaliztions Using a Bayesian Hierarchical Nonlinear 
-Model with Discrepancy
+% Forecasting Influenza Hospitaliztions Using a Bayesian Hierarchical Nonlinear 
+% Model with Discrepancy
+% 
+% 
+% This manuscript seems to be a very reasonable attempt at forecasting influenza 
+% induced hospitalizations. The
+% manuscript also puts the methodology in the context of the FluSight competition 
+% organised by the CDC in the US.
+% Apologies, I have found it hard to review as it is conceptually very hard to do 
+% so in my opinion. Firstly,
+% I do not find comfortable modelling the proportion, rather a natural outcome to 
+% model will be the number of hospitalization using a Generalized Linear model 
+% (Poisson distribution). Secondly, I find it un-intuitive to model the
+% flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model 
+% we focus on two aspects: the systematic part and the random part. Here, as far 
+% as  I understand the systematic part is the ASG function. The random
+% part has several components, which are hidden in Equation (5). My difficulty in 
+% modelling using the ASG
+% function lies in my inability to relate any meaningful meteorological variables 
+% (which are very much drivers of the seasonal flu outbreaks). See for example 
+% the paper,  ``A hierarchical Bayesian model for improving short-term forecasting 
+% of hospital demand by including meteorological information`` by  SK Sahu, 
+% B Baffour, PR Harper, JH Minty, C Sarran
+% Journal of the Royal Statistical Society, Series A (2014).
 
-This manuscript seems to be a very reasonable attempt at forecasting influenza 
-induced hospitalizations. The
-manuscript also puts the methodology in the context of the FluSight competition 
-organised by the CDC in the US.
-Apologies, I have found it hard to review as it is conceptually very hard to do 
-so in my opinion. Firstly,
-I do not find comfortable modelling the proportion, rather a natural outcome to 
-model will be the number of hospitalization using a Generalized Linear model 
-(Poisson distribution). Secondly, I find it un-intuitive to model the
-flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model 
-we focus on two aspects: the systematic part and the random part. Here, as far 
-as  I understand the systematic part is the ASG function. The random
-part has several components, which are hidden in Equation (5). My difficulty in 
-modelling using the ASG
-function lies in my inability to relate any meaningful meteorological variables 
-(which are very much drivers of the seasonal flu outbreaks). See for example 
-the paper,  "A hierarchical Bayesian model for improving short-term forecasting 
-of hospital demand by including meteorological information" by  SK Sahu, 
-B Baffour, PR Harper, JH Minty, C Sarran
-Journal of the Royal Statistical Society, Series A (2014).
-
-In spite of the above criticisms, I appreciate the hierarchical modeling 
+% In spite of the above criticisms, 
+I appreciate the hierarchical modeling 
 approach of fitting the two models separately in components. But I wonder how 
@@
 suggest one way to propagate this uncertainty is by using MCMC samples 
-(and methods), see for example, the paper,  "A rigorous statistical framework 
+(and methods), see for example, the paper,  ``A rigorous statistical framework 
 for spatio-temporal pollution prediction and estimation of its long-term impact 
-on health" by 
+on health`` by 
 D Lee, S Mukhopadhyay, A Rushworth, and SK Sahu, in Biostatistics (2017).
+\\~\\
+\emph{We appreciated this comment and the included reference. We used the reference
+to assist in our new explanation, and made sure to cite it. 
+To clarify how uncertainty was propogated, we consildated sections
+  3.7 and 3.8, reordered the main ideas, and removed redundant information. We
+  also added some advantages/disadvantages of our approach. The changes now
+  read:}
+  \begin{quote}
+        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
+        \label{sec:implementation_posterior}
+        \emph{
+        Rather than combine the ILI and hopsitalization models into a single model,
+        (1) and 
+        (6) were fit separately via Bayesian posterior updating 
+        in two stages and then fused together to produce hospitalization 
+        forecasts. In the first stage, future ILI 
+        forecasts were obtained from the posterior predictive distribution where for
+        week $w$, the predictive distribution was obtained by integrating over the 
+        parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
+        as in (\ref{eq:ili_post}) where 
+        $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
+        is the density function of the posterior distribution for the model parameters
+        and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$. 
+        The desired forecasts were for weeks $w + i$ 
+        where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
+        predicted or forecasted ILI at $i$ weeks in the future.}
+
+        \begin{equation}
+            \label{eq:ili_post}
+            \tag{7}
+            p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
+            p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
+            \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
+            \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
+            d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+        \end{equation}
+        % \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
+        % \spencer{Isn't that captured in $d \pi_s$?}
+
+        \emph{\indent In the second stage, the posterior distribution
+        for parameters of model (6) given observed
+        hospitalization data $\textbf{H}_w$, observed ILI data, and
+        denoted as 
+        $p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
+        was obtained.
+        The posterior predictive in (\ref{eq:ili_post}) was then combined with 
+        the hospitalization parameter posterior distribution
+        to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
+        for week $w + i$.
+        Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
+        parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
+        That is, the first stage component model produced $K$ draws
+        $(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
+        and the second stage component model produced
+        $K$ parameter draws 
+        $(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
+        $(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
+        ILI and parameter posterior distribution draws was then plugged into equation
+        (6) to produce}
+
+        \[
+         \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
+                                          \alpha_{1s}^{(k)} 
+                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
+                                          \alpha_{2s}^{(k)} 
+                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
+                                          \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
+                                          \epsilon_{s,w + i}^{(k)}
+        \]
+        \emph{where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
+        $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
+        from which
+        the probabilistic hospitalization forecast was estimated. In practice we set
+        $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
+        \\~\\
+        \indent Occasionally $\{H^{(k)}_{s,w + i}\}^K$ included a small number of negative 
+        values,
+        which do not make sense when the distribution is meant to forecast
+        hospitalizations, a nonnegative number. We attempted to alleviate this problem
+        by modeling log-hospitalizations or by assuming hospitalizations followed a
+        distribution truncated at 0, but these models led to poorer forecasts
+        and issues drawing from the posterior distribution.
+        For the forecast competition analysis in 
+        section 5, all negative values of $\{H^{(k)}_{s,w + i}\}^K$ 
+        were 
+        set to 0 to reflect realistic values of hospitalizations and comply with the 
+        FluSight forecasting rules.}
+        \\~\\
+        \emph{
+        \indent In the two-stage scheme, the posterior distribution of 
+        model (1) parameters was not directly influenced by the 
+        hospitalization data and the posterior distribution of model 
+        (6) parameters was not directly influenced by model 
+        (1) parameters. Hence there was limited feedback between the
+        two components, but propagation of uncertainty in ILI forecasts was still 
+        permitted in the estimating of hospitalization forecasts.
+        This two-stage modeling is remeniscent of the Bayesian cut 
+        described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
+        into two component models or ``modules`` may be justified for several reasons
+        including to reduce time to fit the model, avoiding mixing issues of posterior
+        sampling distributions, improving predictive performance or to prevent
+        model misspecification in one model to influence the other 
+        \cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, 
+        lee2017rigorous, plummer2015cuts}. A drawback
+        from cutting a model is potential loss of useful information. 
+        In our current 
+        application for example, the ILI model parameters receive no potentially 
+        useful information from hospiatlization data.
+        \\~\\
+        \indent The decision to cut the forecast model into two component models
+        was made for three major reasons.
+        When fitting the fully Bayesian joint model, the time to fit a single model 
+        using
+        MCMC sampling and
+        obtain forecasts was sometimes many hours whereas fitting the two models
+        separately took a fraction of the time. When assessing convergence of model
+        parameters, it was rare that posterior sampling chains for the joint model
+        showed convergence even after 55,000 iterations, whereas the component model 
+        fits gave no signs which raised concern for convergence. And most importantly,
+        the forecasts from the component models generally outperformed those of the 
+        joint model in terms of minimizing proper scoring rules.}
+        \\~\\
+  \end{quote}
 
@@
 of the book
-"Bayesian modeling of spatio-temporal data with R" by Sahu 
+``Bayesian modeling of spatio-temporal data with R`` by Sahu 
 (Chapman and Hall, 2022). 
@@
 \begin{quote}
+\emph{
 Model comparison was done by calculating a proper scoring rule for each 
@@
 The WIS is 
-defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
+defined in (8) where $Q$ is a forecast represented by all 
 included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
@@
 $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
-single interval, defined in (\ref{eq:is}). Here 
+single interval, defined in (9). Here 
 $\vmathbb{1}\{\cdot\}$ is the indicator function.
@@
 the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
-\cite[]{bracher2021evaluating}.}
+\cite[]{bracher2021evaluating}.}}
 \end{quote}

### 2025-12-01 manuscript_res.tex
fix up arxiv manuscript
Fix up reviewer comments

@@
 \subsection{Hospitalization Forecasts Using the Bayesian Cut}
-
-Rather than combining the ILI model in equation (\ref{eq:ili_model}) with the 
-hospitalization model in equation (\ref{eq:final_hosp}) and fitting a single
-posterior distribution, both models were fit separately and 1 to 4 week ahead
-ILI forecasts from model (\ref{eq:ili_model}) were combined with the posterior
-predictive distribution of model (\ref{eq:final_hosp}) to obtain 1 to 4 week
-ahead hospitalization forecasts. In this scheme, the posterior distribution of 
+\label{sec:implementation_posterior}
+
+Rather than combine the ILI and hopsitalization models into a single model,
+(\ref{eq:ili_model}) and 
+(\ref{eq:final_hosp}) were fit separately via Bayesian posterior updating 
+in two stages and then fused together to produce hospitalization 
+forecasts. In the first stage, future ILI 
+forecasts were obtained from the posterior predictive distribution where for
+week $w$, the predictive distribution was obtained by integrating over the 
+parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
+as in (\ref{eq:ili_post}) where 
+$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
+is the density function of the posterior distribution for the model parameters
+and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$. 
+The desired forecasts were for weeks $w + i$ 
+where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
+predicted or forecasted ILI at $i$ weeks in the future.
+
+\begin{equation}
+    \label{eq:ili_post}
+    p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
+    p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
+    \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
+    \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
+    d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+\end{equation}
+% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
+% \spencer{Isn't that captured in $d \pi_s$?}
+
+In the second stage, the posterior distribution
+for parameters of model (\ref{eq:final_hosp}) given observed
+hospitalization data $\textbf{H}_w$, observed ILI data, and
+denoted as 
+$p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
+was obtained.
+The posterior predictive in (\ref{eq:ili_post}) was then combined with 
+the hospitalization parameter posterior distribution
+to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
+for week $w + i$.
+Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
+parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
+That is, the first stage component model produced $K$ draws
+$(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
+and the second stage component model produced
+$K$ parameter draws 
+$(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
+$(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
+ILI and parameter posterior distribution draws was then plugged into equation
+(\ref{eq:final_hosp}) to produce
+
+\[
+ \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
+                                  \alpha_{1s}^{(k)} 
+                                  (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
+                                  \alpha_{2s}^{(k)} 
+                                  (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
+                                  \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
+                                  \epsilon_{s,w + i}^{(k)}
+\]
+where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
+$\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
+from which
+the probabilistic hospitalization forecast was estimated. In practice we set
+$K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
+
+Occasionally $\{H^{(k)}_{s,w + i}\}^K$ included a small number of negative 
+values,
+which do not make sense when the distribution is meant to forecast
+hospitalizations, a nonnegative number. We attempted to alleviate this problem
+by modeling log-hospitalizations or by assuming hospitalizations followed a
+distribution truncated at 0, but these models led to poorer forecasts
+and issues drawing from the posterior distribution.
+For the forecast competition analysis in 
+section \ref{sec:analysis}, all negative values of $\{H^{(k)}_{s,w + i}\}^K$ 
+were 
+set to 0 to reflect realistic values of hospitalizations and comply with the 
+FluSight forecasting rules.
+
+In the two-stage scheme, the posterior distribution of 
 model (\ref{eq:ili_model}) parameters was not directly influenced by the 
@@
 permitted in the estimating of hospitalization forecasts.
-
-This two-component modeling is remeniscent of the Bayesian cut 
+This two-stage modeling is remeniscent of the Bayesian cut 
 described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
-into two component models or ``modules`` may be justfied for several reasons
+into two component models or ``modules`` may be justified for several reasons
 including to reduce time to fit the model, avoiding mixing issues of posterior
@@
 model misspecification in one model to influence the other 
-\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, plummer2015cuts}.
+\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, 
+lee2017rigorous, plummer2015cuts}. A drawback
+from cutting a model is potential loss of useful information. 
+In our current 
+application for example, the ILI model parameters receive no potentially 
+useful information from hospiatlization data.
 
 The decision to cut the forecast model into two component models
-was made for several reasons.
-When fitting the fully Bayesian joint model, the time to fit a single model using
-Markov chain Monte Carlo (MCMC) sampling and
+was made for three major reasons.
+When fitting the fully Bayesian joint model, the time to fit a single model 
+using
+MCMC sampling and
 obtain forecasts was sometimes many hours whereas fitting the two models
@@
 
-Model (\ref{eq:ili_model}) was fit via Bayesian posterior updating. Future ILI 
-forecasts were obtained via the posterior predictive distribution where for
-week $w$, the predictive distribution was obtained by integrating over the 
-parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
-as in (\ref{eq:ili_post}) where 
-$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ 
-is the density function of the posterior distribution for the model parameters. 
-If the current week was $w$ then the desired forecasts were for weeks $w + i$ 
-where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
-predicted or forecasted ILI at $i$ weeks in the future.
-
-\begin{equation}
-    \label{eq:ili_post}
-    p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}) = \int \int \int \int 
-    p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
-    \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
-    \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) 
-    d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
-\end{equation}
-% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
-% \spencer{Isn't that captured in $d \pi_s$?}
-Hospitalization forecasts $\widetilde{H}_{s,w + i}$ at week $w + i$ were then
-obtained by combining the posterior predictive distribution of model 
-(\ref{eq:final_hosp}) with (\ref{eq:ili_post}) and integrating over 
-$\widetilde{ILI}_{s,w + i}$ as in (\ref{eq:hosp_post_forc}). Here,
-$\boldsymbol{\alpha}$ represents all regression and variance parameters in
-model (\ref{eq:final_hosp}).
-
-\begin{equation}
-\begin{aligned}
-\label{eq:hosp_post_forc}
-    p(\tilde{H}_{s,w + i} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) &= 
-       \int p(\tilde{H}_{s,w + i} | \textbf{H}, \textbf{ILI}, 
-       \widetilde{ILI}_{s,w + i}, \boldsymbol{\alpha}) 
-       p(\widetilde{ILI}_{s,w + i} | \textbf{ILI}) d\widetilde{ILI}_{s,w + i}
-\end{aligned}
-\end{equation}
-
-\subsection{Parameter estimation and posterior predictive sampling}
-\label{sec:implementation_posterior}
-
-The models were fit via MCMC sampling using the 
+Both component models were fit via MCMC sampling using the 
 \texttt{cmdstanr} package which was developed and is maintained by the Stan 
@@
 hospitalization models are included in the supplementary material 
-\cite[]{wadsworth2024bas}.
+\cite[]{wadsworth2025bas}.
 
@@
 recommend threshold of 1.05, and the smallest $ESS$ was over 500.
-% Tables \ref{tab:ili_diagnostics} and 
-% \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum 
-% $ESS$ over all parameters for ILI and hospitalization models respectively. 
-% Forecast models for all other weeks of the season were fit using one chain of 
-% 60,000 draws where the first 10,000 draws were discarded as a burn-in. 
-% % \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
-% For parameters of the ASG models that were prone to cause trouble in posterior 
-% sampling, the starting values were set to be the MLEs.
-
-% \begin{table}
-% \caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI 
-% models fitted on US data for week 14 of the 2023 season}
-% \label{tab:ili_diagnostics}
-% \begin{center}
-%     \begin{tabular}{c|cccc}
-%      & ASG & ASGD & SIR & SIRD \\
-%      \hline
-%      $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
-%      $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
-%     \end{tabular}
-%     \end{center}
-% \end{table}
-
-% \begin{table}
-% \caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six 
-% hospitalization models fitted on US data for week 14 of the 2023 season}
-% \label{tab:hsop_diagnostics}
-% \begin{center}
-%     \begin{tabular}{c|cccccc}
-%      & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
-%      \hline
-%      $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & 
-%      $< 1.001$\\
-%      $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
-%     \end{tabular}
-%     \end{center}
-% \end{table}
-
-To obtain forecast distributions of hospitalizations, draws from the posterior 
-predictive distribution from the ILI model were used in conjunction with the 
-posterior distribution of the hospitalizations model. When fitting model 
-(\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were 
-saved. Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal 
-distributions for the model parameters were saved. To obtain forecast 
-distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following 
-steps were repeated $K$ times where $K$ is an integer for the number of desired 
-samples. We set $K = 50,000$.
-
-\begin{quote}
-\begin{enumerate}[Step 1:]
-  \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
-  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, 
-  $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior 
-  distributions
-  \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
-  \begin{itemize}
-    \item[] $\mu_{s,w + i}^* = \alpha_{0s}^* + \alpha_{1s}^* (ILI_{s,w + i}^* \times P) + \alpha_{2s}^* (ILI_{s,w + i}^* \times P)^2 + \phi^* H^*_{s,w + i - 1}$
-  \end{itemize}
-   \item Repeat step 3 for $i \in \{1,2,3,4\}$ to obtain $H^*_{s,(w + 1):(w + 4)}$
-  \item Repeat steps 1-4 $K$ times
-\end{enumerate}
-\end{quote}
-The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
-for hospitalizations at week $w + i$. 
-Occasionally $\{H^*_{s,w + i}\}^K$ included a small number of negative values,
-which do not make sense when the distribution is meant to forecast
-hospitalizations, a nonnegative number. We attempted to alleviate this problem
-by modeling log-hospitalizations or by assuming hospitalizations followed a
-distribution truncated at 0, but these models led to poorer forecasts
-and issues drawing from the posterior distribution.
-For the forecast competition analysis in 
-section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
-set to 0 to reflect realistic values of hospitalizations and comply with the 
-FluSight forecasting rules.
+
+% \subsection{Parameter estimation and posterior predictive sampling}
+% \label{sec:implementation_posterior}
+% 
+% The models were fit via MCMC sampling using the 
+% \texttt{cmdstanr} package which was developed and is maintained by the Stan 
+% Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
+% implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn 
+% sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
+% several diagnostic statistics for assessing the sampler.
+% Plots of prior and posterior distributions for select parameters from ILI and 
+% hospitalization models are included in the supplementary material 
+% \cite[]{wadsworth2024bas}.
+% 
+% Model fit was assessed for four ILI models and for 
+% the hospitalization model. The ILI models included the SIR and 
+% ASG models and models with and without discrepancy modeling. When discrepancy 
+% is included, the models are denoted as SIRD and ASGD. These models were fit 
+% using US national data from 2010 to 2023 flu seasons, where data from the 
+% 2020 season was excluded because of the unique behavior during that season. 
+% % Assessment for hospitalization modeling was done for six different models. 
+% % These include NORM, LNORM, and LST models, and models where a quadratic ILI 
+% % term is included or excluded. 
+% Sampling was done with four chains where from each chain 
+% 60,000 posterior draws were sampled, and the first 10,000 draws were discarded 
+% as a burn-in. The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the 
+% effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for 
+% each parameter, and for all models fit, the largest $\hat{R}$ was below the 
+% recommend threshold of 1.05, and the smallest $ESS$ was over 500.
+% % Tables \ref{tab:ili_diagnostics} and 
+% % \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum 
+% % $ESS$ over all parameters for ILI and hospitalization models respectively. 
+% % Forecast models for all other weeks of the season were fit using one chain of 
+% % 60,000 draws where the first 10,000 draws were discarded as a burn-in. 
+% % % \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
+% % For parameters of the ASG models that were prone to cause trouble in posterior 
+% % sampling, the starting values were set to be the MLEs. 
+% 
+% % \begin{table}
+% % \caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI 
+% % models fitted on US data for week 14 of the 2023 season}
+% % \label{tab:ili_diagnostics}
+% % \begin{center}
+% %     \begin{tabular}{c|cccc}
+% %      & ASG & ASGD & SIR & SIRD \\
+% %      \hline
+% %      $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
+% %      $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
+% %     \end{tabular}
+% %     \end{center}
+% % \end{table}
+% 
+% 
+% % \begin{table}
+% % \caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six 
+% % hospitalization models fitted on US data for week 14 of the 2023 season}
+% % \label{tab:hsop_diagnostics}
+% % \begin{center}
+% %     \begin{tabular}{c|cccccc}
+% %      & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
+% %      \hline
+% %      $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & 
+% %      $< 1.001$\\
+% %      $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
+% %     \end{tabular}
+% %     \end{center}
+% % \end{table}
+% 
+% 
+% 
+% 
+% 
+% 
+% To obtain forecast distributions of hospitalizations, draws from the posterior 
+% predictive distribution from the ILI model were used in conjunction with the 
+% posterior distribution of the hospitalizations model. When fitting model 
+% (\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were 
+% saved. Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal 
+% distributions for the model parameters were saved. To obtain forecast 
+% distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following 
+% steps were repeated $K$ times where $K$ is an integer for the number of desired 
+% samples. We set $K = 50,000$.
+% 
+% \begin{quote}
+% \begin{enumerate}[Step 1:]
+%   \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
+%   \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, 
+%   $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior 
+%   distributions
+%   \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
+%   \begin{itemize}
+%     \item[] $\mu_{s,w + i}^* = \alpha_{0s}^* + \alpha_{1s}^* (ILI_{s,w + i}^* \times P) + \alpha_{2s}^* (ILI_{s,w + i}^* \times P)^2 + \phi^* H^*_{s,w + i - 1}$
+%   \end{itemize}
+%    \item Repeat step 3 for $i \in \{1,2,3,4\}$ to obtain $H^*_{s,(w + 1):(w + 4)}$
+%   \item Repeat steps 1-4 $K$ times
+% \end{enumerate}
+% \end{quote}
+% The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
+% for hospitalizations at week $w + i$. 
+% Occasionally $\{H^*_{s,w + i}\}^K$ included a small number of negative values,
+% which do not make sense when the distribution is meant to forecast
+% hospitalizations, a nonnegative number. We attempted to alleviate this problem
+% by modeling log-hospitalizations or by assuming hospitalizations followed a
+% distribution truncated at 0, but these models led to poorer forecasts
+% and issues drawing from the posterior distribution.
+% For the forecast competition analysis in 
+% section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
+% set to 0 to reflect realistic values of hospitalizations and comply with the 
+% FluSight forecasting rules.
 

### 2025-12-01 old stuff/manuscript.tex
move old files (probably to be deleted later)

--- 2024-12-18 nonlinear_forecast/manuscript.tex
@@
 
-To better inform public decision making regarding the flu epidemic, in 2013 the United States Centers for Disease Control and Prevention (CDC) organized a national flu forecasting competition, also known as FluSight \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}.
-Originally, over a dozen teams of researchers from academic and industry backgrounds participated in FluSight by contributing their own forecast models.
-Besides the 2020 season --or the flu season spanning the fall of 2020 and the winter of 2021-- FluSight has been operated annually and researchers outside the CDC have been invited to participate.
-Initially the target data for forecasts was influenza-like illness (ILI) data.
-ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}.
-The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, Osthus et al.
-\cite[]{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework.
-An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
+To better inform public decision making regarding the flu epidemic, in 2013 the 
+United States Centers for Disease Control and Prevention (CDC) organized a 
+national flu forecasting competition, also known as FluSight 
+\cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. 
+Originally, over a dozen teams of researchers from academic and industry 
+backgrounds participated in FluSight by contributing their own forecast models. 
+Besides the 2020 season --or the flu season spanning the fall of 2020 and the 
+winter of 2021-- FluSight has been operated annually and researchers outside 
+the CDC have been invited to participate. Initially the target data for 
+forecasts was influenza-like illness (ILI) data. ILI is the proportion of 
+patients who meet a healthcare provider and who display flu like symptoms, 
+and ILI data has been available at the state and national level since the 2010 
+flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}. 
+The collaborative ILI forecasting effort has led to a number of modeling 
+developments in flu forecasting 
+\cite[see references therein for more examples]{mcandrew2021adaptively, 
+osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's 
+introduction, Osthus et al.  \cite[]{osthus2019dynamic} categorized the most 
+commonly used flu forecasting models into four classes including mechanistic 
+models based on differential equation compartmental models, agent based models 
+based on population simulation, machine learning/regression models including 
+data driven machine learning and statistical models, and data assimilation 
+models which are constructed by assimilating mechanistic models into a 
+probabilistic framework. An additional forecast model used in FluSight 
+involves the combination of several forecasts into a single ensemble forecast, 
+which has been shown to perform well relative to individual models 
+\cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.
 
-The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications.
-As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}.
-The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded.
-The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}.
-Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition.
-Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}.
-This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.
+The administration of FluSight saw few changes during the first seven seasons, 
+but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 
+forecasting led to major modifications. As a result of the COVID-19 pandemic 
+which began during the 2019 flu season, the typical flu outbreak behavior was 
+altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}. 
+The COVID-19 pandemic led to the creation of the Health and Human Services 
+(HHS) Patient Impact and Hospital Capacity Data System 
+\cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization 
+data, and the COVID-19 Forecast Hub was founded. The COVID-19 Forecast Hub 
+was based on FluSight but with certain major adjustments including how the 
+forecast uncertainty is represented and the addition of the weekly publication 
+of a multi-model ensemble forecast as the official forecast of the CDC 
+\cite[]{bracher2021evaluating, Cramer2022-hub-dataset}. Using estimated 
+quantiles for representing forecast uncertainty and creating a multi-model 
+ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by 
+the flu forecast competition. Additionally the target of the flu forecasts 
+changed from being ILI data to being HHS hospitalization data, which reports 
+the number of hospitalizations due to a laboratory confirmed flu infection  
+\cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of 
+having COVID-19 cases in the population making ILI data, already only a proxy 
+for flu behavior, more difficult to interpret.
 
-The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models.
-The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
-Herein we use ILI models similar to those in Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
-The model of Osthus et al.
-\cite[]{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework.
-Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by Kennedy and O'Hagan \cite[]{kennedy2001bayesian}.
-The model in Ulloa \cite[]{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data.
-Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.
+The contribution of this manuscript is to introduce a two component framework 
+for modeling HHS hospitalization forecasts where hospitalization data and 
+years of ILI data are used to inform forecast models. The first modeling 
+component is a model of ILI data and the second is a model of hospitalization 
+data with ILI as a predictive covariate.
+Herein we use ILI models similar to those in Osthus et al. 
+\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
+The model of Osthus et al. \cite[]{osthus2019dynamic} is a combined data 
+assimilation and statistical regression model which involves a compartmental 
+model in a probabilistic framework. Their model also includes an additional 
+component for capturing a systematic discrepancy between the deterministic 
+part of the model and the actual data, an idea which was first introduced by 
+Kennedy and O'Hagan \cite[]{kennedy2001bayesian}. The model in Ulloa 
+\cite[]{ulloa2019} is a Bayesian hierarchical regression model with an 
+underlying function intended to capture the trajectory of the seasonal 
+ILI data. Herein, we provide a framework under which discrepancy modeling may 
+be used along with a general function modeling ILI data, and we show the 
+effectiveness of including discrepancy modeling during certain periods of 
+the flu season.
 
-In line with the newer FluSight standard of forecasting hospitalizations, we model hospitalizations as a linear function of ILI.
-Thus forecasts produced herein target flu hospitalizations and are a mapping of ILI forecasts to hospitalizations.
-This allows for ILI data from many seasons to be exploited and for ILI forecasts to assist in forecasting hospitalizations, which has fewer seasons of data than ILI.
-Several modeling schemes and their forecasts for the 2023 flu hospitalization season are compared, and it is shown that the modeling decisions produce good forecast results for different states or times during the flu season.
+In line with the newer FluSight standard of forecasting hospitalizations, we 
+model hospitalizations as a linear function of ILI. Thus forecasts produced 
+herein target flu hospitalizations and are a mapping of ILI forecasts to 
+hospitalizations. This allows for ILI data from many seasons to be exploited 
+and for ILI forecasts to assist in forecasting hospitalizations, which has 
+fewer seasons of data than ILI. Several modeling schemes and their forecasts 
+for the 2023 flu hospitalization season are compared, and it is shown that the 
+modeling decisions produce good forecast results for different states or times 
+during the flu season.
 
-In section \ref{sec:data} we review the ILI and hospitalization data provided by the CDC and targeted by FluSight. 
-In section \ref{sec:functions} the modeling framework contributed by this manuscript is given.
-In the same section, functions similar to those used by Osthus et al.
-\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} are defined.
- These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively.
-Model fitting and implementation are described at the end of the section.
-Section \ref{sec:simulation2} includes a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared.
-Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts.
-Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}.
- Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.
+In section \ref{sec:data} we review the ILI and hospitalization data provided 
+by the CDC and targeted by FluSight. 
+In section \ref{sec:functions} the modeling framework contributed by this 
+manuscript is given. In the same section, functions similar to those used by 
+Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} are 
+defined.
+ These functions are the susceptible-infectious-recovered (SIR) compartmental 
+ model and the asymmetric Gaussian (ASG) function  respectively. Model fitting 
+ and implementation are described at the end of the section. Section 
+ \ref{sec:simulation2} includes a simulation study where four ILI forecast 
+ models and their use in forecasting hospitalizations are compared. Commonly 
+ used proper scoring rules \cite[]{gneiting2007strictly}, which are also 
+ introduced and defined in section \ref{sec:simulation2}, are used for 
+ comparing the forecasts. Forecasting of the 2023 flu outbreak along with 
+ assessment and comparison under several selected models is performed in 
+ section \ref{sec:analysis}. 
+ Finally, the manuscript is concluded in section \ref{sec:conclusion} with 
+ general observations and some discussion.
 
  \section{Flu outbreak data} \label{sec:data}
-In this section we introduce and define ILI and hospitalization data and visually evaluate the data.
-ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season.
-Both of these data were collected at the state, territorial, and national level and were reported at least weekly.
-Overall the data is reported for 53 locations including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), and at the US national level.
-We will refer to forecast targets throughout this manuscript.
-A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a specific location and week during the season.
+In this section we introduce and define ILI and hospitalization data and 
+visually evaluate the data. ILI and hospitalization data have been the object 
+of forecasting for FluSight with ILI being the target for the first seven 
+seasons and hospitalizations being the target since the 2022 season. Both of 
+these data were collected at the state, territorial, and national level and 
+were reported at least weekly. Overall the data is reported for 53 locations 
+including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), 
+and at the US national level. We will refer to forecast targets throughout this 
+manuscript. A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a 
+specific location and week during the season.
 
 \subsection{Influenza-like illness data}
-The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers.
-Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases.
-An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}.
-Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza.
-To know whether or not a sick patient is infected with influenza requires a laboratory test.
+The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects 
+information on respiratory illness from outpatient visits to health care 
+providers. Over 3,400 outpatient health care providers in all 50 US states, PR, 
+DC, and the US Virgin Islands report each week the total number of outpatient 
+visits along with the number of ILI cases. An ILI case is defined as a 
+``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
+and/or a sore throat." Prior to the 2021 season, the definition included 
+``without a known cause other than influenza" \cite[]{cdcfluview2023}. Because 
+other illnesses such as COVID-19, RSV, and the common cold may induce similar 
+respiratory symptoms, ILI may include patients infected with some disease 
+other than influenza. To know whether or not a sick patient is infected with 
+influenza requires a laboratory test.
 
 In 2013, when FluSight began, the ILI data was the object of the forecasts. 
-The data was released publicly at HHS region levels, and forecast teams were asked to provide forecasts of several ILI targets on the regional levels including season onset, 1-4 week ahead ILI levels, and the week of peak ILI activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}.
-Currently, the ILI data is collected by the CDC and published on an online portal for viewing at the national, HHS region, census, and state levels \cite[]{cdc2024fluviewportal}.
-To obtain ILI data, we used the \texttt{R} package \texttt{cdcfluview} which provides functions for downloading the data \cite[]{rudis2021cdcfluview}.
-Weekly ILI data from the national, HHS region, and census levels are available from the 1997 flu season until the current season.
-At the state level, data is available from the 2010 flu season to the current season.
+The data was released publicly at HHS region levels, and forecast teams were 
+asked to provide forecasts of several ILI targets on the regional levels 
+including season onset, 1-4 week ahead ILI levels, and the week of peak ILI 
+activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}. Currently, 
+the ILI data is collected by the CDC and published on an online portal for 
+viewing at the national, HHS region, census, and state levels 
+\cite[]{cdc2024fluviewportal}. To obtain ILI data, we used the \texttt{R} 
+package \texttt{cdcfluview} which provides functions for downloading the data 
+\cite[]{rudis2021cdcfluview}. Weekly ILI data from the national, HHS region, 
+and census levels are available from the 1997 flu season until the current 
+season. At the state level, data is available from the 2010 flu season to the 
+current season.
 
-The top of figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023.
-For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons.
-To better align with the flu behavior, week 1 is set as the first week of August and week 52 or 53 is the last week in July of the following year.
-For example week 1 of the 2013 season corresponds to the first week of August 2013, and week 52 of the same season corresponds to the last week of July 2014.
-This convention is used for the remainder of this manuscript.
+The top of figure \ref{fig:us_ili} shows the ILI data at the national level for 
+flu seasons 2010 to 2023. For most seasons there are 52 weeks listed, but for 
+the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 
+Sundays during those seasons. 
+To better align with the flu behavior, week 1 is set as the first week of 
+August and week 52 or 53 is the last week in July of the following year. For 
+example week 1 of the 2013 season corresponds to the first week of August 2013, 
+and week 52 of the same season corresponds to the last week of July 2014. This 
+convention is used for the remainder of this manuscript.
 
@@
 
-Hospital admission data, used as the object of FluSight forecasting for the 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported Patient Impact and Hospital Capacity by State Timeseries}.
-Several targets of respiratory illnesses including COVID-19, RSV, and influenza are reported weekly by most hospitals in the US.
-In February 2022 it became mandatory for all hospitals to report the number of COVID-19 and influenza hospitalizations, and since then reporting of hospitalizations has become widespread.
-These data were updated every Wednesday and Friday according to NHSN guidelines \cite[]{healthdata2024covidts}.
+Hospital admission data, used as the object of FluSight forecasting for the 
+2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
+Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
+Patient Impact and Hospital Capacity by State Timeseries}. Several targets of 
+respiratory illnesses including COVID-19, RSV, and influenza are reported 
+weekly by most hospitals in the US. In February 2022 it became mandatory for 
+all hospitals to report the number of COVID-19 and influenza hospitalizations, 
+and since then reporting of hospitalizations has become widespread. These data 
+were updated every Wednesday and Friday according to NHSN guidelines 
+\cite[]{healthdata2024covidts}.
 
-The bottom of figure \ref{fig:us_ili} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons.
-These plots show similarities to the ILI plots in the top of the figure in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends.
-For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22.
-Additional plots of ILI data and hospitalization
+The bottom of figure \ref{fig:us_ili} shows the weekly national 
+hospitalizations for the 2022 and 2023 flu seasons. These plots show 
+similarities to the ILI plots in the top of the figure in that at the early 
+weeks of the season hospitalizations are low, but they increase in the fall 
+to a peak after which they decrease until the flu outbreak ends. For both the 
+2022 and 2023 seasons, the hospitalizations peaked during the same week as 
+ILI, and in 2023 that peak occurred during the holiday week 22. Additional 
+plots of ILI data and hospitalization
 data for select states are included in the supplementary material 
-\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar to 
+\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar 
+to 
 those in figure \ref{fig:us_ili} with activity increasing
@@
 
-The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows a similar trajectory.
-Compartmental models are standard mathematical models used for modelling disease outbreaks.
-One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}.
-Ulloa \cite[]{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data.
-The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy.
-In the first part of this section, we present an ILI model similar to the model in Osthus et al.
-\cite[]{osthus2019dynamic}.
-With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.
+The typical behavior of the ILI data which starts at lower values in the late 
+summer and fall but which increases to a peak, usually in December or January, 
+followed by a decline motivates the use of a nonlinear function which follows 
+a similar trajectory. Compartmental models are standard mathematical models 
+used for modelling disease outbreaks. One important compartmental model is the 
+susceptible-infectious-recovered (SIR) model, which is used by some to model 
+ILI data \cite[]{osthus2019dynamic, allen2017primer}. Ulloa \cite[]{ulloa2019} 
+chose to use the asymmetric Gaussian (ASG) function to model ILI data. The 
+SIR and ASG models may both be appropriate to describe ILI behavior over the 
+course of a flu season, however there may also be systematic behavior not 
+captured by either, necessitating an additional model component to capture 
+the discrepancy. In the first part of this section, we present an ILI model 
+similar to the model in Osthus et al. \cite[]{osthus2019dynamic}. With some 
+generalization, the model may incorporate any appropriate nonlinear function, 
+though the focus here is on the SIR and ASG functions which are defined.
 
-With the aim of forecasting hospitalizations, we also introduce a linear model of hospitalizations data with ILI as a predictive covariate.
-To forecast hospitalizations, ILI data is first forecasted and the forecast is then plugged in as a covariate in the hospitalization model, thus producing hospitalization forecasts.
-The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.
+With the aim of forecasting hospitalizations, we also introduce a linear 
+model of hospitalizations data with ILI as a predictive covariate. To forecast 
+hospitalizations, ILI data is first forecasted and the forecast is then 
+plugged in as a covariate in the hospitalization model, thus producing 
+hospitalization forecasts. The hospitalization model is also defined in this 
+section, and the section is concluded with descriptions of selected prior 
+distributions, model implementation, and posterior sampling.
 
@@
 \end{equation}
-The idea for using the reverse random walk is that there are several previous seasons of ILI data, and assuming the random walk captures systematic behavior, fitting it hierarchically over seasons can assist in predicting future behavior in the current season.
-Reverse random walks have also been used with success in election forecasting and other flu forecasting models \cite[]{osthus2021multiscale, osthus2019dynamic, linzer2013dynamic}.
-The sum to zero constraint $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve identifiability.
+The idea for using the reverse random walk is that there are several previous 
+seasons of ILI data, and assuming the random walk captures systematic behavior, 
+fitting it hierarchically over seasons can assist in predicting future behavior 
+in the current season. 
+Reverse random walks have also been used with success in election forecasting 
+and other flu forecasting models \cite[]{osthus2021multiscale, 
+osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint 
+$-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
+identifiability.
 
 \subsection{ILI forecasts}
-Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating.
-Future ILI forecasts are obtained via the posterior predictive distribution where for week $w$, the predictive distribution is obtained by integrating over the parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ as in (\ref{eq:ili_post}) where $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ is the density function of the posterior distribution for the model parameters.
-If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ where $i$ is a positive integer.
+Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating. Future ILI 
+forecasts are obtained via the posterior predictive distribution where for
+week $w$, the predictive distribution is obtained by integrating over the 
+parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
+as in (\ref{eq:ili_post}) where 
+$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ 
+is the density function of the posterior distribution for the model parameters. 
+If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ 
+where $i$ is a positive integer.
 
@@
 
-The second component for forecast modeling is based on the relationship between hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
-This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}. Figure \ref{fig:lag_scatter}
+The second component for forecast modeling is based on the relationship between 
+hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
+This is an example of an autoregressive model with exogenous variables where 
+the autoregressive lag is one (ARX(1)) 
+\cite[]{raftery2010online,ljung1987system}. Figure \ref{fig:lag_scatter}
 shows scatterplots of the difference between hospitalizations and scaled 1 week 
@@
     \includegraphics[scale=.46]{Images/lagdiff_vs_ili.png}
-    \caption{Scatterplots of national hospitalizations (left) or log hospitalizations (right) lag differences by ILI \% where the lag is scaled by $\phi$.
-Points are colored by season.
-For hospitalizations $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$.
-}
+    \caption{Scatterplots of national hospitalizations (left) or log 
+    hospitalizations (right) lag differences by ILI \% where the lag is 
+    scaled by $\phi$. Points are colored by season. For hospitalizations 
+    $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$. }
     \label{fig:lag_scatter}
@@
     \label{eq:final_hosp}
-    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
-    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, \omega_s) %N(0, \sigma_{\epsilon}^2)
+    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} 
+    (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
+    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, 
+    \omega_s) %N(0, \sigma_{\epsilon}^2)
     \end{aligned}
 \end{equation}
-In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
+In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for 
+week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed 
+according to some distribution $D_s$ with mean parameter 0, scale parameter 
+$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees 
+of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.
 
-This model is for any location, and for fitting purposes $ILI_{s,w}$ is always multiplied by $P$ which is proportional to the population of the state or territory, in this case the total population divided by 50,000.
-This is done as a means of scaling so that the prior distribution assigned to  $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, $\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all states.
+This model is for any location, and for fitting purposes $ILI_{s,w}$ is always 
+multiplied by $P$ which is proportional to the population of the state or 
+territory, in this case the total population divided by 50,000. This is done 
+as a means of scaling so that the prior distribution assigned to  
+$\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, 
+$\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all 
+states.
 
-Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also fit via Bayesian posterior updating.
-To obtain forecasts for $H_{s, w^* + i}$, the ILI posterior predictive distribution is used along with the posterior distribution for the parameters in (\ref{eq:final_hosp}).
-We considered three scenarios for model (\ref{eq:final_hosp}).
-A model where $D_s$ belongs to the normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, or $H_{s,w} + 1$ is lognormally distributed (LNORM).
-The population value $P$ is excluded from the LNORM model, and we set $\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting.
-In the LNORM model, if the linear parameters are not set the same for seasons 22 and 23, the final variance was more prone to be extreme.
-Besides varying the distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ or there is no quadratic ILI term.
+Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also 
+fit via Bayesian posterior updating. To obtain forecasts for $H_{s, w^* + i}$, 
+the ILI posterior predictive distribution is used along with the posterior 
+distribution for the parameters in (\ref{eq:final_hosp}). We considered three 
+scenarios for model (\ref{eq:final_hosp}). A model where $D_s$ belongs to the 
+normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ 
+is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, 
+or $H_{s,w} + 1$ is lognormally distributed (LNORM). The population value $P$ 
+is excluded from the LNORM model, and we set 
+$\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting. In 
+the LNORM model, if the linear parameters are not set the same for seasons 22 
+and 23, the final variance was more prone to be extreme. Besides varying the 
+distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ 
+or there is no quadratic ILI term.
 
@@
 
-The priors selected for the ILI data model under both the SIR and ASG models largely follow the prior selections in Osthus et al.
-\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved numerical stability and/or we felt the adjusted prior made more sense for the problem.
-For model (\ref{eq:ili_model}), parameters which are common even when using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, $\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$.
-For the SIR function $\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$.
-For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
+The priors selected for the ILI data model under both the SIR and ASG models 
+largely follow the prior selections in Osthus et al. \cite[]{osthus2019dynamic} 
+and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved 
+numerical stability and/or we felt the adjusted prior made more sense for the 
+problem. For model (\ref{eq:ili_model}), parameters which are common even when 
+using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
+$\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$. For the SIR function 
+$\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
+function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
+For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
+estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
 
-The priors assigned were mostly noninformative, though in certain cases the prior distributions were selected for numerical stability as was the case for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$.
-For these two scale parameters only, rather than assigning a half-normal prior to the standard deviation parameters, as recommended by Gelman \cite[]{gelman2006prior}, the priors were assigned to the variance parameters.
-Univariate parameters were assigned either a normal distribution prior if the support is on $\mathbb{R}$, a half-normal prior if the support is nonnegative, or a truncated-normal prior to match a more specific support.
-Under the ASG model, $\theta_s$ is modeled hierarchically over seasons so that for each season the transformed parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are assigned to $\theta$ and $\Sigma$.
+The priors assigned were mostly noninformative, though in certain cases the 
+prior distributions were selected for numerical stability as was the case 
+for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$. For these two scale 
+parameters only, rather than assigning a half-normal prior to the standard 
+deviation parameters, as recommended by Gelman \cite[]{gelman2006prior}, the 
+priors were assigned to the variance parameters. Univariate parameters were 
+assigned either a normal distribution prior if the support is on $\mathbb{R}$, 
+a half-normal prior if the support is nonnegative, or a truncated-normal 
+prior to match a more specific support. Under the ASG model, $\theta_s$ is 
+modeled hierarchically over seasons so that for each season the transformed 
+parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are 
+assigned to $\theta$ and $\Sigma$.
 
 Additional prior constraints were made to improve parameter identifiability. 
-In Osthus et al.
-\cite[]{osthus2019dynamic} the initial value of the susceptible population compartment of the SIR model was set to $S_0 = 0.9$.
-The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative priors.
-To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the parameters.
-A modular Bayesian approach involves multiple steps of parameter fitting where some parameters may be estimated without priors via maximum likelihood estimation or other means.
-Fitting the rest of the model parameters involves assigning priors and conditioning on the previously fit parameters.
-This has been done in computer modeling to improve identifiability and other issues, though Liu et al.
-\cite[]{liu2009modularization} warn this approach is not probabilistically sound if parameter inference is a priority \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}.
-We carried out the modular fit by first estimating the maximum likelihood estimate (MLE) for the parameter $\lambda_s$ for each season in (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
+In Osthus et al. \cite[]{osthus2019dynamic} the initial value of the 
+susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
+The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
+priors. To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
+in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the 
+parameters. A modular Bayesian approach involves multiple steps of parameter 
+fitting where some parameters may be estimated without priors via maximum 
+likelihood estimation or other means. Fitting the rest of the model parameters 
+involves assigning priors and conditioning on the previously fit parameters. 
+This has been done in computer modeling to improve identifiability and other 
+issues, though Liu et al. \cite[]{liu2009modularization} warn this approach is 
+not probabilistically sound if parameter inference is a priority 
+\cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. 
+We carried out the modular fit by first estimating the maximum likelihood 
+estimate (MLE) for the parameter $\lambda_s$ for each season in 
+(\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.
 
@@
 
-The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the Stan Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}.
-Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}.
-The \texttt{cmdstanr} package provides several diagnostic statistics for assessing the sampler.
-As mentioned, most model parameter prior distributions were intended to be uninformative.
-Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material \cite[]{wadsworth2024bas}.
+The models were fit via Markov chain Monte Carlo (MCMC) sampling using the 
+\texttt{cmdstanr} package which was developed and is maintained by the Stan 
+Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
+implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn 
+sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
+several diagnostic statistics for assessing the sampler.
+As mentioned, most model parameter prior distributions were intended to be 
+uninformative.
+Plots of posterior distributions for select parameters from ILI and 
+hospitalization models are shown in the supplementary material 
+\cite[]{wadsworth2024bas}.
 
-We assessed the model fit for four ILI models.
-These included the SIR and ASG models and models with and without discrepancy modeling.
-When discrepancy is included, the models are denoted as SIRD and ASGD.
-These models were fit using US national data from 2010 to 2023 flu seasons, where data from the 2020 season was excluded because of the unique behavior during that season.
+We assessed the model fit for four ILI models. These included the SIR and 
+ASG models and models with and without discrepancy modeling. When discrepancy 
+is included, the models are denoted as SIRD and ASGD. These models were fit 
+using US national data from 2010 to 2023 flu seasons, where data from the 
+2020 season was excluded because of the unique behavior during that season. 
 Assessment for hospitalization modeling was done for six different models. 
-These include NORM, LNORM, and LST models, and models where a quadratic ILI term is included or excluded.
-To assess posterior sampling convergence, models were fit to data where ILI and hospitalization data up to week 14 of the 2023 season was included.
-Sampling was done with four chains where from each chain 60,000 posterior draws were sampled, and the first 10,000 draws were discarded as a burn-in.
-The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for each parameter.
-Tables \ref{tab:ili_diagnostics} and \ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum $ESS$ over all parameters for ILI and hospitalization models respectively.
-Forecast models for all other weeks of the season were fit using one chain of 60,000 draws where the first 10,000 draws were discarded as a burn-in.
+These include NORM, LNORM, and LST models, and models where a quadratic ILI 
+term is included or excluded. To assess posterior sampling convergence, models 
+were fit to data where ILI and hospitalization data up to week 14 of the 2023 
+season was included. Sampling was done with four chains where from each chain 
+60,000 posterior draws were sampled, and the first 10,000 draws were discarded 
+as a burn-in. The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the 
+effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for 
+each parameter. Tables \ref{tab:ili_diagnostics} and 
+\ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum 
+$ESS$ over all parameters for ILI and hospitalization models respectively. 
+Forecast models for all other weeks of the season were fit using one chain of 
+60,000 draws where the first 10,000 draws were discarded as a burn-in. 
 % \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
-For parameters of the ASG models that were prone to cause trouble in posterior sampling, the starting values were set to be the MLEs.
+For parameters of the ASG models that were prone to cause trouble in posterior 
+sampling, the starting values were set to be the MLEs.
 
 \begin{table}
-\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI models fitted on US data for week 14 of the 2023 season}
+\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI 
+models fitted on US data for week 14 of the 2023 season}
 \label{tab:ili_diagnostics}
@@
 \end{quote}
-The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast for hospitalizations at week $w + i$.
-For the forecast competition analysis in section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were set to 0 to reflect realistic values of hospitalizations and comply with the FluSight forecasting rules.
+The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
+for hospitalizations at week $w + i$. For the forecast competition analysis in 
+section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
+set to 0 to reflect realistic values of hospitalizations and comply with the 
+FluSight forecasting rules.
 
 \section{Simulation Study} \label{sec:simulation2}
-In this section, we present a simulation study conducted for comparing ILI models and further assessing the hospitalization forecast model. US ILI data is used, and hospitalization data is simulated. 
-A leave-one-season-out (LOSO) approach was combined with a Monte Carlo simulation approach.
-For each replication, we simulated log-hospitalizations for all weeks during seasons 2010, ..., 2022, excluding 2020, using the existing ILI data as a predictive covariate.
-Each season was in turn ``left-out" and treated as if it was the most recent season which we desired to forecast.
-Fitting and forecasting was then done for weeks 14, 20, 26, 32, and 38 of the left out season, giving two weeks that tend to occur as flu cases increase, two as cases decrease, and one that occurs when the cases may be increasing or decreasing.
-Week 20 is a week leading up to the holiday week 22 where ILI typically has a local peak.
-We were particularly interested in how important modeling discrepancy is for forecasting at week 20.
+In this section, we present a simulation study conducted for comparing ILI 
+models and further assessing the hospitalization forecast model. US ILI data 
+is used, and hospitalization data is simulated. 
+A leave-one-season-out (LOSO) approach was combined with a Monte Carlo 
+simulation approach. For each replication, we simulated log-hospitalizations 
+for all weeks during seasons 2010, ..., 2022, excluding 2020, using the 
+existing ILI data as a predictive covariate. Each season was in turn 
+``left-out" and treated as if it was the most recent season which we desired 
+to forecast. Fitting and forecasting was then done for weeks 14, 20, 26, 32, 
+and 38 of the left out season, giving two weeks that tend to occur as flu 
+cases increase, two as cases decrease, and one that occurs when the cases may 
+be increasing or decreasing. Week 20 is a week leading up to the holiday week 
+22 where ILI typically has a local peak. We were particularly interested in 
+how important modeling discrepancy is for forecasting at week 20.
 
-For the simulation of hospitalizations, the parameters $\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and $\sigma^2_{\epsilon_s}$ from the hospitalization model in (\ref{eq:final_hosp}) were considered the same across all seasons so that all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$.
-The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization data from the 2022 season.
-For fitting, the hospitalization data was first log-transformed.
-We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
-and assigned the noninformative prior $p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$.
-The marginal posterior distribution $\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ was then the established posterior multivariate normal distribution and $\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior distribution \cite[]{gelman2013bayesian}.
-The posterior means of those parameters were used as the values from which log-hospitalizations were simulated.
-The number of replicates in the simulation was 500.
+For the simulation of hospitalizations, the parameters 
+$\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and 
+$\sigma^2_{\epsilon_s}$ from the hospitalization model in 
+(\ref{eq:final_hosp}) were considered the same across all seasons so that 
+all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$. 
+The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were 
+estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization 
+data from the 2022 season. For fitting, the hospitalization data was first 
+log-transformed. We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
+and assigned the noninformative prior 
+$p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$. 
+The marginal posterior distribution 
+$\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ 
+was then the established posterior multivariate normal distribution and 
+$\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior 
+distribution \cite[]{gelman2013bayesian}. The posterior means of those 
+parameters were used as the values from which log-hospitalizations were 
+simulated. The number of replicates in the simulation was 500.
 
-Model comparison was done by calculating the continuous ranked probability score (CRPS) for each forecast.
-The CRPS is a proper scoring rule which evaluates the forecast distribution function.
-Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}.
-Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}.
-The CRPS is defined in (\ref{eq:crps2}).
-Here $F(\cdot)$ is the forecast distribution function, and $y^*$ is the observed targeted value of the forecast.
-The orientation of the CRPS is negative, meaning the smaller the score the better.
-Here $\vmathbb{1}\{\cdot\}$ is the indicator function.
-The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.
+Model comparison was done by calculating the continuous ranked probability 
+score (CRPS) for each forecast. The CRPS is a proper scoring rule which 
+evaluates the forecast distribution function. Proper scoring rules are the 
+current standard for comparing performance between probabilistic forecasts 
+and selecting the best forecasts according to the notion of maximizing 
+sharpness subject to (auto-)calibration 
+\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
+rules are commonly used in forecast comparison and have the property that a 
+forecaster is incentivized to be honest in the reporting of their forecasts 
+\cite[]{gneiting2007strictly, gneiting2014probabilistic}. The CRPS is defined 
+in (\ref{eq:crps2}). Here $F(\cdot)$ is the forecast distribution function, 
+and $y^*$ is the observed targeted value of the forecast. The orientation of 
+the CRPS is negative, meaning the smaller the score the better. Here 
+$\vmathbb{1}\{\cdot\}$ is the indicator function.
+The CRPS is calculated using a quantile decomposition from 
+\cite[]{laio2007verification}.
 
@@
 Figure \ref{fig:crps_by_week_horizon} shows 
-boxplots of the CRPS for the 500 simulation replicates faceted by week and horizon. 
+boxplots of the CRPS for the 500 simulation replicates faceted by week and 
+horizon. 
 % Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
-Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with lower medians and lower variation than the two models not accounting for ILI discrepancy.
-This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough.
-This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22.
-Among the four models, ASG tends to have the lowest CRPS values though this is not always the case.
-This suggests that the additional flexibility of the five parameter ASG function is often better for forecasting than that of the three parameter SIR function.
-Additional
-plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and another common proper scoring rule known as the logarithmic score.
+Notable plots here are for week 20 where the two models accounting for the 
+ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with 
+lower medians and lower variation than the two models not accounting for ILI 
+discrepancy. This makes sense given the forecasts at week 20 are forecasting 
+weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and 
+trough. This confirms that there is value in modeling discrepancy at least 
+during around the holiday weeks near week 22. Among the four models, ASG 
+tends to have the lowest CRPS values though this is not always the case. 
+This suggests that the additional flexibility of the five parameter ASG 
+function is often better for forecasting than that of the three parameter SIR 
+function. Additional
+plots in the supplemtary material \cite[]{wadsworth2024bas} further show that 
+the ASG models often but not always outperform the SIR models according to the 
+CRPS and another common proper scoring rule known as the logarithmic score.
 
@@
     \includegraphics[scale=.4]{Images/crps_by_week_horizon.png}
-    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis).
-Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
+    \caption{Boxplots of continuous ranked probability score (CRPS) for the 
+    four ILI models over all seasons in the simulation study faceted by horizon 
+    (x-axis) and week (y-axis). Horizons include 1-4 week ahead forecasts and 
+    weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
     \label{fig:crps_by_week_horizon}
@@
 
-In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations.
-The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$.
-A single forecast is thus comprised of 11 predictive intervals and a median.
-Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, and forecasts were made at the state and national level.
-The first week of forecasting took place during the week of October 7, 2023, and the final week was the week of April 27, 2024 making 29 total weeks of forecasts.
- The same format was used during the 2021 and 2022 seasons and for the COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
+In this section we apply the forecast models to make forecasts of the 2023 
+flu season weekly hospitalizations. The scoring of the forecasts is in the 
+context of the FluSight competition where each competing forecast was 
+submitted as a set of 23 quantiles corresponding to the given probability 
+levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$. A 
+single forecast is thus comprised of 11 predictive intervals and a median.
+Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
+and forecasts were made at the state and national level. The first week of 
+forecasting took place during the week of October 7, 2023, and the final week 
+was the week of April 27, 2024 making 29 total weeks of forecasts.
+ The same format was used during the 2021 and 2022 seasons and for the 
+ COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
 % Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
- Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS).
-The focus in this section will be on the LWIS (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
+ Primary scores for evaluating each forecast were the weighted interval score 
+ (WIS), the log-weighted interval score (LWIS), and the relative weighted 
+ interval score (RWIS). The focus in this section will be on the LWIS 
+ (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
 
-The WIS is a proper scoring rule used for scoring quantile or interval forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all included quantiles, $B$ is the number of intervals,  $y^*$ is the observed value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ interval.
-$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a single interval as defined in (\ref{eq:is}).
-The goal of the forecaster is to minimize the WIS.
-The LWIS is the same as the WIS except that it is evaluated over the log of quantiles and the log of the observed value.
+The WIS is a proper scoring rule used for scoring quantile or interval 
+forecasts 
+\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} 
+and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
+included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
+value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
+weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
+interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
+single interval as defined in (\ref{eq:is}). The goal of the forecaster is to 
+minimize the WIS. The LWIS is the same as the WIS except that it is evaluated 
+over the log of quantiles and the log of the observed value.
 
@@
 
-We fit 24 forecast models for each location for all 29 weeks, and for each week forecast 1-4 week ahead hospitalizations.
-The 24 models included all combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST hospitalization models, and both quadratic and linear hospitalization models.
+We fit 24 forecast models for each location for all 29 weeks, and for each 
+week forecast 1-4 week ahead hospitalizations. The 24 models included all 
+combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
+hospitalization models, and both quadratic and linear hospitalization models. 
 The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
-Because we set $S_{0s} = 0.9$, prior distributions were assigned only to $I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery rate $\delta_s = \rho_s \beta_s$.
-Here $\vmathbb{1}_{A}$ represents the indicator function for values within the set $A$.
+Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
+$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
+rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
+indicator function for values within the set $A$.
 
@@
 
- In this manuscript we introduce a statistical modeling framework which allows for the incorporation of several ILI forecast modeling methods.
-Specifically, we built upon Osthus et al.
-\cite[]{osthus2019dynamic} and introduced a framework for modeling ILI which includes the use of an arbitrary function for modeling the main trajectory of ILI along with modeling the discrepancy.
-We model flu hospitalizations by incorporating the ILI forecast model into a model forecasting hospitalizations where hospitalization predictions are a linear or quadratic function of ILI.
+ In this manuscript we introduce a statistical modeling framework which allows 
+ for the incorporation of several ILI forecast modeling methods. Specifically, 
+ we built upon Osthus et al. \cite[]{osthus2019dynamic} and introduced a 
+ framework for modeling ILI which includes the use of an arbitrary function 
+ for modeling the main trajectory of ILI along with modeling the discrepancy. 
+ We model flu hospitalizations by incorporating the ILI forecast model into a 
+ model forecasting hospitalizations where hospitalization predictions are a 
+ linear or quadratic function of ILI.
 
-The simulation study in section \ref{sec:simulation2} suggests the ASG function in ILI modeling may slightly outperform the SIR model according to the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season forecasts, the SIR model was overall superior.
-The results from both the simulation study and the real data analysis suggest that the addition of a discrepancy component in ILI modeling may improve forecasts especially near the holiday week between Christmas and New Years day.
-It should not be assumed that these conclusions may be generalized for all locations in the US, weeks of a season, or for future flu seasons.
+The simulation study in section \ref{sec:simulation2} suggests the ASG 
+function in ILI modeling may slightly outperform the SIR model according to 
+the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season 
+forecasts, the SIR model was overall superior. The results from both the 
+simulation study and the real data analysis suggest that the addition of a 
+discrepancy component in ILI modeling may improve forecasts especially near 
+the holiday week between Christmas and New Years day. It should not be assumed 
+that these conclusions may be generalized for all locations in the US, weeks 
+of a season, or for future flu seasons. 
 % Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.
 
-Forecasting the seasonal influenza outbreak remains a challenging task for forecasters.
-The general modeling framework in this manuscript is successful under diverse modeling conditions for all locations in the US and may contribute to future forecasting efforts.
-All forecast models were presented as separate from one another, but in the various locations and times of the flu season, different models perform better than others.
-To build on the work done here, a natural step forward would be to combine all or a few selected forecasts into an ensemble forecast.
-Such an ensemble may work to cancel out certain model biases or highlight model strengths, leading to more robust forecasts.
-
-\section{Acknowledgments}
-This work is partially supported by the National Science Foundation under Grant No.
-2152117.
-Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
-
-\section{Supplementary Material}
-Supplement to "Forecasting Influenza Hospitalizations Using a
-Bayesian Hierarchical Nonlinear Model with
-Discrepancy"\\
-A. Additional plots for ILI \& Hospitalization data for select regions, results from simulation study, and 2023-24 flu analysis\\
-% \end{supplement}
-B. Posterior distribution plots for select parameters\\
-% \sdescription{2. Plots of posterior distributions for parameters in both ILI and hospitalization models}
-C. FluSight forecast competition scoring results
-
-%Bibliography
-% \bibliographystyle{unsrt}  
-% \bibliography{references}
-
-\bibliographystyle{apa}
-\bibliography{master_bib}
-
-\end{document}
+Forecasting the seasonal influenza outbreak remains a challenging task for 
+forecasters. The general modeling framework in this manuscript is successful 
+under diverse modeling conditions for all locations in the US and may 
+contribute to future forecasting efforts. All forecast models were presented 
+as separate from one another, but in the various locations and times of the 
+flu season, different models perform better than others. To build on the work 
+done here, a natural step forward would be to combine all or a few selected 
+forecasts into an ensemble forecast. Such an ensemble may work to cancel out 
+certain model biases or highlight model strengths, leading to more robust 
+forecasts.

### 2025-12-01 response_round1.tex
change italics

@@
 uncertainty propogation was done. Specifics are found below and in the 
-revised manuscript. Our responses to each reviewers particular comments are
-\emph{italicized}, and specific changes/additions to the main manuscript are 
+revised manuscript. Reviewer comments are \emph{italicized}, our responses to 
+each reviewer's comments are in normal script, 
+and specific changes/additions to the main manuscript are 
 indented.
@@
 
-I think the authors did a great job in this newly submitted version and 
-addressed most of the earlier raised comments in a nice and convincing way. 
-The paper is well written and interesting to read.
-
-I only have two remaining comments.
+% \emph{
+% I think the authors did a great job in this newly submitted version and 
+% addressed most of the earlier raised comments in a nice and convincing way. 
+% The paper is well written and interesting to read. 
+% }
+% I only have two remaining comments.
 
@@
 
-  \item Section 3.7. End of of first paragraph: ``..., but propagation of 
+  \item \emph{Section 3.7. End of of first paragraph: ``..., but propagation of 
   uncertainty   in ILI forecasts was stil permitted in the estimating of 
   hospitalisation forecasts.``
-
+  }
+  \emph{
   Please clarify more explicitly how uncertainty was propagated between the two 
@@
   already at this stage and consider adding a brief discussion of the advantages 
-  and disadvantages of this approach.
+  and disadvantages of this approach.}
   \\~\\
-  \emph{To clarify how uncertainty was propogated, we consildated sections
+  To clarify how uncertainty was propogated, we consildated sections
   3.7 and 3.8, reordered the main ideas, and removed redundant information. We
   also added some advantages/disadvantages of our approach. The changes now
-  read:}
+  read:
   \begin{quote}
@@
         \label{sec:implementation_posterior}
-        \emph{
+
         Rather than combine the ILI and hopsitalization models into a single model,
@@
         where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
-        predicted or forecasted ILI at $i$ weeks in the future.}
+        predicted or forecasted ILI at $i$ weeks in the future.
 
@@
 
-        \emph{\indent In the second stage, the posterior distribution
+        \indent In the second stage, the posterior distribution
         for parameters of model (6) given observed
@@
         ILI and parameter posterior distribution draws was then plugged into equation
-        (6) to produce}
+        (6) to produce
 
@@
         \]
-        \emph{where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
+        where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
         $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
@@
         set to 0 to reflect realistic values of hospitalizations and comply with the 
-        FluSight forecasting rules.}
+        FluSight forecasting rules.
         \\~\\
-        \emph{
+
         \indent In the two-stage scheme, the posterior distribution of 
@@
         the forecasts from the component models generally outperformed those of the 
-        joint model in terms of minimizing proper scoring rules.}
+        joint model in terms of minimizing proper scoring rules.
         \\~\\
   \end{quote}
-
+  \emph{
   \item I recommend including the statement raised by the reviewer in the first 
@@
   using WIS instead of CRPS is well explained, especially given that WIS is the 
-  primary evaluation measure used in the FluSight competition 
+  primary evaluation measure used in the FluSight competition} 
   \\~\\
 
-  \emph{The following statements were added to page 17.}
+  The following statements were added to page 17.
   \begin{quote}
-  \emph{
+
       Commonly used proper scoring rules for continuous distribution forecasts
       include the log score and the continuously
-      ranked probability score (CRPS).}
+      ranked probability score (CRPS).
       \\~\\
-  \emph{
+
       We note that as the number of forecast intervals $B$ increases, the WIS becomes 
@@
       the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
-      \cite[]{bracher2021evaluating}.}
+      \cite[]{bracher2021evaluating}.
       \\~\\
@@
 \\~\\
-\noindent The authors have addressed all my comments from the first round. I only have two remaining points
-on the extended empirical analysis on FluSight.
+\noindent \emph{The authors have addressed all my comments from the first round. I only have two remaining points
+on the extended empirical analysis on FluSight.}
 
@@
 
-  \item The authors mention the difficulty of missing submissions during the FluSight season (hence the
+  \item \emph{The authors mention the difficulty of missing submissions during the FluSight season (hence the
     display chosen for Figure 10). How was this handled for Table 2? Do the averages refer to different
@@
     Models with very small submission fractions (maybe those with less than 80\%, definitely those with
-    less than 50\%) in my opinion should just be removed altogether.
+    less than 50\%) in my opinion should just be removed altogether.}
     \\~\\
-    \emph{The columns RLWIS and RWIS in table 2 correspond directly to the 
+          The columns RLWIS and RWIS in table 2 correspond directly to the 
           values plotted in figure 10. We rearranged the table so that RLWIS and
@@
           To clarify what's given in table 2 we stated the
-          following.}
-     \emph{     
+          following.
+
     \begin{quote}
@@
     \end{quote}
-    }
 
     \begin{quote}
-    \emph{
+
       During the FluSight competition, most participating teams would occasionally 
@@
       FluSight targets during the 2023 season.
-      }
-      \emph{
+
       Because forecast submissions for some targets were occassionally missed
@@
       targets for which forecasts were included or exlcuded. This makes a
-      simple average of scores is an imperfect measure for comparison.}
+      simple average of scores is an imperfect measure for comparison.
     \end{quote}
 
-  \item It is uncommon to report median scores as their statistical properties are unclear. As mean score
+  \item \emph{It is uncommon to report median scores as their statistical properties are unclear. As mean score
     are also there I'm not fundamentally opposed to reporting them, but in my opinion mean scores
-    alone are sufficient.  
+    alone are sufficient.}  
     \\~\\
-    \emph{Thank you for this comment. Because of it, we removed the median scores
+    Thank you for this comment. Because of it, we removed the median scores
     from the table which improved the aesthetic of the table as well as simplified
-    it. See the new table below.}
+    it. See the new table below.
 \\~\\    
@@
 % In spite of the above criticisms, 
-I appreciate the hierarchical modeling 
+\emph{I appreciate the hierarchical modeling 
 approach of fitting the two models separately in components. But I wonder how 
@@
 on health`` by 
-D Lee, S Mukhopadhyay, A Rushworth, and SK Sahu, in Biostatistics (2017).
+D Lee, S Mukhopadhyay, A Rushworth, and SK Sahu, in Biostatistics (2017).}
 \\~\\
-\emph{We appreciated this comment and the included reference. We used the reference
+We appreciated this comment and the included reference. We used the reference
 to assist in our new explanation, and made sure to cite it. 
@@
   also added some advantages/disadvantages of our approach. The changes now
-  read:}
+  read:
   \begin{quote}
@@
         the forecasts from the component models generally outperformed those of the 
-        joint model in terms of minimizing proper scoring rules.}
+        joint model in terms of minimizing proper scoring rules.
         \\~\\
@@
 
-In conclusion, I do think that the authors have done a good job of analysis and 
+\emph{In conclusion, I do think that the authors have done a good job of analysis and 
 simulation based demonstrations of the results. These can be further improved 
@@
 ``Bayesian modeling of spatio-temporal data with R`` by Sahu 
-(Chapman and Hall, 2022). 
+(Chapman and Hall, 2022).} 
 \\~\\
-\emph{In the simulation study in section 4 of the first version of this paper, 
+In the simulation study in section 4 of the first version of this paper, 
 scoring was
@@
 below and the added statements are bolded.
-}
 
 \begin{quote}
-\emph{
+
 Model comparison was done by calculating a proper scoring rule for each 
@@
 the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
-\cite[]{bracher2021evaluating}.}}
+\cite[]{bracher2021evaluating}.}
 \end{quote}

### 2025-12-01 response_round1.tex
fix spelling in response file

@@
   \item \emph{Section 3.7. End of of first paragraph: ``..., but propagation of 
-  uncertainty   in ILI forecasts was stil permitted in the estimating of 
+  uncertainty   in ILI forecasts was still permitted in the estimating of 
   hospitalisation forecasts.``
@@
   \\~\\
-  To clarify how uncertainty was propogated, we consildated sections
+  To clarify how uncertainty was propogated, we consolidated sections
   3.7 and 3.8, reordered the main ideas, and removed redundant information. We

### 2025-12-03 response_round1.tex
clean up section 3.7

@@
         Rather than combine the ILI and hopsitalization models into a single model,
-        (1) and 
+        the ILI model (1) and hospitalization model
         (6) were fit separately via Bayesian posterior updating 
@@
         forecasts. In the first stage, future ILI 
-        forecasts were obtained from the posterior predictive distribution where for
-        week $w$, the predictive distribution was obtained by integrating over the 
-        parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
-        as in (\ref{eq:ili_post}) where 
-        $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
-        is the density function of the posterior distribution for the model parameters
+        forecasts were obtained from the ILI forecast model in (7).
+        Here 
+        $p(\xi | \textbf{ILI}_w)$ 
+        is posterior distribution from the ILI model (1)
         and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$.
-        The desired forecasts were for weeks $w + i$ 
-        where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
+        The ILI forecast model was obtained by integrating over the 
+        parameters 
+        $\xi = (\pi$, $\kappa$, $\sigma^2_{\gamma}$, $\sigma^2_{\gamma_W})$.
+        The desired forecasts are for weeks $w + i$ 
+        where $i$ is a positive integer, and $\widetilde{ILI}_{w + i}$ represents the 
         predicted or forecasted ILI at $i$ weeks in the future.
+        Note that
+        in this section
+        the parameter subscript for season $s$ is dropped to simplify notation.
+
+        % \begin{equation}
+        %     \label{eq:ili_post}
+        %     p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
+        %     p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
+        %     \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
+        %     \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
+        %     d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+        % \end{equation}
 
         \begin{equation}
+        \tag{7}
             \label{eq:ili_post}
-            \tag{7}
-            p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
-            p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
-            \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
-            \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
-            d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+            p(\widetilde{ILI}_{w +i} | \textbf{ILI}_w) = \int  
+            p(\widetilde{ILI}_{w + i} | \xi) 
+            p(\xi | \textbf{ILI}_w) 
+            d\xi
         \end{equation}
@@
 
-        \indent In the second stage, the posterior distribution
-        for parameters of model (6) given observed
-        hospitalization data $\textbf{H}_w$, observed ILI data, and
-        denoted as 
-        $p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
-        was obtained.
-        The posterior predictive in (\ref{eq:ili_post}) was then combined with 
-        the hospitalization parameter posterior distribution
-        to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
+        In the second stage, the posterior distribution
+        for parameters of the hospitalization 
+        model (6) given observed
+        hospitalization data $\textbf{H}_w$ and $\textbf{ILI}_w$ was obtained.
+        The posterior distribution is denoted as
+        $p(\psi | \textbf{H}_w, \textbf{ILI}_w)$.
+        The ILI forecast model in (7) was then combined with 
+        the hospitalization posterior distribution
+        to produce a hospitalization forecast $\widetilde{H}_{w + i}$ 
         for week $w + i$.
-        Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
-        parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
-        That is, the first stage component model produced $K$ draws
-        $(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
-        and the second stage component model produced
-        $K$ parameter draws 
-        $(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
-        $(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
-        ILI and parameter posterior distribution draws was then plugged into equation
-        (6) to produce
+        Both the ILI forecast model in (7) and the hospitalization
+        posterior distributions were estimated via Markov chain Monte Carlo 
+        (MCMC) sampling.
+        In the first or ILI stage, $K$ MCMC draws
+        $(\widetilde{ILI}_{w +i}^{(1)}, ..., \widetilde{ILI}_{w +i}^{(K)})$
+        from the forecast model were produced,
+        and in the second or hospitalization stage
+        $K$ draws $(\psi^{(1)}, ..., \psi^{(K)})$ 
+        from the hospitalization posterior 
+        were produced. Each of the
+        ILI forecast and hospitalization draws was then plugged into 
+        the hospitalization model
+        (6) to produce the approximate hospitalization forecast
+        distribution in the bottom line of(8). 
+        Here $k \in \{1, ..., K\}$, 
+        and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
+        In the analyses in sections \ref{sec:simulation2} and \ref{sec:analysis}
+        we set $K = 50,000$ draws.
 
-        \[
-         \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
-                                          \alpha_{1s}^{(k)} 
-                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
-                                          \alpha_{2s}^{(k)} 
-                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
-                                          \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
-                                          \epsilon_{s,w + i}^{(k)}
-        \]
-        where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
-        $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
-        from which
-        the probabilistic hospitalization forecast was estimated. In practice we set
-        $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
-        \\~\\
-        \indent Occasionally $\{H^{(k)}_{s,w + i}\}^K$ included a small number of negative 
-        values,
-        which do not make sense when the distribution is meant to forecast
-        hospitalizations, a nonnegative number. We attempted to alleviate this problem
-        by modeling log-hospitalizations or by assuming hospitalizations followed a
-        distribution truncated at 0, but these models led to poorer forecasts
-        and issues drawing from the posterior distribution.
-        For the forecast competition analysis in 
-        section 5, all negative values of $\{H^{(k)}_{s,w + i}\}^K$ 
-        were 
-        set to 0 to reflect realistic values of hospitalizations and comply with the 
-        FluSight forecasting rules.
-        \\~\\
+        \begin{equation}
+        \tag{8}
+        \begin{aligned}
+        \label{eq:hosp_post_pred}
+          &p(\widetilde{H}_{w + i} | \textbf{H}_w, \textbf{ILI}_w)\\
+          &= \int p(\widetilde{H}_{w +i} | 
+               \psi, \widetilde{ILI}_{w +i}, \textbf{H}_w)
+               p(\psi | \textbf{H}_w, \textbf{ILI}_w) d \psi d\widetilde{ILI} \\
+          &\approx \frac{1}{K} p(\widetilde{H}_{w + i}^{(k)} | \psi^{(k)},  
+                              \tilde{ILI}_{w +i}^{(k)}, \textbf{H}_w, \textbf{ILI}_w) 
+        \end{aligned}
+        \end{equation}
+        % \[
+        %  \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
+        %                                   \alpha_{1s}^{(k)} 
+        %                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
+        %                                   \alpha_{2s}^{(k)} 
+        %                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
+        %                                   \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
+        %                                   \epsilon_{s,w + i}^{(k)}
+        % \]
+        % where $k \in \{1, ..., K\}$. 
+        % The hospitalization draws were then combined into
+        % $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
+        % from which
+        % the probabilistic hospitalization forecast was estimated.
+        % In practice we set
+        % $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
 
-        \indent In the two-stage scheme, the posterior distribution of 
+        % \[ \begin{array}{rl}
+        % p(\tilde{H}_w | H_w, ILI_w) 
+        % &=               \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | H_w, ILI_w)  d\tilde{ILI}_w \\
+        % &\approx  \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | ILI_w)  d\tilde{ILI}_w
+        % \end{array} \]
+
+        In the two-stage scheme, the posterior distribution of the ILI
         model (1) parameters was not directly influenced by the 
-        hospitalization data and the posterior distribution of model 
-        (6) parameters was not directly influenced by model 
-        (1) parameters. Hence there was limited feedback between the
-        two components, but propagation of uncertainty in ILI forecasts was still 
-        permitted in the estimating of hospitalization forecasts.
-        This two-stage modeling is remeniscent of the Bayesian cut 
+        hospitalization data and the posterior distribution of the hospitalization model 
+        (6) was not directly influenced by ILI model 
+        parameters. Hence there was limited feedback between the
+        two components, but the propagation of uncertainty in ILI forecasts was still 
+        accomplished in the estimating of hospitalization forecasts through
+        the forecast model (8).
+        This two-stage modeling is reminiscent of the Bayesian cut 
         described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
-        into two component models or ``modules`` may be justified for several reasons
+        into two component models or ``modules'' may be justified for several reasons
         including to reduce time to fit the model, avoiding mixing issues of posterior
@@
         useful information from hospiatlization data.
-        \\~\\
-        \indent The decision to cut the forecast model into two component models
+
+        The decision to cut the forecast model into two component models
         was made for three major reasons.

### 2025-12-03 manuscript_res.tex
clean up section 3.7

@@
 ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
-and/or a sore throat.`` Prior to the 2021 season, the definition included 
-``without a known cause other than influenza`` \cite[]{cdcfluview2023}. Because 
+and/or a sore throat.'' Prior to the 2021 season, the definition included 
+``without a known cause other than influenza'' \cite[]{cdcfluview2023}. Because 
 other illnesses such as COVID-19, RSV, and the common cold may induce similar 
@@
 For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
-estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
+estimated is 
+$\psi_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.
 
@@
 Rather than combine the ILI and hopsitalization models into a single model,
-(\ref{eq:ili_model}) and 
+the ILI model (\ref{eq:ili_model}) and hospitalization model
 (\ref{eq:final_hosp}) were fit separately via Bayesian posterior updating 
@@
 forecasts. In the first stage, future ILI 
-forecasts were obtained from the posterior predictive distribution where for
-week $w$, the predictive distribution was obtained by integrating over the 
-parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
-as in (\ref{eq:ili_post}) where 
-$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
-is the density function of the posterior distribution for the model parameters
+forecasts were obtained from the ILI forecast model in (\ref{eq:ili_post}).
+Here 
+$p(\xi | \textbf{ILI}_w)$ 
+is posterior distribution from the ILI model (\ref{eq:ili_model})
 and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$.
-The desired forecasts were for weeks $w + i$ 
-where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
+The ILI forecast model was obtained by integrating over the 
+parameters 
+$\xi = (\pi$, $\kappa$, $\sigma^2_{\gamma}$, $\sigma^2_{\gamma_W})$.
+The desired forecasts are for weeks $w + i$ 
+where $i$ is a positive integer, and $\widetilde{ILI}_{w + i}$ represents the 
 predicted or forecasted ILI at $i$ weeks in the future.
+Note that
+in this section
+the parameter subscript for season $s$ is dropped to simplify notation.
+
+% \begin{equation}
+%     \label{eq:ili_post}
+%     p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
+%     p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
+%     \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
+%     \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
+%     d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+% \end{equation}
 
@@
     \label{eq:ili_post}
-    p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
-    p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
-    \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
-    \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
-    d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
+    p(\widetilde{ILI}_{w +i} | \textbf{ILI}_w) = \int  
+    p(\widetilde{ILI}_{w + i} | \xi) 
+    p(\xi | \textbf{ILI}_w) 
+    d\xi
 \end{equation}
@@
 In the second stage, the posterior distribution
-for parameters of model (\ref{eq:final_hosp}) given observed
-hospitalization data $\textbf{H}_w$, observed ILI data, and
-denoted as 
-$p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
-was obtained.
-The posterior predictive in (\ref{eq:ili_post}) was then combined with 
-the hospitalization parameter posterior distribution
-to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
+for parameters of the hospitalization 
+model (\ref{eq:final_hosp}) given observed
+hospitalization data $\textbf{H}_w$ and $\textbf{ILI}_w$ was obtained.
+The posterior distribution is denoted as
+$p(\psi | \textbf{H}_w, \textbf{ILI}_w)$.
+The ILI forecast model in (\ref{eq:ili_post}) was then combined with 
+the hospitalization posterior distribution
+to produce a hospitalization forecast $\widetilde{H}_{w + i}$ 
 for week $w + i$.
-Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
-parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
-That is, the first stage component model produced $K$ draws
-$(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
-and the second stage component model produced
-$K$ parameter draws 
-$(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
-$(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
-ILI and parameter posterior distribution draws was then plugged into equation
-(\ref{eq:final_hosp}) to produce
-
-\[
- \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
-                                  \alpha_{1s}^{(k)} 
-                                  (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
-                                  \alpha_{2s}^{(k)} 
-                                  (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
-                                  \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
-                                  \epsilon_{s,w + i}^{(k)}
-\]
-where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
-$\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
-from which
-the probabilistic hospitalization forecast was estimated. In practice we set
-$K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
-
-Occasionally $\{H^{(k)}_{s,w + i}\}^K$ included a small number of negative 
-values,
-which do not make sense when the distribution is meant to forecast
-hospitalizations, a nonnegative number. We attempted to alleviate this problem
-by modeling log-hospitalizations or by assuming hospitalizations followed a
-distribution truncated at 0, but these models led to poorer forecasts
-and issues drawing from the posterior distribution.
-For the forecast competition analysis in 
-section \ref{sec:analysis}, all negative values of $\{H^{(k)}_{s,w + i}\}^K$ 
-were 
-set to 0 to reflect realistic values of hospitalizations and comply with the 
-FluSight forecasting rules.
-
-In the two-stage scheme, the posterior distribution of 
+Both the ILI forecast model in (\ref{eq:ili_post}) and the hospitalization
+posterior distributions were estimated via Markov chain Monte Carlo 
+(MCMC) sampling.
+In the first or ILI stage, $K$ MCMC draws
+$(\widetilde{ILI}_{w +i}^{(1)}, ..., \widetilde{ILI}_{w +i}^{(K)})$
+from the forecast model were produced,
+and in the second or hospitalization stage
+$K$ draws $(\psi^{(1)}, ..., \psi^{(K)})$ 
+from the hospitalization posterior 
+were produced. Each of the
+ILI forecast and hospitalization draws was then plugged into 
+the hospitalization model
+(\ref{eq:final_hosp}) to produce the approximate hospitalization forecast
+distribution in the bottom line of(\ref{eq:hosp_post_pred}). 
+Here $k \in \{1, ..., K\}$, 
+and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
+In the analyses in sections \ref{sec:simulation2} and \ref{sec:analysis}
+we set $K = 50,000$ draws.
+
+\begin{equation}
+\begin{aligned}
+\label{eq:hosp_post_pred}
+  &p(\widetilde{H}_{w + i} | \textbf{H}_w, \textbf{ILI}_w)\\
+  &= \int p(\widetilde{H}_{w +i} | 
+       \psi, \widetilde{ILI}_{w +i}, \textbf{H}_w)
+       p(\psi | \textbf{H}_w, \textbf{ILI}_w) d \psi d\widetilde{ILI} \\
+  &\approx \frac{1}{K} p(\widetilde{H}_{w + i}^{(k)} | \psi^{(k)},  
+                      \tilde{ILI}_{w +i}^{(k)}, \textbf{H}_w, \textbf{ILI}_w) 
+\end{aligned}
+\end{equation}
+% \[
+%  \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
+%                                   \alpha_{1s}^{(k)} 
+%                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
+%                                   \alpha_{2s}^{(k)} 
+%                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
+%                                   \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
+%                                   \epsilon_{s,w + i}^{(k)}
+% \]
+% where $k \in \{1, ..., K\}$. 
+% The hospitalization draws were then combined into
+% $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
+% from which
+% the probabilistic hospitalization forecast was estimated.
+% In practice we set
+% $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
+
+% \[ \begin{array}{rl}
+% p(\tilde{H}_w | H_w, ILI_w) 
+% &=               \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | H_w, ILI_w)  d\tilde{ILI}_w \\
+% &\approx  \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | ILI_w)  d\tilde{ILI}_w
+% \end{array} \]
+
+In the two-stage scheme, the posterior distribution of the ILI
 model (\ref{eq:ili_model}) parameters was not directly influenced by the 
-hospitalization data and the posterior distribution of model 
-(\ref{eq:final_hosp}) parameters was not directly influenced by model 
-(\ref{eq:ili_model}) parameters. Hence there was limited feedback between the
-two components, but propagation of uncertainty in ILI forecasts was still 
-permitted in the estimating of hospitalization forecasts.
-This two-stage modeling is remeniscent of the Bayesian cut 
+hospitalization data and the posterior distribution of the hospitalization model 
+(\ref{eq:final_hosp}) was not directly influenced by ILI model 
+parameters. Hence there was limited feedback between the
+two components, but the propagation of uncertainty in ILI forecasts was still 
+accomplished in the estimating of hospitalization forecasts through
+the forecast model (\ref{eq:hosp_post_pred}).
+This two-stage modeling is reminiscent of the Bayesian cut 
 described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
-into two component models or ``modules`` may be justified for several reasons
+into two component models or ``modules'' may be justified for several reasons
 including to reduce time to fit the model, avoiding mixing issues of posterior
@@
 Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
-implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn 
+implements Hamiltonian Monte Carlo sampling with the No-U-turn 
 sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
@@
 \cite[]{wadsworth2025bas}.
+
+Occasionally the draws making up the approximate forecast
+distribution
+% $\{H^{(k)}_{s,w + i}\}^K$ 
+included a small number of negative 
+values,
+which do not make sense when the forecast is meant to predict total
+hospitalizations, a nonnegative number. We attempted to alleviate this problem
+by modeling log-hospitalizations or by assuming hospitalizations followed a
+distribution truncated at 0, but these models led to poorer forecasts
+and issues drawing from the posterior distribution.
+For the forecast competition analysis in 
+section \ref{sec:analysis}, all negative values from 
+the hospitalization model (\ref{eq:hosp_post_pred})
+were 
+set to 0 to reflect realistic values of hospitalizations and comply with the 
+FluSight forecasting rules.
 
@@
 existing ILI data as a predictive covariate. Each season was in turn 
-``left-out`` and treated as if it was the most recent season for which forecasts
+``left-out'' and treated as if it was the most recent season for which forecasts
 are desired. Fitting and forecasting was then done for weeks 14, 20, 26, 32, 

### 2025-12-03 manuscript_res.tex
clean up section 3.7

@@
 
-\begin{table}[ht]
-\centering
-% \begin{adjustwidth}{-3cm}{-1.5cm}
-\caption{Overall summary scores for each of the 13 non-ensemble FluSight
-models with over 80\% submissions and 
-the ASGD\_NORM2 and SIRD\_NORM2 models. 
-Summaries include mean weighted interval
-and log-weighted interval scores (MWIS/MLWIS), 
-mean squared error
-difference between predictive model empirical and theoretical coverage (MSEC)
-see figure \ref{fig:lwis_cover_sum} c)), percentage of forecasts for all 
-targets during the 2023 season (\% Forcs), and relative weighted interval
-and log-weighted interval scores (RWIS/RLWIS).
-\\}
-\small
-\begin{tabular}{lrrrrrrrr}
+% \begin{table}[ht]
+% \centering
+% % \begin{adjustwidth}{-3cm}{-1.5cm}
+% \caption{Overall summary scores for each of the 13 non-ensemble FluSight
+% models with over 80\% submissions and 
+% the ASGD\_NORM2 and SIRD\_NORM2 models. 
+% Summaries include mean weighted interval
+% and log-weighted interval scores (MWIS/MLWIS), 
+% mean squared error
+% difference between predictive model empirical and theoretical coverage (MSEC)
+% see figure \ref{fig:lwis_cover_sum} c)), percentage of forecasts for all 
+% targets during the 2023 season (\% Forcs), and relative weighted interval
+% and log-weighted interval scores (RWIS/RLWIS).
+% \\}
+% \small
+% \begin{tabular}{lrrrrrrrr}
+%   % \hline
+%                               & MWIS & MLWIS & MSEC  & \% Forcs & RWIS & RLWIS\\ 
 %   \hline 
-                              & MWIS & MLWIS & MSEC  & \% Forcs & RWIS & RLWIS\\ 
-  \hline 
-ASGD\_NORM2                   & 62   & 0.34  & 0.001 & 100      & 1    & 1 \\ 
-  PSI-PROF                    & 63   & 0.35  & 0.003 & 100      & 0.90 & 0.97 \\ 
-  SIRD\_NORM2                 & 67   & 0.36  & 0.001 & 100      & 0.93 & 0.95 \\
-  CEPH-Rtrend\_fluH           & 69   & 0.39  & 0.004 & 100      & 0.81 & 0.86 \\ 
-  FluSight-baseline           & 85   & 0.41  & 0.023 & 100      & 0.63 & 0.82 \\ 
-  UM-DeepOutbreak             & 83   & 0.46  & 0.010 & 99       & 0.64 & 0.72 \\ 
-  cfarenewal-cfaepimlight     & 76   & 0.37  & 0.001 & 99       & 0.72 & 0.92 \\ 
-  MOBS-GLEAM\_FLUH            & 66   & 0.35  & 0.021 & 98       & 0.87 & 0.95 \\ 
-  LosAlamos\_NAU-CModel\_Flu  & 242  & 0.64  & 0.153 & 93       & 0.21 & 0.53 \\ 
-  UGuelph-CompositeCurve      & 132  & 0.52  & 0.018 & 93       & 0.43 & 0.62 \\ 
-  LUcompUncertLab-chimera     & 84   & 0.43  & 0.004 & 92       & 0.67 & 0.71 \\ 
-  UGA\_flucast-INFLAenza      & 74   & 0.32  & 0.003 & 90       & 0.77 & 1.07 \\ 
-  cfa-flumech                 & 101  & 0.41  & 0.019 & 89       & 0.58 & 0.85 \\ 
-  UGA\_flucast-Copycat        & 76   & 0.35  & 0.012 & 87       & 0.81 & 0.98 \\ 
-  GT-FluFNP                   & 105  & 0.48  & 0.046 & 86       & 0.45 & 0.73 \\ 
-  % UNC\_IDD-InfluPaint         & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
-  % NU\_UCSD-GLEAM\_AI\_FLUH    & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
-  % SigSci-CREG                 & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
-  % PSI-PROF\_beta              & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
-  % NIH-Flu\_ARIMA              & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
-  % GH-model                    & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
-  % UGA\_flucast-OKeeffe        & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
-\end{tabular}
-% \end{adjustwidth}
-\label{tab:fin_analysis_stats}
-\end{table}
+% ASGD\_NORM2                   & 62   & 0.34  & 0.001 & 100      & 1    & 1 \\ 
+%   PSI-PROF                    & 63   & 0.35  & 0.003 & 100      & 0.90 & 0.97 \\ 
+%   SIRD\_NORM2                 & 67   & 0.36  & 0.001 & 100      & 0.93 & 0.95 \\
+%   CEPH-Rtrend\_fluH           & 69   & 0.39  & 0.004 & 100      & 0.81 & 0.86 \\ 
+%   FluSight-baseline           & 85   & 0.41  & 0.023 & 100      & 0.63 & 0.82 \\ 
+%   UM-DeepOutbreak             & 83   & 0.46  & 0.010 & 99       & 0.64 & 0.72 \\ 
+%   cfarenewal-cfaepimlight     & 76   & 0.37  & 0.001 & 99       & 0.72 & 0.92 \\ 
+%   MOBS-GLEAM\_FLUH            & 66   & 0.35  & 0.021 & 98       & 0.87 & 0.95 \\ 
+%   LosAlamos\_NAU-CModel\_Flu  & 242  & 0.64  & 0.153 & 93       & 0.21 & 0.53 \\ 
+%   UGuelph-CompositeCurve      & 132  & 0.52  & 0.018 & 93       & 0.43 & 0.62 \\ 
+%   LUcompUncertLab-chimera     & 84   & 0.43  & 0.004 & 92       & 0.67 & 0.71 \\ 
+%   UGA\_flucast-INFLAenza      & 74   & 0.32  & 0.003 & 90       & 0.77 & 1.07 \\ 
+%   cfa-flumech                 & 101  & 0.41  & 0.019 & 89       & 0.58 & 0.85 \\ 
+%   UGA\_flucast-Copycat        & 76   & 0.35  & 0.012 & 87       & 0.81 & 0.98 \\ 
+%   GT-FluFNP                   & 105  & 0.48  & 0.046 & 86       & 0.45 & 0.73 \\ 
+%   % UNC\_IDD-InfluPaint         & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
+%   % NU\_UCSD-GLEAM\_AI\_FLUH    & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
+%   % SigSci-CREG                 & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
+%   % PSI-PROF\_beta              & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
+%   % NIH-Flu\_ARIMA              & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
+%   % GH-model                    & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
+%   % UGA\_flucast-OKeeffe        & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
+% \end{tabular}
+% % \end{adjustwidth}
+% \label{tab:fin_analysis_stats}
+% \end{table}
 

### 2025-12-05 response_round1.tex
completed revisions/responses

@@
 questions. Overall, they made a huge difference in improving the content of
-our manuscript. In the most recent revision, we are particularly grateful for 
+the manuscript. In the most recent revision, we are particularly grateful for 
 the questions on uncertainty propogation. Obviously the explanation we gave
-for how uncertainty propogation wasn't clear enough, so in this iteration of 
+for how uncertainty was propogated wasn't clear enough, so in this iteration of 
 we reworked sections 3.7 and 3.8 to make more clear how 
-uncertainty propogation was done. Specifics are found below and in the 
+it was done. Specifics are found below and in the 
 revised manuscript. Reviewer comments are \emph{italicized}, our responses to 
@@
   \begin{quote}
-        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
+        \subsection{3.7 Hospitalization Forecasts Using the Bayesian Cut}
         \label{sec:implementation_posterior}
@@
         Rather than combine the ILI and hopsitalization models into a single model,
+        forecasts were made via a two-stage fitting process. In the first stage,
         the ILI model (1) and hospitalization model
-        (6) were fit separately via Bayesian posterior updating 
-        in two stages and then fused together to produce hospitalization 
-        forecasts. In the first stage, future ILI 
-        forecasts were obtained from the ILI forecast model in (7).
+        (6) were fit separately via Bayesian posterior updating. 
+        In the second stage, the two model fits were
+        fused together to produce hospitalization 
+        forecasts. Future ILI 
+        forecasts were obtained from the ILI forecast model in (\ref{eq:ili_post}).
         Here 
         $p(\xi | \textbf{ILI}_w)$ 
-        is posterior distribution from the ILI model (1)
+        is the posterior distribution of the ILI model (1)
         and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$.
@@
         % \spencer{Isn't that captured in $d \pi_s$?}
-
-        In the second stage, the posterior distribution
-        for parameters of the hospitalization 
+        Also in the first stage, the posterior distribution of the hospitalization 
         model (6) given observed
@@
         $p(\psi | \textbf{H}_w, \textbf{ILI}_w)$.
-        The ILI forecast model in (7) was then combined with 
+        In the second stage,
+        the ILI forecast model in (\ref{eq:ili_post}) was then combined with 
         the hospitalization posterior distribution
@@
         for week $w + i$.
-        Both the ILI forecast model in (7) and the hospitalization
+
+        Both the ILI forecast model in (\ref{eq:ili_post}) and the hospitalization
         posterior distributions were estimated via Markov chain Monte Carlo 
         (MCMC) sampling.
-        In the first or ILI stage, $K$ MCMC draws
+        From the fit ILI model, $K$ MCMC draws
         $(\widetilde{ILI}_{w +i}^{(1)}, ..., \widetilde{ILI}_{w +i}^{(K)})$
         from the forecast model were produced, 
-        and in the second or hospitalization stage
         $K$ draws $(\psi^{(1)}, ..., \psi^{(K)})$
-        from the hospitalization posterior 
+        from the hospitalization model posterior distribution
         were produced. Each of the
@@
         (6) to produce the approximate hospitalization forecast
-        distribution in the bottom line of(8). 
+        distribution in the bottom line of (\ref{eq:hosp_post_pred}). 
         Here $k \in \{1, ..., K\}$, 
         and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
-        In the analyses in sections \ref{sec:simulation2} and \ref{sec:analysis}
+        In the analyses in sections 4 and 5
         we set $K = 50,000$ draws.
@@
         accomplished in the estimating of hospitalization forecasts through
-        the forecast model (8).
+        the forecast model (\ref{eq:hosp_post_pred}).
         This two-stage modeling is reminiscent of the Bayesian cut 
@@
         showed convergence even after 55,000 iterations, whereas the component model 
-        fits gave no signs which raised concern for convergence. And most importantly,
+        fits gave no signs which raised concern about convergence. And most importantly,
         the forecasts from the component models generally outperformed those of the 
@@
     \\~\\
-          The columns RLWIS and RWIS in table 2 correspond directly to the 
-          values plotted in figure 10. We rearranged the table so that RLWIS and
+          The columns RLWIS and RWIS in table 2 correspond to the 
+          values plotted in figure 10, except that in the figure the scores
+          are faceted by region or week. 
+          We rearranged the table so that RLWIS and
           RWIS come first. The updated table, shown as "Table 1" is below.
           To clarify what's given in table 2 we stated the
-          following.
+          following on page 22
 
@@
 
+    We also removed comparisons with any competing models with less than 
+    80\% submissions and further clarified the need to use relative scores.
+    The following 2 snippets were added on page 20.
+
     \begin{quote}
@@
       FluSight targets during the 2023 season. 
+      \\~\\
 
@@
     Thank you for this comment. Because of it, we removed the median scores
-    from the table which improved the aesthetic of the table as well as simplified
-    it. See the new table below.
+    from the table. This simplified the table and also greatly 
+    improved the aesthetic when compared to the previous table. 
+    See the new table below.
 \\~\\    
@@
 using the WIS because the available forecasts are not amenable to being scored
-the with the CRPS. A recommendation from one of the reviewers was to simplify 
-by choosing one or the other scores, and we chose the WIS. To respond to this 
-point, however, we modified the first paragraph on page 17. The paragraph is 
-below and the added statements are bolded.
+the with the CRPS. A recommendation from one of the reviewers led us to simplify 
+by choosing one or the other scores, and we chose the WIS.
+In response to your comment, 
+we modified the first paragraph on page 17. The paragraph is 
+below and the added statements are in bold.
 

### 2025-12-05 manuscript_res.tex
completed revisions/responses

@@
 Rather than combine the ILI and hopsitalization models into a single model,
+forecasts were made via a two-stage fitting process. In the first stage,
 the ILI model (\ref{eq:ili_model}) and hospitalization model
-(\ref{eq:final_hosp}) were fit separately via Bayesian posterior updating 
-in two stages and then fused together to produce hospitalization 
-forecasts. In the first stage, future ILI 
+(\ref{eq:final_hosp}) were fit separately via Bayesian posterior updating. 
+In the second stage, the two model fits were
+fused together to produce hospitalization 
+forecasts. Future ILI 
 forecasts were obtained from the ILI forecast model in (\ref{eq:ili_post}).
@@
 $p(\xi | \textbf{ILI}_w)$ 
-is posterior distribution from the ILI model (\ref{eq:ili_model})
+is the posterior distribution of the ILI model (\ref{eq:ili_model})
 and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$.
@@
 % \spencer{Isn't that captured in $d \pi_s$?}
-
-In the second stage, the posterior distribution
-for parameters of the hospitalization 
+Also in the first stage, the posterior distribution of the hospitalization 
 model (\ref{eq:final_hosp}) given observed
@@
 $p(\psi | \textbf{H}_w, \textbf{ILI}_w)$.
-The ILI forecast model in (\ref{eq:ili_post}) was then combined with 
+In the second stage,
+the ILI forecast model in (\ref{eq:ili_post}) was then combined with 
 the hospitalization posterior distribution
@@
 for week $w + i$.
+
 Both the ILI forecast model in (\ref{eq:ili_post}) and the hospitalization
@@
 (MCMC) sampling.
-In the first or ILI stage, $K$ MCMC draws
+From the fit ILI model, $K$ MCMC draws
 $(\widetilde{ILI}_{w +i}^{(1)}, ..., \widetilde{ILI}_{w +i}^{(K)})$
 from the forecast model were produced, 
-and in the second or hospitalization stage
 $K$ draws $(\psi^{(1)}, ..., \psi^{(K)})$
-from the hospitalization posterior 
+from the hospitalization model posterior distribution
 were produced. Each of the
@@
 (\ref{eq:final_hosp}) to produce the approximate hospitalization forecast
-distribution in the bottom line of(\ref{eq:hosp_post_pred}). 
+distribution in the bottom line of (\ref{eq:hosp_post_pred}). 
 Here $k \in \{1, ..., K\}$, 
@@
 showed convergence even after 55,000 iterations, whereas the component model 
-fits gave no signs which raised concern for convergence. And most importantly,
+fits gave no signs which raised concern about convergence. And most importantly,
 the forecasts from the component models generally outperformed those of the 

### 2025-12-12 response_round1.tex
adjust manuscript

@@
 We are grateful to the three reviewers and their insightful comments and 
-questions. Overall, they made a huge difference in improving the content of
+questions. We apologize for the omission of responses to some comments in our
+resubmission on 12/5/25, and are embarassed for having somehow overlooked that.
+We thank the editor for the chance to correct that mistake, and we have 
+corrected it.
+\\~\\
+Overall, the reviewer comments
+made a huge impact on improving the content of
 the manuscript. In the most recent revision, we are particularly grateful for 
-the questions on uncertainty propogation. Obviously the explanation we gave
-for how uncertainty was propogated wasn't clear enough, so in this iteration of 
-we reworked sections 3.7 and 3.8 to make more clear how 
-it was done. Specifics are found below and in the 
+the questions on uncertainty propagation. Obviously the explanation we gave
+for how uncertainty was propagated wasn't clear enough, so in this revision
+we reworked sections 3.7 and 3.8 to clarify the procedure.
+Specifics are found below and in the 
 revised manuscript. Reviewer comments are \emph{italicized}, our responses to 
@@
 
-  \item \emph{Section 3.7. End of of first paragraph: ``..., but propagation of 
+  \item \emph{Section 3.7. End of first paragraph: ``..., but propagation of 
   uncertainty   in ILI forecasts was still permitted in the estimating of 
@@
   \\~\\
-  To clarify how uncertainty was propogated, we consolidated sections
+  To clarify how uncertainty was propagated, we consolidated sections
   3.7 and 3.8, reordered the main ideas, and removed redundant information. We
@@
   \begin{quote}
-        \subsection{3.7 Hospitalization Forecasts Using the Bayesian Cut}
+        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
         \label{sec:implementation_posterior}
 
-        Rather than combine the ILI and hopsitalization models into a single model,
+        Rather than combine the ILI and hospitalization models into a single model,
         forecasts were made via a two-stage fitting process. In the first stage,
@@
         distribution in the bottom line of (\ref{eq:hosp_post_pred}). 
+        This has 
+        similarities to the procedure used in section 3.2.2 of 
+        \cite[]{sahu2014hierarchical} who use it to propagate the uncertainty of 
+        temperature forecasts into forecast model for hospitalizations.
         Here $k \in \{1, ..., K\}$, 
@@
         two components, but the propagation of uncertainty in ILI forecasts was still 
-        accomplished in the estimating of hospitalization forecasts through
+        accomplished in estimating hospitalization forecasts through
         the forecast model (\ref{eq:hosp_post_pred}).
         This two-stage modeling is reminiscent of the Bayesian cut 
-        described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
+        described in \cite[]{plummer2015cuts} and \cite[]{nott2023bayesian}. Cutting a model
         into two component models or ``modules'' may be justified for several reasons
@@
         application for example, the ILI model parameters receive no potentially 
-        useful information from hospiatlization data.
+        useful information from hospitalization data.
 
@@
         showed convergence even after 55,000 iterations, whereas the component model 
-        fits gave no signs which raised concern about convergence. And most importantly,
+        fits gave no signs which raised concerns about convergence. And most importantly,
         the forecasts from the component models generally outperformed those of the 
@@
 
-  The following statements were added to page 17.
+  The following statements were added on page 17.
   \begin{quote}
@@
           We rearranged the table so that RLWIS and
-          RWIS come first. The updated table, shown as "Table 1" is below.
+          RWIS come first. The updated table, shown as "Table 1" below.
           To clarify what's given in table 2 we stated the
@@
 
-      Because forecast submissions for some targets were occassionally missed
+      Because forecast submissions for some targets were occasionally missed
       by participating teams, for any two models there was often a mismatch between
-      targets for which forecasts were included or exlcuded. This makes a
-      simple average of scores is an imperfect measure for comparison.
+      targets for which forecasts were included or excluded. This makes a
+      simple average of scores an imperfect measure for comparison.
     \end{quote}
@@
 % 
-% This manuscript seems to be a very reasonable attempt at forecasting influenza 
-% induced hospitalizations. The
-% manuscript also puts the methodology in the context of the FluSight competition 
-% organised by the CDC in the US.
-% Apologies, I have found it hard to review as it is conceptually very hard to do 
-% so in my opinion. Firstly,
-% I do not find comfortable modelling the proportion, rather a natural outcome to 
-% model will be the number of hospitalization using a Generalized Linear model 
-% (Poisson distribution). Secondly, I find it un-intuitive to model the
-% flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model 
-% we focus on two aspects: the systematic part and the random part. Here, as far 
-% as  I understand the systematic part is the ASG function. The random
-% part has several components, which are hidden in Equation (5). My difficulty in 
-% modelling using the ASG
-% function lies in my inability to relate any meaningful meteorological variables 
-% (which are very much drivers of the seasonal flu outbreaks). See for example 
-% the paper,  ``A hierarchical Bayesian model for improving short-term forecasting 
-% of hospital demand by including meteorological information`` by  SK Sahu, 
-% B Baffour, PR Harper, JH Minty, C Sarran
-% Journal of the Royal Statistical Society, Series A (2014).
+\emph{This manuscript seems to be a very reasonable attempt at forecasting influenza
+induced hospitalizations. The
+manuscript also puts the methodology in the context of the FluSight competition
+organised by the CDC in the US.
+Apologies, I have found it hard to review as it is conceptually very hard to do
+so in my opinion.}
+
+\emph{Firstly,
+I do not find comfortable modelling the proportion, rather a natural outcome to
+model will be the number of hospitalization using a Generalized Linear model
+(Poisson distribution).}
+
+The influenza like illness (ILI) data model in equation (1) is indeed 
+proportion data, and is reported as the proportion of patients visiting
+a provider who display flu like symptoms. This is stated in the introduction
+on page 2, but in the revision we restate it again as below on page 4
+
+\begin{quote}
+An ILI case is defined as a 
+``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
+and/or a sore throat,'' and ILI data is reported as the proportion 
+or percentage of patients who display these symptoms.
+\end{quote}
+
+For that reason, we argue that 
+model (1) is appropriate for ILI data. 
+To clarify that the purpose of this paper is
+to use historically accurate ILI forecasts as a predictor of hospitalizations,
+we added the following two statements; the first in the introduction on page
+3 and the second in section 3 on page 7.
+
+\begin{quote}
+The contribution of this manuscript is to introduce a two component 
+Bayesian modeling framework 
+for modeling HHS hospitalization forecasts where hospitalization data and 
+years of ILI data are used to inform forecast models. 
+As is shown in this manuscript,
+ILI is a good predictor of hospitalizations, 
+and years of experience have led
+to effective ILI forecast methods. Thus producing accurate ILI forecasts 
+and using them to inform hospitalizations greatly assists in producing
+accurate hospitalization forecasts.
+\end{quote}
+
+\begin{quote}
+Presented in this section is methodology for forecasting hospitalizations
+using ILI forecasts as a predictor of hospitalization forecasts. 
+The availability of nearly a decade of ILI data along with accurate forecast
+models allows us to make accurate forecasts of ILI which are used as a linear
+predictor for hospitalization forecasts.
+\end{quote}
+
+Hospitalization data which we are trying to forecast is
+count data. Thus a Poisson is certainly a natural modeling choice. To 
+keep the model simple, however, we modeled hospitalizations using a normal
+distribution, a location-scale t distribution,
+or log transformed hospitalizations and modeled the transformed data with
+a normal linear model. See model (6).
+This choice was made because the forecasts were accurate,
+the modeling is simpler, and fitting the model using a Poisson was much slower
+making it more difficult to meet the forecast deadlines. This argument was 
+added on page 12.
+
+\begin{quote}
+Another natural choice would have been to
+model hospitalizations with a Poisson response generalized linear model. 
+However, a generalized linear model adds additional complexity and we found that
+fitting the model required significantly more time than
+model (6) making it more difficult to produce forecasts
+for all 54 locations and meet the FluSight 
+submission deadline which was roughly 12 hours after data were released.
+Also to avoid complexity, the authors of \cite[]{sahu2014hierarchical} opted to 
+forecast hospitalizations using a normal response model rather than a Poisson
+response model.
+\end{quote}
+
+\emph{Secondly, I find it un-intuitive to model the
+flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model
+we focus on two aspects: the systematic part and the random part. Here, as far
+as  I understand the systematic part is the ASG function. The random
+part has several components, which are hidden in Equation (5). My difficulty in
+modelling using the ASG
+function lies in my inability to relate any meaningful meteorological variables
+(which are very much drivers of the seasonal flu outbreaks). See for example
+the paper,  ``A hierarchical Bayesian model for improving short-term forecasting
+of hospital demand by including meteorological information`` by  SK Sahu,
+B Baffour, PR Harper, JH Minty, C Sarran
+Journal of the Royal Statistical Society, Series A (2014).}
+
+The ASG function, and likewise the susceptible-infectious-recovered (SIR)
+mechanistic model, are used to model
+ILI data because of known annual trajectory, the relationship to 
+mathematical infection disease models, and most importantly
+the success they've had in 
+producing accurate ILI forecasts.
+ILI data and forecasts are used as 
+linear predictors for hospitalizations in model (6).
+Neither model (1) nor (6) preclude the use of additional linear predictors
+for prediciting ILI or hospitalizations, but if additional predictors 
+were to be used it would make more sense to add them to model (6).
+We added the following two statements to discuss this important point.
+The first is near the beginning of section 3.5 on page 12, and the second is 
+in the discussion on page 24.
+
+\begin{quote}
+In model (6), the data is limited to the ILI and 
+hospitalization data provided by FluSight, but including additional
+linear predictors would be straightforward. In fact it has been shown that 
+meteorological predictors and forecasts such as temperature can lead
+to improved forecasts as well as geographic variables such as patient
+age and sex \cite[]{sahu2014hierarchical}.  
+\end{quote}
+
+\begin{quote}
+Additionally, it has been shown that meteorological variables
+such as local temperature forecasts and geographic variables such as age can 
+lead to improved hospitalization forecasts \cite[]{sahu2014hierarchical}. 
+A straightforward and useful extension of the methodology herein would be to 
+collect such variables and incorporate them as added hospitalization 
+predictors.
+\end{quote}
 
@@
 to assist in our new explanation, and made sure to cite it. 
-To clarify how uncertainty was propogated, we consildated sections
+To clarify how uncertainty was propagated, we consolidated sections
   3.7 and 3.8, reordered the main ideas, and removed redundant information. We
@@
         distribution in the bottom line of (\ref{eq:hosp_post_pred}). 
+        This has 
+        similarities to the procedure used in section 3.2.2 of 
+        \cite[]{sahu2014hierarchical}, which uses it to 
+        propagate the uncertainty of 
+        temperature forecasts into forecast model for hospitalizations.
         Here $k \in \{1, ..., K\}$, 
@@
 COVID-19 Forecast Hub--}
-or forecasts made up of predicitive intervals of several nominal 
+or forecasts made up of predictive intervals of several nominal 
 levels

### 2025-12-12 manuscript_res.tex
adjust manuscript

@@
 the CDC have been invited to participate. From hereon, we denote a 
-flu season by the year in which the outbreak began, i.e. the 2020 season 
+flu season by the year in which the outbreak began, i.e. the 2020 
 season outbreak began in 2020 and lasted until spring 2021.
@@
 for modeling HHS hospitalization forecasts where hospitalization data and 
-years of ILI data are used to inform forecast models. The first modeling 
+years of ILI data are used to inform forecast models. 
+As is shown in this manuscript,
+ILI is a good predictor of hospitalizations, 
+and years of experience have led
+to effective ILI forecast methods. Thus producing accurate ILI forecasts 
+and using them to inform hospitalizations greatly assists in producing
+accurate hospitalization forecasts.
+The first modeling 
 component is a hierarchical model of ILI data. 
@@
 ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
-and/or a sore throat.'' Prior to the 2021 season, the definition included 
+and/or a sore throat,'' and ILI data is reported as the proportion 
+or percentage of patients who display these symptoms.
+Prior to the 2021 season, the definition included 
 ``without a known cause other than influenza'' \cite[]{cdcfluview2023}. Because 
@@
 of influenza. From hereon, influenza hospitalizations
-will be referrred to as hospitalizations.
+will be referred to as hospitalizations.
 
@@
 data for all 50 US states are included in the supplementary material 
-\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories 
+\cite[]{wadsworth2025bas}. These plots show flu outbreak trajectories 
 similar to those in figure \ref{fig:us_ili}.
@@
 Later you have a section called ILI Model, but isn't this the ILI model?}
+Presented in this section is methodology for forecasting hospitalizations
+using ILI forecasts as a predictor of hospitalization forecasts. 
+The availability of nearly a decade of ILI data along with accurate forecast
+models allows us to make accurate forecasts of ILI which are used as a linear
+predictor for hospitalization forecasts.
 
@@
 a similar trajectory. Compartmental models are standard mathematical models 
-used for modelling disease outbreaks which may capture ILI behavior. 
+used for modeling disease outbreaks which may capture ILI behavior. 
 One important compartmental model is the 
@@
 \cite[]{raftery2010online,ljung1987system}. 
+In model (\ref{eq:final_hosp}), the data is limited to the ILI and 
+hospitalization data provided by FluSight, but including additional
+linear predictors would be straightforward. In fact it has been shown that 
+meteorological predictors and forecasts such as temperature can lead
+to improved forecasts as well as geographic variables such as patient
+age and sex \cite[]{sahu2014hierarchical}.
 
@@
 when $D_s$ belongs to the location-scale t (LST) family. 
+In the simulation study performed in section \ref{sec:simulation2} 
+$D_s$ belongs to 
+the normal distribution family, and for for analyzing the 2023 hospitalization
+data presented in section \ref{sec:analysis} models were considered where 
+$D_s$ was LST or where hospitalization counts were log transformed 
+and the transformed
+data were modeled as normally distributed.
+Another natural choice would have been to
+model hospitalizations with a Poisson response generalized linear model. 
+However, a generalized linear model adds additional complexity and we found that
+fitting the model required significantly more time than
+model (\ref{eq:final_hosp}) making it more difficult to produce forecasts
+for all 54 locations and meet the FluSight 
+submission deadline which was roughly 12 hours after data were released.
+Also to avoid complexity, the authors of \cite[]{sahu2014hierarchical} opted to 
+forecast hospitalizations using a normal response model rather than a Poisson
+response model.
 
@@
 however, we use empirical Bayes and center the first component of the 
-hierarichal prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
+hierarchical prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
 all seasons and set the variance parameter so the prior was tight around the
@@
 
-Rather than combine the ILI and hopsitalization models into a single model,
+Rather than combine the ILI and hospitalization models into a single model,
 forecasts were made via a two-stage fitting process. In the first stage,
@@
 (\ref{eq:final_hosp}) to produce the approximate hospitalization forecast
-distribution in the bottom line of (\ref{eq:hosp_post_pred}). 
+distribution in the bottom line of (\ref{eq:hosp_post_pred}). This has 
+similarities to the procedure used in section 3.2.2 of 
+\cite[]{sahu2014hierarchical}, which uses it to propagate the uncertainty of 
+temperature forecasts into forecast model for hospitalizations.
 Here $k \in \{1, ..., K\}$, 
@@
 two components, but the propagation of uncertainty in ILI forecasts was still 
-accomplished in the estimating of hospitalization forecasts through
+accomplished in estimating hospitalization forecasts through
 the forecast model (\ref{eq:hosp_post_pred}).
 This two-stage modeling is reminiscent of the Bayesian cut 
-described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
+described in \cite[]{plummer2015cuts} and \cite[]{nott2023bayesian}. Cutting a model
 into two component models or ``modules'' may be justified for several reasons
@@
 application for example, the ILI model parameters receive no potentially 
-useful information from hospiatlization data.
+useful information from hospitalization data.
 
@@
 showed convergence even after 55,000 iterations, whereas the component model 
-fits gave no signs which raised concern about convergence. And most importantly,
+fits gave no signs which raised concerns about convergence. And most importantly,
 the forecasts from the component models generally outperformed those of the 
@@
 % hospitalization models are included in the supplementary material 
-% \cite[]{wadsworth2024bas}.
+% \cite[]{wadsworth2025bas}.
 % 
@@
 fitting criterion. Both the BASE and ARIMA ILI models were fit using the 
-\texttt{forecast} package in \texttt{R} \cite{hyndman2008forecast}.
+\texttt{forecast} package in \texttt{R} \cite[]{hyndman2008forecast}.
 
@@
 all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$. Note that this is not
-a model assumption we make but only a simplyfying assumption for simulating
+a model assumption we make but only a simplifying assumption for simulating
 the data.
@@
 COVID-19 Forecast Hub--
-or forecasts made up of predicitive intervals of several nominal 
+or forecasts made up of predictive intervals of several nominal 
 levels
@@
 10\% to 98\% nominal levels. Each model is undercalibrated, but the ASGD 
-model is the closest to being well calibrated, especially at the hightest 
+model is the closest to being well calibrated, especially at the highest 
 nominal levels. Table \ref{tab:sim_wis_res} shows summaries of the mean WIS and
@@
 % Additional
-% plots in the supplemtary material \cite[]{wadsworth2024bas} further show that 
+% plots in the supplemtary material \cite[]{wadsworth2025bas} further show that 
 % the ASG models often but not always outperform the SIR models according to the 
@@
  Primary scores for evaluating each forecast were the WIS and the 
- log-weighted intveral score (LWIS). The LWIS is defined the same as the WIS
+ log-weighted interval score (LWIS). The LWIS is defined the same as the WIS
  but where the log of the observation as well as the 
@@
 FluSight targets during the 2023 season. 
-The submitted forecasts may be found at \cite{mathis2023flusight}.
+The submitted forecasts may be found at \cite[]{mathis2023flusight}.
 Among the models
@@
 
-Because forecast submissions for some targets were occassionally missed
+Because forecast submissions for some targets were occasionally missed
 by participating teams, for any two models there was often a mismatch between
-targets for which forecasts were included or exlcuded. This makes a
-simple average of scores is an imperfect measure for comparison.
+targets for which forecasts were included or excluded. This makes a
+simple average of scores an imperfect measure for comparison.
 Thus in figure \ref{fig:state_and_date_lwis}, 
@@
 1 indicates the ASGD\_NORM2 model outperforms the competing model. A value 
-higher than 1 indicateds the ASGD\_NORM2 is outperformed. A grey tile indicates
+higher than 1 indicates the ASGD\_NORM2 is outperformed. A grey tile indicates
 the competing model had no forecasts matching the specified target.
@@
  model forecasting hospitalizations where hospitalization predictions are a 
- linear or quadratic function of ILI. Utlizing the ILI data which has been
+ linear or quadratic function of ILI. Utilizing the ILI data which has been
  available for many more years than flu hospitalization, greatly improves
@@
 independent from all other locations. It is reasonable to assume that a
-disease outbreak in one state will share characterstics with the outbreak in 
-neighboring states, and indeed the behavior of ILI and hospitalizations is more
+disease outbreak in one state will share characteristics with the outbreak in 
+neighboring states, and indeed the behavior of ILI and hospitalizations are more
 closely related for neighboring states as shown in data plots in the
@@
 forecasts which were all outperformed by the ensemble forecasts in FluSight.
-Futher extensions of the models herein could include explicit modeling of 
+Further extensions of the models herein could include explicit modeling of 
 spatial relationships between states and inclusion in an ensemble of other
-forecast models.
+forecast models. Additionally, it has been shown that meteorological variables
+such as local temperature forecasts and geographic variables such as age can 
+lead to improved hospitalization forecasts \cite[]{sahu2014hierarchical}. 
+A straightforward and useful extension of the methodology herein would be to 
+collect such variables and incorporate them as added hospitalization 
+predictors.
